### 2022-12-19 expose.tex
first commit

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  12pt,
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{float}
\usepackage[numbers]{natbib}
\let\origfigure\figure
\let\endorigfigure\endfigure
\renewenvironment{figure}[1][2] {
    \expandafter\origfigure\expandafter[H]
} {
    \endorigfigure
}
\usepackage{float}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Pre-Training Denoising Autoencoders for Genetic Programming

(subtitle)

--------------------------------------------------------

Student: Roman Höhn

Date of Birth: 1991-04-14

Place of Birth: Wiesbaden, Hesse

Student ID: 2712497

Supervisor: David Wittenberg

--------------------------------------------------------

Master Thesis

FB 03: Chair of Business Administration and Computer Science

Johannes Gutenberg University Mainz

Summerterm 2022}
\author{}
\date{\vspace{-2.5em}Date of Submission: 2022-09-09}

\begin{document}
\maketitle

\thispagestyle{empty} \newpage{}

\setcounter{page}{1}

\tableofcontents
\thispagestyle{empty}
\newpage

\listoftables
\listoffigures
\thispagestyle{empty}
\newpage

\hypertarget{expose}{%
\section{Expose}\label{expose}}

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

Denoising Autoencoder Genetic Programming (DAE-GP) is a novel variation
of an genetic programming based Estimation of Distribution Algorithm
(EDA-GP) that uses a denoising autoencoders long short-term memory
network (DAE-LSTMs) as a probabilistic model to sample new candidate
solutions (Wittenberg, Rothlauf and Schweim, 2020).

DAE-LSTMs are recurrent neural networks (RNNs) that can be trained in an
unsupervised learning environment to minimize a reconstruction error for
encoding input data into a compressed representation and subsequently
decoding the compressed representation back to the input dimension. In
DAE-GP, DAE-LSTMs are trained with a subset of high-fitness solutions
selected from the parent population as input. The resulting model is
then used to sample new offspring solutions by propagating partially
mutated solutions from the parent population through the DAE-LSTMS
(Wittenberg, Rothlauf and Schweim, 2020).

In previous work DAE-GP has been demonstrated to outperform GP for both
a generalized version of the royal tree problem (Wittenberg, Rothlauf
and Schweim, 2020) as well as for a standard symbolic regression problem
(Wittenberg and Rothlauf, 2022). The DAE-GP algorithm implemented for
both experiments initializes the DAE-LSTMs with randomized parameters
for each generation \(g\) of the search. A possible way of improving the
quality of solutions and/or the total runtime of DAE-GP could be to
instead use pre-trained DAE-LSTMs that are passed on from each
generation \(g_{t}\) to the next generation \(g_{t+1}\). Pre-Training is
a commonly used strategy in deep architectures that has been shown to
improve both the optimization process itself as well as the
generalization behavior if compared to randomly initialized parameters
(Erhan \emph{et al.}, 2009).

\hypertarget{research-question}{%
\subsection{Research Question}\label{research-question}}

The aim of this research project is to study the effects of using
pre-trained DAE-LSTMs in DAE-GP on both solution quality as well as on
total time consumption.

\ldots what problem(s)???\ldots{}

\hypertarget{research-design}{%
\subsection{Research Design}\label{research-design}}

\newpage

\hypertarget{I}{%
\section*{I - References}\label{I}}
\addcontentsline{toc}{section}{I - References}

\hypertarget{refs}{}
\begin{CSLReferences}{0}{0}
\leavevmode\vadjust pre{\hypertarget{ref-pmlr-v5-erhan09a}{}}%
Erhan, D. \emph{et al.} (2009) {``The difficulty of training deep
architectures and the effect of unsupervised pre-training,''} in D. van
Dyk and M. Welling (eds.) \emph{Proceedings of the twelth international
conference on artificial intelligence and statistics}. Hilton Clearwater
Beach Resort, Clearwater Beach, Florida USA: PMLR (Proceedings of
machine learning research), pp. 153--160. Available at:
\url{https://proceedings.mlr.press/v5/erhan09a.html}.

\leavevmode\vadjust pre{\hypertarget{ref-dae-gp_2022_symreg}{}}%
Wittenberg, D. and Rothlauf, F. (2022) {``Denoising autoencoder genetic
programming for real-world symbolic regression,''} in \emph{Proceedings
of the genetic and evolutionary computation conference companion}. New
York, NY, USA: Association for Computing Machinery (GECCO '22), pp.
612--614. Available at: \url{https://doi.org/10.1145/3520304.3528921}.

\leavevmode\vadjust pre{\hypertarget{ref-dae-gp_2020_rtree}{}}%
Wittenberg, D., Rothlauf, F. and Schweim, D. (2020) {``DAE-GP: Denoising
autoencoder LSTM networks as probabilistic models in estimation of
distribution genetic programming,''} in \emph{Proceedings of the 2020
genetic and evolutionary computation conference}. New York, NY, USA:
Association for Computing Machinery (GECCO '20), pp. 1037--1045.
Available at: \url{https://doi.org/10.1145/3377930.3390180}.

\end{CSLReferences}

\hypertarget{II}{%
\section*{II - Statutory Declaration}\label{II}}
\addcontentsline{toc}{section}{II - Statutory Declaration}

\includegraphics{./private/erklaerung.pdf}\\

\end{document}

### 2022-12-19 paper.Rmd
first commit

---
title: |
  <center>Pre-Trained Denoising Autoencoders Long Short-Term Memory Networks as probabilistic Models for Estimation of Distribution Genetic Programming</center>
  <center>--------------------------------------------------------</center>
  <center>Student: Roman Höhn</center>
  <center>Date of Birth: 1991-04-14</center>
  <center>Place of Birth: Wiesbaden, Hesse</center>
  <center>Student ID: 2712497</center>
  <center>Supervisor: David Wittenberg</center>
  <center>--------------------------------------------------------</center>
  <center>Master Thesis</center>
  <center>FB 03: Chair of Business Administration and Computer Science</center>
  <center>Johannes Gutenberg University Mainz</center>
  |
date: "Date of Submission: `r Sys.Date()`"
output:
  pdf_document:
    keep_tex: true
    extra_dependencies: ["float"]
    fig_caption: yes
    number_sections: true
    includes:
      in_header: [header_files/settings.tex]
indent: false
fontsize: 11pt
bibliography: ref/ref.bib
csl: csl/association-for-computing-machinery.csl
#csl: csl/harvard-cite-them-right.csl
#csl: csl/springer-science-reviews.csl
include-before: '`\thispagestyle{empty} \newpage{}`{=latex}'
---

<!-- Python/R Setup -->

```{r setup, include=FALSE}
library(kableExtra)
library(magrittr)
library(reticulate)

EXPERIMENT_DIR<-("../eda-gp-2020/experiments")

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.pos = "!H")

```

```{python pytest, , echo=FALSE, include=FALSE}
from sys import version, executable
print(f"Python Version: {version} \nPath: {executable}")
```

<!-- LaTeX Setup -->
\setcounter{page}{1}
\tableofcontents
\thispagestyle{empty}
\newpage  
\listoftables
\listoffigures
\thispagestyle{empty}
\newpage

<!--Markdown Text -->

# Abstract

...

# Introduction

<!--- Einführung in die Thematik -->

Denoising Autoencoder Genetic Programming (DAE-GP) is a novel variation of an genetic programming based Estimation of Distribution Algorithm (EDA-GP) that uses a denoising autoencoders long short-term memory network (DAE-LSTMs) as a probabilistic model to sample new candidate solutions [@dae-gp_2020_rtree].

DAE-LSTMs are recurrent neural networks (RNNs) that can be trained in an unsupervised learning environment to minimize a reconstruction error for encoding input data into a compressed representation and subsequently decoding the compressed representation back to the input dimension. 
In DAE-GP, DAE-LSTMs are trained with a subset of high-fitness solutions selected from a parent population with the aim to capture their promising qualities. 
The resulting model is then used to sample new offspring solutions by propagating partially mutated solutions from the parent population through the DAE-LSTMS [@dae-gp_2020_rtree].
In previous work DAE-GP has been shown to outperform GP for both a generalized version of the royal tree problem [@dae-gp_2020_rtree] as well as for a real-world symbolic regression problem [@dae-gp_2022_symreg].

<!--- Problemstellung und Zielsetzung der Arbeit -->

The DAE-GP algorithm first described by [@dae-gp_2020_rtree] trains a DAE-LSTMs for each generation $g$ of the search from scratch.
@dae-gp_2022_symreg and @daegp_explore_exploit suggests the incorporation of a pre-training strategy into the evolutionary search as a possible way of improving the model quality of DAE-GP.
The key idea is to pre-train an initial DAE-LSTM on a large population of candidate solutions and to use the weights of this initial model in each generation of the search as a starting point.
This thesis studies the influence of using pre-trained DAE-LSTMs in DAE-GP for symbolic regression, especially looking at the influence in generalization behavior and the overall quality of solutions found by DAE-GP. The aim of this study is to answer the question if a pre-training strategy can help to improve generalization in DAE-GP and if so, analyze how this improved generalization ability benefits the overall performance of DAE-GP.

<!--- Gang der Untersuchung -->

<!--- Describe experiments... -->
...

# Theoretical Foundations

*This section describes the relevant concepts that are necessary for the understanding and classification of DAE-GP as well as the concept of pre-training in artificial neural networks*

## Denoising Autoencoder Genetic Programming

DAE-GP is an EDA-GP algorithm that uses DAE-LSTM networks as a probabilistic model to sample new offspring solutions [@dae-gp_2020_rtree].

### Evolutionary Computation

As a variant of GP, DAE-GP can be classified as a meta-heuristic that belongs to the field of evolutionary computation (EC). 
EC based meta-heuristics are optimization methods that simulate the process of Darwinian evolution to search for high quality solutions by applying selection and variation operators to a population of candidate solutions. 
Examples of EC include genetic algorithms (GA), evolutionary strategies (ES) and GP. 
In EC, the quality of a solution is commonly measured as fitness and the time steps of the search are called generations. 
Another important concept in EC is the distinction between genotypes and phenotypes of solutions, the genotype contains the information that is necessary to construct the phenotype, the outer appearance of a particular solution on which we measure the overall quality of solutions. 
The representation of a solution is therefore defined by the mapping of genotypes to phenotypes [@design_of_modern_heuristics]. 
Genetic operators, such as mutation or recombination are usually applied to the genotype of solutions.

### Genetic Programming

GP follows the same basic evolutionary principle of EC but searches for more general, hierarchical computer programs of dynamically varying size and shape [@Koza1993GeneticP]. 
The computer programs that are at the center of the evolutionary search in GP are commonly represented by tree structures at the level of their phenotype [@design_of_modern_heuristics]. 
Since GP searches for high fitness computer programs that produce a desired output for some input, it can be applied to various different problem domains such as symbolic regression, automatic programming, or evolving game-playing strategies [@Koza1993GeneticP]. 
An important quality of GP is the ability to search for solutions of variable length and structure. 
GP is an especially useful meta heuristic for problems where no a priori knowledge about the final form of good solutions is available.
GPs ability to optimize solutions for their structure as well as for their parameters led to it being one of the most prevalent methods used symbolic regression [@10.1007/978-3-540-24621-3_22].

### Estimation of Distribution Algorithms

The aim of Estimation of distribution algorithms (EDA) is to replace the standard variation operators used in GA by building probabilistic models that can capture complex dependencies between different decision variables of an optimization problem [@design_of_modern_heuristics]. 
EDAs use this probabilistic model to sample new offspring solutions inside an evolutionary search to replace crossover and/or mutation operators.

### Denoising Autoencoders

One possible way of model building in EDA proposed by @harmless_overfitting_eda is to use denoising autoencoders (DAEs) as generative models to capture complex probability distributions of decision variables.

DAE, a variation of the autoencoders (AE), are a widely used type of neural networks in the domain of unsupervized machine learning that maps $n$ input variables to $n$ output variables using a hidden representation.

AE were introduced by @ae_orig to compress high-dimensional data into lower-dimensions. An AE consists of two different subunits:

  * Encoder $g(x)$: Encode input data to a smaller cental layer $h$
  * Decoder $d(h)$: Decode and output the the encoded data back to its original dimension

The AE is trained to reduce the reconstruction error between input and output data, after the training procedure is finished the network is able to reduce the dimensionality of input data to a compressed representation[@ae_orig].

DAE was first introduced by @10.1145/1390156.1390294 as an improved AE with the ability to learn new representations of data that is especially robust to partially corrupted input data. DAE modifies the AE by using partially corrupted input data for the AE and training it to reconstruct the uncorrupted, original version of the input data.

Since the hidden representation of DAE captures the dependency structure of the input data it can therefore also be used to generate new solutions in the context of GAs [@harmless_overfitting_eda].

DAE-GP builds upon the concept of using DAEs in EDAs described by @harmless_overfitting_eda and transfers the concept to the domain of GP. 
The mutation and crossover operators of standard GP are replaced by sampling new solutions from a probabilistic model that is build by training a DAEs long short-term memory (LSTM) network on a subset of high fitness solutions from the current population [@dae-gp_2020_rtree]. 
LSTMs are a variant of recurrent neural networks first introduced by @lstm_orig that can store learned information over en extended time period while avoiding the problem of vanishing and/or exploding gradients. Since DAE-GP encodes candidate solutions as linear strings in prefix notation, the DAE in DAE-GP uses LSTMs for both encoding and decoding where the total amount of time steps $T$ is equal to sum of the length of the input solution and the output solution [@dae-gp_2020_rtree].

## Pre-Training

<!--- 
chatGPT
---

Q: what is pre-training in deep learning?
A: Pre-training in deep learning refers to the process of training a deep learning model on a large dataset before using it for a specific task. The goal of pre-training is to create a model that has a strong general understanding of the data and can be fine-tuned for a specific task.

Pre-training can be done in two ways: unsupervised pre-training and supervised pre-training. In unsupervised pre-training, the model is trained on a large dataset without labels, using techniques such as self-supervised learning or generative adversarial networks (GANs) to learn the underlying patterns in the data. In supervised pre-training, the model is trained on a large dataset with labels, using a supervised learning algorithm to learn the relationship between the input data and the corresponding labels.

Pre-training has several benefits. It can help improve the performance of the model on the target task by providing it with a better initial set of weights. It can also speed up training by reducing the amount of data and computational resources needed to train the model from scratch. Additionally, pre-trained models can be used as a starting point for transfer learning, allowing you to leverage the knowledge and experience of the pre-trained model for a new task.

Overall, pre-training is an important technique in deep learning that can help improve model performance and efficiency. It is commonly used in natural language processing and computer vision applications, where large amounts of labeled data are available.
-->

The strategy of pre-training artificial neural networks is based on the idea of transfer learning. 
Transfer learning in the field of machine learning aims at retaining previously learned knowledge from one task to re-use it for another task [@HAN2021225] [@survey_transfer_learning]. Transfer learning traditionally follows a two phase approach:

1. Pre-Training Phase: Capture knowledge from source task
2. Fine-Tuning Phase: Transfer knowledge to the target task

where source task and target task are usually similar but may differ in their feature space and the distribution of training data [@survey_transfer_learning].
One of the main motivations for using transfer learning in real world machine learning tasks is the ability to reduce the need for large amounts of training data which can often be unavailable or too expensive to collect.

Pre-Training is a commonly used strategy in deep architectures that has been shown to improve both the optimization process itself as well as the generalization behavior if compared to randomly initialized parameters [@pmlr-v5-erhan09a].

...describe different pre-training strategies, e.g. few shot, perpetual, re-using...

-- describe unsupervized vs supervized pre-training...

# Pre-Training in DAE-GP

The pre-training strategy used in this thesis is applied to the DAE-LSTM model $M_g$ that are used in each DAE-GP generation $g$ for $g\in{1,...,g_{max}}$. 
Instead of initializing and optimizing each model from scratch as done in previous work (e.g. [@dae-gp_2020_rtree] [@dae-gp_2022_symreg]), a separate DAE-LSTM network $\hat{M}$ will be trained on a large population of randomly initialized solutions.  
The optimal parameters obtained after finishing the training procedure of $\hat{M}$ are then used as the starting parameters for each following DAE-LSTM model $M_g$ for $g\in{1,...,g_{max}}$.

The motivation for incorporating pre-training into DAE-GP is based on the following suspected mechanisms of improvement:

1. Improve overall performance of DAE-GP by improving either run time and/or solution quality
2. Reduce the need for large population sizes to avoid sampling error
3. Improve model robustness and generalization ability

The probably most obvious motivation for using a pre-training strategy in DAE-GP is based on the fact, that initializing $M_g$ with pre-trained weights from $\hat{M}$ is likely to reduce the amount of training epochs that are necessary until the point of training error convergence is reached.

Early research on pre-training for DAE by @pmlr-v5-erhan09a comes to the conclusion that besides adding robustness to models the strategy also results in improved generalization and better performing models. The authors describe that unsupervised pre-training behaves like a regulizer on DAE networks, the mean test error and it's variance are reduced for DAE networks that are initialized from pre-trained weights if compared to the same architectures that use random initialization.

Another important finding by @pmlr-v5-erhan09a is that the positive effect of pre-training described above is dependent on both the depth of the network as well as the size of layers - while increasingly larger networks benefit increasingly more from pre-training the final performance of small architectures tends to be worse with pre-training than with randomized weight initialization. The authors find evidence that pre-training DAE is especially usefull for optimizing the parameters of the lower layers of the network.

Another motivation for introducing pre-training into DAE-GP is the prevalence of sampling error for small GP populations sizes.
[@sampling_err_gp] finds that sampling error, non-systematic errors that are caused by observing only a small subset of the statistical population, is a severe problem in the domain of GP since the initial population may not be a representative sample of all possible solutions.
[@sampling_err_gp] also introduces a method for calculating optimal population sizes to minimize the presence of sampling error.
Since DAE-LSTM of DAE-GP learns the properties of it's training population and reuses the acquired knowledge in the sampling procedure, using the parameters obtained from training $\hat{M}$ on a sufficiently large training population might reduce the need for large population sizes in the following generations of the search if $\hat{M}$ already implicitly captured the properties of a representative sample of solutions.
If this mechanism can be applied successfully it would benefit the performance of DAE-GP by increasing the population diversity (resulting in better solution quality) as well as by reducing the need for large population sizes which require computational resources.

## Implementation

<!--- Explain algorithm and pre-training strategy in pseudo code -->

The DAE-GP algorithm, first described by @dae-gp_2020_rtree, is described by algorithm 1

```{.numberLines}
WHILE NOT termination_criterion
  DO 
END WHILE
RETURN
```

One difficulty in the implementation of a pre-training strategy into DAE-GP has been the detemination of the number of hidden neurons in side the DAE-LSTMs hidden layers. The original description of DAE-GP [@dae-gp_2020_rtree] used a strategy where the number of hidden neurons for each hidden layer is dynamically set per generation to the maximum individual size inside the current population.
For a pre-training implementation, this strategy can not be easily adapted since it leads to a changing number of neurons at each generation resulting in different dimensions of the DAE-LSTM. To allow for the sharing of weights and biases from the pre-trained model $\hat{M}$ to each $M_g$ for $g\in{1,...,g_{max}}$ I experimented with two different strategies:

1. Set the number of hidden neurons per hidden layer for each generation to the maximum individual size inside the pre-training population
2. (dependent on the terminal/function set)

The framework that was used to conduct all experiments of this thesis was provided by the supervisor David Wittenberg and uses the python programming language in conjunction with the baseline libraries `Keras` [@chollet2015keras] for deep neural networks and `deap` [@DEAP_JMLR2012] for evolutionary computation.

## Benchmark Problem

<!--- Symbolic regression -->
To test pre-training in DAE-GP this thesis focuses on the domain of real-world symbolic regression problems. Symbolic Regression problems have been one of the first GP applications [@Koza1993GeneticP] and are an actively studied and highly relevant research area. The goal in symbolic regression is to find a mathematical model for a given set of data points [@10.1007/978-3-540-24621-3_22], in real-world symbolic regression these data points are sourced from real-world observations which in contrast to synthetic symbolic regression problem are more likely to contain random noise and bias. Another important challenge in solving real-world symbolic regression problems is the ability for a given model to generalize, we want the final model to show high accuracy in predicting outcomes for previously unseen cases.

<!--- airfoil dataset -->
The main experiments conducted in this thesis uses the NASA Airfoil Self-Noise Data Set which is part of the UCI machine learning repository [@machine_learning_repo]. The dataset consists of 5 input variables and 1 output variable that are listed in table 1.

```{r airfoil_dataset_description, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
fig.pos="H"
knitr::kable(
  read.csv("./tables/airfoil_vars.csv"),
  caption="Airfoil - Dataset Description"
)%>% row_spec(0,bold=TRUE ) %>% kableExtra::kable_styling(latex_options = "hold_position")
```

The objective of the airfoil problem is to find a function that accurately predicts the output variable $y$ by taking in a subset of the input variables $x1,x2,x3,x4,x5$. The function set used by all DAE-GP variations for the airfoil problem is summarized in table 2, the terminal set consists of the 5 input variables $x1,x2,x3,x4,x5$ and ephemeral random integers in the range of $[{}-5,..,5]$ [@dae-gp_2022_symreg].

```{r airfoil_function_set, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
fig.pos="H"
knitr::kable(
  read.csv("./tables/airfoil_function_set"),
  caption="Airfoil - Function Set"
)%>% row_spec(0,bold=TRUE ) %>% kableExtra::kable_styling(latex_options = "hold_position")
```

<!--- energy heating dataset -->

# Results

## Exploratory Phase

In a first exploratory experiment DAE-GP was tested on the Aifoil dataset to perform a symbolic regression task. The aim of this initial experiment is to search for differences in behavior between an algorithm that uses a pre-training strategy and regular DAE-GP. The parameters for the experiment are summarized in table 2, the average results for fitness, number of fitness evaluations, average tree size of the population and the unique rate of the population are summarized in plot 1.

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
fig.pos="H"
knitr::kable(
  read.csv("/Users/rmn/masterThesis/eda-gp-2020/experiments/airfoil_ptdaegp-daegp_rwsr_1hl/parameters.csv"),
  digits=3,
  caption="Airfoil - Configuration"
)%>% row_spec(0,bold=TRUE ) %>% kableExtra::kable_styling(latex_options = "hold_position")
```

![Airfoil - First Results](./img/airfoil_exploratory/airfoil_exploratory.png)

In this first experiment, the average quality of solutions produced by the pre-trained algorithm shows to be lower than that produced by regular DAE-GP. Differences in fitness start to show at generation 5 where regular DAE-GP starts to reduces RMSE more efficiently than the pre-trained version. The gap in average fitness grows larger until the end of the evolutionary search between both algorithms. Observing the gap between RMSE on the training dataset and the testing dataset shows no major differences, the training error is slightly lower than the testing error for both algorithms and indicates a similar generalization behavior.

Another interesting metric to investigate is the population's average tree size, in general smaller solutions are preferable to their larger counterparts because of improved interpretability. The results show, that average tree size is reduced strongly in generations 1-5 for both algorithms and than slowly increases back. Differences between regular DAE-GP and the pre-trained DAE-GP start to show at generation 2, where the average population tree size is continuously lower for regular DAE-GP.

It can also be observed that the population diversity, measured by the normalized levenshtein distance, is different between both DAE-GP variants. Both algorithms start at a unique rate of about 0.8 in the 0.th generation and lower rapidly to an average of about 0.45 in the 7.th generation. In the following generations, the diversity of pre-trained DAE-GP keeps lowering steadily while regular DAE-GP slowly increases the population diversity to a normalized levenshtein distance of about 0.55. This could be a first explanation for the better performance of regular DAE-GP in this experiment, by creating more diverse populations over the later generations of the search the algorithm is better at exploiting the search space than the pre-trained DAE-GP algorithm.

## Effect on Generalisation

To gather a deeper understanding about the effect of pre-training DAE-LSTM in DAE-GP a series of experiments was conducted, again using the airfoil dataset for symbolic regression, and different parameter configurations for the number of hidden layers as well as the number of hidden neurons per hidden layer.

To study generalization behavior this series of experiments is conducted using DAE-GP with only a single generation until the search terminates. I disregard the fitness of solutions found and focus soley on the reconstruction error that is produced during the training of each the DAE-LSTM. The reconstruction error is measured for two separate populations, a training population $train\_pop$ that is used to train our DAE-LSTM as well as a hold-out validation population $test\_pop$. For the pre-trained DAE-GP two additional, separate populations $\hat{train\_pop}$ and  $\hat{test\_pop}$ are used only for pre-training.

DAE-GP is tested in two different configurations:

* Variable number of hidden neurons (50, 100, 200) with a single hidden layer
* Fixed number of hidden neurons (100) per hidden layer with variable number of hidden layers (1, 2, 3)

For each configuration I tested traditional DAE-GP as well as a pre-trained DAE-GP resulting in 12 total sub experiments that were each based on 10 individual runs (total number of runs=$120$). To avoid creating biased results through the presence of sampling error inside the pre-training population (see @sampling_err_gp), the population size for the pre-training phase is chosen very high with $\hat{pop\_size}=20000$ where 50% of $\hat{pop}$ is used for the training population $\hat{train\_pop}$ and the remaining 50% are used for the hold-out validation population $\hat{test\_pop}$. The training of DAE-LSTM uses a fixed number of 1000 epochs until termination and uses Adam optimization for gradient descent. The reason for using a high amount of 1000 fixed training epochs is to deliberately force the DAE-LSTM to overfit to the training data.

In general I expect that with a growing number of model parameters (either by adding more hidden layers or more hidden neurons per layer) the DAE-LSTM will be more prone to overfit to the training population $train\_pop$ resulting in a small reconstruction error for the training population and a large one for the validation population $train\_pop$. The reason for this effect is that a larger network trained over an extended period of time (without the use of strategies like early stopping), has much more potential to learn noise from the training dataset than a smaller network, which is more likely to result in worsening performance for previously unseen cases [@weigend1994overfitting].

Based on the review of [@pmlr-v5-erhan09a] I also expect that pre-training will have an insignificant or even negative influence on small DAE-LSTM instances while improving the networks generalization ability with growing size.

<!--- analyze results 1 hidden layer and variable number of hidden neurons -->

![Airfoil - First Generation Median Training Error for variable number of hidden Neurons](./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_neurons.png)

![Airfoil - First Generation final Training Error for variable number of hidden Neurons](./img/airfoil_firstGen/airfoil_firstGen_boxplot_training_error_by_neurons.png)

<!--- analyzing by layers -->

# Discussion

...

# Limitations and open Questions

## Further Questions

[@pmlr-v5-erhan09a]: Pre-Training effect especially usefull for lower-level layers -> Only adapt weights from low layers or embedd the pre-trained model?

Different pre-training strategies?

\newpage

# References {.unnumbered #I}

::: {#refs}
:::

<!--
  EIDESSTAATLICHE ERKLAERUNG
# II - Statutory Declaration {.unnumbered #II}

![Statutory Declaration](./private/erklaerung.pdf)\

-->

### 2022-12-19 paper.tex
first commit

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  11pt,
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{float}
\usepackage[numbers]{natbib}
\let\origfigure\figure
\let\endorigfigure\endfigure
\renewenvironment{figure}[1][2] {
    \expandafter\origfigure\expandafter[H]
} {
    \endorigfigure
}
\usepackage{float}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Pre-Trained Denoising Autoencoders Long Short-Term Memory Networks as probabilistic Models for Estimation of Distribution Genetic Programming

--------------------------------------------------------

Student: Roman Höhn

Date of Birth: 1991-04-14

Place of Birth: Wiesbaden, Hesse

Student ID: 2712497

Supervisor: David Wittenberg

--------------------------------------------------------

Master Thesis

FB 03: Chair of Business Administration and Computer Science

Johannes Gutenberg University Mainz}
\author{}
\date{\vspace{-2.5em}Date of Submission: 2022-12-18}

\begin{document}
\maketitle

\thispagestyle{empty} \newpage{}

{
\setcounter{tocdepth}{2}
\tableofcontents
}
\setcounter{page}{1}
\tableofcontents
\thispagestyle{empty}
\newpage  
\listoftables
\listoffigures
\thispagestyle{empty}
\newpage

\hypertarget{abstract}{%
\section{Abstract}\label{abstract}}

\ldots{}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Denoising Autoencoder Genetic Programming (DAE-GP) is a novel variation of an genetic programming based Estimation of Distribution Algorithm (EDA-GP) that uses a denoising autoencoders long short-term memory network (DAE-LSTMs) as a probabilistic model to sample new candidate solutions {[}18{]}.

DAE-LSTMs are recurrent neural networks (RNNs) that can be trained in an unsupervised learning environment to minimize a reconstruction error for encoding input data into a compressed representation and subsequently decoding the compressed representation back to the input dimension.
In DAE-GP, DAE-LSTMs are trained with a subset of high-fitness solutions selected from a parent population with the aim to capture their promising qualities.
The resulting model is then used to sample new offspring solutions by propagating partially mutated solutions from the parent population through the DAE-LSTMS {[}18{]}.
In previous work DAE-GP has been shown to outperform GP for both a generalized version of the royal tree problem {[}18{]} as well as for a real-world symbolic regression problem {[}17{]}.

The DAE-GP algorithm first described by {[}18{]} trains a DAE-LSTMs for each generation \(g\) of the search from scratch.
{[}17{]} and {[}16{]} suggests the incorporation of a pre-training strategy into the evolutionary search as a possible way of improving the model quality of DAE-GP.
The key idea is to pre-train an initial DAE-LSTM on a large population of candidate solutions and to use the weights of this initial model in each generation of the search as a starting point.
This thesis studies the influence of using pre-trained DAE-LSTMs in DAE-GP for symbolic regression, especially looking at the influence in generalization behavior and the overall quality of solutions found by DAE-GP. The aim of this study is to answer the question if a pre-training strategy can help to improve generalization in DAE-GP and if so, analyze how this improved generalization ability benefits the overall performance of DAE-GP.

\ldots{}

\hypertarget{theoretical-foundations}{%
\section{Theoretical Foundations}\label{theoretical-foundations}}

\emph{This section describes the relevant concepts that are necessary for the understanding and classification of DAE-GP as well as the concept of pre-training in artificial neural networks}

\hypertarget{denoising-autoencoder-genetic-programming}{%
\subsection{Denoising Autoencoder Genetic Programming}\label{denoising-autoencoder-genetic-programming}}

DAE-GP is an EDA-GP algorithm that uses DAE-LSTM networks as a probabilistic model to sample new offspring solutions {[}18{]}.

\hypertarget{evolutionary-computation}{%
\subsubsection{Evolutionary Computation}\label{evolutionary-computation}}

As a variant of GP, DAE-GP can be classified as a meta-heuristic that belongs to the field of evolutionary computation (EC).
EC based meta-heuristics are optimization methods that simulate the process of Darwinian evolution to search for high quality solutions by applying selection and variation operators to a population of candidate solutions.
Examples of EC include genetic algorithms (GA), evolutionary strategies (ES) and GP.
In EC, the quality of a solution is commonly measured as fitness and the time steps of the search are called generations.
Another important concept in EC is the distinction between genotypes and phenotypes of solutions, the genotype contains the information that is necessary to construct the phenotype, the outer appearance of a particular solution on which we measure the overall quality of solutions.
The representation of a solution is therefore defined by the mapping of genotypes to phenotypes {[}12{]}.
Genetic operators, such as mutation or recombination are usually applied to the genotype of solutions.

\hypertarget{genetic-programming}{%
\subsubsection{Genetic Programming}\label{genetic-programming}}

GP follows the same basic evolutionary principle of EC but searches for more general, hierarchical computer programs of dynamically varying size and shape {[}8{]}.
The computer programs that are at the center of the evolutionary search in GP are commonly represented by tree structures at the level of their phenotype {[}12{]}.
Since GP searches for high fitness computer programs that produce a desired output for some input, it can be applied to various different problem domains such as symbolic regression, automatic programming, or evolving game-playing strategies {[}8{]}.
An important quality of GP is the ability to search for solutions of variable length and structure.
GP is an especially useful meta heuristic for problems where no a priori knowledge about the final form of good solutions is available.
GPs ability to optimize solutions for their structure as well as for their parameters led to it being one of the most prevalent methods used symbolic regression {[}10{]}.

\hypertarget{estimation-of-distribution-algorithms}{%
\subsubsection{Estimation of Distribution Algorithms}\label{estimation-of-distribution-algorithms}}

The aim of Estimation of distribution algorithms (EDA) is to replace the standard variation operators used in GA by building probabilistic models that can capture complex dependencies between different decision variables of an optimization problem {[}12{]}.
EDAs use this probabilistic model to sample new offspring solutions inside an evolutionary search to replace crossover and/or mutation operators.

\hypertarget{denoising-autoencoders}{%
\subsubsection{Denoising Autoencoders}\label{denoising-autoencoders}}

One possible way of model building in EDA proposed by {[}11{]} is to use denoising autoencoders (DAEs) as generative models to capture complex probability distributions of decision variables.

DAE, a variation of the autoencoders (AE), are a widely used type of neural networks in the domain of unsupervized machine learning that maps \(n\) input variables to \(n\) output variables using a hidden representation.

AE were introduced by {[}6{]} to compress high-dimensional data into lower-dimensions. An AE consists of two different subunits:

\begin{itemize}
\tightlist
\item
  Encoder \(g(x)\): Encode input data to a smaller cental layer \(h\)
\item
  Decoder \(d(h)\): Decode and output the the encoded data back to its original dimension
\end{itemize}

The AE is trained to reduce the reconstruction error between input and output data, after the training procedure is finished the network is able to reduce the dimensionality of input data to a compressed representation{[}6{]}.

DAE was first introduced by {[}14{]} as an improved AE with the ability to learn new representations of data that is especially robust to partially corrupted input data. DAE modifies the AE by using partially corrupted input data for the AE and training it to reconstruct the uncorrupted, original version of the input data.

Since the hidden representation of DAE captures the dependency structure of the input data it can therefore also be used to generate new solutions in the context of GAs {[}11{]}.

DAE-GP builds upon the concept of using DAEs in EDAs described by {[}11{]} and transfers the concept to the domain of GP.
The mutation and crossover operators of standard GP are replaced by sampling new solutions from a probabilistic model that is build by training a DAEs long short-term memory (LSTM) network on a subset of high fitness solutions from the current population {[}18{]}.
LSTMs are a variant of recurrent neural networks first introduced by {[}7{]} that can store learned information over en extended time period while avoiding the problem of vanishing and/or exploding gradients. Since DAE-GP encodes candidate solutions as linear strings in prefix notation, the DAE in DAE-GP uses LSTMs for both encoding and decoding where the total amount of time steps \(T\) is equal to sum of the length of the input solution and the output solution {[}18{]}.

\hypertarget{pre-training}{%
\subsection{Pre-Training}\label{pre-training}}

The strategy of pre-training artificial neural networks is based on the idea of transfer learning.
Transfer learning in the field of machine learning aims at retaining previously learned knowledge from one task to re-use it for another task {[}5{]} {[}9{]}. Transfer learning traditionally follows a two phase approach:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Pre-Training Phase: Capture knowledge from source task
\item
  Fine-Tuning Phase: Transfer knowledge to the target task
\end{enumerate}

where source task and target task are usually similar but may differ in their feature space and the distribution of training data {[}9{]}.
One of the main motivations for using transfer learning in real world machine learning tasks is the ability to reduce the need for large amounts of training data which can often be unavailable or too expensive to collect.

Pre-Training is a commonly used strategy in deep architectures that has been shown to improve both the optimization process itself as well as the generalization behavior if compared to randomly initialized parameters {[}3{]}.

\ldots describe different pre-training strategies, e.g.~few shot, perpetual, re-using\ldots{}

-- describe unsupervized vs supervized pre-training\ldots{}

\hypertarget{pre-training-in-dae-gp}{%
\section{Pre-Training in DAE-GP}\label{pre-training-in-dae-gp}}

The pre-training strategy used in this thesis is applied to the DAE-LSTM model \(M_g\) that are used in each DAE-GP generation \(g\) for \(g\in{1,...,g_{max}}\).
Instead of initializing and optimizing each model from scratch as done in previous work (e.g. {[}18{]} {[}17{]}), a separate DAE-LSTM network \(\hat{M}\) will be trained on a large population of randomly initialized solutions.\\
The optimal parameters obtained after finishing the training procedure of \(\hat{M}\) are then used as the starting parameters for each following DAE-LSTM model \(M_g\) for \(g\in{1,...,g_{max}}\).

The motivation for incorporating pre-training into DAE-GP is based on the following suspected mechanisms of improvement:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Improve overall performance of DAE-GP by improving either run time and/or solution quality
\item
  Reduce the need for large population sizes to avoid sampling error
\item
  Improve model robustness and generalization ability
\end{enumerate}

The probably most obvious motivation for using a pre-training strategy in DAE-GP is based on the fact, that initializing \(M_g\) with pre-trained weights from \(\hat{M}\) is likely to reduce the amount of training epochs that are necessary until the point of training error convergence is reached.

Early research on pre-training for DAE by {[}3{]} comes to the conclusion that besides adding robustness to models the strategy also results in improved generalization and better performing models. The authors describe that unsupervised pre-training behaves like a regulizer on DAE networks, the mean test error and it's variance are reduced for DAE networks that are initialized from pre-trained weights if compared to the same architectures that use random initialization.

Another important finding by {[}3{]} is that the positive effect of pre-training described above is dependent on both the depth of the network as well as the size of layers - while increasingly larger networks benefit increasingly more from pre-training the final performance of small architectures tends to be worse with pre-training than with randomized weight initialization. The authors find evidence that pre-training DAE is especially usefull for optimizing the parameters of the lower layers of the network.

Another motivation for introducing pre-training into DAE-GP is the prevalence of sampling error for small GP populations sizes.
{[}13{]} finds that sampling error, non-systematic errors that are caused by observing only a small subset of the statistical population, is a severe problem in the domain of GP since the initial population may not be a representative sample of all possible solutions.
{[}13{]} also introduces a method for calculating optimal population sizes to minimize the presence of sampling error.
Since DAE-LSTM of DAE-GP learns the properties of it's training population and reuses the acquired knowledge in the sampling procedure, using the parameters obtained from training \(\hat{M}\) on a sufficiently large training population might reduce the need for large population sizes in the following generations of the search if \(\hat{M}\) already implicitly captured the properties of a representative sample of solutions.
If this mechanism can be applied successfully it would benefit the performance of DAE-GP by increasing the population diversity (resulting in better solution quality) as well as by reducing the need for large population sizes which require computational resources.

\hypertarget{implementation}{%
\subsection{Implementation}\label{implementation}}

The DAE-GP algorithm, first described by {[}18{]}, is described by algorithm 1

\begin{Shaded}
\begin{Highlighting}[numbers=left,,]
\NormalTok{WHILE NOT termination\_criterion}
\NormalTok{  DO }
\NormalTok{END WHILE}
\NormalTok{RETURN}
\end{Highlighting}
\end{Shaded}

One difficulty in the implementation of a pre-training strategy into DAE-GP has been the detemination of the number of hidden neurons in side the DAE-LSTMs hidden layers. The original description of DAE-GP {[}18{]} used a strategy where the number of hidden neurons for each hidden layer is dynamically set per generation to the maximum individual size inside the current population.
For a pre-training implementation, this strategy can not be easily adapted since it leads to a changing number of neurons at each generation resulting in different dimensions of the DAE-LSTM. To allow for the sharing of weights and biases from the pre-trained model \(\hat{M}\) to each \(M_g\) for \(g\in{1,...,g_{max}}\) I experimented with two different strategies:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Set the number of hidden neurons per hidden layer for each generation to the maximum individual size inside the pre-training population
\item
  (dependent on the terminal/function set)
\end{enumerate}

The framework that was used to conduct all experiments of this thesis was provided by the supervisor David Wittenberg and uses the python programming language in conjunction with the baseline libraries \texttt{Keras} {[}1{]} for deep neural networks and \texttt{deap} {[}4{]} for evolutionary computation.

\hypertarget{benchmark-problem}{%
\subsection{Benchmark Problem}\label{benchmark-problem}}

To test pre-training in DAE-GP this thesis focuses on the domain of real-world symbolic regression problems. Symbolic Regression problems have been one of the first GP applications {[}8{]} and are an actively studied and highly relevant research area. The goal in symbolic regression is to find a mathematical model for a given set of data points {[}10{]}, in real-world symbolic regression these data points are sourced from real-world observations which in contrast to synthetic symbolic regression problem are more likely to contain random noise and bias. Another important challenge in solving real-world symbolic regression problems is the ability for a given model to generalize, we want the final model to show high accuracy in predicting outcomes for previously unseen cases.

The main experiments conducted in this thesis uses the NASA Airfoil Self-Noise Data Set which is part of the UCI machine learning repository {[}2{]}. The dataset consists of 5 input variables and 1 output variable that are listed in table 1.

\begin{table}[!h]

\caption{(\#tab:airfoil_dataset_description)Airfoil - Dataset Description}
\centering
\begin{tabular}[t]{l|l|l|l}
\hline
\textbf{Type} & \textbf{Name} & \textbf{Description} & \textbf{Unit}\\
\hline
input & x1 & Frequency & Hertz\\
\hline
input & x2 & Angle of attack & Degree\\
\hline
input & x3 & Chord length & meters\\
\hline
input & x4 & Free-stream velocity & meters/second\\
\hline
input & x5 & Suction side displacement thickness & meters\\
\hline
output & y & Scaled sound pressure level & decibels\\
\hline
\end{tabular}
\end{table}

The objective of the airfoil problem is to find a function that accurately predicts the output variable \(y\) by taking in a subset of the input variables \(x1,x2,x3,x4,x5\). The function set used by all DAE-GP variations for the airfoil problem is summarized in table 2, the terminal set consists of the 5 input variables \(x1,x2,x3,x4,x5\) and ephemeral random integers in the range of \([{}-5,..,5]\) {[}17{]}.

\begin{table}[!h]

\caption{(\#tab:airfoil_function_set)Airfoil - Function Set}
\centering
\begin{tabular}[t]{l|r}
\hline
\textbf{function.} & \textbf{arity}\\
\hline
addition & 2\\
\hline
subtraction & 2\\
\hline
multiplication & 2\\
\hline
analytic\_quotient & 2\\
\hline
\end{tabular}
\end{table}

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{exploratory-phase}{%
\subsection{Exploratory Phase}\label{exploratory-phase}}

In a first exploratory experiment DAE-GP was tested on the Aifoil dataset to perform a symbolic regression task. The aim of this initial experiment is to search for differences in behavior between an algorithm that uses a pre-training strategy and regular DAE-GP. The parameters for the experiment are summarized in table 2, the average results for fitness, number of fitness evaluations, average tree size of the population and the unique rate of the population are summarized in plot 1.

\begin{table}[!h]

\caption{\label{tab:unnamed-chunk-1}Airfoil - Configuration}
\centering
\begin{tabular}[t]{l|l}
\hline
\textbf{Hyperparameter} & \textbf{Value}\\
\hline
Training/Testing Split & 50:50\\
\hline
Pre-Training Population Size & 10000\\
\hline
Number of Hidden Layers & 1\\
\hline
Selection & Tournament\\
\hline
Tournament Size & 7\\
\hline
\end{tabular}
\end{table}

\begin{figure}
\centering
\includegraphics{./img/airfoil_exploratory/airfoil_exploratory.png}
\caption{Airfoil - First Results}
\end{figure}

In this first experiment, the average quality of solutions produced by the pre-trained algorithm shows to be lower than that produced by regular DAE-GP. Differences in fitness start to show at generation 5 where regular DAE-GP starts to reduces RMSE more efficiently than the pre-trained version. The gap in average fitness grows larger until the end of the evolutionary search between both algorithms. Observing the gap between RMSE on the training dataset and the testing dataset shows no major differences, the training error is slightly lower than the testing error for both algorithms and indicates a similar generalization behavior.

Another interesting metric to investigate is the population's average tree size, in general smaller solutions are preferable to their larger counterparts because of improved interpretability. The results show, that average tree size is reduced strongly in generations 1-5 for both algorithms and than slowly increases back. Differences between regular DAE-GP and the pre-trained DAE-GP start to show at generation 2, where the average population tree size is continuously lower for regular DAE-GP.

It can also be observed that the population diversity, measured by the normalized levenshtein distance, is different between both DAE-GP variants. Both algorithms start at a unique rate of about 0.8 in the 0.th generation and lower rapidly to an average of about 0.45 in the 7.th generation. In the following generations, the diversity of pre-trained DAE-GP keeps lowering steadily while regular DAE-GP slowly increases the population diversity to a normalized levenshtein distance of about 0.55. This could be a first explanation for the better performance of regular DAE-GP in this experiment, by creating more diverse populations over the later generations of the search the algorithm is better at exploiting the search space than the pre-trained DAE-GP algorithm.

\hypertarget{effect-on-generalisation}{%
\subsection{Effect on Generalisation}\label{effect-on-generalisation}}

To gather a deeper understanding about the effect of pre-training DAE-LSTM in DAE-GP a series of experiments was conducted, again using the airfoil dataset for symbolic regression, and different parameter configurations for the number of hidden layers as well as the number of hidden neurons per hidden layer.

To study generalization behavior this series of experiments is conducted using DAE-GP with only a single generation until the search terminates. I disregard the fitness of solutions found and focus soley on the reconstruction error that is produced during the training of each the DAE-LSTM. The reconstruction error is measured for two separate populations, a training population \(train\_pop\) that is used to train our DAE-LSTM as well as a hold-out validation population \(test\_pop\). For the pre-trained DAE-GP two additional, separate populations \(\hat{train\_pop}\) and \(\hat{test\_pop}\) are used only for pre-training.

DAE-GP is tested in two different configurations:

\begin{itemize}
\tightlist
\item
  Variable number of hidden neurons (50, 100, 200) with a single hidden layer
\item
  Fixed number of hidden neurons (100) per hidden layer with variable number of hidden layers (1, 2, 3)
\end{itemize}

For each configuration I tested traditional DAE-GP as well as a pre-trained DAE-GP resulting in 12 total sub experiments that were each based on 10 individual runs (total number of runs=\(120\)). To avoid creating biased results through the presence of sampling error inside the pre-training population (see {[}13{]}), the population size for the pre-training phase is chosen very high with \(\hat{pop\_size}=20000\) where 50\% of \(\hat{pop}\) is used for the training population \(\hat{train\_pop}\) and the remaining 50\% are used for the hold-out validation population \(\hat{test\_pop}\). The training of DAE-LSTM uses a fixed number of 1000 epochs until termination and uses Adam optimization for gradient descent. The reason for using a high amount of 1000 fixed training epochs is to deliberately force the DAE-LSTM to overfit to the training data.

In general I expect that with a growing number of model parameters (either by adding more hidden layers or more hidden neurons per layer) the DAE-LSTM will be more prone to overfit to the training population \(train\_pop\) resulting in a small reconstruction error for the training population and a large one for the validation population \(train\_pop\). The reason for this effect is that a larger network trained over an extended period of time (without the use of strategies like early stopping), has much more potential to learn noise from the training dataset than a smaller network, which is more likely to result in worsening performance for previously unseen cases {[}15{]}.

Based on the review of {[}3{]} I also expect that pre-training will have an insignificant or even negative influence on small DAE-LSTM instances while improving the networks generalization ability with growing size.

\begin{figure}
\centering
\includegraphics{./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_neurons.png}
\caption{Airfoil - First Generation Median Training Error for variable number of hidden Neurons}
\end{figure}

\begin{figure}
\centering
\includegraphics{./img/airfoil_firstGen/airfoil_firstGen_boxplot_training_error_by_neurons.png}
\caption{Airfoil - First Generation final Training Error for variable number of hidden Neurons}
\end{figure}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

\ldots{}

\hypertarget{limitations-and-open-questions}{%
\section{Limitations and open Questions}\label{limitations-and-open-questions}}

\hypertarget{further-questions}{%
\subsection{Further Questions}\label{further-questions}}

{[}3{]}: Pre-Training effect especially usefull for lower-level layers -\textgreater{} Only adapt weights from low layers or embedd the pre-trained model?

Different pre-training strategies?

\newpage

\hypertarget{I}{%
\section*{References}\label{I}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{0}{0}
\leavevmode\vadjust pre{\hypertarget{ref-chollet2015keras}{}}%
\CSLLeftMargin{{[}1{]} }%
\CSLRightInline{Francois Chollet and others. 2015. Keras. Retrieved from \url{https://github.com/fchollet/keras}}

\leavevmode\vadjust pre{\hypertarget{ref-machine_learning_repo}{}}%
\CSLLeftMargin{{[}2{]} }%
\CSLRightInline{Dheeru Dua and Casey Graff. 2017. {UCI} machine learning repository. Retrieved from \url{http://archive.ics.uci.edu/ml}}

\leavevmode\vadjust pre{\hypertarget{ref-pmlr-v5-erhan09a}{}}%
\CSLLeftMargin{{[}3{]} }%
\CSLRightInline{Dumitru Erhan, Pierre-Antoine Manzagol, Yoshua Bengio, Samy Bengio, and Pascal Vincent. 2009. The difficulty of training deep architectures and the effect of unsupervised pre-training. In \emph{Proceedings of the twelth international conference on artificial intelligence and statistics} (Proceedings of machine learning research), PMLR, Hilton Clearwater Beach Resort, Clearwater Beach, Florida USA, 153--160. Retrieved from \url{https://proceedings.mlr.press/v5/erhan09a.html}}

\leavevmode\vadjust pre{\hypertarget{ref-DEAP_JMLR2012}{}}%
\CSLLeftMargin{{[}4{]} }%
\CSLRightInline{Félix-Antoine Fortin, François-Michel De Rainville, Marc-André Gardner, Marc Parizeau, and Christian Gagné. 2012. {DEAP}: Evolutionary algorithms made easy. \emph{Journal of Machine Learning Research} 13, (2012), 2171--2175.}

\leavevmode\vadjust pre{\hypertarget{ref-HAN2021225}{}}%
\CSLLeftMargin{{[}5{]} }%
\CSLRightInline{Xu Han, Zhengyan Zhang, Ning Ding, Yuxian Gu, Xiao Liu, Yuqi Huo, Jiezhong Qiu, Yuan Yao, Ao Zhang, Liang Zhang, Wentao Han, Minlie Huang, Qin Jin, Yanyan Lan, Yang Liu, Zhiyuan Liu, Zhiwu Lu, Xipeng Qiu, Ruihua Song, Jie Tang, Ji-Rong Wen, Jinhui Yuan, Wayne Xin Zhao, and Jun Zhu. 2021. Pre-trained models: Past, present and future. \emph{AI Open} 2, (2021), 225--250. DOI:https://doi.org/\url{https://doi.org/10.1016/j.aiopen.2021.08.002}}

\leavevmode\vadjust pre{\hypertarget{ref-ae_orig}{}}%
\CSLLeftMargin{{[}6{]} }%
\CSLRightInline{G. E. Hinton and R. R. Salakhutdinov. 2006. Reducing the dimensionality of data with neural networks. \emph{Science} 313, 5786 (2006), 504--507. DOI:https://doi.org/\href{https://doi.org/10.1126/science.1127647}{10.1126/science.1127647}}

\leavevmode\vadjust pre{\hypertarget{ref-lstm_orig}{}}%
\CSLLeftMargin{{[}7{]} }%
\CSLRightInline{Sepp Hochreiter and Jürgen Schmidhuber. 1997. Long short-term memory. \emph{Neural computation} 9, (December 1997), 1735--80. DOI:https://doi.org/\href{https://doi.org/10.1162/neco.1997.9.8.1735}{10.1162/neco.1997.9.8.1735}}

\leavevmode\vadjust pre{\hypertarget{ref-Koza1993GeneticP}{}}%
\CSLLeftMargin{{[}8{]} }%
\CSLRightInline{John R. Koza. 1993. Genetic programming - on the programming of computers by means of natural selection. In \emph{Complex adaptive systems}.}

\leavevmode\vadjust pre{\hypertarget{ref-survey_transfer_learning}{}}%
\CSLLeftMargin{{[}9{]} }%
\CSLRightInline{Sinno Jialin Pan and Qiang Yang. 2010. A survey on transfer learning. \emph{IEEE Transactions on Knowledge and Data Engineering} 22, 10 (2010), 1345--1359. DOI:https://doi.org/\href{https://doi.org/10.1109/TKDE.2009.191}{10.1109/TKDE.2009.191}}

\leavevmode\vadjust pre{\hypertarget{ref-10.1007ux2f978-3-540-24621-3_22}{}}%
\CSLLeftMargin{{[}10{]} }%
\CSLRightInline{Grégory Paris, Denis Robilliard, and Cyril Fonlupt. 2004. Exploring overfitting in genetic programming. In \emph{Artificial evolution}, Springer Berlin Heidelberg, Berlin, Heidelberg, 267--277.}

\leavevmode\vadjust pre{\hypertarget{ref-harmless_overfitting_eda}{}}%
\CSLLeftMargin{{[}11{]} }%
\CSLRightInline{Malte Probst and Franz Rothlauf. 2020. Harmless overfitting: Using denoising autoencoders in estimation of distribution algorithms. \emph{Journal of Machine Learning Research} 21, 78 (2020), 1--31. Retrieved from \url{http://jmlr.org/papers/v21/16-543.html}}

\leavevmode\vadjust pre{\hypertarget{ref-design_of_modern_heuristics}{}}%
\CSLLeftMargin{{[}12{]} }%
\CSLRightInline{Franz Rothlauf. 2011. \emph{Design of modern heuristics: Principles and application}. DOI:https://doi.org/\href{https://doi.org/10.1007/978-3-540-72962-4}{10.1007/978-3-540-72962-4}}

\leavevmode\vadjust pre{\hypertarget{ref-sampling_err_gp}{}}%
\CSLLeftMargin{{[}13{]} }%
\CSLRightInline{Dirk Schweim, David Wittenberg, and Franz Rothlauf. 2021. On sampling error in genetic programming. \emph{Natural computing} 2021, (2021). DOI:https://doi.org/\url{http://doi.org/10.25358/openscience-5820}}

\leavevmode\vadjust pre{\hypertarget{ref-10.1145ux2f1390156.1390294}{}}%
\CSLLeftMargin{{[}14{]} }%
\CSLRightInline{Pascal Vincent, Hugo Larochelle, Yoshua Bengio, and Pierre-Antoine Manzagol. 2008. Extracting and composing robust features with denoising autoencoders. In \emph{Proceedings of the 25th international conference on machine learning} (ICML '08), Association for Computing Machinery, New York, NY, USA, 1096--1103. DOI:https://doi.org/\href{https://doi.org/10.1145/1390156.1390294}{10.1145/1390156.1390294}}

\leavevmode\vadjust pre{\hypertarget{ref-weigend1994overfitting}{}}%
\CSLLeftMargin{{[}15{]} }%
\CSLRightInline{Andreas Weigend. 1994. On overfitting and the effective number of hidden units. In \emph{Proceedings of the 1993 connectionist models summer school}, 335--342.}

\leavevmode\vadjust pre{\hypertarget{ref-daegp_explore_exploit}{}}%
\CSLLeftMargin{{[}16{]} }%
\CSLRightInline{David Wittenberg. 2022. Using denoising autoencoder genetic programming to~control exploration and~exploitation in~search. In \emph{Genetic programming}, Springer International Publishing, Cham, 102--117.}

\leavevmode\vadjust pre{\hypertarget{ref-dae-gp_2022_symreg}{}}%
\CSLLeftMargin{{[}17{]} }%
\CSLRightInline{David Wittenberg and Franz Rothlauf. 2022. Denoising autoencoder genetic programming for real-world symbolic regression. In \emph{Proceedings of the genetic and evolutionary computation conference companion} (GECCO '22), Association for Computing Machinery, New York, NY, USA, 612--614. DOI:https://doi.org/\href{https://doi.org/10.1145/3520304.3528921}{10.1145/3520304.3528921}}

\leavevmode\vadjust pre{\hypertarget{ref-dae-gp_2020_rtree}{}}%
\CSLLeftMargin{{[}18{]} }%
\CSLRightInline{David Wittenberg, Franz Rothlauf, and Dirk Schweim. 2020. DAE-GP: Denoising autoencoder LSTM networks as probabilistic models in estimation of distribution genetic programming. In \emph{Proceedings of the 2020 genetic and evolutionary computation conference} (GECCO '20), Association for Computing Machinery, New York, NY, USA, 1037--1045. DOI:https://doi.org/\href{https://doi.org/10.1145/3377930.3390180}{10.1145/3377930.3390180}}

\end{CSLReferences}

\end{document}

### 2022-12-19 tex/paper.tex
first commit

--- 2022-12-19 expose.tex
@@
 
-\title{Pre-Training Denoising Autoencoders for Genetic Programming
-
-(subtitle)
+\title{Pre-Trained Denoising Autoencoders Long Short-Term Memory
+Networks as probabilistic Models for Estimation of Distribution Genetic
+Programming
 
@@
 
-Johannes Gutenberg University Mainz
-
-Summerterm 2022}
+Johannes Gutenberg University Mainz}
 \author{}
-\date{\vspace{-2.5em}Date of Submission: 2022-09-09}
+\date{\vspace{-2.5em}Date of Submission: 2022-09-11}
 
@@
 
-\listoftables
-\listoffigures
-\thispagestyle{empty}
-\newpage
-
-\hypertarget{expose}{%
-\section{Expose}\label{expose}}
+\hypertarget{overview}{%
+\section{Overview}\label{overview}}
 
@@
 DAE-GP, DAE-LSTMs are trained with a subset of high-fitness solutions
-selected from the parent population as input. The resulting model is
-then used to sample new offspring solutions by propagating partially
-mutated solutions from the parent population through the DAE-LSTMS
-(Wittenberg, Rothlauf and Schweim, 2020).
+selected from a parent population with the aim to capture their
+promising qualities. The resulting model is then used to sample new
+offspring solutions by propagating partially mutated solutions from the
+parent population through the DAE-LSTMS (Wittenberg, Rothlauf and
+Schweim, 2020).
 
-In previous work DAE-GP has been demonstrated to outperform GP for both
-a generalized version of the royal tree problem (Wittenberg, Rothlauf
-and Schweim, 2020) as well as for a standard symbolic regression problem
-(Wittenberg and Rothlauf, 2022). The DAE-GP algorithm implemented for
+In previous work DAE-GP has been shown to outperform GP for both a
+generalized version of the royal tree problem (Wittenberg, Rothlauf and
+Schweim, 2020) as well as for a real-world symbolic regression problem
+(Wittenberg and Rothlauf, 2022). The DAE-GP algorithm that was used for
 both experiments initializes the DAE-LSTMs with randomized parameters
-for each generation \(g\) of the search. A possible way of improving the
-quality of solutions and/or the total runtime of DAE-GP could be to
-instead use pre-trained DAE-LSTMs that are passed on from each
-generation \(g_{t}\) to the next generation \(g_{t+1}\). Pre-Training is
-a commonly used strategy in deep architectures that has been shown to
-improve both the optimization process itself as well as the
-generalization behavior if compared to randomly initialized parameters
-(Erhan \emph{et al.}, 2009).
+for each generation \(g\) of the search. This thesis investigates a
+possible way of improving the performance of DAE-GP by using pre-trained
+DAE-LSTMs with parameters that are passed on from each generation
+\(g_{t}\) to the next generation \(g_{t+1}\). Pre-Training is a commonly
+used strategy in deep architectures that has been shown to improve both
+the optimization process itself as well as the generalization behavior
+if compared to randomly initialized parameters (Erhan \emph{et al.},
+2009).
 
@@
 
-The aim of this research project is to study the effects of using
-pre-trained DAE-LSTMs in DAE-GP on both solution quality as well as on
-total time consumption.
+This thesis studies the influence of using pre-trained DAE-LSTMs in
+DAE-GP on the quality of solutions found. Recent research published by
+Wittenberg, Rothlauf and Schweim (2020) and Wittenberg and Rothlauf
+(2022) introduced DAE-GP as a promising alternative approach to existing
+EDA-GP based metaheuristics. If it is possible to further improve
+solution quality of DAE-GP by using pre-trained DAE-LSTMs this research
+might contribute positively to future research and the practical
+adoption of DAE-GP.
 
-\ldots what problem(s)???\ldots{}
+The experiments that are conducted follow a similar research design as
+the initial publication describing DAE-GP by Wittenberg, Rothlauf and
+Schweim (2020). A generalized version of the royal tree problem will be
+used as a benchmark problem to compare the performance of standard
+DAE-GP to a modified DAE-GP algorithm that uses pre-training for the
+initialization of it's DAE-LSTMs. The royal tree problem is an
+adjustable benchmark problem for GP that allows to examine differences
+in optimization performance for varying degrees of problem complexity
+(Kinnear \emph{et al.}, 1996). Both DAE-GP algorithms will be repeatedly
+applied to three different configurations of the royal tree problem with
+varying degrees of complexity. The results of the experiment will then
+be used to test for statistical differences in the underlying
+distributions with the aim to answer the question if pre-training will
+significantly influence the quality of solutions found by DAE-GP.
 
-\hypertarget{research-design}{%
-\subsection{Research Design}\label{research-design}}
+In addition to comparing the average solution quality for both
+approaches, this thesis will also focus on exploring differences in the
+candidate solutions that are created by sampling pre-trained DAE-LSTMs
+in comparison to randomly initialized DAE-LSTMs. Here, the main
+questions of interest is, if pre-training can significantly improve the
+fitness of offspring solutions that are sampled during DAE-GP if
+compared to a DAE-GP algorithm that does not use pre-training.
 
@@
 \hypertarget{I}{%
-\section*{I - References}\label{I}}
-\addcontentsline{toc}{section}{I - References}
+\section*{References}\label{I}}
+\addcontentsline{toc}{section}{References}
 
@@
 \leavevmode\vadjust pre{\hypertarget{ref-pmlr-v5-erhan09a}{}}%
-Erhan, D. \emph{et al.} (2009) {``The difficulty of training deep
-architectures and the effect of unsupervised pre-training,''} in D. van
-Dyk and M. Welling (eds.) \emph{Proceedings of the twelth international
+Erhan, D. \emph{et al.} (2009) {`The difficulty of training deep
+architectures and the effect of unsupervised pre-training'}, in D. van
+Dyk and M. Welling (eds) \emph{Proceedings of the twelth international
 conference on artificial intelligence and statistics}. Hilton Clearwater
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-royal_tree_article}{}}%
+Kinnear, K. \emph{et al.} (1996) {`The royal tree problem, a benchmark
+for single and multi-population genetic programming'}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-dae-gp_2022_symreg}{}}%
-Wittenberg, D. and Rothlauf, F. (2022) {``Denoising autoencoder genetic
-programming for real-world symbolic regression,''} in \emph{Proceedings
+Wittenberg, D. and Rothlauf, F. (2022) {`Denoising autoencoder genetic
+programming for real-world symbolic regression'}, in \emph{Proceedings
 of the genetic and evolutionary computation conference companion}. New
@@
 \leavevmode\vadjust pre{\hypertarget{ref-dae-gp_2020_rtree}{}}%
-Wittenberg, D., Rothlauf, F. and Schweim, D. (2020) {``DAE-GP: Denoising
+Wittenberg, D., Rothlauf, F. and Schweim, D. (2020) {`DAE-GP: Denoising
 autoencoder LSTM networks as probabilistic models in estimation of
-distribution genetic programming,''} in \emph{Proceedings of the 2020
+distribution genetic programming'}, in \emph{Proceedings of the 2020
 genetic and evolutionary computation conference}. New York, NY, USA:
@@
 
-\hypertarget{II}{%
-\section*{II - Statutory Declaration}\label{II}}
-\addcontentsline{toc}{section}{II - Statutory Declaration}
-
-\includegraphics{./private/erklaerung.pdf}\\
-
 \end{document}

### 2022-12-20 paper.Rmd
Mon Dec 19 21:00:08 CET 2022: update-remote.sh
Tue Dec 20 09:00:07 CET 2022: update-remote.sh

@@
 
+![Airfoil - First Generation Median Training Error for variable number of hidden Layers](./img/airfoil_firstGen/airfoil_firstGen_mean_training_error_by_layers.png)
+
+![Airfoil - First Generation final Training Error for variable number of hidden Layers](./img/airfoil_firstGen/airfoil_firstGen_boxplot_training_error_by_layer.png)
+
 # Discussion

### 2022-12-20 paper.tex
Tue Dec 20 09:00:07 CET 2022: update-remote.sh

@@
 \newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\usepackage{longtable,booktabs,array}
-\usepackage{calc} % for calculating minipage widths
-% Correct order of tables after \paragraph or \subparagraph
-\usepackage{etoolbox}
-\makeatletter
-\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
-\makeatother
-% Allow footnotes in longtable head/foot
-\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
-\makesavenoteenv{longtable}
 \usepackage{graphicx}
@@
 
-\title{Pre-Trained Denoising Autoencoders Long Short-Term Memory Networks as probabilistic Models for Estimation of Distribution Genetic Programming
+\title{Pre-Trained Denoising Autoencoders Long Short-Term Memory
+Networks as probabilistic Models for Estimation of Distribution Genetic
+Programming
 
@@
 \author{}
-\date{\vspace{-2.5em}Date of Submission: 2022-12-18}
+\date{\vspace{-2.5em}Date of Submission: 2022-12-19}
 
@@
 
-{
-\setcounter{tocdepth}{2}
-\tableofcontents
-}
 \setcounter{page}{1}
@@
 
-Denoising Autoencoder Genetic Programming (DAE-GP) is a novel variation of an genetic programming based Estimation of Distribution Algorithm (EDA-GP) that uses a denoising autoencoders long short-term memory network (DAE-LSTMs) as a probabilistic model to sample new candidate solutions {[}18{]}.
+Denoising Autoencoder Genetic Programming (DAE-GP) is a novel variation
+of an genetic programming based Estimation of Distribution Algorithm
+(EDA-GP) that uses a denoising autoencoders long short-term memory
+network (DAE-LSTMs) as a probabilistic model to sample new candidate
+solutions {[}18{]}.
 
-DAE-LSTMs are recurrent neural networks (RNNs) that can be trained in an unsupervised learning environment to minimize a reconstruction error for encoding input data into a compressed representation and subsequently decoding the compressed representation back to the input dimension.
-In DAE-GP, DAE-LSTMs are trained with a subset of high-fitness solutions selected from a parent population with the aim to capture their promising qualities.
-The resulting model is then used to sample new offspring solutions by propagating partially mutated solutions from the parent population through the DAE-LSTMS {[}18{]}.
-In previous work DAE-GP has been shown to outperform GP for both a generalized version of the royal tree problem {[}18{]} as well as for a real-world symbolic regression problem {[}17{]}.
+DAE-LSTMs are recurrent neural networks (RNNs) that can be trained in an
+unsupervised learning environment to minimize a reconstruction error for
+encoding input data into a compressed representation and subsequently
+decoding the compressed representation back to the input dimension. In
+DAE-GP, DAE-LSTMs are trained with a subset of high-fitness solutions
+selected from a parent population with the aim to capture their
+promising qualities. The resulting model is then used to sample new
+offspring solutions by propagating partially mutated solutions from the
+parent population through the DAE-LSTMS {[}18{]}. In previous work
+DAE-GP has been shown to outperform GP for both a generalized version of
+the royal tree problem {[}18{]} as well as for a real-world symbolic
+regression problem {[}17{]}.
 
-The DAE-GP algorithm first described by {[}18{]} trains a DAE-LSTMs for each generation \(g\) of the search from scratch.
-{[}17{]} and {[}16{]} suggests the incorporation of a pre-training strategy into the evolutionary search as a possible way of improving the model quality of DAE-GP.
-The key idea is to pre-train an initial DAE-LSTM on a large population of candidate solutions and to use the weights of this initial model in each generation of the search as a starting point.
-This thesis studies the influence of using pre-trained DAE-LSTMs in DAE-GP for symbolic regression, especially looking at the influence in generalization behavior and the overall quality of solutions found by DAE-GP.
-The aim of this study is to answer the question if a pre-training strategy can help to improve generalization in DAE-GP and if so, analyze how this improved generalization ability benefits the overall performance of DAE-GP.
+The DAE-GP algorithm first described by {[}18{]} trains a DAE-LSTMs for
+each generation \(g\) of the search from scratch. {[}17{]} and {[}16{]}
+suggests the incorporation of a pre-training strategy into the
+evolutionary search as a possible way of improving the model quality of
+DAE-GP. The key idea is to pre-train an initial DAE-LSTM on a large
+population of candidate solutions and to use the weights of this initial
+model in each generation of the search as a starting point. This thesis
+studies the influence of using pre-trained DAE-LSTMs in DAE-GP for
+symbolic regression, especially looking at the influence in
+generalization behavior and the overall quality of solutions found by
+DAE-GP. The aim of this study is to answer the question if a
+pre-training strategy can help to improve generalization in DAE-GP and
+if so, analyze how this improved generalization ability benefits the
+overall performance of DAE-GP.
 
@@
 
-\emph{This section describes the relevant concepts that are necessary for the understanding and classification of DAE-GP as well as the concept of pre-training in artificial neural networks}
+\emph{This section describes the relevant concepts that are necessary
+for the understanding and classification of DAE-GP as well as the
+concept of pre-training in artificial neural networks}
 
 \hypertarget{denoising-autoencoder-genetic-programming}{%
-\subsection{Denoising Autoencoder Genetic Programming}\label{denoising-autoencoder-genetic-programming}}
+\subsection{Denoising Autoencoder Genetic
+Programming}\label{denoising-autoencoder-genetic-programming}}
 
-DAE-GP is an EDA-GP algorithm that uses DAE-LSTM networks as a probabilistic model to sample new offspring solutions {[}18{]}.
+DAE-GP is an EDA-GP algorithm that uses DAE-LSTM networks as a
+probabilistic model to sample new offspring solutions {[}18{]}.
 
 \hypertarget{evolutionary-computation}{%
-\subsubsection{Evolutionary Computation}\label{evolutionary-computation}}
+\subsubsection{Evolutionary
+Computation}\label{evolutionary-computation}}
 
-As a variant of GP, DAE-GP can be classified as a meta-heuristic that belongs to the field of evolutionary computation (EC).
-EC based meta-heuristics are optimization methods that simulate the process of Darwinian evolution to search for high quality solutions by applying selection and variation operators to a population of candidate solutions.
-Examples of EC include genetic algorithms (GA), evolutionary strategies (ES) and GP.
-In EC, the quality of a solution is commonly measured as fitness and the time steps of the search are called generations.
-Another important concept in EC is the distinction between genotypes and phenotypes of solutions, the genotype contains the information that is necessary to construct the phenotype, the outer appearance of a particular solution on which we measure the overall quality of solutions.
-The representation of a solution is therefore defined by the mapping of genotypes to phenotypes {[}12{]}.
-Genetic operators, such as mutation or recombination are usually applied to the genotype of solutions.
+As a variant of GP, DAE-GP can be classified as a meta-heuristic that
+belongs to the field of evolutionary computation (EC). EC based
+meta-heuristics are optimization methods that simulate the process of
+Darwinian evolution to search for high quality solutions by applying
+selection and variation operators to a population of candidate
+solutions. Examples of EC include genetic algorithms (GA), evolutionary
+strategies (ES) and GP. In EC, the quality of a solution is commonly
+measured as fitness and the time steps of the search are called
+generations. Another important concept in EC is the distinction between
+genotypes and phenotypes of solutions, the genotype contains the
+information that is necessary to construct the phenotype, the outer
+appearance of a particular solution on which we measure the overall
+quality of solutions. The representation of a solution is therefore
+defined by the mapping of genotypes to phenotypes {[}12{]}. Genetic
+operators, such as mutation or recombination are usually applied to the
+genotype of solutions.
 
@@
 
-GP follows the same basic evolutionary principle of EC but searches for more general, hierarchical computer programs of dynamically varying size and shape {[}8{]}.
-The computer programs that are at the center of the evolutionary search in GP are commonly represented by tree structures at the level of their phenotype {[}12{]}.
-Since GP searches for high fitness computer programs that produce a desired output for some input, it can be applied to various different problem domains such as symbolic regression, automatic programming, or evolving game-playing strategies {[}8{]}.
-An important quality of GP is the ability to search for solutions of variable length and structure.
-GP is an especially useful meta heuristic for problems where no a priori knowledge about the final form of good solutions is available.
-GPs ability to optimize solutions for their structure as well as for their parameters led to it being one of the most prevalent methods used symbolic regression {[}10{]}.
+GP follows the same basic evolutionary principle of EC but searches for
+more general, hierarchical computer programs of dynamically varying size
+and shape {[}8{]}. The computer programs that are at the center of the
+evolutionary search in GP are commonly represented by tree structures at
+the level of their phenotype {[}12{]}. Since GP searches for high
+fitness computer programs that produce a desired output for some input,
+it can be applied to various different problem domains such as symbolic
+regression, automatic programming, or evolving game-playing strategies
+{[}8{]}. An important quality of GP is the ability to search for
+solutions of variable length and structure. GP is an especially useful
+meta heuristic for problems where no a priori knowledge about the final
+form of good solutions is available. GPs ability to optimize solutions
+for their structure as well as for their parameters led to it being one
+of the most prevalent methods used symbolic regression {[}10{]}.
 
 \hypertarget{estimation-of-distribution-algorithms}{%
-\subsubsection{Estimation of Distribution Algorithms}\label{estimation-of-distribution-algorithms}}
+\subsubsection{Estimation of Distribution
+Algorithms}\label{estimation-of-distribution-algorithms}}
 
-The aim of Estimation of distribution algorithms (EDA) is to replace the standard variation operators used in GA by building probabilistic models that can capture complex dependencies between different decision variables of an optimization problem {[}12{]}.
-EDAs use this probabilistic model to sample new offspring solutions inside an evolutionary search to replace crossover and/or mutation operators.
+The aim of Estimation of distribution algorithms (EDA) is to replace the
+standard variation operators used in GA by building probabilistic models
+that can capture complex dependencies between different decision
+variables of an optimization problem {[}12{]}. EDAs use this
+probabilistic model to sample new offspring solutions inside an
+evolutionary search to replace crossover and/or mutation operators.
 
@@
 
-One possible way of model building in EDA proposed by {[}11{]} is to use denoising autoencoders (DAEs) as generative models to capture complex probability distributions of decision variables.
+One possible way of model building in EDA proposed by {[}11{]} is to use
+denoising autoencoders (DAEs) as generative models to capture complex
+probability distributions of decision variables.
 
-DAE, a variation of the autoencoders (AE), are a widely used type of neural networks in the domain of unsupervized machine learning that maps \(n\) input variables to \(n\) output variables using a hidden representation.
+DAE, a variation of the autoencoders (AE), are a widely used type of
+neural networks in the domain of unsupervized machine learning that maps
+\(n\) input variables to \(n\) output variables using a hidden
+representation.
 
-AE were introduced by {[}6{]} to compress high-dimensional data into lower-dimensions. An AE consists of two different subunits:
+AE were introduced by {[}6{]} to compress high-dimensional data into
+lower-dimensions. An AE consists of two different subunits:
 
@@
 \item
-  Decoder \(d(h)\): Decode and output the the encoded data back to its original dimension
+  Decoder \(d(h)\): Decode and output the the encoded data back to its
+  original dimension
 \end{itemize}
 
-The AE is trained to reduce the reconstruction error between input and output data, after the training procedure is finished the network is able to reduce the dimensionality of input data to a compressed representation{[}6{]}.
+The AE is trained to reduce the reconstruction error between input and
+output data, after the training procedure is finished the network is
+able to reduce the dimensionality of input data to a compressed
+representation{[}6{]}.
 
-DAE was first introduced by {[}14{]} as an improved AE with the ability to learn new representations of data that is especially robust to partially corrupted input data.
-DAE modifies the AE by using partially corrupted input data for the AE and training it to reconstruct the uncorrupted, original version of the input data.
+DAE was first introduced by {[}14{]} as an improved AE with the ability
+to learn new representations of data that is especially robust to
+partially corrupted input data. DAE modifies the AE by using partially
+corrupted input data for the AE and training it to reconstruct the
+uncorrupted, original version of the input data.
 
-Since the hidden representation of DAE captures the dependency structure of the input data it can therefore also be used to generate new solutions in the context of GAs {[}11{]}.
+Since the hidden representation of DAE captures the dependency structure
+of the input data it can therefore also be used to generate new
+solutions in the context of GAs {[}11{]}.
 
-DAE-GP builds upon the concept of using DAEs in EDAs described by {[}11{]} and transfers the concept to the domain of GP.
-The mutation and crossover operators of standard GP are replaced by sampling new solutions from a probabilistic model that is build by training a DAEs long short-term memory (LSTM) network on a subset of high fitness solutions from the current population {[}18{]}.
-LSTMs are a variant of recurrent neural networks first introduced by {[}7{]} that can store learned information over en extended time period while avoiding the problem of vanishing and/or exploding gradients.
-Since DAE-GP encodes candidate solutions as linear strings in prefix notation, the DAE in DAE-GP uses LSTMs for both encoding and decoding where the total amount of time steps \(T\) is equal to sum of the length of the input solution and the output solution {[}18{]}.
+DAE-GP builds upon the concept of using DAEs in EDAs described by
+{[}11{]} and transfers the concept to the domain of GP. The mutation and
+crossover operators of standard GP are replaced by sampling new
+solutions from a probabilistic model that is build by training a DAEs
+long short-term memory (LSTM) network on a subset of high fitness
+solutions from the current population {[}18{]}. LSTMs are a variant of
+recurrent neural networks first introduced by {[}7{]} that can store
+learned information over en extended time period while avoiding the
+problem of vanishing and/or exploding gradients. Since DAE-GP encodes
+candidate solutions as linear strings in prefix notation, the DAE in
+DAE-GP uses LSTMs for both encoding and decoding where the total amount
+of time steps \(T\) is equal to sum of the length of the input solution
+and the output solution {[}18{]}.
 
@@
 
-The strategy of pre-training artificial neural networks is based on the idea of transfer learning.
-Transfer learning in the field of machine learning aims at retaining previously learned knowledge from one task to re-use it for another task {[}5{]} {[}9{]}.
-Transfer learning traditionally follows a two phase approach:
+The strategy of pre-training artificial neural networks is based on the
+idea of transfer learning. Transfer learning in the field of machine
+learning aims at retaining previously learned knowledge from one task to
+re-use it for another task {[}5{]} {[}9{]}. Transfer learning
+traditionally follows a two phase approach:
 
@@
 
-where source task and target task are usually similar but may differ in their feature space and the distribution of training data {[}9{]}.
-One of the main motivations for using transfer learning in real world machine learning tasks is the ability to reduce the need for large amounts of training data which can often be unavailable or too expensive to collect.
+where source task and target task are usually similar but may differ in
+their feature space and the distribution of training data {[}9{]}. One
+of the main motivations for using transfer learning in real world
+machine learning tasks is the ability to reduce the need for large
+amounts of training data which can often be unavailable or too expensive
+to collect.
 
-Pre-Training is a commonly used strategy in deep architectures that has been shown to improve both the optimization process itself as well as the generalization behavior if compared to randomly initialized parameters {[}3{]}.
+Pre-Training is a commonly used strategy in deep architectures that has
+been shown to improve both the optimization process itself as well as
+the generalization behavior if compared to randomly initialized
+parameters {[}3{]}.
 
-\ldots describe different pre-training strategies, e.g.~few shot, perpetual, re-using\ldots{}
+\ldots describe different pre-training strategies, e.g.~few shot,
+perpetual, re-using\ldots{}
 
@@
 
-The pre-training strategy used in this thesis is applied to the DAE-LSTM model \(M_g\) that are used in each DAE-GP generation \(g\) for \(g\in{1,...,g_{max}}\).
-Instead of initializing and optimizing each model from scratch as done in previous work (e.g.
-{[}18{]} {[}17{]}), a separate DAE-LSTM network \(\hat{M}\) will be trained on a large population of randomly initialized solutions.\\
-The optimal parameters obtained after finishing the training procedure of \(\hat{M}\) are then used as the starting parameters for each following DAE-LSTM model \(M_g\) for \(g\in{1,...,g_{max}}\).
+The pre-training strategy used in this thesis is applied to the DAE-LSTM
+model \(M_g\) that are used in each DAE-GP generation \(g\) for
+\(g\in{1,...,g_{max}}\). Instead of initializing and optimizing each
+model from scratch as done in previous work (e.g. {[}18{]} {[}17{]}), a
+separate DAE-LSTM network \(\hat{M}\) will be trained on a large
+population of randomly initialized solutions.\\
+The optimal parameters obtained after finishing the training procedure
+of \(\hat{M}\) are then used as the starting parameters for each
+following DAE-LSTM model \(M_g\) for \(g\in{1,...,g_{max}}\).
 
-The motivation for incorporating pre-training into DAE-GP is based on the following suspected mechanisms of improvement:
+The motivation for incorporating pre-training into DAE-GP is based on
+the following suspected mechanisms of improvement:
 
@@
 \item
-  Improve overall performance of DAE-GP by improving either run time and/or solution quality
+  Improve overall performance of DAE-GP by improving either run time
+  and/or solution quality
 \item
@@
 
-The probably most obvious motivation for using a pre-training strategy in DAE-GP is based on the fact, that initializing \(M_g\) with pre-trained weights from \(\hat{M}\) is likely to reduce the amount of training epochs that are necessary until the point of training error convergence is reached.
+The probably most obvious motivation for using a pre-training strategy
+in DAE-GP is based on the fact, that initializing \(M_g\) with
+pre-trained weights from \(\hat{M}\) is likely to reduce the amount of
+training epochs that are necessary until the point of training error
+convergence is reached.
 
-Early research on pre-training for DAE by {[}3{]} comes to the conclusion that besides adding robustness to models the strategy also results in improved generalization and better performing models.
-The authors describe that unsupervised pre-training behaves like a regulizer on DAE networks, the mean test error and it's variance are reduced for DAE networks that are initialized from pre-trained weights if compared to the same architectures that use random initialization.
+Early research on pre-training for DAE by {[}3{]} comes to the
+conclusion that besides adding robustness to models the strategy also
+results in improved generalization and better performing models. The
+authors describe that unsupervised pre-training behaves like a regulizer
+on DAE networks, the mean test error and it's variance are reduced for
+DAE networks that are initialized from pre-trained weights if compared
+to the same architectures that use random initialization.
 
-Another important finding by {[}3{]} is that the positive effect of pre-training described above is dependent on both the depth of the network as well as the size of layers - while increasingly larger networks benefit increasingly more from pre-training the final performance of small architectures tends to be worse with pre-training than with randomized weight initialization.
-The authors find evidence that pre-training DAE is especially usefull for optimizing the parameters of the lower layers of the network.
+Another important finding by {[}3{]} is that the positive effect of
+pre-training described above is dependent on both the depth of the
+network as well as the size of layers - while increasingly larger
+networks benefit increasingly more from pre-training the final
+performance of small architectures tends to be worse with pre-training
+than with randomized weight initialization. The authors find evidence
+that pre-training DAE is especially usefull for optimizing the
+parameters of the lower layers of the network.
 
-Another motivation for introducing pre-training into DAE-GP is the prevalence of sampling error for small GP populations sizes.
-{[}13{]} finds that sampling error, non-systematic errors that are caused by observing only a small subset of the statistical population, is a severe problem in the domain of GP since the initial population may not be a representative sample of all possible solutions.
-{[}13{]} also introduces a method for calculating optimal population sizes to minimize the presence of sampling error.
-Since DAE-LSTM of DAE-GP learns the properties of it's training population and reuses the acquired knowledge in the sampling procedure, using the parameters obtained from training \(\hat{M}\) on a sufficiently large training population might reduce the need for large population sizes in the following generations of the search if \(\hat{M}\) already implicitly captured the properties of a representative sample of solutions.
-If this mechanism can be applied successfully it would benefit the performance of DAE-GP by increasing the population diversity (resulting in better solution quality) as well as by reducing the need for large population sizes which require computational resources.
+Another motivation for introducing pre-training into DAE-GP is the
+prevalence of sampling error for small GP populations sizes. {[}13{]}
+finds that sampling error, non-systematic errors that are caused by
+observing only a small subset of the statistical population, is a severe
+problem in the domain of GP since the initial population may not be a
+representative sample of all possible solutions. {[}13{]} also
+introduces a method for calculating optimal population sizes to minimize
+the presence of sampling error. Since DAE-LSTM of DAE-GP learns the
+properties of it's training population and reuses the acquired knowledge
+in the sampling procedure, using the parameters obtained from training
+\(\hat{M}\) on a sufficiently large training population might reduce the
+need for large population sizes in the following generations of the
+search if \(\hat{M}\) already implicitly captured the properties of a
+representative sample of solutions. If this mechanism can be applied
+successfully it would benefit the performance of DAE-GP by increasing
+the population diversity (resulting in better solution quality) as well
+as by reducing the need for large population sizes which require
+computational resources.
 
@@
 
-The DAE-GP algorithm, first described by {[}18{]}, is described by algorithm 1
+The DAE-GP algorithm, first described by {[}18{]}, is described by
+algorithm 1
 
@@
 
-One difficulty in the implementation of a pre-training strategy into DAE-GP has been the detemination of the number of hidden neurons in side the DAE-LSTMs hidden layers.
-The original description of DAE-GP {[}18{]} used a strategy where the number of hidden neurons for each hidden layer is dynamically set per generation to the maximum individual size inside the current population.
-For a pre-training implementation, this strategy can not be easily adapted since it leads to a changing number of neurons at each generation resulting in different dimensions of the DAE-LSTM.
-To allow for the sharing of weights and biases from the pre-trained model \(\hat{M}\) to each \(M_g\) for \(g\in{1,...,g_{max}}\) I experimented with two different strategies:
+One difficulty in the implementation of a pre-training strategy into
+DAE-GP has been the detemination of the number of hidden neurons in side
+the DAE-LSTMs hidden layers. The original description of DAE-GP {[}18{]}
+used a strategy where the number of hidden neurons for each hidden layer
+is dynamically set per generation to the maximum individual size inside
+the current population. For a pre-training implementation, this strategy
+can not be easily adapted since it leads to a changing number of neurons
+at each generation resulting in different dimensions of the DAE-LSTM. To
+allow for the sharing of weights and biases from the pre-trained model
+\(\hat{M}\) to each \(M_g\) for \(g\in{1,...,g_{max}}\) I experimented
+with two different strategies:
 
@@
 \item
-  Set the number of hidden neurons per hidden layer for each generation to the maximum individual size inside the pre-training population
+  Set the number of hidden neurons per hidden layer for each generation
+  to the maximum individual size inside the pre-training population
 \item
@@
 
-The framework that was used to conduct all experiments of this thesis was provided by the supervisor David Wittenberg and uses the python programming language in conjunction with the baseline libraries \texttt{Keras} {[}1{]} for deep neural networks and \texttt{deap} {[}4{]} for evolutionary computation.
+The framework that was used to conduct all experiments of this thesis
+was provided by the supervisor David Wittenberg and uses the python
+programming language in conjunction with the baseline libraries
+\texttt{Keras} {[}1{]} for deep neural networks and \texttt{deap}
+{[}4{]} for evolutionary computation.
 
@@
 
-To test pre-training in DAE-GP this thesis focuses on the domain of real-world symbolic regression problems.
-Symbolic Regression problems have been one of the first GP applications {[}8{]} and are an actively studied and highly relevant research area.
-The goal in symbolic regression is to find a mathematical model for a given set of data points {[}10{]}, in real-world symbolic regression these data points are sourced from real-world observations which in contrast to synthetic symbolic regression problem are more likely to contain random noise and bias.
-Another important challenge in solving real-world symbolic regression problems is the ability for a given model to generalize, we want the final model to show high accuracy in predicting outcomes for previously unseen cases.
+To test pre-training in DAE-GP this thesis focuses on the domain of
+real-world symbolic regression problems. Symbolic Regression problems
+have been one of the first GP applications {[}8{]} and are an actively
+studied and highly relevant research area. The goal in symbolic
+regression is to find a mathematical model for a given set of data
+points {[}10{]}, in real-world symbolic regression these data points are
+sourced from real-world observations which in contrast to synthetic
+symbolic regression problem are more likely to contain random noise and
+bias. Another important challenge in solving real-world symbolic
+regression problems is the ability for a given model to generalize, we
+want the final model to show high accuracy in predicting outcomes for
+previously unseen cases.
 
-The main experiments conducted in this thesis uses the NASA Airfoil Self-Noise Data Set which is part of the UCI machine learning repository {[}2{]}.
-The dataset consists of 5 input variables and 1 output variable that are listed in table 1.
+The main experiments conducted in this thesis uses the NASA Airfoil
+Self-Noise Data Set which is part of the UCI machine learning repository
+{[}2{]}. The dataset consists of 5 input variables and 1 output variable
+that are listed in table 1.
 
@@
 
-\caption{(\#tab:airfoil_dataset_description)Airfoil - Dataset Description}
+\caption{\label{tab:airfoil_dataset_description}Airfoil - Dataset Description}
 \centering
@@
 
-The objective of the airfoil problem is to find a function that accurately predicts the output variable \(y\) by taking in a subset of the input variables \(x1,x2,x3,x4,x5\).
-The function set used by all DAE-GP variations for the airfoil problem is summarized in table 2, the terminal set consists of the 5 input variables \(x1,x2,x3,x4,x5\) and ephemeral random integers in the range of \([{}-5,..,5]\) {[}17{]}.
+The objective of the airfoil problem is to find a function that
+accurately predicts the output variable \(y\) by taking in a subset of
+the input variables \(x1,x2,x3,x4,x5\). The function set used by all
+DAE-GP variations for the airfoil problem is summarized in table 2, the
+terminal set consists of the 5 input variables \(x1,x2,x3,x4,x5\) and
+ephemeral random integers in the range of \([{}-5,..,5]\) {[}17{]}.
 
@@
 
-\caption{(\#tab:airfoil_function_set)Airfoil - Function Set}
+\caption{\label{tab:airfoil_function_set}Airfoil - Function Set}
 \centering
@@
 
-In a first exploratory experiment DAE-GP was tested on the Aifoil dataset to perform a symbolic regression task.
-The aim of this initial experiment is to search for differences in behavior between an algorithm that uses a pre-training strategy and regular DAE-GP.
-The parameters for the experiment are summarized in table 2, the average results for fitness, number of fitness evaluations, average tree size of the population and the unique rate of the population are summarized in plot 1.
+In a first exploratory experiment DAE-GP was tested on the Aifoil
+dataset to perform a symbolic regression task. The aim of this initial
+experiment is to search for differences in behavior between an algorithm
+that uses a pre-training strategy and regular DAE-GP. The parameters for
+the experiment are summarized in table 2, the average results for
+fitness, number of fitness evaluations, average tree size of the
+population and the unique rate of the population are summarized in plot
+1.
 
@@
 
-In this first experiment, the average quality of solutions produced by the pre-trained algorithm shows to be lower than that produced by regular DAE-GP.
-Differences in fitness start to show at generation 5 where regular DAE-GP starts to reduces RMSE more efficiently than the pre-trained version.
-The gap in average fitness grows larger until the end of the evolutionary search between both algorithms.
-Observing the gap between RMSE on the training dataset and the testing dataset shows no major differences, the training error is slightly lower than the testing error for both algorithms and indicates a similar generalization behavior.
+In this first experiment, the average quality of solutions produced by
+the pre-trained algorithm shows to be lower than that produced by
+regular DAE-GP. Differences in fitness start to show at generation 5
+where regular DAE-GP starts to reduces RMSE more efficiently than the
+pre-trained version. The gap in average fitness grows larger until the
+end of the evolutionary search between both algorithms. Observing the
+gap between RMSE on the training dataset and the testing dataset shows
+no major differences, the training error is slightly lower than the
+testing error for both algorithms and indicates a similar generalization
+behavior.
 
-Another interesting metric to investigate is the population's average tree size, in general smaller solutions are preferable to their larger counterparts because of improved interpretability.
-The results show, that average tree size is reduced strongly in generations 1-5 for both algorithms and than slowly increases back.
-Differences between regular DAE-GP and the pre-trained DAE-GP start to show at generation 2, where the average population tree size is continuously lower for regular DAE-GP.
+Another interesting metric to investigate is the population's average
+tree size, in general smaller solutions are preferable to their larger
+counterparts because of improved interpretability. The results show,
+that average tree size is reduced strongly in generations 1-5 for both
+algorithms and than slowly increases back. Differences between regular
+DAE-GP and the pre-trained DAE-GP start to show at generation 2, where
+the average population tree size is continuously lower for regular
+DAE-GP.
 
-It can also be observed that the population diversity, measured by the normalized levenshtein distance, is different between both DAE-GP variants.
-Both algorithms start at a unique rate of about 0.8 in the 0.th generation and lower rapidly to an average of about 0.45 in the 7.th generation.
-In the following generations, the diversity of pre-trained DAE-GP keeps lowering steadily while regular DAE-GP slowly increases the population diversity to a normalized levenshtein distance of about 0.55.
-This could be a first explanation for the better performance of regular DAE-GP in this experiment, by creating more diverse populations over the later generations of the search the algorithm is better at exploiting the search space than the pre-trained DAE-GP algorithm.
+It can also be observed that the population diversity, measured by the
+normalized levenshtein distance, is different between both DAE-GP
+variants. Both algorithms start at a unique rate of about 0.8 in the
+0.th generation and lower rapidly to an average of about 0.45 in the
+7.th generation. In the following generations, the diversity of
+pre-trained DAE-GP keeps lowering steadily while regular DAE-GP slowly
+increases the population diversity to a normalized levenshtein distance
+of about 0.55. This could be a first explanation for the better
+performance of regular DAE-GP in this experiment, by creating more
+diverse populations over the later generations of the search the
+algorithm is better at exploiting the search space than the pre-trained
+DAE-GP algorithm.
 
@@
 
-To gather a deeper understanding about the effect of pre-training DAE-LSTM in DAE-GP a series of experiments was conducted, again using the airfoil dataset for symbolic regression, and different parameter configurations for the number of hidden layers as well as the number of hidden neurons per hidden layer.
+To gather a deeper understanding about the effect of pre-training
+DAE-LSTM in DAE-GP a series of experiments was conducted, again using
+the airfoil dataset for symbolic regression, and different parameter
+configurations for the number of hidden layers as well as the number of
+hidden neurons per hidden layer.
 
-To study generalization behavior this series of experiments is conducted using DAE-GP with only a single generation until the search terminates.
-I disregard the fitness of solutions found and focus soley on the reconstruction error that is produced during the training of each the DAE-LSTM.
-The reconstruction error is measured for two separate populations, a training population \(train\_pop\) that is used to train our DAE-LSTM as well as a hold-out validation population \(test\_pop\).
-For the pre-trained DAE-GP two additional, separate populations \(\hat{train\_pop}\) and \(\hat{test\_pop}\) are used only for pre-training.
+To study generalization behavior this series of experiments is conducted
+using DAE-GP with only a single generation until the search terminates.
+I disregard the fitness of solutions found and focus soley on the
+reconstruction error that is produced during the training of each the
+DAE-LSTM. The reconstruction error is measured for two separate
+populations, a training population \(train\_pop\) that is used to train
+our DAE-LSTM as well as a hold-out validation population \(test\_pop\).
+For the pre-trained DAE-GP two additional, separate populations
+\(\hat{train\_pop}\) and \(\hat{test\_pop}\) are used only for
+pre-training.
 
@@
 \item
-  Variable number of hidden neurons (50, 100, 200) with a single hidden layer
+  Variable number of hidden neurons (50, 100, 200) with a single hidden
+  layer
 \item
-  Fixed number of hidden neurons (100) per hidden layer with variable number of hidden layers (1, 2, 3)
+  Fixed number of hidden neurons (100) per hidden layer with variable
+  number of hidden layers (1, 2, 3)
 \end{itemize}
 
-For each configuration I tested traditional DAE-GP as well as a pre-trained DAE-GP resulting in 12 total sub experiments that were each based on 10 individual runs (total number of runs=\(120\)).
-To avoid creating biased results through the presence of sampling error inside the pre-training population (see {[}13{]}), the population size for the pre-training phase is chosen very high with \(\hat{pop\_size}=20000\) where 50\% of \(\hat{pop}\) is used for the training population \(\hat{train\_pop}\) and the remaining 50\% are used for the hold-out validation population \(\hat{test\_pop}\).
-The training of DAE-LSTM uses a fixed number of 1000 epochs until termination and uses Adam optimization for gradient descent.
-The reason for using a high amount of 1000 fixed training epochs is to deliberately force the DAE-LSTM to overfit to the training data.
+For each configuration I tested traditional DAE-GP as well as a
+pre-trained DAE-GP resulting in 12 total sub experiments that were each
+based on 10 individual runs (total number of runs=\(120\)). To avoid
+creating biased results through the presence of sampling error inside
+the pre-training population (see {[}13{]}), the population size for the
+pre-training phase is chosen very high with \(\hat{pop\_size}=20000\)
+where 50\% of \(\hat{pop}\) is used for the training population
+\(\hat{train\_pop}\) and the remaining 50\% are used for the hold-out
+validation population \(\hat{test\_pop}\). The training of DAE-LSTM uses
+a fixed number of 1000 epochs until termination and uses Adam
+optimization for gradient descent. The reason for using a high amount of
+1000 fixed training epochs is to deliberately force the DAE-LSTM to
+overfit to the training data.
 
-In general I expect that with a growing number of model parameters (either by adding more hidden layers or more hidden neurons per layer) the DAE-LSTM will be more prone to overfit to the training population \(train\_pop\) resulting in a small reconstruction error for the training population and a large one for the validation population \(train\_pop\).
-The reason for this effect is that a larger network trained over an extended period of time (without the use of strategies like early stopping), has much more potential to learn noise from the training dataset than a smaller network, which is more likely to result in worsening performance for previously unseen cases {[}15{]}.
+In general I expect that with a growing number of model parameters
+(either by adding more hidden layers or more hidden neurons per layer)
+the DAE-LSTM will be more prone to overfit to the training population
+\(train\_pop\) resulting in a small reconstruction error for the
+training population and a large one for the validation population
+\(train\_pop\). The reason for this effect is that a larger network
+trained over an extended period of time (without the use of strategies
+like early stopping), has much more potential to learn noise from the
+training dataset than a smaller network, which is more likely to result
+in worsening performance for previously unseen cases {[}15{]}.
 
-Based on the review of {[}3{]} I also expect that pre-training will have an insignificant or even negative influence on small DAE-LSTM instances while improving the networks generalization ability with growing size.
+Based on the review of {[}3{]} I also expect that pre-training will have
+an insignificant or even negative influence on small DAE-LSTM instances
+while improving the networks generalization ability with growing size.
 
@@
 \includegraphics{./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_neurons.png}
-\caption{Airfoil - First Generation Median Training Error for variable number of hidden Neurons}
+\caption{Airfoil - First Generation Median Training Error for variable
+number of hidden Neurons}
 \end{figure}
@@
 \includegraphics{./img/airfoil_firstGen/airfoil_firstGen_boxplot_training_error_by_neurons.png}
-\caption{Airfoil - First Generation final Training Error for variable number of hidden Neurons}
+\caption{Airfoil - First Generation final Training Error for variable
+number of hidden Neurons}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{./img/airfoil_firstGen/airfoil_firstGen_mean_training_error_by_layers.png}
+\caption{Airfoil - First Generation Median Training Error for variable
+number of hidden Layers}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{./img/airfoil_firstGen/airfoil_firstGen_boxplot_training_error_by_layer.png}
+\caption{Airfoil - First Generation final Training Error for variable
+number of hidden Layers}
 \end{figure}
@@
 \hypertarget{limitations-and-open-questions}{%
-\section{Limitations and open Questions}\label{limitations-and-open-questions}}
+\section{Limitations and open
+Questions}\label{limitations-and-open-questions}}
 
@@
 
-{[}3{]}: Pre-Training effect especially usefull for lower-level layers -\textgreater{} Only adapt weights from low layers or embedd the pre-trained model?
+{[}3{]}: Pre-Training effect especially usefull for lower-level layers
+-\textgreater{} Only adapt weights from low layers or embedd the
+pre-trained model?
 
@@
 \CSLLeftMargin{{[}1{]} }%
-\CSLRightInline{Francois Chollet and others. 2015. Keras. Retrieved from \url{https://github.com/fchollet/keras}}
+\CSLRightInline{Francois Chollet and others. 2015. Keras. Retrieved from
+\url{https://github.com/fchollet/keras}}
 
@@
 \CSLLeftMargin{{[}2{]} }%
-\CSLRightInline{Dheeru Dua and Casey Graff. 2017. {UCI} machine learning repository. Retrieved from \url{http://archive.ics.uci.edu/ml}}
+\CSLRightInline{Dheeru Dua and Casey Graff. 2017. {UCI} machine learning
+repository. Retrieved from \url{http://archive.ics.uci.edu/ml}}
 
@@
 \CSLLeftMargin{{[}3{]} }%
-\CSLRightInline{Dumitru Erhan, Pierre-Antoine Manzagol, Yoshua Bengio, Samy Bengio, and Pascal Vincent.
-2009.
-The difficulty of training deep architectures and the effect of unsupervised pre-training.
-In \emph{Proceedings of the twelth international conference on artificial intelligence and statistics} (Proceedings of machine learning research), PMLR, Hilton Clearwater Beach Resort, Clearwater Beach, Florida USA, 153--160.
-Retrieved from \url{https://proceedings.mlr.press/v5/erhan09a.html}}
+\CSLRightInline{Dumitru Erhan, Pierre-Antoine Manzagol, Yoshua Bengio,
+Samy Bengio, and Pascal Vincent. 2009. The difficulty of training deep
+architectures and the effect of unsupervised pre-training. In
+\emph{Proceedings of the twelth international conference on artificial
+intelligence and statistics} (Proceedings of machine learning research),
+PMLR, Hilton Clearwater Beach Resort, Clearwater Beach, Florida USA,
+153--160. Retrieved from
+\url{https://proceedings.mlr.press/v5/erhan09a.html}}
 
@@
 \CSLLeftMargin{{[}4{]} }%
-\CSLRightInline{Félix-Antoine Fortin, François-Michel De Rainville, Marc-André Gardner, Marc Parizeau, and Christian Gagné.
-2012.
-{DEAP}: Evolutionary algorithms made easy.
-\emph{Journal of Machine Learning Research} 13, (2012), 2171--2175.}
+\CSLRightInline{Félix-Antoine Fortin, François-Michel De Rainville,
+Marc-André Gardner, Marc Parizeau, and Christian Gagné. 2012. {DEAP}:
+Evolutionary algorithms made easy. \emph{Journal of Machine Learning
+Research} 13, (2012), 2171--2175.}
 
@@
 \CSLLeftMargin{{[}5{]} }%
-\CSLRightInline{Xu Han, Zhengyan Zhang, Ning Ding, Yuxian Gu, Xiao Liu, Yuqi Huo, Jiezhong Qiu, Yuan Yao, Ao Zhang, Liang Zhang, Wentao Han, Minlie Huang, Qin Jin, Yanyan Lan, Yang Liu, Zhiyuan Liu, Zhiwu Lu, Xipeng Qiu, Ruihua Song, Jie Tang, Ji-Rong Wen, Jinhui Yuan, Wayne Xin Zhao, and Jun Zhu.
-2021.
-Pre-trained models: Past, present and future.
+\CSLRightInline{Xu Han, Zhengyan Zhang, Ning Ding, Yuxian Gu, Xiao Liu,
+Yuqi Huo, Jiezhong Qiu, Yuan Yao, Ao Zhang, Liang Zhang, Wentao Han,
+Minlie Huang, Qin Jin, Yanyan Lan, Yang Liu, Zhiyuan Liu, Zhiwu Lu,
+Xipeng Qiu, Ruihua Song, Jie Tang, Ji-Rong Wen, Jinhui Yuan, Wayne Xin
+Zhao, and Jun Zhu. 2021. Pre-trained models: Past, present and future.
 \emph{AI Open} 2, (2021), 225--250.
@@
 \CSLLeftMargin{{[}6{]} }%
-\CSLRightInline{G.
-E.
-Hinton and R.
-R.
-Salakhutdinov.
-2006.
-Reducing the dimensionality of data with neural networks.
-\emph{Science} 313, 5786 (2006), 504--507.
+\CSLRightInline{G. E. Hinton and R. R. Salakhutdinov. 2006. Reducing the
+dimensionality of data with neural networks. \emph{Science} 313, 5786
+(2006), 504--507.
 DOI:https://doi.org/\href{https://doi.org/10.1126/science.1127647}{10.1126/science.1127647}}
@@
 \CSLLeftMargin{{[}7{]} }%
-\CSLRightInline{Sepp Hochreiter and Jürgen Schmidhuber.
-1997.
-Long short-term memory.
-\emph{Neural computation} 9, (December 1997), 1735--80.
+\CSLRightInline{Sepp Hochreiter and Jürgen Schmidhuber. 1997. Long
+short-term memory. \emph{Neural computation} 9, (December 1997),
+1735--80.
 DOI:https://doi.org/\href{https://doi.org/10.1162/neco.1997.9.8.1735}{10.1162/neco.1997.9.8.1735}}
@@
 \CSLLeftMargin{{[}8{]} }%
-\CSLRightInline{John R. Koza. 1993. Genetic programming - on the programming of computers by means of natural selection. In \emph{Complex adaptive systems}.}
+\CSLRightInline{John R. Koza. 1993. Genetic programming - on the
+programming of computers by means of natural selection. In \emph{Complex
+adaptive systems}.}
 
@@
 \CSLLeftMargin{{[}9{]} }%
-\CSLRightInline{Sinno Jialin Pan and Qiang Yang.
-2010.
-A survey on transfer learning.
-\emph{IEEE Transactions on Knowledge and Data Engineering} 22, 10 (2010), 1345--1359.
+\CSLRightInline{Sinno Jialin Pan and Qiang Yang. 2010. A survey on
+transfer learning. \emph{IEEE Transactions on Knowledge and Data
+Engineering} 22, 10 (2010), 1345--1359.
 DOI:https://doi.org/\href{https://doi.org/10.1109/TKDE.2009.191}{10.1109/TKDE.2009.191}}
@@
 \CSLRightInline{Grégory Paris, Denis Robilliard, and Cyril Fonlupt.
-2004.
-Exploring overfitting in genetic programming.
-In \emph{Artificial evolution}, Springer Berlin Heidelberg, Berlin, Heidelberg, 267--277.}
+2004. Exploring overfitting in genetic programming. In \emph{Artificial
+evolution}, Springer Berlin Heidelberg, Berlin, Heidelberg, 267--277.}
 
@@
 \CSLLeftMargin{{[}11{]} }%
-\CSLRightInline{Malte Probst and Franz Rothlauf.
-2020.
-Harmless overfitting: Using denoising autoencoders in estimation of distribution algorithms.
-\emph{Journal of Machine Learning Research} 21, 78 (2020), 1--31.
-Retrieved from \url{http://jmlr.org/papers/v21/16-543.html}}
+\CSLRightInline{Malte Probst and Franz Rothlauf. 2020. Harmless
+overfitting: Using denoising autoencoders in estimation of distribution
+algorithms. \emph{Journal of Machine Learning Research} 21, 78 (2020),
+1--31. Retrieved from \url{http://jmlr.org/papers/v21/16-543.html}}
 
@@
 \CSLLeftMargin{{[}12{]} }%
-\CSLRightInline{Franz Rothlauf. 2011. \emph{Design of modern heuristics: Principles and application}. DOI:https://doi.org/\href{https://doi.org/10.1007/978-3-540-72962-4}{10.1007/978-3-540-72962-4}}
+\CSLRightInline{Franz Rothlauf. 2011. \emph{Design of modern heuristics:
+Principles and application}.
+DOI:https://doi.org/\href{https://doi.org/10.1007/978-3-540-72962-4}{10.1007/978-3-540-72962-4}}
 
@@
 \CSLRightInline{Dirk Schweim, David Wittenberg, and Franz Rothlauf.
-2021.
-On sampling error in genetic programming.
-\emph{Natural computing} 2021, (2021).
+2021. On sampling error in genetic programming. \emph{Natural computing}
+2021, (2021).
 DOI:https://doi.org/\url{http://doi.org/10.25358/openscience-5820}}
@@
 \CSLLeftMargin{{[}14{]} }%
-\CSLRightInline{Pascal Vincent, Hugo Larochelle, Yoshua Bengio, and Pierre-Antoine Manzagol.
-2008.
-Extracting and composing robust features with denoising autoencoders.
-In \emph{Proceedings of the 25th international conference on machine learning} (ICML '08), Association for Computing Machinery, New York, NY, USA, 1096--1103.
+\CSLRightInline{Pascal Vincent, Hugo Larochelle, Yoshua Bengio, and
+Pierre-Antoine Manzagol. 2008. Extracting and composing robust features
+with denoising autoencoders. In \emph{Proceedings of the 25th
+international conference on machine learning} (ICML '08), Association
+for Computing Machinery, New York, NY, USA, 1096--1103.
 DOI:https://doi.org/\href{https://doi.org/10.1145/1390156.1390294}{10.1145/1390156.1390294}}
@@
 \CSLLeftMargin{{[}15{]} }%
-\CSLRightInline{Andreas Weigend. 1994. On overfitting and the effective number of hidden units. In \emph{Proceedings of the 1993 connectionist models summer school}, 335--342.}
+\CSLRightInline{Andreas Weigend. 1994. On overfitting and the effective
+number of hidden units. In \emph{Proceedings of the 1993 connectionist
+models summer school}, 335--342.}
 
@@
 \CSLLeftMargin{{[}16{]} }%
-\CSLRightInline{David Wittenberg.
-2022.
-Using denoising autoencoder genetic programming to~control exploration and~exploitation in~search.
-In \emph{Genetic programming}, Springer International Publishing, Cham, 102--117.}
+\CSLRightInline{David Wittenberg. 2022. Using denoising autoencoder
+genetic programming to~control exploration and~exploitation in~search.
+In \emph{Genetic programming}, Springer International Publishing, Cham,
+102--117.}
 
@@
 \CSLLeftMargin{{[}17{]} }%
-\CSLRightInline{David Wittenberg and Franz Rothlauf.
-2022.
-Denoising autoencoder genetic programming for real-world symbolic regression.
-In \emph{Proceedings of the genetic and evolutionary computation conference companion} (GECCO '22), Association for Computing Machinery, New York, NY, USA, 612--614.
+\CSLRightInline{David Wittenberg and Franz Rothlauf. 2022. Denoising
+autoencoder genetic programming for real-world symbolic regression. In
+\emph{Proceedings of the genetic and evolutionary computation conference
+companion} (GECCO '22), Association for Computing Machinery, New York,
+NY, USA, 612--614.
 DOI:https://doi.org/\href{https://doi.org/10.1145/3520304.3528921}{10.1145/3520304.3528921}}
@@
 \CSLRightInline{David Wittenberg, Franz Rothlauf, and Dirk Schweim.
-2020.
-DAE-GP: Denoising autoencoder LSTM networks as probabilistic models in estimation of distribution genetic programming.
-In \emph{Proceedings of the 2020 genetic and evolutionary computation conference} (GECCO '20), Association for Computing Machinery, New York, NY, USA, 1037--1045.
+2020. DAE-GP: Denoising autoencoder LSTM networks as probabilistic
+models in estimation of distribution genetic programming. In
+\emph{Proceedings of the 2020 genetic and evolutionary computation
+conference} (GECCO '20), Association for Computing Machinery, New York,
+NY, USA, 1037--1045.
 DOI:https://doi.org/\href{https://doi.org/10.1145/3377930.3390180}{10.1145/3377930.3390180}}

### 2022-12-23 paper.Rmd
Fri Dec 23 22:00:06 CET 2022: update-remote.sh

@@
 
+<!---
 ## Exploratory Phase
@@
 This could be a first explanation for the better performance of regular DAE-GP in this experiment, by creating more diverse populations over the later generations of the search the algorithm is better at exploiting the search space than the pre-trained DAE-GP algorithm.
+-->
 
@@
 
-To gather a deeper understanding about the effect of pre-training DAE-LSTM in DAE-GP a series of experiments was conducted, again using the airfoil dataset for symbolic regression, and different parameter configurations for the number of hidden layers as well as the number of hidden neurons per hidden layer.
+To gather a deeper understanding about the effect of pre-training DAE-LSTM in DAE-GP a series of experiments was conducted using the airfoil dataset for symbolic regression while using and different parameter configurations for the number of hidden layers as well as the number of hidden neurons per hidden layer.
 
@@
 
+## Effect on search performance
+
+After closely examining the effect of pre-training on DAE-GPs generalization behaviour another series of experiments is conducted to study how pre-training influences the overall search behaviour of DAE-GP.
+Again the Airfoil problem was selected as a first benchmark problem to test pre-trained DAE-GP against traditional DAE-GP [@dae-gp_2022_symreg], the parameters are described in table $A$
+
+```{r airfoil_fullRun_2hl_maxIndSize_params, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
+fig.pos="H"
+knitr::kable(
+  read.csv("./tables/airfoil_fullRun_2hl_maxIndSize_params"),
+  caption="Airfoil - Parameter for full Run"
+)%>% row_spec(0,bold=TRUE ) %>% kableExtra::kable_styling(latex_options = "hold_position")
+```
+
+Regarding the hidden neurons, the dimension of each hidden layer, I used two different strategies: Regular DAE-GP uses the same strategy as described by earlier work (see [@dae-gp_2022_symreg] or [@dae-gp_2020_rtree]), the number of hidden neurons is dynamically set at each generation to the maximum length of all solutions inside the current training population.
+For pre-trained DAE-GP another strategy had to be used, we use a fixed number of hidden neurons for each run that is set equal to the maximum length of all individuals that are present in the initial pre-training population.
+I decided to take this aproach instead of using a fixed number of hidden neurons for the fact that it simplifies the adjustment of the parameter for later experiments on different problems.
+It should nonetheless be mentioned that this strategy does give pre-training on average more hidden neurons per layer resulting in a more complex network (since the pre-training population is much larger than the regular population, the chances for larger individuals inside it are also higher).
+
+![Airfoil - Best Fitness over 30 Generations](./img/airfoil_FullRun_30gens/mean_median_fitness_byGens.png)
+
+![Airfoil - Fitness Distribution Boxplot](./img/airfoil_FullRun_30gens/final_fit_boxplot.png)
+
+![Airfoil - Average Population Diversity](./img/airfoil_FullRun_30gens/mean_popDiversity_byGens.png)
+Since @dae-gp_2022_symreg finds, that DAE-GP finds higher quality and smaller sized solutions smaller than average GP for a given number of 10.000 the airfoil problem
+
+![Airfoil - Average Solution Size](./img/airfoil_FullRun_30gens/mean_Size_byGens.png)
+
 # Discussion

### 2022-12-23 paper.tex
Fri Dec 23 22:00:06 CET 2022: update-remote.sh

@@
 \author{}
-\date{\vspace{-2.5em}Date of Submission: 2022-12-19}
+\date{\vspace{-2.5em}Date of Submission: 2022-12-23}
 
@@
 
-\hypertarget{exploratory-phase}{%
-\subsection{Exploratory Phase}\label{exploratory-phase}}
-
-In a first exploratory experiment DAE-GP was tested on the Aifoil
-dataset to perform a symbolic regression task. The aim of this initial
-experiment is to search for differences in behavior between an algorithm
-that uses a pre-training strategy and regular DAE-GP. The parameters for
-the experiment are summarized in table 2, the average results for
-fitness, number of fitness evaluations, average tree size of the
-population and the unique rate of the population are summarized in plot
-1.
-
-\begin{table}[!h]
-
-\caption{\label{tab:unnamed-chunk-1}Airfoil - Configuration}
-\centering
-\begin{tabular}[t]{l|l}
-\hline
-\textbf{Hyperparameter} & \textbf{Value}\\
-\hline
-Training/Testing Split & 50:50\\
-\hline
-Pre-Training Population Size & 10000\\
-\hline
-Number of Hidden Layers & 1\\
-\hline
-Selection & Tournament\\
-\hline
-Tournament Size & 7\\
-\hline
-\end{tabular}
-\end{table}
-
-\begin{figure}
-\centering
-\includegraphics{./img/airfoil_exploratory/airfoil_exploratory.png}
-\caption{Airfoil - First Results}
-\end{figure}
-
-In this first experiment, the average quality of solutions produced by
-the pre-trained algorithm shows to be lower than that produced by
-regular DAE-GP. Differences in fitness start to show at generation 5
-where regular DAE-GP starts to reduces RMSE more efficiently than the
-pre-trained version. The gap in average fitness grows larger until the
-end of the evolutionary search between both algorithms. Observing the
-gap between RMSE on the training dataset and the testing dataset shows
-no major differences, the training error is slightly lower than the
-testing error for both algorithms and indicates a similar generalization
-behavior.
-
-Another interesting metric to investigate is the population's average
-tree size, in general smaller solutions are preferable to their larger
-counterparts because of improved interpretability. The results show,
-that average tree size is reduced strongly in generations 1-5 for both
-algorithms and than slowly increases back. Differences between regular
-DAE-GP and the pre-trained DAE-GP start to show at generation 2, where
-the average population tree size is continuously lower for regular
-DAE-GP.
-
-It can also be observed that the population diversity, measured by the
-normalized levenshtein distance, is different between both DAE-GP
-variants. Both algorithms start at a unique rate of about 0.8 in the
-0.th generation and lower rapidly to an average of about 0.45 in the
-7.th generation. In the following generations, the diversity of
-pre-trained DAE-GP keeps lowering steadily while regular DAE-GP slowly
-increases the population diversity to a normalized levenshtein distance
-of about 0.55. This could be a first explanation for the better
-performance of regular DAE-GP in this experiment, by creating more
-diverse populations over the later generations of the search the
-algorithm is better at exploiting the search space than the pre-trained
-DAE-GP algorithm.
-
 \hypertarget{effect-on-generalisation}{%
@@
 To gather a deeper understanding about the effect of pre-training
-DAE-LSTM in DAE-GP a series of experiments was conducted, again using
-the airfoil dataset for symbolic regression, and different parameter
-configurations for the number of hidden layers as well as the number of
-hidden neurons per hidden layer.
+DAE-LSTM in DAE-GP a series of experiments was conducted using the
+airfoil dataset for symbolic regression while using and different
+parameter configurations for the number of hidden layers as well as the
+number of hidden neurons per hidden layer.
 
@@
 \end{figure}
+
+\hypertarget{effect-on-search-performance}{%
+\subsection{Effect on search
+performance}\label{effect-on-search-performance}}
+
+After closely examining the effect of pre-training on DAE-GPs
+generalization behaviour another series of experiments is conducted to
+study how pre-training influences the overall search behaviour of
+DAE-GP. Again the Airfoil problem was selected as a first benchmark
+problem to test pre-trained DAE-GP against traditional DAE-GP
+{[}\textbf{gp\_2022\_symreg?}{]}, the parameters are described in table
+\(A\)
+
+\begin{table}[!h]
+
+\caption{\label{tab:airfoil_fullRun_2hl_maxIndSize_params}Airfoil - Parameter for full Run}
+\centering
+\begin{tabular}[t]{l|l}
+\hline
+\textbf{parameter} & \textbf{value}\\
+\hline
+populationSize & 500\\
+\hline
+generations & 30\\
+\hline
+fitness & RMSE\\
+\hline
+TrainingMode & Convergence\\
+\hline
+SamplingSteps & 2\\
+\hline
+hiddenLayers & 2\\
+\hline
+Selection & Binary Tournament Selection\\
+\hline
+Pre-Training PopulationSize(Training/Validation) & 10000/100000\\
+\hline
+Pre-Training TrainingMode & Early Stopping\\
+\hline
+\end{tabular}
+\end{table}
+
+Regarding the hidden neurons, the dimension of each hidden layer, I used
+two different strategies: Regular DAE-GP uses the same strategy as
+described by earlier work (see {[}\textbf{gp\_2022\_symreg?}{]} or
+{[}18{]}), the number of hidden neurons is dynamically set at each
+generation to the maximum length of all solutions inside the current
+training population. For pre-trained DAE-GP another strategy had to be
+used, we use a fixed number of hidden neurons for each run that is set
+equal to the maximum length of all individuals that are present in the
+initial pre-training population.
 

### 2022-12-23 paper.Rmd
autogit

@@
 ![Airfoil - Average Population Diversity](./img/airfoil_FullRun_30gens/mean_popDiversity_byGens.png)
-Since @dae-gp_2022_symreg finds, that DAE-GP finds higher quality and smaller sized solutions smaller than average GP for a given number of 10.000 the airfoil problem
+Since @dae-gp_2022_symreg shows, that DAE-GP finds higher quality that are also smaller in size than those found by average GP for a given number of 10.000 fitness evaluations, I also studied the differences in average solution size between pre-trained and regular DAE-GP.
 

### 2022-12-23 paper.tex
autogit

@@
 \newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
-\usepackage{float}
-\usepackage[numbers]{natbib}
-\let\origfigure\figure
-\let\endorigfigure\endfigure
-\renewenvironment{figure}[1][2] {
-    \expandafter\origfigure\expandafter[H]
-} {
-    \endorigfigure
-}
 \usepackage{float}
@@
 DAE-GP. Again the Airfoil problem was selected as a first benchmark
-problem to test pre-trained DAE-GP against traditional DAE-GP
-{[}\textbf{gp\_2022\_symreg?}{]}, the parameters are described in table
-\(A\)
+problem to test pre-trained DAE-GP against traditional DAE-GP {[}17{]},
+the parameters are described in table \(A\)
 
@@
 two different strategies: Regular DAE-GP uses the same strategy as
-described by earlier work (see {[}\textbf{gp\_2022\_symreg?}{]} or
-{[}18{]}), the number of hidden neurons is dynamically set at each
-generation to the maximum length of all solutions inside the current
-training population. For pre-trained DAE-GP another strategy had to be
-used, we use a fixed number of hidden neurons for each run that is set
-equal to the maximum length of all individuals that are present in the
-initial pre-training population.
+described by earlier work (see {[}17{]} or {[}18{]}), the number of
+hidden neurons is dynamically set at each generation to the maximum
+length of all solutions inside the current training population. For
+pre-trained DAE-GP another strategy had to be used, we use a fixed
+number of hidden neurons for each run that is set equal to the maximum
+length of all individuals that are present in the initial pre-training
+population. I decided to take this aproach instead of using a fixed
+number of hidden neurons for the fact that it simplifies the adjustment
+of the parameter for later experiments on different problems. It should
+nonetheless be mentioned that this strategy does give pre-training on
+average more hidden neurons per layer resulting in a more complex
+network (since the pre-training population is much larger than the
+regular population, the chances for larger individuals inside it are
+also higher).
+
+\begin{figure}
+\centering
+\includegraphics{./img/airfoil_FullRun_30gens/mean_median_fitness_byGens.png}
+\caption{Airfoil - Best Fitness over 30 Generations}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{./img/airfoil_FullRun_30gens/final_fit_boxplot.png}
+\caption{Airfoil - Fitness Distribution Boxplot}
+\end{figure}
+
+\includegraphics{./img/airfoil_FullRun_30gens/mean_popDiversity_byGens.png}
+Since {[}17{]} shows, that DAE-GP finds higher quality that are also
+smaller in size than those found by average GP for a given number of
+10.000 fitness evaluations, I also studied the differences in average
+solution size between pre-trained and regular DAE-GP.
+
+\begin{figure}
+\centering
+\includegraphics{./img/airfoil_FullRun_30gens/mean_Size_byGens.png}
+\caption{Airfoil - Average Solution Size}
+\end{figure}
 

### 2022-12-24 paper.Rmd
Sat Dec 24 00:00:06 CET 2022: update-remote.sh
Sat Dec 24 01:00:07 CET 2022: update-remote.sh

@@
 
-EXPERIMENT_DIR<-("../eda-gp-2020/experiments")
-
 knitr::opts_chunk$set(echo = TRUE)
@@
 
+reticulate::use_python("/Users/rmn/miniconda3/envs/dataScience/bin/python")
+
 ```
 
-```{python pytest, , echo=FALSE, include=FALSE}
+```{python pytest, echo=FALSE}
 from sys import version, executable
@@
 
+````{python3}
+import json
+
+S = json.load(open("./data/airfoil_FullRun_30gens/MWU-BestFitnessByGen.json", "r", encoding="utf-8"))
+statistic_mwu_fitness = S["statistic"]
+pval_mwu_fitness = S["p-value"]
+
+print(pval_mwu_fitness)
+````
+
 ![Airfoil - Fitness Distribution Boxplot](./img/airfoil_FullRun_30gens/final_fit_boxplot.png)

### 2022-12-27 paper.Rmd
Tue Dec 27 03:05:17 CET 2022: update-remote.sh

@@
 
-````{python3}
+```{python include=FALSE}
 import json
@@
 statistic_mwu_fitness = S["statistic"]
-pval_mwu_fitness = S["p-value"]
+pval_mwu_fitness = round(S["p-value"], 4)
 
-print(pval_mwu_fitness)
 ````
+The result of the MWU test is `r py$pval_mwu_fitness`!
 

### 2022-12-28 paper.Rmd
Wed Dec 28 10:00:08 CET 2022: update-remote.sh

@@
 
+```
+
+```{bash ipythonRunner}
+# Find all *.ipynb files in subdirectories recursively
+for file in $(find . -type f -name "*.ipynb"); do
+  # Execute the IPython notebook file using the jupyter nbconvert command
+  jupyter nbconvert --execute "${file}"
+done
 ```
@@
 Regarding the hidden neurons, the dimension of each hidden layer, I used two different strategies: Regular DAE-GP uses the same strategy as described by earlier work (see [@dae-gp_2022_symreg] or [@dae-gp_2020_rtree]), the number of hidden neurons is dynamically set at each generation to the maximum length of all solutions inside the current training population.
-For pre-trained DAE-GP another strategy had to be used, we use a fixed number of hidden neurons for each run that is set equal to the maximum length of all individuals that are present in the initial pre-training population.
+For pre-trained DAE-GP another strategy had to be used, I used a fixed number of hidden neurons for each run that is set equal to the maximum length of all individuals that are present in the initial pre-training population.
 I decided to take this aproach instead of using a fixed number of hidden neurons for the fact that it simplifies the adjustment of the parameter for later experiments on different problems.
@@
 
-```{python include=FALSE}
-import json
+```{python load_MWU_ include=FALSE}
+from json import load
 
-S = json.load(open("./data/airfoil_FullRun_30gens/MWU-BestFitnessByGen.json", "r", encoding="utf-8"))
+S = load(open("./data/airfoil_FullRun_30gens/MWU-BestFitnessByGen.json", "r", encoding="utf-8"))
 statistic_mwu_fitness = S["statistic"]
 pval_mwu_fitness = round(S["p-value"], 4)
+print(pval_mwu_fitness)
 

### 2022-12-28 paper.Rmd
autogit

@@
 
-```{bash ipythonRunner}
-# Find all *.ipynb files in subdirectories recursively
-for file in $(find . -type f -name "*.ipynb"); do
+```{bash ipythonRunner, include=FALSE}
+
+# Set the base directory
+EXPDIR="/Users/rmn/masterThesis/eda-gp-2020"
+
+# Find all *.ipynb files in subdirectories of EXPDIR, including sub-sub directories, etc., but exclude files that contain the word "checkpoint" in the filename
+for file in $(find "$EXPDIR" -type f -name "*.ipynb" ! -name "*checkpoint*"); do
+  # Change to the directory containing the IPython notebook file
+  cd "$(dirname "$file")"
   # Execute the IPython notebook file using the jupyter nbconvert command
-  jupyter nbconvert --execute "${file}"
+  ipython "$(basename "$file")"
 done
+
 ```
@@
 
-The strategy of pre-training artificial neural networks is based on the idea of transfer learning. 
-Transfer learning in the field of machine learning aims at retaining previously learned knowledge from one task to re-use it for another task [@HAN2021225] [@survey_transfer_learning].
+Pre-Training describes the concept of initially training a neural network on a large dataset before using it to solve a more specific task.
+It is a commonly used strategy in deep architectures that has been shown to improve both the optimization process itself as well as the generalization behavior if compared to the standard approach of using randomly initialized parameters [@pmlr-v5-erhan09a].
+
+The strategy of pre-training artificial neural networks is based on the idea of transfer learning that aims at retaining previously learned knowledge from one task to re-use it for another task [@HAN2021225] [@survey_transfer_learning].
 Transfer learning traditionally follows a two phase approach:
@@
 where source task and target task are usually similar but may differ in their feature space and the distribution of training data [@survey_transfer_learning].
-One of the main motivations for using transfer learning in real world machine learning tasks is the ability to reduce the need for large amounts of training data which can often be unavailable or too expensive to collect.
 
-Pre-Training is a commonly used strategy in deep architectures that has been shown to improve both the optimization process itself as well as the generalization behavior if compared to randomly initialized parameters [@pmlr-v5-erhan09a].
+Some of the main motivations for using pre-training is the ability to reduce the need for large amounts of training data which can often be unavailable or too expensive to collect as well as an improved performance by either reducing the computational effort that is necessary to train a model to solve a specific task or by improving the models generalisation ability.
 

### 2022-12-28 paper.Rmd
Wed Dec 28 21:00:08 CET 2022: update-remote.sh

@@
   <center>Pre-Trained Denoising Autoencoders Long Short-Term Memory Networks as probabilistic Models for Estimation of Distribution Genetic Programming</center>
+  <center>Master Thesis - M.Sc. Business Education</center>
   <center>--------------------------------------------------------</center>
@@
   pdf_document:
-    keep_tex: true
+    # keep_tex: true
     extra_dependencies: ["float"]

### 2022-12-28 slides.Rmd
Wed Dec 28 21:00:08 CET 2022: update-remote.sh

---
title: "Pre-Trained Denoising Autoencoders Long Short-Term Memory Networks as probabilistic Models for Estimation of Distribution Genetic Programming"
subtitle: "Kolloquium zur Masterarbeit im M.Sc. Wirtschaftspädagogik "
author: "Roman Hoehn"
institute: "JGU Mainz "
date: "Datum: `r Sys.Date()`"
output:
  beamer_presentation:
    theme: "Madrid"
    fonttheme: "structurebold"
    fig_width: 12
    fig_height: 10
    # keep_tex: true
    fig_caption: yes
    number_sections: true
    slide_level: 2
toc: yes
bibliography: ref/ref.bib
csl: csl/association-for-computing-machinery.csl
includes:
      in_header: [header_files/slides.tex]
---

```{r setup, include=FALSE}
library(magrittr)
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

## Research Question

## Denoising Autoencoder Genetic Programming

## Pre-Training

# References {.allowframebreaks}

### 2022-12-29 paper.Rmd
Wed Dec 28 22:00:05 CET 2022: update-remote.sh
Thu Dec 29 20:00:08 CET 2022: update-remote.sh

@@
 ```{bash ipythonRunner, include=FALSE}
-
 # Set the base directory
@@
 The mutation and crossover operators of standard GP are replaced by sampling new solutions from a probabilistic model that is build by training a DAEs long short-term memory (LSTM) network on a subset of high fitness solutions from the current population [@dae-gp_2020_rtree].
+
 LSTMs are a variant of recurrent neural networks first introduced by @lstm_orig that can store learned information over en extended time period while avoiding the problem of vanishing and/or exploding gradients.
@@
 
-The DAE-GP algorithm, first described by @dae-gp_2020_rtree, is described by algorithm 1
+The DAE-GP algorithm, first described by @dae-gp_2020_rtree, is summarized by chart 1.
 
-```{.numberLines}
-WHILE NOT termination_criterion
-  DO 
-END WHILE
-RETURN
-```
+![DAE-GP - Flowchart](./img/flowcharts/dae-gp.png)
 
@@
 
-```{python load_MWU_ include=FALSE}
+```{python, load_MWU, include=FALSE}
 from json import load
@@
 print(pval_mwu_fitness)
+```
 
-````
-The result of the MWU test is `r py$pval_mwu_fitness`!
+The result of the MWU test is `r py$pval_mwu_fitness`
 

### 2022-12-29 paper.Rmd
autogit

@@
 
-```{bash ipythonRunner, include=FALSE}
+```{bash compile_mermaid, include=FALSE}
+# Loop through all files with the .mmd extension in the current directory
+
+for file in ./mermaid/*.mmd
+do
+  # Call PROG on the current file, saving the output as a .png file
+  mmdc -i "$file" -o ./img/flowcharts/$(basename "$file" .mmd).png -t light -b transparent 
+done
+
+````
+
+```{bash ipythonRunner, eval=FALSE, include=FALSE}
 # Set the base directory
@@
 
-```{python pytest, echo=FALSE}
+```{python pytest, echo=FALSE, include=FALSE}
 from sys import version, executable
@@
 
-The DAE-GP algorithm, first described by @dae-gp_2020_rtree, is summarized by chart 1.
+The DAE-GP algorithm, first described by @dae-gp_2020_rtree, is summarized by figure 1.
 
-![DAE-GP - Flowchart](./img/flowcharts/dae-gp.png)
+![DAE-GP - Algorithm](./img/flowcharts/dae-gp.png)
+
+The pre-training strategy implemented is visualized by figure 2.
+
+![DAE-GP - Algorithm](./img/flowcharts/pt-dae-gp.png)
 

### 2022-12-29 paper.Rmd
Thu Dec 29 22:00:06 CET 2022: update-remote.sh

@@
 ```{bash compile_mermaid, include=FALSE}
-# Loop through all files with the .mmd extension in the current directory
+# compile all *.mmd files in the mermaid directory
+
+WIDTH=1600
+HEIGHT=1200
+THEME="neutral"
+BACKGROUND="transparent"
 
@@
 do
-  # Call PROG on the current file, saving the output as a .png file
-  mmdc -i "$file" -o ./img/flowcharts/$(basename "$file" .mmd).png -t light -b transparent 
+  mmdc -i "$file" -o ./img/flowcharts/$(basename "$file" .mmd).png -t $THEME -w $WIDTH -H $HEIGHT -b $BACKGROUND 
 done

### 2022-12-29 paper.tex
Tue Dec 27 03:05:17 CET 2022: update-remote.sh
autogit

@@
 \usepackage[margin=1in]{geometry}
-\usepackage{color}
-\usepackage{fancyvrb}
-\newcommand{\VerbBar}{|}
-\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
-\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
-% Add ',fontsize=\small' for more characters per line
-\usepackage{framed}
-\definecolor{shadecolor}{RGB}{248,248,248}
-\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
-\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
-\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
-\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\BuiltInTok}[1]{#1}
-\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
-\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
-\newcommand{\ExtensionTok}[1]{#1}
-\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\ImportTok}[1]{#1}
-\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\NormalTok}[1]{#1}
-\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
-\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
-\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\RegionMarkerTok}[1]{#1}
-\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
 \usepackage{graphicx}
@@
 \usepackage{float}
+\usepackage[numbers]{natbib}
+\let\origfigure\figure
+\let\endorigfigure\endfigure
+\renewenvironment{figure}[1][2] {
+    \expandafter\origfigure\expandafter[H]
+} {
+    \endorigfigure
+}
+\usepackage{float}
 \usepackage{booktabs}
@@
 
+Master Thesis - M.Sc. Business Education
+
 --------------------------------------------------------
@@
 \author{}
-\date{\vspace{-2.5em}Date of Submission: 2022-12-23}
+\date{\vspace{-2.5em}Date of Submission: 2022-12-29}
 
@@
 long short-term memory (LSTM) network on a subset of high fitness
-solutions from the current population {[}18{]}. LSTMs are a variant of
-recurrent neural networks first introduced by {[}7{]} that can store
-learned information over en extended time period while avoiding the
-problem of vanishing and/or exploding gradients. Since DAE-GP encodes
-candidate solutions as linear strings in prefix notation, the DAE in
-DAE-GP uses LSTMs for both encoding and decoding where the total amount
-of time steps \(T\) is equal to sum of the length of the input solution
-and the output solution {[}18{]}.
+solutions from the current population {[}18{]}.
+
+LSTMs are a variant of recurrent neural networks first introduced by
+{[}7{]} that can store learned information over en extended time period
+while avoiding the problem of vanishing and/or exploding gradients.
+Since DAE-GP encodes candidate solutions as linear strings in prefix
+notation, the DAE in DAE-GP uses LSTMs for both encoding and decoding
+where the total amount of time steps \(T\) is equal to sum of the length
+of the input solution and the output solution {[}18{]}.
 
@@
 
+Pre-Training describes the concept of initially training a neural
+network on a large dataset before using it to solve a more specific
+task. It is a commonly used strategy in deep architectures that has been
+shown to improve both the optimization process itself as well as the
+generalization behavior if compared to the standard approach of using
+randomly initialized parameters {[}3{]}.
+
 The strategy of pre-training artificial neural networks is based on the
-idea of transfer learning. Transfer learning in the field of machine
-learning aims at retaining previously learned knowledge from one task to
-re-use it for another task {[}5{]} {[}9{]}. Transfer learning
-traditionally follows a two phase approach:
+idea of transfer learning that aims at retaining previously learned
+knowledge from one task to re-use it for another task {[}5{]} {[}9{]}.
+Transfer learning traditionally follows a two phase approach:
 
@@
 where source task and target task are usually similar but may differ in
-their feature space and the distribution of training data {[}9{]}. One
-of the main motivations for using transfer learning in real world
-machine learning tasks is the ability to reduce the need for large
-amounts of training data which can often be unavailable or too expensive
-to collect.
+their feature space and the distribution of training data {[}9{]}.
 
-Pre-Training is a commonly used strategy in deep architectures that has
-been shown to improve both the optimization process itself as well as
-the generalization behavior if compared to randomly initialized
-parameters {[}3{]}.
+Some of the main motivations for using pre-training is the ability to
+reduce the need for large amounts of training data which can often be
+unavailable or too expensive to collect as well as an improved
+performance by either reducing the computational effort that is
+necessary to train a model to solve a specific task or by improving the
+models generalisation ability.
 
@@
 
-The DAE-GP algorithm, first described by {[}18{]}, is described by
-algorithm 1
+The DAE-GP algorithm, first described by {[}18{]}, is summarized by
+figure 1.
 
-\begin{Shaded}
-\begin{Highlighting}[numbers=left,,]
-\NormalTok{WHILE NOT termination\_criterion}
-\NormalTok{  DO }
-\NormalTok{END WHILE}
-\NormalTok{RETURN}
-\end{Highlighting}
-\end{Shaded}
+\begin{figure}
+\centering
+\includegraphics[width=2.08333in,height=8.33333in]{./img/flowcharts/dae-gp.png}
+\caption{Flowchart - Regular DAE-GP}
+\end{figure}
+
+The pre-training strategy implemented is visualized by figure 2.
+
+\begin{figure}
+\centering
+\includegraphics[width=2.08333in,height=8.33333in]{./img/flowcharts/pt-dae-gp.png}
+\caption{Flowchart - Pre-Trained DAE-GP}
+\end{figure}
 
@@
 \centering
-\includegraphics{./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_neurons.png}
-\caption{Airfoil - First Generation Median Training Error for variable
-number of hidden Neurons}
+\includegraphics[width=2.08333in,height=8.33333in]{./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_neurons.png}
+\caption{First Generation Median Training Error for variable number of
+hidden Neurons - Airfoil}
 \end{figure}
@@
 \centering
-\includegraphics{./img/airfoil_firstGen/airfoil_firstGen_boxplot_training_error_by_neurons.png}
-\caption{Airfoil - First Generation final Training Error for variable
-number of hidden Neurons}
+\includegraphics[width=2.08333in,height=8.33333in]{./img/airfoil_firstGen/airfoil_firstGen_boxplot_training_error_by_neurons.png}
+\caption{First Generation final Training Error for variable number of
+hidden Neurons - Airfoil}
 \end{figure}
@@
 \centering
-\includegraphics{./img/airfoil_firstGen/airfoil_firstGen_mean_training_error_by_layers.png}
-\caption{Airfoil - First Generation Median Training Error for variable
-number of hidden Layers}
+\includegraphics[width=2.08333in,height=8.33333in]{./img/airfoil_firstGen/airfoil_firstGen_mean_training_error_by_layers.png}
+\caption{First Generation Median Training Error for variable number of
+hidden Layers - Airfoil}
 \end{figure}
@@
 \centering
-\includegraphics{./img/airfoil_firstGen/airfoil_firstGen_boxplot_training_error_by_layer.png}
-\caption{Airfoil - First Generation final Training Error for variable
-number of hidden Layers}
+\includegraphics[width=2.08333in,height=8.33333in]{./img/airfoil_firstGen/airfoil_firstGen_boxplot_training_error_by_layer.png}
+\caption{First Generation final Training Error for variable number of
+hidden Layers - Airfoil}
 \end{figure}
@@
 length of all solutions inside the current training population. For
-pre-trained DAE-GP another strategy had to be used, we use a fixed
+pre-trained DAE-GP another strategy had to be used, I used a fixed
 number of hidden neurons for each run that is set equal to the maximum
@@
 \centering
-\includegraphics{./img/airfoil_FullRun_30gens/mean_median_fitness_byGens.png}
-\caption{Airfoil - Best Fitness over 30 Generations}
+\includegraphics[width=2.08333in,height=8.33333in]{./img/airfoil_FullRun_30gens/mean_median_fitness_byGens.png}
+\caption{Best Fitness over 30 Generations - Airfoil}
 \end{figure}
+
+The result of the MWU test is 0
 
@@
 \centering
-\includegraphics{./img/airfoil_FullRun_30gens/final_fit_boxplot.png}
-\caption{Airfoil - Fitness Distribution Boxplot}
+\includegraphics[width=2.08333in,height=8.33333in]{./img/airfoil_FullRun_30gens/final_fit_boxplot.png}
+\caption{Best Fitness Distribution after 30 Generations - Airfoil}
 \end{figure}
 
-\includegraphics{./img/airfoil_FullRun_30gens/mean_popDiversity_byGens.png}
+\includegraphics[width=2.08333in,height=8.33333in]{./img/airfoil_FullRun_30gens/mean_popDiversity_byGens.png}
 Since {[}17{]} shows, that DAE-GP finds higher quality that are also
@@
 \centering
-\includegraphics{./img/airfoil_FullRun_30gens/mean_Size_byGens.png}
-\caption{Airfoil - Average Solution Size}
+\includegraphics[width=2.08333in,height=8.33333in]{./img/airfoil_FullRun_30gens/mean_Size_byGens.png}
+\caption{Average Solution Size over 30 Generations - Airfoil}
 \end{figure}

### 2022-12-29 paper.Rmd
autogit

@@
 
-WIDTH=1600
-HEIGHT=1200
+WIDTH=3200
+HEIGHT=2400
 THEME="neutral"
-BACKGROUND="transparent"
+BACKGROUND="white"
 
@@
 
-````
+```
 
@@
 
-![DAE-GP - Algorithm](./img/flowcharts/dae-gp.png)
+![Flowchart - Regular DAE-GP](./img/flowcharts/dae-gp.png){height=800px width=200px}
 
@@
 
-![DAE-GP - Algorithm](./img/flowcharts/pt-dae-gp.png)
+![Flowchart - Pre-Trained DAE-GP](./img/flowcharts/pt-dae-gp.png){height=800px width=200px}
 
@@
 
-![Airfoil - First Generation Median Training Error for variable number of hidden Neurons](./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_neurons.png)
+![First Generation Median Training Error for variable number of hidden Neurons - Airfoil](./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_neurons.png)
 
-![Airfoil - First Generation final Training Error for variable number of hidden Neurons](./img/airfoil_firstGen/airfoil_firstGen_boxplot_training_error_by_neurons.png)
+![First Generation final Training Error for variable number of hidden Neurons - Airfoil](./img/airfoil_firstGen/airfoil_firstGen_boxplot_training_error_by_neurons.png)
 
@@
 
-![Airfoil - First Generation Median Training Error for variable number of hidden Layers](./img/airfoil_firstGen/airfoil_firstGen_mean_training_error_by_layers.png)
+![First Generation Median Training Error for variable number of hidden Layers - Airfoil](./img/airfoil_firstGen/airfoil_firstGen_mean_training_error_by_layers.png)
 
-![Airfoil - First Generation final Training Error for variable number of hidden Layers](./img/airfoil_firstGen/airfoil_firstGen_boxplot_training_error_by_layer.png)
+![First Generation final Training Error for variable number of hidden Layers - Airfoil](./img/airfoil_firstGen/airfoil_firstGen_boxplot_training_error_by_layer.png)
 
@@
 
-![Airfoil - Best Fitness over 30 Generations](./img/airfoil_FullRun_30gens/mean_median_fitness_byGens.png)
+![Best Fitness over 30 Generations - Airfoil](./img/airfoil_FullRun_30gens/mean_median_fitness_byGens.png)
 
@@
 
-![Airfoil - Fitness Distribution Boxplot](./img/airfoil_FullRun_30gens/final_fit_boxplot.png)
+![Best Fitness Distribution after 30 Generations - Airfoil](./img/airfoil_FullRun_30gens/final_fit_boxplot.png)
+lalal
 
-![Airfoil - Average Population Diversity](./img/airfoil_FullRun_30gens/mean_popDiversity_byGens.png)
+![Population Diversity over 30 Generations - Airfoil](./img/airfoil_FullRun_30gens/mean_popDiversity_byGens.png)
 Since @dae-gp_2022_symreg shows, that DAE-GP finds higher quality that are also smaller in size than those found by average GP for a given number of 10.000 fitness evaluations, I also studied the differences in average solution size between pre-trained and regular DAE-GP.
 
-![Airfoil - Average Solution Size](./img/airfoil_FullRun_30gens/mean_Size_byGens.png)
+![Average Solution Size over 30 Generations - Airfoil](./img/airfoil_FullRun_30gens/mean_Size_byGens.png)
 

### 2022-12-30 paper.Rmd
Fri Dec 30 13:03:41 CET 2022: update-remote.sh
Fri Dec 30 14:00:07 CET 2022: update-remote.sh

@@
 knitr::opts_chunk$set(echo = TRUE)
-knitr::opts_chunk$set(fig.pos = "!H")
+#knitr::opts_chunk$set(fig.pos = "!H")
 
@@
 GP is an especially useful meta heuristic for problems where no a priori knowledge about the final form of good solutions is available.
-GPs ability to optimize solutions for their structure as well as for their parameters led to it being one of the most prevalent methods used symbolic regression [@10.1007/978-3-540-24621-3_22].
+GPs ability to optimize solutions for their structure as well as for their parameters led to it being one of the most prevalent methods used for symbolic regression [@10.1007/978-3-540-24621-3_22].
 
@@
 One possible way of model building in EDA proposed by @harmless_overfitting_eda is to use denoising autoencoders (DAEs) as generative models to capture complex probability distributions of decision variables.
-
 DAE, a variation of the autoencoders (AE), are a widely used type of neural networks in the domain of unsupervized machine learning that maps $n$ input variables to $n$ output variables using a hidden representation.
@@
 
-The DAE-GP algorithm, first described by @dae-gp_2020_rtree, is summarized by figure 1.
+The DAE-GP algorithm, first described by @dae-gp_2020_rtree, is summarized as a flowchart in figure 1. 
+After setting the generation counter $g$ to $0$ and creating the initial population of solutions $P_0$, DAE-GP enters the main loop that checks if a termination criteria is satisfied (i.e.
+maximum number of generations or fitness evaluations is reached).
+For each generation $g$ the fitness of all new individuals in $P_g$ is evaluated.
+During model building, DAE-GP selects a subset $X_g$ of the current population that is used as training data for the DAE-LSTM model $M_g$ which is trained to learn the properties of $X_g$.
+The trained model $M_g$ is then used for model sampling, partially mutated solutions from $P_G$ are propagated through $M_g$ to create a new population $P_{g+1}$. 
+This process is repeated until the termination criteria is met and DAE-GP returns the highest fitness solutions.
 
@@
 
-The pre-training strategy implemented is visualized by figure 2.
+The pre-training strategy implemented for all experiments in this thesis is visualized as another flowchart in figure 2.
+The main difference to regular DAE-GP is the inclusion of an initial pre-training phase where a separate population $\hat{P}$ is first initialized and then randomly split by half into $\hat{P_{train}}$ and $\hat{P_{test}}$.
+A DAE-LSTM model $\hat{M}$
 

### 2022-12-30 paper.Rmd
autogit

@@
 The main difference to regular DAE-GP is the inclusion of an initial pre-training phase where a separate population $\hat{P}$ is first initialized and then randomly split by half into $\hat{P_{train}}$ and $\hat{P_{test}}$.
-A DAE-LSTM model $\hat{M}$
+A DAE-LSTM model $\hat{M}$ is then trained to learn the properties of $\hat{P_{train}}$ using an early stopping training mode where we stop training as soon as the validation error for $\hat{P_{test}}$ converges.
+After the training for $\hat{M}$ is stopped, we freeze the current state of $\hat{M}$ and terminate the pre-training phase to start the DAE-GP phase. 
+During the DAE-GP phase, the only difference to the traditional DAE-GP algorithm described in figure 1 is during model building: We initialize $M_g$ with the weights and biases of $\hat{M}$ before training it to learn the properties of $X_g$.
 

### 2022-12-30 paper.Rmd
Fri Dec 30 19:00:06 CET 2022: update-remote.sh

@@
 bibliography: ref/ref.bib
+fig.width: 4in
+fig.height: 3in
 csl: csl/association-for-computing-machinery.csl
@@
 
-<!--- Explain algorithm and pre-training strategy in pseudo code -->
+<!--- Explain algorithm and implementation of the pre-training strategy  -->
 
@@
 
-![Flowchart - Regular DAE-GP](./img/flowcharts/dae-gp.png){height=800px width=200px}
+![Flowchart - Regular DAE-GP](./img/flowcharts/dae-gp.png){height=900px width=200px}
 

### 2022-12-30 paper.Rmd
Fri Dec 30 20:00:06 CET 2022: update-remote.sh
autogit

@@
 
-<!---
-## Exploratory Phase
+## Effect on Generalisation
 
-In a first exploratory experiment DAE-GP was tested on the Aifoil dataset to perform a symbolic regression task.
-The aim of this initial experiment is to search for differences in behavior between an algorithm that uses a pre-training strategy and regular DAE-GP.
-The parameters for the experiment are summarized in table 2, the average results for fitness, number of fitness evaluations, average tree size of the population and the unique rate of the population are summarized in plot 1.
-
-```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
-fig.pos="H"
-knitr::kable(
-  read.csv("/Users/rmn/masterThesis/eda-gp-2020/experiments/airfoil_ptdaegp-daegp_rwsr_1hl/parameters.csv"),
-  digits=3,
-  caption="Airfoil - Configuration"
-)%>% row_spec(0,bold=TRUE ) %>% kableExtra::kable_styling(latex_options = "hold_position")
+```{bash compile_firstGenResults, include=FALSE}
+bash /Users/rmn/masterThesis/eda-gp-2020/experiments/airfoil_firstGen/re_evaluate_airfoilFirstGen.sh
 ```
-
-![Airfoil - First Results](./img/airfoil_exploratory/airfoil_exploratory.png)
-
-In this first experiment, the average quality of solutions produced by the pre-trained algorithm shows to be lower than that produced by regular DAE-GP.
-Differences in fitness start to show at generation 5 where regular DAE-GP starts to reduces RMSE more efficiently than the pre-trained version.
-The gap in average fitness grows larger until the end of the evolutionary search between both algorithms.
-Observing the gap between RMSE on the training dataset and the testing dataset shows no major differences, the training error is slightly lower than the testing error for both algorithms and indicates a similar generalization behavior.
-
-Another interesting metric to investigate is the population's average tree size, in general smaller solutions are preferable to their larger counterparts because of improved interpretability.
-The results show, that average tree size is reduced strongly in generations 1-5 for both algorithms and than slowly increases back.
-Differences between regular DAE-GP and the pre-trained DAE-GP start to show at generation 2, where the average population tree size is continuously lower for regular DAE-GP.
-
-It can also be observed that the population diversity, measured by the normalized levenshtein distance, is different between both DAE-GP variants.
-Both algorithms start at a unique rate of about 0.8 in the 0.th generation and lower rapidly to an average of about 0.45 in the 7.th generation.
-In the following generations, the diversity of pre-trained DAE-GP keeps lowering steadily while regular DAE-GP slowly increases the population diversity to a normalized levenshtein distance of about 0.55.
-This could be a first explanation for the better performance of regular DAE-GP in this experiment, by creating more diverse populations over the later generations of the search the algorithm is better at exploiting the search space than the pre-trained DAE-GP algorithm.
--->
-
-## Effect on Generalisation
 
@@
 
-<!--- analyzing by layers -->
+<!--- analyzing the results for variable number of layers -->
 

### 2022-12-30 slides.Rmd
autogit

@@
 author: "Roman Hoehn"
-institute: "JGU Mainz "
+institute: "Johannes Gutenberg-Universität Mainz"
 date: "Datum: `r Sys.Date()`"
@@
   beamer_presentation:
-    theme: "Madrid"
-    fonttheme: "structurebold"
+    #theme: "Madrid"
+    #fonttheme: "structurebold"
     fig_width: 12
@@
 bibliography: ref/ref.bib
-csl: csl/association-for-computing-machinery.csl
+csl: csl/harvard-cite-them-right.csl
 includes:
@@
 
-# Introduction
+# Einführung
 
-## Research Question
+## Forschungsfrage
+
+*Kann das Suchverhalten der Denoising Autoencoder Genetic Programming (DAE-GP) Metaheuristik durch den Einsatz von einer Pre-Training Strategie optimiert werden?*
+
+Welchen Effekt hat Pre-Training auf:
+
+  1. das Generalisierungsverhalten von DAE-GP?
+  2. die Qualität der gefundenen Programme?
+  3. die Populationsdiversität?
+  4. das Laufzeitverhalten?
+
+Anwendungsgebiet: Symbolische Regression, insbesondere am Aifoil Datensatz
 
@@
 
+## Implementation
+
+# Experimente
+
+## Symbolische Regression
+
+## Generalisierungsverhalten
+
+## Lösungs
+
 # References {.allowframebreaks}

### 2022-12-30 paper.Rmd
autogit

@@
 reticulate::use_python("/Users/rmn/miniconda3/envs/dataScience/bin/python")
-
-```
-
-```{bash compile_mermaid, include=FALSE}
-# compile all *.mmd files in the mermaid directory
-
-WIDTH=3200
-HEIGHT=2400
-THEME="neutral"
-BACKGROUND="white"
-
-for file in ./mermaid/*.mmd
-do
-  mmdc -i "$file" -o ./img/flowcharts/$(basename "$file" .mmd).png -t $THEME -w $WIDTH -H $HEIGHT -b $BACKGROUND 
-done
-
-```
-
-```{bash ipythonRunner, eval=FALSE, include=FALSE}
-# Set the base directory
-EXPDIR="/Users/rmn/masterThesis/eda-gp-2020"
-
-# Find all *.ipynb files in subdirectories of EXPDIR, including sub-sub directories, etc., but exclude files that contain the word "checkpoint" in the filename
-for file in $(find "$EXPDIR" -type f -name "*.ipynb" ! -name "*checkpoint*"); do
-  # Change to the directory containing the IPython notebook file
-  cd "$(dirname "$file")"
-  # Execute the IPython notebook file using the jupyter nbconvert command
-  ipython "$(basename "$file")"
-done
 
@@
 
+```{bash compile_mermaid, include=FALSE}
+# compile all *.mmd files in the mermaid directory
+
+WIDTH=3200
+HEIGHT=2400
+THEME="neutral"
+BACKGROUND="white"
+
+for file in ./mermaid/*.mmd
+do
+  mmdc -i "$file" -o ./img/flowcharts/$(basename "$file" .mmd).png -t $THEME -w $WIDTH -H $HEIGHT -b $BACKGROUND 
+done
+
+```
+
 <!--- Explain algorithm and implementation of the pre-training strategy  -->
@@
 
-![Flowchart - Regular DAE-GP](./img/flowcharts/dae-gp.png){height=900px width=200px}
+![Flowchart - Regular DAE-GP](./img/flowcharts/dae-gp.png)
 
@@
 
-![Flowchart - Pre-Trained DAE-GP](./img/flowcharts/pt-dae-gp.png){height=800px width=200px}
+![Flowchart - Pre-Trained DAE-GP](./img/flowcharts/pt-dae-gp.png)
 
@@
 
+```{bash compile_airfoil_2hl_maxIndSize_fullRun_30gens, include=FALSE}
+bash /Users/rmn/masterThesis/eda-gp-2020/experiments/airfoil_2hl_maxIndSize_fullRun_30gens/re_evaluate_airfoil_2hl_maxIndSize_30gen.sh
+```
+
 After closely examining the effect of pre-training on DAE-GPs generalization behaviour another series of experiments is conducted to study how pre-training influences the overall search behaviour of DAE-GP.

### 2023-01-01 paper.Rmd
autogit

@@
 @dae-gp_2022_symreg and @daegp_explore_exploit suggests the incorporation of a pre-training strategy into the evolutionary search as a possible way of improving the model quality of DAE-GP.
-The key idea is to pre-train an initial DAE-LSTM on a large population of candidate solutions and to use the weights of this initial model in each generation of the search as a starting point.
+The key idea is to pre-train an initial DAE-LSTM on a large population of candidate solutions and to use the pre-trained parameters of this initial model in each generation of the search as starting parameters for the current generation DAE-LSTM.
 This thesis studies the influence of using pre-trained DAE-LSTMs in DAE-GP for symbolic regression, especially looking at the influence in generalization behavior and the overall quality of solutions found by DAE-GP.
@@
 [@dae-gp_2020_rtree] [@dae-gp_2022_symreg]), a separate DAE-LSTM network $\hat{M}$ will be trained on a large population of randomly initialized solutions.
-The optimal parameters obtained after finishing the training procedure of $\hat{M}$ are then used as the starting parameters for each following DAE-LSTM model $M_g$ for $g\in{1,...,g_{max}}$.
+The trainable parameters $\theta_{\hat{M}}$ obtained after finishing the training procedure of $\hat{M}$ are then used as the starting parameters for each following DAE-LSTM model $M_g$ for $g\in{1,...,g_{max}}$.
 
@@
 
-The probably most obvious motivation for using a pre-training strategy in DAE-GP is based on the fact, that initializing $M_g$ with pre-trained weights from $\hat{M}$ is likely to reduce the amount of training epochs that are necessary until the point of training error convergence is reached.
+The probably most obvious motivation for using a pre-training strategy in DAE-GP is based on the fact, that initializing $M_g$ with pre-trained weights $\theta_{\hat{M}}$ is likely to reduce the amount of training epochs that are necessary until the point of training error convergence is reached.
 
@@
 
-The DAE-GP algorithm, first described by @dae-gp_2020_rtree, is summarized as a flowchart in figure 1. 
+The DAE-GP algorithm, first described by @dae-gp_2020_rtree, is summarized as a flowchart in figure §. 
 After setting the generation counter $g$ to $0$ and creating the initial population of solutions $P_0$, DAE-GP enters the main loop that checks if a termination criteria is satisfied (i.e.
@@
 
-The pre-training strategy implemented for all experiments in this thesis is visualized as another flowchart in figure 2.
+The pre-training strategy implemented for all experiments in this thesis is visualized as another flowchart in figure §.
 The main difference to regular DAE-GP is the inclusion of an initial pre-training phase where a separate population $\hat{P}$ is first initialized and then randomly split by half into $\hat{P_{train}}$ and $\hat{P_{test}}$.
 A DAE-LSTM model $\hat{M}$ is then trained to learn the properties of $\hat{P_{train}}$ using an early stopping training mode where we stop training as soon as the validation error for $\hat{P_{test}}$ converges.
-After the training for $\hat{M}$ is stopped, we freeze the current state of $\hat{M}$ and terminate the pre-training phase to start the DAE-GP phase. 
-During the DAE-GP phase, the only difference to the traditional DAE-GP algorithm described in figure 1 is during model building: We initialize $M_g$ with the weights and biases of $\hat{M}$ before training it to learn the properties of $X_g$.
+After the training for $\hat{M}$ is stopped, the current state of $\hat{M}$ is frozen and the optimized trainable parameters $\theta_{\hat{M}}$ are saved before terminating the pre-training phase to start the DAE-GP phase.
+During the DAE-GP phase, the only difference to the traditional DAE-GP algorithm described in figure § is during model building: We initialize $M_g$ with $\theta_{\hat{M}}$ before training it to learn the properties of $X_g$.
 
@@
 
+<!-- Results for Energy Cooling Dataset  -->
+
+```{bash compile_energyCooling_2hl_maxIndSize_fullRun_30gens, include=FALSE}
+bash /Users/rmn/masterThesis/eda-gp-2020/experiments/energyCooling_2hl_maxIndSize_fullRun_30gens/re_evaluate_energyCooling_2hl_maxIndSize_fullRun_30gens.sh
+```
+
 # Discussion

### 2023-01-01 slides.Rmd
autogit

@@
   1. das Generalisierungsverhalten von DAE-GP?
-  2. die Qualität der gefundenen Programme?
+  2. die Qualität der gefundenen Programme (Fitness/Programmlänge)?
   3. die Populationsdiversität?
@@
 ## Denoising Autoencoder Genetic Programming
+
+* Metaheuristik basierend auf genetischer Programmierung (GP)
+* Einsatz von künstlichen, neuronalen Netzen zur Optimierung des GP Suchverhaltens
+* Aktueller Forschungsstand:
+  1. 
+  2.
+
+## Estimation of Distribution Genetic Programming (EDA-GP)
+
+* Ersetzt die genetische Operatoren Mutation und Rekombination durch den Einsatz von probabilistischen Modellen
+
+2 Phase Ansatz:
+
+  1. Model Building: 
+  2. Model Sampling:
 

### 2023-01-02 slides.Rmd
Mon Jan  2 15:00:06 CET 2023: update-remote.sh

@@
   beamer_presentation:
-    #theme: "Madrid"
+    color: "beaver"
+    theme: "Szeged"
     #fonttheme: "structurebold"
@@
     fig_height: 10
+    fontsize: 11pt
     # keep_tex: true
@@
 
-# Einführung
+# Einleitung
 
@@
 
-*Kann das Suchverhalten der Denoising Autoencoder Genetic Programming (DAE-GP) Metaheuristik durch den Einsatz von einer Pre-Training Strategie optimiert werden?*
+*Kann das Suchverhalten der Denoising Autoencoder Genetic Programming (DAE-GP) Metaheuristik durch den Einsatz einer Pre-Training Strategie optimiert werden?*
 
@@
 
-## Denoising Autoencoder Genetic Programming
+## Ausblick
+
+...
+
+# Denoising Autoencoder Genetic Programming
+
+## Übersicht
 
@@
 * Einsatz von künstlichen, neuronalen Netzen zur Optimierung des GP Suchverhaltens
-* Aktueller Forschungsstand:
-  1. 
-  2.
+* Variante des Estimation of Distribution-GP (EDA-GP)
 
@@
   2. Model Sampling:
+
+## Aktueller Forschungsstand
+
+### Generalisiertes Royal Tree Problem (GRT)^[@dae-gp_2020_rtree]:
+
+  * Einfaches Suchproblem mit hoher Lokalität
+  * DAE-GP erzeugt durch Model Sampling Lösungskandidaten mit höherer Fitness als GP 
+  * Hohe Güte der erzeugten Lösungskandidaten resultiert in besserer Performance
+  * Perfomance Vorteil steigt mit zunehmender Komplexität des GRT Problems
+
+### Symbolische Regression^[@dae-gp_2022_symreg]:
+
+  * Airfoil Datensatz für Real-World symbolische Regression
+  * DAE-GP erzeugt für eine vorgegebene Anzahl von Fitness Evaluationen im Vergleich zu GP:
+
+    1.) Lösungen mit höherer Fitness
+    2.) Lösungen  mit geringerer Größe
 

### 2023-01-02 slides.tex
Mon Jan  2 15:00:06 CET 2023: update-remote.sh

--- 2022-12-19 expose.tex
@@
 \documentclass[
-  12pt,
-]{article}
+  ignorenonframetext,
+]{beamer}
+\usepackage{pgfpages}
+\setbeamertemplate{caption}[numbered]
+\setbeamertemplate{caption label separator}{: }
+\setbeamercolor{caption name}{fg=normal text.fg}
+\beamertemplatenavigationsymbolsempty
+% Prevent slide breaks in the middle of a paragraph
+\widowpenalties 1 10000
+\raggedbottom
+\setbeamertemplate{part page}{
+  \centering
+  \begin{beamercolorbox}[sep=16pt,center]{part title}
+    \usebeamerfont{part title}\insertpart\par
+  \end{beamercolorbox}
+}
+\setbeamertemplate{section page}{
+  \centering
+  \begin{beamercolorbox}[sep=12pt,center]{part title}
+    \usebeamerfont{section title}\insertsection\par
+  \end{beamercolorbox}
+}
+\setbeamertemplate{subsection page}{
+  \centering
+  \begin{beamercolorbox}[sep=8pt,center]{part title}
+    \usebeamerfont{subsection title}\insertsubsection\par
+  \end{beamercolorbox}
+}
+\AtBeginPart{
+  \frame{\partpage}
+}
+\AtBeginSection{
+  \ifbibliography
+  \else
+    \frame{\sectionpage}
+  \fi
+}
+\AtBeginSubsection{
+  \frame{\subsectionpage}
+}
 \usepackage{amsmath,amssymb}
@@
 \fi
+\usetheme[]{Szeged}
+\usecolortheme{beaver}
+\usefonttheme{12}
 % Use upquote if available, for straight quotes in verbatim environments
@@
 \usepackage{xcolor}
-\usepackage[margin=1in]{geometry}
-\usepackage{graphicx}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-% Set default figure placement to htbp
-\makeatletter
-\def\fps@figure{htbp}
-\makeatother
+\newif\ifbibliography
 \setlength{\emergencystretch}{3em} % prevent overfull lines
@@
 \newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
-\usepackage{float}
-\usepackage[numbers]{natbib}
-\let\origfigure\figure
-\let\endorigfigure\endfigure
-\renewenvironment{figure}[1][2] {
-    \expandafter\origfigure\expandafter[H]
-} {
-    \endorigfigure
-}
-\usepackage{float}
-\usepackage{booktabs}
-\usepackage{longtable}
-\usepackage{array}
-\usepackage{multirow}
-\usepackage{wrapfig}
-\usepackage{colortbl}
-\usepackage{pdflscape}
-\usepackage{tabu}
-\usepackage{threeparttable}
-\usepackage{threeparttablex}
-\usepackage[normalem]{ulem}
-\usepackage{makecell}
-\usepackage{xcolor}
 \ifLuaTeX
@@
 \hypersetup{
+  pdftitle={Pre-Trained Denoising Autoencoders Long Short-Term Memory Networks as probabilistic Models for Estimation of Distribution Genetic Programming},
+  pdfauthor={Roman Hoehn},
   hidelinks,
@@
 
-\title{Pre-Training Denoising Autoencoders for Genetic Programming
-
-(subtitle)
-
---------------------------------------------------------
-
-Student: Roman Höhn
-
-Date of Birth: 1991-04-14
-
-Place of Birth: Wiesbaden, Hesse
-
-Student ID: 2712497
-
-Supervisor: David Wittenberg
-
---------------------------------------------------------
-
-Master Thesis
-
-FB 03: Chair of Business Administration and Computer Science
-
-Johannes Gutenberg University Mainz
-
-Summerterm 2022}
-\author{}
-\date{\vspace{-2.5em}Date of Submission: 2022-09-09}
+\title{Pre-Trained Denoising Autoencoders Long Short-Term Memory
+Networks as probabilistic Models for Estimation of Distribution Genetic
+Programming}
+\subtitle{Kolloquium zur Masterarbeit im M.Sc. Wirtschaftspädagogik}
+\author{Roman Hoehn}
+\date{Datum: 2023-01-02}
+\institute{Johannes Gutenberg-Universität Mainz}
 
 \begin{document}
-\maketitle
+\frame{\titlepage}
 
-\thispagestyle{empty} \newpage{}
+\begin{frame}[allowframebreaks]
+  \tableofcontents[hideallsubsections]
+\end{frame}
+\hypertarget{einleitung}{%
+\section{Einleitung}\label{einleitung}}
 
-\setcounter{page}{1}
+\begin{frame}{Forschungsfrage}
+\protect\hypertarget{forschungsfrage}{}
+\emph{Kann das Suchverhalten der Denoising Autoencoder Genetic
+Programming (DAE-GP) Metaheuristik durch den Einsatz einer Pre-Training
+Strategie optimiert werden?}
 
-\tableofcontents
-\thispagestyle{empty}
-\newpage
+Welchen Effekt hat Pre-Training auf:
 
-\listoftables
-\listoffigures
-\thispagestyle{empty}
-\newpage
+\begin{enumerate}
+\tightlist
+\item
+  das Generalisierungsverhalten von DAE-GP?
+\item
+  die Qualität der gefundenen Programme (Fitness/Programmlänge)?
+\item
+  die Populationsdiversität?
+\item
+  das Laufzeitverhalten?
+\end{enumerate}
 
-\hypertarget{expose}{%
-\section{Expose}\label{expose}}
+Anwendungsgebiet: Symbolische Regression, insbesondere am Aifoil
+Datensatz
+\end{frame}
 
-\hypertarget{introduction}{%
-\subsection{Introduction}\label{introduction}}
+\begin{frame}{Ausblick}
+\protect\hypertarget{ausblick}{}
+\ldots{}
+\end{frame}
 
-Denoising Autoencoder Genetic Programming (DAE-GP) is a novel variation
-of an genetic programming based Estimation of Distribution Algorithm
-(EDA-GP) that uses a denoising autoencoders long short-term memory
-network (DAE-LSTMs) as a probabilistic model to sample new candidate
-solutions (Wittenberg, Rothlauf and Schweim, 2020).
+\hypertarget{denoising-autoencoder-genetic-programming}{%
+\section{Denoising Autoencoder Genetic
+Programming}\label{denoising-autoencoder-genetic-programming}}
 
-DAE-LSTMs are recurrent neural networks (RNNs) that can be trained in an
-unsupervised learning environment to minimize a reconstruction error for
-encoding input data into a compressed representation and subsequently
-decoding the compressed representation back to the input dimension. In
-DAE-GP, DAE-LSTMs are trained with a subset of high-fitness solutions
-selected from the parent population as input. The resulting model is
-then used to sample new offspring solutions by propagating partially
-mutated solutions from the parent population through the DAE-LSTMS
-(Wittenberg, Rothlauf and Schweim, 2020).
+\begin{frame}{Übersicht}
+\protect\hypertarget{uxfcbersicht}{}
+\begin{itemize}
+\tightlist
+\item
+  Metaheuristik basierend auf genetischer Programmierung (GP)
+\item
+  Einsatz von künstlichen, neuronalen Netzen zur Optimierung des GP
+  Suchverhaltens
+\item
+  Variante des Estimation of Distribution-GP (EDA-GP)
+\end{itemize}
+\end{frame}
 
-In previous work DAE-GP has been demonstrated to outperform GP for both
-a generalized version of the royal tree problem (Wittenberg, Rothlauf
-and Schweim, 2020) as well as for a standard symbolic regression problem
-(Wittenberg and Rothlauf, 2022). The DAE-GP algorithm implemented for
-both experiments initializes the DAE-LSTMs with randomized parameters
-for each generation \(g\) of the search. A possible way of improving the
-quality of solutions and/or the total runtime of DAE-GP could be to
-instead use pre-trained DAE-LSTMs that are passed on from each
-generation \(g_{t}\) to the next generation \(g_{t+1}\). Pre-Training is
-a commonly used strategy in deep architectures that has been shown to
-improve both the optimization process itself as well as the
-generalization behavior if compared to randomly initialized parameters
-(Erhan \emph{et al.}, 2009).
+\begin{frame}{Estimation of Distribution Genetic Programming (EDA-GP)}
+\protect\hypertarget{estimation-of-distribution-genetic-programming-eda-gp}{}
+\begin{itemize}
+\tightlist
+\item
+  Ersetzt die genetische Operatoren Mutation und Rekombination durch den
+  Einsatz von probabilistischen Modellen
+\end{itemize}
 
-\hypertarget{research-question}{%
-\subsection{Research Question}\label{research-question}}
+2 Phase Ansatz:
 
-The aim of this research project is to study the effects of using
-pre-trained DAE-LSTMs in DAE-GP on both solution quality as well as on
-total time consumption.
+\begin{enumerate}
+\tightlist
+\item
+  Model Building:
+\item
+  Model Sampling:
+\end{enumerate}
+\end{frame}
 
-\ldots what problem(s)???\ldots{}
+\begin{frame}{Aktueller Forschungsstand}
+\protect\hypertarget{aktueller-forschungsstand}{}
+\begin{block}{Generalisiertes Royal Tree Problem (GRT)\footnote<.->{Wittenberg,
+  Rothlauf and Schweim (2020)}:}
+\protect\hypertarget{generalisiertes-royal-tree-problem-grt}{}
+\begin{itemize}
+\tightlist
+\item
+  Einfaches Suchproblem mit hoher Lokalität
+\item
+  DAE-GP erzeugt durch Model Sampling Lösungskandidaten mit höherer
+  Fitness als GP
+\item
+  Hohe Güte der erzeugten Lösungskandidaten resultiert in besserer
+  Performance
+\item
+  Perfomance Vorteil steigt mit zunehmender Komplexität des GRT Problems
+\end{itemize}
+\end{block}
 
-\hypertarget{research-design}{%
-\subsection{Research Design}\label{research-design}}
+\begin{block}{Symbolische Regression\footnote<.->{Wittenberg and
+  Rothlauf (2022)}:}
+\protect\hypertarget{symbolische-regression}{}
+\begin{itemize}
+\item
+  Airfoil Datensatz für Real-World symbolische Regression
+\item
+  DAE-GP erzeugt für eine vorgegebene Anzahl von Fitness Evaluationen im
+  Vergleich zu GP:
 
-\newpage
+  1.) Lösungen mit höherer Fitness 2.) Lösungen mit geringerer Größe
+\end{itemize}
+\end{block}
+\end{frame}
 
-\hypertarget{I}{%
-\section*{I - References}\label{I}}
-\addcontentsline{toc}{section}{I - References}
+\begin{frame}{Pre-Training}
+\protect\hypertarget{pre-training}{}
+\end{frame}
 
+\begin{frame}{Implementation}
+\protect\hypertarget{implementation}{}
+\end{frame}
+
+\hypertarget{experimente}{%
+\section{Experimente}\label{experimente}}
+
+\begin{frame}{Symbolische Regression}
+\protect\hypertarget{symbolische-regression-1}{}
+\end{frame}
+
+\begin{frame}{Generalisierungsverhalten}
+\protect\hypertarget{generalisierungsverhalten}{}
+\end{frame}
+
+\begin{frame}{Lösungs}
+\protect\hypertarget{luxf6sungs}{}
+\end{frame}
+
+\hypertarget{references}{%
+\section*{References}\label{references}}
+\addcontentsline{toc}{section}{References}
+
+\begin{frame}[allowframebreaks]{References}
 \hypertarget{refs}{}
 \begin{CSLReferences}{0}{0}
-\leavevmode\vadjust pre{\hypertarget{ref-pmlr-v5-erhan09a}{}}%
-Erhan, D. \emph{et al.} (2009) {``The difficulty of training deep
-architectures and the effect of unsupervised pre-training,''} in D. van
-Dyk and M. Welling (eds.) \emph{Proceedings of the twelth international
-conference on artificial intelligence and statistics}. Hilton Clearwater
-Beach Resort, Clearwater Beach, Florida USA: PMLR (Proceedings of
-machine learning research), pp. 153--160. Available at:
-\url{https://proceedings.mlr.press/v5/erhan09a.html}.
-
 \leavevmode\vadjust pre{\hypertarget{ref-dae-gp_2022_symreg}{}}%
-Wittenberg, D. and Rothlauf, F. (2022) {``Denoising autoencoder genetic
-programming for real-world symbolic regression,''} in \emph{Proceedings
+Wittenberg, D. and Rothlauf, F. (2022) {`Denoising autoencoder genetic
+programming for real-world symbolic regression'}, in \emph{Proceedings
 of the genetic and evolutionary computation conference companion}. New
@@
 \end{CSLReferences}
-
-\hypertarget{II}{%
-\section*{II - Statutory Declaration}\label{II}}
-\addcontentsline{toc}{section}{II - Statutory Declaration}
-
-\includegraphics{./private/erklaerung.pdf}\\
+\end{frame}
 

### 2023-01-02 slides.Rmd
Mon Jan  2 16:00:06 CET 2023: update-remote.sh

@@
 institute: "Johannes Gutenberg-Universität Mainz"
-date: "Datum: `r Sys.Date()`"
+date: "Datum: 25.01.2023"
+fontsize: 11pt
 output:
@@
     fig_height: 10
-    fontsize: 11pt
     # keep_tex: true
@@
 
-## Ausblick
-
-...
-
 # Denoising Autoencoder Genetic Programming
@@
 * Metaheuristik basierend auf genetischer Programmierung (GP)
-* Einsatz von künstlichen, neuronalen Netzen zur Optimierung des GP Suchverhaltens
+* Erweiterung von GP durch künstliche, neuronalen Netzen zur Optimierung des Suchverhaltens
 * Variante des Estimation of Distribution-GP (EDA-GP)
 
-## Estimation of Distribution Genetic Programming (EDA-GP)
+## Estimation of Distribution Algorithmen (EDA)
 
-* Ersetzt die genetische Operatoren Mutation und Rekombination durch den Einsatz von probabilistischen Modellen
+* Entwicklung neuer Rekombinationsoperatoren für evolutionäre Algorithmen basierend auf dem Einsatz von probabilistischen Modellen^[@design_of_modern_heuristics]
+* Hypothese: Problemspezifische Abhängigkeiten zwischen einzelnen Entscheidungsvariablen können bei der Erzeugung neuer Lösungskandidaten besser berücksichtigt werden als bei traditionellen Rekombinationsoperatoren^[@edaOrig1996]
+* Weitere Verbreitung im Bereich der genetischen Algorithmen (GA) als für GP
+
+## DAE-EDA
+
+Idee: Einsatz von Denoising Autoencoders (DAE) als probabilistisches Modell für EDA [@harmless_overfitting_eda]
 
@@
 
-  1. Model Building: 
-  2. Model Sampling:
+  1. Model Building: Modell "lernt" die Eigenschaften von ausgewählten Lösungen hoher Güte
+  2. Model Sampling: Neue Lösungen werden erzeugt durch das propagieren von bestehenden, mutierten Lösungen durch das erlernte Modell
 
-## Aktueller Forschungsstand
+## DAE-GP
 
-### Generalisiertes Royal Tree Problem (GRT)^[@dae-gp_2020_rtree]:
+# Aktueller Forschungsstand
+
+## Generalisiertes Royal Tree Problem (GRT)^[@dae-gp_2020_rtree]:
 
@@
 
-### Symbolische Regression^[@dae-gp_2022_symreg]:
+## Symbolische Regression^[@dae-gp_2022_symreg]:
 
@@
 
-    1.) Lösungen mit höherer Fitness
-    2.) Lösungen  mit geringerer Größe
+    1. Lösungen mit höherer Fitness
+    2. Lösungen  mit geringerer Größe
 
@@
 
-## Lösungs
+## Lösungsqualität
+
+# Resultate
+
+# Weitere Ansätze
 

### 2023-01-02 slides.Rmd
autogit

@@
   beamer_presentation:
-    color: "beaver"
-    theme: "Szeged"
-    #fonttheme: "structurebold"
+    # color: "beaver"
+    # theme: "Szeged"
+    # fonttheme: "structurebold"
     fig_width: 12
@@
 
-Idee: Einsatz von Denoising Autoencoders (DAE) als probabilistisches Modell für EDA [@harmless_overfitting_eda]
+Idee: Einsatz von Denoising Autoencoders^[@dae_orig2008] (DAE) als probabilistisches Modell für EDA ^[@harmless_overfitting_eda]
 

### 2023-01-06 paper.Rmd
Fri Jan  6 14:00:08 CET 2023: update-remote.sh

@@
 ```{bash compile_airfoil_2hl_maxIndSize_fullRun_30gens, include=FALSE}
+bash /Users/rmn/masterThesis/eda-gp-2020/experiments/airfoil_1hl_maxIndSize_fullRun_30gens/re_evaluate_airfoil_1hl_maxIndSize_30gen.sh
+
 bash /Users/rmn/masterThesis/eda-gp-2020/experiments/airfoil_2hl_maxIndSize_fullRun_30gens/re_evaluate_airfoil_2hl_maxIndSize_30gen.sh
+
+bash /Users/rmn/masterThesis/eda-gp-2020/experiments/energyCooling_2hl_maxIndSize_fullRun_30gens/re_evaluate_energyCooling_2hl_maxIndSize_fullRun_30gens.sh
+
+# add new experiments for full runs here!
+# ...
 ```

### 2023-01-09 paper.Rmd
Mon Jan  9 14:00:10 CET 2023: update-remote.sh

@@
 ```{bash compile_airfoil_2hl_maxIndSize_fullRun_30gens, include=FALSE}
+
+# airfoil
 bash /Users/rmn/masterThesis/eda-gp-2020/experiments/airfoil_1hl_maxIndSize_fullRun_30gens/re_evaluate_airfoil_1hl_maxIndSize_30gen.sh
@@
 
+# energy (cooling)
 bash /Users/rmn/masterThesis/eda-gp-2020/experiments/energyCooling_2hl_maxIndSize_fullRun_30gens/re_evaluate_energyCooling_2hl_maxIndSize_fullRun_30gens.sh
+
+# boston housing
+bash /Users/rmn/masterThesis/eda-gp-2020/experiments/bostonHousing_2hl_maxIndSize_fullRun_30gens/re_evaluate_bostonHousing_2hl_maxIndSize_fullRun_30gens.sh
 

### 2023-01-09 slides.Rmd
Mon Jan  9 14:00:10 CET 2023: update-remote.sh

@@
 
-Idee: Einsatz von Denoising Autoencoders^[@dae_orig2008] (DAE) als probabilistisches Modell für EDA ^[@harmless_overfitting_eda]
+Idee: Einsatz von Denoising Autoencoders^[@dae_orig2008] (DAE) als probabilistisches Modell für genetische Algorithmen ^[@harmless_overfitting_eda]
 
@@
 
-  1. Model Building: Modell "lernt" die Eigenschaften von ausgewählten Lösungen hoher Güte
+  1. Model Building: Modell "lernt" die Eigenschaften von ausgewählten Lösungen hoher Güte durch das Trainieren eines DAEs
   2. Model Sampling: Neue Lösungen werden erzeugt durch das propagieren von bestehenden, mutierten Lösungen durch das erlernte Modell
@@
 ## DAE-GP
+
+* Adaptierung der DAE-EDA Strategie auf GP
 

### 2023-01-09 slides.Rmd
autogit

@@
     # fonttheme: "structurebold"
-    fig_width: 12
-    fig_height: 10
+    # fig_width: 12
+    # fig_height: 10
     # keep_tex: true
@@
 
-## DAE-EDA
+## Denoising Autoencoder Estimation of Distribution Algorithmen (DAE-EDA)
 
@@
 
-## DAE-GP
+## Denoising Autoencoder Genetic Programming (DAE-GP)
 
-* Adaptierung der DAE-EDA Strategie auf GP
+* Adaptierung des DAE-EDA Algorithmus auf GP
+* Linearisierung von GP Individuen als strings in prefix Notation
+* seq2seq learning Problem: Einsatz von DAE - Long Short Term Memory Netzwerken (DAE-LSTM)
+
+## DAE-GP Ablauf
+
+![Flowchart - Regular DAE-GP](./img/flowcharts/dae-gp.png){height=80%}
 
@@
 
+Idee: Modelle werden vor ihrem eigentlichen Einsatz zum Lösen eines Problems auf möglichst großen Datensätzen vortrainiert
+
+Mögliche Vorteile durch Pre-Training^[@pmlr-v5-erhan09a]:
+
+1. Geringere Bedarf an Trainings Daten für vortrainierte Modelle
+2. Reduktion der Trainingszeiten/Laufzeiten
+3. Verbesserung der Güte des Modells
+4. Verbessertes Generalisierungsverhalten des Modells
+
 ## Implementation

### 2023-01-09 paper.Rmd
Mon Jan  9 16:00:17 CET 2023: update-remote.sh

@@
 The pre-training strategy implemented for all experiments in this thesis is visualized as another flowchart in figure §.
-The main difference to regular DAE-GP is the inclusion of an initial pre-training phase where a separate population $\hat{P}$ is first initialized and then randomly split by half into $\hat{P_{train}}$ and $\hat{P_{test}}$.
-A DAE-LSTM model $\hat{M}$ is then trained to learn the properties of $\hat{P_{train}}$ using an early stopping training mode where we stop training as soon as the validation error for $\hat{P_{test}}$ converges.
+The main difference to regular DAE-GP is the inclusion of an initial pre-training phase where a separate population $\hat{P}$ is first initialized and then randomly split by half into $\hat{P}_{train}$ and $\hat{P}_{test}$.
+A DAE-LSTM model $\hat{M}$ is then trained to learn the properties of $\hat{P}_{train}$ using an early stopping training mode where we stop training as soon as the validation error for $\hat{P}_{test}$ converges.
 After the training for $\hat{M}$ is stopped, the current state of $\hat{M}$ is frozen and the optimized trainable parameters $\theta_{\hat{M}}$ are saved before terminating the pre-training phase to start the DAE-GP phase.

### 2023-01-09 slides.Rmd
Mon Jan  9 15:00:17 CET 2023: update-remote.sh
Mon Jan  9 16:00:17 CET 2023: update-remote.sh

@@
 * Variante des Estimation of Distribution-GP (EDA-GP)
+* Der Einsatz einer Pre-Training Strategie wurde in mehreren Publikationen als möglicher Weg für eine weitere Optimierung vorgeschlagen^[@dae-gp_2022_symreg]^[@daegp_explore_exploit)
 
@@
 
+## Pre-Training
+
+Idee: Modelle werden vor ihrem eigentlichen Einsatz zum Lösen eines Problems auf möglichst großen Datensätzen vortrainiert
+
+Mögliche Vorteile durch Pre-Training^[@pmlr-v5-erhan09a]:
+
+1. Geringere Bedarf an Trainings Daten für vortrainierte Modelle
+2. Reduktion der Trainingszeiten/Laufzeiten
+3. Verbesserung der Güte des Modells
+4. Verbessertes Generalisierungsverhalten des Modells
+
 # Aktueller Forschungsstand
@@
 
-## Pre-Training
+## Pre-Training für Denoising Autoencoders^[@pmlr-v5-erhan09a]
 
-Idee: Modelle werden vor ihrem eigentlichen Einsatz zum Lösen eines Problems auf möglichst großen Datensätzen vortrainiert
+### Positive Wirkung von Pre-Training auf DAE
 
-Mögliche Vorteile durch Pre-Training^[@pmlr-v5-erhan09a]:
+1. Gesteigerte Modell Performance (sinkender Testfehler)
+2. Besserer Generalisierungsfähigkeit
+3. Erhöhter Robustness des Algorithmus (sinkende Varianz des Testfehlers)
 
-1. Geringere Bedarf an Trainings Daten für vortrainierte Modelle
-2. Reduktion der Trainingszeiten/Laufzeiten
-3. Verbesserung der Güte des Modells
-4. Verbessertes Generalisierungsverhalten des Modells
+### Einfluss der Modell Architektur
 
-## Implementation
+* Positiver Effekt steigt mit zunehmender Komplexitität des DAE
+* Je mehr versteckte Layer oder die Anzahl an Neuronen pro verstecktem Layer vorhanden sind, desto besserer Effekt des Pre-Trainings
+* Für sehr kleine DAE, zeigt Pre-Training jedoch inverse, negative Auswirkung auf die Modell Performance
+
+# Implementation
+
+## Überblick
+
+Gewählte Pre-Training Strategie:
+
+* (Klassisches) Pre-Training: Einmaliges Training eines DAE-LSTM mit einer großen Population aus Lösungskandidaten $\hat{P}_{train}$
+
+* Trainingsmethode Early Stopping: Abbruch des Trainings sobald der Testfehler für eine seperate Population $\hat{P}_{test}$ konvergiert
+
+**Ausschluss anderer Pre-Training Strategien wie Re-Use Learning, Few-Shot Learning**
+
+## Pre-Trained DAE-GP Ablauf
+
+![Flowchart - Pre-Trained DAE-GP](./img/flowcharts/pt-dae-gp.png){height=75%}
 

### 2023-01-09 paper.Rmd
autogit

@@
 I disregard the fitness of solutions found and focus soley on the reconstruction error that is produced during the training of each the DAE-LSTM.
-The reconstruction error is measured for two separate populations, a training population $train\_pop$ that is used to train our DAE-LSTM as well as a hold-out validation population $test\_pop$.
-For the pre-trained DAE-GP two additional, separate populations $\hat{train\_pop}$ and  $\hat{test\_pop}$ are used only for pre-training.
+The reconstruction error is measured for two separate populations, a training population $P_{train}$ that is used to train our DAE-LSTM as well as a hold-out validation population $P_{test}$.
+For the pre-trained DAE-GP two additional, separate populations $\hat{P}_{train}$ and  $\hat{P}_{test}$ are used exclusively for pre-training.
 
@@
 For each configuration I tested traditional DAE-GP as well as a pre-trained DAE-GP resulting in 12 total sub experiments that were each based on 10 individual runs (total number of runs=$120$).
-To avoid creating biased results through the presence of sampling error inside the pre-training population (see @sampling_err_gp), the population size for the pre-training phase is chosen very high with $\hat{pop\_size}=20000$ where 50% of $\hat{pop}$ is used for the training population $\hat{train\_pop}$ and the remaining 50% are used for the hold-out validation population $\hat{test\_pop}$.
+To avoid creating biased results through the presence of sampling error inside the pre-training population (see @sampling_err_gp), the population size for the pre-training phase is chosen very high with the size of $\hat{P}$ =20000 where 50% of $\hat{P}$ is used for the training population $\hat{P}_{train}$ and the remaining 50% are used for the hold-out validation population $\hat{P}_{test}$.
 The training of DAE-LSTM uses a fixed number of 1000 epochs until termination and uses Adam optimization for gradient descent.
@@
 
-In general I expect that with a growing number of model parameters (either by adding more hidden layers or more hidden neurons per layer) the DAE-LSTM will be more prone to overfit to the training population $train\_pop$ resulting in a small reconstruction error for the training population and a large one for the validation population $train\_pop$.
+In general I expect that with a growing number of model parameters (either by adding more hidden layers or more hidden neurons per layer) the DAE-LSTM will be more prone to overfit to the training population $P_{train}$ resulting in a small reconstruction error for the training population and a large one for the validation population $P_{test}$.
 The reason for this effect is that a larger network trained over an extended period of time (without the use of strategies like early stopping), has much more potential to learn noise from the training dataset than a smaller network, which is more likely to result in worsening performance for previously unseen cases [@weigend1994overfitting].
@@
 
-![First Generation Median Training Error for variable number of hidden Neurons - Airfoil](./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_neurons.png)
-
-![First Generation final Training Error for variable number of hidden Neurons - Airfoil](./img/airfoil_firstGen/airfoil_firstGen_boxplot_training_error_by_neurons.png)
+![First Generation Median Training Error for variable number of hidden Neurons - Airfoil](./img/airfoil_firstGen/airfoil_firstGen_mean_training_error_by_neurons.png)
 
@@
 ![First Generation Median Training Error for variable number of hidden Layers - Airfoil](./img/airfoil_firstGen/airfoil_firstGen_mean_training_error_by_layers.png)
-
-![First Generation final Training Error for variable number of hidden Layers - Airfoil](./img/airfoil_firstGen/airfoil_firstGen_boxplot_training_error_by_layer.png)
 

### 2023-01-09 paper.Rmd
autogit

@@
 bibliography: ref/ref.bib
+link-citations: true
 fig.width: 4in

### 2023-01-09 slides.Rmd
autogit

@@
 fontsize: 11pt
+link-citations: true
 output:
@@
 * Variante des Estimation of Distribution-GP (EDA-GP)
-* Der Einsatz einer Pre-Training Strategie wurde in mehreren Publikationen als möglicher Weg für eine weitere Optimierung vorgeschlagen^[@dae-gp_2022_symreg]^[@daegp_explore_exploit)
+* Der Einsatz einer Pre-Training Strategie wurde in mehreren Publikationen als möglicher Weg für eine weitere Optimierung vorgeschlagen^[@dae-gp_2022_symreg] ^[@daegp_explore_exploit]
 
@@
 
-**Ausschluss anderer Pre-Training Strategien wie Re-Use Learning, Few-Shot Learning**
+*Ausschluss anderer Pre-Training Strategien wie Re-Use Learning, Few-Shot Learning*
 
@@
 ![Flowchart - Pre-Trained DAE-GP](./img/flowcharts/pt-dae-gp.png){height=75%}
+
+## Herausforderungen
+
+... hidden neurons ...
 

### 2023-01-10 paper.Rmd
Tue Jan 10 10:00:09 CET 2023: update-remote.sh

@@
 
+<!-- Overview Real World SymReg  -->
+
+To gather more confidence in the previous results, the pre-training strategy for DAE-GP was tested for various real world symbolic regression problems taken from @machine_learning_repo.
+Are brief summary of the benchmark problems is given in table X
+
+```{r full_run_realWorldSymReg_problems, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
+fig.pos="H"
+knitr::kable(
+  read.csv("./tables/full_run_realWorldSymReg_problems.csv"),
+  caption="Overview Real World Symbolic Regression Benchmark Problems"
+)%>% row_spec(0,bold=TRUE ) %>% kableExtra::kable_styling(latex_options = "hold_position")
+```
+Table X summarizes the median best fitness achieved after running each algorithm for 30 runs using a population size of 500 individuals.
+Statistically significant differences are marked by using asteriks for ($\alpha$) 0.1, 0.05 and 0.01.
+
+```{python createCSV_final_fitness_benchmark}
+DIRS=[
+
+]
+
+```
+
+```{r summarize_final_fitness_benchmarkSymReg, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
+fig.pos="H"
+knitr::kable(
+  read.csv("./tables/full_run_realWorldSymReg_problems.csv"),
+  caption="Overview Real World Symbolic Regression Benchmark Problems"
+)%>% row_spec(0,bold=TRUE ) %>% kableExtra::kable_styling(latex_options = "hold_position")
+```
+
 <!-- Results for Energy Cooling Dataset  -->

### 2023-01-10 data/summarize_finalFitness.py

@@ -1,0 +1,19 @@
+import json
+from scipy.stats import mannwhitneyu
+from statistics import median
+
+RESULTS = {
+    "airfoil_1hl" : "/Users/rmn/github/master_thesis/data/airfoil_1hl_maxIndSize_fullRun_30gens",
+    "airfoil_2hl" : "/Users/rmn/github/master_thesis/data/airfoil_2hl_maxIndSize_fullRun_30gens",
+    "bostonHousing": "/Users/rmn/github/master_thesis/data/bostonHousing_2hl_maxIndSize_fullRun_30gens",
+    "energyCooling": "/Users/rmn/github/master_thesis/data/energyCooling_2hl_FullRun_30gens"
+}
+
+FILENAME="best_final_fitness.json"
+
+
+csv_string = "Problem?Set?Median_RMSE?P_Value?Significance\n"
+
+for problem,path in RESULTS.keys():
+    d = json.load()
+    csv_string += f"{problem},"

### 2023-01-10 paper.Rmd
Tue Jan 10 11:00:11 CET 2023: update-remote.sh

--- 2023-01-09 paper.Rmd
@@
 
-```{python, load_MWU, include=FALSE}
-from json import load
-
-S = load(open("./data/airfoil_FullRun_30gens/MWU-BestFitnessByGen.json", "r", encoding="utf-8"))
-statistic_mwu_fitness = S["statistic"]
-pval_mwu_fitness = round(S["p-value"], 4)
-print(pval_mwu_fitness)
-```
-
-The result of the MWU test is `r py$pval_mwu_fitness`
-
 ![Best Fitness Distribution after 30 Generations - Airfoil](./img/airfoil_FullRun_30gens/final_fit_boxplot.png)
@@
 ![Average Solution Size over 30 Generations - Airfoil](./img/airfoil_FullRun_30gens/mean_Size_byGens.png)
+
+<!-- Overview Real World SymReg  -->
+
+To gather more confidence in the previous results, the pre-training strategy for DAE-GP was tested for various real world symbolic regression problems taken from @machine_learning_repo.
+Are brief summary of the benchmark problems is given in table X
+
+```{r full_run_realWorldSymReg_problems, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
+fig.pos="H"
+knitr::kable(
+  read.csv("./tables/full_run_realWorldSymReg_problems.csv"),
+  caption="Overview - Real World Symbolic Regression Benchmark Problems"
+)%>% row_spec(0,bold=TRUE ) %>% kableExtra::kable_styling(latex_options = "hold_position")
+```
+Table X summarizes the median best fitness achieved after running each algorithm for 30 runs using a population size of 500 individuals.^[P-Values are marked with asteriks (1,2,3) for ($\alpha$) (0.1, 0.05 and 0.01)].
+
+``` {bash recreate_csv_file_benchmarkRWSymReg, include=FALSE}
+/Users/rmn/miniconda3/envs/dataScience/bin/python /Users/rmn/github/master_thesis/data/summarize_finalFitness.py
+```
+
+```{r full_run_realWorldSymReg_performance, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
+fig.pos="H"
+knitr::kable(
+  read.csv("/Users/rmn/github/master_thesis/data/summary_table_final_fit.csv", sep=","),
+  caption="Performance - Real World Symbolic Regression Benchmark Problems"
+)%>% row_spec(0,bold=TRUE ) %>% kableExtra::kable_styling(latex_options = "hold_position")
+```
 

### 2023-01-10 data/summarize_finalFitness.py

@@ -1,6 +1,7 @@
 import json
-from scipy.stats import mannwhitneyu
+from scipy.stats import mannwhitneyu as mwu
 from statistics import median
+from os.path import join
 
 RESULTS = {
     "airfoil_1hl" : "/Users/rmn/github/master_thesis/data/airfoil_1hl_maxIndSize_fullRun_30gens",
@@ -11,9 +12,36 @@ RESULTS = {
 
 FILENAME="best_final_fitness.json"
 
+def getPVal(sampleA, sampleB):
+    stats, pval = mwu(sampleA, sampleB)
+    print(pval)
+    if pval <= 0.01:
+        return f"{pval:.6f}***"
+    elif pval <= 0.05:
+        return f"{pval:.6f}**"
+    elif pval <= 0.1:
+        return f"{pval:.6f}*"
+    else:
+        return f"{pval:.6f}"
 
-csv_string = "Problem?Set?Median_RMSE?P_Value?Significance\n"
 
-for problem,path in RESULTS.keys():
-    d = json.load()
-    csv_string += f"{problem},"
+csv_string = "Problem,Set,DAE-GP,Pre-Trained_DAE-GP,P_Value\n"
+
+for problem,path in RESULTS.items():
+    d = json.load(open(join(path, FILENAME),"r",encoding="utf-8"))
+
+    reg_med_train = median(d["DAE-GP (train)"])
+    pt_med_train = median(d["Pre-Trained (train)"])
+    reg_med_test = median(d["DAE-GP (test)"])
+    pt_med_test = median(d["Pre-Trained (test)"])
+
+    assert d["DAE-GP (train)"] != d["Pre-Trained (train)"]
+    assert d["DAE-GP (test)"] != d["Pre-Trained (test)"]
+
+
+
+    csv_string += f"{problem},Train,{reg_med_train},{pt_med_train},{getPVal(d['DAE-GP (train)'], d['Pre-Trained (train)'])}\n,Test,{reg_med_test},{pt_med_test},{getPVal(d['DAE-GP (test)'], d['Pre-Trained (test)'])}\n"
+
+
+with open("/Users/rmn/github/master_thesis/data/summary_table_final_fit.csv", "w", encoding="utf-8") as f:
+    f.write(csv_string)

### 2023-01-10 paper.Rmd
new evaluations

@@
 ```
-Table X summarizes the median best fitness achieved after running each algorithm for 30 runs using a population size of 500 individuals.^[P-Values are marked with asteriks (1,2,3) for ($\alpha$) (0.1, 0.05 and 0.01)].
+Table X summarizes the median best fitness achieved after running each algorithm for 10 runs using a population size of 500 individuals for 30 generations.^[P-Values are marked with asteriks (1,2,3) for ($\alpha$)-levels of $0.1$, $0.05$ and $0.01$].
 

### 2023-01-10 data/summarize_finalFitness.py

@@ -14,7 +14,6 @@ FILENAME="best_final_fitness.json"
 
 def getPVal(sampleA, sampleB):
     stats, pval = mwu(sampleA, sampleB)
-    print(pval)
     if pval <= 0.01:
         return f"{pval:.6f}***"
     elif pval <= 0.05:
@@ -25,7 +24,7 @@ def getPVal(sampleA, sampleB):
         return f"{pval:.6f}"
 
 
-csv_string = "Problem,Set,DAE-GP,Pre-Trained_DAE-GP,P_Value\n"
+csv_string = "Problem,Hidden_Layer,Dataset,DAE-GP,Pre-Trained_DAE-GP,P_Value\n"
 
 for problem,path in RESULTS.items():
     d = json.load(open(join(path, FILENAME),"r",encoding="utf-8"))
@@ -38,9 +37,11 @@ for problem,path in RESULTS.items():
     assert d["DAE-GP (train)"] != d["Pre-Trained (train)"]
     assert d["DAE-GP (test)"] != d["Pre-Trained (test)"]
 
+    d["hiddenLayer"]
 
 
-    csv_string += f"{problem},Train,{reg_med_train},{pt_med_train},{getPVal(d['DAE-GP (train)'], d['Pre-Trained (train)'])}\n,Test,{reg_med_test},{pt_med_test},{getPVal(d['DAE-GP (test)'], d['Pre-Trained (test)'])}\n"
+
+    csv_string += f"{d['problem']},{d['hiddenLayer']},Train,{reg_med_train},{pt_med_train},{getPVal(d['DAE-GP (train)'], d['Pre-Trained (train)'])}\n,{d['hiddenLayer']},Test,{reg_med_test},{pt_med_test},{getPVal(d['DAE-GP (test)'], d['Pre-Trained (test)'])}\n"
 
 
 with open("/Users/rmn/github/master_thesis/data/summary_table_final_fit.csv", "w", encoding="utf-8") as f:

### 2023-01-10 slides.Rmd
Tue Jan 10 09:00:11 CET 2023: update-remote.sh
new evaluations
Tue Jan 10 15:00:10 CET 2023: update-remote.sh

@@
 
-* (Klassisches) Pre-Training: Einmaliges Training eines DAE-LSTM mit einer großen Population aus Lösungskandidaten $\hat{P}_{train}$
+* (Klassisches) Pre-Training: Einmaliges Pre-Training eines DAE-LSTM mit einer großen Population aus Lösungskandidaten $\hat{P}_{train}$
 
@@
 
-## Symbolische Regression
+## Generalisierungsverhalten
+
+Fragestellung: Welchen Einfluss hat die Dimension der eingesetzten DAE-LSTMs auf das Generalisierungsverhalten von DAE-GP in Kombination mit Pre-Training?
+
+Experimenteller Aufbau:
+
+* Ausschließliche Betrachtung des Rekonstruktionsfehlers der ersten Generation der DAE-GP Läufe  
+* DAE-LSTM Training erfolgt über eine fixe Anzahl von $5000$ Epochen (kein frühzeitiger Abbruch bei Konvergenz des Testfehlers)
+
+Insgesamt 12 Subexperimente mit jeweils 10 Durchläufen ($120$ Gesamtdurchläufe):
+
+* Variable Anzahl von Hidden Neurons (50, 100, 200) mit einem Hidden Layer
+* Variable Anzahl von Hidden Layers (1, 2, 3) mit 100 Hidden Neurons
+* (normales) DAE-GP und pre-trained DAE-GP
+
+## Einfluss der Anzahl von Hidden Layers
+
+![First Generation Median Training Error for variable number of hidden Layers - Airfoil](./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_layers.png){height=75%}
+
+## Einfluss der Anzahl von Hidden Neurons
+
+![First Generation Median Training Error for variable number of hidden Neurons - Airfoil](./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_neurons.png){height=75%}
+
+Anwendung
 

### 2023-01-10 slides.Rmd
Tue Jan 10 16:00:09 CET 2023: update-remote.sh

--- 2023-01-09 slides.Rmd
@@
 ```{r setup, include=FALSE}
+library(kableExtra)
 library(magrittr)
+library(reticulate)
+
 knitr::opts_chunk$set(echo = TRUE)
+#knitr::opts_chunk$set(fig.pos = "!H")
+
+reticulate::use_python("/Users/rmn/miniconda3/envs/dataScience/bin/python")
+
+```
+
+```{python pytest, echo=FALSE, include=FALSE}
+from sys import version, executable
+print(f"Python Version: {version} \nPath: {executable}")
 ```
@@
 
-# Experimente
+# Experiment Generalisierungsverhalten
 
-## Symbolische Regression
+## Fragestellung
+
+Welchen Einfluss hat die Dimension der eingesetzten DAE-LSTMs auf das Generalisierungsverhalten von DAE-GP in Kombination mit Pre-Training?
+
+Experimenteller Aufbau:
+
+* Ausschließliche Betrachtung des Rekonstruktionsfehlers der ersten Generation (Nutzung von seperaten Test- und Trainingspopulationen)
+* DAE-LSTM Training erfolgt über eine fixe Anzahl von $5000$ Epochen (kein frühzeitiger Abbruch bei Konvergenz des Testfehlers)
+
+Insgesamt 12 Subexperimente mit jeweils 10 Durchläufen ($120$ Gesamtdurchläufe):
+
+* Variable Anzahl von Hidden Neurons (50, 100, 200) mit einem Hidden Layer
+* Variable Anzahl von Hidden Layers (1, 2, 3) mit 100 Hidden Neurons
+* (normales) DAE-GP und pre-trained DAE-GP
+
+## Einfluss der Anzahl von Hidden Layers
+
+![Airfoil - Erste Generation Median Trainingsfehler - Variable Anzahl Hidden Layer](./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_layers.png){height=85%}
+
+## Einfluss der Anzahl von Hidden Neurons
+
+![Airfoil - Erste Generation Median Trainingsfehler - Variable Anzahl Hidden Neurons](./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_neurons.png){height=85%}
+
+## Interpretation
+
+# Experiment Suchverhalten Airfoil Datensatz
+
+## Fragestellung
+
+Welchen Einfluss hat die Verwendung einer Pre-Training Strategie auf das Suchverhalten von DAE-GP bei der Anwendung auf dem Airfoil Datensatz für symbolische Regressionsprobleme?
+
+Aufbau: Betrachtung des Suchverhalten über je 10 Gesamtduchläufe für DAE-GP, pre-trained DAE-GP und reguläres GP ($30$ Durchläufe gesamt)
+
+Fokus insbesondere auf:
+
+* Lösungsgüte
+* Größe der gefundenen Lösungen
+* Populationsdiversität
+
+## Hyperparameter
+
+```{r airfoil_fullRun_2hl_maxIndSize_params, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
+fig.pos="H"
+knitr::kable(
+  read.csv("./tables/airfoil_fullRun_2hl_maxIndSize_params"),
+  caption="Hyperparameter"
+)%>% row_spec(0,bold=TRUE ) %>% kableExtra::kable_styling(latex_options = "hold_position")
+```
+## Funktions Set
+
+```{r airfoil_function_set, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
+fig.pos="H"
+knitr::kable(
+  read.csv("./tables/airfoil_function_set"),
+  caption="Funktions Set"
+)%>% row_spec(0,bold=TRUE ) %>% kableExtra::kable_styling(latex_options = "hold_position")
+```
+
+## Lösungsgüte Airfoil
+
+![Airfoil Datensatz - Fitness über Generationen](~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens_withGP/mean_median_fitness_byGens.png)
+
+## Verteilung Lösungsgüte Airfoil
+
+![Airfoil Datensatz - Fitness über Generationen](~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens/final_fit_boxplot.png)
+
+## Lösungsgröße Airfoil
+
+![Airfoil Datensatz - Größe der besten Lösung über Generationen](~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens_withGP/mean_Size_byGens.png)
+
+Anwendung
 

### 2023-01-10 slides.Rmd
Tue Jan 10 17:00:11 CET 2023: update-remote.sh

@@
     # fonttheme: "structurebold"
-    # fig_width: 12
-    # fig_height: 10
+    # fig_width: 7
+    # fig_height: 5
     # keep_tex: true
-    fig_caption: yes
+    fig_caption: no
     number_sections: true
@@
 
-![Airfoil Datensatz - Fitness über Generationen](~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens_withGP/mean_median_fitness_byGens.png)
+![Airfoil Datensatz - Fitness über Generationen](~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens_withGP/mean_median_fitness_byGens.png){height=90%}
 
@@
 
-![Airfoil Datensatz - Fitness über Generationen](~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens/final_fit_boxplot.png)
+![Airfoil Datensatz - Fitness über Generationen](~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens/final_fit_boxplot.png){height=90%}
 
@@
 ![Airfoil Datensatz - Größe der besten Lösung über Generationen](~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens_withGP/mean_Size_byGens.png)
+
+## Kontrollexperiment: Reduzierung der DAE-LSTM Dimension Airfoil
+
+Frage: Welchen Einfluss hat die Reduktion der DAE-LSTM auf 1 hidden Layer?
+
+## Lösungsgüte Airfoil mit einem hidden Layer
+
+![Airfoil Datensatz - 1 HL - Lösungsgüte über Generationen](~/masterThesis/master_thesis/img/airfoil_1hl_maxIndSize_fullRun_30gens/mean_median_fitness_byGens.png)
+
+## Verteilung Lösungsgüte Airfoil mit einem hidden Layer
+
+![Airfoil Datensatz - 1 HL - Lösungsgüte über Generationen](~/masterThesis/master_thesis/img/airfoil_1hl_maxIndSize_fullRun_30gens/final_fit_boxplot.png)
 

### 2023-01-10 slides.Rmd
Tue Jan 10 18:00:07 CET 2023: update-remote.sh

@@
 
-# Experiment Suchverhalten Airfoil Datensatz
+# Experimente Suchverhalten Symbolische Regression
 
@@
 
-Anwendung
+## Anwendung auf weiteren Datensätzen
+
+Airfoil Datensatz: Ergebnis deuten bei einer ausreichenden Anzahl von Hidden Layern auf einen postitiven Effekt der Pre-Training Strategie hin:
+
+* Höhere Fitness der gefundenen Lösungen
+* Kleinere Größe der gefundenen Lösungen
+
+Daher: Ausweitung des Experiments auf weitere Datensätze
+
+## Übersicht Datensätze
+
+```{r full_run_realWorldSymReg_problems, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
+fig.pos="H"
+knitr::kable(
+  read.csv("./tables/full_run_realWorldSymReg_problems.csv"),
+  #caption="Overview - Real World Symbolic Regression Benchmark Problems"
+)%>% row_spec(0,bold=TRUE ) %>% kableExtra::kable_styling(latex_options = "hold_position")
+```
+
+## Zusammenfassung Experimente zur symbolischen Regression
+
+```{r full_run_realWorldSymReg_performance, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
+library(knitr)
+library(kableExtra)
+
+# Create a sample data frame
+df <- read.csv("/Users/rmn/github/master_thesis/data/summary_table_final_fit.csv", sep=",")
+print(df)
+
+# Create a function to format the values
+bold_min <- function(x, y){
+  if(x <= y) return(paste0("**", x, "**"))
+  else return(x)
+}
+
+# Apply the function to the data frame
+df$DAE.GP <- mapply(bold_min, df$DAE_GP, df$Pre_Trained_DAE_GP)
+)
+
+fig.pos="H"
+knitr::kable(
+  df,
+  #caption="Performance - Real World Symbolic Regression Benchmark Problems"
+)%>% row_spec(0,bold=TRUE ) %>% kableExtra::kable_styling(latex_options = "hold_position", font_size = 7)
+```
 

### 2023-01-10 slides.Rmd
Tue Jan 10 19:00:08 CET 2023: update-remote.sh

@@
 df <- read.csv("/Users/rmn/github/master_thesis/data/summary_table_final_fit.csv", sep=",")
-print(df)
 
@@
 # Apply the function to the data frame
-df$DAE.GP <- mapply(bold_min, df$DAE_GP, df$Pre_Trained_DAE_GP)
-)
+df <- mapply(bold_min, df$DAE_GP, df$Pre_Trained_DAE_GP)
+#print(df)
 

### 2023-01-10 slides.Rmd
Tue Jan 10 20:00:09 CET 2023: update-remote.sh

@@
 library(kableExtra)
+library(dplyr)
 
-# Create a sample data frame
+# Create the data frame from csv
 df <- read.csv("/Users/rmn/github/master_thesis/data/summary_table_final_fit.csv", sep=",")
 
-# Create a function to format the values
-bold_min <- function(x, y){
-  if(x <= y) return(paste0("**", x, "**"))
-  else return(x)
+# Create a new dataframe with the modified values of the two columns,
+df_bold <- df
+
+for(i in 1:nrow(df)){
+  if(df[i, "DAE.GP"] < df[i, "Pre.Trained_DAE.GP"])
+    df_bold[i, "DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(df_bold[i, "DAE.GP"]))
+  else if (df[i, "DAE.GP"] > df[i, "Pre.Trained_DAE.GP"])
+    df_bold[i, "Pre.Trained_DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(df_bold[i, "Pre.Trained_DAE.GP"]))
 }
 
-# Apply the function to the data frame
-df <- mapply(bold_min, df$DAE_GP, df$Pre_Trained_DAE_GP)
-#print(df)
+# 
+knitr::kable(
+ df_bold,
+ format = "markdown",
+ digits = 3,
+ caption="Performance - Real World Symbolic Regression Benchmark Problems"
+)
+%>% row_spec(0,bold=FALSE ) %>% kableExtra::kable_styling(latex_options = "hold_position", font_size = 7)
 
-fig.pos="H"
-knitr::kable(
-  df,
-  #caption="Performance - Real World Symbolic Regression Benchmark Problems"
-)%>% row_spec(0,bold=TRUE ) %>% kableExtra::kable_styling(latex_options = "hold_position", font_size = 7)
 ```

### 2023-01-10 paper.Rmd
Tue Jan 10 21:00:08 CET 2023: update-remote.sh

@@
 ```{r full_run_realWorldSymReg_performance, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
-fig.pos="H"
+library(knitr)
+library(kableExtra)
+library(dplyr)
+
+# Create the data frame from csv
+df <- read.csv(
+  "/Users/rmn/github/master_thesis/data/summary_table_final_fit.csv",
+  sep=",",
+  colClasses = c("character", "numeric", "character", "numeric", "numeric", "character")
+)
+
+# Create a new dataframe with the modified values of the two columns,
+df_bold <- df
+
+for(i in 1:nrow(df)){
+  if(df[i, "DAE.GP"] < df[i, "Pre.Trained_DAE.GP"]){
+    df_bold[i, "DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(round(df[i, "DAE.GP"],4)))
+    df_bold[i, "Pre.Trained_DAE.GP"] <- as.character(round(df[i, "Pre.Trained_DAE.GP"],4))
+    }
+  else if (df[i, "DAE.GP"] > df[i, "Pre.Trained_DAE.GP"]) {
+    df_bold[i, "Pre.Trained_DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(round(df[i, "Pre.Trained_DAE.GP"],4)))
+  df_bold[i, "DAE.GP"] <- as.character(round(df[i, "DAE.GP"],4))
+  }
+}
+# shorten column names
+col_names <- c("Problem", "Hid.Layers", "Set", "DAE-GP", "Pre-Trained", "P-Value")
+
 knitr::kable(
-  read.csv("/Users/rmn/github/master_thesis/data/summary_table_final_fit.csv", sep=","),
-  caption="Performance - Real World Symbolic Regression Benchmark Problems"
-)%>% row_spec(0,bold=TRUE ) %>% kableExtra::kable_styling(latex_options = "hold_position")
+ df_bold,
+ format = "markdown",
+ col.names = col_names,
+ align = "c",
+ caption = "Symbolic Regression Overview - Mean final Fitness"
+) %>% row_spec(0,bold=FALSE ) %>% kableExtra::kable_styling(latex_options = "hold_position", font_size = 7)
+
 ```

### 2023-01-10 slides.Rmd
Tue Jan 10 21:00:08 CET 2023: update-remote.sh

@@
 ```
-## Funktions Set
+## Funktions Set Symbolische Regression
 
@@
   read.csv("./tables/airfoil_function_set"),
-  caption="Funktions Set"
+  #caption="Funktions Set"
 )%>% row_spec(0,bold=TRUE ) %>% kableExtra::kable_styling(latex_options = "hold_position")
@@
 
-## Zusammenfassung Experimente zur symbolischen Regression
+## Mean RMSE - Übersicht
 
@@
 # Create the data frame from csv
-df <- read.csv("/Users/rmn/github/master_thesis/data/summary_table_final_fit.csv", sep=",")
+df <- read.csv(
+  "/Users/rmn/github/master_thesis/data/summary_table_final_fit.csv",
+  sep=",",
+  colClasses = c("character", "numeric", "character", "numeric", "numeric", "character")
+)
 
@@
 for(i in 1:nrow(df)){
-  if(df[i, "DAE.GP"] < df[i, "Pre.Trained_DAE.GP"])
-    df_bold[i, "DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(df_bold[i, "DAE.GP"]))
-  else if (df[i, "DAE.GP"] > df[i, "Pre.Trained_DAE.GP"])
-    df_bold[i, "Pre.Trained_DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(df_bold[i, "Pre.Trained_DAE.GP"]))
+  if(df[i, "DAE.GP"] < df[i, "Pre.Trained_DAE.GP"]){
+    df_bold[i, "DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(round(df[i, "DAE.GP"],4)))
+    df_bold[i, "Pre.Trained_DAE.GP"] <- as.character(round(df[i, "Pre.Trained_DAE.GP"],4))
     }
+  else if (df[i, "DAE.GP"] > df[i, "Pre.Trained_DAE.GP"]) {
+    df_bold[i, "Pre.Trained_DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(round(df[i, "Pre.Trained_DAE.GP"],4)))
+  df_bold[i, "DAE.GP"] <- as.character(round(df[i, "DAE.GP"],4))
+  }
+}
+# shorten column names
+col_names <- c("Problem", "Hid.Layers", "Set", "DAE-GP", "Pre-Trained", "P-Value")
 
-# 
 knitr::kable(
@@
  format = "markdown",
- digits = 3,
- caption="Performance - Real World Symbolic Regression Benchmark Problems"
-)
-%>% row_spec(0,bold=FALSE ) %>% kableExtra::kable_styling(latex_options = "hold_position", font_size = 7)
+ col.names = col_names,
+ align = "c"
+ #caption = "Symbolic Regression Overview - Mean final Fitness"
+) %>% row_spec(0,bold=FALSE ) %>% kableExtra::kable_styling(latex_options = "hold_position", font_size = 7)
 
 ```
+## Fitness - Energy Cooling Datensatz
 
-## Generalisierungsverhalten
+![Energy (Cooling) Datensatz - Fitness über Generationen](~/masterThesis/master_thesis/img/energyCooling_2hl_FullRun_30gens/mean_median_fitness_byGens.png)
 
-## Lösungsqualität
+## Fitness - Boston Housing Datensatz
 
-# Resultate
+![Boston Housing Datensatz - Fitness über Generationen](~/masterThesis/master_thesis/img/bostonHousing_2hl_maxIndSize_fullRun_30gens/mean_median_fitness_byGens.png)
+
+## Auswertung
+
+Keine Evidenz für einen statistisch signifikanten Einfluss von Pre-Training auf die erzielte Lösungsqualität in den durchgeführten Experimenten!
 

### 2023-01-10 data/summarize_finalFitness.py

@@ -1,6 +1,6 @@
 import json
 from scipy.stats import mannwhitneyu as mwu
-from statistics import median
+from statistics import median, mean
 from os.path import join
 
 RESULTS = {
@@ -15,13 +15,13 @@ FILENAME="best_final_fitness.json"
 def getPVal(sampleA, sampleB):
     stats, pval = mwu(sampleA, sampleB)
     if pval <= 0.01:
-        return f"{pval:.6f}***"
+        return f"{pval:.4f}***"
     elif pval <= 0.05:
-        return f"{pval:.6f}**"
+        return f"{pval:.4f}**"
     elif pval <= 0.1:
-        return f"{pval:.6f}*"
+        return f"{pval:.4f}*"
     else:
-        return f"{pval:.6f}"
+        return f"{pval:.4f}"
 
 
 csv_string = "Problem,Hidden_Layer,Dataset,DAE-GP,Pre-Trained_DAE-GP,P_Value\n"
@@ -29,16 +29,13 @@ csv_string = "Problem,Hidden_Layer,Dataset,DAE-GP,Pre-Trained_DAE-GP,P_Value\n"
 for problem,path in RESULTS.items():
     d = json.load(open(join(path, FILENAME),"r",encoding="utf-8"))
 
-    reg_med_train = median(d["DAE-GP (train)"])
-    pt_med_train = median(d["Pre-Trained (train)"])
-    reg_med_test = median(d["DAE-GP (test)"])
-    pt_med_test = median(d["Pre-Trained (test)"])
+    reg_med_train = mean(d["DAE-GP (train)"])
+    pt_med_train = mean(d["Pre-Trained (train)"])
+    reg_med_test = mean(d["DAE-GP (test)"])
+    pt_med_test = mean(d["Pre-Trained (test)"])
 
     assert d["DAE-GP (train)"] != d["Pre-Trained (train)"]
     assert d["DAE-GP (test)"] != d["Pre-Trained (test)"]
-
-    d["hiddenLayer"]
-
 
 
     csv_string += f"{d['problem']},{d['hiddenLayer']},Train,{reg_med_train},{pt_med_train},{getPVal(d['DAE-GP (train)'], d['Pre-Trained (train)'])}\n,{d['hiddenLayer']},Test,{reg_med_test},{pt_med_test},{getPVal(d['DAE-GP (test)'], d['Pre-Trained (test)'])}\n"

### 2023-01-10 paper.Rmd
Tue Jan 10 22:00:10 CET 2023: update-remote.sh

@@
 df <- read.csv(
-  "/Users/rmn/github/master_thesis/data/summary_table_final_fit.csv",
+  "/Users/rmn/github/master_thesis/data/summary_table_final_fit_mean.csv",
+  # "/Users/rmn/github/master_thesis/data/summary_table_final_fit_median.csv",  -> Use this path to summarize by median
   sep=",",
@@
  col.names = col_names,
- align = "c",
- caption = "Symbolic Regression Overview - Mean final Fitness"
+ align = "c"
+ #caption = "Symbolic Regression Overview - Mean final Fitness"
 ) %>% row_spec(0,bold=FALSE ) %>% kableExtra::kable_styling(latex_options = "hold_position", font_size = 7)

### 2023-01-10 slides.Rmd
Tue Jan 10 22:00:10 CET 2023: update-remote.sh

@@
 ---
-title: "Pre-Trained Denoising Autoencoders Long Short-Term Memory Networks as probabilistic Models for Estimation of Distribution Genetic Programming"
+#title: "Pre-Trained Denoising Autoencoders Long Short-Term Memory Networks as probabilistic Models for Estimation of Distribution Genetic Programming"
 subtitle: "Kolloquium zur Masterarbeit im M.Sc. Wirtschaftspädagogik "
@@
   beamer_presentation:
-    # color: "beaver"
-    # theme: "Szeged"
-    # fonttheme: "structurebold"
-    # fig_width: 7
-    # fig_height: 5
+    theme: "Madrid"
+    fonttheme: "structurebold"
+    fig_width: 12
+    fig_height: 10
     # keep_tex: true
@@
       in_header: [header_files/slides.tex]
+
 ---

### 2023-01-10 data/summarize_finalFitness.py

@@ -2,6 +2,7 @@ import json
 from scipy.stats import mannwhitneyu as mwu
 from statistics import median, mean
 from os.path import join
+import numpy as np
 
 RESULTS = {
     "airfoil_1hl" : "/Users/rmn/github/master_thesis/data/airfoil_1hl_maxIndSize_fullRun_30gens",
@@ -14,6 +15,9 @@ FILENAME="best_final_fitness.json"
 
 def getPVal(sampleA, sampleB):
     stats, pval = mwu(sampleA, sampleB)
+    assert len(sampleA) > 6 and len(sampleB) > 6
+
+    print(pval)
     if pval <= 0.01:
         return f"{pval:.4f}***"
     elif pval <= 0.05:
@@ -24,15 +28,40 @@ def getPVal(sampleA, sampleB):
         return f"{pval:.4f}"
 
 
+# summarize by mean
 csv_string = "Problem,Hidden_Layer,Dataset,DAE-GP,Pre-Trained_DAE-GP,P_Value\n"
 
 for problem,path in RESULTS.items():
+
     d = json.load(open(join(path, FILENAME),"r",encoding="utf-8"))
 
-    reg_med_train = mean(d["DAE-GP (train)"])
-    pt_med_train = mean(d["Pre-Trained (train)"])
-    reg_med_test = mean(d["DAE-GP (test)"])
-    pt_med_test = mean(d["Pre-Trained (test)"])
+    reg_mean_train = np.mean(d["DAE-GP (train)"])
+    pt_mean_train = np.mean(d["Pre-Trained (train)"])
+    reg_mean_test = np.mean(d["DAE-GP (test)"])
+    pt_mean_test = np.mean(d["Pre-Trained (test)"])
+
+    assert d["DAE-GP (train)"] != d["Pre-Trained (train)"]
+    assert d["DAE-GP (test)"] != d["Pre-Trained (test)"]
+
+
+    csv_string += f"{d['problem']},{d['hiddenLayer']},Train,{reg_mean_train},{pt_mean_train},{getPVal(d['DAE-GP (train)'], d['Pre-Trained (train)'])}\n,{d['hiddenLayer']},Test,{reg_mean_test},{pt_mean_test},{getPVal(d['DAE-GP (test)'], d['Pre-Trained (test)'])}\n"
+
+
+with open("/Users/rmn/github/master_thesis/data/summary_table_final_fit_mean.csv", "w", encoding="utf-8") as f:
+    f.write(csv_string)
[…]

### 2023-01-10 slides.Rmd
autogit

@@
 ---
-#title: "Pre-Trained Denoising Autoencoders Long Short-Term Memory Networks as probabilistic Models for Estimation of Distribution Genetic Programming"
+title: "Pre-Trained Denoising Autoencoders Long Short-Term Memory Networks as probabilistic Models for Estimation of Distribution Genetic Programming"
 subtitle: "Kolloquium zur Masterarbeit im M.Sc. Wirtschaftspädagogik "
@@
   beamer_presentation:
-    theme: "Madrid"
+    #theme: "Madrid"
     fonttheme: "structurebold"
@@
 knitr::opts_chunk$set(echo = TRUE)
-#knitr::opts_chunk$set(fig.pos = "!H")
+knitr::opts_chunk$set(fig.pos = "!H")
 
@@
 
-2 Phase Ansatz:
+2 Phasen Ansatz:
 

### 2023-01-11 data/summarize_finalFitness.py
Wed Jan 11 09:00:09 CET 2023: update-remote.sh

@@ -8,7 +8,8 @@ RESULTS = {
     "airfoil_1hl" : "/Users/rmn/github/master_thesis/data/airfoil_1hl_maxIndSize_fullRun_30gens",
     "airfoil_2hl" : "/Users/rmn/github/master_thesis/data/airfoil_2hl_maxIndSize_fullRun_30gens",
     "bostonHousing": "/Users/rmn/github/master_thesis/data/bostonHousing_2hl_maxIndSize_fullRun_30gens",
-    "energyCooling": "/Users/rmn/github/master_thesis/data/energyCooling_2hl_FullRun_30gens"
+    "energyCooling": "/Users/rmn/github/master_thesis/data/energyCooling_2hl_FullRun_30gens",
+    #"concrete": "/Users/rmn/github/master_thesis/data/concrete_2hl_FullRun_30gens"
 }
 
 FILENAME="best_final_fitness.json"

### 2023-01-11 slides.Rmd
Wed Jan 11 10:00:11 CET 2023: update-remote.sh

@@
 * Metaheuristik basierend auf genetischer Programmierung (GP)
-* Erweiterung von GP durch künstliche, neuronalen Netzen zur Optimierung des Suchverhaltens
+* Ersetzung der Variationsoperatoren von GP durch künstliche, neuronalen Netzen zur Optimierung des Suchverhaltens^[@dae-gp_2020_rtree]
 * Variante des Estimation of Distribution-GP (EDA-GP)
-* Der Einsatz einer Pre-Training Strategie wurde in mehreren Publikationen als möglicher Weg für eine weitere Optimierung vorgeschlagen^[@dae-gp_2022_symreg] ^[@daegp_explore_exploit]
+* Einsatz von Pre-Training in mehreren Publikationen als möglicher Weg für eine weitere Optimierung vorgeschlagen^[@dae-gp_2022_symreg] ^[@daegp_explore_exploit]
 
@@
 * Entwicklung neuer Rekombinationsoperatoren für evolutionäre Algorithmen basierend auf dem Einsatz von probabilistischen Modellen^[@design_of_modern_heuristics]
-* Hypothese: Problemspezifische Abhängigkeiten zwischen einzelnen Entscheidungsvariablen können bei der Erzeugung neuer Lösungskandidaten besser berücksichtigt werden als bei traditionellen Rekombinationsoperatoren^[@edaOrig1996]
+* Hypothese: Problemspezifische Abhängigkeiten zwischen Entscheidungsvariablen können bei der Erzeugung neuer Individuen besser berücksichtigt werden als bei traditionellen Rekombinationsoperatoren^[@edaOrig1996]
 * Weitere Verbreitung im Bereich der genetischen Algorithmen (GA) als für GP
@@
 * Adaptierung des DAE-EDA Algorithmus auf GP
-* Linearisierung von GP Individuen als strings in prefix Notation
+* Darstellung von Individuen als Zeichenketten in prefix Notation
 * seq2seq learning Problem: Einsatz von DAE - Long Short Term Memory Netzwerken (DAE-LSTM)
@@
 
+...
+
 # Experimente Suchverhalten Symbolische Regression
@@
 
-Welchen Einfluss hat die Verwendung einer Pre-Training Strategie auf das Suchverhalten von DAE-GP bei der Anwendung auf dem Airfoil Datensatz für symbolische Regressionsprobleme?
+Welchen Einfluss hat die Verwendung einer Pre-Training Strategie auf das Suchverhalten von DAE-GP bei der Anwendung auf symbolische Regressionsprobleme?
 

### 2023-01-11 slides.Rmd
Wed Jan 11 11:00:10 CET 2023: update-remote.sh

@@
 * Lösungsgüte
-* Größe der gefundenen Lösungen
+* Größe der gefundenen Lösungen (=Anzahl an Knoten)
 * Populationsdiversität
@@
 
-## Mean RMSE - Übersicht
+## Auswertung Fitness - Symbolische Regression
 
-```{r full_run_realWorldSymReg_performance, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
+```{r full_run_realWorldSymReg_fitness, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
 library(knitr)
@@
 
-## Auswertung
+## Zusammenfassung Fitness - Symbolische Regression
 
 Keine Evidenz für einen statistisch signifikanten Einfluss von Pre-Training auf die erzielte Lösungsqualität in den durchgeführten Experimenten!
+
+* Airfoil:
+* Energy Cooling:
+* Boston Housing: 
+* Concrete:
+
+## Auswertung Lösungsgröße - Symbolische Regression
+
+```{r full_run_realWorldSymReg_best_solution_size, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
+library(knitr)
+library(kableExtra)
+library(dplyr)
+
+# Create the data frame from csv
+df <- read.csv(
+  "/Users/rmn/github/master_thesis/data/summary_table_best_size_mean.csv",
+  sep=",",
+  colClasses = c("character", "numeric", "numeric", "numeric", "character")
+)
+
+# Create a new dataframe with the modified values of the two columns,
+df_bold <- df
+
+for(i in 1:nrow(df)){
+  if(df[i, "DAE.GP"] < df[i, "Pre.Trained_DAE.GP"]){
+    df_bold[i, "DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(round(df[i, "DAE.GP"],4)))
+    df_bold[i, "Pre.Trained_DAE.GP"] <- as.character(round(df[i, "Pre.Trained_DAE.GP"],4))
+    }
+  else if (df[i, "DAE.GP"] > df[i, "Pre.Trained_DAE.GP"]) {
+    df_bold[i, "Pre.Trained_DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(round(df[i, "Pre.Trained_DAE.GP"],4)))
+  df_bold[i, "DAE.GP"] <- as.character(round(df[i, "DAE.GP"],4))
+  }
+}
+# shorten column names
+col_names <- c("Problem", "Hid.Layers", "DAE-GP", "Pre-Trained", "P-Value")
+
+knitr::kable(
+ df_bold,
+ format = "markdown",
+ col.names = col_names,
+ align = "c"
+ #caption = "Symbolic Regression Overview - Mean final Fitness"
+) %>% row_spec(0,bold=FALSE ) %>% kableExtra::kable_styling(latex_options = "hold_position", font_size = 7)
+
+```
 

### 2023-01-11 data/summarize_csv_files.py

@@ -1,6 +1,7 @@
 import json
 from scipy.stats import mannwhitneyu as mwu
-from statistics import median, mean
+from scipy.stats import normaltest
+
 from os.path import join
 import numpy as np
 
@@ -12,13 +13,14 @@ RESULTS = {
     #"concrete": "/Users/rmn/github/master_thesis/data/concrete_2hl_FullRun_30gens"
 }
 
+
 FILENAME="best_final_fitness.json"
 
 def getPVal(sampleA, sampleB):
     stats, pval = mwu(sampleA, sampleB)
-    assert len(sampleA) > 6 and len(sampleB) > 6
 
-    print(pval)
+    assert len(sampleA) > 6 and len(sampleB) >= 8
+
     if pval <= 0.01:
         return f"{pval:.4f}***"
     elif pval <= 0.05:
@@ -29,12 +31,26 @@ def getPVal(sampleA, sampleB):
         return f"{pval:.4f}"
 
 
+# Fitness
+# ---
+
 # summarize by mean
 csv_string = "Problem,Hidden_Layer,Dataset,DAE-GP,Pre-Trained_DAE-GP,P_Value\n"
 
 for problem,path in RESULTS.items():
 
     d = json.load(open(join(path, FILENAME),"r",encoding="utf-8"))
+
+    # test for normal distribution ?
+    # ----
+    # reg_train = d["DAE-GP (train)"]
+    # pt_train = d["Pre-Trained (train)"]
+    # reg_test = d["DAE-GP (test)"]
+    # pt_test = d["Pre-Trained (test)"]
+
+    # for vals in (reg_train, pt_train, reg_test, pt_test):
+    #     s, p = normaltest(vals)
+    #     print("p_val:", p)
 
     reg_mean_train = np.mean(d["DAE-GP (train)"])
     pt_mean_train = np.mean(d["Pre-Trained (train)"])
@@ -73,3 +89,29 @@ for problem,path in RESULTS.items():
 
 with open("/Users/rmn/github/master_thesis/data/summary_table_final_fit_median.csv", "w", encoding="utf-8") as f:
     f.write(csv_string)
+
+
+
+FILENAME_SIZE = "size_best_solution.json"
+
+# Solution Size
+# ---
+
+# summarize by mean
+csv_string = "Problem,Hidden_Layer,DAE-GP,Pre-Trained_DAE-GP,P_Value\n"
+
[…]

### 2023-01-11 slides.Rmd
autogit

@@
   read.csv("./tables/airfoil_fullRun_2hl_maxIndSize_params"),
-  caption="Hyperparameter"
+  # caption="Hyperparameter"
 )%>% row_spec(0,bold=TRUE ) %>% kableExtra::kable_styling(latex_options = "hold_position")
@@
   read.csv("./tables/airfoil_function_set"),
-  #caption="Funktions Set"
+  # caption="Funktions Set"
 )%>% row_spec(0,bold=TRUE ) %>% kableExtra::kable_styling(latex_options = "hold_position")

### 2023-01-11 slides.Rmd
Wed Jan 11 13:00:11 CET 2023: update-remote.sh
Wed Jan 11 14:00:08 CET 2023: update-remote.sh

@@
 ```
+*P-Werte sind markiert mit (1,2,3) Sternen für $\alpha$ gleich ($0.1$, $0.05$ and $0.01$), Statistischer Test: Mann-Whitney U [@]*
+
 ## Fitness - Energy Cooling Datensatz
@@
 
-## Auswertung Lösungsgröße - Symbolische Regression
+## Auswertung Mean der Größe der besten Individuen - Symbolische Regression
 
@@
 ```
+## Durchschnittliche Lösungsgröße - Symbolische Regression
+
+\begin{tabular}{cc}
+\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/energyCooling_2hl_maxIndSize_fullRun_30gens/mean_Size_byGens.png} &
+\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/bostonHousing_2hl_maxIndSize_fullRun_30gens/mean_Size_byGens.png} \\
+\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens/mean_Size_byGens.png} &
+\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens/mean_Size_byGens.png}
+\end{tabular}
 

### 2023-01-11 data/summarize_csv_files.py

@@ -1,6 +1,7 @@
 import json
 from scipy.stats import mannwhitneyu as mwu
 from scipy.stats import normaltest
+from vargha_delahni_A import VD_A
 
 from os.path import join
 import numpy as np
@@ -35,7 +36,7 @@ def getPVal(sampleA, sampleB):
 # ---
 
 # summarize by mean
-csv_string = "Problem,Hidden_Layer,Dataset,DAE-GP,Pre-Trained_DAE-GP,P_Value\n"
+csv_string = "Problem,Hidden_Layer,Dataset,DAE-GP,Pre-Trained_DAE-GP,P_Value,VDA\n"
 
 for problem,path in RESULTS.items():
 
@@ -61,7 +62,7 @@ for problem,path in RESULTS.items():
     assert d["DAE-GP (test)"] != d["Pre-Trained (test)"]
 
 
-    csv_string += f"{d['problem']},{d['hiddenLayer']},Train,{reg_mean_train},{pt_mean_train},{getPVal(d['DAE-GP (train)'], d['Pre-Trained (train)'])}\n,{d['hiddenLayer']},Test,{reg_mean_test},{pt_mean_test},{getPVal(d['DAE-GP (test)'], d['Pre-Trained (test)'])}\n"
+    csv_string += f"{d['problem']},{d['hiddenLayer']},Train,{reg_mean_train},{pt_mean_train},{getPVal(d['DAE-GP (train)'], d['Pre-Trained (train)'])},{VD_A(d['DAE-GP (train)'], d['Pre-Trained (train)'])}\n,{d['hiddenLayer']},Test,{reg_mean_test},{pt_mean_test},{getPVal(d['DAE-GP (test)'], d['Pre-Trained (test)'])},{VD_A(d['DAE-GP (test)'], d['Pre-Trained (test)'])}\n"
 
 
 with open("/Users/rmn/github/master_thesis/data/summary_table_final_fit_mean.csv", "w", encoding="utf-8") as f:

### 2023-01-11 data/vargha_delahni_A.py

@@ -1,0 +1,45 @@
+# code taken from https://gist.github.com/jacksonpradolima/f9b19d65b7f16603c837024d5f8c8a65
+
+import itertools as it
+
+from bisect import bisect_left
+from typing import List
+import scipy.stats as ss
+
+from pandas import Categorical
+
+
+def VD_A(treatment: List[float], control: List[float]):
+    """
+    Computes Vargha and Delaney A index
+    A. Vargha and H. D. Delaney.
+    A critique and improvement of the CL common language
+    effect size statistics of McGraw and Wong.
+    Journal of Educational and Behavioral Statistics, 25(2):101-132, 2000
+    The formula to compute A has been transformed to minimize accuracy errors
+    See: http://mtorchiano.wordpress.com/2014/05/19/effect-size-of-r-precision/
+    :param treatment: a numeric list
+    :param control: another numeric list
+    :returns the value estimate and the magnitude
+    """
+    m = len(treatment)
+    n = len(control)
+
+    if m != n:
+        raise ValueError("Data d and f must have the same length")
+
+    r = ss.rankdata(treatment + control)
+    r1 = sum(r[0:m])
+
+    # Compute the measure
+    # A = (r1/m - (m+1)/2)/n # formula (14) in Vargha and Delaney, 2000
+    A = (2 * r1 - m * (m + 1)) / (2 * n * m)  # equivalent formula to avoid accuracy errors
+
+    levels = [0.147, 0.33, 0.474]  # effect sizes from Hess and Kromrey, 2004
+    magnitude = ["negligible", "small", "medium", "large"]
+    scaled_A = (A - 0.5) * 2
+
+    magnitude = magnitude[bisect_left(levels, abs(scaled_A))]
+    estimate = A
+
+    return estimate, magnitude

### 2023-01-11 slides.Rmd
Wed Jan 11 15:00:10 CET 2023: update-remote.sh

@@
 date: "Datum: 25.01.2023"
-fontsize: 11pt
+#fontsize: 11pt
 link-citations: true
@@
 knitr::opts_chunk$set(echo = TRUE)
-knitr::opts_chunk$set(fig.pos = "!H")
+knitr::opts_chunk$set(fig.pos = "H")
 
@@
 ```
-
-## Lösungsgüte Airfoil
+## Airfoil Vergleich GP
 
 ![Airfoil Datensatz - Fitness über Generationen](~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens_withGP/mean_median_fitness_byGens.png){height=90%}
+
+## Ergebnisse - Lösungsgüte Airfoil
+
+\begin{tabular}{c}
+\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens/mean_meadian_fitness_byGens.png} &
+\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens/final_fit_boxplot.png} \\
+
+\end{tabular}
 
@@
   sep=",",
-  colClasses = c("character", "numeric", "character", "numeric", "numeric", "character")
+  colClasses = c("character", "numeric", "character", "numeric", "numeric", "character", "character")
 )
@@
 # shorten column names
-col_names <- c("Problem", "Hid.Layers", "Set", "DAE-GP", "Pre-Trained", "P-Value")
+col_names <- c("Problem", "Hidden-Layers", "Set", "DAE-GP", "Pre-Trained", "P-Value", "Cliffs-Delta")
 
@@
  #caption = "Symbolic Regression Overview - Mean final Fitness"
-) %>% row_spec(0,bold=FALSE ) %>% kableExtra::kable_styling(latex_options = "hold_position", font_size = 7)
+) %>% row_spec(0,bold=FALSE, font_size=7 ) %>% kableExtra::kable_styling(latex_options = "hold_position", font_size=7)
 
 ```
-*P-Werte sind markiert mit (1,2,3) Sternen für $\alpha$ gleich ($0.1$, $0.05$ and $0.01$), Statistischer Test: Mann-Whitney U [@]*
 
@@
   sep=",",
-  colClasses = c("character", "numeric", "numeric", "numeric", "character")
+  colClasses = c("character", "numeric", "numeric", "numeric", "character", "character")
 )
@@
 # shorten column names
-col_names <- c("Problem", "Hid.Layers", "DAE-GP", "Pre-Trained", "P-Value")
+col_names <- c("Problem", "Hid.Layers", "DAE-GP", "Pre-Trained", "P-Value", "Cliffs-Delta")
 
@@
  #caption = "Symbolic Regression Overview - Mean final Fitness"
-) %>% row_spec(0,bold=FALSE ) %>% kableExtra::kable_styling(latex_options = "hold_position", font_size = 7)
+) %>% row_spec(0,bold=FALSE) %>% kableExtra::kable_styling(latex_options = "hold_position", font_size = 7)
 

### 2023-01-11 slides.tex
Wed Jan 11 15:00:10 CET 2023: update-remote.sh

@@
 \fi
-\usetheme[]{Szeged}
-\usecolortheme{beaver}
-\usefonttheme{12}
+\usefonttheme{structurebold}
 % Use upquote if available, for straight quotes in verbatim environments
@@
 \newif\ifbibliography
+\usepackage{longtable,booktabs,array}
+\usepackage{calc} % for calculating minipage widths
+\usepackage{caption}
+% Make caption package work with longtable
+\makeatletter
+\def\fnum@table{\tablename~\thetable}
+\makeatother
+\usepackage{graphicx}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
 \setlength{\emergencystretch}{3em} % prevent overfull lines
@@
 \author{Roman Hoehn}
-\date{Datum: 2023-01-02}
+\date{Datum: 25.01.2023}
 \institute{Johannes Gutenberg-Universität Mainz}
@@
 
-\begin{frame}{Ausblick}
-\protect\hypertarget{ausblick}{}
-\ldots{}
-\end{frame}
-
 \hypertarget{denoising-autoencoder-genetic-programming}{%
@@
 \item
-  Einsatz von künstlichen, neuronalen Netzen zur Optimierung des GP
-  Suchverhaltens
+  Ersetzung der Variationsoperatoren von GP durch künstliche, neuronalen
+  Netzen zur Optimierung des Suchverhaltens\footnote<.->{Wittenberg,
+    Rothlauf and Schweim
+    (\protect\hyperlink{ref-dae-gp_2020_rtree}{2020})}
 \item
   Variante des Estimation of Distribution-GP (EDA-GP)
+\item
+  Einsatz von Pre-Training in mehreren Publikationen als möglicher Weg
+  für eine weitere Optimierung vorgeschlagen\footnote<.->{Wittenberg and
+    Rothlauf (\protect\hyperlink{ref-dae-gp_2022_symreg}{2022})}
+  \footnote<.->{Wittenberg
+    (\protect\hyperlink{ref-daegp_explore_exploit}{2022})}
 \end{itemize}
@@
 
-\begin{frame}{Estimation of Distribution Genetic Programming (EDA-GP)}
-\protect\hypertarget{estimation-of-distribution-genetic-programming-eda-gp}{}
+\begin{frame}{Estimation of Distribution Algorithmen (EDA)}
+\protect\hypertarget{estimation-of-distribution-algorithmen-eda}{}
 \begin{itemize}
@@
 \item
-  Ersetzt die genetische Operatoren Mutation und Rekombination durch den
-  Einsatz von probabilistischen Modellen
+  Entwicklung neuer Rekombinationsoperatoren für evolutionäre
+  Algorithmen basierend auf dem Einsatz von probabilistischen
+  Modellen\footnote<.->{Rothlauf
+    (\protect\hyperlink{ref-design_of_modern_heuristics}{2011})}
+\item
+  Hypothese: Problemspezifische Abhängigkeiten zwischen
+  Entscheidungsvariablen können bei der Erzeugung neuer Individuen
+  besser berücksichtigt werden als bei traditionellen
+  Rekombinationsoperatoren\footnote<.->{Mühlenbein and Paass
+    (\protect\hyperlink{ref-edaOrig1996}{1996})}
+\item
+  Weitere Verbreitung im Bereich der genetischen Algorithmen (GA) als
+  für GP
 \end{itemize}
+\end{frame}
 
-2 Phase Ansatz:
+\begin{frame}{Denoising Autoencoder Estimation of Distribution
+Algorithmen (DAE-EDA)}
+\protect\hypertarget{denoising-autoencoder-estimation-of-distribution-algorithmen-dae-eda}{}
+Idee: Einsatz von Denoising Autoencoders\footnote<.->{Vincent \emph{et
+  al.} (\protect\hyperlink{ref-dae_orig2008}{2008})} (DAE) als
+probabilistisches Modell für genetische Algorithmen \footnote<.->{Probst
+  and Rothlauf (\protect\hyperlink{ref-harmless_overfitting_eda}{2020})}
+
+2 Phasen Ansatz:
 
@@
 \item
-  Model Building:
+  Model Building: Modell ``lernt'' die Eigenschaften von ausgewählten
+  Lösungen hoher Güte durch das Trainieren eines DAEs
 \item
-  Model Sampling:
+  Model Sampling: Neue Lösungen werden erzeugt durch das propagieren von
+  bestehenden, mutierten Lösungen durch das erlernte Modell
 \end{enumerate}
@@
 
-\begin{frame}{Aktueller Forschungsstand}
-\protect\hypertarget{aktueller-forschungsstand}{}
-\begin{block}{Generalisiertes Royal Tree Problem (GRT)\footnote<.->{Wittenberg,
-  Rothlauf and Schweim (2020)}:}
+\begin{frame}{Denoising Autoencoder Genetic Programming (DAE-GP)}
+\protect\hypertarget{denoising-autoencoder-genetic-programming-dae-gp}{}
+\begin{itemize}
+\tightlist
+\item
+  Adaptierung des DAE-EDA Algorithmus auf GP
+\item
+  Darstellung von Individuen als Zeichenketten in prefix Notation
+\item
+  seq2seq learning Problem: Einsatz von DAE - Long Short Term Memory
+  Netzwerken (DAE-LSTM)
+\end{itemize}
+\end{frame}
+
+\begin{frame}{DAE-GP Ablauf}
+\protect\hypertarget{dae-gp-ablauf}{}
+\includegraphics[width=\textwidth,height=0.8\textheight]{./img/flowcharts/dae-gp.png}
+\end{frame}
+
+\begin{frame}{Pre-Training}
+\protect\hypertarget{pre-training}{}
+Idee: Modelle werden vor ihrem eigentlichen Einsatz zum Lösen eines
+Problems auf möglichst großen Datensätzen vortrainiert
+
+Mögliche Vorteile durch Pre-Training\footnote<.->{Erhan \emph{et al.}
+  (\protect\hyperlink{ref-pmlr-v5-erhan09a}{2009})}:
+
+\begin{enumerate}
+\tightlist
+\item
+  Geringere Bedarf an Trainings Daten für vortrainierte Modelle
+\item
+  Reduktion der Trainingszeiten/Laufzeiten
+\item
+  Verbesserung der Güte des Modells
+\item
+  Verbessertes Generalisierungsverhalten des Modells
+\end{enumerate}
+\end{frame}
+
+\hypertarget{aktueller-forschungsstand}{%
+\section{Aktueller Forschungsstand}\label{aktueller-forschungsstand}}
+
+\begin{frame}{Generalisiertes Royal Tree Problem (GRT)\footnote<.->{Wittenberg,
+  Rothlauf and Schweim (\protect\hyperlink{ref-dae-gp_2020_rtree}{2020})}:}
 \protect\hypertarget{generalisiertes-royal-tree-problem-grt}{}
@@
 \end{itemize}
-\end{block}
+\end{frame}
 
-\begin{block}{Symbolische Regression\footnote<.->{Wittenberg and
-  Rothlauf (2022)}:}
+\begin{frame}{Symbolische Regression\footnote<.->{Wittenberg and
+  Rothlauf (\protect\hyperlink{ref-dae-gp_2022_symreg}{2022})}:}
 \protect\hypertarget{symbolische-regression}{}
@@
 
-  1.) Lösungen mit höherer Fitness 2.) Lösungen mit geringerer Größe
+  \begin{enumerate}
+  \tightlist
+  \item
+    Lösungen mit höherer Fitness
+  \item
+    Lösungen mit geringerer Größe
+  \end{enumerate}
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Pre-Training für Denoising Autoencoders\footnote<.->{Erhan
+  \emph{et al.} (\protect\hyperlink{ref-pmlr-v5-erhan09a}{2009})}}
+\protect\hypertarget{pre-training-fuxfcr-denoising-autoencoders}{}
+\begin{block}{Positive Wirkung von Pre-Training auf DAE}
+\protect\hypertarget{positive-wirkung-von-pre-training-auf-dae}{}
+\begin{enumerate}
+\tightlist
+\item
+  Gesteigerte Modell Performance (sinkender Testfehler)
+\item
+  Besserer Generalisierungsfähigkeit
+\item
+  Erhöhter Robustness des Algorithmus (sinkende Varianz des Testfehlers)
+\end{enumerate}
+\end{block}
+
+\begin{block}{Einfluss der Modell Architektur}
+\protect\hypertarget{einfluss-der-modell-architektur}{}
+\begin{itemize}
+\tightlist
+\item
+  Positiver Effekt steigt mit zunehmender Komplexitität des DAE
+\item
+  Je mehr versteckte Layer oder die Anzahl an Neuronen pro verstecktem
+  Layer vorhanden sind, desto besserer Effekt des Pre-Trainings
+\item
+  Für sehr kleine DAE, zeigt Pre-Training jedoch inverse, negative
+  Auswirkung auf die Modell Performance
 \end{itemize}
@@
 
-\begin{frame}{Pre-Training}
-\protect\hypertarget{pre-training}{}
+\hypertarget{implementation}{%
+\section{Implementation}\label{implementation}}
+
+\begin{frame}{Überblick}
+\protect\hypertarget{uxfcberblick}{}
+Gewählte Pre-Training Strategie:
+
+\begin{itemize}
+\item
+  (Klassisches) Pre-Training: Einmaliges Pre-Training eines DAE-LSTM mit
+  einer großen Population aus Lösungskandidaten \(\hat{P}_{train}\)
+\item
+  Trainingsmethode Early Stopping: Abbruch des Trainings sobald der
+  Testfehler für eine seperate Population \(\hat{P}_{test}\) konvergiert
+\end{itemize}
+
+\emph{Ausschluss anderer Pre-Training Strategien wie Re-Use Learning,
+Few-Shot Learning}
 \end{frame}
 
-\begin{frame}{Implementation}
-\protect\hypertarget{implementation}{}
+\begin{frame}{Pre-Trained DAE-GP Ablauf}
+\protect\hypertarget{pre-trained-dae-gp-ablauf}{}
+\includegraphics[width=\textwidth,height=0.75\textheight]{./img/flowcharts/pt-dae-gp.png}
 \end{frame}
 
-\hypertarget{experimente}{%
-\section{Experimente}\label{experimente}}
-
-\begin{frame}{Symbolische Regression}
-\protect\hypertarget{symbolische-regression-1}{}
+\begin{frame}{Herausforderungen}
+\protect\hypertarget{herausforderungen}{}
+\ldots{} hidden neurons \ldots{}
 \end{frame}
 
-\begin{frame}{Generalisierungsverhalten}
-\protect\hypertarget{generalisierungsverhalten}{}
+\hypertarget{experiment-generalisierungsverhalten}{%
+\section{Experiment
+Generalisierungsverhalten}\label{experiment-generalisierungsverhalten}}
+
+\begin{frame}{Fragestellung}
+\protect\hypertarget{fragestellung}{}
+Welchen Einfluss hat die Dimension der eingesetzten DAE-LSTMs auf das
+Generalisierungsverhalten von DAE-GP in Kombination mit Pre-Training?
+
+Experimenteller Aufbau:
+
+\begin{itemize}
+\tightlist
+\item
+  Ausschließliche Betrachtung des Rekonstruktionsfehlers der ersten
+  Generation (Nutzung von seperaten Test- und Trainingspopulationen)
+\item
+  DAE-LSTM Training erfolgt über eine fixe Anzahl von \(5000\) Epochen
+  (kein frühzeitiger Abbruch bei Konvergenz des Testfehlers)
+\end{itemize}
+
+Insgesamt 12 Subexperimente mit jeweils 10 Durchläufen (\(120\)
+Gesamtdurchläufe):
+
+\begin{itemize}
+\tightlist
+\item
+  Variable Anzahl von Hidden Neurons (50, 100, 200) mit einem Hidden
+  Layer
+\item
+  Variable Anzahl von Hidden Layers (1, 2, 3) mit 100 Hidden Neurons
+\item
+  (normales) DAE-GP und pre-trained DAE-GP
+\end{itemize}
 \end{frame}
 
-\begin{frame}{Lösungs}
-\protect\hypertarget{luxf6sungs}{}
+\begin{frame}{Einfluss der Anzahl von Hidden Layers}
+\protect\hypertarget{einfluss-der-anzahl-von-hidden-layers}{}
+\includegraphics[width=\textwidth,height=0.85\textheight]{./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_layers.png}
 \end{frame}
+
+\begin{frame}{Einfluss der Anzahl von Hidden Neurons}
+\protect\hypertarget{einfluss-der-anzahl-von-hidden-neurons}{}
+\includegraphics[width=\textwidth,height=0.85\textheight]{./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_neurons.png}
+\end{frame}
+
+\begin{frame}{Interpretation}
+\protect\hypertarget{interpretation}{}
+\ldots{}
+\end{frame}
+
+\hypertarget{experimente-suchverhalten-symbolische-regression}{%
+\section{Experimente Suchverhalten Symbolische
+Regression}\label{experimente-suchverhalten-symbolische-regression}}
+
+\begin{frame}{Fragestellung}
+\protect\hypertarget{fragestellung-1}{}
+Welchen Einfluss hat die Verwendung einer Pre-Training Strategie auf das
+Suchverhalten von DAE-GP bei der Anwendung auf symbolische
+Regressionsprobleme?
+
+Aufbau: Betrachtung des Suchverhalten über je 10 Gesamtduchläufe für
+DAE-GP, pre-trained DAE-GP und reguläres GP (\(30\) Durchläufe gesamt)
+
+Fokus insbesondere auf:
+
+\begin{itemize}
+\tightlist
+\item
+  Lösungsgüte
+\item
+  Größe der gefundenen Lösungen (=Anzahl an Knoten)
+\item
+  Populationsdiversität
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Hyperparameter}
+\protect\hypertarget{hyperparameter}{}
+\begin{table}[!h]
+\centering
+\begin{tabular}{l|l}
+\hline
+\textbf{parameter} & \textbf{value}\\
+\hline
+populationSize & 500\\
+\hline
+generations & 30\\
+\hline
+fitness & RMSE\\
+\hline
+TrainingMode & Convergence\\
+\hline
+SamplingSteps & 1\\
+\hline
+hiddenLayers & 2\\
+\hline
+Selection & Binary Tournament Selection\\
+\hline
+Pre-Training PopulationSize(Training/Validation) & 10000/100000\\
+\hline
+Pre-Training TrainingMode & Early Stopping\\
+\hline
+\end{tabular}
+\end{table}
+\end{frame}
+
+\begin{frame}{Funktions Set Symbolische Regression}
+\protect\hypertarget{funktions-set-symbolische-regression}{}
+\begin{table}[!h]
+\centering
+\begin{tabular}{l|r}
+\hline
+\textbf{function.} & \textbf{arity}\\
+\hline
+addition & 2\\
+\hline
+subtraction & 2\\
+\hline
+multiplication & 2\\
+\hline
+analytic\_quotient & 2\\
+\hline
+\end{tabular}
+\end{table}
+\end{frame}
+
+\begin{frame}{Airfoil Vergleich GP}
+\protect\hypertarget{airfoil-vergleich-gp}{}
+\includegraphics[width=\textwidth,height=0.9\textheight]{~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens_withGP/mean_median_fitness_byGens.png}
+\end{frame}
+
+\begin{frame}{Ergebnisse - Lösungsgüte Airfoil}
+\protect\hypertarget{ergebnisse---luxf6sungsguxfcte-airfoil}{}
+\begin{tabular}{c}
+\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens/mean_meadian_fitness_byGens.png} &
+\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens/final_fit_boxplot.png} \\
+
+\end{tabular}
+\end{frame}
+
+\begin{frame}{Verteilung Lösungsgüte Airfoil}
+\protect\hypertarget{verteilung-luxf6sungsguxfcte-airfoil}{}
+\includegraphics[width=\textwidth,height=0.9\textheight]{~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens/final_fit_boxplot.png}
+\end{frame}
+
+\begin{frame}{Lösungsgröße Airfoil}
+\protect\hypertarget{luxf6sungsgruxf6uxdfe-airfoil}{}
+\includegraphics{~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens_withGP/mean_Size_byGens.png}
+\end{frame}
+
+\begin{frame}{Kontrollexperiment: Reduzierung der DAE-LSTM Dimension
+Airfoil}
+\protect\hypertarget{kontrollexperiment-reduzierung-der-dae-lstm-dimension-airfoil}{}
+Frage: Welchen Einfluss hat die Reduktion der DAE-LSTM auf 1 hidden
+Layer?
+\end{frame}
+
+\begin{frame}{Lösungsgüte Airfoil mit einem hidden Layer}
+\protect\hypertarget{luxf6sungsguxfcte-airfoil-mit-einem-hidden-layer}{}
+\includegraphics{~/masterThesis/master_thesis/img/airfoil_1hl_maxIndSize_fullRun_30gens/mean_median_fitness_byGens.png}
+\end{frame}
+
+\begin{frame}{Verteilung Lösungsgüte Airfoil mit einem hidden Layer}
+\protect\hypertarget{verteilung-luxf6sungsguxfcte-airfoil-mit-einem-hidden-layer}{}
+\includegraphics{~/masterThesis/master_thesis/img/airfoil_1hl_maxIndSize_fullRun_30gens/final_fit_boxplot.png}
+\end{frame}
+
+\begin{frame}{Anwendung auf weiteren Datensätzen}
+\protect\hypertarget{anwendung-auf-weiteren-datensuxe4tzen}{}
+Airfoil Datensatz: Ergebnis deuten bei einer ausreichenden Anzahl von
+Hidden Layern auf einen postitiven Effekt der Pre-Training Strategie
+hin:
+
+\begin{itemize}
+\tightlist
+\item
+  Höhere Fitness der gefundenen Lösungen
+\item
+  Kleinere Größe der gefundenen Lösungen
+\end{itemize}
+
+Daher: Ausweitung des Experiments auf weitere Datensätze
+\end{frame}
+
+\begin{frame}{Übersicht Datensätze}
+\protect\hypertarget{uxfcbersicht-datensuxe4tze}{}
+\begin{table}[!h]
+\centering
+\begin{tabular}{l|r|r}
+\hline
+\textbf{Problem} & \textbf{Observations} & \textbf{Features}\\
+\hline
+Airfoil & 1503 & 5\\
+\hline
+Boston\_Housing & 506 & 13\\
+\hline
+Energy\_Cooling & 768 & 8\\
+\hline
+Concrete & 1030 & 8\\
+\hline
+\end{tabular}
+\end{table}
+\end{frame}
+
+\begin{frame}{Auswertung Fitness - Symbolische Regression}
+\protect\hypertarget{auswertung-fitness---symbolische-regression}{}
+\begin{longtable}[]{@{}
+  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1910}}
+  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1685}}
+  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.0787}}
+  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1461}}
+  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1461}}
+  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1124}}
+  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1573}}@{}}
+\toprule()
+\begin{minipage}[b]{\linewidth}\centering
+Problem
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+Hidden-Layers
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+Set
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+DAE-GP
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+Pre-Trained
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+P-Value
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+Cliffs-Delta
+\end{minipage} \\
+\midrule()
+\endhead
+Airfoil & 1 & Train & \textbf{33.3674} & 35.3454 & 0.0171** & 0.64 \\
+& 1 & Test & \textbf{33.7271} & 35.612 & 0.1400 & 0.4 \\
+Airfoil & 2 & Train & 17.4762 & \textbf{14.7233} & 0.3073 & -0.28 \\
+& 2 & Test & 17.5299 & \textbf{14.7603} & 0.5706 & -0.16 \\
+Boston\_Housing & 2 & Train & 8.1922 & \textbf{8.0763} & 0.2894 &
+-0.29 \\
+& 2 & Test & 8.0501 & \textbf{7.9692} & 0.7051 & -0.11 \\
+Energy(Cooling) & 2 & Train & \textbf{4.5271} & 4.7782 & 0.0188** &
+0.63 \\
+& 2 & Test & \textbf{4.67} & 4.9196 & 0.2886 & 0.29 \\
+\bottomrule()
+\end{longtable}
+\end{frame}
+
+\begin{frame}{Fitness - Energy Cooling Datensatz}
+\protect\hypertarget{fitness---energy-cooling-datensatz}{}
+\includegraphics{~/masterThesis/master_thesis/img/energyCooling_2hl_FullRun_30gens/mean_median_fitness_byGens.png}
+\end{frame}
+
+\begin{frame}{Fitness - Boston Housing Datensatz}
+\protect\hypertarget{fitness---boston-housing-datensatz}{}
+\includegraphics{~/masterThesis/master_thesis/img/bostonHousing_2hl_maxIndSize_fullRun_30gens/mean_median_fitness_byGens.png}
+\end{frame}
+
+\begin{frame}{Zusammenfassung Fitness - Symbolische Regression}
+\protect\hypertarget{zusammenfassung-fitness---symbolische-regression}{}
+Keine Evidenz für einen statistisch signifikanten Einfluss von
+Pre-Training auf die erzielte Lösungsqualität in den durchgeführten
+Experimenten!
+
+\begin{itemize}
+\tightlist
+\item
+  Airfoil:
+\item
+  Energy Cooling:
+\item
+  Boston Housing:
+\item
+  Concrete:
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Auswertung Mean der Größe der besten Individuen -
+Symbolische Regression}
+\protect\hypertarget{auswertung-mean-der-gruxf6uxdfe-der-besten-individuen---symbolische-regression}{}
+\begin{longtable}[]{@{}
+  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2152}}
+  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1519}}
+  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1646}}
+  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1646}}
+  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1266}}
+  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1772}}@{}}
+\toprule()
+\begin{minipage}[b]{\linewidth}\centering
+Problem
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+Hid.Layers
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+DAE-GP
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+Pre-Trained
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+P-Value
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+Cliffs-Delta
+\end{minipage} \\
+\midrule()
+\endhead
+Airfoil & 1 & \textbf{33.3674} & 35.3454 & 0.0171** & 0.64 \\
+Airfoil & 2 & 17.4762 & \textbf{14.7233} & 0.3073 & -0.28 \\
+Boston\_Housing & 2 & 8.1922 & \textbf{8.0763} & 0.2894 & -0.29 \\
+Energy(Cooling) & 2 & \textbf{4.5271} & 4.7782 & 0.0188** & 0.63 \\
+\bottomrule()
+\end{longtable}
+\end{frame}
+
+\begin{frame}{Durchschnittliche Lösungsgröße - Symbolische Regression}
+\protect\hypertarget{durchschnittliche-luxf6sungsgruxf6uxdfe---symbolische-regression}{}
+\begin{tabular}{cc}
+\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/energyCooling_2hl_maxIndSize_fullRun_30gens/mean_Size_byGens.png} &
+\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/bostonHousing_2hl_maxIndSize_fullRun_30gens/mean_Size_byGens.png} \\
+\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens/mean_Size_byGens.png} &
+\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens/mean_Size_byGens.png}
+\end{tabular}
+\end{frame}
+
+\hypertarget{weitere-ansuxe4tze}{%
+\section{Weitere Ansätze}\label{weitere-ansuxe4tze}}
 
@@
 \begin{CSLReferences}{0}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-pmlr-v5-erhan09a}{}}%
+Erhan, D. \emph{et al.} (2009) {`The difficulty of training deep
+architectures and the effect of unsupervised pre-training'}, in D. van
+Dyk and M. Welling (eds) \emph{Proceedings of the twelth international
+conference on artificial intelligence and statistics}. Hilton Clearwater
+Beach Resort, Clearwater Beach, Florida USA: PMLR (Proceedings of
+machine learning research), pp. 153--160. Available at:
+\url{https://proceedings.mlr.press/v5/erhan09a.html}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-edaOrig1996}{}}%
+Mühlenbein, H. and Paass, G. (1996) {`From recombination of genes to the
+estimation of distributions i. Binary parameters.'}, in \emph{From
+Recombination of Genes to the Estimation of Distributions I. Binary
+Parameters}, pp. 178--187.
+
+\leavevmode\vadjust pre{\hypertarget{ref-harmless_overfitting_eda}{}}%
+Probst, M. and Rothlauf, F. (2020) {`Harmless overfitting: Using
+denoising autoencoders in estimation of distribution algorithms'},
+\emph{Journal of Machine Learning Research}, 21(78), pp. 1--31.
+Available at: \url{http://jmlr.org/papers/v21/16-543.html}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-design_of_modern_heuristics}{}}%
+Rothlauf, F. (2011) \emph{Design of modern heuristics: Principles and
+application}, \emph{Natural Computing Series}. Available at:
+\url{https://doi.org/10.1007/978-3-540-72962-4}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-dae_orig2008}{}}%
+Vincent, P. \emph{et al.} (2008) {`Extracting and composing robust
+features with denoising autoencoders'}, in \emph{Proceedings of the 25th
+International Conference on Machine Learning}, pp. 1096--1103. Available
+at: \url{https://doi.org/10.1145/1390156.1390294}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-daegp_explore_exploit}{}}%
+Wittenberg, D. (2022) {`Using denoising autoencoder genetic programming
+to~control exploration and~exploitation in~search'}, in E. Medvet, G.
+Pappa, and B. Xue (eds) \emph{Genetic programming}. Cham: Springer
+International Publishing, pp. 102--117.
+
 \leavevmode\vadjust pre{\hypertarget{ref-dae-gp_2022_symreg}{}}%

### 2023-01-11 data/summarize_csv_files.py

@@ -2,6 +2,7 @@ import json
 from scipy.stats import mannwhitneyu as mwu
 from scipy.stats import normaltest
 from vargha_delahni_A import VD_A
+from cliffs_delta import cliffs_delta
 
 from os.path import join
 import numpy as np
@@ -16,6 +17,10 @@ RESULTS = {
 
 
 FILENAME="best_final_fitness.json"
+
+def cliffsDeltaPretty(*args):
+    val,desc = cliffs_delta(*args)
+    return f"{val}"
 
 def getPVal(sampleA, sampleB):
     stats, pval = mwu(sampleA, sampleB)
@@ -36,7 +41,7 @@ def getPVal(sampleA, sampleB):
 # ---
 
 # summarize by mean
-csv_string = "Problem,Hidden_Layer,Dataset,DAE-GP,Pre-Trained_DAE-GP,P_Value,VDA\n"
+csv_string = "Problem,Hidden_Layer,Dataset,DAE-GP,Pre-Trained_DAE-GP,P_Value,Cliffs_Delta\n"
 
 for problem,path in RESULTS.items():
 
@@ -62,7 +67,7 @@ for problem,path in RESULTS.items():
     assert d["DAE-GP (test)"] != d["Pre-Trained (test)"]
 
 
-    csv_string += f"{d['problem']},{d['hiddenLayer']},Train,{reg_mean_train},{pt_mean_train},{getPVal(d['DAE-GP (train)'], d['Pre-Trained (train)'])},{VD_A(d['DAE-GP (train)'], d['Pre-Trained (train)'])}\n,{d['hiddenLayer']},Test,{reg_mean_test},{pt_mean_test},{getPVal(d['DAE-GP (test)'], d['Pre-Trained (test)'])},{VD_A(d['DAE-GP (test)'], d['Pre-Trained (test)'])}\n"
+    csv_string += f"{d['problem']},{d['hiddenLayer']},Train,{reg_mean_train},{pt_mean_train},{getPVal(d['DAE-GP (train)'], d['Pre-Trained (train)'])},{cliffsDeltaPretty(d['Pre-Trained (train)'], d['DAE-GP (train)'])}\n,{d['hiddenLayer']},Test,{reg_mean_test},{pt_mean_test},{getPVal(d['DAE-GP (test)'], d['Pre-Trained (test)'])},{cliffsDeltaPretty(d['Pre-Trained (test)'], d['DAE-GP (test)'])}\n"
 
 
[…]

### 2023-01-11 data/vargha_delahni_A.py

@@ -42,4 +42,4 @@ def VD_A(treatment: List[float], control: List[float]):
     magnitude = magnitude[bisect_left(levels, abs(scaled_A))]
     estimate = A
 
-    return estimate, magnitude
+    return f"{estimate}({magnitude})"

### 2023-01-11 slides.Rmd
Wed Jan 11 16:00:14 CET 2023: update-remote.sh

@@
 ```
-## Airfoil Vergleich GP
+## Airfoil Übersicht
 
@@
 
-## Ergebnisse - Lösungsgüte Airfoil
-
-\begin{tabular}{c}
-\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens/mean_meadian_fitness_byGens.png} &
-\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens/final_fit_boxplot.png} \\
-
-\end{tabular}
-
 ## Verteilung Lösungsgüte Airfoil
 
-![Airfoil Datensatz - Fitness über Generationen](~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens/final_fit_boxplot.png){height=90%}
+![Airfoil Datensatz - Fitness über Generationen](~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens_withGP/final_fit_boxplot.png){height=90%}
 
@@
 
-![Airfoil Datensatz - 1 HL - Lösungsgüte über Generationen](~/masterThesis/master_thesis/img/airfoil_1hl_maxIndSize_fullRun_30gens/final_fit_boxplot.png)
+![Airfoil Datensatz - 1 HL - Lösungsgüte über Generationen](~/masterThesis/master_thesis/img/airfoil_1hl_maxIndSize_fullRun_30gens/final_fit_boxplot.png){height=90%}
 

### 2023-01-11 data/summarize_csv_files.py
Wed Jan 11 17:00:11 CET 2023: update-remote.sh

@@ -12,7 +12,7 @@ RESULTS = {
     "airfoil_2hl" : "/Users/rmn/github/master_thesis/data/airfoil_2hl_maxIndSize_fullRun_30gens",
     "bostonHousing": "/Users/rmn/github/master_thesis/data/bostonHousing_2hl_maxIndSize_fullRun_30gens",
     "energyCooling": "/Users/rmn/github/master_thesis/data/energyCooling_2hl_FullRun_30gens",
-    #"concrete": "/Users/rmn/github/master_thesis/data/concrete_2hl_FullRun_30gens"
+    "concrete": "/Users/rmn/github/master_thesis/data/concrete_2hl_FullRun_30gens"
 }
 
 
@@ -25,7 +25,7 @@ def cliffsDeltaPretty(*args):
 def getPVal(sampleA, sampleB):
     stats, pval = mwu(sampleA, sampleB)
 
-    assert len(sampleA) > 6 and len(sampleB) >= 8
+    #assert len(sampleA) > 6 and len(sampleB) >= 8
 
     if pval <= 0.01:
         return f"{pval:.4f}***"

### 2023-01-11 paper.Rmd
Wed Jan 11 18:00:11 CET 2023: update-remote.sh

--- 2023-01-10 paper.Rmd
@@
 
+```{bash executeFirstGenNBs, include=FALSE}
+bash /Users/rmn/masterThesis/eda-gp-2020/experiments/airfoil_firstGen/re_evaluate_airfoilFirstGen.sh
+```
+
 <!--- analyze results 1 hidden layer and variable number of hidden neurons -->
 
-![First Generation Median Training Error for variable number of hidden Neurons - Airfoil](./img/airfoil_firstGen/airfoil_firstGen_mean_training_error_by_neurons.png)
+![First Generation Median Training Error for variable number of hidden Neurons - Airfoil](./img/airfoil_firstGen/airfoil_firstGen_mean_training_error_by_neurons_6plots.png)
 
@@
 
-![First Generation Median Training Error for variable number of hidden Layers - Airfoil](./img/airfoil_firstGen/airfoil_firstGen_mean_training_error_by_layers.png)
+![First Generation Median Training Error for variable number of hidden Layers - Airfoil](./img/airfoil_firstGen/airfoil_firstGen_mean_training_error_by_layers_6plot.png)
 
@@
 
-# add new experiments for full runs here!
+# concrete
+bash /Users/rmn/masterThesis/eda-gp-2020/experiments/concrete_2hl_maxIndSize_fullRun_30gens/re_evaluate_concrete_2hl_maxIndSize_fullRun_30gens.sh
+
 # ...
@@
 ``` {bash recreate_csv_file_benchmarkRWSymReg, include=FALSE}
-/Users/rmn/miniconda3/envs/dataScience/bin/python /Users/rmn/github/master_thesis/data/summarize_finalFitness.py
+/Users/rmn/miniconda3/envs/dataScience/bin/python /Users/rmn/github/master_thesis/data/summarize_csv_files.py
 ```
 
-```{r full_run_realWorldSymReg_performance, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
-fig.pos="H"
+```{r full_run_realWorldSymReg_fitness, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
+library(knitr)
+library(kableExtra)
+library(dplyr)
+
+# Create the data frame from csv
+df <- read.csv(
+  "/Users/rmn/github/master_thesis/data/summary_table_final_fit_mean.csv",
+  # "/Users/rmn/github/master_thesis/data/summary_table_final_fit_median.csv",  -> Use this path to summarize by median
+  sep=",",
+  colClasses = c("character", "numeric", "character", "numeric", "numeric", "character", "character")
+)
+
+# Create a new dataframe with the modified values of the two columns,
+df_bold <- df
+
+for(i in 1:nrow(df)){
+  if(df[i, "DAE.GP"] < df[i, "Pre.Trained_DAE.GP"]){
+    df_bold[i, "DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(round(df[i, "DAE.GP"],4)))
+    df_bold[i, "Pre.Trained_DAE.GP"] <- as.character(round(df[i, "Pre.Trained_DAE.GP"],4))
+    }
+  else if (df[i, "DAE.GP"] > df[i, "Pre.Trained_DAE.GP"]) {
+    df_bold[i, "Pre.Trained_DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(round(df[i, "Pre.Trained_DAE.GP"],4)))
+  df_bold[i, "DAE.GP"] <- as.character(round(df[i, "DAE.GP"],4))
+  }
+}
+# shorten column names
+col_names <- c("Problem", "Hidden-Layers", "Set", "DAE-GP", "Pre-Trained", "P-Value", "Cliffs-Delta")
+
 knitr::kable(
-  read.csv("/Users/rmn/github/master_thesis/data/summary_table_final_fit.csv", sep=","),
-  caption="Performance - Real World Symbolic Regression Benchmark Problems"
-)%>% row_spec(0,bold=TRUE ) %>% kableExtra::kable_styling(latex_options = "hold_position")
+ df_bold,
+ format = "markdown",
+ col.names = col_names,
+ align = "c"
+ #caption = "Symbolic Regression Overview - Mean final Fitness"
+) %>% row_spec(0,bold=TRUE ) %>% kableExtra::kable_styling(latex_options = "hold_position")
+
 ```
 
-<!-- Results for Energy Cooling Dataset  -->
+## Solution Size in Symbolic Regression
 
-```{bash compile_energyCooling_2hl_maxIndSize_fullRun_30gens, include=FALSE}
-bash /Users/rmn/masterThesis/eda-gp-2020/experiments/energyCooling_2hl_maxIndSize_fullRun_30gens/re_evaluate_energyCooling_2hl_maxIndSize_fullRun_30gens.sh
+<!-- Overview Table  -->
+
+```{r full_run_realWorldSymReg_best_solution_size, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
+library(knitr)
+library(kableExtra)
+library(dplyr)
+
+# Create the data frame from csv
+df <- read.csv(
+  "/Users/rmn/github/master_thesis/data/summary_table_best_size_mean.csv",
+  sep=",",
+  colClasses = c("character", "numeric", "numeric", "numeric", "character", "character")
+)
+
+# Create a new dataframe with the modified values of the two columns,
+df_bold <- df
+
+for(i in 1:nrow(df)){
+  if(df[i, "DAE.GP"] < df[i, "Pre.Trained_DAE.GP"]){
+    df_bold[i, "DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(round(df[i, "DAE.GP"],4)))
+    df_bold[i, "Pre.Trained_DAE.GP"] <- as.character(round(df[i, "Pre.Trained_DAE.GP"],4))
+    }
+  else if (df[i, "DAE.GP"] > df[i, "Pre.Trained_DAE.GP"]) {
+    df_bold[i, "Pre.Trained_DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(round(df[i, "Pre.Trained_DAE.GP"],4)))
+  df_bold[i, "DAE.GP"] <- as.character(round(df[i, "DAE.GP"],4))
+  }
+}
+# shorten column names
+col_names <- c("Problem", "Hid.Layers", "DAE-GP", "Pre-Trained", "P-Value", "Cliffs-Delta")
+
+knitr::kable(
+ df_bold,
+ format = "markdown",
+ col.names = col_names,
+ align = "c"
+ #caption = "Symbolic Regression Overview - Mean final Fitness"
+) %>% row_spec(0,bold=FALSE) %>% kableExtra::kable_styling(latex_options = "hold_position", font_size = 7)
+
 ```
+
+<!-- Overview Plots  -->
+
+\begin{tabular}{cc}
+\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/energyCooling_2hl_maxIndSize_fullRun_30gens/mean_Size_byGens.png} &
+\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/bostonHousing_2hl_maxIndSize_fullRun_30gens/mean_Size_byGens.png} \\
+\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens/mean_Size_byGens.png} &
+\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/concrete_2hl_maxIndSize_fullRun_30gens/mean_Size_byGens.png}
+\end{tabular}
 

### 2023-01-11 paper.Rmd
Wed Jan 11 19:00:10 CET 2023: update-remote.sh

@@
 ## Pre-Training
-
-<!--- 
-chatGPT
----
-
-Q: what is pre-training in deep learning?
-A: Pre-training in deep learning refers to the process of training a deep learning model on a large dataset before using it for a specific task.
-The goal of pre-training is to create a model that has a strong general understanding of the data and can be fine-tuned for a specific task.
-
-Pre-training can be done in two ways: unsupervised pre-training and supervised pre-training.
-In unsupervised pre-training, the model is trained on a large dataset without labels, using techniques such as self-supervised learning or generative adversarial networks (GANs) to learn the underlying patterns in the data.
-In supervised pre-training, the model is trained on a large dataset with labels, using a supervised learning algorithm to learn the relationship between the input data and the corresponding labels.
-
-Pre-training has several benefits.
-It can help improve the performance of the model on the target task by providing it with a better initial set of weights.
-It can also speed up training by reducing the amount of data and computational resources needed to train the model from scratch.
-Additionally, pre-trained models can be used as a starting point for transfer learning, allowing you to leverage the knowledge and experience of the pre-trained model for a new task.
-
-Overall, pre-training is an important technique in deep learning that can help improve model performance and efficiency.
-It is commonly used in natural language processing and computer vision applications, where large amounts of labeled data are available.
--->
 
@@
 
-```{bash compile_airfoil_2hl_maxIndSize_fullRun_30gens, include=FALSE}
+```{bash compile_symReg_fullRun_30gens, include=FALSE}
 
@@
 
-![Best Fitness over 30 Generations - Airfoil](./img/airfoil_FullRun_30gens/mean_median_fitness_byGens.png)
+![Best Fitness over 30 Generations - Airfoil](./img/airfoil_2hl_maxIndSize_fullRun_30gens_withGP/mean_median_fitness_byGens.png)
 
-![Best Fitness Distribution after 30 Generations - Airfoil](./img/airfoil_FullRun_30gens/final_fit_boxplot.png)
+![Best Fitness Distribution after 30 Generations - Airfoil](./img/airfoil_2hl_maxIndSize_fullRun_30gens/final_fit_boxplot.png)
 lalal
@@
 ![Average Solution Size over 30 Generations - Airfoil](./img/airfoil_FullRun_30gens/mean_Size_byGens.png)
+
+The experiments was repeated using the same setup with the only difference being a reduction from two hidden layers to 1 hidden layer.
+
+![Best Fitness over 30 Generations - 1 hidden Layer - Airfoil](./img/airfoil_1hl_maxIndSize_fullRun_30gens/mean_median_fitness_byGens.png)
+
+...
 
@@
 ![Statutory Declaration](./private/erklaerung.pdf)\
-
 -->

### 2023-01-11 paper.tex
Wed Jan 11 19:00:10 CET 2023: update-remote.sh

@@
 \usepackage[margin=1in]{geometry}
+\usepackage{longtable,booktabs,array}
+\usepackage{calc} % for calculating minipage widths
+% Correct order of tables after \paragraph or \subparagraph
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
+\makeatother
+% Allow footnotes in longtable head/foot
+\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
+\makesavenoteenv{longtable}
 \usepackage{graphicx}
@@
 \author{}
-\date{\vspace{-2.5em}Date of Submission: 2022-12-29}
+\date{\vspace{-2.5em}Date of Submission: 2023-01-11}
 
@@
 network (DAE-LSTMs) as a probabilistic model to sample new candidate
-solutions {[}18{]}.
+solutions {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]}.
 
@@
 offspring solutions by propagating partially mutated solutions from the
-parent population through the DAE-LSTMS {[}18{]}. In previous work
+parent population through the DAE-LSTMS
+{[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]}. In previous work
 DAE-GP has been shown to outperform GP for both a generalized version of
-the royal tree problem {[}18{]} as well as for a real-world symbolic
-regression problem {[}17{]}.
+the royal tree problem
+{[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]} as well as for a
+real-world symbolic regression problem
+{[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]}.
 
-The DAE-GP algorithm first described by {[}18{]} trains a DAE-LSTMs for
-each generation \(g\) of the search from scratch. {[}17{]} and {[}16{]}
-suggests the incorporation of a pre-training strategy into the
-evolutionary search as a possible way of improving the model quality of
-DAE-GP. The key idea is to pre-train an initial DAE-LSTM on a large
-population of candidate solutions and to use the weights of this initial
-model in each generation of the search as a starting point. This thesis
-studies the influence of using pre-trained DAE-LSTMs in DAE-GP for
-symbolic regression, especially looking at the influence in
-generalization behavior and the overall quality of solutions found by
-DAE-GP. The aim of this study is to answer the question if a
-pre-training strategy can help to improve generalization in DAE-GP and
-if so, analyze how this improved generalization ability benefits the
-overall performance of DAE-GP.
+The DAE-GP algorithm first described by
+{[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]} trains a DAE-LSTMs
+for each generation \(g\) of the search from scratch.
+{[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]} and
+{[}\protect\hyperlink{ref-daegp_explore_exploit}{16}{]} suggests the
+incorporation of a pre-training strategy into the evolutionary search as
+a possible way of improving the model quality of DAE-GP. The key idea is
+to pre-train an initial DAE-LSTM on a large population of candidate
+solutions and to use the pre-trained parameters of this initial model in
+each generation of the search as starting parameters for the current
+generation DAE-LSTM. This thesis studies the influence of using
+pre-trained DAE-LSTMs in DAE-GP for symbolic regression, especially
+looking at the influence in generalization behavior and the overall
+quality of solutions found by DAE-GP. The aim of this study is to answer
+the question if a pre-training strategy can help to improve
+generalization in DAE-GP and if so, analyze how this improved
+generalization ability benefits the overall performance of DAE-GP.
 
@@
 DAE-GP is an EDA-GP algorithm that uses DAE-LSTM networks as a
-probabilistic model to sample new offspring solutions {[}18{]}.
+probabilistic model to sample new offspring solutions
+{[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]}.
 
@@
 quality of solutions. The representation of a solution is therefore
-defined by the mapping of genotypes to phenotypes {[}12{]}. Genetic
+defined by the mapping of genotypes to phenotypes
+{[}\protect\hyperlink{ref-design_of_modern_heuristics}{12}{]}. Genetic
 operators, such as mutation or recombination are usually applied to the
@@
 more general, hierarchical computer programs of dynamically varying size
-and shape {[}8{]}. The computer programs that are at the center of the
-evolutionary search in GP are commonly represented by tree structures at
-the level of their phenotype {[}12{]}. Since GP searches for high
-fitness computer programs that produce a desired output for some input,
-it can be applied to various different problem domains such as symbolic
-regression, automatic programming, or evolving game-playing strategies
-{[}8{]}. An important quality of GP is the ability to search for
-solutions of variable length and structure. GP is an especially useful
-meta heuristic for problems where no a priori knowledge about the final
-form of good solutions is available. GPs ability to optimize solutions
-for their structure as well as for their parameters led to it being one
-of the most prevalent methods used symbolic regression {[}10{]}.
+and shape {[}\protect\hyperlink{ref-Koza1993GeneticP}{8}{]}. The
+computer programs that are at the center of the evolutionary search in
+GP are commonly represented by tree structures at the level of their
+phenotype {[}\protect\hyperlink{ref-design_of_modern_heuristics}{12}{]}.
+Since GP searches for high fitness computer programs that produce a
+desired output for some input, it can be applied to various different
+problem domains such as symbolic regression, automatic programming, or
+evolving game-playing strategies
+{[}\protect\hyperlink{ref-Koza1993GeneticP}{8}{]}. An important quality
+of GP is the ability to search for solutions of variable length and
+structure. GP is an especially useful meta heuristic for problems where
+no a priori knowledge about the final form of good solutions is
+available. GPs ability to optimize solutions for their structure as well
+as for their parameters led to it being one of the most prevalent
+methods used for symbolic regression
+{[}\protect\hyperlink{ref-10.1007ux2f978-3-540-24621-3_22}{10}{]}.
 
@@
 that can capture complex dependencies between different decision
-variables of an optimization problem {[}12{]}. EDAs use this
-probabilistic model to sample new offspring solutions inside an
+variables of an optimization problem
+{[}\protect\hyperlink{ref-design_of_modern_heuristics}{12}{]}. EDAs use
+this probabilistic model to sample new offspring solutions inside an
 evolutionary search to replace crossover and/or mutation operators.
@@
 
-One possible way of model building in EDA proposed by {[}11{]} is to use
+One possible way of model building in EDA proposed by
+{[}\protect\hyperlink{ref-harmless_overfitting_eda}{11}{]} is to use
 denoising autoencoders (DAEs) as generative models to capture complex
-probability distributions of decision variables.
+probability distributions of decision variables. DAE, a variation of the
+autoencoders (AE), are a widely used type of neural networks in the
+domain of unsupervized machine learning that maps \(n\) input variables
+to \(n\) output variables using a hidden representation.
 
-DAE, a variation of the autoencoders (AE), are a widely used type of
-neural networks in the domain of unsupervized machine learning that maps
-\(n\) input variables to \(n\) output variables using a hidden
-representation.
-
-AE were introduced by {[}6{]} to compress high-dimensional data into
-lower-dimensions. An AE consists of two different subunits:
+AE were introduced by {[}\protect\hyperlink{ref-ae_orig}{6}{]} to
+compress high-dimensional data into lower-dimensions. An AE consists of
+two different subunits:
 
@@
 able to reduce the dimensionality of input data to a compressed
-representation{[}6{]}.
+representation{[}\protect\hyperlink{ref-ae_orig}{6}{]}.
 
-DAE was first introduced by {[}14{]} as an improved AE with the ability
-to learn new representations of data that is especially robust to
-partially corrupted input data. DAE modifies the AE by using partially
-corrupted input data for the AE and training it to reconstruct the
-uncorrupted, original version of the input data.
+DAE was first introduced by
+{[}\protect\hyperlink{ref-10.1145ux2f1390156.1390294}{14}{]} as an
+improved AE with the ability to learn new representations of data that
+is especially robust to partially corrupted input data. DAE modifies the
+AE by using partially corrupted input data for the AE and training it to
+reconstruct the uncorrupted, original version of the input data.
 
@@
 of the input data it can therefore also be used to generate new
-solutions in the context of GAs {[}11{]}.
+solutions in the context of GAs
+{[}\protect\hyperlink{ref-harmless_overfitting_eda}{11}{]}.
 
 DAE-GP builds upon the concept of using DAEs in EDAs described by
-{[}11{]} and transfers the concept to the domain of GP. The mutation and
-crossover operators of standard GP are replaced by sampling new
-solutions from a probabilistic model that is build by training a DAEs
-long short-term memory (LSTM) network on a subset of high fitness
-solutions from the current population {[}18{]}.
+{[}\protect\hyperlink{ref-harmless_overfitting_eda}{11}{]} and transfers
+the concept to the domain of GP. The mutation and crossover operators of
+standard GP are replaced by sampling new solutions from a probabilistic
+model that is build by training a DAEs long short-term memory (LSTM)
+network on a subset of high fitness solutions from the current
+population {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]}.
 
 LSTMs are a variant of recurrent neural networks first introduced by
-{[}7{]} that can store learned information over en extended time period
-while avoiding the problem of vanishing and/or exploding gradients.
-Since DAE-GP encodes candidate solutions as linear strings in prefix
-notation, the DAE in DAE-GP uses LSTMs for both encoding and decoding
-where the total amount of time steps \(T\) is equal to sum of the length
-of the input solution and the output solution {[}18{]}.
+{[}\protect\hyperlink{ref-lstm_orig}{7}{]} that can store learned
+information over en extended time period while avoiding the problem of
+vanishing and/or exploding gradients. Since DAE-GP encodes candidate
+solutions as linear strings in prefix notation, the DAE in DAE-GP uses
+LSTMs for both encoding and decoding where the total amount of time
+steps \(T\) is equal to sum of the length of the input solution and the
+output solution {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]}.
 
@@
 generalization behavior if compared to the standard approach of using
-randomly initialized parameters {[}3{]}.
+randomly initialized parameters
+{[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{3}{]}.
 
@@
 idea of transfer learning that aims at retaining previously learned
-knowledge from one task to re-use it for another task {[}5{]} {[}9{]}.
-Transfer learning traditionally follows a two phase approach:
+knowledge from one task to re-use it for another task
+{[}\protect\hyperlink{ref-HAN2021225}{5}{]}
+{[}\protect\hyperlink{ref-survey_transfer_learning}{9}{]}. Transfer
+learning traditionally follows a two phase approach:
 
@@
 where source task and target task are usually similar but may differ in
-their feature space and the distribution of training data {[}9{]}.
+their feature space and the distribution of training data
+{[}\protect\hyperlink{ref-survey_transfer_learning}{9}{]}.
 
@@
 \(g\in{1,...,g_{max}}\). Instead of initializing and optimizing each
-model from scratch as done in previous work (e.g. {[}18{]} {[}17{]}), a
-separate DAE-LSTM network \(\hat{M}\) will be trained on a large
-population of randomly initialized solutions.\\
-The optimal parameters obtained after finishing the training procedure
-of \(\hat{M}\) are then used as the starting parameters for each
-following DAE-LSTM model \(M_g\) for \(g\in{1,...,g_{max}}\).
+model from scratch as done in previous work (e.g.
+{[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]}
+{[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]}), a separate
+DAE-LSTM network \(\hat{M}\) will be trained on a large population of
+randomly initialized solutions.\\
+The trainable parameters \(\theta_{\hat{M}}\) obtained after finishing
+the training procedure of \(\hat{M}\) are then used as the starting
+parameters for each following DAE-LSTM model \(M_g\) for
+\(g\in{1,...,g_{max}}\).
 
@@
 in DAE-GP is based on the fact, that initializing \(M_g\) with
-pre-trained weights from \(\hat{M}\) is likely to reduce the amount of
-training epochs that are necessary until the point of training error
+pre-trained weights \(\theta_{\hat{M}}\) is likely to reduce the amount
+of training epochs that are necessary until the point of training error
 convergence is reached.
 
-Early research on pre-training for DAE by {[}3{]} comes to the
+Early research on pre-training for DAE by
+{[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{3}{]} comes to the
 conclusion that besides adding robustness to models the strategy also
@@
 
-Another important finding by {[}3{]} is that the positive effect of
-pre-training described above is dependent on both the depth of the
-network as well as the size of layers - while increasingly larger
+Another important finding by
+{[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{3}{]} is that the positive
+effect of pre-training described above is dependent on both the depth of
+the network as well as the size of layers - while increasingly larger
 networks benefit increasingly more from pre-training the final
@@
 Another motivation for introducing pre-training into DAE-GP is the
-prevalence of sampling error for small GP populations sizes. {[}13{]}
-finds that sampling error, non-systematic errors that are caused by
-observing only a small subset of the statistical population, is a severe
-problem in the domain of GP since the initial population may not be a
-representative sample of all possible solutions. {[}13{]} also
-introduces a method for calculating optimal population sizes to minimize
-the presence of sampling error. Since DAE-LSTM of DAE-GP learns the
-properties of it's training population and reuses the acquired knowledge
-in the sampling procedure, using the parameters obtained from training
+prevalence of sampling error for small GP populations sizes.
+{[}\protect\hyperlink{ref-sampling_err_gp}{13}{]} finds that sampling
+error, non-systematic errors that are caused by observing only a small
+subset of the statistical population, is a severe problem in the domain
+of GP since the initial population may not be a representative sample of
+all possible solutions.
+{[}\protect\hyperlink{ref-sampling_err_gp}{13}{]} also introduces a
+method for calculating optimal population sizes to minimize the presence
+of sampling error. Since DAE-LSTM of DAE-GP learns the properties of
+it's training population and reuses the acquired knowledge in the
+sampling procedure, using the parameters obtained from training
 \(\hat{M}\) on a sufficiently large training population might reduce the
@@
 
-The DAE-GP algorithm, first described by {[}18{]}, is summarized by
-figure 1.
+The DAE-GP algorithm, first described by
+{[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]}, is summarized as a
+flowchart in figure §. After setting the generation counter \(g\) to
+\(0\) and creating the initial population of solutions \(P_0\), DAE-GP
+enters the main loop that checks if a termination criteria is satisfied
+(i.e.~maximum number of generations or fitness evaluations is reached).
+For each generation \(g\) the fitness of all new individuals in \(P_g\)
+is evaluated. During model building, DAE-GP selects a subset \(X_g\) of
+the current population that is used as training data for the DAE-LSTM
+model \(M_g\) which is trained to learn the properties of \(X_g\). The
+trained model \(M_g\) is then used for model sampling, partially mutated
+solutions from \(P_G\) are propagated through \(M_g\) to create a new
+population \(P_{g+1}\). This process is repeated until the termination
+criteria is met and DAE-GP returns the highest fitness solutions.
 
@@
 \centering
-\includegraphics[width=2.08333in,height=8.33333in]{./img/flowcharts/dae-gp.png}
+\includegraphics{./img/flowcharts/dae-gp.png}
 \caption{Flowchart - Regular DAE-GP}
@@
 
-The pre-training strategy implemented is visualized by figure 2.
+The pre-training strategy implemented for all experiments in this thesis
+is visualized as another flowchart in figure §. The main difference to
+regular DAE-GP is the inclusion of an initial pre-training phase where a
+separate population \(\hat{P}\) is first initialized and then randomly
+split by half into \(\hat{P}_{train}\) and \(\hat{P}_{test}\). A
+DAE-LSTM model \(\hat{M}\) is then trained to learn the properties of
+\(\hat{P}_{train}\) using an early stopping training mode where we stop
+training as soon as the validation error for \(\hat{P}_{test}\)
+converges. After the training for \(\hat{M}\) is stopped, the current
+state of \(\hat{M}\) is frozen and the optimized trainable parameters
+\(\theta_{\hat{M}}\) are saved before terminating the pre-training phase
+to start the DAE-GP phase. During the DAE-GP phase, the only difference
+to the traditional DAE-GP algorithm described in figure § is during
+model building: We initialize \(M_g\) with \(\theta_{\hat{M}}\) before
+training it to learn the properties of \(X_g\).
 
@@
 \centering
-\includegraphics[width=2.08333in,height=8.33333in]{./img/flowcharts/pt-dae-gp.png}
+\includegraphics{./img/flowcharts/pt-dae-gp.png}
 \caption{Flowchart - Pre-Trained DAE-GP}
@@
 DAE-GP has been the detemination of the number of hidden neurons in side
-the DAE-LSTMs hidden layers. The original description of DAE-GP {[}18{]}
-used a strategy where the number of hidden neurons for each hidden layer
-is dynamically set per generation to the maximum individual size inside
-the current population. For a pre-training implementation, this strategy
-can not be easily adapted since it leads to a changing number of neurons
-at each generation resulting in different dimensions of the DAE-LSTM. To
-allow for the sharing of weights and biases from the pre-trained model
+the DAE-LSTMs hidden layers. The original description of DAE-GP
+{[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]} used a strategy
+where the number of hidden neurons for each hidden layer is dynamically
+set per generation to the maximum individual size inside the current
+population. For a pre-training implementation, this strategy can not be
+easily adapted since it leads to a changing number of neurons at each
+generation resulting in different dimensions of the DAE-LSTM. To allow
+for the sharing of weights and biases from the pre-trained model
 \(\hat{M}\) to each \(M_g\) for \(g\in{1,...,g_{max}}\) I experimented
@@
 programming language in conjunction with the baseline libraries
-\texttt{Keras} {[}1{]} for deep neural networks and \texttt{deap}
-{[}4{]} for evolutionary computation.
+\texttt{Keras} {[}\protect\hyperlink{ref-chollet2015keras}{1}{]} for
+deep neural networks and \texttt{deap}
+{[}\protect\hyperlink{ref-DEAP_JMLR2012}{4}{]} for evolutionary
+computation.
 
@@
 real-world symbolic regression problems. Symbolic Regression problems
-have been one of the first GP applications {[}8{]} and are an actively
+have been one of the first GP applications
+{[}\protect\hyperlink{ref-Koza1993GeneticP}{8}{]} and are an actively
 studied and highly relevant research area. The goal in symbolic
 regression is to find a mathematical model for a given set of data
-points {[}10{]}, in real-world symbolic regression these data points are
-sourced from real-world observations which in contrast to synthetic
-symbolic regression problem are more likely to contain random noise and
-bias. Another important challenge in solving real-world symbolic
-regression problems is the ability for a given model to generalize, we
-want the final model to show high accuracy in predicting outcomes for
-previously unseen cases.
+points
+{[}\protect\hyperlink{ref-10.1007ux2f978-3-540-24621-3_22}{10}{]}, in
+real-world symbolic regression these data points are sourced from
+real-world observations which in contrast to synthetic symbolic
+regression problem are more likely to contain random noise and bias.
+Another important challenge in solving real-world symbolic regression
+problems is the ability for a given model to generalize, we want the
+final model to show high accuracy in predicting outcomes for previously
+unseen cases.
 
@@
 Self-Noise Data Set which is part of the UCI machine learning repository
-{[}2{]}. The dataset consists of 5 input variables and 1 output variable
-that are listed in table 1.
+{[}\protect\hyperlink{ref-machine_learning_repo}{2}{]}. The dataset
+consists of 5 input variables and 1 output variable that are listed in
+table 1.
 
@@
 terminal set consists of the 5 input variables \(x1,x2,x3,x4,x5\) and
-ephemeral random integers in the range of \([{}-5,..,5]\) {[}17{]}.
+ephemeral random integers in the range of \([{}-5,..,5]\)
+{[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]}.
 
@@
 DAE-LSTM. The reconstruction error is measured for two separate
-populations, a training population \(train\_pop\) that is used to train
-our DAE-LSTM as well as a hold-out validation population \(test\_pop\).
+populations, a training population \(P_{train}\) that is used to train
+our DAE-LSTM as well as a hold-out validation population \(P_{test}\).
 For the pre-trained DAE-GP two additional, separate populations
-\(\hat{train\_pop}\) and \(\hat{test\_pop}\) are used only for
+\(\hat{P}_{train}\) and \(\hat{P}_{test}\) are used exclusively for
 pre-training.
@@
 creating biased results through the presence of sampling error inside
-the pre-training population (see {[}13{]}), the population size for the
-pre-training phase is chosen very high with \(\hat{pop\_size}=20000\)
-where 50\% of \(\hat{pop}\) is used for the training population
-\(\hat{train\_pop}\) and the remaining 50\% are used for the hold-out
-validation population \(\hat{test\_pop}\). The training of DAE-LSTM uses
-a fixed number of 1000 epochs until termination and uses Adam
-optimization for gradient descent. The reason for using a high amount of
-1000 fixed training epochs is to deliberately force the DAE-LSTM to
-overfit to the training data.
+the pre-training population (see
+{[}\protect\hyperlink{ref-sampling_err_gp}{13}{]}), the population size
+for the pre-training phase is chosen very high with the size of
+\(\hat{P}\) =20000 where 50\% of \(\hat{P}\) is used for the training
+population \(\hat{P}_{train}\) and the remaining 50\% are used for the
+hold-out validation population \(\hat{P}_{test}\). The training of
+DAE-LSTM uses a fixed number of 1000 epochs until termination and uses
+Adam optimization for gradient descent. The reason for using a high
+amount of 1000 fixed training epochs is to deliberately force the
+DAE-LSTM to overfit to the training data.
 
@@
 the DAE-LSTM will be more prone to overfit to the training population
-\(train\_pop\) resulting in a small reconstruction error for the
-training population and a large one for the validation population
-\(train\_pop\). The reason for this effect is that a larger network
-trained over an extended period of time (without the use of strategies
-like early stopping), has much more potential to learn noise from the
-training dataset than a smaller network, which is more likely to result
-in worsening performance for previously unseen cases {[}15{]}.
+\(P_{train}\) resulting in a small reconstruction error for the training
+population and a large one for the validation population \(P_{test}\).
+The reason for this effect is that a larger network trained over an
+extended period of time (without the use of strategies like early
+stopping), has much more potential to learn noise from the training
+dataset than a smaller network, which is more likely to result in
+worsening performance for previously unseen cases
+{[}\protect\hyperlink{ref-weigend1994overfitting}{15}{]}.
 
-Based on the review of {[}3{]} I also expect that pre-training will have
-an insignificant or even negative influence on small DAE-LSTM instances
-while improving the networks generalization ability with growing size.
+Based on the review of {[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{3}{]}
+I also expect that pre-training will have an insignificant or even
+negative influence on small DAE-LSTM instances while improving the
+networks generalization ability with growing size.
 
@@
 \centering
-\includegraphics[width=2.08333in,height=8.33333in]{./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_neurons.png}
+\includegraphics{./img/airfoil_firstGen/airfoil_firstGen_mean_training_error_by_neurons_6plots.png}
 \caption{First Generation Median Training Error for variable number of
@@
 \centering
-\includegraphics[width=2.08333in,height=8.33333in]{./img/airfoil_firstGen/airfoil_firstGen_boxplot_training_error_by_neurons.png}
-\caption{First Generation final Training Error for variable number of
-hidden Neurons - Airfoil}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=2.08333in,height=8.33333in]{./img/airfoil_firstGen/airfoil_firstGen_mean_training_error_by_layers.png}
+\includegraphics{./img/airfoil_firstGen/airfoil_firstGen_mean_training_error_by_layers_6plot.png}
 \caption{First Generation Median Training Error for variable number of
-hidden Layers - Airfoil}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=2.08333in,height=8.33333in]{./img/airfoil_firstGen/airfoil_firstGen_boxplot_training_error_by_layer.png}
-\caption{First Generation final Training Error for variable number of
 hidden Layers - Airfoil}
@@
 DAE-GP. Again the Airfoil problem was selected as a first benchmark
-problem to test pre-trained DAE-GP against traditional DAE-GP {[}17{]},
-the parameters are described in table \(A\)
+problem to test pre-trained DAE-GP against traditional DAE-GP
+{[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]}, the parameters are
+described in table \(A\)
 
@@
 \hline
-SamplingSteps & 2\\
+SamplingSteps & 1\\
 \hline
@@
 two different strategies: Regular DAE-GP uses the same strategy as
-described by earlier work (see {[}17{]} or {[}18{]}), the number of
+described by earlier work (see
+{[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]} or
+{[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]}), the number of
 hidden neurons is dynamically set at each generation to the maximum
@@
 \centering
-\includegraphics[width=2.08333in,height=8.33333in]{./img/airfoil_FullRun_30gens/mean_median_fitness_byGens.png}
+\includegraphics{./img/airfoil_2hl_maxIndSize_fullRun_30gens_withGP/mean_median_fitness_byGens.png}
 \caption{Best Fitness over 30 Generations - Airfoil}
@@
 
-The result of the MWU test is 0
+\includegraphics{./img/airfoil_2hl_maxIndSize_fullRun_30gens/final_fit_boxplot.png}
+lalal
+
+\includegraphics{./img/airfoil_FullRun_30gens/mean_popDiversity_byGens.png}
+Since {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]} shows, that
+DAE-GP finds higher quality that are also smaller in size than those
+found by average GP for a given number of 10.000 fitness evaluations, I
+also studied the differences in average solution size between
+pre-trained and regular DAE-GP.
 
@@
 \centering
-\includegraphics[width=2.08333in,height=8.33333in]{./img/airfoil_FullRun_30gens/final_fit_boxplot.png}
-\caption{Best Fitness Distribution after 30 Generations - Airfoil}
+\includegraphics{./img/airfoil_FullRun_30gens/mean_Size_byGens.png}
+\caption{Average Solution Size over 30 Generations - Airfoil}
 \end{figure}
 
-\includegraphics[width=2.08333in,height=8.33333in]{./img/airfoil_FullRun_30gens/mean_popDiversity_byGens.png}
-Since {[}17{]} shows, that DAE-GP finds higher quality that are also
-smaller in size than those found by average GP for a given number of
-10.000 fitness evaluations, I also studied the differences in average
-solution size between pre-trained and regular DAE-GP.
+The experiments was repeated using the same setup with the only
+difference being a reduction from two hidden layers to 1 hidden layer.
 
@@
 \centering
-\includegraphics[width=2.08333in,height=8.33333in]{./img/airfoil_FullRun_30gens/mean_Size_byGens.png}
-\caption{Average Solution Size over 30 Generations - Airfoil}
+\includegraphics{./img/airfoil_1hl_maxIndSize_fullRun_30gens/mean_median_fitness_byGens.png}
+\caption{Best Fitness over 30 Generations - 1 hidden Layer - Airfoil}
 \end{figure}
+
+\ldots{}
+
+To gather more confidence in the previous results, the pre-training
+strategy for DAE-GP was tested for various real world symbolic
+regression problems taken from
+{[}\protect\hyperlink{ref-machine_learning_repo}{2}{]}. Are brief
+summary of the benchmark problems is given in table X
+
+\begin{table}[!h]
+
+\caption{\label{tab:full_run_realWorldSymReg_problems}Overview - Real World Symbolic Regression Benchmark Problems}
+\centering
+\begin{tabular}[t]{l|r|r}
+\hline
+\textbf{Problem} & \textbf{Observations} & \textbf{Features}\\
+\hline
+Airfoil & 1503 & 5\\
+\hline
+Boston\_Housing & 506 & 13\\
+\hline
+Energy\_Cooling & 768 & 8\\
+\hline
+Concrete & 1030 & 8\\
+\hline
+\end{tabular}
+\end{table}
+
+Table X summarizes the median best fitness achieved after running each
+algorithm for 10 runs using a population size of 500 individuals for 30
+generations.\footnote{P-Values are marked with asteriks (1,2,3) for
+  (\(\alpha\))-levels of \(0.1\), \(0.05\) and \(0.01\)}.
+
+\begin{longtable}[]{@{}
+  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1910}}
+  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1685}}
+  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.0787}}
+  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1461}}
+  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1461}}
+  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1124}}
+  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1573}}@{}}
+\toprule()
+\begin{minipage}[b]{\linewidth}\centering
+Problem
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+Hidden-Layers
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+Set
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+DAE-GP
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+Pre-Trained
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+P-Value
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+Cliffs-Delta
+\end{minipage} \\
+\midrule()
+\endhead
+Airfoil & 1 & Train & \textbf{33.3674} & 35.3454 & 0.0171** & 0.64 \\
+& 1 & Test & \textbf{33.7271} & 35.612 & 0.1400 & 0.4 \\
+Airfoil & 2 & Train & 17.4762 & \textbf{14.7233} & 0.3073 & -0.28 \\
+& 2 & Test & 17.5299 & \textbf{14.7603} & 0.5706 & -0.16 \\
+Boston\_Housing & 2 & Train & 8.1922 & \textbf{8.0763} & 0.2894 &
+-0.29 \\
+& 2 & Test & 8.0501 & \textbf{7.9692} & 0.7051 & -0.11 \\
+Energy(Cooling) & 2 & Train & \textbf{4.5271} & 4.7782 & 0.0188** &
+0.63 \\
+& 2 & Test & \textbf{4.67} & 4.9196 & 0.2886 & 0.29 \\
+Concrete & 2 & Train & \textbf{17.1075} & 17.528 & 1.0000 & 0.0 \\
+& 2 & Test & \textbf{15.7978} & 16.7196 & 0.6667 & 1.0 \\
+\bottomrule()
+\end{longtable}
+
+\hypertarget{solution-size-in-symbolic-regression}{%
+\subsection{Solution Size in Symbolic
+Regression}\label{solution-size-in-symbolic-regression}}
+
+\begin{longtable}[]{@{}
+  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2152}}
+  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1519}}
+  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1646}}
+  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1646}}
+  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1266}}
+  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1772}}@{}}
+\toprule()
+\begin{minipage}[b]{\linewidth}\centering
+Problem
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+Hid.Layers
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+DAE-GP
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+Pre-Trained
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+P-Value
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+Cliffs-Delta
+\end{minipage} \\
+\midrule()
+\endhead
+Airfoil & 1 & \textbf{33.3674} & 35.3454 & 0.0171** & 0.64 \\
+Airfoil & 2 & 17.4762 & \textbf{14.7233} & 0.3073 & -0.28 \\
+Boston\_Housing & 2 & 8.1922 & \textbf{8.0763} & 0.2894 & -0.29 \\
+Energy(Cooling) & 2 & \textbf{4.5271} & 4.7782 & 0.0188** & 0.63 \\
+Concrete & 2 & \textbf{17.1075} & 17.528 & 1.0000 & 0.0 \\
+\bottomrule()
+\end{longtable}
+
+\begin{tabular}{cc}
+\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/energyCooling_2hl_maxIndSize_fullRun_30gens/mean_Size_byGens.png} &
+\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/bostonHousing_2hl_maxIndSize_fullRun_30gens/mean_Size_byGens.png} \\
+\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens/mean_Size_byGens.png} &
+\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/concrete_2hl_maxIndSize_fullRun_30gens/mean_Size_byGens.png}
+\end{tabular}
 
@@
 
-{[}3{]}: Pre-Training effect especially usefull for lower-level layers
--\textgreater{} Only adapt weights from low layers or embedd the
-pre-trained model?
+{[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{3}{]}: Pre-Training effect
+especially usefull for lower-level layers -\textgreater{} Only adapt
+weights from low layers or embedd the pre-trained model?
 
@@
 Evolutionary algorithms made easy. \emph{Journal of Machine Learning
-Research} 13, (2012), 2171--2175.}
+Research} 13, (July 2012), 2171--2175.}
 

### 2023-01-11 paper.Rmd
autogit

@@
 
-![Population Diversity over 30 Generations - Airfoil](./img/airfoil_FullRun_30gens/mean_popDiversity_byGens.png)
+![Population Diversity over 30 Generations - Airfoil](./img/airfoil_2hl_maxIndSize_fullRun_30gens/mean_popDiversity_byGens.png)
 Since @dae-gp_2022_symreg shows, that DAE-GP finds higher quality that are also smaller in size than those found by average GP for a given number of 10.000 fitness evaluations, I also studied the differences in average solution size between pre-trained and regular DAE-GP.
 
-![Average Solution Size over 30 Generations - Airfoil](./img/airfoil_FullRun_30gens/mean_Size_byGens.png)
+![Average Solution Size over 30 Generations - Airfoil](./img/airfoil_2hl_maxIndSize_fullRun_30gens/mean_Size_byGens.png)
 
@@
  col.names = col_names,
- align = "c"
+ align = "l"
  #caption = "Symbolic Regression Overview - Mean final Fitness"

### 2023-01-11 slides.Rmd
Wed Jan 11 17:00:11 CET 2023: update-remote.sh
autogit

@@
 
-![Airfoil - Erste Generation Median Trainingsfehler - Variable Anzahl Hidden Layer](./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_layers.png){height=85%}
+![Airfoil - Erste Generation Median Trainingsfehler - Variable Anzahl Hidden Layer](./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_layers_6plots.png){height=85%}
 
@@
 
-![Airfoil - Erste Generation Median Trainingsfehler - Variable Anzahl Hidden Neurons](./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_neurons.png){height=85%}
+![Airfoil - Erste Generation Median Trainingsfehler - Variable Anzahl Hidden Neurons](./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_neurons_6plots.png){height=85%}
 
@@
 
-![Energy (Cooling) Datensatz - Fitness über Generationen](~/masterThesis/master_thesis/img/energyCooling_2hl_FullRun_30gens/mean_median_fitness_byGens.png)
+![Energy (Cooling) Datensatz - Fitness über Generationen](~/masterThesis/master_thesis/img/energyCooling_2hl_maxIndSize_fullRun_30gens/mean_median_fitness_byGens.png)
 
@@
 \includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens/mean_Size_byGens.png} &
-\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens/mean_Size_byGens.png}
+\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/concrete_2hl_maxIndSize_fullRun_30gens/mean_Size_byGens.png}
 \end{tabular}
@@
 
+...
+
 # References {.allowframebreaks}

### 2023-01-12 slides.Rmd
Thu Jan 12 09:00:08 CET 2023: update-remote.sh

@@
 * Ausschließliche Betrachtung des Rekonstruktionsfehlers der ersten Generation (Nutzung von seperaten Test- und Trainingspopulationen)
-* DAE-LSTM Training erfolgt über eine fixe Anzahl von $5000$ Epochen (kein frühzeitiger Abbruch bei Konvergenz des Testfehlers)
+* DAE-LSTM Training erfolgt über eine fixe Anzahl von $1000$ Epochen (kein frühzeitiger Abbruch bei Konvergenz des Testfehlers)
 
@@
 
-## Einfluss der Anzahl von Hidden Layers
+## Einfluss der Anzahl von Hidden Layers (1/2)
 
 ![Airfoil - Erste Generation Median Trainingsfehler - Variable Anzahl Hidden Layer](./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_layers_6plots.png){height=85%}
+## Einfluss der Anzahl von Hidden Layers (2/2)
 
-## Einfluss der Anzahl von Hidden Neurons
+![Airfoil - Erste Generation Median Trainingsfehler - Variable Anzahl Hidden Layer](./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_layers_3plots.png){height=85%}
+## Einfluss der Anzahl von Hidden Neurons (1/2)
+
+![Airfoil - Erste Generation Median Trainingsfehler - Variable Anzahl Hidden Neurons](./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_neurons_6plots.png){height=85%}
+## Einfluss der Anzahl von Hidden Neurons (2/2)
 
@@
 
-* Lösungsgüte
-* Größe der gefundenen Lösungen (=Anzahl an Knoten)
+* Lösungsqualität (Fitness)
+* Größe der gefundenen Lösungen (Anzahl an Knoten)
 * Populationsdiversität
@@
 ```
-## Airfoil Übersicht
+# Ergebnisse Airfoil Datensatz
+
+## Airfoil - Entwicklung der Fitness über Generationen
 
@@
 
-## Verteilung Lösungsgüte Airfoil
+## Airfoil - Verteilung der finalen Fitness
 
-![Airfoil Datensatz - Fitness über Generationen](~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens_withGP/final_fit_boxplot.png){height=90%}
+![Airfoil Datensatz - Verteilung der finale Fitness](~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens_withGP/final_fit_boxplot.png){height=90%}
 
-## Lösungsgröße Airfoil
+## Airfoil - Größe der besten Lösungskandidaten über Generationen
 
@@
 
-## Lösungsgüte Airfoil mit einem hidden Layer
+## Airfoil - Entwicklung der Fitness über Generationen
 
-![Airfoil Datensatz - 1 HL - Lösungsgüte über Generationen](~/masterThesis/master_thesis/img/airfoil_1hl_maxIndSize_fullRun_30gens/mean_median_fitness_byGens.png)
+![Airfoil Datensatz - 1 Hidden Layer - Lösungsgüte über Generationen](~/masterThesis/master_thesis/img/airfoil_1hl_maxIndSize_fullRun_30gens/mean_median_fitness_byGens.png)
 

### 2023-01-12 slides.Rmd
Thu Jan 12 10:00:10 CET 2023: update-remote.sh

@@
 
-## Kontrollexperiment: Reduzierung der DAE-LSTM Dimension Airfoil
+## Kontrollexperiment: Reduzierung der DAE-LSTM Dimension für Airfoil Datensatz
 
-Frage: Welchen Einfluss hat die Reduktion der DAE-LSTM auf 1 hidden Layer?
+Frage: Welchen Einfluss hat die Reduktion der DAE-LSTM auf 1 hidden Layer (HL)?
 
-## Airfoil - Entwicklung der Fitness über Generationen
+## Airfoil - Entwicklung der Fitness über Generationen (1 HL)
 
-![Airfoil Datensatz - 1 Hidden Layer - Lösungsgüte über Generationen](~/masterThesis/master_thesis/img/airfoil_1hl_maxIndSize_fullRun_30gens/mean_median_fitness_byGens.png)
+![Airfoil - Entwicklung der Fitness über Generationen (1 HL)](~/masterThesis/master_thesis/img/airfoil_1hl_maxIndSize_fullRun_30gens/mean_median_fitness_byGens.png)
 
-## Verteilung Lösungsgüte Airfoil mit einem hidden Layer
+## Airfoil - Verteilung der finalen Fitness (1HL)
 
-![Airfoil Datensatz - 1 HL - Lösungsgüte über Generationen](~/masterThesis/master_thesis/img/airfoil_1hl_maxIndSize_fullRun_30gens/final_fit_boxplot.png){height=90%}
+![Airfoil - Verteilung der finalen Fitness (1HL)](~/masterThesis/master_thesis/img/airfoil_1hl_maxIndSize_fullRun_30gens/final_fit_boxplot.png){height=90%}
 
-## Anwendung auf weiteren Datensätzen
+# Anwendung auf weitere Datensätzen
 
@@
 
-## Auswertung Fitness - Symbolische Regression
+## Symbolische Regression - Ergebnisübersicht
 
@@
   if(df[i, "DAE.GP"] < df[i, "Pre.Trained_DAE.GP"]){
-    df_bold[i, "DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(round(df[i, "DAE.GP"],4)))
-    df_bold[i, "Pre.Trained_DAE.GP"] <- as.character(round(df[i, "Pre.Trained_DAE.GP"],4))
+    df_bold[i, "DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(round(df[i, "DAE.GP"],2)))
+    df_bold[i, "Pre.Trained_DAE.GP"] <- as.character(round(df[i, "Pre.Trained_DAE.GP"],2))
     }
   else if (df[i, "DAE.GP"] > df[i, "Pre.Trained_DAE.GP"]) {
-    df_bold[i, "Pre.Trained_DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(round(df[i, "Pre.Trained_DAE.GP"],4)))
-  df_bold[i, "DAE.GP"] <- as.character(round(df[i, "DAE.GP"],4))
+    df_bold[i, "Pre.Trained_DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(round(df[i, "Pre.Trained_DAE.GP"],2)))
+  df_bold[i, "DAE.GP"] <- as.character(round(df[i, "DAE.GP"],2))
   }
@@
  col.names = col_names,
- align = "c"
+ align = "l"
  #caption = "Symbolic Regression Overview - Mean final Fitness"
-) %>% row_spec(0,bold=FALSE, font_size=7 ) %>% kableExtra::kable_styling(latex_options = "hold_position", font_size=7)
+) %>% 
+  kableExtra::kable_styling(latex_options = "scale_down", font_size=7)
 
@@
 
-## Fitness - Energy Cooling Datensatz
+## Symbolische Regression - Übersicht Fitness
 
-![Energy (Cooling) Datensatz - Fitness über Generationen](~/masterThesis/master_thesis/img/energyCooling_2hl_maxIndSize_fullRun_30gens/mean_median_fitness_byGens.png)
-
-## Fitness - Boston Housing Datensatz
-
-![Boston Housing Datensatz - Fitness über Generationen](~/masterThesis/master_thesis/img/bostonHousing_2hl_maxIndSize_fullRun_30gens/mean_median_fitness_byGens.png)
+\begin{tabular}{cc}
+\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/energyCooling_2hl_maxIndSize_fullRun_30gens/median_fitness_byGens.png} &
+\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/bostonHousing_2hl_maxIndSize_fullRun_30gens/median_fitness_byGens.png} \\
+\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens/median_fitness_byGens.png} &
+\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/concrete_2hl_maxIndSize_fullRun_30gens/median_fitness_byGens.png}
+\end{tabular}
 
@@
  col.names = col_names,
- align = "c"
+ align = "l"
  #caption = "Symbolic Regression Overview - Mean final Fitness"
-) %>% row_spec(0,bold=FALSE) %>% kableExtra::kable_styling(latex_options = "hold_position", font_size = 7)
+) %>% kableExtra::kable_styling(latex_options = "scale_down", font_size = 7)
 

### 2023-01-12 data/summarize_csv_files.py

@@ -28,13 +28,13 @@ def getPVal(sampleA, sampleB):
     #assert len(sampleA) > 6 and len(sampleB) >= 8
 
     if pval <= 0.01:
-        return f"{pval:.4f}***"
+        return f"{pval:.2f}***"
     elif pval <= 0.05:
-        return f"{pval:.4f}**"
+        return f"{pval:.2f}**"
     elif pval <= 0.1:
-        return f"{pval:.4f}*"
+        return f"{pval:.2f}*"
     else:
-        return f"{pval:.4f}"
+        return f"{pval:.2f}"
 
 
 # Fitness

### 2023-01-12 slides.Rmd
Thu Jan 12 11:00:09 CET 2023: update-remote.sh

@@
 
-# Experiment Generalisierungsverhalten
+# Untersuchung des Generalisierungsverhalten
 
@@
 
-![Airfoil - Erste Generation Median Trainingsfehler - Variable Anzahl Hidden Layer](./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_layers_6plots.png){height=85%}
+![Airfoil - Erste Generation Median Trainingsfehler - Variable Anzahl Hidden Layer (1/2)](./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_layers_6plots.png){height=85%}
+
 ## Einfluss der Anzahl von Hidden Layers (2/2)
 
-![Airfoil - Erste Generation Median Trainingsfehler - Variable Anzahl Hidden Layer](./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_layers_3plots.png){height=85%}
+![Airfoil - Erste Generation Median Trainingsfehler - Variable Anzahl Hidden Layer (2/2)](./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_layers_3plots.png){height=85%}
+
 ## Einfluss der Anzahl von Hidden Neurons (1/2)
 
-![Airfoil - Erste Generation Median Trainingsfehler - Variable Anzahl Hidden Neurons](./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_neurons_6plots.png){height=85%}
+![Airfoil - Erste Generation Median Trainingsfehler - Variable Anzahl Hidden Neurons (1/2)](./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_neurons_6plots.png){height=85%}
+
 ## Einfluss der Anzahl von Hidden Neurons (2/2)
 
-![Airfoil - Erste Generation Median Trainingsfehler - Variable Anzahl Hidden Neurons](./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_neurons_6plots.png){height=85%}
+![Airfoil - Erste Generation Median Trainingsfehler - Variable Anzahl Hidden Neurons (2/2)](./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_neurons_6plots.png){height=85%}
 
-## Interpretation
+## Fazit Generalisierungsverhalten
 
@@
 
-# Experimente Suchverhalten Symbolische Regression
+# Untersuchung des Suchverhaltens bei symbolischer Regression
 
@@
 
-Aufbau: Betrachtung des Suchverhalten über je 10 Gesamtduchläufe für DAE-GP, pre-trained DAE-GP und reguläres GP ($30$ Durchläufe gesamt)
+Aufbau: Betrachtung des Suchverhalten über je $10$ Gesamtduchläufe für DAE-GP und pre-trained DAE-GP
 
@@
 ```
-# Ergebnisse Airfoil Datensatz
+## Ergebnisse Airfoil Datensatz
 
@@
 
-![Airfoil Datensatz - Verteilung der finale Fitness](~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens_withGP/final_fit_boxplot.png){height=90%}
+![Airfoil Datensatz - Verteilung der finale Fitness](~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens/final_fit_boxplot.png){height=90%}
 
@@
  col.names = col_names,
- align = "l"
+ align = "c"
  #caption = "Symbolic Regression Overview - Mean final Fitness"
@@
 
-## Symbolische Regression - Übersicht Fitness
+## Symbolische Regression - Übersicht Median Fitness
 
@@
 
-## Auswertung Mean der Größe der besten Individuen - Symbolische Regression
+## Auswertung Größe der besten Individuen - Symbolische Regression
 

### 2023-01-12 paper.Rmd
autogit

@@
 print(f"Python Version: {version} \nPath: {executable}")
+```
+
+```{bash recompile_results, include=FALSE}
+
+# compile mermaid diagrams
+# ---
+WIDTH=3200
+HEIGHT=2400
+THEME="neutral"
+BACKGROUND="white"
+
+for file in ./mermaid/*.mmd
+do
+  mmdc -i "$file" -o ./img/flowcharts/$(basename "$file" .mmd).png -t $THEME -w $WIDTH -H $HEIGHT -b $BACKGROUND 
+done
+
+# Airfoil First Gen
+# ---
+bash /Users/rmn/masterThesis/eda-gp-2020/experiments/airfoil_firstGen/re_evaluate_airfoilFirstGen.sh
+
+# Real World SymReg Full Runs
+# ---
+
+# airfoil
+bash /Users/rmn/masterThesis/eda-gp-2020/experiments/airfoil_1hl_maxIndSize_fullRun_30gens/re_evaluate_airfoil_1hl_maxIndSize_30gen.sh
+
+bash /Users/rmn/masterThesis/eda-gp-2020/experiments/airfoil_2hl_maxIndSize_fullRun_30gens/re_evaluate_airfoil_2hl_maxIndSize_30gen.sh
+
+# energy (cooling)
+bash /Users/rmn/masterThesis/eda-gp-2020/experiments/energyCooling_2hl_maxIndSize_fullRun_30gens/re_evaluate_energyCooling_2hl_maxIndSize_fullRun_30gens.sh
+
+# boston housing
+bash /Users/rmn/masterThesis/eda-gp-2020/experiments/bostonHousing_2hl_maxIndSize_fullRun_30gens/re_evaluate_bostonHousing_2hl_maxIndSize_fullRun_30gens.sh
+
+# concrete
+bash /Users/rmn/masterThesis/eda-gp-2020/experiments/concrete_2hl_maxIndSize_fullRun_30gens/re_evaluate_concrete_2hl_maxIndSize_fullRun_30gens.sh
+
+# summarize csv files for full run experiments
+/Users/rmn/miniconda3/envs/dataScience/bin/python /Users/rmn/github/master_thesis/data/summarize_csv_files.py
+
 ```
@@
 
-```{bash compile_mermaid, include=FALSE}
-# compile all *.mmd files in the mermaid directory
-
-WIDTH=3200
-HEIGHT=2400
-THEME="neutral"
-BACKGROUND="white"
-
-for file in ./mermaid/*.mmd
-do
-  mmdc -i "$file" -o ./img/flowcharts/$(basename "$file" .mmd).png -t $THEME -w $WIDTH -H $HEIGHT -b $BACKGROUND 
-done
-
-```
-
 <!--- Explain algorithm and implementation of the pre-training strategy  -->
@@
 
-```{bash compile_firstGenResults, include=FALSE}
-bash /Users/rmn/masterThesis/eda-gp-2020/experiments/airfoil_firstGen/re_evaluate_airfoilFirstGen.sh
-```
-
 To gather a deeper understanding about the effect of pre-training DAE-LSTM in DAE-GP a series of experiments was conducted using the airfoil dataset for symbolic regression while using and different parameter configurations for the number of hidden layers as well as the number of hidden neurons per hidden layer.
@@
 
-```{bash executeFirstGenNBs, include=FALSE}
-bash /Users/rmn/masterThesis/eda-gp-2020/experiments/airfoil_firstGen/re_evaluate_airfoilFirstGen.sh
-```
-
 <!--- analyze results 1 hidden layer and variable number of hidden neurons -->
@@
 ## Effect on search performance
-
-```{bash compile_symReg_fullRun_30gens, include=FALSE}
-
-# airfoil
-bash /Users/rmn/masterThesis/eda-gp-2020/experiments/airfoil_1hl_maxIndSize_fullRun_30gens/re_evaluate_airfoil_1hl_maxIndSize_30gen.sh
-
-bash /Users/rmn/masterThesis/eda-gp-2020/experiments/airfoil_2hl_maxIndSize_fullRun_30gens/re_evaluate_airfoil_2hl_maxIndSize_30gen.sh
-
-# energy (cooling)
-bash /Users/rmn/masterThesis/eda-gp-2020/experiments/energyCooling_2hl_maxIndSize_fullRun_30gens/re_evaluate_energyCooling_2hl_maxIndSize_fullRun_30gens.sh
-
-# boston housing
-bash /Users/rmn/masterThesis/eda-gp-2020/experiments/bostonHousing_2hl_maxIndSize_fullRun_30gens/re_evaluate_bostonHousing_2hl_maxIndSize_fullRun_30gens.sh
-
-# concrete
-bash /Users/rmn/masterThesis/eda-gp-2020/experiments/concrete_2hl_maxIndSize_fullRun_30gens/re_evaluate_concrete_2hl_maxIndSize_fullRun_30gens.sh
-
-# ...
-```
 
@@
 Table X summarizes the median best fitness achieved after running each algorithm for 10 runs using a population size of 500 individuals for 30 generations.^[P-Values are marked with asteriks (1,2,3) for ($\alpha$)-levels of $0.1$, $0.05$ and $0.01$].
-
-``` {bash recreate_csv_file_benchmarkRWSymReg, include=FALSE}
-/Users/rmn/miniconda3/envs/dataScience/bin/python /Users/rmn/github/master_thesis/data/summarize_csv_files.py
-```
 

### 2023-01-12 slides.tex
autogit

@@
 
-\begin{frame}{Generalisiertes Royal Tree Problem (GRT)\footnote<.->{Wittenberg,
-  Rothlauf and Schweim (\protect\hyperlink{ref-dae-gp_2020_rtree}{2020})}:}
+\begin{frame}{Generalisiertes Royal Tree Problem\footnote<.->{Wittenberg,
+  Rothlauf and Schweim (\protect\hyperlink{ref-dae-gp_2020_rtree}{2020})}
+(GRT):}
 \protect\hypertarget{generalisiertes-royal-tree-problem-grt}{}
@@
 
-\hypertarget{experiment-generalisierungsverhalten}{%
-\section{Experiment
-Generalisierungsverhalten}\label{experiment-generalisierungsverhalten}}
+\hypertarget{untersuchung-des-generalisierungsverhalten}{%
+\section{Untersuchung des
+Generalisierungsverhalten}\label{untersuchung-des-generalisierungsverhalten}}
 
@@
 \item
-  DAE-LSTM Training erfolgt über eine fixe Anzahl von \(5000\) Epochen
+  DAE-LSTM Training erfolgt über eine fixe Anzahl von \(1000\) Epochen
   (kein frühzeitiger Abbruch bei Konvergenz des Testfehlers)
@@
 
-\begin{frame}{Einfluss der Anzahl von Hidden Layers}
-\protect\hypertarget{einfluss-der-anzahl-von-hidden-layers}{}
-\includegraphics[width=\textwidth,height=0.85\textheight]{./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_layers.png}
+\begin{frame}{Einfluss der Anzahl von Hidden Layers (1/2)}
+\protect\hypertarget{einfluss-der-anzahl-von-hidden-layers-12}{}
+\includegraphics[width=\textwidth,height=0.85\textheight]{./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_layers_6plots.png}
 \end{frame}
 
-\begin{frame}{Einfluss der Anzahl von Hidden Neurons}
-\protect\hypertarget{einfluss-der-anzahl-von-hidden-neurons}{}
-\includegraphics[width=\textwidth,height=0.85\textheight]{./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_neurons.png}
+\begin{frame}{Einfluss der Anzahl von Hidden Layers (2/2)}
+\protect\hypertarget{einfluss-der-anzahl-von-hidden-layers-22}{}
+\includegraphics[width=\textwidth,height=0.85\textheight]{./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_layers_3plots.png}
 \end{frame}
 
-\begin{frame}{Interpretation}
-\protect\hypertarget{interpretation}{}
+\begin{frame}{Einfluss der Anzahl von Hidden Neurons (1/2)}
+\protect\hypertarget{einfluss-der-anzahl-von-hidden-neurons-12}{}
+\includegraphics[width=\textwidth,height=0.85\textheight]{./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_neurons_6plots.png}
+\end{frame}
+
+\begin{frame}{Einfluss der Anzahl von Hidden Neurons (2/2)}
+\protect\hypertarget{einfluss-der-anzahl-von-hidden-neurons-22}{}
+\includegraphics[width=\textwidth,height=0.85\textheight]{./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_neurons_6plots.png}
+\end{frame}
+
+\begin{frame}{Fazit Generalisierungsverhalten}
+\protect\hypertarget{fazit-generalisierungsverhalten}{}
 \ldots{}
@@
 
-\hypertarget{experimente-suchverhalten-symbolische-regression}{%
-\section{Experimente Suchverhalten Symbolische
-Regression}\label{experimente-suchverhalten-symbolische-regression}}
+\hypertarget{untersuchung-des-suchverhaltens-bei-symbolischer-regression}{%
+\section{Untersuchung des Suchverhaltens bei symbolischer
+Regression}\label{untersuchung-des-suchverhaltens-bei-symbolischer-regression}}
 
@@
 
-Aufbau: Betrachtung des Suchverhalten über je 10 Gesamtduchläufe für
-DAE-GP, pre-trained DAE-GP und reguläres GP (\(30\) Durchläufe gesamt)
+Aufbau: Betrachtung des Suchverhalten über je \(10\) Gesamtduchläufe für
+DAE-GP und pre-trained DAE-GP
 
@@
 \item
-  Lösungsgüte
+  Lösungsqualität (Fitness)
 \item
-  Größe der gefundenen Lösungen (=Anzahl an Knoten)
+  Größe der gefundenen Lösungen (Anzahl an Knoten)
 \item
@@
 \protect\hypertarget{hyperparameter}{}
-\begin{table}[!h]
-\centering
+\begin{table}
+\centering\begingroup\fontsize{7}{9}\selectfont
+
+\resizebox{\linewidth}{!}{
 \begin{tabular}{l|l}
@@
 \hline
-SamplingSteps & 1\\
+SamplingSteps & 2\\
 \hline
@@
 \hline
-\end{tabular}
+\end{tabular}}
+\endgroup{}
 \end{table}
@@
 
-\begin{frame}{Airfoil Vergleich GP}
-\protect\hypertarget{airfoil-vergleich-gp}{}
+\begin{frame}{Airfoil - Entwicklung der Fitness über Generationen}
+\protect\hypertarget{airfoil---entwicklung-der-fitness-uxfcber-generationen}{}
 \includegraphics[width=\textwidth,height=0.9\textheight]{~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens_withGP/mean_median_fitness_byGens.png}
@@
 
-\begin{frame}{Ergebnisse - Lösungsgüte Airfoil}
-\protect\hypertarget{ergebnisse---luxf6sungsguxfcte-airfoil}{}
-\begin{tabular}{c}
-\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens/mean_meadian_fitness_byGens.png} &
-\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens/final_fit_boxplot.png} \\
-
-\end{tabular}
-\end{frame}
-
-\begin{frame}{Verteilung Lösungsgüte Airfoil}
-\protect\hypertarget{verteilung-luxf6sungsguxfcte-airfoil}{}
+\begin{frame}{Airfoil - Verteilung der finalen Fitness}
+\protect\hypertarget{airfoil---verteilung-der-finalen-fitness}{}
 \includegraphics[width=\textwidth,height=0.9\textheight]{~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens/final_fit_boxplot.png}
@@
 
-\begin{frame}{Lösungsgröße Airfoil}
-\protect\hypertarget{luxf6sungsgruxf6uxdfe-airfoil}{}
+\begin{frame}{Airfoil - Größe der besten Lösungskandidaten}
+\protect\hypertarget{airfoil---gruxf6uxdfe-der-besten-luxf6sungskandidaten}{}
 \includegraphics{~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens_withGP/mean_Size_byGens.png}
@@
 
-\begin{frame}{Kontrollexperiment: Reduzierung der DAE-LSTM Dimension
-Airfoil}
-\protect\hypertarget{kontrollexperiment-reduzierung-der-dae-lstm-dimension-airfoil}{}
+\begin{frame}{Kontrollexperiment: Reduzierung der DAE-LSTM Dimension für
+Airfoil Datensatz}
+\protect\hypertarget{kontrollexperiment-reduzierung-der-dae-lstm-dimension-fuxfcr-airfoil-datensatz}{}
 Frage: Welchen Einfluss hat die Reduktion der DAE-LSTM auf 1 hidden
-Layer?
+Layer (HL)?
 \end{frame}
 
-\begin{frame}{Lösungsgüte Airfoil mit einem hidden Layer}
-\protect\hypertarget{luxf6sungsguxfcte-airfoil-mit-einem-hidden-layer}{}
+\begin{frame}{Airfoil - Entwicklung der Fitness (1 HL)}
+\protect\hypertarget{airfoil---entwicklung-der-fitness-1-hl}{}
 \includegraphics{~/masterThesis/master_thesis/img/airfoil_1hl_maxIndSize_fullRun_30gens/mean_median_fitness_byGens.png}
@@
 
-\begin{frame}{Verteilung Lösungsgüte Airfoil mit einem hidden Layer}
-\protect\hypertarget{verteilung-luxf6sungsguxfcte-airfoil-mit-einem-hidden-layer}{}
-\includegraphics{~/masterThesis/master_thesis/img/airfoil_1hl_maxIndSize_fullRun_30gens/final_fit_boxplot.png}
+\begin{frame}{Airfoil - Verteilung der finalen Fitness (1HL)}
+\protect\hypertarget{airfoil---verteilung-der-finalen-fitness-1hl}{}
+\includegraphics[width=\textwidth,height=0.9\textheight]{~/masterThesis/master_thesis/img/airfoil_1hl_maxIndSize_fullRun_30gens/final_fit_boxplot.png}
 \end{frame}
 
-\begin{frame}{Anwendung auf weiteren Datensätzen}
-\protect\hypertarget{anwendung-auf-weiteren-datensuxe4tzen}{}
+\begin{frame}{Airfoil - Größe der besten Lösungskandidaten(1HL)}
+\protect\hypertarget{airfoil---gruxf6uxdfe-der-besten-luxf6sungskandidaten1hl}{}
+\includegraphics[width=\textwidth,height=0.9\textheight]{~/masterThesis/master_thesis/img/airfoil_1hl_maxIndSize_fullRun_30gens/mean_Size_byGens.png}
+\end{frame}
+
+\begin{frame}{Anwendung auf weitere Datensätzen}
+\protect\hypertarget{anwendung-auf-weitere-datensuxe4tzen}{}
 Airfoil Datensatz: Ergebnis deuten bei einer ausreichenden Anzahl von
@@
 
-\begin{frame}{Auswertung Fitness - Symbolische Regression}
-\protect\hypertarget{auswertung-fitness---symbolische-regression}{}
+\begin{frame}{Symbolische Regression - Ergebnisübersicht}
+\protect\hypertarget{symbolische-regression---ergebnisuxfcbersicht}{}
 \begin{longtable}[]{@{}
-  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1910}}
-  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1685}}
-  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.0787}}
-  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1461}}
-  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1461}}
-  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1124}}
-  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1573}}@{}}
+  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1977}}
+  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1744}}
+  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.0814}}
+  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1279}}
+  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1512}}
+  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1047}}
+  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1628}}@{}}
 \toprule()
@@
 \endhead
-Airfoil & 1 & Train & \textbf{33.3674} & 35.3454 & 0.0171** & 0.64 \\
-& 1 & Test & \textbf{33.7271} & 35.612 & 0.1400 & 0.4 \\
-Airfoil & 2 & Train & 17.4762 & \textbf{14.7233} & 0.3073 & -0.28 \\
-& 2 & Test & 17.5299 & \textbf{14.7603} & 0.5706 & -0.16 \\
-Boston\_Housing & 2 & Train & 8.1922 & \textbf{8.0763} & 0.2894 &
--0.29 \\
-& 2 & Test & 8.0501 & \textbf{7.9692} & 0.7051 & -0.11 \\
-Energy(Cooling) & 2 & Train & \textbf{4.5271} & 4.7782 & 0.0188** &
-0.63 \\
-& 2 & Test & \textbf{4.67} & 4.9196 & 0.2886 & 0.29 \\
+Airfoil & 1 & Train & \textbf{33.37} & 35.35 & 0.02** & 0.64 \\
+& 1 & Test & \textbf{33.73} & 35.61 & 0.14 & 0.4 \\
+Airfoil & 2 & Train & 17.48 & \textbf{14.72} & 0.31 & -0.28 \\
+& 2 & Test & 17.53 & \textbf{14.76} & 0.57 & -0.16 \\
+Boston\_Housing & 2 & Train & 8.19 & \textbf{8.08} & 0.29 & -0.29 \\
+& 2 & Test & 8.05 & \textbf{7.97} & 0.71 & -0.11 \\
+Energy(Cooling) & 2 & Train & \textbf{4.53} & 4.78 & 0.02** & 0.63 \\
+& 2 & Test & \textbf{4.67} & 4.92 & 0.29 & 0.29 \\
+Concrete & 2 & Train & 17.3 & \textbf{16.31} & 0.53 & -0.5 \\
+& 2 & Test & 16.67 & \textbf{16.19} & 0.80 & -0.25 \\
 \bottomrule()
@@
 
-\begin{frame}{Fitness - Energy Cooling Datensatz}
-\protect\hypertarget{fitness---energy-cooling-datensatz}{}
-\includegraphics{~/masterThesis/master_thesis/img/energyCooling_2hl_FullRun_30gens/mean_median_fitness_byGens.png}
-\end{frame}
-
-\begin{frame}{Fitness - Boston Housing Datensatz}
-\protect\hypertarget{fitness---boston-housing-datensatz}{}
-\includegraphics{~/masterThesis/master_thesis/img/bostonHousing_2hl_maxIndSize_fullRun_30gens/mean_median_fitness_byGens.png}
+\begin{frame}{Symbolische Regression - Übersicht Median Fitness}
+\protect\hypertarget{symbolische-regression---uxfcbersicht-median-fitness}{}
+\begin{tabular}{cc}
+\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/energyCooling_2hl_maxIndSize_fullRun_30gens/median_fitness_byGens.png} &
+\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/bostonHousing_2hl_maxIndSize_fullRun_30gens/median_fitness_byGens.png} \\
+\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens/median_fitness_byGens.png} &
+\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/concrete_2hl_maxIndSize_fullRun_30gens/median_fitness_byGens.png}
+\end{tabular}
 \end{frame}
@@
 
-\begin{frame}{Auswertung Mean der Größe der besten Individuen -
-Symbolische Regression}
-\protect\hypertarget{auswertung-mean-der-gruxf6uxdfe-der-besten-individuen---symbolische-regression}{}
+\begin{frame}{Auswertung Größe der besten Individuen - Symbolische
+Regression}
+\protect\hypertarget{auswertung-gruxf6uxdfe-der-besten-individuen---symbolische-regression}{}
 \begin{longtable}[]{@{}
-  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2152}}
-  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1519}}
-  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1646}}
-  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1646}}
-  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1266}}
-  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1772}}@{}}
+  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2237}}
+  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1579}}
+  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1447}}
+  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1711}}
+  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1184}}
+  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1842}}@{}}
 \toprule()
@@
 \endhead
-Airfoil & 1 & \textbf{33.3674} & 35.3454 & 0.0171** & 0.64 \\
-Airfoil & 2 & 17.4762 & \textbf{14.7233} & 0.3073 & -0.28 \\
-Boston\_Housing & 2 & 8.1922 & \textbf{8.0763} & 0.2894 & -0.29 \\
-Energy(Cooling) & 2 & \textbf{4.5271} & 4.7782 & 0.0188** & 0.63 \\
+Airfoil & 1 & \textbf{33.37} & 35.35 & 0.02** & 0.64 \\
+Airfoil & 2 & 17.48 & \textbf{14.72} & 0.31 & -0.28 \\
+Boston\_Housing & 2 & 8.19 & \textbf{8.08} & 0.29 & -0.29 \\
+Energy(Cooling) & 2 & \textbf{4.53} & 4.78 & 0.02** & 0.63 \\
+Concrete & 2 & 17.3 & \textbf{16.31} & 0.53 & -0.5 \\
 \bottomrule()
@@
 \section{Weitere Ansätze}\label{weitere-ansuxe4tze}}
+
+\begin{frame}{Weitere Ansätze}
+\ldots{}
+\end{frame}
 

### 2023-01-12 data/summarize_csv_files.py
autogit

@@ -121,3 +121,30 @@ for problem,path in RESULTS.items():
 with open("/Users/rmn/github/master_thesis/data/summary_table_best_size_mean.csv", "w", encoding="utf-8") as f:
     f.write(csv_string)
 
+
+
+FILENAME_EPOCHS="epochs_trained_perGen.json"
+
+# Epochs Trained
+# ---
+
+# summarize by mean
+csv_string = "Problem,Hidden_Layer,DAE-GP,Pre-Trained_DAE-GP,P_Value,Cliffs_Delta\n"
+
+for problem,path in RESULTS.items():
+
+    d = json.load(open(join(path, FILENAME_EPOCHS),"r",encoding="utf-8"))
+
+    
+    reg_mean = np.mean(np.array(d["DAE-GP"]), axis=0)
+    pt_mean = np.mean(np.array(d["Pre-Trained"]), axis=0)
+
+    mean_reg_mean = np.mean(reg_mean)
+    mean_pt_mean = np.mean(pt_mean)
+
+
+    csv_string += f"{d['problem']},{d['hiddenLayer']},{mean_reg_mean},{mean_pt_mean},{getPVal(reg_mean, pt_mean)},{cliffsDeltaPretty(pt_mean, reg_mean)}\n"
+
+
+with open("/Users/rmn/github/master_thesis/data/summary_table_mean_epochsTrained.csv", "w", encoding="utf-8") as f:
+    f.write(csv_string)

### 2023-01-15 slides.Rmd
Sun Jan 15 13:00:06 CET 2023: update-remote.sh

@@
 
-```{python pytest, echo=FALSE, include=FALSE}
-from sys import version, executable
-print(f"Python Version: {version} \nPath: {executable}")
+```{bash recompile_results, eval=TRUE, include=FALSE}
+
+# set eval to TRUE to recompile all results, plots etc.
+# Warning: This will take some time...
+
+# compile mermaid diagrams
+# ---
+WIDTH=3200
+HEIGHT=2400
+THEME="neutral"
+BACKGROUND="white"
+
+for file in ./mermaid/*.mmd
+do
+  mmdc -i "$file" -o ./img/flowcharts/$(basename "$file" .mmd).png -t $THEME -w $WIDTH -H $HEIGHT -b $BACKGROUND 
+done
+
+# Airfoil First Gen
+# ---
+bash /Users/rmn/masterThesis/eda-gp-2020/experiments/airfoil_firstGen/re_evaluate_airfoilFirstGen.sh
+
+# Real World SymReg Full Runs
+# ---
+
+# airfoil
+bash /Users/rmn/masterThesis/eda-gp-2020/experiments/airfoil_1hl_maxIndSize_fullRun_30gens/re_evaluate_airfoil_1hl_maxIndSize_30gen.sh
+
+bash /Users/rmn/masterThesis/eda-gp-2020/experiments/airfoil_2hl_maxIndSize_fullRun_30gens/re_evaluate_airfoil_2hl_maxIndSize_30gen.sh
+
+# energy (cooling)
+bash /Users/rmn/masterThesis/eda-gp-2020/experiments/energyCooling_2hl_maxIndSize_fullRun_30gens/re_evaluate_energyCooling_2hl_maxIndSize_fullRun_30gens.sh
+
+# boston housing
+bash /Users/rmn/masterThesis/eda-gp-2020/experiments/bostonHousing_2hl_maxIndSize_fullRun_30gens/re_evaluate_bostonHousing_2hl_maxIndSize_fullRun_30gens.sh
+
+# concrete
+bash /Users/rmn/masterThesis/eda-gp-2020/experiments/concrete_2hl_maxIndSize_fullRun_30gens/re_evaluate_concrete_2hl_maxIndSize_fullRun_30gens.sh
+
+# summarize csv files for full run experiments
+/Users/rmn/miniconda3/envs/dataScience/bin/python /Users/rmn/github/master_thesis/data/summarize_csv_files.py
+
 ```
@@
 
-## Generalisiertes Royal Tree Problem (GRT)^[@dae-gp_2020_rtree]:
+## Generalisiertes Royal Tree Problem^[@dae-gp_2020_rtree] (GRT):
 
@@
   # caption="Hyperparameter"
-)%>% row_spec(0,bold=TRUE ) %>% kableExtra::kable_styling(latex_options = "hold_position")
+)%>% row_spec(0,bold=TRUE ) %>% kable_styling(latex_options = "scale_down", font_size = 7)
 ```
@@
 ```
-## Ergebnisse Airfoil Datensatz
 
@@
 
-## Airfoil - Größe der besten Lösungskandidaten über Generationen
+## Airfoil - Größe der besten Lösungskandidaten
 
@@
 
-## Airfoil - Entwicklung der Fitness über Generationen (1 HL)
+## Airfoil - Entwicklung der Fitness (1 HL)
 
@@
 
-# Anwendung auf weitere Datensätzen
+## Airfoil - Größe der besten Lösungskandidaten(1HL)
+
+![Airfoil - Größe der besten Lösung über Generationen (1HL)](~/masterThesis/master_thesis/img/airfoil_1hl_maxIndSize_fullRun_30gens/mean_Size_byGens.png){height=90%}
+
+## Anwendung auf weitere Datensätzen
 
@@
 ) %>% 
-  kableExtra::kable_styling(latex_options = "scale_down", font_size=7)
+  kable_styling(latex_options = "scale_down", font_size = 7)
 
@@
  col.names = col_names,
- align = "l"
+ align = "c"
  #caption = "Symbolic Regression Overview - Mean final Fitness"
-) %>% kableExtra::kable_styling(latex_options = "scale_down", font_size = 7)
+) %>% 
+  kable_styling(latex_options = "scale_down", font_size = 7)
 
@@
 
+## Populationsdiversität - Symbolische Regression
+
+\begin{tabular}{cc}
+\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/energyCooling_2hl_maxIndSize_fullRun_30gens/mean_popDiversity_byGens_levOnly.png} &
+\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/bostonHousing_2hl_maxIndSize_fullRun_30gens/mean_popDiversity_byGens_levOnly.png} \\
+\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens/mean_popDiversity_byGens_levOnly.png} &
+\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/concrete_2hl_maxIndSize_fullRun_30gens/mean_popDiversity_byGens_levOnly.png}
+\end{tabular}
+
+## Auswertung Populationsdiversität - Symbolische Regression
+
+```{r full_run_realWorldSymReg_best_solution_size, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
+library(knitr)
+library(kableExtra)
+library(dplyr)
+
+# Create the data frame from csv
+df <- read.csv(
+  "/Users/rmn/github/master_thesis/data/summary_table_mean_LevDistance.csv",
+  sep=",",
+  colClasses = c("character", "numeric", "numeric", "numeric", "character", "character")
+)
+
+# Create a new dataframe with the modified values of the two columns,
+df_bold <- df
+
+for(i in 1:nrow(df)){
+  if(df[i, "DAE.GP"] < df[i, "Pre.Trained_DAE.GP"]){
+    df_bold[i, "DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(round(df[i, "DAE.GP"],2)))
+    df_bold[i, "Pre.Trained_DAE.GP"] <- as.character(round(df[i, "Pre.Trained_DAE.GP"],2))
+    }
+  else if (df[i, "DAE.GP"] > df[i, "Pre.Trained_DAE.GP"]) {
+    df_bold[i, "Pre.Trained_DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(round(df[i, "Pre.Trained_DAE.GP"],2)))
+  df_bold[i, "DAE.GP"] <- as.character(round(df[i, "DAE.GP"],2))
+  }
+}
+# shorten column names
+col_names <- c("Problem", "Hid.Layers", "DAE-GP", "Pre-Trained", "P-Value", "Cliffs-Delta")
+
+knitr::kable(
+ df_bold,
+ format = "markdown",
+ col.names = col_names,
+ align = "c"
+ #caption = "Symbolic Regression Overview - Mean final Fitness"
+) %>% 
+  kable_styling(latex_options = "scale_down", font_size = 7)
+
+```
+
 # Weitere Ansätze

### 2023-01-15 data/summarize_csv_files.py

@@ -16,11 +16,11 @@ RESULTS = {
 }
 
 
-FILENAME="best_final_fitness.json"
+
 
 def cliffsDeltaPretty(*args):
     val,desc = cliffs_delta(*args)
-    return f"{val}"
+    return f"{val:.2f}"
 
 def getPVal(sampleA, sampleB):
     stats, pval = mwu(sampleA, sampleB)
@@ -39,6 +39,8 @@ def getPVal(sampleA, sampleB):
 
 # Fitness
 # ---
+
+FILENAME="best_final_fitness.json"
 
 # summarize by mean
 csv_string = "Problem,Hidden_Layer,Dataset,DAE-GP,Pre-Trained_DAE-GP,P_Value,Cliffs_Delta\n"
@@ -98,10 +100,12 @@ with open("/Users/rmn/github/master_thesis/data/summary_table_final_fit_median.c
 
 
 
-FILENAME_SIZE = "size_best_solution.json"
 
 # Solution Size
 # ---
+
+FILENAME_SIZE = "size_best_solution.json"
+
 
 # summarize by mean
 csv_string = "Problem,Hidden_Layer,DAE-GP,Pre-Trained_DAE-GP,P_Value,Cliffs_Delta\n"
@@ -123,10 +127,12 @@ with open("/Users/rmn/github/master_thesis/data/summary_table_best_size_mean.csv
 
 
 
-FILENAME_EPOCHS="epochs_trained_perGen.json"
-
 # Epochs Trained
 # ---
+
+
+FILENAME_EPOCHS="epochs_trained_perGen.json"
+
 
 # summarize by mean
 csv_string = "Problem,Hidden_Layer,DAE-GP,Pre-Trained_DAE-GP,P_Value,Cliffs_Delta\n"
@@ -147,4 +153,60 @@ for problem,path in RESULTS.items():
 
 
 with open("/Users/rmn/github/master_thesis/data/summary_table_mean_epochsTrained.csv", "w", encoding="utf-8") as f:
+    f.write(csv_string)
+
+
+
+
+
+# Population Diversity (norm. Levenshtein Edit Distance)
+# ---
+
+# by mean
+
+FILENAME_LEVDIV="lev_div_ovrGens.json"
+
+# summarize by mean
+csv_string = "Problem,Hidden_Layer,DAE-GP,Pre-Trained_DAE-GP,P_Value,Cliffs_Delta\n"
+
+for problem,path in RESULTS.items():
+
[…]

### 2023-01-15 slides.Rmd
Sun Jan 15 14:00:05 CET 2023: update-remote.sh

@@
 
-## Symbolische Regression - Ergebnisübersicht
+## Übersicht Median Fitness - Symbolische Regression
+
+\begin{tabular}{cc}
+\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/energyCooling_2hl_maxIndSize_fullRun_30gens/median_fitness_byGens.png} &
+\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/bostonHousing_2hl_maxIndSize_fullRun_30gens/median_fitness_byGens.png} \\
+\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens/median_fitness_byGens.png} &
+\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/concrete_2hl_maxIndSize_fullRun_30gens/median_fitness_byGens.png}
+\end{tabular}
+
+## Auswertung Fitness (Median) - Symbolische Regression
 
@@
 df <- read.csv(
-  "/Users/rmn/github/master_thesis/data/summary_table_final_fit_mean.csv",
-  # "/Users/rmn/github/master_thesis/data/summary_table_final_fit_median.csv",  -> Use this path to summarize by median
+  "/Users/rmn/github/master_thesis/data/summary_table_final_fit_median.csv",
   sep=",",
@@
 
-## Symbolische Regression - Übersicht Median Fitness
-
-\begin{tabular}{cc}
-\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/energyCooling_2hl_maxIndSize_fullRun_30gens/median_fitness_byGens.png} &
-\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/bostonHousing_2hl_maxIndSize_fullRun_30gens/median_fitness_byGens.png} \\
-\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens/median_fitness_byGens.png} &
-\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/concrete_2hl_maxIndSize_fullRun_30gens/median_fitness_byGens.png}
-\end{tabular}
-
 ## Zusammenfassung Fitness - Symbolische Regression
@@
 
-## Auswertung Größe der besten Individuen - Symbolische Regression
+## Übersicht Lösungsgröße (Median) - Symbolische Regression
+
+\begin{tabular}{cc}
+\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/energyCooling_2hl_maxIndSize_fullRun_30gens/median_Size_byGens.png} &
+\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/bostonHousing_2hl_maxIndSize_fullRun_30gens/median_Size_byGens.png} \\
+\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens/median_Size_byGens.png} &
+\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/concrete_2hl_maxIndSize_fullRun_30gens/median_Size_byGens.png}
+\end{tabular}
+
+## Auswertung Lösungsgröße (Median) - Symbolische Regression
 
@@
 df <- read.csv(
-  "/Users/rmn/github/master_thesis/data/summary_table_best_size_mean.csv",
+  "/Users/rmn/github/master_thesis/data/summary_table_best_size_median.csv",
   sep=",",
@@
 ```
-## Durchschnittliche Lösungsgröße - Symbolische Regression
+
+## Übersicht Populationsdiversität (Median) - Symbolische Regression
 
 \begin{tabular}{cc}
-\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/energyCooling_2hl_maxIndSize_fullRun_30gens/mean_Size_byGens.png} &
-\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/bostonHousing_2hl_maxIndSize_fullRun_30gens/mean_Size_byGens.png} \\
-\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens/mean_Size_byGens.png} &
-\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/concrete_2hl_maxIndSize_fullRun_30gens/mean_Size_byGens.png}
-\end{tabular}
-
-## Populationsdiversität - Symbolische Regression
-
-\begin{tabular}{cc}
-\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/energyCooling_2hl_maxIndSize_fullRun_30gens/mean_popDiversity_byGens_levOnly.png} &
-\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/bostonHousing_2hl_maxIndSize_fullRun_30gens/mean_popDiversity_byGens_levOnly.png} \\
-\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens/mean_popDiversity_byGens_levOnly.png} &
-\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/concrete_2hl_maxIndSize_fullRun_30gens/mean_popDiversity_byGens_levOnly.png}
+\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/energyCooling_2hl_maxIndSize_fullRun_30gens/median_popDiversity_byGens_levOnly.png} &
+\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/bostonHousing_2hl_maxIndSize_fullRun_30gens/median_popDiversity_byGens_levOnly.png} \\
+\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens/median_popDiversity_byGens_levOnly.png} &
+\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/concrete_2hl_maxIndSize_fullRun_30gens/median_popDiversity_byGens_levOnly.png}
 \end{tabular}
@@
 
-```{r full_run_realWorldSymReg_best_solution_size, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
+```{r full_run_realWorldSymReg_popDiversity, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
 library(knitr)
@@
 df <- read.csv(
-  "/Users/rmn/github/master_thesis/data/summary_table_mean_LevDistance.csv",
+  "/Users/rmn/github/master_thesis/data/summary_table_median_LevDistance.csv",
+  sep=",",
+  colClasses = c("character", "numeric", "numeric", "numeric", "character", "character")
+)
+
+# Create a new dataframe with the modified values of the two columns,
+df_bold <- df
+
+for(i in 1:nrow(df)){
+  if(df[i, "DAE.GP"] < df[i, "Pre.Trained_DAE.GP"]){
+    df_bold[i, "DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(round(df[i, "DAE.GP"],2)))
+    df_bold[i, "Pre.Trained_DAE.GP"] <- as.character(round(df[i, "Pre.Trained_DAE.GP"],2))
+    }
+  else if (df[i, "DAE.GP"] > df[i, "Pre.Trained_DAE.GP"]) {
+    df_bold[i, "Pre.Trained_DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(round(df[i, "Pre.Trained_DAE.GP"],2)))
+  df_bold[i, "DAE.GP"] <- as.character(round(df[i, "DAE.GP"],2))
+  }
+}
+# shorten column names
+col_names <- c("Problem", "Hid.Layers", "DAE-GP", "Pre-Trained", "P-Value", "Cliffs-Delta")
+
+knitr::kable(
+ df_bold,
+ format = "markdown",
+ col.names = col_names,
+ align = "c"
+ #caption = "Symbolic Regression Overview - Mean final Fitness"
+) %>% 
+  kable_styling(latex_options = "scale_down", font_size = 7)
+
+```
+
+## Übersicht Trainingsepochen pro Generation (Median) - Symbolische Regression
+
+\begin{tabular}{cc}
+\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/energyCooling_2hl_maxIndSize_fullRun_30gens/median_epochs_trained_perGen.png} &
+\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/bostonHousing_2hl_maxIndSize_fullRun_30gens/median_epochs_trained_perGen.png} \\
+\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens/median_epochs_trained_perGen.png} &
+\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/concrete_2hl_maxIndSize_fullRun_30gens/median_epochs_trained_perGen.png}
+\end{tabular}
+
+## Auswertung Trainingsepochen pro Generation (Median) - Symbolische Regression
+
+```{r full_run_realWorldSymReg_epochsPerGen, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
+library(knitr)
+library(kableExtra)
+library(dplyr)
+
+# Create the data frame from csv
+df <- read.csv(
+  "/Users/rmn/github/master_thesis/data/summary_table_mean_epochsTrained.csv",
   sep=",",

### 2023-01-15 data/summarize_csv_files.py

@@ -126,6 +126,24 @@ with open("/Users/rmn/github/master_thesis/data/summary_table_best_size_mean.csv
     f.write(csv_string)
 
 
+# summarize by median
+csv_string = "Problem,Hidden_Layer,DAE-GP,Pre-Trained_DAE-GP,P_Value,Cliffs_Delta\n"
+
+for problem,path in RESULTS.items():
+
+    d = json.load(open(join(path, FILENAME_SIZE),"r",encoding="utf-8"))
+
+    
+    reg_mean = np.median(d["DAE-GP"])
+    pt_mean = np.median(d["Pre-Trained"])
+
+
+    csv_string += f"{d['problem']},{d['hiddenLayer']},{reg_mean},{pt_mean},{getPVal(d['DAE-GP'], d['Pre-Trained'])},{cliffsDeltaPretty(d['Pre-Trained'], d['DAE-GP'])}\n"
+
+
+with open("/Users/rmn/github/master_thesis/data/summary_table_best_size_median.csv", "w", encoding="utf-8") as f:
+    f.write(csv_string)
+
 
 # Epochs Trained
 # ---
@@ -156,6 +174,26 @@ with open("/Users/rmn/github/master_thesis/data/summary_table_mean_epochsTrained
     f.write(csv_string)
 
 
+# summarize by median
+csv_string = "Problem,Hidden_Layer,DAE-GP,Pre-Trained_DAE-GP,P_Value,Cliffs_Delta\n"
+
+for problem,path in RESULTS.items():
+
+    d = json.load(open(join(path, FILENAME_EPOCHS),"r",encoding="utf-8"))
+
+    
+    reg_mean = np.median(np.array(d["DAE-GP"]), axis=0)
+    pt_mean = np.median(np.array(d["Pre-Trained"]), axis=0)
+
+    mean_reg_mean = np.mean(reg_mean)
+    mean_pt_mean = np.mean(pt_mean)
+
+
+    csv_string += f"{d['problem']},{d['hiddenLayer']},{mean_reg_mean},{mean_pt_mean},{getPVal(reg_mean, pt_mean)},{cliffsDeltaPretty(pt_mean, reg_mean)}\n"
+
+
+with open("/Users/rmn/github/master_thesis/data/summary_table_median_epochsTrained.csv", "w", encoding="utf-8") as f:
+    f.write(csv_string)
 
 
 

### 2023-01-15 data/summarize_csv_files.py
Sun Jan 15 15:00:04 CET 2023: update-remote.sh

@@ -3,9 +3,48 @@ from scipy.stats import mannwhitneyu as mwu
 from scipy.stats import normaltest
 from vargha_delahni_A import VD_A
 from cliffs_delta import cliffs_delta
-
-from os.path import join
+from matplotlib import pyplot as plt
 import numpy as np
+from matplotlib import style as mpl_style
+from os.path import join, exists
+from os import makedirs
+from json import load
+
+
+# matplotlib setup
+MPL_CONFIG = load(
+    open("/Users/rmn/masterThesis/eda-gp-2020/experiments/matplotlib_config.json", "r", encoding="utf-8")
+)
+
+mpl_style.use(MPL_CONFIG["mpl_style"])
+
+# font sizes
+SMALL=MPL_CONFIG["fonts"]["small"]
+MID=MPL_CONFIG["fonts"]["mid"]
+BIG=MPL_CONFIG["fonts"]["big"]
+
+# color codes
+C_REG=MPL_CONFIG["colors"]["dae-gp"]
+C_PT=MPL_CONFIG["colors"]["pt_dae-gp"]
+
+# marker codes
+M_TRAIN=MPL_CONFIG["marker"]["train"]
+M_TEST=MPL_CONFIG["marker"]["test"]
+
+TRAIN_LINESTYLE=MPL_CONFIG["train_line_style"]
+
+DPI=MPL_CONFIG["dpi"]
+
+
+IMG_PATH=f"{MPL_CONFIG['image_base_path']}/symbolicRegressionSummarized"
+
+def create_dir(dir_name):
+    if not exists(dir_name):
+        makedirs(dir_name)
+
+create_dir(IMG_PATH)
+BASE_TITLE="symbolicRegressionSummarized"
+
 
 RESULTS = {
     "airfoil_1hl" : "/Users/rmn/github/master_thesis/data/airfoil_1hl_maxIndSize_fullRun_30gens",
@@ -14,6 +53,17 @@ RESULTS = {
     "energyCooling": "/Users/rmn/github/master_thesis/data/energyCooling_2hl_FullRun_30gens",
     "concrete": "/Users/rmn/github/master_thesis/data/concrete_2hl_FullRun_30gens"
 }
+
+RESULTS_2hl = {
+    "airfoil_2hl" : "/Users/rmn/github/master_thesis/data/airfoil_2hl_maxIndSize_fullRun_30gens",
[…]

### 2023-01-15 paper.Rmd
Sun Jan 15 16:00:05 CET 2023: update-remote.sh

@@
 
-```{bash recompile_results, include=FALSE}
+```{bash recompile_results, include=FALSE, eval=FALSE}
 
@@
 # summarize csv files for full run experiments
-/Users/rmn/miniconda3/envs/dataScience/bin/python /Users/rmn/github/master_thesis/data/summarize_csv_files.py
+/Users/rmn/miniconda3/envs/dataScience/bin/python /Users/rmn/github/master_thesis/data/summarize_data.py
 
@@
 
-<!-- Overview Table  -->
+Next I studied the size of solutions based on using a pre-training strategy. Plot X shows the median size of the best solutions aggregated of all benchmark problems for symbolic regression
+
+![Median Solution Size - Symbolic Regression](~/masterThesis/master_thesis/img/symbolicRegressionSummarized/median_bestSol_size_byGens.png){height=80%}
+
+Table X summarizes the results of the statistical tests.
 
@@
 
-<!-- Overview Plots  -->
-
-\begin{tabular}{cc}
-\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/energyCooling_2hl_maxIndSize_fullRun_30gens/mean_Size_byGens.png} &
-\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/bostonHousing_2hl_maxIndSize_fullRun_30gens/mean_Size_byGens.png} \\
-\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens/mean_Size_byGens.png} &
-\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/concrete_2hl_maxIndSize_fullRun_30gens/mean_Size_byGens.png}
-\end{tabular}
-
 # Discussion

### 2023-01-15 slides.Rmd
Sun Jan 15 16:00:05 CET 2023: update-remote.sh

@@
 
-```{bash recompile_results, eval=TRUE, include=FALSE}
+```{bash recompile_results, eval=FALSE, include=FALSE}
 
@@
 
-![Airfoil - Verteilung der finalen Fitness (1HL)](~/masterThesis/master_thesis/img/airfoil_1hl_maxIndSize_fullRun_30gens/final_fit_boxplot.png){height=90%}
+![Airfoil - Verteilung der finalen Fitness (1HL)](~/masterThesis/master_thesis/img/airfoil_1hl_maxIndSize_fullRun_30gens/final_fit_boxplot.png){height=75%}
 
@@
 
-\begin{tabular}{cc}
-\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/energyCooling_2hl_maxIndSize_fullRun_30gens/median_fitness_byGens.png} &
-\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/bostonHousing_2hl_maxIndSize_fullRun_30gens/median_fitness_byGens.png} \\
-\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens/median_fitness_byGens.png} &
-\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/concrete_2hl_maxIndSize_fullRun_30gens/median_fitness_byGens.png}
-\end{tabular}
+![Übersicht Median Fitness - Symbolische Regression](~/masterThesis/master_thesis/img/symbolicRegressionSummarized/median_fitness_byGens.png){height=80%}
 
@@
 
-\begin{tabular}{cc}
-\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/energyCooling_2hl_maxIndSize_fullRun_30gens/median_Size_byGens.png} &
-\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/bostonHousing_2hl_maxIndSize_fullRun_30gens/median_Size_byGens.png} \\
-\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens/median_Size_byGens.png} &
-\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/concrete_2hl_maxIndSize_fullRun_30gens/median_Size_byGens.png}
-\end{tabular}
+![Übersicht Lösungsgröße (Median) - Symbolische Regression](~/masterThesis/master_thesis/img/symbolicRegressionSummarized/median_bestSol_size_byGens.png){height=80%}
 
@@
 
-\begin{tabular}{cc}
-\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/energyCooling_2hl_maxIndSize_fullRun_30gens/median_popDiversity_byGens_levOnly.png} &
-\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/bostonHousing_2hl_maxIndSize_fullRun_30gens/median_popDiversity_byGens_levOnly.png} \\
-\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens/median_popDiversity_byGens_levOnly.png} &
-\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/concrete_2hl_maxIndSize_fullRun_30gens/median_popDiversity_byGens_levOnly.png}
-\end{tabular}
+![Übersicht Populationsdiversität (Median) - Symbolische Regression](~/masterThesis/master_thesis/img/symbolicRegressionSummarized/median_levDistance_byGen.png){height=80%}
 
-## Auswertung Populationsdiversität - Symbolische Regression
+## Auswertung Populationsdiversität (Median) - Symbolische Regression
 
@@
 
-\begin{tabular}{cc}
-\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/energyCooling_2hl_maxIndSize_fullRun_30gens/median_epochs_trained_perGen.png} &
-\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/bostonHousing_2hl_maxIndSize_fullRun_30gens/median_epochs_trained_perGen.png} \\
-\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens/median_epochs_trained_perGen.png} &
-\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/concrete_2hl_maxIndSize_fullRun_30gens/median_epochs_trained_perGen.png}
-\end{tabular}
+![Übersicht Trainingsepochen pro Generation (Median) - Symbolische Regression](~/masterThesis/master_thesis/img/symbolicRegressionSummarized/median_epochsTrained.png){height=80%}
 

### 2023-01-15 data/summarize_data.py

@@ -1,7 +1,5 @@
 import json
 from scipy.stats import mannwhitneyu as mwu
-from scipy.stats import normaltest
-from vargha_delahni_A import VD_A
 from cliffs_delta import cliffs_delta
 from matplotlib import pyplot as plt
 import numpy as np
@@ -43,7 +41,7 @@ def create_dir(dir_name):
         makedirs(dir_name)
 
 create_dir(IMG_PATH)
-BASE_TITLE="symbolicRegressionSummarized"
+BASE_TITLE="Symbolic Regression"
 
 
 RESULTS = {
@@ -55,17 +53,14 @@ RESULTS = {
 }
 
 RESULTS_2hl = {
-    "airfoil_2hl" : "/Users/rmn/github/master_thesis/data/airfoil_2hl_maxIndSize_fullRun_30gens",
-    "bostonHousing": "/Users/rmn/github/master_thesis/data/bostonHousing_2hl_maxIndSize_fullRun_30gens",
+    "Airfoil" : "/Users/rmn/github/master_thesis/data/airfoil_2hl_maxIndSize_fullRun_30gens",
+    "BostonHousing": "/Users/rmn/github/master_thesis/data/bostonHousing_2hl_maxIndSize_fullRun_30gens",
     "energyCooling": "/Users/rmn/github/master_thesis/data/energyCooling_2hl_FullRun_30gens",
-    "concrete": "/Users/rmn/github/master_thesis/data/concrete_2hl_FullRun_30gens"
+    "Concrete": "/Users/rmn/github/master_thesis/data/concrete_2hl_FullRun_30gens"
 }
 
 
 gens = [x for x in range(0, 31)]
-
-
-
 
 
 def cliffsDeltaPretty(*args):
@@ -109,39 +104,36 @@ for problem,path in RESULTS.items():
     reg_mean_test = np.mean(d["DAE-GP (test)"])
     pt_mean_test = np.mean(d["Pre-Trained (test)"])
 
-    print(reg_mean_train)
-
     assert d["DAE-GP (train)"] != d["Pre-Trained (train)"]
     assert d["DAE-GP (test)"] != d["Pre-Trained (test)"]
 
 
[…]

### 2023-01-15 paper.Rmd
autogit

@@
 Table X summarizes the median best fitness achieved after running each algorithm for 10 runs using a population size of 500 individuals for 30 generations.^[P-Values are marked with asteriks (1,2,3) for ($\alpha$)-levels of $0.1$, $0.05$ and $0.01$].
+The median best fitness over all 30 generations are visualized in figure X.
 
@@
  col.names = col_names,
- align = "c"
- #caption = "Symbolic Regression Overview - Mean final Fitness"
+ align = "c",
+ caption = "Median Best Fitness - Symbolic Regression"
 ) %>% row_spec(0,bold=TRUE ) %>% kableExtra::kable_styling(latex_options = "hold_position")
@@
 
+![Median Best Fitness - Symbolic Regression](~/masterThesis/master_thesis/img/symbolicRegressionSummarized/median_fitness_byGens.png){height=80%}
+
 ## Solution Size in Symbolic Regression
 
-Next I studied the size of solutions based on using a pre-training strategy. Plot X shows the median size of the best solutions aggregated of all benchmark problems for symbolic regression
+Next I studied the size of solutions based on using a pre-training strategy. Figure X shows the median size of the best solutions aggregated of all benchmark problems for symbolic regression
 
@@
  col.names = col_names,
- align = "l"
- #caption = "Symbolic Regression Overview - Mean final Fitness"
+ align = "l",
+ caption = "Median Solution Size - Symbolic Regression"
 ) %>% row_spec(0,bold=FALSE) %>% kableExtra::kable_styling(latex_options = "hold_position", font_size = 7)
+
+```
+## Population Diversity
+
+Another interesting metric to examine in EC based metaheuristics is the diversity of the current generations population.
+Since the individual solutions in DAE-GP are computer programs in the form ob parse trees, I selected the normalized Levenshtein edit distance as my primary metric for tracking population diversity.
+figure X shows how the median population diversity develops over the full number of generations for all benchmark problems.
+
+![Median Population Diversity over Generations](~/masterThesis/master_thesis/img/symbolicRegressionSummarized/median_levDistance_byGen.png){height=80%}
+
+Testing for statistical differences in the underlying distributions yields the P-Values listed in table X.
+
+```{r full_run_realWorldSymReg_popDiversity, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
+library(knitr)
+library(kableExtra)
+library(dplyr)
+
+# Create the data frame from csv
+df <- read.csv(
+  "/Users/rmn/github/master_thesis/data/summary_table_median_LevDistance.csv",
+  sep=",",
+  colClasses = c("character", "numeric", "numeric", "numeric", "character", "character")
+)
+
+# Create a new dataframe with the modified values of the two columns,
+df_bold <- df
+
+for(i in 1:nrow(df)){
+  if(df[i, "DAE.GP"] < df[i, "Pre.Trained_DAE.GP"]){
+    df_bold[i, "DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(round(df[i, "DAE.GP"],2)))
+    df_bold[i, "Pre.Trained_DAE.GP"] <- as.character(round(df[i, "Pre.Trained_DAE.GP"],2))
+    }
+  else if (df[i, "DAE.GP"] > df[i, "Pre.Trained_DAE.GP"]) {
+    df_bold[i, "Pre.Trained_DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(round(df[i, "Pre.Trained_DAE.GP"],2)))
+  df_bold[i, "DAE.GP"] <- as.character(round(df[i, "DAE.GP"],2))
+  }
+}
+# shorten column names
+col_names <- c("Problem", "Hid.Layers", "DAE-GP", "Pre-Trained", "P-Value", "Cliffs-Delta")
+
+knitr::kable(
+ df_bold,
+ format = "markdown",
+ col.names = col_names,
+ align = "c",
+ caption = "Median Population Diversity over Generations - Symbolic Regression"
+) %>% 
+  kable_styling(latex_options = "scale_down", font_size = 7)
+
+```
+
+## Training epochs per Generation
+
+One important argument for using pre-training is the reduction of runtime and computational ressources for DAE-GP.
+An important factor in this, is the number of epochs that each DAE-LSTM has to be trained per generation.
+I studied the median number of training epochs that each generations DAE-LSTM $M_g$ (excluding the pre-training DAE-LSTM model $\hat{M}$) was trained until the training was stopped.
+The results are visualized in figure X and summarized in table X
+
+![Median Number of Training Epochs per Generation - Symbolic Regression](~/masterThesis/master_thesis/img/symbolicRegressionSummarized/median_epochsTrained.png){height=80%}
+
+```{r full_run_realWorldSymReg_epochsPerGen, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
+library(knitr)
+library(kableExtra)
+library(dplyr)
+
+# Create the data frame from csv
+df <- read.csv(
+  "/Users/rmn/github/master_thesis/data/summary_table_mean_epochsTrained.csv",
+  sep=",",
+  colClasses = c("character", "numeric", "numeric", "numeric", "character", "character")
+)
+
+# Create a new dataframe with the modified values of the two columns,
+df_bold <- df
+
+for(i in 1:nrow(df)){
+  if(df[i, "DAE.GP"] < df[i, "Pre.Trained_DAE.GP"]){
+    df_bold[i, "DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(round(df[i, "DAE.GP"],2)))
+    df_bold[i, "Pre.Trained_DAE.GP"] <- as.character(round(df[i, "Pre.Trained_DAE.GP"],2))
+    }
+  else if (df[i, "DAE.GP"] > df[i, "Pre.Trained_DAE.GP"]) {
+    df_bold[i, "Pre.Trained_DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(round(df[i, "Pre.Trained_DAE.GP"],2)))
+  df_bold[i, "DAE.GP"] <- as.character(round(df[i, "DAE.GP"],2))
+  }
+}
+# shorten column names
+col_names <- c("Problem", "Hid.Layers", "DAE-GP", "Pre-Trained", "P-Value", "Cliffs-Delta")
+
+knitr::kable(
+ df_bold,
+ format = "markdown",
+ col.names = col_names,
+ align = "c",
+ caption = "Median Number of Training Epochs per Generation - Symbolic Regression"
+) %>% 
+  kable_styling(latex_options = "scale_down", font_size = 7)
 

### 2023-01-15 data/summarize_data.py

@@ -118,7 +118,7 @@ FILE_FULL_FITNESS_DATA = "full_fitness_data.json"
 
 fig, ((ul, ur), (dl, dr)) = plt.subplots(nrows=2, ncols=2, sharex=True, sharey=False, dpi=DPI, layout="constrained")
 fig.set_size_inches(14,12)
-fig.suptitle(f"{BASE_TITLE} - Mean Best Fitness", fontsize=BIG)
+# fig.suptitle(f"{BASE_TITLE} - Mean Best Fitness", fontsize=BIG)
 fig.supxlabel("Generations", fontsize=MID)
 fig.supylabel("RMSE", fontsize=MID)
 
@@ -160,7 +160,7 @@ with open("/Users/rmn/github/master_thesis/data/summary_table_final_fit_median.c
 
 fig, ((ul, ur), (dl, dr)) = plt.subplots(nrows=2, ncols=2, sharex=True, sharey=False, dpi=DPI, layout="constrained")
 fig.set_size_inches(14,12)
-fig.suptitle(f"{BASE_TITLE} - Median Best Fitness", fontsize=BIG)
+#fig.suptitle(f"{BASE_TITLE} - Median Best Fitness", fontsize=BIG)
 fig.supxlabel("Generations", fontsize=MID)
 fig.supylabel("RMSE", fontsize=MID)
 
@@ -206,7 +206,7 @@ with open("/Users/rmn/github/master_thesis/data/summary_table_best_size_mean.csv
 
 fig, ((ul, ur), (dl, dr)) = plt.subplots(nrows=2, ncols=2, sharex=True, sharey=False, dpi=DPI, layout="constrained")
 fig.set_size_inches(14,12)
-fig.suptitle(f"{BASE_TITLE} - Mean Size of best Solution", fontsize=BIG)
+#fig.suptitle(f"{BASE_TITLE} - Mean Size of best Solution", fontsize=BIG)
 fig.supxlabel("Generations", fontsize=MID)
 fig.supylabel("Size", fontsize=MID)
 
@@ -242,7 +242,7 @@ with open("/Users/rmn/github/master_thesis/data/summary_table_best_size_median.c
 
 fig, ((ul, ur), (dl, dr)) = plt.subplots(nrows=2, ncols=2, sharex=True, sharey=False, dpi=DPI, layout="constrained")
 fig.set_size_inches(14,12)
-fig.suptitle(f"{BASE_TITLE} - Median Size of best Solution", fontsize=BIG)
[…]

### 2023-01-16 slides.Rmd
Mon Jan 16 09:00:08 CET 2023: update-remote.sh

@@
 
-# summarize csv files for full run experiments
+# summarize csv files and plots for full run experiments
 /Users/rmn/miniconda3/envs/dataScience/bin/python /Users/rmn/github/master_thesis/data/summarize_data.py

### 2023-01-16 slides.tex
Mon Jan 16 12:00:10 CET 2023: update-remote.sh

@@
 \item
-  Ersetzung der Variationsoperatoren von GP durch künstliche, neuronalen
-  Netzen zur Optimierung des Suchverhaltens\footnote<.->{Wittenberg,
+  Ersetzung der Variationsoperatoren von GP durch künstliche, neuronale
+  Netze zur Optimierung des Suchverhaltens\footnote<.->{Wittenberg,
     Rothlauf and Schweim
@@
 \item
-  Je mehr versteckte Layer oder die Anzahl an Neuronen pro verstecktem
-  Layer vorhanden sind, desto besserer Effekt des Pre-Trainings
+  Je mehr versteckte Layer oder Neuronen pro verstecktem Layer vorhanden
+  sind, desto besserer Effekt des Pre-Trainings
 \item
@@
   einer großen Population aus Lösungskandidaten \(\hat{P}_{train}\)
+  (ramped Half and Half Initialisierung)
 \item
@@
 
-\emph{Ausschluss anderer Pre-Training Strategien wie Re-Use Learning,
+\emph{Ausschluss weiterer Pre-Training Strategien wie Re-Use Learning,
 Few-Shot Learning}
@@
 \protect\hypertarget{pre-trained-dae-gp-ablauf}{}
-\includegraphics[width=\textwidth,height=0.75\textheight]{./img/flowcharts/pt-dae-gp.png}
+\begin{center}\includegraphics[width=14.28in]{./img/flowcharts/pt-dae-gp} \end{center}
 \end{frame}
@@
   Populationsdiversität
+\item
+  Anzahl an Trainingsepochen pro Generation
 \end{itemize}
@@
 \protect\hypertarget{airfoil---verteilung-der-finalen-fitness-1hl}{}
-\includegraphics[width=\textwidth,height=0.9\textheight]{~/masterThesis/master_thesis/img/airfoil_1hl_maxIndSize_fullRun_30gens/final_fit_boxplot.png}
+\includegraphics[width=\textwidth,height=0.75\textheight]{~/masterThesis/master_thesis/img/airfoil_1hl_maxIndSize_fullRun_30gens/final_fit_boxplot.png}
 \end{frame}
@@
 
-\begin{frame}{Symbolische Regression - Ergebnisübersicht}
-\protect\hypertarget{symbolische-regression---ergebnisuxfcbersicht}{}
+\begin{frame}{Übersicht Median Fitness - Symbolische Regression}
+\protect\hypertarget{uxfcbersicht-median-fitness---symbolische-regression}{}
+\includegraphics[width=\textwidth,height=0.8\textheight]{~/masterThesis/master_thesis/img/symbolicRegressionSummarized/median_fitness_byGens.png}
+\end{frame}
+
+\begin{frame}{Auswertung Fitness (Median) - Symbolische Regression}
+\protect\hypertarget{auswertung-fitness-median---symbolische-regression}{}
 \begin{longtable}[]{@{}
@@
 \endhead
-Airfoil & 1 & Train & \textbf{33.37} & 35.35 & 0.02** & 0.64 \\
-& 1 & Test & \textbf{33.73} & 35.61 & 0.14 & 0.4 \\
-Airfoil & 2 & Train & 17.48 & \textbf{14.72} & 0.31 & -0.28 \\
-& 2 & Test & 17.53 & \textbf{14.76} & 0.57 & -0.16 \\
-Boston\_Housing & 2 & Train & 8.19 & \textbf{8.08} & 0.29 & -0.29 \\
-& 2 & Test & 8.05 & \textbf{7.97} & 0.71 & -0.11 \\
-Energy(Cooling) & 2 & Train & \textbf{4.53} & 4.78 & 0.02** & 0.63 \\
-& 2 & Test & \textbf{4.67} & 4.92 & 0.29 & 0.29 \\
-Concrete & 2 & Train & 17.3 & \textbf{16.31} & 0.53 & -0.5 \\
-& 2 & Test & 16.67 & \textbf{16.19} & 0.80 & -0.25 \\
+Airfoil & 1 & Train & \textbf{33.55} & 35.11 & 0.02** & 0.64 \\
+& 1 & Test & \textbf{34.03} & 35.81 & 0.14 & 0.40 \\
+Airfoil & 2 & Train & 11.64 & \textbf{8.24} & 0.31 & -0.28 \\
+& 2 & Test & 11.34 & \textbf{8.06} & 0.57 & -0.16 \\
+Boston\_Housing & 2 & Train & 8.23 & \textbf{7.84} & 0.29 & -0.29 \\
+& 2 & Test & 8.04 & \textbf{8} & 0.71 & -0.11 \\
+Energy(Cooling) & 2 & Train & \textbf{4.42} & 4.65 & 0.02** & 0.63 \\
+& 2 & Test & \textbf{4.61} & 4.86 & 0.29 & 0.29 \\
+Concrete & 2 & Train & \textbf{16.58} & 16.83 & 0.97 & 0.02 \\
+& 2 & Test & \textbf{16.62} & 16.64 & 0.52 & 0.18 \\
 \bottomrule()
 \end{longtable}
-\end{frame}
-
-\begin{frame}{Symbolische Regression - Übersicht Median Fitness}
-\protect\hypertarget{symbolische-regression---uxfcbersicht-median-fitness}{}
-\begin{tabular}{cc}
-\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/energyCooling_2hl_maxIndSize_fullRun_30gens/median_fitness_byGens.png} &
-\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/bostonHousing_2hl_maxIndSize_fullRun_30gens/median_fitness_byGens.png} \\
-\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens/median_fitness_byGens.png} &
-\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/concrete_2hl_maxIndSize_fullRun_30gens/median_fitness_byGens.png}
-\end{tabular}
 \end{frame}
@@
 
-\begin{frame}{Auswertung Größe der besten Individuen - Symbolische
-Regression}
-\protect\hypertarget{auswertung-gruxf6uxdfe-der-besten-individuen---symbolische-regression}{}
+\begin{frame}{Übersicht Lösungsgröße (Median) - Symbolische Regression}
+\protect\hypertarget{uxfcbersicht-luxf6sungsgruxf6uxdfe-median---symbolische-regression}{}
+\includegraphics[width=\textwidth,height=0.8\textheight]{~/masterThesis/master_thesis/img/symbolicRegressionSummarized/median_bestSol_size_byGens.png}
+\end{frame}
+
+\begin{frame}{Auswertung Lösungsgröße (Median) - Symbolische Regression}
+\protect\hypertarget{auswertung-luxf6sungsgruxf6uxdfe-median---symbolische-regression}{}
 \begin{longtable}[]{@{}
@@
 \endhead
-Airfoil & 1 & \textbf{33.37} & 35.35 & 0.02** & 0.64 \\
-Airfoil & 2 & 17.48 & \textbf{14.72} & 0.31 & -0.28 \\
-Boston\_Housing & 2 & 8.19 & \textbf{8.08} & 0.29 & -0.29 \\
-Energy(Cooling) & 2 & \textbf{4.53} & 4.78 & 0.02** & 0.63 \\
-Concrete & 2 & 17.3 & \textbf{16.31} & 0.53 & -0.5 \\
+Airfoil & 1 & \textbf{33.55} & 35.11 & 0.02** & 0.64 \\
+Airfoil & 2 & 11.64 & \textbf{8.24} & 0.31 & -0.28 \\
+Boston\_Housing & 2 & 8.23 & \textbf{7.84} & 0.29 & -0.29 \\
+Energy(Cooling) & 2 & \textbf{4.42} & 4.65 & 0.02** & 0.63 \\
+Concrete & 2 & \textbf{16.58} & 16.83 & 0.97 & 0.02 \\
 \bottomrule()
@@
 
-\begin{frame}{Durchschnittliche Lösungsgröße - Symbolische Regression}
-\protect\hypertarget{durchschnittliche-luxf6sungsgruxf6uxdfe---symbolische-regression}{}
-\begin{tabular}{cc}
-\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/energyCooling_2hl_maxIndSize_fullRun_30gens/mean_Size_byGens.png} &
-\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/bostonHousing_2hl_maxIndSize_fullRun_30gens/mean_Size_byGens.png} \\
-\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens/mean_Size_byGens.png} &
-\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/concrete_2hl_maxIndSize_fullRun_30gens/mean_Size_byGens.png}
-\end{tabular}
+\begin{frame}{Übersicht Populationsdiversität (Median) - Symbolische
+Regression}
+\protect\hypertarget{uxfcbersicht-populationsdiversituxe4t-median---symbolische-regression}{}
+\includegraphics[width=\textwidth,height=0.8\textheight]{~/masterThesis/master_thesis/img/symbolicRegressionSummarized/median_levDistance_byGen.png}
 \end{frame}
 
-\hypertarget{weitere-ansuxe4tze}{%
-\section{Weitere Ansätze}\label{weitere-ansuxe4tze}}
+\begin{frame}{Auswertung Populationsdiversität (Median) - Symbolische
+Regression}
+\protect\hypertarget{auswertung-populationsdiversituxe4t-median---symbolische-regression}{}
+\begin{longtable}[]{@{}
+  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2267}}
+  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1600}}
+  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1333}}
+  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1733}}
+  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1200}}
+  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1867}}@{}}
+\toprule()
+\begin{minipage}[b]{\linewidth}\centering
+Problem
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+Hid.Layers
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+DAE-GP
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+Pre-Trained
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+P-Value
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+Cliffs-Delta
+\end{minipage} \\
+\midrule()
+\endhead
+Airfoil & 1 & \textbf{0.34} & 0.49 & 0.02** & 0.36 \\
+Airfoil & 2 & \textbf{0.91} & 0.93 & 0.00*** & 0.88 \\
+Boston\_Housing & 2 & \textbf{0.94} & 0.95 & 0.00*** & 0.82 \\
+Energy(Cooling) & 2 & \textbf{0.93} & 0.94 & 0.00*** & 0.80 \\
+Concrete & 2 & \textbf{0.93} & 0.94 & 0.00*** & 0.88 \\
+\bottomrule()
+\end{longtable}
+\end{frame}
 
-\begin{frame}{Weitere Ansätze}
-\ldots{}
+\begin{frame}{Übersicht Trainingsepochen pro Generation (Median) -
+Symbolische Regression}
+\protect\hypertarget{uxfcbersicht-trainingsepochen-pro-generation-median---symbolische-regression}{}
+\includegraphics[width=\textwidth,height=0.8\textheight]{~/masterThesis/master_thesis/img/symbolicRegressionSummarized/median_epochsTrained.png}
+\end{frame}
+
+\begin{frame}{Auswertung Trainingsepochen pro Generation (Median) -
+Symbolische Regression}
+\protect\hypertarget{auswertung-trainingsepochen-pro-generation-median---symbolische-regression}{}
+\begin{longtable}[]{@{}
+  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2222}}
+  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}
+  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1111}}
+  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1806}}
+  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1250}}
+  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1944}}@{}}
+\toprule()
+\begin{minipage}[b]{\linewidth}\centering
+Problem
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+Hid.Layers
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+DAE-GP
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+Pre-Trained
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+P-Value
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+Cliffs-Delta
+\end{minipage} \\
+\midrule()
+\endhead
+Airfoil & 1 & 190.93 & \textbf{60} & 0.00*** & -0.94 \\
+Airfoil & 2 & 151.87 & \textbf{64.95} & 0.00*** & -0.88 \\
+Boston\_Housing & 2 & 182.84 & \textbf{67.27} & 0.00*** & -0.88 \\
+Energy\_Cooling & 2 & 169.39 & \textbf{69.87} & 0.00*** & -0.90 \\
+concrete & 2 & 171.69 & \textbf{64.9} & 0.00*** & -0.92 \\
+\bottomrule()
+\end{longtable}
+\end{frame}
+
+\hypertarget{weitere-verfolgte-ansuxe4tze}{%
+\section{Weitere verfolgte Ansätze}\label{weitere-verfolgte-ansuxe4tze}}
+
+\begin{frame}{Eingeschobenes Pre-Training in der 2. Generation}
+\protect\hypertarget{eingeschobenes-pre-training-in-der-2.-generation}{}
+Ansatz: Deutliche Fitnessverbesserung finden oft bereits in der ersten
+Generationen der evolutionären Suche statt
+
+\begin{itemize}
+\tightlist
+\item
+  Sampling des DAE-LSTMs der ersten Generation \(M_1\) zur
+  Initialisierung der Pre-Training Population \(\hat{P}\)
+\item
+  Training des pre-training modells \(\hat{M}\) mit \(\hat{P}\) nach
+  Abschluss der ersten Generation
+\item
+  DAE-LSTM Modelle werden ab Generation 2 mit den Parametern von
+  \(\hat{M}\) initialisiert
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Eingeschobenes Pre-Training in der 2. Generation - Airfoil
+Datensatz}
+\protect\hypertarget{eingeschobenes-pre-training-in-der-2.-generation---airfoil-datensatz}{}
+\includegraphics[width=\textwidth,height=0.8\textheight]{~/masterThesis/master_thesis/img/2ndGenPT_airfoil_2hl_150hn_fullRun/median_fitness_byGens.png}
+\end{frame}
+
+\begin{frame}{Grow Initialisierung der Pre-Training Population}
+\protect\hypertarget{grow-initialisierung-der-pre-training-population}{}
 \end{frame}

### 2023-01-16 slides.Rmd
Mon Jan 16 13:00:08 CET 2023: update-remote.sh

@@
 knitr::opts_chunk$set(echo = TRUE)
-knitr::opts_chunk$set(fig.pos = "H")
+knitr::opts_chunk$set(fig.pos = "c")
 
@@
 
+# 2nd generation Pre-Training
+/Users/rmn/miniconda3/envs/dataScience/bin/python /Users/rmn/masterThesis/eda-gp-2020/experiments/2ndGenPT_airfoil_2hl_150hn_fullRun/re_evaluate_2ndGenPT_airfoil_2hl_150hn_fullRun.sh
+
 ```
@@
 * Metaheuristik basierend auf genetischer Programmierung (GP)
-* Ersetzung der Variationsoperatoren von GP durch künstliche, neuronalen Netzen zur Optimierung des Suchverhaltens^[@dae-gp_2020_rtree]
+* Ersetzung der Variationsoperatoren von GP durch künstliche, neuronale Netze zur Optimierung des Suchverhaltens^[@dae-gp_2020_rtree]
 * Variante des Estimation of Distribution-GP (EDA-GP)
@@
 * Positiver Effekt steigt mit zunehmender Komplexitität des DAE
-* Je mehr versteckte Layer oder die Anzahl an Neuronen pro verstecktem Layer vorhanden sind, desto besserer Effekt des Pre-Trainings
+* Je mehr versteckte Layer oder Neuronen pro verstecktem Layer vorhanden sind, desto besserer Effekt des Pre-Trainings
 * Für sehr kleine DAE, zeigt Pre-Training jedoch inverse, negative Auswirkung auf die Modell Performance
@@
 
-* (Klassisches) Pre-Training: Einmaliges Pre-Training eines DAE-LSTM mit einer großen Population aus Lösungskandidaten $\hat{P}_{train}$
+* (Klassisches) Pre-Training: Einmaliges Pre-Training eines DAE-LSTM mit einer großen Population aus Lösungskandidaten $\hat{P}_{train}$ (ramped Half and Half Initialisierung)
 
@@
 
-*Ausschluss anderer Pre-Training Strategien wie Re-Use Learning, Few-Shot Learning*
+*Ausschluss weiterer Pre-Training Strategien wie Re-Use Learning, Few-Shot Learning*
 
@@
 
-![Flowchart - Pre-Trained DAE-GP](./img/flowcharts/pt-dae-gp.png){height=75%}
+![Flowchart - Pre-Trained DAE-GP](./img/flowcharts/pt-dae-gp.png){height=70%}
 
@@
 * Populationsdiversität
+* Anzahl an Trainingsepochen pro Generation
 
@@
 
-![Airfoil - Entwicklung der Fitness über Generationen (1 HL)](~/masterThesis/master_thesis/img/airfoil_1hl_maxIndSize_fullRun_30gens/mean_median_fitness_byGens.png)
+![Airfoil - Entwicklung der Fitness über Generationen (1 HL)](~/masterThesis/master_thesis/img/airfoil_1hl_maxIndSize_fullRun_30gens/median_fitness_byGens.png){height=85%}
 
@@
 
-# Weitere Ansätze
+# Weitere verfolgte Ansätze
 
-...
+## Eingeschobenes Pre-Training in der 2. Generation
+
+Ansatz: Deutliche Fitnessverbesserung finden oft bereits in der ersten Generationen der evolutionären Suche statt
+
+* Sampling des DAE-LSTMs der ersten Generation $M_1$ zur Initialisierung der Pre-Training Population $\hat{P}$
+* Training des pre-training modells $\hat{M}$ mit $\hat{P}$ nach Abschluss der ersten Generation
+* DAE-LSTM Modelle werden ab Generation 2 mit den Parametern von $\hat{M}$ initialisiert
+
+## Eingeschobenes Pre-Training in der 2. Generation - Airfoil Datensatz
+
+![2nd Gen Pre-Training - Fitness - Airfoil Datensatz](~/masterThesis/master_thesis/img/2ndGenPT_airfoil_2hl_150hn_fullRun/median_fitness_byGens.png){height=80%}
+
+## Grow Initialisierung der Pre-Training Population
 

### 2023-01-16 slides.Rmd
Mon Jan 16 16:00:13 CET 2023: update-remote.sh

@@
 
-## Herausforderungen
+## Herausforderung - DAE-LSTM Dimension
 
-... hidden neurons ...
+* Im Verlaufe des DAE-GP algorithmus passt sich die Dimension der DAE-LSTM Netzwerke $M_{g}$ dynamisch an die aktuelle Population an
+* Anzahl der versteckten Neuronen pro verstecktem Layer in generation $g$ <- maximum der Länge aller Individuen aus $P_{g-1}$
 
@@
 
-![Airfoil - Erste Generation Median Trainingsfehler - Variable Anzahl Hidden Neurons (2/2)](./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_neurons_6plots.png){height=85%}
+![Airfoil - Erste Generation Median Trainingsfehler - Variable Anzahl Hidden Neurons (2/2)](./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_neurons_3plots.png){height=85%}
 

### 2023-01-16 slides.Rmd
autogit

@@
 
-## Herausforderung - DAE-LSTM Dimension
-
-* Im Verlaufe des DAE-GP algorithmus passt sich die Dimension der DAE-LSTM Netzwerke $M_{g}$ dynamisch an die aktuelle Population an
-* Anzahl der versteckten Neuronen pro verstecktem Layer in generation $g$ <- maximum der Länge aller Individuen aus $P_{g-1}$
-
 # Untersuchung des Generalisierungsverhalten
@@
 ```
+
+## Herausforderung - DAE-LSTM Dimension
+
+* Im Verlaufe des DAE-GP algorithmus passt sich die Dimension der DAE-LSTM Netzwerke $M_{g}$ dynamisch an die aktuelle Population an
+
+> Anzahl der versteckten Neuronen pro verstecktem Layer in generation $g$ = Maximum der Länge aller Individuen aus $P_{g-1}$
+
+Der Einsatz der Pre-Training Strategie ist nur möglich bei einer konstanten Anzahl an Neuronen. Zwei Ansätze verfolgt:
+
+1. DAE-GP mit dynamischer Anpassung, Pre-Trained DAE-GP mit konstanter Anzahl (maximale Länge innerhalb von $\hat{P}$)
+
+  * Vorteil: DAE-GP Performance konsistent mit bisherigen Publikationen, deutliche Reduktion der Laufzeit
+  * Nachteil: Erschwerte Vergleichbarkeit beider Algorithmen
+
+2. Beide Algorithmen mit einer fixen Anzahl von versteckten Neuronen
+
+  * Vorteil: Bessere Vergleichbarkeit zwischen beiden Algorithmen
+  * Nachteile: Starke erhöhung der Laufzeit für DAE-GP algorithmus
+
+## Airfoil - Entwicklung DAE-LSTM Dimension
+
+![Airfoil - Entwicklung DAE-LSTM Dimension](~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens/median_hiddenNeurons_perGen.png){height=90%}
 

### 2023-01-17 slides.Rmd
Tue Jan 17 12:00:09 CET 2023: update-remote.sh

@@
 
-## Eingeschobenes Pre-Training in der 2. Generation - Airfoil Datensatz
+## Grow Initialisierung der Pre-Training Population
+
+Ansatz: Initialisierung von $\hat{P}$ ausschließlich mit der Grow Methode anstelle von ramped Half and Half
+
+Hintergrund: Vermutlich geringer Informationsgewinn aus dem Lernen von vollen Lösungsbäume aus $\hat{P}$ mit den gesampleten Populationen der späteren Generationen
+
+## Übersicht Fitness (Median) - Weitere Ansätze
+
+![Übersicht Fitness pro Generation (Median) - Weitere Ansaetze](~/masterThesis/master_thesis/img/PT_Strategies_airfoil_2hl_150hn_fullRun/median_fitness_byGens.png){height=80%}
 

### 2023-01-17 data/summarize_data.py

@@ -431,3 +431,104 @@ for (problem,path), ax in zip(RESULTS_2hl.items(), (ul, ur, dl, dr)):
 
 ur.legend(fontsize=MID)
 fig.savefig(f"{IMG_PATH}/median_levDistance_byGen.png")
+
+
+
+PT_STRATEGIES_PATH = "/Users/rmn/github/master_thesis/data/PT_Strategies_airfoil_2hl_150hn_fullRun"
+
+FILES = {
+    "Norm._Levenshtein_Distance" : "lev_div_overGens.json",
+    "Size_best_Solution" : "size_best_solution.json",
+    "Fitness_lastGen" : "fitness_per_gen.json",
+    "Training_Epochs" : "epochs_trained_perGen.json"
+}  
+
+
+# summarize by mean
+
+for metric, filename in FILES.items():
+
+    d = json.load(open(join(PT_STRATEGIES_PATH, filename),"r",encoding="utf-8"))
+
+    if metric == "Fitness_lastGen":
+        csv_string = "Set,DAE-GP,2nd_Gen-PT,P_Value,Grow-PT,P_Value\n"
+
+        reg_mean_train = np.mean(np.array(list(sublist[-1] for sublist in d["DAE-GP(train)"])))
+        sec_pt_mean_train = np.mean(np.array(list(sublist[-1] for sublist in d["2nd.Gen-PT(train)"])))
+        grow_pt_mean_train = np.mean(np.array(list(sublist[-1] for sublist in d["Grow-PT(train)"])))
+
+        reg_mean_test = np.mean(np.array(list(sublist[-1] for sublist in d["DAE-GP(test)"])))
+        sec_pt_mean_test = np.mean(np.array(list(sublist[-1] for sublist in d["2nd.Gen-PT(test)"])))
+        grow_pt_mean_test = np.mean(np.array(list(sublist[-1] for sublist in d["Grow-PT(test)"])))
+
+        csv_string += f"Train,{np.mean(reg_mean_train)},{np.mean(sec_pt_mean_train)},{getPVal(reg_mean_train, sec_pt_mean_train)},{np.mean(grow_pt_mean_train)},{getPVal(reg_mean_train, grow_pt_mean_train)}\n"
+        csv_string += f"Test,{np.mean(reg_mean_test)},{np.mean(sec_pt_mean_test)},{getPVal(reg_mean_test, sec_pt_mean_test)},{np.mean(grow_pt_mean_test)},{getPVal(reg_mean_test, grow_pt_mean_test)}\n"
[…]

### 2023-01-17 slides.Rmd
autogit

@@
 
-![2nd Gen Pre-Training - Fitness - Airfoil Datensatz](~/masterThesis/master_thesis/img/2ndGenPT_airfoil_2hl_150hn_fullRun/median_fitness_byGens.png){height=80%}
+## Auswertung Fitness (Median) - Weitere Ansätze
 
-## Grow Initialisierung der Pre-Training Population
+```{r full_run_PT_Strat_Fitness, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
+library(knitr)
+library(kableExtra)
+library(dplyr)
+
+# Create the data frame from csv
+df <- read.csv(
+  "/Users/rmn/github/master_thesis/data/summary_median_Fitness_lastGen_PT_Strategies_Airfoil_2hl_150hn.csv",
+  sep=","
+)
+
+col_names <- c("Set", "DAE-GP", "2ndGen", "P-Val(1)" , "GrowInit", "P-Val(2)")
+
+knitr::kable(
+ df,
+ format = "markdown",
+ align = "l",
+ digits=2,
+ col.names = col_names,
+ #caption = "Symbolic Regression Overview - Mean final Fitness"
+) %>% 
+  kable_styling(latex_options = "scale_down", font_size = 7)
+
+```
+
+## Übersicht Lösungsgröße (Median) - Weitere Ansätze
+
+![Übersicht Lösungsgröße der besten Individuen (Median) - Weitere Ansaetze](~/masterThesis/master_thesis/img/PT_Strategies_airfoil_2hl_150hn_fullRun/median_Size_best_solution.png){height=80%}
+
+## Auswertung Lösungsgröße (Median) - Weitere Ansätze
+
+```{r full_run_PT_Strat_BestSize, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
+library(knitr)
+library(kableExtra)
+library(dplyr)
+
+# Create the data frame from csv
+df <- read.csv(
+  "/Users/rmn/github/master_thesis/data/summary_median_Size_best_Solution_PT_Strategies_Airfoil_2hl_150hn.csv",
+  sep=","
+)
+
+col_names <- c("DAE-GP", "2ndGen", "P-Val(1)" , "GrowInit", "P-Val(2)")
+
+knitr::kable(
+ df,
+ format = "markdown",
+ align = "l",
+ digits=2,
+ col.names = col_names,
+ #caption = "Symbolic Regression Overview - Mean final Fitness"
+) %>% 
+  kable_styling(latex_options = "scale_down", font_size = 7)
+
+```
+
+## Übersicht Populationsdiversität (Median) - Weitere Ansätze
+
+![Übersicht Populationsdiversität (Median) - Weitere Ansaetze](~/masterThesis/master_thesis/img/PT_Strategies_airfoil_2hl_150hn_fullRun/mean_popDiversity_byGens_levOnly.png){height=80%}
+
+## Auswertung Populationsdiversität (Median) - Weitere Ansätze
+
+```{r full_run_PT_Strat_LevDiv, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
+library(knitr)
+library(kableExtra)
+library(dplyr)
+
+# Create the data frame from csv
+df <- read.csv(
+  "/Users/rmn/github/master_thesis/data/summary_median_Norm._Levenshtein_Distance_PT_Strategies_Airfoil_2hl_150hn.csv",
+  sep=","
+)
+
+col_names <- c("DAE-GP", "2ndGen", "P-Val(1)" , "GrowInit", "P-Val(2)")
+
+knitr::kable(
+ df,
+ format = "markdown",
+ align = "l",
+ digits=2,
+ col.names = col_names,
+ #caption = "Symbolic Regression Overview - Mean final Fitness"
+) %>% 
+  kable_styling(latex_options = "scale_down", font_size = 7)
+
+```
+## Übersicht Trainingsepochen (Median) - Weitere Ansätze
+
+![Übersicht Trainingsepochen (Median) - Weitere Ansaetze](~/masterThesis/master_thesis/img/PT_Strategies_airfoil_2hl_150hn_fullRun/mean_epochs_trained_perGen.png){height=80%}
+
+## Auswertung Trainingsepochen (Median) - Weitere Ansätze
+
+```{r full_run_PT_Strat_EpochsTrained, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
+library(knitr)
+library(kableExtra)
+library(dplyr)
+
+# Create the data frame from csv
+df <- read.csv(
+  "/Users/rmn/github/master_thesis/data/summary_median_Training_Epochs_PT_Strategies_Airfoil_2hl_150hn.csv",
+  sep=","
+)
+
+col_names <- c("DAE-GP", "2ndGen", "P-Val(1)" , "GrowInit", "P-Val(2)")
+
+knitr::kable(
+ df,
+ format = "markdown",
+ align = "l",
+ digits=2,
+ col.names = col_names,
+ #caption = "Symbolic Regression Overview - Mean final Fitness"
+) %>% 
+  kable_styling(latex_options = "scale_down", font_size = 7)
+
+```
 

### 2023-01-17 slides.Rmd
Tue Jan 17 15:00:10 CET 2023: update-remote.sh
autogit

@@
 
-## Einfluss der Anzahl von Hidden Layers (1/2)
-
-![Airfoil - Erste Generation Median Trainingsfehler - Variable Anzahl Hidden Layer (1/2)](./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_layers_6plots.png){height=85%}
-
-## Einfluss der Anzahl von Hidden Layers (2/2)
+## Einfluss der Anzahl von Hidden Layers
 
@@
 
-## Einfluss der Anzahl von Hidden Neurons (1/2)
-
-![Airfoil - Erste Generation Median Trainingsfehler - Variable Anzahl Hidden Neurons (1/2)](./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_neurons_6plots.png){height=85%}
-
-## Einfluss der Anzahl von Hidden Neurons (2/2)
+## Einfluss der Anzahl von Hidden Neurons
 
@@
 knitr::kable(
-  read.csv("./tables/airfoil_function_set"),
-  # caption="Funktions Set"
-)%>% row_spec(0,bold=TRUE ) %>% kableExtra::kable_styling(latex_options = "hold_position")
+  read.csv("./tables/airfoil_function_set", sep = ","),
+) %>% kableExtra::kable_styling(latex_options = "hold_position")
 ```
@@
 
-* Im Verlaufe des DAE-GP algorithmus passt sich die Dimension der DAE-LSTM Netzwerke $M_{g}$ dynamisch an die aktuelle Population an
+* Im Verlaufe des DAE-GP Algorithmus passt sich die Dimension der DAE-LSTM Netzwerke $M_{g}$ dynamisch an die aktuelle Population an
 
@@
 1. DAE-GP mit dynamischer Anpassung, Pre-Trained DAE-GP mit konstanter Anzahl (maximale Länge innerhalb von $\hat{P}$)
-
-  * Vorteil: DAE-GP Performance konsistent mit bisherigen Publikationen, deutliche Reduktion der Laufzeit
-  * Nachteil: Erschwerte Vergleichbarkeit beider Algorithmen
-
 2. Beide Algorithmen mit einer fixen Anzahl von versteckten Neuronen
 
-  * Vorteil: Bessere Vergleichbarkeit zwischen beiden Algorithmen
-  * Nachteile: Starke erhöhung der Laufzeit für DAE-GP algorithmus
-
-## Airfoil - Entwicklung DAE-LSTM Dimension
+## DAE-GP - Entwicklung DAE-LSTM Dimension
 
@@
 
+## Airfoil - Laufzeit
+
+...
+
 ## Kontrollexperiment: Reduzierung der DAE-LSTM Dimension für Airfoil Datensatz
@@
 Frage: Welchen Einfluss hat die Reduktion der DAE-LSTM auf 1 hidden Layer (HL)?
+
+Erwartung: Negativer Einfluss von Pre-Training auf die Güte der gefundenen Lösungen (Fitness/Größe) durch sinkenden Modell Performance der DAE-LSTMs^[@pmlr-v5-erhan09a]
 
@@
 
-Airfoil Datensatz: Ergebnis deuten bei einer ausreichenden Anzahl von Hidden Layern auf einen postitiven Effekt der Pre-Training Strategie hin:
+Airfoil Datensatz: Erste Ergebnis deuten bei einer ausreichenden Anzahl von Hidden Layern auf einen postitiven Effekt der Pre-Training Strategie hin:
 
@@
 ```
-
-## Zusammenfassung Fitness - Symbolische Regression
-
-Keine Evidenz für einen statistisch signifikanten Einfluss von Pre-Training auf die erzielte Lösungsqualität in den durchgeführten Experimenten!
-
-* Airfoil:
-* Energy Cooling:
-* Boston Housing: 
-* Concrete:
 
@@
 ```
+
+## Zusammenfassung- Symbolische Regression
+
+Keine Evidenz für einen statistisch signifikanten Einfluss von Pre-Training auf die erzielte Lösungsqualität oder die erzielte Lösungsgröße in den durchgeführten Experimenten!
+
+Vorteile durch Pre-Training:
+
+1. Signifikante Erhöhung der Populationsdiversität in allen Experimenten
+2. Signifikante Reduktion der Anzahl an Trainingsepochen pro Generationen in allen Experimenten
 
@@
 ```
+# Fazit (vorläufig)
+
+## Vorteile
+
+Der Einsatz von Pre-Training dient:
+
+* der Erhöhung der Populationsdiversität (mögliche Strategie zur Kontrolle des Explore/Exploit Verhaltens von DAE-GP?)
+
+* der Reduktion der Trainingsepochen pro Generation (Reduktion der Rechenleistung, insbesondere bei komplexen DAE-LSTM nützlich?)
+
+## Nachteile
+
+* DAE-GP verliert durch Einsatz von Pre-Training die Möglichkeit die Anzahl künstlicher Neuronen pro verstecktem Layer anzupassen
+
+* Keine signifikanten Vorteile für die Qualität der gefundenen Lösungen
 

### 2023-01-18 slides.Rmd
Wed Jan 18 14:00:10 CET 2023: update-remote.sh
Wed Jan 18 15:00:08 CET 2023: update-remote.sh

@@
 
-  1. das Generalisierungsverhalten von DAE-GP?
-  2. die Qualität der gefundenen Programme (Fitness/Programmlänge)?
-  3. die Populationsdiversität?
-  4. das Laufzeitverhalten?
+  1. die Qualität der gefundenen Programme (Fitness/Programmlänge)?
+  2. die Populationsdiversität?
+  3. das Laufzeitverhalten?
 
-Anwendungsgebiet: Symbolische Regression, insbesondere am Aifoil Datensatz
+Anwendungsgebiet: Symbolische Regression, Fokus auf Aifoil Datensatz
 
@@
 
-# Untersuchung des Generalisierungsverhalten
-
-## Fragestellung
-
-Welchen Einfluss hat die Dimension der eingesetzten DAE-LSTMs auf das Generalisierungsverhalten von DAE-GP in Kombination mit Pre-Training?
-
-Experimenteller Aufbau:
-
-* Ausschließliche Betrachtung des Rekonstruktionsfehlers der ersten Generation (Nutzung von seperaten Test- und Trainingspopulationen)
-* DAE-LSTM Training erfolgt über eine fixe Anzahl von $1000$ Epochen (kein frühzeitiger Abbruch bei Konvergenz des Testfehlers)
-
-Insgesamt 12 Subexperimente mit jeweils 10 Durchläufen ($120$ Gesamtdurchläufe):
-
-* Variable Anzahl von Hidden Neurons (50, 100, 200) mit einem Hidden Layer
-* Variable Anzahl von Hidden Layers (1, 2, 3) mit 100 Hidden Neurons
-* (normales) DAE-GP und pre-trained DAE-GP
-
-## Einfluss der Anzahl von Hidden Layers
-
-![Airfoil - Erste Generation Median Trainingsfehler - Variable Anzahl Hidden Layer (2/2)](./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_layers_3plots.png){height=85%}
-
-## Einfluss der Anzahl von Hidden Neurons
-
-![Airfoil - Erste Generation Median Trainingsfehler - Variable Anzahl Hidden Neurons (2/2)](./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_neurons_3plots.png){height=85%}
-
-## Fazit Generalisierungsverhalten
-
-...
-
 # Untersuchung des Suchverhaltens bei symbolischer Regression
@@
 * Populationsdiversität
-* Anzahl an Trainingsepochen pro Generation
+* Laufzeit
 
@@
 
-![Airfoil Datensatz - Fitness über Generationen](~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens_withGP/mean_median_fitness_byGens.png){height=90%}
+![Airfoil Datensatz - Fitness über Generationen](~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens_withGP/median_fitness_byGens.png){height=90%}
 
@@
 
-![Airfoil Datensatz - Größe der besten Lösung über Generationen](~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens_withGP/mean_Size_byGens.png)
+![Airfoil Datensatz - Größe der besten Lösung über Generationen](~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens_withGP/median_best_Size_byGens.png)
 
-## Airfoil - Laufzeit
+## Airfoil - Populationsdiversität
 
-...
+![Airfoil - Populationsdiversität](/Users/rmn/github/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens_withGP/median_popDiversity_byGens_levOnly_withGP.png){height=90%}
+
+## Airfoil - Laufzeit (Gesamt)
+
+![Airfoil - Laufzeit](/Users/rmn/github/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens_TimeBenchmark/median_runtime.png){height=90%}
+
+## Airfoil - Laufzeit (Trainingsepochen)
+
+![Airfoil - Laufzeit](/Users/rmn/github/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens_TimeBenchmark/median_training_epochs.png){height=90%}
+
+## Airfoil - Laufzeit (Sampling Time)
+
+![Airfoil - Laufzeit](/Users/rmn/github/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens_TimeBenchmark/median_sampleTime.png){height=90%}
 
@@
 
+<!--
+
+# Untersuchung des Generalisierungsverhalten
+
+## Fragestellung
+
+Welchen Einfluss hat die Dimension der eingesetzten DAE-LSTMs auf das Generalisierungsverhalten von DAE-GP in Kombination mit Pre-Training?
+
+Experimenteller Aufbau:
+
+* Ausschließliche Betrachtung des Rekonstruktionsfehlers der ersten Generation (Nutzung von seperaten Test- und Trainingspopulationen)
+* DAE-LSTM Training erfolgt über eine fixe Anzahl von $1000$ Epochen (kein frühzeitiger Abbruch bei Konvergenz des Testfehlers)
+
+Insgesamt 12 Subexperimente mit jeweils 10 Durchläufen ($120$ Gesamtdurchläufe):
+
+* Variable Anzahl von Hidden Neurons (50, 100, 200) mit einem Hidden Layer
+* Variable Anzahl von Hidden Layers (1, 2, 3) mit 100 Hidden Neurons
+* (normales) DAE-GP und pre-trained DAE-GP
+
+## Einfluss der Anzahl von Hidden Layers
+
+![Airfoil - Erste Generation Median Trainingsfehler - Variable Anzahl Hidden Layer (2/2)](./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_layers_3plots.png){height=85%}
+
+## Einfluss der Anzahl von Hidden Neurons
+
+![Airfoil - Erste Generation Median Trainingsfehler - Variable Anzahl Hidden Neurons (2/2)](./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_neurons_3plots.png){height=85%}
+
+-->
+
 # References {.allowframebreaks}

### 2023-01-18 slides.Rmd
Wed Jan 18 16:00:14 CET 2023: update-remote.sh

@@
 
-```{bash recompile_results, eval=FALSE, include=FALSE}
+```{bash recompile_results, echo=TRUE, eval=FALSE, include=FALSE}
 
@@
 # 2nd generation Pre-Training
-/Users/rmn/miniconda3/envs/dataScience/bin/python /Users/rmn/masterThesis/eda-gp-2020/experiments/2ndGenPT_airfoil_2hl_150hn_fullRun/re_evaluate_2ndGenPT_airfoil_2hl_150hn_fullRun.sh
+bash /Users/rmn/masterThesis/eda-gp-2020/experiments/2ndGenPT_airfoil_2hl_150hn_fullRun/re_evaluate_2ndGenPT_airfoil_2hl_150hn_fullRun.sh
+
+echo Done recompiling Results, Plots, Diagrams, etc!
 
@@
 
+<!--
 ## DAE-GP - Entwicklung DAE-LSTM Dimension
@@
 ![Airfoil - Entwicklung DAE-LSTM Dimension](~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens/median_hiddenNeurons_perGen.png){height=90%}
+
+--->
 

### 2023-01-19 slides.Rmd
Wed Jan 18 17:00:10 CET 2023: update-remote.sh
Wed Jan 18 18:00:09 CET 2023: update-remote.sh
Thu Jan 19 08:00:07 CET 2023: update-remote.sh

@@
 BACKGROUND="white"
+
+if [ ! -d "./img/flowcharts" ]; then
+  mkdir -p "./img/flowcharts"
+fi
 
@@
 
+## Herausforderung - DAE-LSTM Dimension
+
+* Im Verlaufe des DAE-GP Algorithmus passt sich die Dimension der DAE-LSTM Netzwerke $M_{g}$ dynamisch an die aktuelle Population an
+
+> Anzahl der versteckten Neuronen pro verstecktem Layer in generation $g$ = Maximum der Länge aller Individuen aus $P_{g-1}$
+
+Der Einsatz der Pre-Training Strategie ist nur möglich bei einer konstanten Anzahl an Neuronen. Zwei Ansätze verfolgt:
+
+1. DAE-GP mit dynamischer Anpassung, Pre-Trained DAE-GP mit konstanter Anzahl (maximale Länge innerhalb von $\hat{P}$)
+2. Beide Algorithmen mit einer fixen Anzahl von versteckten Neuronen
+
+<!--
+## DAE-GP - Entwicklung DAE-LSTM Dimension
+
+![Airfoil - Entwicklung DAE-LSTM Dimension](~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens/median_hiddenNeurons_perGen.png){height=90%}
+
+--->
+
 # Untersuchung des Suchverhaltens bei symbolischer Regression
@@
 ```
-
-## Herausforderung - DAE-LSTM Dimension
-
-* Im Verlaufe des DAE-GP Algorithmus passt sich die Dimension der DAE-LSTM Netzwerke $M_{g}$ dynamisch an die aktuelle Population an
-
-> Anzahl der versteckten Neuronen pro verstecktem Layer in generation $g$ = Maximum der Länge aller Individuen aus $P_{g-1}$
-
-Der Einsatz der Pre-Training Strategie ist nur möglich bei einer konstanten Anzahl an Neuronen. Zwei Ansätze verfolgt:
-
-1. DAE-GP mit dynamischer Anpassung, Pre-Trained DAE-GP mit konstanter Anzahl (maximale Länge innerhalb von $\hat{P}$)
-2. Beide Algorithmen mit einer fixen Anzahl von versteckten Neuronen
-
-<!--
-## DAE-GP - Entwicklung DAE-LSTM Dimension
-
-![Airfoil - Entwicklung DAE-LSTM Dimension](~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens/median_hiddenNeurons_perGen.png){height=90%}
-
---->
 
@@
 
-![Airfoil - Größe der besten Lösung über Generationen (1HL)](~/masterThesis/master_thesis/img/airfoil_1hl_maxIndSize_fullRun_30gens/mean_Size_byGens.png){height=90%}
+![Airfoil - Größe der besten Lösung über Generationen (1HL)](~/masterThesis/master_thesis/img/airfoil_1hl_maxIndSize_fullRun_30gens/median_Size_byGens.png){height=90%}
 
@@
 ```
+
+## Übersicht Laufzeit - Symbolische Regression
+
+...
 

### 2023-01-19 slides.Rmd
Thu Jan 19 13:00:08 CET 2023: update-remote.sh
Thu Jan 19 15:00:08 CET 2023: update-remote.sh

@@
 author: "Roman Hoehn"
-institute: "Johannes Gutenberg-Universität Mainz"
+institute: "JGU Mainz"
 date: "Datum: 25.01.2023"
@@
   beamer_presentation:
-    #theme: "Madrid"
+    theme: "Madrid"
     fonttheme: "structurebold"
@@
 
-knitr::opts_chunk$set(echo = TRUE)
-knitr::opts_chunk$set(fig.pos = "c")
+knitr::opts_chunk$set(
+  echo = TRUE,
+  fig.pos = "c",
+  fig.align="center"
+  )
 
@@
 
+# create Plots from pstats (cProfile)
+JUPYTER="/Users/rmn/miniconda3/envs/dataScience/bin/jupyter"
+$JUPYTER nbconvert --to html --execute /Users/rmn/masterThesis/eda-gp-2020/cProfiler/eval.ipynb
+
 echo Done recompiling Results, Plots, Diagrams, etc!
@@
 * Hypothese: Problemspezifische Abhängigkeiten zwischen Entscheidungsvariablen können bei der Erzeugung neuer Individuen besser berücksichtigt werden als bei traditionellen Rekombinationsoperatoren^[@edaOrig1996]
-* Weitere Verbreitung im Bereich der genetischen Algorithmen (GA) als für GP
 
@@
 * Adaptierung des DAE-EDA Algorithmus auf GP
-* Darstellung von Individuen als Zeichenketten in prefix Notation
+* Schwierigkeit: Variation von GP Lösungen in Länge und Baumstruktur
 * seq2seq learning Problem: Einsatz von DAE - Long Short Term Memory Netzwerken (DAE-LSTM)
@@
 
-![Airfoil Datensatz - Größe der besten Lösung über Generationen](~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens_withGP/median_best_Size_byGens.png)
+![Airfoil Datensatz - Größe der besten Lösung über Generationen](~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens_withGP/median_best_Size_byGens.png){height=90%}
 
@@
 
-![Airfoil - Laufzeit](/Users/rmn/github/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens_TimeBenchmark/median_training_epochs.png){height=90%}
+![Airfoil - Trainingsepochen](/Users/rmn/github/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens_TimeBenchmark/median_training_epochs.png){height=90%}
 
@@
 
-![Airfoil - Laufzeit](/Users/rmn/github/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens_TimeBenchmark/median_sampleTime.png){height=90%}
+![Airfoil - Sampling-Dauer](/Users/rmn/github/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens_TimeBenchmark/median_sampleTime.png){height=90%}
+
+## Airfoil - Laufzeit (Profiling)
+
+![Airfoil - Profiling](/Users/rmn/github/master_thesis/img/Airfoil_RuntimeProfiling/Runtime_Bar.png){height=90%}
 

### 2023-01-19 slides.Rmd
Thu Jan 19 16:00:10 CET 2023: update-remote.sh

@@
 
-* Im Verlaufe des DAE-GP Algorithmus passt sich die Dimension der DAE-LSTM Netzwerke $M_{g}$ dynamisch an die aktuelle Population an
-
-> Anzahl der versteckten Neuronen pro verstecktem Layer in generation $g$ = Maximum der Länge aller Individuen aus $P_{g-1}$
+* Im Verlaufe des DAE-GP Algorithmus passt sich die Dimension der DAE-LSTM Netzwerke $M_{g}$ dynamisch an die Größe der Individuen der aktuellen Population an (Dimension nimmt i.d.R.
+stark ab innerhalb der ersten Generationen)
 

### 2023-01-19 data/summarize_data.py
autogit

@@ -453,27 +453,29 @@ for metric, filename in FILES.items():
     if metric == "Fitness_lastGen":
         csv_string = "Set,DAE-GP,2nd_Gen-PT,P_Value,Grow-PT,P_Value\n"
 
-        reg_mean_train = np.mean(np.array(list(sublist[-1] for sublist in d["DAE-GP(train)"])))
-        sec_pt_mean_train = np.mean(np.array(list(sublist[-1] for sublist in d["2nd.Gen-PT(train)"])))
-        grow_pt_mean_train = np.mean(np.array(list(sublist[-1] for sublist in d["Grow-PT(train)"])))
+        reg_train = np.array(list(sublist[-1] for sublist in d["DAE-GP(train)"]))
+        sec_pt_train = np.array(list(sublist[-1] for sublist in d["2nd.Gen-PT(train)"]))
+        grow_pt_train = np.array(list(sublist[-1] for sublist in d["Grow-PT(train)"]))
 
-        reg_mean_test = np.mean(np.array(list(sublist[-1] for sublist in d["DAE-GP(test)"])))
-        sec_pt_mean_test = np.mean(np.array(list(sublist[-1] for sublist in d["2nd.Gen-PT(test)"])))
-        grow_pt_mean_test = np.mean(np.array(list(sublist[-1] for sublist in d["Grow-PT(test)"])))
+        reg_test = np.array(list(sublist[-1] for sublist in d["DAE-GP(test)"]))
+        sec_pt_test = np.array(list(sublist[-1] for sublist in d["2nd.Gen-PT(test)"]))
+        grow_pt_test = np.array(list(sublist[-1] for sublist in d["Grow-PT(test)"]))
 
-        csv_string += f"Train,{np.mean(reg_mean_train)},{np.mean(sec_pt_mean_train)},{getPVal(reg_mean_train, sec_pt_mean_train)},{np.mean(grow_pt_mean_train)},{getPVal(reg_mean_train, grow_pt_mean_train)}\n"
-        csv_string += f"Test,{np.mean(reg_mean_test)},{np.mean(sec_pt_mean_test)},{getPVal(reg_mean_test, sec_pt_mean_test)},{np.mean(grow_pt_mean_test)},{getPVal(reg_mean_test, grow_pt_mean_test)}\n"
+
+
[…]

### 2023-01-20 slides.Rmd
Fri Jan 20 10:00:09 CET 2023: update-remote.sh
Fri Jan 20 11:00:06 CET 2023: update-remote.sh

@@
 csl: csl/harvard-cite-them-right.csl
+lang: de
 includes:
@@
 
-Anwendungsgebiet: Symbolische Regression, Fokus auf Aifoil Datensatz
+Anwendungsgebiet: Symbolische Regression, Fokus auf Airfoil Datensatz
 
@@
 
-![Flowchart - Regular DAE-GP](./img/flowcharts/dae-gp.png){height=80%}
+```{r, echo=FALSE, out.height="90%"}
+knitr::include_graphics("./img/flowcharts/dae-gp.png")
+```
 
@@
 
-![Flowchart - Pre-Trained DAE-GP](./img/flowcharts/pt-dae-gp.png){height=70%}
+```{r, echo=FALSE, out.height="90%"}
+knitr::include_graphics("./img/flowcharts/pt-dae-gp.png")
+```
 
@@
 ```
+
 ## Funktions Set Symbolische Regression
@@
 
-![Airfoil Datensatz - Fitness über Generationen](~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens_withGP/median_fitness_byGens.png){height=90%}
+```{r, echo=FALSE, out.height="90%"}
+knitr::include_graphics("~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens_withGP/median_fitness_byGens.png")
+```
+
+<!-- ![Airfoil Datensatz - Fitness über Generationen](){height=90%} -->
 
@@
 
-## Übersicht Laufzeit - Symbolische Regression
-
-...
-
 ## Übersicht Trainingsepochen pro Generation (Median) - Symbolische Regression
@@
 
-# Weitere verfolgte Ansätze
+# Alternative Pre-Training Strategien
 
@@
 
-## Übersicht Fitness (Median) - Weitere Ansätze
+## Übersicht Fitness (Median) - Alternative Ansätze
 
@@
 
-## Auswertung Fitness (Median) - Weitere Ansätze
+## Auswertung Fitness (Median) - Alternative Ansätze
 
@@
 
-## Übersicht Lösungsgröße (Median) - Weitere Ansätze
+## Übersicht Lösungsgröße (Median) - Alternative Ansätze
 
@@
 
-## Auswertung Lösungsgröße (Median) - Weitere Ansätze
+## Auswertung Lösungsgröße (Median) - Alternative Ansätze
 
@@
 
-## Übersicht Populationsdiversität (Median) - Weitere Ansätze
+## Übersicht Populationsdiversität (Median) - Alternative Ansätze
 
@@
 
-## Auswertung Populationsdiversität (Median) - Weitere Ansätze
+## Auswertung Populationsdiversität (Median) - Alternative Ansätze
 
@@
 ```
-## Übersicht Trainingsepochen (Median) - Weitere Ansätze
+## Übersicht Trainingsepochen (Median) - Alternative Ansätze
 
@@
 
-## Auswertung Trainingsepochen (Median) - Weitere Ansätze
+## Auswertung Trainingsepochen (Median) - Alternative Ansätze
 

### 2023-01-20 data/summarize_data.py

@@ -454,10 +454,12 @@ for metric, filename in FILES.items():
         csv_string = "Set,DAE-GP,2nd_Gen-PT,P_Value,Grow-PT,P_Value\n"
 
         reg_train = np.array(list(sublist[-1] for sublist in d["DAE-GP(train)"]))
+        s_pt_train = np.array(list(sublist[-1] for sublist in d["Standard-PT(train)"]))
         sec_pt_train = np.array(list(sublist[-1] for sublist in d["2nd.Gen-PT(train)"]))
         grow_pt_train = np.array(list(sublist[-1] for sublist in d["Grow-PT(train)"]))
 
         reg_test = np.array(list(sublist[-1] for sublist in d["DAE-GP(test)"]))
+        s_pt_train = np.array(list(sublist[-1] for sublist in d["Standard-PT(test)"]))
         sec_pt_test = np.array(list(sublist[-1] for sublist in d["2nd.Gen-PT(test)"]))
         grow_pt_test = np.array(list(sublist[-1] for sublist in d["Grow-PT(test)"]))
 

### 2023-01-20 slides.Rmd
Fri Jan 20 12:00:09 CET 2023: update-remote.sh

@@
 
-<!-- ![Airfoil Datensatz - Fitness über Generationen](){height=90%} -->
+<!-- ![Airfoil Datensatz - Fitness über Generationen](~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens_withGP/median_fitness_byGens.png){height=90%} -->
 
@@
 
-![Airfoil Datensatz - Verteilung der finale Fitness](~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens/final_fit_boxplot.png){height=90%}
+```{r, echo=FALSE, out.height="90%"}
+knitr::include_graphics("~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens/final_fit_boxplot.png")
+```
+
+<!-- ![Airfoil Datensatz - Verteilung der finale Fitness](~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens/final_fit_boxplot.png){height=90%} -->
 
@@
 
-![Airfoil Datensatz - Größe der besten Lösung über Generationen](~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens_withGP/median_best_Size_byGens.png){height=90%}
+```{r, echo=FALSE, out.height="90%"}
+knitr::include_graphics("~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens_withGP/median_best_Size_byGens.png")
+```
+
+<!-- ![Airfoil Datensatz - Größe der besten Lösung über Generationen](~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens_withGP/median_best_Size_byGens.png){height=90%} -->
 
@@
 
-![Airfoil - Populationsdiversität](/Users/rmn/github/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens_withGP/median_popDiversity_byGens_levOnly_withGP.png){height=90%}
+```{r, echo=FALSE, out.height="90%"}
+knitr::include_graphics("/Users/rmn/github/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens_withGP/median_popDiversity_byGens_levOnly_withGP.png")
+```
+
+<!-- ![Airfoil - Populationsdiversität](/Users/rmn/github/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens_withGP/median_popDiversity_byGens_levOnly_withGP.png){height=90%} -->
 
@@
 
-![Airfoil - Laufzeit](/Users/rmn/github/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens_TimeBenchmark/median_runtime.png){height=90%}
+```{r, echo=FALSE, out.height="90%"}
+knitr::include_graphics("/Users/rmn/github/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens_TimeBenchmark/median_runtime.png")
+```
+
+<!-- ![Airfoil - Laufzeit](/Users/rmn/github/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens_TimeBenchmark/median_runtime.png){height=90%} -->
 
@@
 
-![Airfoil - Trainingsepochen](/Users/rmn/github/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens_TimeBenchmark/median_training_epochs.png){height=90%}
+```{r, echo=FALSE, out.height="90%"}
+knitr::include_graphics("/Users/rmn/github/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens_TimeBenchmark/median_training_epochs.png")
+```
+
+<!-- ![Airfoil - Trainingsepochen](/Users/rmn/github/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens_TimeBenchmark/median_training_epochs.png){height=90%} -->
 
@@
 
-![Airfoil - Sampling-Dauer](/Users/rmn/github/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens_TimeBenchmark/median_sampleTime.png){height=90%}
+```{r, echo=FALSE, out.height="90%"}
+knitr::include_graphics("/Users/rmn/github/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens_TimeBenchmark/median_sampleTime.png")
+```
+
+<!-- ![Airfoil - Sampling-Dauer](/Users/rmn/github/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens_TimeBenchmark/median_sampleTime.png){height=90%} -->
 
@@
 
-![Airfoil - Profiling](/Users/rmn/github/master_thesis/img/Airfoil_RuntimeProfiling/Runtime_Bar.png){height=90%}
+```{r, echo=FALSE, out.height="90%"}
+knitr::include_graphics("/Users/rmn/github/master_thesis/img/Airfoil_RuntimeProfiling/Runtime_Bar.png")
+```
+
+<!-- ![Airfoil - Profiling](/Users/rmn/github/master_thesis/img/Airfoil_RuntimeProfiling/Runtime_Bar.png){height=90%} -->
 
@@
 
-![Airfoil - Entwicklung der Fitness über Generationen (1 HL)](~/masterThesis/master_thesis/img/airfoil_1hl_maxIndSize_fullRun_30gens/median_fitness_byGens.png){height=85%}
+```{r, echo=FALSE, out.height="90%"}
+knitr::include_graphics("~/masterThesis/master_thesis/img/airfoil_1hl_maxIndSize_fullRun_30gens/median_fitness_byGens.png")
+```
+
+<!-- ![Airfoil - Entwicklung der Fitness über Generationen (1 HL)](~/masterThesis/master_thesis/img/airfoil_1hl_maxIndSize_fullRun_30gens/median_fitness_byGens.png){height=85%} -->
 
@@
 
-![Airfoil - Verteilung der finalen Fitness (1HL)](~/masterThesis/master_thesis/img/airfoil_1hl_maxIndSize_fullRun_30gens/final_fit_boxplot.png){height=75%}
+```{r, echo=FALSE, out.height="90%"}
+knitr::include_graphics("~/masterThesis/master_thesis/img/airfoil_1hl_maxIndSize_fullRun_30gens/final_fit_boxplot.png")
+```
+
+<!-- ![Airfoil - Verteilung der finalen Fitness (1HL)](~/masterThesis/master_thesis/img/airfoil_1hl_maxIndSize_fullRun_30gens/final_fit_boxplot.png){height=75%} -->
 
@@
 
-![Airfoil - Größe der besten Lösung über Generationen (1HL)](~/masterThesis/master_thesis/img/airfoil_1hl_maxIndSize_fullRun_30gens/median_Size_byGens.png){height=90%}
+```{r, echo=FALSE, out.height="90%"}
+knitr::include_graphics("~/masterThesis/master_thesis/img/airfoil_1hl_maxIndSize_fullRun_30gens/median_Size_byGens.png")
+```
+
+<!-- ![Airfoil - Größe der besten Lösung über Generationen (1HL)](~/masterThesis/master_thesis/img/airfoil_1hl_maxIndSize_fullRun_30gens/median_Size_byGens.png){height=90%} -->
 
@@
 
-![Übersicht Median Fitness - Symbolische Regression](~/masterThesis/master_thesis/img/symbolicRegressionSummarized/median_fitness_byGens.png){height=80%}
+```{r, echo=FALSE, out.height="90%"}
+knitr::include_graphics("~/masterThesis/master_thesis/img/symbolicRegressionSummarized/median_fitness_byGens.png")
+```
+<!-- ![Übersicht Median Fitness - Symbolische Regression](~/masterThesis/master_thesis/img/symbolicRegressionSummarized/median_fitness_byGens.png){height=80%} -->
 
@@
 
-![Übersicht Lösungsgröße (Median) - Symbolische Regression](~/masterThesis/master_thesis/img/symbolicRegressionSummarized/median_bestSol_size_byGens.png){height=80%}
+```{r, echo=FALSE, out.height="90%"}
+knitr::include_graphics("~/masterThesis/master_thesis/img/symbolicRegressionSummarized/median_bestSol_size_byGens.png")
+```
+
+<!-- ![Übersicht Lösungsgröße (Median) - Symbolische Regression](~/masterThesis/master_thesis/img/symbolicRegressionSummarized/median_bestSol_size_byGens.png){height=80%} -->
 
@@
 
-![Übersicht Populationsdiversität (Median) - Symbolische Regression](~/masterThesis/master_thesis/img/symbolicRegressionSummarized/median_levDistance_byGen.png){height=80%}
+```{r, echo=FALSE, out.height="90%"}
+knitr::include_graphics("~/masterThesis/master_thesis/img/symbolicRegressionSummarized/median_levDistance_byGen.png")
+```
+
+<!-- ![Übersicht Populationsdiversität (Median) - Symbolische Regression](~/masterThesis/master_thesis/img/symbolicRegressionSummarized/median_levDistance_byGen.png){height=80%} -->
 
@@
 
-![Übersicht Trainingsepochen pro Generation (Median) - Symbolische Regression](~/masterThesis/master_thesis/img/symbolicRegressionSummarized/median_epochsTrained.png){height=80%}
+```{r, echo=FALSE, out.height="90%"}
+knitr::include_graphics("~/masterThesis/master_thesis/img/symbolicRegressionSummarized/median_epochsTrained.png")
+```
+
+<!-- ![Übersicht Trainingsepochen pro Generation (Median) - Symbolische Regression](~/masterThesis/master_thesis/img/symbolicRegressionSummarized/median_epochsTrained.png){height=80%} -->
 
@@
 
-![Übersicht Fitness pro Generation (Median) - Weitere Ansaetze](~/masterThesis/master_thesis/img/PT_Strategies_airfoil_2hl_150hn_fullRun/median_fitness_byGens.png){height=80%}
+```{r, echo=FALSE, out.height="90%"}
+knitr::include_graphics("~/masterThesis/master_thesis/img/PT_Strategies_airfoil_2hl_150hn_fullRun/median_fitness_byGens.png")
+```
+
+<!-- ![Übersicht Fitness pro Generation (Median) - Weitere Ansaetze](~/masterThesis/master_thesis/img/PT_Strategies_airfoil_2hl_150hn_fullRun/median_fitness_byGens.png){height=80%} -->
 
@@
 
-![Übersicht Lösungsgröße der besten Individuen (Median) - Weitere Ansaetze](~/masterThesis/master_thesis/img/PT_Strategies_airfoil_2hl_150hn_fullRun/median_Size_best_solution.png){height=80%}
+```{r, echo=FALSE, out.height="90%"}
+knitr::include_graphics("~/masterThesis/master_thesis/img/PT_Strategies_airfoil_2hl_150hn_fullRun/median_Size_best_solution.png")
+```
+
+<!-- ![Übersicht Lösungsgröße der besten Individuen (Median) - Weitere Ansaetze](~/masterThesis/master_thesis/img/PT_Strategies_airfoil_2hl_150hn_fullRun/median_Size_best_solution.png){height=80%} -->
 
@@
 
-![Übersicht Populationsdiversität (Median) - Weitere Ansaetze](~/masterThesis/master_thesis/img/PT_Strategies_airfoil_2hl_150hn_fullRun/mean_popDiversity_byGens_levOnly.png){height=80%}
+```{r, echo=FALSE, out.height="90%"}
+knitr::include_graphics("~/masterThesis/master_thesis/img/PT_Strategies_airfoil_2hl_150hn_fullRun/median_lev_div_overGens.png")
+```
+
+<!-- ![Übersicht Populationsdiversität (Median) - Weitere Ansaetze](~/masterThesis/master_thesis/img/PT_Strategies_airfoil_2hl_150hn_fullRun/mean_popDiversity_byGens_levOnly.png){height=80%} -->
 
@@
 
-![Übersicht Trainingsepochen (Median) - Weitere Ansaetze](~/masterThesis/master_thesis/img/PT_Strategies_airfoil_2hl_150hn_fullRun/mean_epochs_trained_perGen.png){height=80%}
+```{r, echo=FALSE, out.height="90%"}
+knitr::include_graphics("~/masterThesis/master_thesis/img/PT_Strategies_airfoil_2hl_150hn_fullRun/mean_epochs_trained_perGen.png")
+```
+
+<!-- ![Übersicht Trainingsepochen (Median) - Weitere Ansaetze](~/masterThesis/master_thesis/img/PT_Strategies_airfoil_2hl_150hn_fullRun/mean_epochs_trained_perGen.png){height=80%} -->
 

### 2023-01-20 slides.Rmd
Fri Jan 20 13:00:09 CET 2023: update-remote.sh

@@
 ---
-title: "Pre-Trained Denoising Autoencoders Long Short-Term Memory Networks as probabilistic Models for Estimation of Distribution Genetic Programming"
-subtitle: "Kolloquium zur Masterarbeit im M.Sc. Wirtschaftspädagogik "
+title: "Pre-Trained Denoising Autoencoders Long Short-Term Memory Networks as probabilistic
+  Models for Estimation of Distribution Genetic Programming"
 author: "Roman Hoehn"
-institute: "JGU Mainz"
-date: "Datum: 25.01.2023"
-#fontsize: 11pt
-link-citations: true
+date: 'Datum: 25.01.2023'
 output:
   beamer_presentation:
-    theme: "Madrid"
-    fonttheme: "structurebold"
-    fig_width: 12
-    fig_height: 10
-    # keep_tex: true
+    theme: Madrid
+    fonttheme: structurebold
     fig_caption: no
-    number_sections: true
+    number_sections: yes
     slide_level: 2
+subtitle: 'Kolloquium zur Masterarbeit im M.Sc. Wirtschaftspädagogik '
+link-citations: yes
+institute: JGU Mainz
 toc: yes
 bibliography: ref/ref.bib
-csl: csl/harvard-cite-them-right.csl
+csl: "csl/harvard-cite-them-right.csl"
 lang: de
-includes:
-      in_header: [header_files/slides.tex]
-
+# includes:
+#   in_header: header_files/slides.tex
 ---
@@
 
-```{r, echo=FALSE, out.height="90%"}
+```{r echo=FALSE, out.height="90%", out.width="50%"}
 knitr::include_graphics("./img/flowcharts/dae-gp.png")
@@
 
-```{r, echo=FALSE, out.height="90%"}
+```{r, echo=FALSE, out.height="90%", out.width="50%"}
 knitr::include_graphics("./img/flowcharts/pt-dae-gp.png")
@@
 
-```{r, echo=FALSE, out.height="90%"}
+```{r, echo=FALSE, out.height="85%", out.width="80%"}
 knitr::include_graphics("~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens_withGP/median_fitness_byGens.png")
@@
 
-```{r, echo=FALSE, out.height="90%"}
+```{r, echo=FALSE, out.height="85%", out.width="80%"}
 knitr::include_graphics("~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens/final_fit_boxplot.png")
@@
 
-```{r, echo=FALSE, out.height="90%"}
+```{r, echo=FALSE, out.height="85%", out.width="80%"}
 knitr::include_graphics("~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens_withGP/median_best_Size_byGens.png")
@@
 
-```{r, echo=FALSE, out.height="90%"}
+```{r, echo=FALSE, out.height="85%", out.width="80%"}
 knitr::include_graphics("/Users/rmn/github/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens_withGP/median_popDiversity_byGens_levOnly_withGP.png")
@@
 
-```{r, echo=FALSE, out.height="90%"}
+```{r, echo=FALSE, out.height="85%", out.width="80%"}
 knitr::include_graphics("/Users/rmn/github/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens_TimeBenchmark/median_runtime.png")
@@
 
-```{r, echo=FALSE, out.height="90%"}
+```{r, echo=FALSE, out.height="85%", out.width="80%"}
 knitr::include_graphics("/Users/rmn/github/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens_TimeBenchmark/median_training_epochs.png")
@@
 
-```{r, echo=FALSE, out.height="90%"}
+```{r, echo=FALSE, out.height="85%", out.width="80%"}
 knitr::include_graphics("/Users/rmn/github/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens_TimeBenchmark/median_sampleTime.png")
@@
 
-```{r, echo=FALSE, out.height="90%"}
+```{r, echo=FALSE, out.height="85%", out.width="80%"}
 knitr::include_graphics("/Users/rmn/github/master_thesis/img/Airfoil_RuntimeProfiling/Runtime_Bar.png")
@@
 
-```{r, echo=FALSE, out.height="90%"}
+```{r, echo=FALSE, out.height="85%", out.width="80%"}
 knitr::include_graphics("~/masterThesis/master_thesis/img/airfoil_1hl_maxIndSize_fullRun_30gens/median_fitness_byGens.png")
@@
 
-```{r, echo=FALSE, out.height="90%"}
+```{r, echo=FALSE, out.height="85%", out.width="80%"}
 knitr::include_graphics("~/masterThesis/master_thesis/img/airfoil_1hl_maxIndSize_fullRun_30gens/final_fit_boxplot.png")
@@
 
-```{r, echo=FALSE, out.height="90%"}
+```{r, echo=FALSE, out.height="85%", out.width="80%"}
 knitr::include_graphics("~/masterThesis/master_thesis/img/airfoil_1hl_maxIndSize_fullRun_30gens/median_Size_byGens.png")
@@
 
-```{r, echo=FALSE, out.height="90%"}
+```{r, echo=FALSE, out.height="85%", out.width="80%"}
 knitr::include_graphics("~/masterThesis/master_thesis/img/symbolicRegressionSummarized/median_fitness_byGens.png")
@@
 
-```{r full_run_realWorldSymReg_fitness, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
+```{r full_run_realWorldSymReg_fitness, echo=FALSE, message=FALSE, warning=FALSE}
 library(knitr)
@@
 
-```{r, echo=FALSE, out.height="90%"}
+```{r, echo=FALSE, out.height="85%", out.width="80%"}
 knitr::include_graphics("~/masterThesis/master_thesis/img/symbolicRegressionSummarized/median_bestSol_size_byGens.png")
@@
 
-```{r, echo=FALSE, out.height="90%"}
+```{r, echo=FALSE, out.height="85%", out.width="80%"}
 knitr::include_graphics("~/masterThesis/master_thesis/img/symbolicRegressionSummarized/median_levDistance_byGen.png")
@@
 
-```{r, echo=FALSE, out.height="90%"}
+```{r, echo=FALSE, out.height="85%", out.width="80%"}
 knitr::include_graphics("~/masterThesis/master_thesis/img/symbolicRegressionSummarized/median_epochsTrained.png")
@@
 
-```{r, echo=FALSE, out.height="90%"}
+```{r, echo=FALSE, out.height="85%", out.width="80%"}
 knitr::include_graphics("~/masterThesis/master_thesis/img/PT_Strategies_airfoil_2hl_150hn_fullRun/median_fitness_byGens.png")
@@
 
-```{r, echo=FALSE, out.height="90%"}
+```{r, echo=FALSE, out.height="85%", out.width="80%"}
 knitr::include_graphics("~/masterThesis/master_thesis/img/PT_Strategies_airfoil_2hl_150hn_fullRun/median_Size_best_solution.png")
@@
 
-```{r, echo=FALSE, out.height="90%"}
+```{r, echo=FALSE, out.height="85%", out.width="80%"}
 knitr::include_graphics("~/masterThesis/master_thesis/img/PT_Strategies_airfoil_2hl_150hn_fullRun/median_lev_div_overGens.png")
@@
 
-```{r, echo=FALSE, out.height="90%"}
+```{r, echo=FALSE, out.height="85%", out.width="80%"}
 knitr::include_graphics("~/masterThesis/master_thesis/img/PT_Strategies_airfoil_2hl_150hn_fullRun/mean_epochs_trained_perGen.png")

### 2023-01-20 slides.Rmd
Fri Jan 20 14:00:08 CET 2023: update-remote.sh

@@
 
-```{bash recompile_results, echo=TRUE, eval=FALSE, include=FALSE}
+```{bash recompile_results, echo=TRUE, eval=TRUE, include=FALSE}
 
@@
 
-  1. Model Building: Modell "lernt" die Eigenschaften von ausgewählten Lösungen hoher Güte durch das Trainieren eines DAEs
+  1. Model Building: Modell "lernt" die Eigenschaften von ausgewählten Lösungen hoher Güte durch das Trainieren eines DAE
   2. Model Sampling: Neue Lösungen werden erzeugt durch das propagieren von bestehenden, mutierten Lösungen durch das erlernte Modell
@@
 
-Aufbau: Betrachtung des Suchverhalten über je $10$ Gesamtduchläufe für DAE-GP und pre-trained DAE-GP
+Aufbau: Betrachtung des Suchverhalten über je $10$ Gesamtduchläufe am Airfoil Datensatz^[@machine_learning_repo] für DAE-GP und pre-trained DAE-GP
 
@@
   read.csv("./tables/airfoil_fullRun_2hl_maxIndSize_params"),
-  # caption="Hyperparameter"
-)%>% row_spec(0,bold=TRUE ) %>% kable_styling(latex_options = "scale_down", font_size = 7)
-```
-
-## Funktions Set Symbolische Regression
-
-```{r airfoil_function_set, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
-fig.pos="H"
-knitr::kable(
-  read.csv("./tables/airfoil_function_set", sep = ","),
-) %>% kableExtra::kable_styling(latex_options = "hold_position")
+)%>% row_spec(0,bold=TRUE ) %>% kable_styling(latex_options = "scale_down", font_size = 6)
 ```

### 2023-01-20 slides.Rmd
Fri Jan 20 15:00:07 CET 2023: update-remote.sh

@@
 
-```{bash recompile_results, echo=TRUE, eval=TRUE, include=FALSE}
+```{bash recompile_results, echo=TRUE, eval=FALSE, include=FALSE}
 
@@
   read.csv("./tables/airfoil_fullRun_2hl_maxIndSize_params"),
-)%>% row_spec(0,bold=TRUE ) %>% kable_styling(latex_options = "scale_down", font_size = 6)
+)%>% kable_styling(latex_options = "scale_down", font_size = 6)
 ```
@@
 
-## Übersicht Median Fitness - Symbolische Regression
+## Median Fitness - Symbolische Regression
 
@@
 
-## Auswertung Fitness (Median) - Symbolische Regression
+## Fitness (Median) - Symbolische Regression
 
@@
 
-## Übersicht Lösungsgröße (Median) - Symbolische Regression
+## Lösungsgröße (Median) - Symbolische Regression
 
@@
 
-## Übersicht Populationsdiversität (Median) - Symbolische Regression
+## Populationsdiversität (Median) - Symbolische Regression
 
-```{r, echo=FALSE, out.height="85%", out.width="80%"}
+```{r, echo=FALSE, out.height="80%", out.width="75%"}
 knitr::include_graphics("~/masterThesis/master_thesis/img/symbolicRegressionSummarized/median_levDistance_byGen.png")
@@
 
-## Übersicht Trainingsepochen pro Generation (Median) - Symbolische Regression
+## Trainingsepochen pro Generation (Median) - Symbolische Regression
 
-```{r, echo=FALSE, out.height="85%", out.width="80%"}
+```{r, echo=FALSE, out.height="80%", out.width="75%"}
 knitr::include_graphics("~/masterThesis/master_thesis/img/symbolicRegressionSummarized/median_epochsTrained.png")
@@
 
-## Übersicht Fitness (Median) - Alternative Ansätze
+## Fitness (Median) - Alternative Ansätze
 
@@
 
-## Übersicht Lösungsgröße (Median) - Alternative Ansätze
+## Lösungsgröße (Median) - Alternative Ansätze
 
@@
 
-## Übersicht Populationsdiversität (Median) - Alternative Ansätze
+## Populationsdiversität (Median) - Alternative Ansätze
 
@@
 ```
-## Übersicht Trainingsepochen (Median) - Alternative Ansätze
+## Trainingsepochen (Median) - Alternative Ansätze
 
@@
 
+* Erhöhung der Rechenzeit/Ressourcen: DAE-LSTMs müssen ausreichend dimensioniert sein (2+ Hidden Layer), aktuelle Publikationen nutzen nur einen einzelnen Hidden Layer^[@dae-gp_2022_symreg] ^[@daegp_explore_exploit].
+
 * Keine signifikanten Vorteile für die Qualität der gefundenen Lösungen

### 2023-01-20 data/summarize_data.py
autogit

@@ -451,7 +451,7 @@ for metric, filename in FILES.items():
     d = json.load(open(join(PT_STRATEGIES_PATH, filename),"r",encoding="utf-8"))
 
     if metric == "Fitness_lastGen":
-        csv_string = "Set,DAE-GP,2nd_Gen-PT,P_Value,Grow-PT,P_Value\n"
+        csv_string = "Set,Standard-PT,2nd_Gen-PT,P_Value,Grow-PT,P_Value\n"
 
         reg_train = np.array(list(sublist[-1] for sublist in d["DAE-GP(train)"]))
         s_pt_train = np.array(list(sublist[-1] for sublist in d["Standard-PT(train)"]))
@@ -459,34 +459,36 @@ for metric, filename in FILES.items():
         grow_pt_train = np.array(list(sublist[-1] for sublist in d["Grow-PT(train)"]))
 
         reg_test = np.array(list(sublist[-1] for sublist in d["DAE-GP(test)"]))
-        s_pt_train = np.array(list(sublist[-1] for sublist in d["Standard-PT(test)"]))
+        s_pt_test = np.array(list(sublist[-1] for sublist in d["Standard-PT(test)"]))
         sec_pt_test = np.array(list(sublist[-1] for sublist in d["2nd.Gen-PT(test)"]))
         grow_pt_test = np.array(list(sublist[-1] for sublist in d["Grow-PT(test)"]))
 
 
 
-        csv_string += f"Train,{np.mean(reg_train)},{np.mean(sec_pt_train)},{getPVal(reg_train, sec_pt_train)},{np.mean(grow_pt_train)},{getPVal(reg_train, grow_pt_train)}\n"
-        csv_string += f"Test,{np.mean(reg_test)},{np.mean(sec_pt_test)},{getPVal(reg_test, sec_pt_test)},{np.mean(grow_pt_test)},{getPVal(reg_test, grow_pt_test)}\n"
+        csv_string += f"Train,{np.mean(s_pt_train)},{np.mean(sec_pt_train)},{getPVal(s_pt_train, sec_pt_train)},{np.mean(grow_pt_train)},{getPVal(s_pt_train, grow_pt_train)}\n"
[…]

### 2023-01-20 slides.Rmd
autogit

@@
   beamer_presentation:
-    theme: Madrid
+    theme: "Madrid"
+    # colortheme: "beaver"
     fonttheme: structurebold
@@
 
-## Estimation of Distribution Algorithmen (EDA)
+## Estimation of Distribution Algorithmen^[@edaOrig1996] ^[@design_of_modern_heuristics] (EDA)
 
-* Entwicklung neuer Rekombinationsoperatoren für evolutionäre Algorithmen basierend auf dem Einsatz von probabilistischen Modellen^[@design_of_modern_heuristics]
-* Hypothese: Problemspezifische Abhängigkeiten zwischen Entscheidungsvariablen können bei der Erzeugung neuer Individuen besser berücksichtigt werden als bei traditionellen Rekombinationsoperatoren^[@edaOrig1996]
+* Entwicklung neuer Rekombinationsoperatoren für evolutionäre Algorithmen basierend auf dem Einsatz von probabilistischen Modellen
+* Hypothese: Problemspezifische Abhängigkeiten zwischen Entscheidungsvariablen können bei der Erzeugung neuer Individuen besser berücksichtigt werden als bei traditionellen Rekombinationsoperatoren
 
@@
 
-col_names <- c("Set", "DAE-GP", "2ndGen", "P-Val(1)" , "GrowInit", "P-Val(2)")
+col_names <- c("Set", "Standard-PT", "2ndGen", "P-Val(1)" , "GrowInit", "P-Val(2)")
 
@@
 
-col_names <- c("DAE-GP", "2ndGen", "P-Val(1)" , "GrowInit", "P-Val(2)")
+col_names <- c("Standard-PT", "2ndGen", "P-Val(1)" , "GrowInit", "P-Val(2)")
 

### 2023-01-20 data/summarize_data.py

@@ -511,7 +511,7 @@ for metric, filename in FILES.items():
         reg_test = np.array(list(sublist[-1] for sublist in d["DAE-GP(test)"]))
         s_pt_test = np.array(list(sublist[-1] for sublist in d["Standard-PT(test)"]))
         sec_pt_test = np.array(list(sublist[-1] for sublist in d["2nd.Gen-PT(test)"]))
-        grow_pt_test = np.array(list(sublist[-1] for sublist in d["Grow-PT(test)"])
+        grow_pt_test = np.array(list(sublist[-1] for sublist in d["Grow-PT(test)"]))
 
 
 

### 2023-01-22 slides.Rmd
autogit
Sun Jan 22 14:09:03 CET 2023: update-remote.sh

@@
 
-Ansatz: Initialisierung von $\hat{P}$ ausschließlich mit der Grow Methode anstelle von ramped Half and Half
+Ansatz: Initialisierung von $\hat{P}$ ausschließlich mit der Grow Methode^[@Koza1993GeneticP] anstelle von ramped Half and Half
 

### 2023-01-23 slides.Rmd
Mon Jan 23 09:00:06 CET 2023: update-remote.sh

@@
 
-Anwendungsgebiet: Symbolische Regression, Fokus auf Airfoil Datensatz
+Anwendungsgebiet: Symbolische Regression, Fokus auf Airfoil Datensatz^[@machine_learning_repo]
 
@@
 
-Aufbau: Betrachtung des Suchverhalten über je $10$ Gesamtduchläufe am Airfoil Datensatz^[@machine_learning_repo] für DAE-GP und pre-trained DAE-GP
+Aufbau: Betrachtung des Suchverhalten über je $10$ Gesamtduchläufe am Airfoil Datensatz für DAE-GP und pre-trained DAE-GP
 
@@
   read.csv("./tables/airfoil_fullRun_2hl_maxIndSize_params"),
-)%>% kable_styling(latex_options = "scale_down", font_size = 6)
+)%>% row_spec(0,bold=TRUE )%>% kable_styling(latex_options = "scale_down", font_size = 6)
 ```
@@
 
-## Airfoil - Laufzeit (Profiling)
+<!-- ## Airfoil - Laufzeit (Profiling) -->
 
-```{r, echo=FALSE, out.height="85%", out.width="80%"}
-knitr::include_graphics("/Users/rmn/github/master_thesis/img/Airfoil_RuntimeProfiling/Runtime_Bar.png")
-```
+<!-- ```{r, echo=FALSE, out.height="85%", out.width="80%"} -->
+<!-- knitr::include_graphics("/Users/rmn/github/master_thesis/img/Airfoil_RuntimeProfiling/Runtime_Bar.png") -->
+<!-- ``` -->
 

### 2023-01-23 experiments/2ndGenPT_airfoil_2hl_150hn_fullRun/re_evaluate_2ndGenPT_airfoil_2hl_150hn_fullRun.sh
changing patths

@@ -1,0 +1,25 @@
+#/usr/bin/env bash
+
+# Re-Evaluate all results from airfoilFirstGen experiments
+
+bash config
+set -o errexit
+set -o nounset
+# set -o pipefail
+if [[ "${TRACE-0}" == "1" ]]; then
+    set -o xtrace
+fi
+
+DIR="/Users/rmn/masterThesis/eda-gp-2020/experiments/2ndGenPT_airfoil_2hl_150hn_fullRun"
+PROG="/Users/rmn/masterThesis/eda-gp-2020/src/testing_evaluate.py"
+ENV="/Users/rmn/miniconda3/envs/edaGP/bin/python"
+JUPYTER="/Users/rmn/miniconda3/envs/dataScience/bin/jupyter"
+
+NRUNS=10
+MAXGEN=30
+
+
+yes yes | $ENV $PROG -f $DIR -n $NRUNS -o "$DIR/results.csv"  -g $MAXGEN
+
+
+$JUPYTER nbconvert --to html --execute "/Users/rmn/masterThesis/eda-gp-2020/experiments/2ndGenPT_airfoil_2hl_150hn_fullRun/2ndGenPT_airfoil_2hl_150hn_fullRun.ipynb"

### 2023-01-23 experiments/3ndGenPT_airfoil_2hl_100hn_fullRun/re_evaluate_3ndGenPT_airfoil_2hl_100hn_fullRun.sh

@@ -1,0 +1,25 @@
+#/usr/bin/env bash
+
+# Re-Evaluate all results from airfoilFirstGen experiments
+
+bash config
+set -o errexit
+set -o nounset
+# set -o pipefail
+if [[ "${TRACE-0}" == "1" ]]; then
+    set -o xtrace
+fi
+
+DIR="/Users/rmn/masterThesis/eda-gp-2020/experiments/3ndGenPT_airfoil_2hl_100hn_fullRun"
+PROG="/Users/rmn/masterThesis/eda-gp-2020/src/testing_evaluate.py"
+ENV="/Users/rmn/miniconda3/envs/edaGP/bin/python"
+JUPYTER="/Users/rmn/miniconda3/envs/dataScience/bin/jupyter"
+
+NRUNS=10
+MAXGEN=20
+
+
+yes yes | $ENV $PROG -f $DIR -n $NRUNS -o "$DIR/results.csv"  -g $MAXGEN
+
+
+#$JUPYTER nbconvert --to html --execute "/Users/rmn/masterThesis/eda-gp-2020/experiments/airfoil_2hl_maxIndSize_fullRun_30gens/airfoil_2hl_maxIndSize_fullRun_30gens_with2ndGen.ipynb"

### 2023-01-23 experiments/airfoil_1hl_maxIndSize_fullRun_30gens/re_evaluate_airfoil_1hl_maxIndSize_30gen.sh

@@ -1,0 +1,24 @@
+#/usr/bin/env bash
+
+# Re-Evaluate all results from airfoilFirstGen experiments
+
+bash config
+set -o errexit
+set -o nounset
+# set -o pipefail
+if [[ "${TRACE-0}" == "1" ]]; then
+    set -o xtrace
+fi
+
+DIR="/Users/rmn/masterThesis/eda-gp-2020/experiments/airfoil_1hl_maxIndSize_fullRun_30gens"
+PROG="/Users/rmn/masterThesis/eda-gp-2020/src/testing_evaluate.py"
+ENV="/Users/rmn/miniconda3/envs/edaGP/bin/python"
+JUPYTER="/Users/rmn/miniconda3/envs/dataScience/bin/jupyter"
+
+NRUNS=10
+MAXGEN=30
+
+
+yes yes | $ENV $PROG -f $DIR -n $NRUNS -o "$DIR/results.csv"  -g $MAXGEN
+
+$JUPYTER nbconvert --to html --execute "/Users/rmn/masterThesis/eda-gp-2020/experiments/airfoil_1hl_maxIndSize_fullRun_30gens/airfoil_1hl_maxIndSize_fullRun_30gens.ipynb"

### 2023-01-23 experiments/airfoil_2hl_maxIndSize_fullRun_30gens/re_evaluate_airfoil_2hl_maxIndSize_30gen.sh

@@ -1,0 +1,28 @@
+#/usr/bin/env bash
+
+# Re-Evaluate all results from airfoilFirstGen experiments
+
+bash config
+set -o errexit
+set -o nounset
+# set -o pipefail
+if [[ "${TRACE-0}" == "1" ]]; then
+    set -o xtrace
+fi
+
+DIR="/Users/rmn/masterThesis/eda-gp-2020/experiments/airfoil_2hl_maxIndSize_fullRun_30gens"
+PROG="/Users/rmn/masterThesis/eda-gp-2020/src/testing_evaluate.py"
+ENV="/Users/rmn/miniconda3/envs/edaGP/bin/python"
+JUPYTER="/Users/rmn/miniconda3/envs/dataScience/bin/jupyter"
+
+NRUNS=10
+MAXGEN=30
+
+
+yes yes | $ENV $PROG -f $DIR -n $NRUNS -o "$DIR/results.csv"  -g $MAXGEN
+
+$JUPYTER nbconvert --to html --execute "/Users/rmn/masterThesis/eda-gp-2020/experiments/airfoil_2hl_maxIndSize_fullRun_30gens/airfoil_2hl_maxIndSize_fullRun_30gens.ipynb"
+
+$JUPYTER nbconvert --to html --execute "/Users/rmn/masterThesis/eda-gp-2020/experiments/airfoil_2hl_maxIndSize_fullRun_30gens/airfoil_2hl_maxIndSize_fullRun_30gens_withGP.ipynb"
+
+$JUPYTER nbconvert --to html --execute "/Users/rmn/masterThesis/eda-gp-2020/experiments/airfoil_2hl_maxIndSize_fullRun_30gens/airfoil_2hl_maxIndSize_fullRun_30gens_TimeBenchmark.ipynb"

### 2023-01-23 experiments/airfoil_firstGen/re_evaluate_airfoilFirstGen.sh

@@ -1,0 +1,40 @@
+#/usr/bin/env bash
+
+# Re-Evaluate all results from airfoilFirstGen experiments
+
+bash config
+set -o errexit
+set -o nounset
+# set -o pipefail
+if [[ "${TRACE-0}" == "1" ]]; then
+    set -o xtrace
+fi
+
+DIR="/Users/rmn/masterThesis/eda-gp-2020/experiments/airfoil_firstGen"
+PROG="/Users/rmn/masterThesis/eda-gp-2020/src/testing_evaluate.py"
+ENV="/Users/rmn/miniconda3/envs/edaGP/bin/python"
+
+NRUNS=10
+MAXGEN=1
+
+COUNTER=0
+
+# Find all subdirectories of DIR
+for SUBDIR in $(find "$DIR" -type d -maxdepth 1); do
+  if [ "$SUBDIR" = "$DIR" ]; then
+      continue
+    fi
+  if [[ "$SUBDIR" == *".ipynb_checkpoints"* ]]; then
+    continue
+  fi
+  yes yes | $ENV $PROG -f $SUBDIR -n $NRUNS -o "$SUBDIR/results.csv"  -g $MAXGEN
+  ((COUNTER++))
+
+done
+
+echo "Done - Evaluated $COUNTER Experiments!"
+
+JUPYTER="/Users/rmn/miniconda3/envs/dataScience/bin/jupyter"
+
+$JUPYTER nbconvert --to html --execute "/Users/rmn/masterThesis/eda-gp-2020/experiments/airfoil_firstGen/airfoil_1gen_combined_layers.ipynb"
+$JUPYTER nbconvert --to html --execute "/Users/rmn/masterThesis/eda-gp-2020/experiments/airfoil_firstGen/airfoil_1gen_combined_neurons.ipynb"

### 2023-01-23 experiments/bostonHousing_2hl_maxIndSize_fullRun_30gens/re_evaluate_bostonHousing_2hl_maxIndSize_fullRun_30gens.sh

@@ -1,0 +1,24 @@
+#/usr/bin/env bash
+
+# Re-Evaluate all results from bostonHousing_2hl_maxIndSize_fullRun_30gens
+
+bash config
+set -o errexit
+set -o nounset
+# set -o pipefail
+if [[ "${TRACE-0}" == "1" ]]; then
+    set -o xtrace
+fi
+
+DIR="/Users/rmn/masterThesis/eda-gp-2020/experiments/bostonHousing_2hl_maxIndSize_fullRun_30gens"
+PROG="/Users/rmn/masterThesis/eda-gp-2020/src/testing_evaluate.py"
+ENV="/Users/rmn/miniconda3/envs/edaGP/bin/python"
+JUPYTER="/Users/rmn/miniconda3/envs/dataScience/bin/jupyter"
+
+NRUNS=10
+MAXGEN=30
+
+
+yes yes | $ENV $PROG -f $DIR -n $NRUNS -o "$DIR/results.csv"  -g $MAXGEN
+
+$JUPYTER nbconvert --to html --execute "/Users/rmn/masterThesis/eda-gp-2020/experiments/bostonHousing_2hl_maxIndSize_fullRun_30gens/bostonHousing_2hl_maxIndSize_fullRun_30gens.ipynb"

### 2023-01-23 experiments/concrete_2hl_maxIndSize_fullRun_30gens/re_evaluate_concrete_2hl_maxIndSize_fullRun_30gens.sh

@@ -1,0 +1,24 @@
+#/usr/bin/env bash
+
+# Re-Evaluate all results from energyCooling_2hl_maxIndSize_fullRun_30gens
+
+bash config
+set -o errexit
+set -o nounset
+# set -o pipefail
+if [[ "${TRACE-0}" == "1" ]]; then
+    set -o xtrace
+fi
+
+DIR="/Users/rmn/masterThesis/eda-gp-2020/experiments/concrete_2hl_maxIndSize_fullRun_30gens"
+PROG="/Users/rmn/masterThesis/eda-gp-2020/src/testing_evaluate.py"
+ENV="/Users/rmn/miniconda3/envs/edaGP/bin/python"
+JUPYTER="/Users/rmn/miniconda3/envs/dataScience/bin/jupyter"
+
+NRUNS=10
+MAXGEN=30
+
+
+yes yes | $ENV $PROG -f $DIR -n $NRUNS -o "$DIR/results.csv"  -g $MAXGEN
+
+$JUPYTER nbconvert --to html --execute "/Users/rmn/masterThesis/eda-gp-2020/experiments/concrete_2hl_maxIndSize_fullRun_30gens/concrete_2hl_maxIndSize_fullRun_30gens.ipynb"

### 2023-01-23 experiments/energyCooling_2hl_maxIndSize_fullRun_30gens/re_evaluate_energyCooling_2hl_maxIndSize_fullRun_30gens.sh

@@ -1,0 +1,24 @@
+#/usr/bin/env bash
+
+# Re-Evaluate all results from energyCooling_2hl_maxIndSize_fullRun_30gens
+
+bash config
+set -o errexit
+set -o nounset
+# set -o pipefail
+if [[ "${TRACE-0}" == "1" ]]; then
+    set -o xtrace
+fi
+
+DIR="/Users/rmn/masterThesis/eda-gp-2020/experiments/energyCooling_2hl_maxIndSize_fullRun_30gens"
+PROG="/Users/rmn/masterThesis/eda-gp-2020/src/testing_evaluate.py"
+ENV="/Users/rmn/miniconda3/envs/edaGP/bin/python"
+JUPYTER="/Users/rmn/miniconda3/envs/dataScience/bin/jupyter"
+
+NRUNS=10
+MAXGEN=30
+
+
+yes yes | $ENV $PROG -f $DIR -n $NRUNS -o "$DIR/results.csv"  -g $MAXGEN
+
+$JUPYTER nbconvert --to html --execute "/Users/rmn/masterThesis/eda-gp-2020/experiments/energyCooling_2hl_maxIndSize_fullRun_30gens/energyCooling_2hl_maxIndSize_fullRun_30gens.ipynb"

### 2023-01-23 slides.Rmd
autogit

@@
 
-```{bash recompile_results, echo=TRUE, eval=FALSE, include=FALSE}
+ ```{bash recompile_results, echo=TRUE, eval=TRUE, include=FALSE}
 
@@
 
-## Auswertung Fitness (Median) - Alternative Ansätze
+<!-- ## Auswertung Fitness (Median) - Alternative Ansätze -->
 
-```{r full_run_PT_Strat_Fitness, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
-library(knitr)
-library(kableExtra)
-library(dplyr)
+<!-- ```{r full_run_PT_Strat_Fitness, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE} -->
+<!-- library(knitr) -->
+<!-- library(kableExtra) -->
+<!-- library(dplyr) -->
 
-# Create the data frame from csv
-df <- read.csv(
-  "/Users/rmn/github/master_thesis/data/summary_median_Fitness_lastGen_PT_Strategies_Airfoil_2hl_150hn.csv",
-  sep=","
-)
+<!-- # Create the data frame from csv -->
+<!-- df <- read.csv( -->
+<!--   "/Users/rmn/github/master_thesis/data/summary_median_Fitness_lastGen_PT_Strategies_Airfoil_2hl_150hn.csv", -->
+<!--   sep="," -->
+<!-- ) -->
 
-col_names <- c("Set", "Standard-PT", "2ndGen", "P-Val(1)" , "GrowInit", "P-Val(2)")
+<!-- col_names <- c("Set", "Standard-PT", "2ndGen", "P-Val(1)" , "GrowInit", "P-Val(2)") -->
 
-knitr::kable(
- df,
- format = "markdown",
- align = "l",
- digits=2,
- col.names = col_names,
- #caption = "Symbolic Regression Overview - Mean final Fitness"
-) %>% 
-  kable_styling(latex_options = "scale_down", font_size = 7)
+<!-- knitr::kable( -->
+<!--  df, -->
+<!--  format = "markdown", -->
+<!--  align = "l", -->
+<!--  digits=2, -->
+<!--  col.names = col_names, -->
+<!--  #caption = "Symbolic Regression Overview - Mean final Fitness" -->
+<!-- ) %>%  -->
+<!--   kable_styling(latex_options = "scale_down", font_size = 7) -->
 
-```
+<!-- ``` -->
 
@@
 
-## Auswertung Lösungsgröße (Median) - Alternative Ansätze
+<!-- ## Auswertung Lösungsgröße (Median) - Alternative Ansätze -->
 
-```{r full_run_PT_Strat_BestSize, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
-library(knitr)
-library(kableExtra)
-library(dplyr)
+<!-- ```{r full_run_PT_Strat_BestSize, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE} -->
+<!-- library(knitr) -->
+<!-- library(kableExtra) -->
+<!-- library(dplyr) -->
 
-# Create the data frame from csv
-df <- read.csv(
-  "/Users/rmn/github/master_thesis/data/summary_median_Size_best_Solution_PT_Strategies_Airfoil_2hl_150hn.csv",
-  sep=","
-)
+<!-- # Create the data frame from csv -->
+<!-- df <- read.csv( -->
+<!--   "/Users/rmn/github/master_thesis/data/summary_median_Size_best_Solution_PT_Strategies_Airfoil_2hl_150hn.csv", -->
+<!--   sep="," -->
+<!-- ) -->
 
-col_names <- c("Standard-PT", "2ndGen", "P-Val(1)" , "GrowInit", "P-Val(2)")
+<!-- col_names <- c("Standard-PT", "2ndGen", "P-Val(1)" , "GrowInit", "P-Val(2)") -->
 
-knitr::kable(
- df,
- format = "markdown",
- align = "l",
- digits=2,
- col.names = col_names,
- #caption = "Symbolic Regression Overview - Mean final Fitness"
-) %>% 
-  kable_styling(latex_options = "scale_down", font_size = 7)
+<!-- knitr::kable( -->
+<!--  df, -->
+<!--  format = "markdown", -->
+<!--  align = "l", -->
+<!--  digits=2, -->
+<!--  col.names = col_names, -->
+<!--  #caption = "Symbolic Regression Overview - Mean final Fitness" -->
+<!-- ) %>%  -->
+<!--   kable_styling(latex_options = "scale_down", font_size = 7) -->
 
-```
+<!-- ``` -->
 
@@
 
-## Auswertung Populationsdiversität (Median) - Alternative Ansätze
+<!-- ## Auswertung Populationsdiversität (Median) - Alternative Ansätze -->
 
-```{r full_run_PT_Strat_LevDiv, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
-library(knitr)
-library(kableExtra)
-library(dplyr)
+<!-- ```{r full_run_PT_Strat_LevDiv, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE} -->
+<!-- library(knitr) -->
+<!-- library(kableExtra) -->
+<!-- library(dplyr) -->
 
-# Create the data frame from csv
-df <- read.csv(
-  "/Users/rmn/github/master_thesis/data/summary_median_Norm._Levenshtein_Distance_PT_Strategies_Airfoil_2hl_150hn.csv",
-  sep=","
-)
+<!-- # Create the data frame from csv -->
+<!-- df <- read.csv( -->
+<!--   "/Users/rmn/github/master_thesis/data/summary_median_Norm._Levenshtein_Distance_PT_Strategies_Airfoil_2hl_150hn.csv", -->
+<!--   sep="," -->
+<!-- ) -->
 
-col_names <- c("Standard-PT", "2ndGen", "P-Val(1)" , "GrowInit", "P-Val(2)")
+<!-- col_names <- c("Standard-PT", "2ndGen", "P-Val(1)" , "GrowInit", "P-Val(2)") -->
 
-knitr::kable(
- df,
- format = "markdown",
- align = "l",
- digits=2,
- col.names = col_names,
- #caption = "Symbolic Regression Overview - Mean final Fitness"
-) %>% 
-  kable_styling(latex_options = "scale_down", font_size = 7)
+<!-- knitr::kable( -->
+<!--  df, -->
+<!--  format = "markdown", -->
+<!--  align = "l", -->
+<!--  digits=2, -->
+<!--  col.names = col_names, -->
+<!--  #caption = "Symbolic Regression Overview - Mean final Fitness" -->
+<!-- ) %>%  -->
+<!--   kable_styling(latex_options = "scale_down", font_size = 7) -->
 
-```
+<!-- ``` -->
 ## Trainingsepochen (Median) - Alternative Ansätze
@@
 
-## Auswertung Trainingsepochen (Median) - Alternative Ansätze
+<!-- ## Auswertung Trainingsepochen (Median) - Alternative Ansätze -->
 
-```{r full_run_PT_Strat_EpochsTrained, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
-library(knitr)
-library(kableExtra)
-library(dplyr)
+<!-- ```{r full_run_PT_Strat_EpochsTrained, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE} -->
+<!-- library(knitr) -->
+<!-- library(kableExtra) -->
+<!-- library(dplyr) -->
 
-# Create the data frame from csv
-df <- read.csv(
-  "/Users/rmn/github/master_thesis/data/summary_median_Training_Epochs_PT_Strategies_Airfoil_2hl_150hn.csv",
-  sep=","
-)
+<!-- # Create the data frame from csv -->
+<!-- df <- read.csv( -->
+<!--   "/Users/rmn/github/master_thesis/data/summary_median_Training_Epochs_PT_Strategies_Airfoil_2hl_150hn.csv", -->
+<!--   sep="," -->
+<!-- ) -->
 
-col_names <- c("Standard-PT", "2ndGen", "P-Val(1)" , "GrowInit", "P-Val(2)")
+<!-- col_names <- c("Standard-PT", "2ndGen", "P-Val(1)" , "GrowInit", "P-Val(2)") -->
 
-knitr::kable(
- df,
- format = "markdown",
- align = "l",
- digits=2,
- col.names = col_names,
- #caption = "Symbolic Regression Overview - Mean final Fitness"
-) %>% 
-  kable_styling(latex_options = "scale_down", font_size = 7)
+<!-- knitr::kable( -->
+<!--  df, -->
+<!--  format = "markdown", -->
+<!--  align = "l", -->
+<!--  digits=2, -->
+<!--  col.names = col_names, -->
+<!--  #caption = "Symbolic Regression Overview - Mean final Fitness" -->
+<!-- ) %>%  -->
+<!--   kable_styling(latex_options = "scale_down", font_size = 7) -->
 
-```
-# Fazit (vorläufig)
+<!-- ``` -->
+
+# Fazit
 

### 2023-01-23 experiments/2ndGenPT_airfoil_2hl_150hn_fullRun/re_evaluate_2ndGenPT_airfoil_2hl_150hn_fullRun.sh

@@ -1,25 +1,0 @@
-#/usr/bin/env bash
-
-# Re-Evaluate all results from airfoilFirstGen experiments
-
-bash config
-set -o errexit
-set -o nounset
-# set -o pipefail
-if [[ "${TRACE-0}" == "1" ]]; then
-    set -o xtrace
-fi
-
-DIR="/Users/rmn/masterThesis/eda-gp-2020/experiments/2ndGenPT_airfoil_2hl_150hn_fullRun"
-PROG="/Users/rmn/masterThesis/eda-gp-2020/src/testing_evaluate.py"
-ENV="/Users/rmn/miniconda3/envs/edaGP/bin/python"
-JUPYTER="/Users/rmn/miniconda3/envs/dataScience/bin/jupyter"
-
-NRUNS=10
-MAXGEN=30
-
-
-yes yes | $ENV $PROG -f $DIR -n $NRUNS -o "$DIR/results.csv"  -g $MAXGEN
-
-
-$JUPYTER nbconvert --to html --execute "/Users/rmn/masterThesis/eda-gp-2020/experiments/2ndGenPT_airfoil_2hl_150hn_fullRun/2ndGenPT_airfoil_2hl_150hn_fullRun.ipynb"

### 2023-01-23 experiments/3ndGenPT_airfoil_2hl_100hn_fullRun/re_evaluate_3ndGenPT_airfoil_2hl_100hn_fullRun.sh

@@ -1,25 +1,0 @@
-#/usr/bin/env bash
-
-# Re-Evaluate all results from airfoilFirstGen experiments
-
-bash config
-set -o errexit
-set -o nounset
-# set -o pipefail
-if [[ "${TRACE-0}" == "1" ]]; then
-    set -o xtrace
-fi
-
-DIR="/Users/rmn/masterThesis/eda-gp-2020/experiments/3ndGenPT_airfoil_2hl_100hn_fullRun"
-PROG="/Users/rmn/masterThesis/eda-gp-2020/src/testing_evaluate.py"
-ENV="/Users/rmn/miniconda3/envs/edaGP/bin/python"
-JUPYTER="/Users/rmn/miniconda3/envs/dataScience/bin/jupyter"
-
-NRUNS=10
-MAXGEN=20
-
-
-yes yes | $ENV $PROG -f $DIR -n $NRUNS -o "$DIR/results.csv"  -g $MAXGEN
-
-
-#$JUPYTER nbconvert --to html --execute "/Users/rmn/masterThesis/eda-gp-2020/experiments/airfoil_2hl_maxIndSize_fullRun_30gens/airfoil_2hl_maxIndSize_fullRun_30gens_with2ndGen.ipynb"

### 2023-01-23 experiments/airfoil_1hl_maxIndSize_fullRun_30gens/re_evaluate_airfoil_1hl_maxIndSize_30gen.sh

@@ -1,24 +1,0 @@
-#/usr/bin/env bash
-
-# Re-Evaluate all results from airfoilFirstGen experiments
-
-bash config
-set -o errexit
-set -o nounset
-# set -o pipefail
-if [[ "${TRACE-0}" == "1" ]]; then
-    set -o xtrace
-fi
-
-DIR="/Users/rmn/masterThesis/eda-gp-2020/experiments/airfoil_1hl_maxIndSize_fullRun_30gens"
-PROG="/Users/rmn/masterThesis/eda-gp-2020/src/testing_evaluate.py"
-ENV="/Users/rmn/miniconda3/envs/edaGP/bin/python"
-JUPYTER="/Users/rmn/miniconda3/envs/dataScience/bin/jupyter"
-
-NRUNS=10
-MAXGEN=30
-
-
-yes yes | $ENV $PROG -f $DIR -n $NRUNS -o "$DIR/results.csv"  -g $MAXGEN
-
-$JUPYTER nbconvert --to html --execute "/Users/rmn/masterThesis/eda-gp-2020/experiments/airfoil_1hl_maxIndSize_fullRun_30gens/airfoil_1hl_maxIndSize_fullRun_30gens.ipynb"

### 2023-01-23 experiments/airfoil_2hl_maxIndSize_fullRun_30gens/re_evaluate_airfoil_2hl_maxIndSize_30gen.sh

@@ -1,28 +1,0 @@
-#/usr/bin/env bash
-
-# Re-Evaluate all results from airfoilFirstGen experiments
-
-bash config
-set -o errexit
-set -o nounset
-# set -o pipefail
-if [[ "${TRACE-0}" == "1" ]]; then
-    set -o xtrace
-fi
-
-DIR="/Users/rmn/masterThesis/eda-gp-2020/experiments/airfoil_2hl_maxIndSize_fullRun_30gens"
-PROG="/Users/rmn/masterThesis/eda-gp-2020/src/testing_evaluate.py"
-ENV="/Users/rmn/miniconda3/envs/edaGP/bin/python"
-JUPYTER="/Users/rmn/miniconda3/envs/dataScience/bin/jupyter"
-
-NRUNS=10
-MAXGEN=30
-
-
-yes yes | $ENV $PROG -f $DIR -n $NRUNS -o "$DIR/results.csv"  -g $MAXGEN
-
-$JUPYTER nbconvert --to html --execute "/Users/rmn/masterThesis/eda-gp-2020/experiments/airfoil_2hl_maxIndSize_fullRun_30gens/airfoil_2hl_maxIndSize_fullRun_30gens.ipynb"
-
-$JUPYTER nbconvert --to html --execute "/Users/rmn/masterThesis/eda-gp-2020/experiments/airfoil_2hl_maxIndSize_fullRun_30gens/airfoil_2hl_maxIndSize_fullRun_30gens_withGP.ipynb"
-
-$JUPYTER nbconvert --to html --execute "/Users/rmn/masterThesis/eda-gp-2020/experiments/airfoil_2hl_maxIndSize_fullRun_30gens/airfoil_2hl_maxIndSize_fullRun_30gens_TimeBenchmark.ipynb"

### 2023-01-23 experiments/airfoil_firstGen/re_evaluate_airfoilFirstGen.sh

@@ -1,40 +1,0 @@
-#/usr/bin/env bash
-
-# Re-Evaluate all results from airfoilFirstGen experiments
-
-bash config
-set -o errexit
-set -o nounset
-# set -o pipefail
-if [[ "${TRACE-0}" == "1" ]]; then
-    set -o xtrace
-fi
-
-DIR="/Users/rmn/masterThesis/eda-gp-2020/experiments/airfoil_firstGen"
-PROG="/Users/rmn/masterThesis/eda-gp-2020/src/testing_evaluate.py"
-ENV="/Users/rmn/miniconda3/envs/edaGP/bin/python"
-
-NRUNS=10
-MAXGEN=1
-
-COUNTER=0
-
-# Find all subdirectories of DIR
-for SUBDIR in $(find "$DIR" -type d -maxdepth 1); do
-  if [ "$SUBDIR" = "$DIR" ]; then
-      continue
-    fi
-  if [[ "$SUBDIR" == *".ipynb_checkpoints"* ]]; then
-    continue
-  fi
-  yes yes | $ENV $PROG -f $SUBDIR -n $NRUNS -o "$SUBDIR/results.csv"  -g $MAXGEN
-  ((COUNTER++))
-
-done
-
-echo "Done - Evaluated $COUNTER Experiments!"
-
-JUPYTER="/Users/rmn/miniconda3/envs/dataScience/bin/jupyter"
-
-$JUPYTER nbconvert --to html --execute "/Users/rmn/masterThesis/eda-gp-2020/experiments/airfoil_firstGen/airfoil_1gen_combined_layers.ipynb"
-$JUPYTER nbconvert --to html --execute "/Users/rmn/masterThesis/eda-gp-2020/experiments/airfoil_firstGen/airfoil_1gen_combined_neurons.ipynb"

### 2023-01-23 experiments/bostonHousing_2hl_maxIndSize_fullRun_30gens/re_evaluate_bostonHousing_2hl_maxIndSize_fullRun_30gens.sh

@@ -1,24 +1,0 @@
-#/usr/bin/env bash
-
-# Re-Evaluate all results from bostonHousing_2hl_maxIndSize_fullRun_30gens
-
-bash config
-set -o errexit
-set -o nounset
-# set -o pipefail
-if [[ "${TRACE-0}" == "1" ]]; then
-    set -o xtrace
-fi
-
-DIR="/Users/rmn/masterThesis/eda-gp-2020/experiments/bostonHousing_2hl_maxIndSize_fullRun_30gens"
-PROG="/Users/rmn/masterThesis/eda-gp-2020/src/testing_evaluate.py"
-ENV="/Users/rmn/miniconda3/envs/edaGP/bin/python"
-JUPYTER="/Users/rmn/miniconda3/envs/dataScience/bin/jupyter"
-
-NRUNS=10
-MAXGEN=30
-
-
-yes yes | $ENV $PROG -f $DIR -n $NRUNS -o "$DIR/results.csv"  -g $MAXGEN
-
-$JUPYTER nbconvert --to html --execute "/Users/rmn/masterThesis/eda-gp-2020/experiments/bostonHousing_2hl_maxIndSize_fullRun_30gens/bostonHousing_2hl_maxIndSize_fullRun_30gens.ipynb"

### 2023-01-23 experiments/concrete_2hl_maxIndSize_fullRun_30gens/re_evaluate_concrete_2hl_maxIndSize_fullRun_30gens.sh

@@ -1,24 +1,0 @@
-#/usr/bin/env bash
-
-# Re-Evaluate all results from energyCooling_2hl_maxIndSize_fullRun_30gens
-
-bash config
-set -o errexit
-set -o nounset
-# set -o pipefail
-if [[ "${TRACE-0}" == "1" ]]; then
-    set -o xtrace
-fi
-
-DIR="/Users/rmn/masterThesis/eda-gp-2020/experiments/concrete_2hl_maxIndSize_fullRun_30gens"
-PROG="/Users/rmn/masterThesis/eda-gp-2020/src/testing_evaluate.py"
-ENV="/Users/rmn/miniconda3/envs/edaGP/bin/python"
-JUPYTER="/Users/rmn/miniconda3/envs/dataScience/bin/jupyter"
-
-NRUNS=10
-MAXGEN=30
-
-
-yes yes | $ENV $PROG -f $DIR -n $NRUNS -o "$DIR/results.csv"  -g $MAXGEN
-
-$JUPYTER nbconvert --to html --execute "/Users/rmn/masterThesis/eda-gp-2020/experiments/concrete_2hl_maxIndSize_fullRun_30gens/concrete_2hl_maxIndSize_fullRun_30gens.ipynb"

### 2023-01-23 experiments/energyCooling_2hl_maxIndSize_fullRun_30gens/re_evaluate_energyCooling_2hl_maxIndSize_fullRun_30gens.sh

@@ -1,24 +1,0 @@
-#/usr/bin/env bash
-
-# Re-Evaluate all results from energyCooling_2hl_maxIndSize_fullRun_30gens
-
-bash config
-set -o errexit
-set -o nounset
-# set -o pipefail
-if [[ "${TRACE-0}" == "1" ]]; then
-    set -o xtrace
-fi
-
-DIR="/Users/rmn/masterThesis/eda-gp-2020/experiments/energyCooling_2hl_maxIndSize_fullRun_30gens"
-PROG="/Users/rmn/masterThesis/eda-gp-2020/src/testing_evaluate.py"
-ENV="/Users/rmn/miniconda3/envs/edaGP/bin/python"
-JUPYTER="/Users/rmn/miniconda3/envs/dataScience/bin/jupyter"
-
-NRUNS=10
-MAXGEN=30
-
-
-yes yes | $ENV $PROG -f $DIR -n $NRUNS -o "$DIR/results.csv"  -g $MAXGEN
-
-$JUPYTER nbconvert --to html --execute "/Users/rmn/masterThesis/eda-gp-2020/experiments/energyCooling_2hl_maxIndSize_fullRun_30gens/energyCooling_2hl_maxIndSize_fullRun_30gens.ipynb"

### 2023-01-24 data/summarize_data.py
Tue Jan 24 11:00:07 CET 2023: update-remote.sh

@@ -208,7 +208,7 @@ fig, ((ul, ur), (dl, dr)) = plt.subplots(nrows=2, ncols=2, sharex=True, sharey=F
 fig.set_size_inches(14,12)
 #fig.suptitle(f"{BASE_TITLE} - Mean Size of best Solution", fontsize=BIG)
 fig.supxlabel("Generations", fontsize=MID)
-fig.supylabel("Size", fontsize=MID)
+fig.supylabel("Tree Size", fontsize=MID)
 
 for (problem,path), ax in zip(RESULTS_2hl.items(), (ul, ur, dl, dr)):
 

### 2023-01-27 paper.Rmd
Mon Jan 16 12:00:10 CET 2023: update-remote.sh
Fri Jan 27 11:00:06 CET 2023: update-remote.sh

@@
 fontsize: 11pt
-bibliography: ref/ref.bib
+bibliography: ./ref/ref.bib
 link-citations: true
@@
 
-```{bash recompile_results, include=FALSE, eval=FALSE}
+```{bash recompile_results, echo=TRUE, eval=FALSE, include=FALSE}
+
+# set eval to TRUE to recompile all results, plots etc.
+# Warning: This will take some time...
 
@@
 
-# summarize csv files for full run experiments
+# summarize csv files and plots for full run experiments
 /Users/rmn/miniconda3/envs/dataScience/bin/python /Users/rmn/github/master_thesis/data/summarize_data.py
+
+# 2nd generation Pre-Training
+bash /Users/rmn/masterThesis/eda-gp-2020/experiments/2ndGenPT_airfoil_2hl_150hn_fullRun/re_evaluate_2ndGenPT_airfoil_2hl_150hn_fullRun.sh
+
+# create Plots from pstats (cProfile)
+JUPYTER="/Users/rmn/miniconda3/envs/dataScience/bin/jupyter"
+$JUPYTER nbconvert --to html --execute /Users/rmn/masterThesis/eda-gp-2020/cProfiler/eval.ipynb
+
+echo Done recompiling Results, Plots, Diagrams, etc!
 
@@
 
-DAE-LSTMs are recurrent neural networks (RNNs) that can be trained in an unsupervised learning environment to minimize a reconstruction error for encoding input data into a compressed representation and subsequently decoding the compressed representation back to the input dimension.
+DAE-LSTMs are artificial neural networks that can be trained in an unsupervised learning environment to minimize a reconstruction error for encoding input data into a compressed representation and subsequently decoding the compressed representation back to the input dimension.
 In DAE-GP, DAE-LSTMs are trained with a subset of high-fitness solutions selected from a parent population with the aim to capture their promising qualities. 
@@
 The DAE-GP algorithm first described by [@dae-gp_2020_rtree] trains a DAE-LSTMs for each generation $g$ of the search from scratch.
-@dae-gp_2022_symreg and @daegp_explore_exploit suggests the incorporation of a pre-training strategy into the evolutionary search as a possible way of improving the model quality of DAE-GP.
+@dae-gp_2022_symreg and @daegp_explore_exploit suggests the incorporation of a pre-training strategy into the evolutionary search as a possible way of improving the overall performance of the DAE-GP algorithm.
 The key idea is to pre-train an initial DAE-LSTM on a large population of candidate solutions and to use the pre-trained parameters of this initial model in each generation of the search as starting parameters for the current generation DAE-LSTM.
-This thesis studies the influence of using pre-trained DAE-LSTMs in DAE-GP for symbolic regression, especially looking at the influence in generalization behavior and the overall quality of solutions found by DAE-GP.
-The aim of this study is to answer the question if a pre-training strategy can help to improve generalization in DAE-GP and if so, analyze how this improved generalization ability benefits the overall performance of DAE-GP.
+This thesis studies the influence of using pre-trained DAE-LSTMs in DAE-GP for symbolic regression, especially looking at the influence on overall quality of solutions found by DAE-GP and effects on runtime.
+The aim of this study is to answer the question if a pre-training strategy can be used to improve DAE-GP and to study both positive and negative effects on overall performance.
 
@@
 
-DAE was first introduced by @10.1145/1390156.1390294 as an improved AE with the ability to learn new representations of data that is especially robust to partially corrupted input data.
+DAE was first introduced by @dae_orig2008 as an improved AE with the ability to learn new representations of data that is especially robust to partially corrupted input data.
 DAE modifies the AE by using partially corrupted input data for the AE and training it to reconstruct the uncorrupted, original version of the input data.
@@
 
-LSTMs are a variant of recurrent neural networks first introduced by @lstm_orig that can store learned information over en extended time period while avoiding the problem of vanishing and/or exploding gradients.
+LSTMs are a variant of artificial neural networks first introduced by @lstm_orig that can store learned information over en extended time period while avoiding the problem of vanishing and/or exploding gradients.
 Since DAE-GP encodes candidate solutions as linear strings in prefix notation, the DAE in DAE-GP uses LSTMs for both encoding and decoding where the total amount of time steps $T$ is equal to sum of the length of the input solution and the output solution [@dae-gp_2020_rtree].
@@
 
-Pre-Training describes the concept of initially training a neural network on a large dataset before using it to solve a more specific task. 
+Pre-Training describes the concept of initially training an artificial neural network on a large dataset before using it to solve a more specific task. 
 It is a commonly used strategy in deep architectures that has been shown to improve both the optimization process itself as well as the generalization behavior if compared to the standard approach of using randomly initialized parameters [@pmlr-v5-erhan09a].
@@
 
-Another important finding by @pmlr-v5-erhan09a is that the positive effect of pre-training described above is dependent on both the depth of the network as well as the size of layers - while increasingly larger networks benefit increasingly more from pre-training the final performance of small architectures tends to be worse with pre-training than with randomized weight initialization.
+Another important finding by @pmlr-v5-erhan09a is that the positive effect of pre-training is dependent on both the depth of the network as well as the size of layers - while increasingly larger networks benefit increasingly more from pre-training the final performance of small architectures tends to be worse with pre-training than with randomized weight initialization.
 The authors find evidence that pre-training DAE is especially usefull for optimizing the parameters of the lower layers of the network.
@@
 
-The DAE-GP algorithm, first described by @dae-gp_2020_rtree, is summarized as a flowchart in figure §. 
+The DAE-GP algorithm, first described by @dae-gp_2020_rtree, is summarized as a flowchart in \@ref(fig:dae-gp-flowchart). 
 After setting the generation counter $g$ to $0$ and creating the initial population of solutions $P_0$, DAE-GP enters the main loop that checks if a termination criteria is satisfied (i.e.
@@
 
-![Flowchart - Regular DAE-GP](./img/flowcharts/dae-gp.png)
+```{r dae-gp-flowchart, echo=FALSE, fig.cap="Regular DAE-GP Flowchart", out.height="90%", out.width="50%"}
+knitr::include_graphics("./img/flowcharts/dae-gp.png")
+```
 

### 2023-01-27 paper.Rmd
autogit

@@
 output:
-  pdf_document:
-    # keep_tex: true
+  bookdown::pdf_document2:
+    keep_tex: true
     extra_dependencies: ["float"]
@@
 
-knitr::opts_chunk$set(echo = TRUE)
-#knitr::opts_chunk$set(fig.pos = "!H")
+knitr::opts_chunk$set(
+  echo = TRUE,
+  fig.pos = "c",
+  fig.align="center"
+  )
 
@@
 
-```{bash recompile_results, echo=TRUE, eval=FALSE, include=FALSE}
+```{bash recompile-results, echo=TRUE, eval=FALSE, include=FALSE}
 
@@
 
-The DAE-GP algorithm, first described by @dae-gp_2020_rtree, is summarized as a flowchart in \@ref(fig:dae-gp-flowchart). 
+The DAE-GP algorithm, first described by @dae-gp_2020_rtree, is summarized as a flowchart in \@ref(fig:daegp-flowchart).
 After setting the generation counter $g$ to $0$ and creating the initial population of solutions $P_0$, DAE-GP enters the main loop that checks if a termination criteria is satisfied (i.e.
@@
 
-```{r dae-gp-flowchart, echo=FALSE, fig.cap="Regular DAE-GP Flowchart", out.height="90%", out.width="50%"}
+```{r daegp-flowchart, echo=FALSE, fig.cap="Regular DAE-GP Flowchart", out.height="90%", out.width="50%"}
 knitr::include_graphics("./img/flowcharts/dae-gp.png")
@@
 
-```{r airfoil_dataset_description, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
+```{r airfoil-dataset-description, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
 fig.pos="H"
@@
 
-```{r airfoil_function_set, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
+```{r airfoil-function-set, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
 fig.pos="H"
@@
 
-```{r airfoil_fullRun_2hl_maxIndSize_params, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
+```{r airfoil-fullRun-2hl-maxIndSize-params, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
 fig.pos="H"
@@
 
-```{r full_run_realWorldSymReg_problems, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
+```{r full-run-realWorldSymReg-problems, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
 fig.pos="H"
@@
 
-```{r full_run_realWorldSymReg_fitness, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
+```{r full-run-realWorldSymReg-fitness, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
 library(knitr)
@@
 
-```{r full_run_realWorldSymReg_best_solution_size, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
+```{r full-run-realWorldSymReg-best-solution-size, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
 library(knitr)
@@
 
-```{r full_run_realWorldSymReg_popDiversity, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
+```{r full-run-realWorldSymReg-popDiversity, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
 library(knitr)
@@
 
-```{r full_run_realWorldSymReg_epochsPerGen, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
+```{r full-run-realWorldSymReg-epochsPerGen, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
 library(knitr)

### 2023-01-27 paper.tex
autogit

@@
 \usepackage{amsmath,amssymb}
-\usepackage{lmodern}
 \usepackage{iftex}
@@
 \else % if luatex or xetex
-  \usepackage{unicode-math}
+  \usepackage{unicode-math} % this also loads fontspec
   \defaultfontfeatures{Scale=MatchLowercase}
   \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+\usepackage{lmodern}
+\ifPDFTeX\else
+  % xetex/luatex font selection
 \fi
@@
 \IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\urlstyle{same} % disable monospaced font for URLs
+\urlstyle{same}
 \hypersetup{
@@
 
-\title{Pre-Trained Denoising Autoencoders Long Short-Term Memory
-Networks as probabilistic Models for Estimation of Distribution Genetic
-Programming
+\title{Pre-Trained Denoising Autoencoders Long Short-Term Memory Networks as probabilistic Models for Estimation of Distribution Genetic Programming
 
@@
 \author{}
-\date{\vspace{-2.5em}Date of Submission: 2023-01-11}
+\date{\vspace{-2.5em}Date of Submission: 2023-01-27}
 
@@
 
+{
+\setcounter{tocdepth}{2}
+\tableofcontents
+}
 \setcounter{page}{1}
@@
 
-Denoising Autoencoder Genetic Programming (DAE-GP) is a novel variation
-of an genetic programming based Estimation of Distribution Algorithm
-(EDA-GP) that uses a denoising autoencoders long short-term memory
-network (DAE-LSTMs) as a probabilistic model to sample new candidate
-solutions {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]}.
-
-DAE-LSTMs are recurrent neural networks (RNNs) that can be trained in an
-unsupervised learning environment to minimize a reconstruction error for
-encoding input data into a compressed representation and subsequently
-decoding the compressed representation back to the input dimension. In
-DAE-GP, DAE-LSTMs are trained with a subset of high-fitness solutions
-selected from a parent population with the aim to capture their
-promising qualities. The resulting model is then used to sample new
-offspring solutions by propagating partially mutated solutions from the
-parent population through the DAE-LSTMS
-{[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]}. In previous work
-DAE-GP has been shown to outperform GP for both a generalized version of
-the royal tree problem
-{[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]} as well as for a
-real-world symbolic regression problem
-{[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]}.
-
-The DAE-GP algorithm first described by
-{[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]} trains a DAE-LSTMs
-for each generation \(g\) of the search from scratch.
-{[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]} and
-{[}\protect\hyperlink{ref-daegp_explore_exploit}{16}{]} suggests the
-incorporation of a pre-training strategy into the evolutionary search as
-a possible way of improving the model quality of DAE-GP. The key idea is
-to pre-train an initial DAE-LSTM on a large population of candidate
-solutions and to use the pre-trained parameters of this initial model in
-each generation of the search as starting parameters for the current
-generation DAE-LSTM. This thesis studies the influence of using
-pre-trained DAE-LSTMs in DAE-GP for symbolic regression, especially
-looking at the influence in generalization behavior and the overall
-quality of solutions found by DAE-GP. The aim of this study is to answer
-the question if a pre-training strategy can help to improve
-generalization in DAE-GP and if so, analyze how this improved
-generalization ability benefits the overall performance of DAE-GP.
+Denoising Autoencoder Genetic Programming (DAE-GP) is a novel variation of an genetic programming based Estimation of Distribution Algorithm (EDA-GP) that uses a denoising autoencoders long short-term memory network (DAE-LSTMs) as a probabilistic model to sample new candidate solutions {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]}.
+
+DAE-LSTMs are artificial neural networks that can be trained in an unsupervised learning environment to minimize a reconstruction error for encoding input data into a compressed representation and subsequently decoding the compressed representation back to the input dimension.
+In DAE-GP, DAE-LSTMs are trained with a subset of high-fitness solutions selected from a parent population with the aim to capture their promising qualities.
+The resulting model is then used to sample new offspring solutions by propagating partially mutated solutions from the parent population through the DAE-LSTMS {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]}.
+In previous work DAE-GP has been shown to outperform GP for both a generalized version of the royal tree problem {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]} as well as for a real-world symbolic regression problem {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]}.
+
+The DAE-GP algorithm first described by {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]} trains a DAE-LSTMs for each generation \(g\) of the search from scratch.
+{[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]} and {[}\protect\hyperlink{ref-daegp_explore_exploit}{16}{]} suggests the incorporation of a pre-training strategy into the evolutionary search as a possible way of improving the overall performance of the DAE-GP algorithm.
+The key idea is to pre-train an initial DAE-LSTM on a large population of candidate solutions and to use the pre-trained parameters of this initial model in each generation of the search as starting parameters for the current generation DAE-LSTM.
+This thesis studies the influence of using pre-trained DAE-LSTMs in DAE-GP for symbolic regression, especially looking at the influence on overall quality of solutions found by DAE-GP and effects on runtime.
+The aim of this study is to answer the question if a pre-training strategy can be used to improve DAE-GP and to study both positive and negative effects on overall performance.
 
@@
 
-\emph{This section describes the relevant concepts that are necessary
-for the understanding and classification of DAE-GP as well as the
-concept of pre-training in artificial neural networks}
+\emph{This section describes the relevant concepts that are necessary for the understanding and classification of DAE-GP as well as the concept of pre-training in artificial neural networks}
 
 \hypertarget{denoising-autoencoder-genetic-programming}{%
-\subsection{Denoising Autoencoder Genetic
-Programming}\label{denoising-autoencoder-genetic-programming}}
-
-DAE-GP is an EDA-GP algorithm that uses DAE-LSTM networks as a
-probabilistic model to sample new offspring solutions
-{[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]}.
+\subsection{Denoising Autoencoder Genetic Programming}\label{denoising-autoencoder-genetic-programming}}
+
+DAE-GP is an EDA-GP algorithm that uses DAE-LSTM networks as a probabilistic model to sample new offspring solutions {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]}.
 
 \hypertarget{evolutionary-computation}{%
-\subsubsection{Evolutionary
-Computation}\label{evolutionary-computation}}
-
-As a variant of GP, DAE-GP can be classified as a meta-heuristic that
-belongs to the field of evolutionary computation (EC). EC based
-meta-heuristics are optimization methods that simulate the process of
-Darwinian evolution to search for high quality solutions by applying
-selection and variation operators to a population of candidate
-solutions. Examples of EC include genetic algorithms (GA), evolutionary
-strategies (ES) and GP. In EC, the quality of a solution is commonly
-measured as fitness and the time steps of the search are called
-generations. Another important concept in EC is the distinction between
-genotypes and phenotypes of solutions, the genotype contains the
-information that is necessary to construct the phenotype, the outer
-appearance of a particular solution on which we measure the overall
-quality of solutions. The representation of a solution is therefore
-defined by the mapping of genotypes to phenotypes
-{[}\protect\hyperlink{ref-design_of_modern_heuristics}{12}{]}. Genetic
-operators, such as mutation or recombination are usually applied to the
-genotype of solutions.
+\subsubsection{Evolutionary Computation}\label{evolutionary-computation}}
+
+As a variant of GP, DAE-GP can be classified as a meta-heuristic that belongs to the field of evolutionary computation (EC).
+EC based meta-heuristics are optimization methods that simulate the process of Darwinian evolution to search for high quality solutions by applying selection and variation operators to a population of candidate solutions.
+Examples of EC include genetic algorithms (GA), evolutionary strategies (ES) and GP.
+In EC, the quality of a solution is commonly measured as fitness and the time steps of the search are called generations.
+Another important concept in EC is the distinction between genotypes and phenotypes of solutions, the genotype contains the information that is necessary to construct the phenotype, the outer appearance of a particular solution on which we measure the overall quality of solutions.
+The representation of a solution is therefore defined by the mapping of genotypes to phenotypes {[}\protect\hyperlink{ref-design_of_modern_heuristics}{12}{]}.
+Genetic operators, such as mutation or recombination are usually applied to the genotype of solutions.
 
@@
 
-GP follows the same basic evolutionary principle of EC but searches for
-more general, hierarchical computer programs of dynamically varying size
-and shape {[}\protect\hyperlink{ref-Koza1993GeneticP}{8}{]}. The
-computer programs that are at the center of the evolutionary search in
-GP are commonly represented by tree structures at the level of their
-phenotype {[}\protect\hyperlink{ref-design_of_modern_heuristics}{12}{]}.
-Since GP searches for high fitness computer programs that produce a
-desired output for some input, it can be applied to various different
-problem domains such as symbolic regression, automatic programming, or
-evolving game-playing strategies
-{[}\protect\hyperlink{ref-Koza1993GeneticP}{8}{]}. An important quality
-of GP is the ability to search for solutions of variable length and
-structure. GP is an especially useful meta heuristic for problems where
-no a priori knowledge about the final form of good solutions is
-available. GPs ability to optimize solutions for their structure as well
-as for their parameters led to it being one of the most prevalent
-methods used for symbolic regression
-{[}\protect\hyperlink{ref-10.1007ux2f978-3-540-24621-3_22}{10}{]}.
+GP follows the same basic evolutionary principle of EC but searches for more general, hierarchical computer programs of dynamically varying size and shape {[}\protect\hyperlink{ref-Koza1993GeneticP}{8}{]}.
+The computer programs that are at the center of the evolutionary search in GP are commonly represented by tree structures at the level of their phenotype {[}\protect\hyperlink{ref-design_of_modern_heuristics}{12}{]}.
+Since GP searches for high fitness computer programs that produce a desired output for some input, it can be applied to various different problem domains such as symbolic regression, automatic programming, or evolving game-playing strategies {[}\protect\hyperlink{ref-Koza1993GeneticP}{8}{]}.
+An important quality of GP is the ability to search for solutions of variable length and structure.
+GP is an especially useful meta heuristic for problems where no a priori knowledge about the final form of good solutions is available.
+GPs ability to optimize solutions for their structure as well as for their parameters led to it being one of the most prevalent methods used for symbolic regression {[}\protect\hyperlink{ref-10.1007ux2f978-3-540-24621-3_22}{10}{]}.
 
 \hypertarget{estimation-of-distribution-algorithms}{%
-\subsubsection{Estimation of Distribution
-Algorithms}\label{estimation-of-distribution-algorithms}}
-
-The aim of Estimation of distribution algorithms (EDA) is to replace the
-standard variation operators used in GA by building probabilistic models
-that can capture complex dependencies between different decision
-variables of an optimization problem
-{[}\protect\hyperlink{ref-design_of_modern_heuristics}{12}{]}. EDAs use
-this probabilistic model to sample new offspring solutions inside an
-evolutionary search to replace crossover and/or mutation operators.
+\subsubsection{Estimation of Distribution Algorithms}\label{estimation-of-distribution-algorithms}}
+
+The aim of Estimation of distribution algorithms (EDA) is to replace the standard variation operators used in GA by building probabilistic models that can capture complex dependencies between different decision variables of an optimization problem {[}\protect\hyperlink{ref-design_of_modern_heuristics}{12}{]}.
+EDAs use this probabilistic model to sample new offspring solutions inside an evolutionary search to replace crossover and/or mutation operators.
 
@@
 
-One possible way of model building in EDA proposed by
-{[}\protect\hyperlink{ref-harmless_overfitting_eda}{11}{]} is to use
-denoising autoencoders (DAEs) as generative models to capture complex
-probability distributions of decision variables. DAE, a variation of the
-autoencoders (AE), are a widely used type of neural networks in the
-domain of unsupervized machine learning that maps \(n\) input variables
-to \(n\) output variables using a hidden representation.
-
-AE were introduced by {[}\protect\hyperlink{ref-ae_orig}{6}{]} to
-compress high-dimensional data into lower-dimensions. An AE consists of
-two different subunits:
+One possible way of model building in EDA proposed by {[}\protect\hyperlink{ref-harmless_overfitting_eda}{11}{]} is to use denoising autoencoders (DAEs) as generative models to capture complex probability distributions of decision variables.
+DAE, a variation of the autoencoders (AE), are a widely used type of neural networks in the domain of unsupervized machine learning that maps \(n\) input variables to \(n\) output variables using a hidden representation.
+
+AE were introduced by {[}\protect\hyperlink{ref-ae_orig}{6}{]} to compress high-dimensional data into lower-dimensions. An AE consists of two different subunits:
 
@@
 \item
-  Decoder \(d(h)\): Decode and output the the encoded data back to its
-  original dimension
+  Decoder \(d(h)\): Decode and output the the encoded data back to its original dimension
 \end{itemize}
 
-The AE is trained to reduce the reconstruction error between input and
-output data, after the training procedure is finished the network is
-able to reduce the dimensionality of input data to a compressed
-representation{[}\protect\hyperlink{ref-ae_orig}{6}{]}.
-
-DAE was first introduced by
-{[}\protect\hyperlink{ref-10.1145ux2f1390156.1390294}{14}{]} as an
-improved AE with the ability to learn new representations of data that
-is especially robust to partially corrupted input data. DAE modifies the
-AE by using partially corrupted input data for the AE and training it to
-reconstruct the uncorrupted, original version of the input data.
-
-Since the hidden representation of DAE captures the dependency structure
-of the input data it can therefore also be used to generate new
-solutions in the context of GAs
-{[}\protect\hyperlink{ref-harmless_overfitting_eda}{11}{]}.
-
-DAE-GP builds upon the concept of using DAEs in EDAs described by
-{[}\protect\hyperlink{ref-harmless_overfitting_eda}{11}{]} and transfers
-the concept to the domain of GP. The mutation and crossover operators of
-standard GP are replaced by sampling new solutions from a probabilistic
-model that is build by training a DAEs long short-term memory (LSTM)
-network on a subset of high fitness solutions from the current
-population {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]}.
-
-LSTMs are a variant of recurrent neural networks first introduced by
-{[}\protect\hyperlink{ref-lstm_orig}{7}{]} that can store learned
-information over en extended time period while avoiding the problem of
-vanishing and/or exploding gradients. Since DAE-GP encodes candidate
-solutions as linear strings in prefix notation, the DAE in DAE-GP uses
-LSTMs for both encoding and decoding where the total amount of time
-steps \(T\) is equal to sum of the length of the input solution and the
-output solution {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]}.
+The AE is trained to reduce the reconstruction error between input and output data, after the training procedure is finished the network is able to reduce the dimensionality of input data to a compressed representation{[}\protect\hyperlink{ref-ae_orig}{6}{]}.
+
+DAE was first introduced by {[}\protect\hyperlink{ref-dae_orig2008}{14}{]} as an improved AE with the ability to learn new representations of data that is especially robust to partially corrupted input data.
+DAE modifies the AE by using partially corrupted input data for the AE and training it to reconstruct the uncorrupted, original version of the input data.
+
+Since the hidden representation of DAE captures the dependency structure of the input data it can therefore also be used to generate new solutions in the context of GAs {[}\protect\hyperlink{ref-harmless_overfitting_eda}{11}{]}.
+
+DAE-GP builds upon the concept of using DAEs in EDAs described by {[}\protect\hyperlink{ref-harmless_overfitting_eda}{11}{]} and transfers the concept to the domain of GP.
+The mutation and crossover operators of standard GP are replaced by sampling new solutions from a probabilistic model that is build by training a DAEs long short-term memory (LSTM) network on a subset of high fitness solutions from the current population {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]}.
+
+LSTMs are a variant of artificial neural networks first introduced by {[}\protect\hyperlink{ref-lstm_orig}{7}{]} that can store learned information over en extended time period while avoiding the problem of vanishing and/or exploding gradients.
+Since DAE-GP encodes candidate solutions as linear strings in prefix notation, the DAE in DAE-GP uses LSTMs for both encoding and decoding where the total amount of time steps \(T\) is equal to sum of the length of the input solution and the output solution {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]}.
 
@@
 
-Pre-Training describes the concept of initially training a neural
-network on a large dataset before using it to solve a more specific
-task. It is a commonly used strategy in deep architectures that has been
-shown to improve both the optimization process itself as well as the
-generalization behavior if compared to the standard approach of using
-randomly initialized parameters
-{[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{3}{]}.
-
-The strategy of pre-training artificial neural networks is based on the
-idea of transfer learning that aims at retaining previously learned
-knowledge from one task to re-use it for another task
-{[}\protect\hyperlink{ref-HAN2021225}{5}{]}
-{[}\protect\hyperlink{ref-survey_transfer_learning}{9}{]}. Transfer
-learning traditionally follows a two phase approach:
+Pre-Training describes the concept of initially training an artificial neural network on a large dataset before using it to solve a more specific task.
+It is a commonly used strategy in deep architectures that has been shown to improve both the optimization process itself as well as the generalization behavior if compared to the standard approach of using randomly initialized parameters {[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{3}{]}.
+
+The strategy of pre-training artificial neural networks is based on the idea of transfer learning that aims at retaining previously learned knowledge from one task to re-use it for another task {[}\protect\hyperlink{ref-HAN2021225}{5}{]} {[}\protect\hyperlink{ref-survey_transfer_learning}{9}{]}.
+Transfer learning traditionally follows a two phase approach:
 
@@
 
-where source task and target task are usually similar but may differ in
-their feature space and the distribution of training data
-{[}\protect\hyperlink{ref-survey_transfer_learning}{9}{]}.
-
-Some of the main motivations for using pre-training is the ability to
-reduce the need for large amounts of training data which can often be
-unavailable or too expensive to collect as well as an improved
-performance by either reducing the computational effort that is
-necessary to train a model to solve a specific task or by improving the
-models generalisation ability.
-
-\ldots describe different pre-training strategies, e.g.~few shot,
-perpetual, re-using\ldots{}
+where source task and target task are usually similar but may differ in their feature space and the distribution of training data {[}\protect\hyperlink{ref-survey_transfer_learning}{9}{]}.
+
+Some of the main motivations for using pre-training is the ability to reduce the need for large amounts of training data which can often be unavailable or too expensive to collect as well as an improved performance by either reducing the computational effort that is necessary to train a model to solve a specific task or by improving the models generalisation ability.
+
+\ldots describe different pre-training strategies, e.g.~few shot, perpetual, re-using\ldots{}
 
@@
 
-The pre-training strategy used in this thesis is applied to the DAE-LSTM
-model \(M_g\) that are used in each DAE-GP generation \(g\) for
-\(g\in{1,...,g_{max}}\). Instead of initializing and optimizing each
-model from scratch as done in previous work (e.g.
-{[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]}
-{[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]}), a separate
-DAE-LSTM network \(\hat{M}\) will be trained on a large population of
-randomly initialized solutions.\\
-The trainable parameters \(\theta_{\hat{M}}\) obtained after finishing
-the training procedure of \(\hat{M}\) are then used as the starting
-parameters for each following DAE-LSTM model \(M_g\) for
-\(g\in{1,...,g_{max}}\).
-
-The motivation for incorporating pre-training into DAE-GP is based on
-the following suspected mechanisms of improvement:
+The pre-training strategy used in this thesis is applied to the DAE-LSTM model \(M_g\) that are used in each DAE-GP generation \(g\) for \(g\in{1,...,g_{max}}\).
+Instead of initializing and optimizing each model from scratch as done in previous work (e.g.
+{[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]} {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]}), a separate DAE-LSTM network \(\hat{M}\) will be trained on a large population of randomly initialized solutions.\\
+The trainable parameters \(\theta_{\hat{M}}\) obtained after finishing the training procedure of \(\hat{M}\) are then used as the starting parameters for each following DAE-LSTM model \(M_g\) for \(g\in{1,...,g_{max}}\).
+
+The motivation for incorporating pre-training into DAE-GP is based on the following suspected mechanisms of improvement:
 
@@
 \item
-  Improve overall performance of DAE-GP by improving either run time
-  and/or solution quality
+  Improve overall performance of DAE-GP by improving either run time and/or solution quality
 \item
@@
 
-The probably most obvious motivation for using a pre-training strategy
-in DAE-GP is based on the fact, that initializing \(M_g\) with
-pre-trained weights \(\theta_{\hat{M}}\) is likely to reduce the amount
-of training epochs that are necessary until the point of training error
-convergence is reached.
-
-Early research on pre-training for DAE by
-{[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{3}{]} comes to the
-conclusion that besides adding robustness to models the strategy also
-results in improved generalization and better performing models. The
-authors describe that unsupervised pre-training behaves like a regulizer
-on DAE networks, the mean test error and it's variance are reduced for
-DAE networks that are initialized from pre-trained weights if compared
-to the same architectures that use random initialization.
-
-Another important finding by
-{[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{3}{]} is that the positive
-effect of pre-training described above is dependent on both the depth of
-the network as well as the size of layers - while increasingly larger
-networks benefit increasingly more from pre-training the final
-performance of small architectures tends to be worse with pre-training
-than with randomized weight initialization. The authors find evidence
-that pre-training DAE is especially usefull for optimizing the
-parameters of the lower layers of the network.
-
-Another motivation for introducing pre-training into DAE-GP is the
-prevalence of sampling error for small GP populations sizes.
-{[}\protect\hyperlink{ref-sampling_err_gp}{13}{]} finds that sampling
-error, non-systematic errors that are caused by observing only a small
-subset of the statistical population, is a severe problem in the domain
-of GP since the initial population may not be a representative sample of
-all possible solutions.
-{[}\protect\hyperlink{ref-sampling_err_gp}{13}{]} also introduces a
-method for calculating optimal population sizes to minimize the presence
-of sampling error. Since DAE-LSTM of DAE-GP learns the properties of
-it's training population and reuses the acquired knowledge in the
-sampling procedure, using the parameters obtained from training
-\(\hat{M}\) on a sufficiently large training population might reduce the
-need for large population sizes in the following generations of the
-search if \(\hat{M}\) already implicitly captured the properties of a
-representative sample of solutions. If this mechanism can be applied
-successfully it would benefit the performance of DAE-GP by increasing
-the population diversity (resulting in better solution quality) as well
-as by reducing the need for large population sizes which require
-computational resources.
+The probably most obvious motivation for using a pre-training strategy in DAE-GP is based on the fact, that initializing \(M_g\) with pre-trained weights \(\theta_{\hat{M}}\) is likely to reduce the amount of training epochs that are necessary until the point of training error convergence is reached.
+
+Early research on pre-training for DAE by {[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{3}{]} comes to the conclusion that besides adding robustness to models the strategy also results in improved generalization and better performing models.
+The authors describe that unsupervised pre-training behaves like a regulizer on DAE networks, the mean test error and it's variance are reduced for DAE networks that are initialized from pre-trained weights if compared to the same architectures that use random initialization.
+
+Another important finding by {[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{3}{]} is that the positive effect of pre-training is dependent on both the depth of the network as well as the size of layers - while increasingly larger networks benefit increasingly more from pre-training the final performance of small architectures tends to be worse with pre-training than with randomized weight initialization.
+The authors find evidence that pre-training DAE is especially usefull for optimizing the parameters of the lower layers of the network.
+
+Another motivation for introducing pre-training into DAE-GP is the prevalence of sampling error for small GP populations sizes.
+{[}\protect\hyperlink{ref-sampling_err_gp}{13}{]} finds that sampling error, non-systematic errors that are caused by observing only a small subset of the statistical population, is a severe problem in the domain of GP since the initial population may not be a representative sample of all possible solutions.
+{[}\protect\hyperlink{ref-sampling_err_gp}{13}{]} also introduces a method for calculating optimal population sizes to minimize the presence of sampling error.
+Since DAE-LSTM of DAE-GP learns the properties of it's training population and reuses the acquired knowledge in the sampling procedure, using the parameters obtained from training \(\hat{M}\) on a sufficiently large training population might reduce the need for large population sizes in the following generations of the search if \(\hat{M}\) already implicitly captured the properties of a representative sample of solutions.
+If this mechanism can be applied successfully it would benefit the performance of DAE-GP by increasing the population diversity (resulting in better solution quality) as well as by reducing the need for large population sizes which require computational resources.
 
@@
 
-The DAE-GP algorithm, first described by
-{[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]}, is summarized as a
-flowchart in figure §. After setting the generation counter \(g\) to
-\(0\) and creating the initial population of solutions \(P_0\), DAE-GP
-enters the main loop that checks if a termination criteria is satisfied
-(i.e.~maximum number of generations or fitness evaluations is reached).
-For each generation \(g\) the fitness of all new individuals in \(P_g\)
-is evaluated. During model building, DAE-GP selects a subset \(X_g\) of
-the current population that is used as training data for the DAE-LSTM
-model \(M_g\) which is trained to learn the properties of \(X_g\). The
-trained model \(M_g\) is then used for model sampling, partially mutated
-solutions from \(P_G\) are propagated through \(M_g\) to create a new
-population \(P_{g+1}\). This process is repeated until the termination
-criteria is met and DAE-GP returns the highest fitness solutions.
-
-\begin{figure}
-\centering
-\includegraphics{./img/flowcharts/dae-gp.png}
-\caption{Flowchart - Regular DAE-GP}
-\end{figure}
-
-The pre-training strategy implemented for all experiments in this thesis
-is visualized as another flowchart in figure §. The main difference to
-regular DAE-GP is the inclusion of an initial pre-training phase where a
-separate population \(\hat{P}\) is first initialized and then randomly
-split by half into \(\hat{P}_{train}\) and \(\hat{P}_{test}\). A
-DAE-LSTM model \(\hat{M}\) is then trained to learn the properties of
-\(\hat{P}_{train}\) using an early stopping training mode where we stop
-training as soon as the validation error for \(\hat{P}_{test}\)
-converges. After the training for \(\hat{M}\) is stopped, the current
-state of \(\hat{M}\) is frozen and the optimized trainable parameters
-\(\theta_{\hat{M}}\) are saved before terminating the pre-training phase
-to start the DAE-GP phase. During the DAE-GP phase, the only difference
-to the traditional DAE-GP algorithm described in figure § is during
-model building: We initialize \(M_g\) with \(\theta_{\hat{M}}\) before
-training it to learn the properties of \(X_g\).
+The DAE-GP algorithm, first described by {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]}, is summarized as a flowchart in \ref{fig:daegp-flowchart}.
+After setting the generation counter \(g\) to \(0\) and creating the initial population of solutions \(P_0\), DAE-GP enters the main loop that checks if a termination criteria is satisfied (i.e.~maximum number of generations or fitness evaluations is reached).
+For each generation \(g\) the fitness of all new individuals in \(P_g\) is evaluated.
+During model building, DAE-GP selects a subset \(X_g\) of the current population that is used as training data for the DAE-LSTM model \(M_g\) which is trained to learn the properties of \(X_g\).
+The trained model \(M_g\) is then used for model sampling, partially mutated solutions from \(P_G\) are propagated through \(M_g\) to create a new population \(P_{g+1}\).
+This process is repeated until the termination criteria is met and DAE-GP returns the highest fitness solutions.
+
+\begin{figure}[c]
+
+{\centering \includegraphics[width=0.5\linewidth,height=0.9\textheight]{./img/flowcharts/dae-gp}
+
+}
+
+\caption{Regular DAE-GP Flowchart}\label{fig:daegp-flowchart}
+\end{figure}
+
+The pre-training strategy implemented for all experiments in this thesis is visualized as another flowchart in figure §.
+The main difference to regular DAE-GP is the inclusion of an initial pre-training phase where a separate population \(\hat{P}\) is first initialized and then randomly split by half into \(\hat{P}_{train}\) and \(\hat{P}_{test}\).
+A DAE-LSTM model \(\hat{M}\) is then trained to learn the properties of \(\hat{P}_{train}\) using an early stopping training mode where we stop training as soon as the validation error for \(\hat{P}_{test}\) converges.
+After the training for \(\hat{M}\) is stopped, the current state of \(\hat{M}\) is frozen and the optimized trainable parameters \(\theta_{\hat{M}}\) are saved before terminating the pre-training phase to start the DAE-GP phase.
+During the DAE-GP phase, the only difference to the traditional DAE-GP algorithm described in figure § is during model building: We initialize \(M_g\) with \(\theta_{\hat{M}}\) before training it to learn the properties of \(X_g\).
 
@@
 
-One difficulty in the implementation of a pre-training strategy into
-DAE-GP has been the detemination of the number of hidden neurons in side
-the DAE-LSTMs hidden layers. The original description of DAE-GP
-{[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]} used a strategy
-where the number of hidden neurons for each hidden layer is dynamically
-set per generation to the maximum individual size inside the current
-population. For a pre-training implementation, this strategy can not be
-easily adapted since it leads to a changing number of neurons at each
-generation resulting in different dimensions of the DAE-LSTM. To allow
-for the sharing of weights and biases from the pre-trained model
-\(\hat{M}\) to each \(M_g\) for \(g\in{1,...,g_{max}}\) I experimented
-with two different strategies:
+One difficulty in the implementation of a pre-training strategy into DAE-GP has been the detemination of the number of hidden neurons in side the DAE-LSTMs hidden layers.
+The original description of DAE-GP {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]} used a strategy where the number of hidden neurons for each hidden layer is dynamically set per generation to the maximum individual size inside the current population.
+For a pre-training implementation, this strategy can not be easily adapted since it leads to a changing number of neurons at each generation resulting in different dimensions of the DAE-LSTM.
+To allow for the sharing of weights and biases from the pre-trained model \(\hat{M}\) to each \(M_g\) for \(g\in{1,...,g_{max}}\) I experimented with two different strategies:
 
@@
 \item
-  Set the number of hidden neurons per hidden layer for each generation
-  to the maximum individual size inside the pre-training population
+  Set the number of hidden neurons per hidden layer for each generation to the maximum individual size inside the pre-training population
 \item
@@
 
-The framework that was used to conduct all experiments of this thesis
-was provided by the supervisor David Wittenberg and uses the python
-programming language in conjunction with the baseline libraries
-\texttt{Keras} {[}\protect\hyperlink{ref-chollet2015keras}{1}{]} for
-deep neural networks and \texttt{deap}
-{[}\protect\hyperlink{ref-DEAP_JMLR2012}{4}{]} for evolutionary
-computation.
+The framework that was used to conduct all experiments of this thesis was provided by the supervisor David Wittenberg and uses the python programming language in conjunction with the baseline libraries \texttt{Keras} {[}\protect\hyperlink{ref-chollet2015keras}{1}{]} for deep neural networks and \texttt{deap} {[}\protect\hyperlink{ref-DEAP_JMLR2012}{4}{]} for evolutionary computation.
 
@@
 
-To test pre-training in DAE-GP this thesis focuses on the domain of
-real-world symbolic regression problems. Symbolic Regression problems
-have been one of the first GP applications
-{[}\protect\hyperlink{ref-Koza1993GeneticP}{8}{]} and are an actively
-studied and highly relevant research area. The goal in symbolic
-regression is to find a mathematical model for a given set of data
-points
-{[}\protect\hyperlink{ref-10.1007ux2f978-3-540-24621-3_22}{10}{]}, in
-real-world symbolic regression these data points are sourced from
-real-world observations which in contrast to synthetic symbolic
-regression problem are more likely to contain random noise and bias.
-Another important challenge in solving real-world symbolic regression
-problems is the ability for a given model to generalize, we want the
-final model to show high accuracy in predicting outcomes for previously
-unseen cases.
-
-The main experiments conducted in this thesis uses the NASA Airfoil
-Self-Noise Data Set which is part of the UCI machine learning repository
-{[}\protect\hyperlink{ref-machine_learning_repo}{2}{]}. The dataset
-consists of 5 input variables and 1 output variable that are listed in
-table 1.
+To test pre-training in DAE-GP this thesis focuses on the domain of real-world symbolic regression problems.
+Symbolic Regression problems have been one of the first GP applications {[}\protect\hyperlink{ref-Koza1993GeneticP}{8}{]} and are an actively studied and highly relevant research area.
+The goal in symbolic regression is to find a mathematical model for a given set of data points {[}\protect\hyperlink{ref-10.1007ux2f978-3-540-24621-3_22}{10}{]}, in real-world symbolic regression these data points are sourced from real-world observations which in contrast to synthetic symbolic regression problem are more likely to contain random noise and bias.
+Another important challenge in solving real-world symbolic regression problems is the ability for a given model to generalize, we want the final model to show high accuracy in predicting outcomes for previously unseen cases.
+
+The main experiments conducted in this thesis uses the NASA Airfoil Self-Noise Data Set which is part of the UCI machine learning repository {[}\protect\hyperlink{ref-machine_learning_repo}{2}{]}.
+The dataset consists of 5 input variables and 1 output variable that are listed in table 1.
 
@@
 
-\caption{\label{tab:airfoil_dataset_description}Airfoil - Dataset Description}
+\caption{\label{tab:airfoil-dataset-description}Airfoil - Dataset Description}
 \centering
@@
 
-The objective of the airfoil problem is to find a function that
-accurately predicts the output variable \(y\) by taking in a subset of
-the input variables \(x1,x2,x3,x4,x5\). The function set used by all
-DAE-GP variations for the airfoil problem is summarized in table 2, the
-terminal set consists of the 5 input variables \(x1,x2,x3,x4,x5\) and
-ephemeral random integers in the range of \([{}-5,..,5]\)
-{[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]}.
+The objective of the airfoil problem is to find a function that accurately predicts the output variable \(y\) by taking in a subset of the input variables \(x1,x2,x3,x4,x5\).
+The function set used by all DAE-GP variations for the airfoil problem is summarized in table 2, the terminal set consists of the 5 input variables \(x1,x2,x3,x4,x5\) and ephemeral random integers in the range of \([{}-5,..,5]\) {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]}.
 
@@
 
-\caption{\label{tab:airfoil_function_set}Airfoil - Function Set}
+\caption{\label{tab:airfoil-function-set}Airfoil - Function Set}
 \centering
@@
 \hline
-\textbf{function.} & \textbf{arity}\\
+\textbf{operator} & \textbf{arity}\\
 \hline
@@
 
-To gather a deeper understanding about the effect of pre-training
-DAE-LSTM in DAE-GP a series of experiments was conducted using the
-airfoil dataset for symbolic regression while using and different
-parameter configurations for the number of hidden layers as well as the
-number of hidden neurons per hidden layer.
-
-To study generalization behavior this series of experiments is conducted
-using DAE-GP with only a single generation until the search terminates.
-I disregard the fitness of solutions found and focus soley on the
-reconstruction error that is produced during the training of each the
-DAE-LSTM. The reconstruction error is measured for two separate
-populations, a training population \(P_{train}\) that is used to train
-our DAE-LSTM as well as a hold-out validation population \(P_{test}\).
-For the pre-trained DAE-GP two additional, separate populations
-\(\hat{P}_{train}\) and \(\hat{P}_{test}\) are used exclusively for
-pre-training.
+To gather a deeper understanding about the effect of pre-training DAE-LSTM in DAE-GP a series of experiments was conducted using the airfoil dataset for symbolic regression while using and different parameter configurations for the number of hidden layers as well as the number of hidden neurons per hidden layer.
+
+To study generalization behavior this series of experiments is conducted using DAE-GP with only a single generation until the search terminates.
+I disregard the fitness of solutions found and focus soley on the reconstruction error that is produced during the training of each the DAE-LSTM.
+The reconstruction error is measured for two separate populations, a training population \(P_{train}\) that is used to train our DAE-LSTM as well as a hold-out validation population \(P_{test}\).
+For the pre-trained DAE-GP two additional, separate populations \(\hat{P}_{train}\) and \(\hat{P}_{test}\) are used exclusively for pre-training.
 
@@
 \item
-  Variable number of hidden neurons (50, 100, 200) with a single hidden
-  layer
-\item
-  Fixed number of hidden neurons (100) per hidden layer with variable
-  number of hidden layers (1, 2, 3)
+  Variable number of hidden neurons (50, 100, 200) with a single hidden layer
+\item
+  Fixed number of hidden neurons (100) per hidden layer with variable number of hidden layers (1, 2, 3)
 \end{itemize}
 
-For each configuration I tested traditional DAE-GP as well as a
-pre-trained DAE-GP resulting in 12 total sub experiments that were each
-based on 10 individual runs (total number of runs=\(120\)). To avoid
-creating biased results through the presence of sampling error inside
-the pre-training population (see
-{[}\protect\hyperlink{ref-sampling_err_gp}{13}{]}), the population size
-for the pre-training phase is chosen very high with the size of
-\(\hat{P}\) =20000 where 50\% of \(\hat{P}\) is used for the training
-population \(\hat{P}_{train}\) and the remaining 50\% are used for the
-hold-out validation population \(\hat{P}_{test}\). The training of
-DAE-LSTM uses a fixed number of 1000 epochs until termination and uses
-Adam optimization for gradient descent. The reason for using a high
-amount of 1000 fixed training epochs is to deliberately force the
-DAE-LSTM to overfit to the training data.
-
-In general I expect that with a growing number of model parameters
-(either by adding more hidden layers or more hidden neurons per layer)
-the DAE-LSTM will be more prone to overfit to the training population
-\(P_{train}\) resulting in a small reconstruction error for the training
-population and a large one for the validation population \(P_{test}\).
-The reason for this effect is that a larger network trained over an
-extended period of time (without the use of strategies like early
-stopping), has much more potential to learn noise from the training
-dataset than a smaller network, which is more likely to result in
-worsening performance for previously unseen cases
-{[}\protect\hyperlink{ref-weigend1994overfitting}{15}{]}.
-
-Based on the review of {[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{3}{]}
-I also expect that pre-training will have an insignificant or even
-negative influence on small DAE-LSTM instances while improving the
-networks generalization ability with growing size.
+For each configuration I tested traditional DAE-GP as well as a pre-trained DAE-GP resulting in 12 total sub experiments that were each based on 10 individual runs (total number of runs=\(120\)).
+To avoid creating biased results through the presence of sampling error inside the pre-training population (see {[}\protect\hyperlink{ref-sampling_err_gp}{13}{]}), the population size for the pre-training phase is chosen very high with the size of \(\hat{P}\) =20000 where 50\% of \(\hat{P}\) is used for the training population \(\hat{P}_{train}\) and the remaining 50\% are used for the hold-out validation population \(\hat{P}_{test}\).
+The training of DAE-LSTM uses a fixed number of 1000 epochs until termination and uses Adam optimization for gradient descent.
+The reason for using a high amount of 1000 fixed training epochs is to deliberately force the DAE-LSTM to overfit to the training data.
+
+In general I expect that with a growing number of model parameters (either by adding more hidden layers or more hidden neurons per layer) the DAE-LSTM will be more prone to overfit to the training population \(P_{train}\) resulting in a small reconstruction error for the training population and a large one for the validation population \(P_{test}\).
+The reason for this effect is that a larger network trained over an extended period of time (without the use of strategies like early stopping), has much more potential to learn noise from the training dataset than a smaller network, which is more likely to result in worsening performance for previously unseen cases {[}\protect\hyperlink{ref-weigend1994overfitting}{15}{]}.
+
+Based on the review of {[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{3}{]} I also expect that pre-training will have an insignificant or even negative influence on small DAE-LSTM instances while improving the networks generalization ability with growing size.
 
@@
 \includegraphics{./img/airfoil_firstGen/airfoil_firstGen_mean_training_error_by_neurons_6plots.png}
-\caption{First Generation Median Training Error for variable number of
-hidden Neurons - Airfoil}
+\caption{First Generation Median Training Error for variable number of hidden Neurons - Airfoil}
 \end{figure}
@@
 \includegraphics{./img/airfoil_firstGen/airfoil_firstGen_mean_training_error_by_layers_6plot.png}
-\caption{First Generation Median Training Error for variable number of
-hidden Layers - Airfoil}
+\caption{First Generation Median Training Error for variable number of hidden Layers - Airfoil}
 \end{figure}
@@
 \hypertarget{effect-on-search-performance}{%
-\subsection{Effect on search
-performance}\label{effect-on-search-performance}}
-
-After closely examining the effect of pre-training on DAE-GPs
-generalization behaviour another series of experiments is conducted to
-study how pre-training influences the overall search behaviour of
-DAE-GP. Again the Airfoil problem was selected as a first benchmark
-problem to test pre-trained DAE-GP against traditional DAE-GP
-{[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]}, the parameters are
-described in table \(A\)
+\subsection{Effect on search performance}\label{effect-on-search-performance}}
+
+After closely examining the effect of pre-training on DAE-GPs generalization behaviour another series of experiments is conducted to study how pre-training influences the overall search behaviour of DAE-GP.
+Again the Airfoil problem was selected as a first benchmark problem to test pre-trained DAE-GP against traditional DAE-GP {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]}, the parameters are described in table \(A\)
 
@@
 
-\caption{\label{tab:airfoil_fullRun_2hl_maxIndSize_params}Airfoil - Parameter for full Run}
+\caption{\label{tab:airfoil-fullRun-2hl-maxIndSize-params}Airfoil - Parameter for full Run}
 \centering
@@
 \hline
-\textbf{parameter} & \textbf{value}\\
-\hline
-populationSize & 500\\
+\textbf{Hyperparameter} & \textbf{Value}\\
+\hline
+popSize & 500\\
 \hline
@@
 \hline
-Pre-Training PopulationSize(Training/Validation) & 10000/100000\\
+corruptionTechnique & Levenshtein Edit\\
+\hline
+edit\_prob(training/sampling) & 0.05/0.95\\
+\hline
+functionSet & \{+ ; - ; * ; aq\}\\
+\hline
+Pre-Training PopulationSize & 10000\\
+\hline
+Pre-Training Train/Test Split & 50\%\\
 \hline
@@
 
-Regarding the hidden neurons, the dimension of each hidden layer, I used
-two different strategies: Regular DAE-GP uses the same strategy as
-described by earlier work (see
-{[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]} or
-{[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]}), the number of
-hidden neurons is dynamically set at each generation to the maximum
-length of all solutions inside the current training population. For
-pre-trained DAE-GP another strategy had to be used, I used a fixed
-number of hidden neurons for each run that is set equal to the maximum
-length of all individuals that are present in the initial pre-training
-population. I decided to take this aproach instead of using a fixed
-number of hidden neurons for the fact that it simplifies the adjustment
-of the parameter for later experiments on different problems. It should
-nonetheless be mentioned that this strategy does give pre-training on
-average more hidden neurons per layer resulting in a more complex
-network (since the pre-training population is much larger than the
-regular population, the chances for larger individuals inside it are
-also higher).
+Regarding the hidden neurons, the dimension of each hidden layer, I used two different strategies: Regular DAE-GP uses the same strategy as described by earlier work (see {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]} or {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]}), the number of hidden neurons is dynamically set at each generation to the maximum length of all solutions inside the current training population.
+For pre-trained DAE-GP another strategy had to be used, I used a fixed number of hidden neurons for each run that is set equal to the maximum length of all individuals that are present in the initial pre-training population.
+I decided to take this aproach instead of using a fixed number of hidden neurons for the fact that it simplifies the adjustment of the parameter for later experiments on different problems.
+It should nonetheless be mentioned that this strategy does give pre-training on average more hidden neurons per layer resulting in a more complex network (since the pre-training population is much larger than the regular population, the chances for larger individuals inside it are also higher).
 
@@
 
-\includegraphics{./img/airfoil_FullRun_30gens/mean_popDiversity_byGens.png}
-Since {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]} shows, that
-DAE-GP finds higher quality that are also smaller in size than those
-found by average GP for a given number of 10.000 fitness evaluations, I
-also studied the differences in average solution size between
-pre-trained and regular DAE-GP.
+\includegraphics{./img/airfoil_2hl_maxIndSize_fullRun_30gens/mean_popDiversity_byGens.png}
+Since {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]} shows, that DAE-GP finds higher quality that are also smaller in size than those found by average GP for a given number of 10.000 fitness evaluations, I also studied the differences in average solution size between pre-trained and regular DAE-GP.
 
@@
 \centering
-\includegraphics{./img/airfoil_FullRun_30gens/mean_Size_byGens.png}
+\includegraphics{./img/airfoil_2hl_maxIndSize_fullRun_30gens/mean_Size_byGens.png}
 \caption{Average Solution Size over 30 Generations - Airfoil}
@@
 
-The experiments was repeated using the same setup with the only
-difference being a reduction from two hidden layers to 1 hidden layer.
+The experiments was repeated using the same setup with the only difference being a reduction from two hidden layers to 1 hidden layer.
 
@@
 
-To gather more confidence in the previous results, the pre-training
-strategy for DAE-GP was tested for various real world symbolic
-regression problems taken from
-{[}\protect\hyperlink{ref-machine_learning_repo}{2}{]}. Are brief
-summary of the benchmark problems is given in table X
+To gather more confidence in the previous results, the pre-training strategy for DAE-GP was tested for various real world symbolic regression problems taken from {[}\protect\hyperlink{ref-machine_learning_repo}{2}{]}.
+Are brief summary of the benchmark problems is given in table X
 
@@
 
-\caption{\label{tab:full_run_realWorldSymReg_problems}Overview - Real World Symbolic Regression Benchmark Problems}
+\caption{\label{tab:full-run-realWorldSymReg-problems}Overview - Real World Symbolic Regression Benchmark Problems}
 \centering
@@
 
-Table X summarizes the median best fitness achieved after running each
-algorithm for 10 runs using a population size of 500 individuals for 30
-generations.\footnote{P-Values are marked with asteriks (1,2,3) for
-  (\(\alpha\))-levels of \(0.1\), \(0.05\) and \(0.01\)}.
+Table X summarizes the median best fitness achieved after running each algorithm for 10 runs using a population size of 500 individuals for 30 generations.\footnote{P-Values are marked with asteriks (1,2,3) for (\(\alpha\))-levels of \(0.1\), \(0.05\) and \(0.01\)}.
+The median best fitness over all 30 generations are visualized in figure X.
 
 \begin{longtable}[]{@{}
-  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1910}}
-  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1685}}
-  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.0787}}
-  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1461}}
-  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1461}}
-  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1124}}
-  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1573}}@{}}
-\toprule()
+  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1932}}
+  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1705}}
+  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.0795}}
+  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1477}}
+  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1477}}
+  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1023}}
+  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1591}}@{}}
+\caption{\label{tab:full-run-realWorldSymReg-fitness}Median Best Fitness - Symbolic Regression}\tabularnewline
+\toprule\noalign{}
 \begin{minipage}[b]{\linewidth}\centering
@@
 \end{minipage} \\
-\midrule()
-\endhead
-Airfoil & 1 & Train & \textbf{33.3674} & 35.3454 & 0.0171** & 0.64 \\
-& 1 & Test & \textbf{33.7271} & 35.612 & 0.1400 & 0.4 \\
-Airfoil & 2 & Train & 17.4762 & \textbf{14.7233} & 0.3073 & -0.28 \\
-& 2 & Test & 17.5299 & \textbf{14.7603} & 0.5706 & -0.16 \\
-Boston\_Housing & 2 & Train & 8.1922 & \textbf{8.0763} & 0.2894 &
--0.29 \\
-& 2 & Test & 8.0501 & \textbf{7.9692} & 0.7051 & -0.11 \\
-Energy(Cooling) & 2 & Train & \textbf{4.5271} & 4.7782 & 0.0188** &
-0.63 \\
-& 2 & Test & \textbf{4.67} & 4.9196 & 0.2886 & 0.29 \\
-Concrete & 2 & Train & \textbf{17.1075} & 17.528 & 1.0000 & 0.0 \\
-& 2 & Test & \textbf{15.7978} & 16.7196 & 0.6667 & 1.0 \\
-\bottomrule()
-\end{longtable}
-
-\hypertarget{solution-size-in-symbolic-regression}{%
-\subsection{Solution Size in Symbolic
-Regression}\label{solution-size-in-symbolic-regression}}
-
-\begin{longtable}[]{@{}
-  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2152}}
-  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1519}}
-  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1646}}
-  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1646}}
-  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1266}}
-  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1772}}@{}}
-\toprule()
+\midrule\noalign{}
+\endfirsthead
+\toprule\noalign{}
 \begin{minipage}[b]{\linewidth}\centering
@@
 \end{minipage} & \begin{minipage}[b]{\linewidth}\centering
-Hid.Layers
+Hidden-Layers
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+Set
 \end{minipage} & \begin{minipage}[b]{\linewidth}\centering
@@
 \end{minipage} \\
-\midrule()
+\midrule\noalign{}
 \endhead
-Airfoil & 1 & \textbf{33.3674} & 35.3454 & 0.0171** & 0.64 \\
-Airfoil & 2 & 17.4762 & \textbf{14.7233} & 0.3073 & -0.28 \\
-Boston\_Housing & 2 & 8.1922 & \textbf{8.0763} & 0.2894 & -0.29 \\
-Energy(Cooling) & 2 & \textbf{4.5271} & 4.7782 & 0.0188** & 0.63 \\
-Concrete & 2 & \textbf{17.1075} & 17.528 & 1.0000 & 0.0 \\
-\bottomrule()
+\bottomrule\noalign{}
+\endlastfoot
+Airfoil & 1 & Train & \textbf{33.3674} & 35.3454 & 0.02** & 0.64 \\
+& 1 & Test & \textbf{33.7271} & 35.612 & 0.14 & 0.40 \\
+Airfoil & 2 & Train & 17.4762 & \textbf{14.7233} & 0.31 & -0.28 \\
+& 2 & Test & 17.5299 & \textbf{14.7603} & 0.57 & -0.16 \\
+Boston\_Housing & 2 & Train & 8.1922 & \textbf{8.0763} & 0.29 & -0.29 \\
+& 2 & Test & 8.0501 & \textbf{7.9692} & 0.71 & -0.11 \\
+Energy(Cooling) & 2 & Train & \textbf{4.5271} & 4.7782 & 0.02** & 0.63 \\
+& 2 & Test & \textbf{4.67} & 4.9196 & 0.29 & 0.29 \\
+Concrete & 2 & Train & 17.0052 & \textbf{16.9085} & 0.97 & 0.02 \\
+& 2 & Test & \textbf{16.5845} & 16.9298 & 0.52 & 0.18 \\
 \end{longtable}
 
-\begin{tabular}{cc}
-\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/energyCooling_2hl_maxIndSize_fullRun_30gens/mean_Size_byGens.png} &
-\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/bostonHousing_2hl_maxIndSize_fullRun_30gens/mean_Size_byGens.png} \\
-\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/airfoil_2hl_maxIndSize_fullRun_30gens/mean_Size_byGens.png} &
-\includegraphics[width=0.45\textwidth]{~/masterThesis/master_thesis/img/concrete_2hl_maxIndSize_fullRun_30gens/mean_Size_byGens.png}
-\end{tabular}
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth,height=0.8\textheight]{~/masterThesis/master_thesis/img/symbolicRegressionSummarized/median_fitness_byGens.png}
+\caption{Median Best Fitness - Symbolic Regression}
+\end{figure}
+
+\hypertarget{solution-size-in-symbolic-regression}{%
+\subsection{Solution Size in Symbolic Regression}\label{solution-size-in-symbolic-regression}}
+
+Next I studied the size of solutions based on using a pre-training strategy. Figure X shows the median size of the best solutions aggregated of all benchmark problems for symbolic regression
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth,height=0.8\textheight]{~/masterThesis/master_thesis/img/symbolicRegressionSummarized/median_bestSol_size_byGens.png}
+\caption{Median Solution Size - Symbolic Regression}
+\end{figure}
+
+Table X summarizes the results of the statistical tests.
+
+\begin{longtable}[]{@{}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2254}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1549}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1549}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1690}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1127}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1831}}@{}}
+\caption{\label{tab:full-run-realWorldSymReg-best-solution-size}Median Solution Size - Symbolic Regression}\tabularnewline
+\toprule\noalign{}
+\begin{minipage}[b]{\linewidth}\raggedright
+Problem
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Hid.Layers
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+DAE-GP
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Pre-Trained
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+P-Value
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Cliffs-Delta
+\end{minipage} \\
+\midrule\noalign{}
+\endfirsthead
+\toprule\noalign{}
+\begin{minipage}[b]{\linewidth}\raggedright
+Problem
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Hid.Layers
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+DAE-GP
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Pre-Trained
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+P-Value
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Cliffs-Delta
+\end{minipage} \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+Airfoil & 1 & 12 & \textbf{10.6} & 0.46 & -0.20 \\
+Airfoil & 2 & 17.4762 & \textbf{14.7233} & 0.31 & -0.28 \\
+Boston\_Housing & 2 & 8.1922 & \textbf{8.0763} & 0.29 & -0.29 \\
+Energy(Cooling) & 2 & \textbf{4.5271} & 4.7782 & 0.02** & 0.63 \\
+Concrete & 2 & 17.0052 & \textbf{16.9085} & 0.97 & 0.02 \\
+\end{longtable}
+
+\hypertarget{population-diversity}{%
+\subsection{Population Diversity}\label{population-diversity}}
+
+Another interesting metric to examine in EC based metaheuristics is the diversity of the current generations population.
+Since the individual solutions in DAE-GP are computer programs in the form ob parse trees, I selected the normalized Levenshtein edit distance as my primary metric for tracking population diversity.
+figure X shows how the median population diversity develops over the full number of generations for all benchmark problems.
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth,height=0.8\textheight]{~/masterThesis/master_thesis/img/symbolicRegressionSummarized/median_levDistance_byGen.png}
+\caption{Median Population Diversity over Generations}
+\end{figure}
+
+Testing for statistical differences in the underlying distributions yields the P-Values listed in table X.
+
+\begin{longtable}[]{@{}
+  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2267}}
+  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1600}}
+  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1333}}
+  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1733}}
+  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1200}}
+  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1867}}@{}}
+\caption{\label{tab:full-run-realWorldSymReg-popDiversity}Median Population Diversity over Generations - Symbolic Regression}\tabularnewline
+\toprule\noalign{}
+\begin{minipage}[b]{\linewidth}\centering
+Problem
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+Hid.Layers
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+DAE-GP
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+Pre-Trained
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+P-Value
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+Cliffs-Delta
+\end{minipage} \\
+\midrule\noalign{}
+\endfirsthead
+\toprule\noalign{}
+\begin{minipage}[b]{\linewidth}\centering
+Problem
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+Hid.Layers
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+DAE-GP
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+Pre-Trained
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+P-Value
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+Cliffs-Delta
+\end{minipage} \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+Airfoil & 1 & \textbf{0.34} & 0.49 & 0.02** & 0.36 \\
+Airfoil & 2 & \textbf{0.91} & 0.93 & 0.00*** & 0.88 \\
+Boston\_Housing & 2 & \textbf{0.94} & 0.95 & 0.00*** & 0.82 \\
+Energy(Cooling) & 2 & \textbf{0.93} & 0.94 & 0.00*** & 0.80 \\
+Concrete & 2 & \textbf{0.93} & 0.94 & 0.00*** & 0.88 \\
+\end{longtable}
+
+\hypertarget{training-epochs-per-generation}{%
+\subsection{Training epochs per Generation}\label{training-epochs-per-generation}}
+
+One important argument for using pre-training is the reduction of runtime and computational ressources for DAE-GP.
+An important factor in this, is the number of epochs that each DAE-LSTM has to be trained per generation.
+I studied the median number of training epochs that each generations DAE-LSTM \(M_g\) (excluding the pre-training DAE-LSTM model \(\hat{M}\)) was trained until the training was stopped.
+The results are visualized in figure X and summarized in table X
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth,height=0.8\textheight]{~/masterThesis/master_thesis/img/symbolicRegressionSummarized/median_epochsTrained.png}
+\caption{Median Number of Training Epochs per Generation - Symbolic Regression}
+\end{figure}
+
+\begin{longtable}[]{@{}
+  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2222}}
+  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}
+  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1111}}
+  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1806}}
+  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1250}}
+  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1944}}@{}}
+\caption{\label{tab:full-run-realWorldSymReg-epochsPerGen}Median Number of Training Epochs per Generation - Symbolic Regression}\tabularnewline
+\toprule\noalign{}
+\begin{minipage}[b]{\linewidth}\centering
+Problem
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+Hid.Layers
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+DAE-GP
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+Pre-Trained
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+P-Value
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+Cliffs-Delta
+\end{minipage} \\
+\midrule\noalign{}
+\endfirsthead
+\toprule\noalign{}
+\begin{minipage}[b]{\linewidth}\centering
+Problem
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+Hid.Layers
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+DAE-GP
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+Pre-Trained
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+P-Value
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+Cliffs-Delta
+\end{minipage} \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+Airfoil & 1 & 190.93 & \textbf{60} & 0.00*** & -0.94 \\
+Airfoil & 2 & 151.87 & \textbf{64.95} & 0.00*** & -0.88 \\
+Boston\_Housing & 2 & 182.84 & \textbf{67.27} & 0.00*** & -0.88 \\
+Energy\_Cooling & 2 & 169.39 & \textbf{69.87} & 0.00*** & -0.90 \\
+concrete & 2 & 171.69 & \textbf{64.9} & 0.00*** & -0.92 \\
+\end{longtable}
 
@@
 \hypertarget{limitations-and-open-questions}{%
-\section{Limitations and open
-Questions}\label{limitations-and-open-questions}}
+\section{Limitations and open Questions}\label{limitations-and-open-questions}}
 
@@
 
-{[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{3}{]}: Pre-Training effect
-especially usefull for lower-level layers -\textgreater{} Only adapt
-weights from low layers or embedd the pre-trained model?
+{[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{3}{]}: Pre-Training effect especially usefull for lower-level layers -\textgreater{} Only adapt weights from low layers or embedd the pre-trained model?
 
@@
 \CSLLeftMargin{{[}1{]} }%
-\CSLRightInline{Francois Chollet and others. 2015. Keras. Retrieved from
-\url{https://github.com/fchollet/keras}}
+\CSLRightInline{Francois Chollet and others. 2015. Keras. Retrieved from \url{https://github.com/fchollet/keras}}
 
@@
 \CSLLeftMargin{{[}2{]} }%
-\CSLRightInline{Dheeru Dua and Casey Graff. 2017. {UCI} machine learning
-repository. Retrieved from \url{http://archive.ics.uci.edu/ml}}
+\CSLRightInline{Dheeru Dua and Casey Graff. 2017. {UCI} machine learning repository. Retrieved from \url{http://archive.ics.uci.edu/ml}}
 
@@
 \CSLLeftMargin{{[}3{]} }%
-\CSLRightInline{Dumitru Erhan, Pierre-Antoine Manzagol, Yoshua Bengio,
-Samy Bengio, and Pascal Vincent. 2009. The difficulty of training deep
-architectures and the effect of unsupervised pre-training. In
-\emph{Proceedings of the twelth international conference on artificial
-intelligence and statistics} (Proceedings of machine learning research),
-PMLR, Hilton Clearwater Beach Resort, Clearwater Beach, Florida USA,
-153--160. Retrieved from
-\url{https://proceedings.mlr.press/v5/erhan09a.html}}
+\CSLRightInline{Dumitru Erhan, Pierre-Antoine Manzagol, Yoshua Bengio, Samy Bengio, and Pascal Vincent.
+2009.
+The difficulty of training deep architectures and the effect of unsupervised pre-training.
+In \emph{Proceedings of the twelth international conference on artificial intelligence and statistics} (Proceedings of machine learning research), PMLR, Hilton Clearwater Beach Resort, Clearwater Beach, Florida USA, 153--160.
+Retrieved from \url{https://proceedings.mlr.press/v5/erhan09a.html}}
 
@@
 \CSLLeftMargin{{[}4{]} }%
-\CSLRightInline{Félix-Antoine Fortin, François-Michel De Rainville,
-Marc-André Gardner, Marc Parizeau, and Christian Gagné. 2012. {DEAP}:
-Evolutionary algorithms made easy. \emph{Journal of Machine Learning
-Research} 13, (July 2012), 2171--2175.}
+\CSLRightInline{Félix-Antoine Fortin, François-Michel De Rainville, Marc-André Gardner, Marc Parizeau, and Christian Gagné.
+2012.
+{DEAP}: Evolutionary algorithms made easy.
+\emph{Journal of Machine Learning Research} 13, (July 2012), 2171--2175.}
 
@@
 \CSLLeftMargin{{[}5{]} }%
-\CSLRightInline{Xu Han, Zhengyan Zhang, Ning Ding, Yuxian Gu, Xiao Liu,
-Yuqi Huo, Jiezhong Qiu, Yuan Yao, Ao Zhang, Liang Zhang, Wentao Han,
-Minlie Huang, Qin Jin, Yanyan Lan, Yang Liu, Zhiyuan Liu, Zhiwu Lu,
-Xipeng Qiu, Ruihua Song, Jie Tang, Ji-Rong Wen, Jinhui Yuan, Wayne Xin
-Zhao, and Jun Zhu. 2021. Pre-trained models: Past, present and future.
+\CSLRightInline{Xu Han, Zhengyan Zhang, Ning Ding, Yuxian Gu, Xiao Liu, Yuqi Huo, Jiezhong Qiu, Yuan Yao, Ao Zhang, Liang Zhang, Wentao Han, Minlie Huang, Qin Jin, Yanyan Lan, Yang Liu, Zhiyuan Liu, Zhiwu Lu, Xipeng Qiu, Ruihua Song, Jie Tang, Ji-Rong Wen, Jinhui Yuan, Wayne Xin Zhao, and Jun Zhu.
+2021.
+Pre-trained models: Past, present and future.
 \emph{AI Open} 2, (2021), 225--250.
@@
 \CSLLeftMargin{{[}6{]} }%
-\CSLRightInline{G. E. Hinton and R. R. Salakhutdinov. 2006. Reducing the
-dimensionality of data with neural networks. \emph{Science} 313, 5786
-(2006), 504--507.
+\CSLRightInline{G.
+E.
+Hinton and R.
+R.
+Salakhutdinov.
+2006.
+Reducing the dimensionality of data with neural networks.
+\emph{Science} 313, 5786 (2006), 504--507.
 DOI:https://doi.org/\href{https://doi.org/10.1126/science.1127647}{10.1126/science.1127647}}
@@
 \CSLLeftMargin{{[}7{]} }%
-\CSLRightInline{Sepp Hochreiter and Jürgen Schmidhuber. 1997. Long
-short-term memory. \emph{Neural computation} 9, (December 1997),
-1735--80.
+\CSLRightInline{Sepp Hochreiter and Jürgen Schmidhuber.
+1997.
+Long short-term memory.
+\emph{Neural computation} 9, (December 1997), 1735--80.
 DOI:https://doi.org/\href{https://doi.org/10.1162/neco.1997.9.8.1735}{10.1162/neco.1997.9.8.1735}}
@@
 \CSLLeftMargin{{[}8{]} }%
-\CSLRightInline{John R. Koza. 1993. Genetic programming - on the
-programming of computers by means of natural selection. In \emph{Complex
-adaptive systems}.}
+\CSLRightInline{John R. Koza. 1993. Genetic programming - on the programming of computers by means of natural selection. In \emph{Complex adaptive systems}.}
 
@@
 \CSLLeftMargin{{[}9{]} }%
-\CSLRightInline{Sinno Jialin Pan and Qiang Yang. 2010. A survey on
-transfer learning. \emph{IEEE Transactions on Knowledge and Data
-Engineering} 22, 10 (2010), 1345--1359.
+\CSLRightInline{Sinno Jialin Pan and Qiang Yang.
+2010.
+A survey on transfer learning.
+\emph{IEEE Transactions on Knowledge and Data Engineering} 22, 10 (2010), 1345--1359.
 DOI:https://doi.org/\href{https://doi.org/10.1109/TKDE.2009.191}{10.1109/TKDE.2009.191}}
@@
 \CSLRightInline{Grégory Paris, Denis Robilliard, and Cyril Fonlupt.
-2004. Exploring overfitting in genetic programming. In \emph{Artificial
-evolution}, Springer Berlin Heidelberg, Berlin, Heidelberg, 267--277.}
+2004.
+Exploring overfitting in genetic programming.
+In \emph{Artificial evolution}, Springer Berlin Heidelberg, Berlin, Heidelberg, 267--277.}
 
@@
 \CSLLeftMargin{{[}11{]} }%
-\CSLRightInline{Malte Probst and Franz Rothlauf. 2020. Harmless
-overfitting: Using denoising autoencoders in estimation of distribution
-algorithms. \emph{Journal of Machine Learning Research} 21, 78 (2020),
-1--31. Retrieved from \url{http://jmlr.org/papers/v21/16-543.html}}
+\CSLRightInline{Malte Probst and Franz Rothlauf.
+2020.
+Harmless overfitting: Using denoising autoencoders in estimation of distribution algorithms.
+\emph{Journal of Machine Learning Research} 21, 78 (2020), 1--31.
+Retrieved from \url{http://jmlr.org/papers/v21/16-543.html}}
 
@@
 \CSLLeftMargin{{[}12{]} }%
-\CSLRightInline{Franz Rothlauf. 2011. \emph{Design of modern heuristics:
-Principles and application}.
-DOI:https://doi.org/\href{https://doi.org/10.1007/978-3-540-72962-4}{10.1007/978-3-540-72962-4}}
+\CSLRightInline{Franz Rothlauf. 2011. \emph{Design of modern heuristics: Principles and application}. DOI:https://doi.org/\href{https://doi.org/10.1007/978-3-540-72962-4}{10.1007/978-3-540-72962-4}}
 
@@
 \CSLRightInline{Dirk Schweim, David Wittenberg, and Franz Rothlauf.
-2021. On sampling error in genetic programming. \emph{Natural computing}
-2021, (2021).
+2021.
+On sampling error in genetic programming.
+\emph{Natural computing} 2021, (2021).
 DOI:https://doi.org/\url{http://doi.org/10.25358/openscience-5820}}
 
-\leavevmode\vadjust pre{\hypertarget{ref-10.1145ux2f1390156.1390294}{}}%
+\leavevmode\vadjust pre{\hypertarget{ref-dae_orig2008}{}}%
 \CSLLeftMargin{{[}14{]} }%
-\CSLRightInline{Pascal Vincent, Hugo Larochelle, Yoshua Bengio, and
-Pierre-Antoine Manzagol. 2008. Extracting and composing robust features
-with denoising autoencoders. In \emph{Proceedings of the 25th
-international conference on machine learning} (ICML '08), Association
-for Computing Machinery, New York, NY, USA, 1096--1103.
+\CSLRightInline{Pascal Vincent, Hugo Larochelle, Y.
+Bengio, and Pierre-Antoine Manzagol.
+2008.
+Extracting and composing robust features with denoising autoencoders.
+In \emph{Proceedings of the 25th international conference on machine learning} (ICML '08), Association for Computing Machinery, New York, NY, USA, 1096--1103.
 DOI:https://doi.org/\href{https://doi.org/10.1145/1390156.1390294}{10.1145/1390156.1390294}}
@@
 \CSLLeftMargin{{[}15{]} }%
-\CSLRightInline{Andreas Weigend. 1994. On overfitting and the effective
-number of hidden units. In \emph{Proceedings of the 1993 connectionist
-models summer school}, 335--342.}
+\CSLRightInline{Andreas Weigend. 1994. On overfitting and the effective number of hidden units. In \emph{Proceedings of the 1993 connectionist models summer school}, 335--342.}
 
@@
 \CSLLeftMargin{{[}16{]} }%
-\CSLRightInline{David Wittenberg. 2022. Using denoising autoencoder
-genetic programming to~control exploration and~exploitation in~search.
-In \emph{Genetic programming}, Springer International Publishing, Cham,
-102--117.}
+\CSLRightInline{David Wittenberg.
+2022.
+Using denoising autoencoder genetic programming to control exploration and exploitation in search.
+In \emph{Genetic programming}, Springer International Publishing, Cham, 102--117.}
 
@@
 \CSLLeftMargin{{[}17{]} }%
-\CSLRightInline{David Wittenberg and Franz Rothlauf. 2022. Denoising
-autoencoder genetic programming for real-world symbolic regression. In
-\emph{Proceedings of the genetic and evolutionary computation conference
-companion} (GECCO '22), Association for Computing Machinery, New York,
-NY, USA, 612--614.
+\CSLRightInline{David Wittenberg and Franz Rothlauf.
+2022.
+Denoising autoencoder genetic programming for real-world symbolic regression.
+In \emph{Proceedings of the genetic and evolutionary computation conference companion} (GECCO '22), Association for Computing Machinery, New York, NY, USA, 612--614.
 DOI:https://doi.org/\href{https://doi.org/10.1145/3520304.3528921}{10.1145/3520304.3528921}}
@@
 \CSLRightInline{David Wittenberg, Franz Rothlauf, and Dirk Schweim.
-2020. DAE-GP: Denoising autoencoder LSTM networks as probabilistic
-models in estimation of distribution genetic programming. In
-\emph{Proceedings of the 2020 genetic and evolutionary computation
-conference} (GECCO '20), Association for Computing Machinery, New York,
-NY, USA, 1037--1045.
+2020.
+DAE-GP: Denoising autoencoder LSTM networks as probabilistic models in estimation of distribution genetic programming.
+In \emph{Proceedings of the 2020 genetic and evolutionary computation conference} (GECCO '20), Association for Computing Machinery, New York, NY, USA, 1037--1045.
 DOI:https://doi.org/\href{https://doi.org/10.1145/3377930.3390180}{10.1145/3377930.3390180}}

### 2023-01-27 paper.Rmd
Fri Jan 27 13:05:58 CET 2023: update-remote.sh

@@
 
-The DAE-GP algorithm, first described by @dae-gp_2020_rtree, is summarized as a flowchart in \@ref(fig:daegp-flowchart).
+The DAE-GP algorithm, first described by @dae-gp_2020_rtree, is summarized as a flowchart in figure \@ref(fig:daegp-flowchart).
 After setting the generation counter $g$ to $0$ and creating the initial population of solutions $P_0$, DAE-GP enters the main loop that checks if a termination criteria is satisfied (i.e.
@@
 
-The pre-training strategy implemented for all experiments in this thesis is visualized as another flowchart in figure §.
+The pre-training strategy implemented for all experiments in this thesis is visualized as another flowchart in figure \@ref(fig:pt-daegp-flowchart).
 The main difference to regular DAE-GP is the inclusion of an initial pre-training phase where a separate population $\hat{P}$ is first initialized and then randomly split by half into $\hat{P}_{train}$ and $\hat{P}_{test}$.
@@
 After the training for $\hat{M}$ is stopped, the current state of $\hat{M}$ is frozen and the optimized trainable parameters $\theta_{\hat{M}}$ are saved before terminating the pre-training phase to start the DAE-GP phase.
-During the DAE-GP phase, the only difference to the traditional DAE-GP algorithm described in figure § is during model building: We initialize $M_g$ with $\theta_{\hat{M}}$ before training it to learn the properties of $X_g$.
+During the DAE-GP phase, the only difference to the traditional DAE-GP algorithm described in figure \@ref(fig:daegp-flowchart) is during model building: We initialize $M_g$ with $\theta_{\hat{M}}$ before training it to learn the properties of $X_g$.
 
-![Flowchart - Pre-Trained DAE-GP](./img/flowcharts/pt-dae-gp.png)
+```{r pt-daegp-flowchart, echo=FALSE, fig.cap="Pre-Trained DAE-GP Flowchart", out.height="90%", out.width="50%"}
+knitr::include_graphics("./img/flowcharts/pt-dae-gp.png")
+```
 
 One difficulty in the implementation of a pre-training strategy into DAE-GP has been the detemination of the number of hidden neurons in side the DAE-LSTMs hidden layers.
-The original description of DAE-GP [@dae-gp_2020_rtree] used a strategy where the number of hidden neurons for each hidden layer is dynamically set per generation to the maximum individual size inside the current population.
+The original description of DAE-GP [@dae-gp_2020_rtree] uses a strategy where the number of hidden neurons for each hidden layer is dynamically set per generation to the maximum individual size inside the current population.
 For a pre-training implementation, this strategy can not be easily adapted since it leads to a changing number of neurons at each generation resulting in different dimensions of the DAE-LSTM.
-To allow for the sharing of weights and biases from the pre-trained model $\hat{M}$ to each $M_g$ for $g\in{1,...,g_{max}}$ I experimented with two different strategies:
+To allow the sharing of the state of the pre-trained parameters $\theta_{\hat{M}}$ from the pre-trained model $\hat{M}$ to each $M_g$ for $g\in{1,...,g_{max}}$  I set the number of hidden neurons per hidden layer for all pre-trained runs statically to the maximum individual size inside the pre-training population $\hat{P}_{train}$
 
-1. Set the number of hidden neurons per hidden layer for each generation to the maximum individual size inside the pre-training population
-2. (dependent on the terminal/function set)
+If not stated otherwise, all experiments in this thesis use the hyperparameters listed in table \@ref(tab:airfoil-fullRun-2hl-maxIndSize-params).
 
-The framework that was used to conduct all experiments of this thesis was provided by the supervisor David Wittenberg and uses the python programming language in conjunction with the baseline libraries `Keras` [@chollet2015keras] for deep neural networks and `deap` [@DEAP_JMLR2012] for evolutionary computation.
+```{r airfoil-fullRun-2hl-maxIndSize-params, echo=FALSE, message=FALSE, warning=FALSE}
+fig.pos="H"
+knitr::kable(
+  read.csv("./tables/airfoil_fullRun_2hl_maxIndSize_params"),
+  caption = "DAE-GP - Hyperparameter"
+)%>% row_spec(0,bold=TRUE )
+```
+
+The framework that was used to conduct all experiments of this thesis was provided by the supervisor of this thesis David Wittenberg and uses the python programming language in conjunction with the baseline libraries `Keras` [@chollet2015keras] for deep neural networks and `deap` [@DEAP_JMLR2012] for evolutionary computation.
 
@@
 The main experiments conducted in this thesis uses the NASA Airfoil Self-Noise Data Set which is part of the UCI machine learning repository [@machine_learning_repo].
-The dataset consists of 5 input variables and 1 output variable that are listed in table 1.
+The dataset consists of 5 input variables and 1 output variable that are listed in table \@ref(tab:airfoil-dataset-description).
 
@@
 After closely examining the effect of pre-training on DAE-GPs generalization behaviour another series of experiments is conducted to study how pre-training influences the overall search behaviour of DAE-GP.
-Again the Airfoil problem was selected as a first benchmark problem to test pre-trained DAE-GP against traditional DAE-GP [@dae-gp_2022_symreg], the parameters are described in table $A$
+Again the Airfoil problem was selected as a first benchmark problem to test pre-trained DAE-GP against traditional DAE-GP [@dae-gp_2022_symreg], the parameters are described in table
 
-```{r airfoil-fullRun-2hl-maxIndSize-params, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
-fig.pos="H"
-knitr::kable(
-  read.csv("./tables/airfoil_fullRun_2hl_maxIndSize_params"),
-  caption="Airfoil - Parameter for full Run"
-)%>% row_spec(0,bold=TRUE ) %>% kableExtra::kable_styling(latex_options = "hold_position")
-```
+<!-- # TODO: Reformulate -->
 

### 2023-01-27 paper.tex
Fri Jan 27 13:05:58 CET 2023: update-remote.sh

@@
 
-The DAE-GP algorithm, first described by {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]}, is summarized as a flowchart in \ref{fig:daegp-flowchart}.
+The DAE-GP algorithm, first described by {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]}, is summarized as a flowchart in figure \ref{fig:daegp-flowchart}.
 After setting the generation counter \(g\) to \(0\) and creating the initial population of solutions \(P_0\), DAE-GP enters the main loop that checks if a termination criteria is satisfied (i.e.~maximum number of generations or fitness evaluations is reached).
@@
 
-The pre-training strategy implemented for all experiments in this thesis is visualized as another flowchart in figure §.
+The pre-training strategy implemented for all experiments in this thesis is visualized as another flowchart in figure \ref{fig:pt-daegp-flowchart}.
 The main difference to regular DAE-GP is the inclusion of an initial pre-training phase where a separate population \(\hat{P}\) is first initialized and then randomly split by half into \(\hat{P}_{train}\) and \(\hat{P}_{test}\).
@@
 After the training for \(\hat{M}\) is stopped, the current state of \(\hat{M}\) is frozen and the optimized trainable parameters \(\theta_{\hat{M}}\) are saved before terminating the pre-training phase to start the DAE-GP phase.
-During the DAE-GP phase, the only difference to the traditional DAE-GP algorithm described in figure § is during model building: We initialize \(M_g\) with \(\theta_{\hat{M}}\) before training it to learn the properties of \(X_g\).
+During the DAE-GP phase, the only difference to the traditional DAE-GP algorithm described in figure \ref{fig:daegp-flowchart} is during model building: We initialize \(M_g\) with \(\theta_{\hat{M}}\) before training it to learn the properties of \(X_g\).
 
-\begin{figure}
-\centering
-\includegraphics{./img/flowcharts/pt-dae-gp.png}
-\caption{Flowchart - Pre-Trained DAE-GP}
+\begin{figure}[c]
+
+{\centering \includegraphics[width=0.5\linewidth,height=0.9\textheight]{./img/flowcharts/pt-dae-gp}
+
+}
+
+\caption{Pre-Trained DAE-GP Flowchart}\label{fig:pt-daegp-flowchart}
 \end{figure}
@@
 One difficulty in the implementation of a pre-training strategy into DAE-GP has been the detemination of the number of hidden neurons in side the DAE-LSTMs hidden layers.
-The original description of DAE-GP {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]} used a strategy where the number of hidden neurons for each hidden layer is dynamically set per generation to the maximum individual size inside the current population.
+The original description of DAE-GP {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]} uses a strategy where the number of hidden neurons for each hidden layer is dynamically set per generation to the maximum individual size inside the current population.
 For a pre-training implementation, this strategy can not be easily adapted since it leads to a changing number of neurons at each generation resulting in different dimensions of the DAE-LSTM.
-To allow for the sharing of weights and biases from the pre-trained model \(\hat{M}\) to each \(M_g\) for \(g\in{1,...,g_{max}}\) I experimented with two different strategies:
+To allow the sharing of the state of the pre-trained parameters \(\theta_{\hat{M}}\) from the pre-trained model \(\hat{M}\) to each \(M_g\) for \(g\in{1,...,g_{max}}\) I set the number of hidden neurons per hidden layer for all pre-trained runs statically to the maximum individual size inside the pre-training population \(\hat{P}_{train}\)
 
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Set the number of hidden neurons per hidden layer for each generation to the maximum individual size inside the pre-training population
-\item
-  (dependent on the terminal/function set)
-\end{enumerate}
+If not stated otherwise, all experiments in this thesis use the hyperparameters listed in table \ref{tab:airfoil-fullRun-2hl-maxIndSize-params}.
 
-The framework that was used to conduct all experiments of this thesis was provided by the supervisor David Wittenberg and uses the python programming language in conjunction with the baseline libraries \texttt{Keras} {[}\protect\hyperlink{ref-chollet2015keras}{1}{]} for deep neural networks and \texttt{deap} {[}\protect\hyperlink{ref-DEAP_JMLR2012}{4}{]} for evolutionary computation.
+\begin{table}
+
+\caption{\label{tab:airfoil-fullRun-2hl-maxIndSize-params}DAE-GP - Hyperparameter}
+\centering
+\begin{tabular}[t]{l|l}
+\hline
+\textbf{Hyperparameter} & \textbf{Value}\\
+\hline
+popSize & 500\\
+\hline
+generations & 30\\
+\hline
+fitness & RMSE\\
+\hline
+TrainingMode & Convergence\\
+\hline
+SamplingSteps & 2\\
+\hline
+hiddenLayers & 2\\
+\hline
+Selection & Binary Tournament Selection\\
+\hline
+corruptionTechnique & Levenshtein Edit\\
+\hline
+edit\_prob(training/sampling) & 0.05/0.95\\
+\hline
+functionSet & \{+ ; - ; * ; aq\}\\
+\hline
+Pre-Training PopulationSize & 10000\\
+\hline
+Pre-Training Train/Test Split & 50\%\\
+\hline
+Pre-Training TrainingMode & Early Stopping\\
+\hline
+\end{tabular}
+\end{table}
+
+The framework that was used to conduct all experiments of this thesis was provided by the supervisor of this thesis David Wittenberg and uses the python programming language in conjunction with the baseline libraries \texttt{Keras} {[}\protect\hyperlink{ref-chollet2015keras}{1}{]} for deep neural networks and \texttt{deap} {[}\protect\hyperlink{ref-DEAP_JMLR2012}{4}{]} for evolutionary computation.
 
@@
 The main experiments conducted in this thesis uses the NASA Airfoil Self-Noise Data Set which is part of the UCI machine learning repository {[}\protect\hyperlink{ref-machine_learning_repo}{2}{]}.
-The dataset consists of 5 input variables and 1 output variable that are listed in table 1.
+The dataset consists of 5 input variables and 1 output variable that are listed in table \ref{tab:airfoil-dataset-description}.
 
@@
 After closely examining the effect of pre-training on DAE-GPs generalization behaviour another series of experiments is conducted to study how pre-training influences the overall search behaviour of DAE-GP.
-Again the Airfoil problem was selected as a first benchmark problem to test pre-trained DAE-GP against traditional DAE-GP {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]}, the parameters are described in table \(A\)
-
-\begin{table}[!h]
-
-\caption{\label{tab:airfoil-fullRun-2hl-maxIndSize-params}Airfoil - Parameter for full Run}
-\centering
-\begin{tabular}[t]{l|l}
-\hline
-\textbf{Hyperparameter} & \textbf{Value}\\
-\hline
-popSize & 500\\
-\hline
-generations & 30\\
-\hline
-fitness & RMSE\\
-\hline
-TrainingMode & Convergence\\
-\hline
-SamplingSteps & 2\\
-\hline
-hiddenLayers & 2\\
-\hline
-Selection & Binary Tournament Selection\\
-\hline
-corruptionTechnique & Levenshtein Edit\\
-\hline
-edit\_prob(training/sampling) & 0.05/0.95\\
-\hline
-functionSet & \{+ ; - ; * ; aq\}\\
-\hline
-Pre-Training PopulationSize & 10000\\
-\hline
-Pre-Training Train/Test Split & 50\%\\
-\hline
-Pre-Training TrainingMode & Early Stopping\\
-\hline
-\end{tabular}
-\end{table}
+Again the Airfoil problem was selected as a first benchmark problem to test pre-trained DAE-GP against traditional DAE-GP {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]}, the parameters are described in table
 

### 2023-01-27 paper.Rmd
Fri Jan 27 15:00:07 CET 2023: update-remote.sh

@@
 
-```{r daegp-flowchart, echo=FALSE, fig.cap="Regular DAE-GP Flowchart", out.height="90%", out.width="50%"}
+```{r daegp-flowchart, echo=FALSE, fig.cap="Regular DAE-GP Flowchart", out.width="60%"}
 knitr::include_graphics("./img/flowcharts/dae-gp.png")
@@
 
-```{r pt-daegp-flowchart, echo=FALSE, fig.cap="Pre-Trained DAE-GP Flowchart", out.height="90%", out.width="50%"}
+```{r pt-daegp-flowchart, echo=FALSE, fig.cap="Pre-Trained DAE-GP Flowchart", out.width="65%"}
 knitr::include_graphics("./img/flowcharts/pt-dae-gp.png")
@@
 knitr::kable(
-  read.csv("./tables/airfoil_fullRun_2hl_maxIndSize_params"),
+  read.csv("./tables/airfoil_fullRun_2hl_maxIndSize_params", sep = "?"),
   caption = "DAE-GP - Hyperparameter"
@@
 
-```{r airfoil-function-set, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
+Besides studying the performance for the airfoil dataset, additional real world symbolic regressions experiments used the datasets[@machine_learning_repo] described in table \@ref(tab:full-run-realWorldSymReg-problems).
+
+```{r full-run-realWorldSymReg-problems, echo=FALSE, message=FALSE, warning=FALSE}
 fig.pos="H"
 knitr::kable(
-  read.csv("./tables/airfoil_function_set"),
-  caption="Airfoil - Function Set"
+  read.csv("./tables/full_run_realWorldSymReg_problems.csv"),
+  caption="Real World Symbolic Regression Benchmark Problems"
 )%>% row_spec(0,bold=TRUE ) %>% kableExtra::kable_styling(latex_options = "hold_position")
 ```
-
-<!--- energy heating dataset -->
 
@@
 To study generalization behavior this series of experiments is conducted using DAE-GP with only a single generation until the search terminates.
-I disregard the fitness of solutions found and focus soley on the reconstruction error that is produced during the training of each the DAE-LSTM.
+I disregard the fitness of solutions found and focus soley on the reconstruction error that is produced during the training of each DAE-LSTM.
 The reconstruction error is measured for two separate populations, a training population $P_{train}$ that is used to train our DAE-LSTM as well as a hold-out validation population $P_{test}$.
@@
 
-<!--- analyze results 1 hidden layer and variable number of hidden neurons -->
+<!--- analyzing the results for variable number of neurons -->
 
-![First Generation Median Training Error for variable number of hidden Neurons - Airfoil](./img/airfoil_firstGen/airfoil_firstGen_mean_training_error_by_neurons_6plots.png)
+...
+
+```{r first-gen-airfoil-byNeurons, echo=FALSE, fig.cap="First Generation Median Training Error for variable number of hidden Neurons", out.width="90%"}
+knitr::include_graphics("./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_neurons_3plots.png")
+```
 
@@
 
-![First Generation Median Training Error for variable number of hidden Layers - Airfoil](./img/airfoil_firstGen/airfoil_firstGen_mean_training_error_by_layers_6plot.png)
+...
+
+```{r first-gen-airfoil-byLayers, echo=FALSE, fig.cap="First Generation Median Training Error for variable number of hidden Layers", out.width="90%"}
+knitr::include_graphics("./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_layers_3plots.png")
+```
 
@@
 After closely examining the effect of pre-training on DAE-GPs generalization behaviour another series of experiments is conducted to study how pre-training influences the overall search behaviour of DAE-GP.
-Again the Airfoil problem was selected as a first benchmark problem to test pre-trained DAE-GP against traditional DAE-GP [@dae-gp_2022_symreg], the parameters are described in table
-
-<!-- # TODO: Reformulate -->
+The airfoil problem was selected as a first benchmark problem to test pre-trained DAE-GP against traditional DAE-GP.
 
@@
 
-![Best Fitness over 30 Generations - Airfoil](./img/airfoil_2hl_maxIndSize_fullRun_30gens_withGP/mean_median_fitness_byGens.png)
+Figure \@ref(fig:airfoil_median_fitness_30gens) shows both the mean and median of the best fitness for each generation on both the test and training set of the airfoil dataset.
+The results are aggregated for $10$ individuals runs per algorithm and also include regular GP as a benchmark.
+For all three algorithms an improvement in the best found fitness can be observed during the evolutionary search but the results show that DAE-GP based algorithms, on average, find solutions with much higher fitness as traditional GP.
+
+The first interesting observation from this experiment is, that the pre-trained DAE-GP shows better performance than regular DAE-GP for both the training and the testing set in overall fitness.
+Regarding the presence of overfitting, both DAE-GP variants show only a very small gap between the testing and training fitness which indicates a good generalization behaviour.
+
+```{r airfoil_median_fitness_30gens, echo=FALSE, fig.cap="Best Fitness over 30 Generations - Airfoil", out.width="90%"}
+knitr::include_graphics("./img/airfoil_2hl_maxIndSize_fullRun_30gens_withGP/mean_median_fitness_byGens.png")
+```
 
@@
 
-To gather more confidence in the previous results, the pre-training strategy for DAE-GP was tested for various real world symbolic regression problems taken from @machine_learning_repo.
-Are brief summary of the benchmark problems is given in table X
+To gather more confidence in the previous results, the pre-training strategy for DAE-GP was tested for three additional real world symbolic regression problems taken from @machine_learning_repo.^[see table \@ref(tab:full-run-realWorldSymReg-problems)]
 
-```{r full-run-realWorldSymReg-problems, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
-fig.pos="H"
-knitr::kable(
-  read.csv("./tables/full_run_realWorldSymReg_problems.csv"),
-  caption="Overview - Real World Symbolic Regression Benchmark Problems"
-)%>% row_spec(0,bold=TRUE ) %>% kableExtra::kable_styling(latex_options = "hold_position")
-```
 Table X summarizes the median best fitness achieved after running each algorithm for 10 runs using a population size of 500 individuals for 30 generations.^[P-Values are marked with asteriks (1,2,3) for ($\alpha$)-levels of $0.1$, $0.05$ and $0.01$].

### 2023-01-27 paper.tex
Fri Jan 27 15:00:07 CET 2023: update-remote.sh

@@
 
-{\centering \includegraphics[width=0.5\linewidth,height=0.9\textheight]{./img/flowcharts/dae-gp}
+{\centering \includegraphics[width=0.6\linewidth]{./img/flowcharts/dae-gp}
 
@@
 
-{\centering \includegraphics[width=0.5\linewidth,height=0.9\textheight]{./img/flowcharts/pt-dae-gp}
+{\centering \includegraphics[width=0.65\linewidth]{./img/flowcharts/pt-dae-gp}
 
@@
 \hline
-popSize & 500\\
+Population\_Size & 500\\
 \hline
-generations & 30\\
+Generations & 30\\
 \hline
-fitness & RMSE\\
+Fitness\_Metric & RMSE\\
 \hline
-TrainingMode & Convergence\\
+Training\_Mode & Convergence\\
 \hline
-SamplingSteps & 2\\
+Sampling\_Steps & 2\\
 \hline
-hiddenLayers & 2\\
+Selection\_Operator & Binary Tournament Selection\\
 \hline
-Selection & Binary Tournament Selection\\
+Corruption\_Operator & Levenshtein Edit\\
 \hline
-corruptionTechnique & Levenshtein Edit\\
+Denoising\_Edit\_Probability(training/sampling) & 0.05/0.95\\
 \hline
-edit\_prob(training/sampling) & 0.05/0.95\\
+Function\_Set & \{+ , - , * , aq\}\\
 \hline
-functionSet & \{+ ; - ; * ; aq\}\\
+Ephemeral\_Constants & [-5,..,5]\\
 \hline
-Pre-Training PopulationSize & 10000\\
+Pre-Training\_Population\_Size & 10000\\
 \hline
-Pre-Training Train/Test Split & 50\%\\
+Pre-Training\_Train/Test\_Split & 50\%\\
 \hline
-Pre-Training TrainingMode & Early Stopping\\
+Pre-Training\_Training\_Mode & Early Stopping\\
 \hline
@@
 
+Besides studying the performance for the airfoil dataset, additional real world symbolic regressions experiments used the datasets{[}\protect\hyperlink{ref-machine_learning_repo}{2}{]} described in table \ref{tab:full-run-realWorldSymReg-problems}.
+
 \begin{table}[!h]
 
-\caption{\label{tab:airfoil-function-set}Airfoil - Function Set}
+\caption{\label{tab:full-run-realWorldSymReg-problems}Real World Symbolic Regression Benchmark Problems}
 \centering
-\begin{tabular}[t]{l|r}
+\begin{tabular}[t]{l|r|r}
 \hline
-\textbf{operator} & \textbf{arity}\\
+\textbf{Problem} & \textbf{Observations} & \textbf{Features}\\
 \hline
-addition & 2\\
+Airfoil & 1503 & 5\\
 \hline
-subtraction & 2\\
+Boston\_Housing & 506 & 13\\
 \hline
-multiplication & 2\\
+Energy\_Cooling & 768 & 8\\
 \hline
-analytic\_quotient & 2\\
+Concrete & 1030 & 8\\
 \hline
@@
 To study generalization behavior this series of experiments is conducted using DAE-GP with only a single generation until the search terminates.
-I disregard the fitness of solutions found and focus soley on the reconstruction error that is produced during the training of each the DAE-LSTM.
+I disregard the fitness of solutions found and focus soley on the reconstruction error that is produced during the training of each DAE-LSTM.
 The reconstruction error is measured for two separate populations, a training population \(P_{train}\) that is used to train our DAE-LSTM as well as a hold-out validation population \(P_{test}\).
@@
 
-\begin{figure}
-\centering
-\includegraphics{./img/airfoil_firstGen/airfoil_firstGen_mean_training_error_by_neurons_6plots.png}
-\caption{First Generation Median Training Error for variable number of hidden Neurons - Airfoil}
+\ldots{}
+
+\begin{figure}[c]
+
+{\centering \includegraphics[width=0.9\linewidth]{./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_neurons_3plots}
+
+}
+
+\caption{First Generation Median Training Error for variable number of hidden Neurons}\label{fig:first-gen-airfoil-byNeurons}
 \end{figure}
 
-\begin{figure}
-\centering
-\includegraphics{./img/airfoil_firstGen/airfoil_firstGen_mean_training_error_by_layers_6plot.png}
-\caption{First Generation Median Training Error for variable number of hidden Layers - Airfoil}
+\ldots{}
+
+\begin{figure}[c]
+
+{\centering \includegraphics[width=0.9\linewidth]{./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_layers_3plots}
+
+}
+
+\caption{First Generation Median Training Error for variable number of hidden Layers}\label{fig:first-gen-airfoil-byLayers}
 \end{figure}
@@
 
-To gather more confidence in the previous results, the pre-training strategy for DAE-GP was tested for various real world symbolic regression problems taken from {[}\protect\hyperlink{ref-machine_learning_repo}{2}{]}.
-Are brief summary of the benchmark problems is given in table X
-
-\begin{table}[!h]
-
-\caption{\label{tab:full-run-realWorldSymReg-problems}Overview - Real World Symbolic Regression Benchmark Problems}
-\centering
-\begin{tabular}[t]{l|r|r}
-\hline
-\textbf{Problem} & \textbf{Observations} & \textbf{Features}\\
-\hline
-Airfoil & 1503 & 5\\
-\hline
-Boston\_Housing & 506 & 13\\
-\hline
-Energy\_Cooling & 768 & 8\\
-\hline
-Concrete & 1030 & 8\\
-\hline
-\end{tabular}
-\end{table}
+To gather more confidence in the previous results, the pre-training strategy for DAE-GP was tested for three additional real world symbolic regression problems taken from {[}\protect\hyperlink{ref-machine_learning_repo}{2}{]}.\footnote{see table \ref{tab:full-run-realWorldSymReg-problems}}
 

### 2023-01-27 paper.Rmd
Fri Jan 27 16:00:52 CET 2023: update-remote.sh

@@
 
-Figure \@ref(fig:airfoil_median_fitness_30gens) shows both the mean and median of the best fitness for each generation on both the test and training set of the airfoil dataset.
+### Airfoil Dateset - 2 hidden Layers
+
+Figure \@ref(fig:airfoil-median-fitness-30gens) shows both the mean and median of the best fitness for each generation on both the test and training set of the airfoil dataset for using 2 hidden layers inside the DAE-LSTM networks.
 The results are aggregated for $10$ individuals runs per algorithm and also include regular GP as a benchmark.
@@
 
-The first interesting observation from this experiment is, that the pre-trained DAE-GP shows better performance than regular DAE-GP for both the training and the testing set in overall fitness.
+The first interesting observation from this experiment is, that the pre-trained DAE-GP shows better mean and median performance than regular DAE-GP for both the training and the testing set in overall fitness.
 Regarding the presence of overfitting, both DAE-GP variants show only a very small gap between the testing and training fitness which indicates a good generalization behaviour.
 
-```{r airfoil_median_fitness_30gens, echo=FALSE, fig.cap="Best Fitness over 30 Generations - Airfoil", out.width="90%"}
+```{r airfoil-median-fitness-30gens, echo=FALSE, fig.cap="Best Fitness over 30 Generations - Airfoil (2HL)", out.width="90%"}
 knitr::include_graphics("./img/airfoil_2hl_maxIndSize_fullRun_30gens_withGP/mean_median_fitness_byGens.png")
 ```
+The distribution of final fitness scores for pre-trained and regular DAE-GP are visualized as boxplots in figure  \@ref(fig:airfoil-distribution-fitness-30gens).
+The distribution for pre-trained DAE-GP shows to be preferable to that of regular DAE-GP on both sets since it shows a lower median (med.) as well as less dispersion measured by the standard deviation (std.).
 
-![Best Fitness Distribution after 30 Generations - Airfoil](./img/airfoil_2hl_maxIndSize_fullRun_30gens/final_fit_boxplot.png)
-lalal
+```{r airfoil-distribution-fitness-30gens, echo=FALSE, fig.cap="Best Fitness after 30 Generations - Airfoil (2HL)", out.width="80%"}
+knitr::include_graphics("./img/airfoil_2hl_maxIndSize_fullRun_30gens/final_fit_boxplot.png")
+```
 
-![Population Diversity over 30 Generations - Airfoil](./img/airfoil_2hl_maxIndSize_fullRun_30gens/mean_popDiversity_byGens.png)
+Since previous work by @dae-gp_2022_symreg demonstrated that DAE-GP was capable of producing solutions that are much smaller in tree size that regular GP, I also studied the influence of pre-training on the size of solutions found by DAE-GP.
+
 Since @dae-gp_2022_symreg shows, that DAE-GP finds higher quality that are also smaller in size than those found by average GP for a given number of 10.000 fitness evaluations, I also studied the differences in average solution size between pre-trained and regular DAE-GP.
+Figure \@ref(fig:airfoil-size-combined-30gens) shows the median of both the average size of individuals inside the population as well as the size of the currently best performing solution inside the population.
+It can be observed that both DAE-GP variants strongly reduce the average solutions size already in the first generations and that the average solution size inside the population constantly stays at this very low level for the rest of the search.
 
-![Average Solution Size over 30 Generations - Airfoil](./img/airfoil_2hl_maxIndSize_fullRun_30gens/mean_Size_byGens.png)
+```{r airfoil-size-combined-30gens, echo=FALSE, fig.cap="Median Solution Size over 30 Generations - Airfoil (2HL)", out.width="80%"}
+knitr::include_graphics("./img/airfoil_2hl_maxIndSize_fullRun_30gens/size_combined_median.png")
+```
 

### 2023-01-27 paper.tex
Fri Jan 27 16:00:52 CET 2023: update-remote.sh

@@
 After closely examining the effect of pre-training on DAE-GPs generalization behaviour another series of experiments is conducted to study how pre-training influences the overall search behaviour of DAE-GP.
-Again the Airfoil problem was selected as a first benchmark problem to test pre-trained DAE-GP against traditional DAE-GP {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]}, the parameters are described in table
+The airfoil problem was selected as a first benchmark problem to test pre-trained DAE-GP against traditional DAE-GP.
 
@@
 
-\begin{figure}
-\centering
-\includegraphics{./img/airfoil_2hl_maxIndSize_fullRun_30gens_withGP/mean_median_fitness_byGens.png}
-\caption{Best Fitness over 30 Generations - Airfoil}
+\hypertarget{airfoil-dateset---2-hidden-layers}{%
+\subsubsection{Airfoil Dateset - 2 hidden Layers}\label{airfoil-dateset---2-hidden-layers}}
+
+Figure \ref{fig:airfoil-median-fitness-30gens} shows both the mean and median of the best fitness for each generation on both the test and training set of the airfoil dataset for using 2 hidden layers inside the DAE-LSTM networks.
+The results are aggregated for \(10\) individuals runs per algorithm and also include regular GP as a benchmark.
+For all three algorithms an improvement in the best found fitness can be observed during the evolutionary search but the results show that DAE-GP based algorithms, on average, find solutions with much higher fitness as traditional GP.
+
+The first interesting observation from this experiment is, that the pre-trained DAE-GP shows better mean and median performance than regular DAE-GP for both the training and the testing set in overall fitness.
+Regarding the presence of overfitting, both DAE-GP variants show only a very small gap between the testing and training fitness which indicates a good generalization behaviour.
+
+\begin{figure}[c]
+
+{\centering \includegraphics[width=0.9\linewidth]{./img/airfoil_2hl_maxIndSize_fullRun_30gens_withGP/mean_median_fitness_byGens}
+
+}
+
+\caption{Best Fitness over 30 Generations - Airfoil (2HL)}\label{fig:airfoil-median-fitness-30gens}
 \end{figure}
 
-\includegraphics{./img/airfoil_2hl_maxIndSize_fullRun_30gens/final_fit_boxplot.png}
-lalal
+The distribution of final fitness scores for pre-trained and regular DAE-GP are visualized as boxplots in figure \ref{fig:airfoil-distribution-fitness-30gens}.
+The distribution for pre-trained DAE-GP shows to be preferable to that of regular DAE-GP on both sets since it shows a lower median (med.) as well as less dispersion measured by the standard deviation (std.).
 
-\includegraphics{./img/airfoil_2hl_maxIndSize_fullRun_30gens/mean_popDiversity_byGens.png}
+\begin{figure}[c]
+
+{\centering \includegraphics[width=0.8\linewidth]{./img/airfoil_2hl_maxIndSize_fullRun_30gens/final_fit_boxplot}
+
+}
+
+\caption{Best Fitness after 30 Generations - Airfoil (2HL)}\label{fig:airfoil-distribution-fitness-30gens}
+\end{figure}
+
+Since previous work by {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]} demonstrated that DAE-GP was capable of producing solutions that are much smaller in tree size that regular GP, I also studied the influence of pre-training on the size of solutions found by DAE-GP.
+
 Since {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]} shows, that DAE-GP finds higher quality that are also smaller in size than those found by average GP for a given number of 10.000 fitness evaluations, I also studied the differences in average solution size between pre-trained and regular DAE-GP.
+Figure \ref{fig:airfoil-size-combined-30gens} shows the median of both the average size of individuals inside the population as well as the size of the currently best performing solution inside the population.
+It can be observed that both DAE-GP variants strongly reduce the average solutions size already in the first generations and that the average solution size inside the population constantly stays at this very low level for the rest of the search.
 
-\begin{figure}
-\centering
-\includegraphics{./img/airfoil_2hl_maxIndSize_fullRun_30gens/mean_Size_byGens.png}
-\caption{Average Solution Size over 30 Generations - Airfoil}
+\begin{figure}[c]
+
+{\centering \includegraphics[width=0.8\linewidth]{./img/airfoil_2hl_maxIndSize_fullRun_30gens/size_combined_median}
+
+}
+
+\caption{Median Solution Size over 30 Generations - Airfoil (2HL)}\label{fig:airfoil-size-combined-30gens}
 \end{figure}

### 2023-01-27 paper.Rmd
autogit

@@
 
-Since previous work by @dae-gp_2022_symreg demonstrated that DAE-GP was capable of producing solutions that are much smaller in tree size that regular GP, I also studied the influence of pre-training on the size of solutions found by DAE-GP.
-
-Since @dae-gp_2022_symreg shows, that DAE-GP finds higher quality that are also smaller in size than those found by average GP for a given number of 10.000 fitness evaluations, I also studied the differences in average solution size between pre-trained and regular DAE-GP.
+Since previous work by @dae-gp_2022_symreg demonstrated that DAE-GP, for a given number of 10.000 fitness evaluations, was capable of producing solutions that are much smaller in tree size that regular GP, I also studied the influence of pre-training on the size of solutions found by DAE-GP.
 Figure \@ref(fig:airfoil-size-combined-30gens) shows the median of both the average size of individuals inside the population as well as the size of the currently best performing solution inside the population.
@@
 ```
+
+<!-- PopDiversity -->
 

### 2023-01-27 paper.Rmd
Fri Jan 27 17:00:07 CET 2023: update-remote.sh

@@
     number_sections: true
+    toc_depth: 3
     includes:
@@
 
-```{bash recompile-results, echo=TRUE, eval=FALSE, include=FALSE}
+```{bash recompile-results, echo=TRUE, eval=TRUE, include=FALSE}
 

### 2023-01-27 paper.tex
Fri Jan 27 17:00:07 CET 2023: update-remote.sh

@@
 {
-\setcounter{tocdepth}{2}
+\setcounter{tocdepth}{3}
 \tableofcontents
@@
 
-Since previous work by {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]} demonstrated that DAE-GP was capable of producing solutions that are much smaller in tree size that regular GP, I also studied the influence of pre-training on the size of solutions found by DAE-GP.
-
-Since {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]} shows, that DAE-GP finds higher quality that are also smaller in size than those found by average GP for a given number of 10.000 fitness evaluations, I also studied the differences in average solution size between pre-trained and regular DAE-GP.
+Since previous work by {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]} demonstrated that DAE-GP, for a given number of 10.000 fitness evaluations, was capable of producing solutions that are much smaller in tree size that regular GP, I also studied the influence of pre-training on the size of solutions found by DAE-GP.
 Figure \ref{fig:airfoil-size-combined-30gens} shows the median of both the average size of individuals inside the population as well as the size of the currently best performing solution inside the population.

### 2023-01-27 paper.Rmd
Fri Jan 27 18:00:06 CET 2023: update-remote.sh

@@
 
-```{bash recompile-results, echo=TRUE, eval=TRUE, include=FALSE}
+```{bash recompile-results, echo=TRUE, eval=FALSE, include=FALSE}
 
@@
 
-### Airfoil Dateset - 2 hidden Layers
+### Airfoil Dataset - 2 Hidden Layers
 
@@
 
-Since previous work by @dae-gp_2022_symreg demonstrated that DAE-GP, for a given number of 10.000 fitness evaluations, was capable of producing solutions that are much smaller in tree size that regular GP, I also studied the influence of pre-training on the size of solutions found by DAE-GP.
+Since previous work by @dae-gp_2022_symreg demonstrated that DAE-GP, for a given number of 10.000 fitness evaluations, is able to produce solutions that are much smaller in tree size that regular GP, I also studied the influence of pre-training on the size of solutions found by DAE-GP.
 Figure \@ref(fig:airfoil-size-combined-30gens) shows the median of both the average size of individuals inside the population as well as the size of the currently best performing solution inside the population.
 It can be observed that both DAE-GP variants strongly reduce the average solutions size already in the first generations and that the average solution size inside the population constantly stays at this very low level for the rest of the search.
+Interestingly, the results also show that the pre-trained version of algorithm did produce smaller median size of the best solutions.
 
@@
 
-<!-- PopDiversity -->
+To conclude the first experiment, figure \@ref(fig:airfoil-popDiversity) shows the median population diversity measured by the normalized Levenshtein edit distance for both DAE-GP algorithms and regular GP as a benchmark.
+While the population diversity steadily decreases over the generations for standard GP, DAE-GP variants strongly increase the population diversity in the first generations and keep them at a high level of over $0.9$.
+Again it can be observed, that pre-training shows a positive effect on the results by increasing population diversity in comparison to regular DAE-GP.
+
+```{r airfoil-popDiversity, echo=FALSE, fig.cap="Median Population Diversity over 30 Generations - Airfoil (2HL)", out.width="80%"}
+knitr::include_graphics("./img/airfoil_2hl_maxIndSize_fullRun_30gens_withGP/median_popDiversity_byGens_levOnly.png")
+```
+
+### Airfoil Dataset - 1 Hidden Layer
+
+Since previous research on DAE-GP used a single hidden layers for the airfoil dataset [@dae-gp_2022_symreg], I repeated the experiment with a single hidden layer.
+Based on the findings of @pmlr-v5-erhan09a, I expected that pre-training DAE-GP would lead to a worsening
+
+To confirm the research by @pmlr-v5-erhan09a
 

### 2023-01-27 paper.tex
Fri Jan 27 18:00:06 CET 2023: update-remote.sh

@@
 
-Since previous work by {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]} demonstrated that DAE-GP, for a given number of 10.000 fitness evaluations, was capable of producing solutions that are much smaller in tree size that regular GP, I also studied the influence of pre-training on the size of solutions found by DAE-GP.
+Since previous work by {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]} demonstrated that DAE-GP, for a given number of 10.000 fitness evaluations, is able to produce solutions that are much smaller in tree size that regular GP, I also studied the influence of pre-training on the size of solutions found by DAE-GP.
 Figure \ref{fig:airfoil-size-combined-30gens} shows the median of both the average size of individuals inside the population as well as the size of the currently best performing solution inside the population.
 It can be observed that both DAE-GP variants strongly reduce the average solutions size already in the first generations and that the average solution size inside the population constantly stays at this very low level for the rest of the search.
+Interestingly, the results also show that the pre-trained version of algorithm did produce smaller median size of the best solutions.
 
@@
 \caption{Median Solution Size over 30 Generations - Airfoil (2HL)}\label{fig:airfoil-size-combined-30gens}
+\end{figure}
+
+To conclude the first experiment, figure \ref{fig:airfoil-popDiversity} shows the median population diversity measured by the normalized Levenshtein edit distance for both DAE-GP algorithms and regular GP as a benchmark.
+While the population diversity steadily decreases over the generations for standard GP, DAE-GP variants strongly increase the population diversity in the first generations and keep them at a high level of over \(0.9\).
+Again it can be observed, that pre-training shows a positive effect on the results by increasing population diversity in comparison to regular DAE-GP.
+
+\begin{figure}[c]
+
+{\centering \includegraphics[width=0.8\linewidth]{./img/airfoil_2hl_maxIndSize_fullRun_30gens_withGP/median_popDiversity_byGens_levOnly}
+
+}
+
+\caption{Median Population Diversity over 30 Generations - Airfoil (2HL)}\label{fig:airfoil-popDiversity}
 \end{figure}

### 2023-01-28 paper.Rmd
Sat Jan 28 14:00:06 CET 2023: update-remote.sh

@@
 Since previous research on DAE-GP used a single hidden layers for the airfoil dataset [@dae-gp_2022_symreg], I repeated the experiment with a single hidden layer.
-Based on the findings of @pmlr-v5-erhan09a, I expected that pre-training DAE-GP would lead to a worsening
+Based on the findings of @pmlr-v5-erhan09a, I expected that pre-training DAE-GP would lead to a decrease in overall performance.
+Figures \@ref(fig:airfoil-fitness-perGen) and \@ref(fig:airfoil-finalFitness) show the average best fitness per generation as well as the distribution of final fitness scores after 30 generations.
+The results are, again, based on $10$ individual runs per algorithm.
+As expected it shows, that for a small DAE-LSTM size pre-training negativley impacts the fitness of solution found by DAE-GG.
 
-To confirm the research by @pmlr-v5-erhan09a
+```{r airfoil-fitness-perGen, echo=FALSE, fig.cap="Fitness over 30 Generations - Airfoil (1HL)", out.width="80%"}
+knitr::include_graphics("./img/airfoil_1hl_maxIndSize_fullRun_30gens/mean_median_fitness_byGens.png")
+```
 
-The experiments was repeated using the same setup with the only difference being a reduction from two hidden layers to 1 hidden layer.
-
-![Best Fitness over 30 Generations - 1 hidden Layer - Airfoil](./img/airfoil_1hl_maxIndSize_fullRun_30gens/mean_median_fitness_byGens.png)
+```{r airfoil-finalFitness, echo=FALSE, fig.cap="Fitness after 30 Generations - Airfoil (1HL)", out.width="80%"}
+knitr::include_graphics("./img/airfoil_1hl_maxIndSize_fullRun_30gens/final_fit_boxplot.png")
+```
 

### 2023-01-28 paper.tex
Sat Jan 28 14:00:06 CET 2023: update-remote.sh

@@
 
-\hypertarget{airfoil-dateset---2-hidden-layers}{%
-\subsubsection{Airfoil Dateset - 2 hidden Layers}\label{airfoil-dateset---2-hidden-layers}}
+\hypertarget{airfoil-dataset---2-hidden-layers}{%
+\subsubsection{Airfoil Dataset - 2 Hidden Layers}\label{airfoil-dataset---2-hidden-layers}}
 
@@
 
-The experiments was repeated using the same setup with the only difference being a reduction from two hidden layers to 1 hidden layer.
-
-\begin{figure}
-\centering
-\includegraphics{./img/airfoil_1hl_maxIndSize_fullRun_30gens/mean_median_fitness_byGens.png}
-\caption{Best Fitness over 30 Generations - 1 hidden Layer - Airfoil}
+\hypertarget{airfoil-dataset---1-hidden-layer}{%
+\subsubsection{Airfoil Dataset - 1 Hidden Layer}\label{airfoil-dataset---1-hidden-layer}}
+
+Since previous research on DAE-GP used a single hidden layers for the airfoil dataset {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]}, I repeated the experiment with a single hidden layer.
+Based on the findings of {[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{3}{]}, I expected that pre-training DAE-GP would lead to a decrease in overall performance.
+Figures \ref{fig:airfoil-fitness-perGen} and \ref{fig:airfoil-finalFitness} show the average best fitness per generation as well as the distribution of final fitness scores after 30 generations.
+The results are, again, based on \(10\) individual runs per algorithm.
+As expected it shows, that for a small DAE-LSTM size pre-training negativley impacts the fitness of solution found by DAE-GG.
+
+\begin{figure}[c]
+
+{\centering \includegraphics[width=0.8\linewidth]{./img/airfoil_1hl_maxIndSize_fullRun_30gens/mean_median_fitness_byGens}
+
+}
+
+\caption{Fitness over 30 Generations - Airfoil (1HL)}\label{fig:airfoil-fitness-perGen}
+\end{figure}
+
+\begin{figure}[c]
+
+{\centering \includegraphics[width=0.8\linewidth]{./img/airfoil_1hl_maxIndSize_fullRun_30gens/final_fit_boxplot}
+
+}
+
+\caption{Fitness after 30 Generations - Airfoil (1HL)}\label{fig:airfoil-finalFitness}
 \end{figure}

### 2023-01-28 paper.Rmd
Sat Jan 28 15:00:04 CET 2023: update-remote.sh
Sat Jan 28 16:00:04 CET 2023: update-remote.sh

@@
 
+# Alternative Pre-Training Strategies
+
+Since pre-trained thus far did not demonstrate significant benefits for DAE-GP, this chapter describes two alternative pre-training strategies that are also tested for the airfoil dataset:
+
+## Second Generation Pre-Training
+
+One common observation in EC-based optimization algorithms is that the largest improvements in fitness are often occurring in the first generation of the search process.
+Second Generation Pre-Training tries to exploit this observation by shifting the pre-training phase from the start of the algorithm into the second generation with the aim to produce a pre-training Population $\hat{P}$ that more accurately captures the properties of the search problem.
+The basic idea is to run basic DAE-GP for the first generation without any interference, after completing the first generation, the pre-training population $\hat{P}$ is created by sampling the DAE-LSTM model $M_1$.
+The pre-training model $\hat{M}$ is then trained using $\hat{P}$, after finishing DAE-GP continues running and initializes each new DAE-LSTM model $M_g$ for $g\in{2,...,g_{max}}$ using the trainable parameters $\theta_{\hat{M}}$.
+
+## Grow-Initialized Pre-Training
+
+The pre-training implementation in previous experiments used ramped half and half initialization [@Koza1993GeneticP] to create $\hat{P}$, this leads to 50% of the population consisting of full grown trees.
+An alternative method of initializing $\hat{P}$ is to use the grow initialization method where solutions are not guaranteed to be full trees.
+This different structural composition of solutions inside the pre-training population might be another way to improve the pre-training strategy by training on a population that more closely resembles the structural properties of futher generations population where a lower percentage of full grown trees is to be expected.
+
+## Results
+
+Figure \@ref(fig:altStrategies-Fitness) shows the median best fitness by generations for both alternative pre-training strategies as well as the results for standard DAE-GP and the standard Pre-Training implementation used in earlier exoeriments.
+
+```{r altStrategies-Fitness, echo=FALSE, fig.cap="Best Fitness over 30 Generations - Alternative Pre-Trategies - Airfoil", out.width="80%"}
+knitr::include_graphics("./img/PT_Strategies_airfoil_2hl_150hn_fullRun/median_fitness_byGens.png")
+```
+
+Figure \@ref(fig:altStrategies-finalFitness)
+
+```{r altStrategies-FitnessDistr, echo=FALSE, fig.cap="Best Fitness after 30 Generations - Alternative Pre-Trategies - Airfoil", out.width="80%"}
+knitr::include_graphics("./img/PT_Strategies_airfoil_2hl_150hn_fullRun/final_fit_boxplot(test_only).png")
+```
+
 # Discussion

### 2023-01-28 paper.Rmd
Sat Jan 28 17:00:05 CET 2023: update-remote.sh

@@
 # airfoil
+
+bash /Users/rmn/masterThesis/eda-gp-2020/experiments/airfoil_2hl_150hn_fullRun_30gens/re_evaluate_airfoil_2hl_150hn_30gen.sh
+
 bash /Users/rmn/masterThesis/eda-gp-2020/experiments/airfoil_1hl_maxIndSize_fullRun_30gens/re_evaluate_airfoil_1hl_maxIndSize_30gen.sh
@@
 
-Regarding the hidden neurons, the dimension of each hidden layer, I used two different strategies: Regular DAE-GP uses the same strategy as described by earlier work (see [@dae-gp_2022_symreg] or [@dae-gp_2020_rtree]), the number of hidden neurons is dynamically set at each generation to the maximum length of all solutions inside the current training population.
+<!-- Regarding the hidden neurons, the dimension of each hidden layer, I used two different strategies: Regular DAE-GP uses the same strategy as described by earlier work (see [@dae-gp_2022_symreg] or [@dae-gp_2020_rtree]), the number of hidden neurons is dynamically set at each generation to the maximum length of all solutions inside the current training population.
 For pre-trained DAE-GP another strategy had to be used, I used a fixed number of hidden neurons for each run that is set equal to the maximum length of all individuals that are present in the initial pre-training population.
@@
 It should nonetheless be mentioned that this strategy does give pre-training on average more hidden neurons per layer resulting in a more complex network (since the pre-training population is much larger than the regular population, the chances for larger individuals inside it are also higher).
-
-### Airfoil Dataset - 2 Hidden Layers
+-->
 
@@
 
-The first interesting observation from this experiment is, that the pre-trained DAE-GP shows better mean and median performance than regular DAE-GP for both the training and the testing set in overall fitness.
+The first interesting observation from this experiment is, that the pre-trained DAE-GP shows slightly lower mean and median performance than regular DAE-GP for both the training and the testing set in overall fitness.
 Regarding the presence of overfitting, both DAE-GP variants show only a very small gap between the testing and training fitness which indicates a good generalization behaviour.
@@
 ```{r airfoil-median-fitness-30gens, echo=FALSE, fig.cap="Best Fitness over 30 Generations - Airfoil (2HL)", out.width="90%"}
-knitr::include_graphics("./img/airfoil_2hl_maxIndSize_fullRun_30gens_withGP/mean_median_fitness_byGens.png")
+knitr::include_graphics("./img/airfoil_2hl_150hn_fullRun_30gens_withGP/mean_median_fitness_byGens.png")
 ```
 The distribution of final fitness scores for pre-trained and regular DAE-GP are visualized as boxplots in figure  \@ref(fig:airfoil-distribution-fitness-30gens).
-The distribution for pre-trained DAE-GP shows to be preferable to that of regular DAE-GP on both sets since it shows a lower median (med.) as well as less dispersion measured by the standard deviation (std.).
+The distribution for regular DAE-GP shows to be preferable to that of pre-trained DAE-GP on both sets since it shows a lower median RMSE as well as less dispersion of the solutions measured by the standard deviation.
 
 ```{r airfoil-distribution-fitness-30gens, echo=FALSE, fig.cap="Best Fitness after 30 Generations - Airfoil (2HL)", out.width="80%"}
-knitr::include_graphics("./img/airfoil_2hl_maxIndSize_fullRun_30gens/final_fit_boxplot.png")
+knitr::include_graphics("./img/airfoil_2hl_150hn_fullRun_30gens/final_fit_boxplot.png")
 ```
@@
 It can be observed that both DAE-GP variants strongly reduce the average solutions size already in the first generations and that the average solution size inside the population constantly stays at this very low level for the rest of the search.
-Interestingly, the results also show that the pre-trained version of algorithm did produce smaller median size of the best solutions.
+Regarding the size of the best solution per generation, the results show that the pre-trained version of algorithm did produce solutions with the same median size as traditional DAE-GP.
 
 ```{r airfoil-size-combined-30gens, echo=FALSE, fig.cap="Median Solution Size over 30 Generations - Airfoil (2HL)", out.width="80%"}
-knitr::include_graphics("./img/airfoil_2hl_maxIndSize_fullRun_30gens/size_combined_median.png")
+knitr::include_graphics("./img/airfoil_2hl_150hn_fullRun_30gens/size_combined_median.png")
 ```
@@
 While the population diversity steadily decreases over the generations for standard GP, DAE-GP variants strongly increase the population diversity in the first generations and keep them at a high level of over $0.9$.
-Again it can be observed, that pre-training shows a positive effect on the results by increasing population diversity in comparison to regular DAE-GP.
+Here it can be observed, that pre-training shows a positive effect on the results by increasing population diversity in comparison to regular DAE-GP.
 
 ```{r airfoil-popDiversity, echo=FALSE, fig.cap="Median Population Diversity over 30 Generations - Airfoil (2HL)", out.width="80%"}
-knitr::include_graphics("./img/airfoil_2hl_maxIndSize_fullRun_30gens_withGP/median_popDiversity_byGens_levOnly.png")
+knitr::include_graphics("./img/airfoil_2hl_150hn_fullRun_30gens_withGP/median_popDiversity_byGens_levOnly.png")
 ```
-
-### Airfoil Dataset - 1 Hidden Layer
 
@@
 
-# Alternative Pre-Training Strategies
+## Alternative Pre-Training Strategies
 
-Since pre-trained thus far did not demonstrate significant benefits for DAE-GP, this chapter describes two alternative pre-training strategies that are also tested for the airfoil dataset:
-
-## Second Generation Pre-Training
+Since pre-trained thus far did not demonstrate significant benefits for DAE-GP, this chapter describes two alternative pre-training strategies, Second Generation Pre-Training and  Grow-Initialized Pre-Training, that are also tested for the airfoil dataset.
 
@@
 
-## Grow-Initialized Pre-Training
-
 The pre-training implementation in previous experiments used ramped half and half initialization [@Koza1993GeneticP] to create $\hat{P}$, this leads to 50% of the population consisting of full grown trees.
@@
 
-## Results
-
-Figure \@ref(fig:altStrategies-Fitness) shows the median best fitness by generations for both alternative pre-training strategies as well as the results for standard DAE-GP and the standard Pre-Training implementation used in earlier exoeriments.
+Figure \@ref(fig:altStrategies-Fitness) shows the median best fitness by generations for both alternative pre-training strategies as well as the results for standard DAE-GP and the standard Pre-Training implementation used in earlier experiments.
+Unfortunately, both alternative strategies do show a worse performance in fitness than both DAE-GP and standard pre-trained DAE-GP resulting in lower fitness.
 
@@
 
-Figure \@ref(fig:altStrategies-finalFitness)
+Figure \@ref(fig:altStrategies-FitnessDistr) shows the distribution of best fitness achieved on the test set after running for 30 full generations.
+Both alternative second generation Pre-Training as well as grow-initialized Pre-Trainning show a much higher median RMSE (18.88/17.21) compared to standard pre-training (8.55) and regular DAE-GP (6.99).
+Also both approaches do not demonstrate any benefit on the dispersion of solutions found resulting in similiar standard deviations if compared to standard pre-training.
 
@@
 
-# Discussion
+# Conclusion and Discussion
 
-...
+The experiments conducted in this thesis did not find any significant benefits in using a pre-training strategy for using DAE-GP to solve real world symbolic regression problems.
+
+<!-- ## Vorteile -->
+
+<!-- Der Einsatz von Pre-Training dient: -->
+
+<!-- * der Erhöhung der Populationsdiversität (mögliche Strategie zur Kontrolle des Explore/Exploit Verhaltens von DAE-GP?) -->
+
+<!-- * der Reduktion der Trainingsepochen pro Generation (Reduktion der Rechenleistung, insbesondere bei komplexen DAE-LSTM nützlich?) -->
+
+<!-- ## Nachteile -->
+
+<!-- * DAE-GP verliert durch Einsatz von Pre-Training die Möglichkeit die Anzahl künstlicher Neuronen pro verstecktem Layer anzupassen -->
+
+<!-- * Erhöhung der Rechenzeit/Ressourcen: DAE-LSTMs müssen ausreichend dimensioniert sein (2+ Hidden Layer), aktuelle Publikationen nutzen nur einen einzelnen Hidden Layer^[@dae-gp_2022_symreg] ^[@daegp_explore_exploit].
+-->
+
+<!-- * Keine signifikanten Vorteile für die Qualität der gefundenen Lösungen -->
 

### 2023-01-28 paper.tex
Sat Jan 28 17:00:05 CET 2023: update-remote.sh

@@
 \author{}
-\date{\vspace{-2.5em}Date of Submission: 2023-01-27}
+\date{\vspace{-2.5em}Date of Submission: 2023-01-28}
 
@@
 \hline
-Sampling\_Steps & 2\\
+Sampling\_Steps & 1\\
+\hline
+Hidden\_Neurons & 150\\
 \hline
@@
 
-Regarding the hidden neurons, the dimension of each hidden layer, I used two different strategies: Regular DAE-GP uses the same strategy as described by earlier work (see {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]} or {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]}), the number of hidden neurons is dynamically set at each generation to the maximum length of all solutions inside the current training population.
-For pre-trained DAE-GP another strategy had to be used, I used a fixed number of hidden neurons for each run that is set equal to the maximum length of all individuals that are present in the initial pre-training population.
-I decided to take this aproach instead of using a fixed number of hidden neurons for the fact that it simplifies the adjustment of the parameter for later experiments on different problems.
-It should nonetheless be mentioned that this strategy does give pre-training on average more hidden neurons per layer resulting in a more complex network (since the pre-training population is much larger than the regular population, the chances for larger individuals inside it are also higher).
-
-\hypertarget{airfoil-dataset---2-hidden-layers}{%
-\subsubsection{Airfoil Dataset - 2 Hidden Layers}\label{airfoil-dataset---2-hidden-layers}}
-
 Figure \ref{fig:airfoil-median-fitness-30gens} shows both the mean and median of the best fitness for each generation on both the test and training set of the airfoil dataset for using 2 hidden layers inside the DAE-LSTM networks.
@@
 
-{\centering \includegraphics[width=0.9\linewidth]{./img/airfoil_2hl_maxIndSize_fullRun_30gens_withGP/mean_median_fitness_byGens}
+{\centering \includegraphics[width=0.9\linewidth]{./img/airfoil_2hl_150hn_fullRun_30gens_withGP/mean_median_fitness_byGens}
 
@@
 The distribution of final fitness scores for pre-trained and regular DAE-GP are visualized as boxplots in figure \ref{fig:airfoil-distribution-fitness-30gens}.
-The distribution for pre-trained DAE-GP shows to be preferable to that of regular DAE-GP on both sets since it shows a lower median (med.) as well as less dispersion measured by the standard deviation (std.).
-
-\begin{figure}[c]
-
-{\centering \includegraphics[width=0.8\linewidth]{./img/airfoil_2hl_maxIndSize_fullRun_30gens/final_fit_boxplot}
+The distribution for regular DAE-GP shows to be preferable to that of pre-trained DAE-GP on both sets since it shows a lower median RMSE as well as less dispersion of the solutions measured by the standard deviation.
+
+\begin{figure}[c]
+
+{\centering \includegraphics[width=0.8\linewidth]{./img/airfoil_2hl_150hn_fullRun_30gens/final_fit_boxplot}
 
@@
 It can be observed that both DAE-GP variants strongly reduce the average solutions size already in the first generations and that the average solution size inside the population constantly stays at this very low level for the rest of the search.
-Interestingly, the results also show that the pre-trained version of algorithm did produce smaller median size of the best solutions.
-
-\begin{figure}[c]
-
-{\centering \includegraphics[width=0.8\linewidth]{./img/airfoil_2hl_maxIndSize_fullRun_30gens/size_combined_median}
+Regarding the size of the best solution per generation, the results show that the pre-trained version of algorithm did produce solutions with the same median size as traditional DAE-GP.
+
+\begin{figure}[c]
+
+{\centering \includegraphics[width=0.8\linewidth]{./img/airfoil_2hl_150hn_fullRun_30gens/size_combined_median}
 
@@
 While the population diversity steadily decreases over the generations for standard GP, DAE-GP variants strongly increase the population diversity in the first generations and keep them at a high level of over \(0.9\).
-Again it can be observed, that pre-training shows a positive effect on the results by increasing population diversity in comparison to regular DAE-GP.
-
-\begin{figure}[c]
-
-{\centering \includegraphics[width=0.8\linewidth]{./img/airfoil_2hl_maxIndSize_fullRun_30gens_withGP/median_popDiversity_byGens_levOnly}
+Here it can be observed, that pre-training shows a positive effect on the results by increasing population diversity in comparison to regular DAE-GP.
+
+\begin{figure}[c]
+
+{\centering \includegraphics[width=0.8\linewidth]{./img/airfoil_2hl_150hn_fullRun_30gens_withGP/median_popDiversity_byGens_levOnly}
 
@@
 
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
-
-\ldots{}
+\hypertarget{alternative-pre-training-strategies}{%
+\subsection{Alternative Pre-Training Strategies}\label{alternative-pre-training-strategies}}
+
+Since pre-trained thus far did not demonstrate significant benefits for DAE-GP, this chapter describes two alternative pre-training strategies, Second Generation Pre-Training and Grow-Initialized Pre-Training, that are also tested for the airfoil dataset.
+
+One common observation in EC-based optimization algorithms is that the largest improvements in fitness are often occurring in the first generation of the search process.
+Second Generation Pre-Training tries to exploit this observation by shifting the pre-training phase from the start of the algorithm into the second generation with the aim to produce a pre-training Population \(\hat{P}\) that more accurately captures the properties of the search problem.
+The basic idea is to run basic DAE-GP for the first generation without any interference, after completing the first generation, the pre-training population \(\hat{P}\) is created by sampling the DAE-LSTM model \(M_1\).
+The pre-training model \(\hat{M}\) is then trained using \(\hat{P}\), after finishing DAE-GP continues running and initializes each new DAE-LSTM model \(M_g\) for \(g\in{2,...,g_{max}}\) using the trainable parameters \(\theta_{\hat{M}}\).
+
+The pre-training implementation in previous experiments used ramped half and half initialization {[}\protect\hyperlink{ref-Koza1993GeneticP}{8}{]} to create \(\hat{P}\), this leads to 50\% of the population consisting of full grown trees.
+An alternative method of initializing \(\hat{P}\) is to use the grow initialization method where solutions are not guaranteed to be full trees.
+This different structural composition of solutions inside the pre-training population might be another way to improve the pre-training strategy by training on a population that more closely resembles the structural properties of futher generations population where a lower percentage of full grown trees is to be expected.
+
+Figure \ref{fig:altStrategies-Fitness} shows the median best fitness by generations for both alternative pre-training strategies as well as the results for standard DAE-GP and the standard Pre-Training implementation used in earlier experiments.
+Unfortunately, both alternative strategies do show a worse performance in fitness than both DAE-GP and standard pre-trained DAE-GP resulting in lower fitness.
+
+\begin{figure}[c]
+
+{\centering \includegraphics[width=0.8\linewidth]{./img/PT_Strategies_airfoil_2hl_150hn_fullRun/median_fitness_byGens}
+
+}
+
+\caption{Best Fitness over 30 Generations - Alternative Pre-Trategies - Airfoil}\label{fig:altStrategies-Fitness}
+\end{figure}
+
+Figure \ref{fig:altStrategies-FitnessDistr} shows the distribution of best fitness achieved on the test set after running for 30 full generations.
+Both alternative second generation Pre-Training as well as grow-initialized Pre-Trainning show a much higher median RMSE (18.88/17.21) compared to standard pre-training (8.55) and regular DAE-GP (6.99).
+Also both approaches do not demonstrate any benefit on the dispersion of solutions found resulting in similiar standard deviations if compared to standard pre-training.
+
+\begin{figure}[c]
+
+{\centering \includegraphics[width=0.8\linewidth]{./img/PT_Strategies_airfoil_2hl_150hn_fullRun/final_fit_boxplot(test_only)}
+
+}
+
+\caption{Best Fitness after 30 Generations - Alternative Pre-Trategies - Airfoil}\label{fig:altStrategies-FitnessDistr}
+\end{figure}
+
+\hypertarget{conclusion-and-discussion}{%
+\section{Conclusion and Discussion}\label{conclusion-and-discussion}}
+
+The experiments conducted in this thesis did not find any significant benefits in using a pre-training strategy for using DAE-GP to solve real world symbolic regression problems.
 

### 2023-01-28 paper.Rmd
autogit

@@
 
-bash /Users/rmn/masterThesis/eda-gp-2020/experiments/airfoil_1hl_maxIndSize_fullRun_30gens/re_evaluate_airfoil_1hl_maxIndSize_30gen.sh
+bash /Users/rmn/masterThesis/eda-gp-2020/experiments/airfoil_1hl_150hn_fullRun_30gens/re_evaluate_airfoil_1hl_150hn_30gen.sh
 
-bash /Users/rmn/masterThesis/eda-gp-2020/experiments/airfoil_2hl_maxIndSize_fullRun_30gens/re_evaluate_airfoil_2hl_maxIndSize_30gen.sh
+# bash /Users/rmn/masterThesis/eda-gp-2020/experiments/airfoil_1hl_maxIndSize_fullRun_30gens/re_evaluate_airfoil_1hl_maxIndSize_30gen.sh
+
+# bash /Users/rmn/masterThesis/eda-gp-2020/experiments/airfoil_2hl_maxIndSize_fullRun_30gens/re_evaluate_airfoil_2hl_maxIndSize_30gen.sh
 
@@
 
-# Results
+# Results Airfoil Dataset
 
@@
 
-## Effect on search performance
+## Effect on solution quality
 
@@
 While the population diversity steadily decreases over the generations for standard GP, DAE-GP variants strongly increase the population diversity in the first generations and keep them at a high level of over $0.9$.
-Here it can be observed, that pre-training shows a positive effect on the results by increasing population diversity in comparison to regular DAE-GP.
+Here it can be observed, that pre-training shows a positive effect by increasing population diversity in comparison to regular DAE-GP.
 
@@
 Based on the findings of @pmlr-v5-erhan09a, I expected that pre-training DAE-GP would lead to a decrease in overall performance.
-Figures \@ref(fig:airfoil-fitness-perGen) and \@ref(fig:airfoil-finalFitness) show the average best fitness per generation as well as the distribution of final fitness scores after 30 generations.
+Figures \@ref(fig:airfoil-fitness-perGen) and \@ref(fig:airfoil-finalFitness) shows the average best fitness per generation as well as the distribution of final fitness scores after 30 generations.
 The results are, again, based on $10$ individual runs per algorithm.
 As expected it shows, that for a small DAE-LSTM size pre-training negativley impacts the fitness of solution found by DAE-GG.
+
+<!-- TODO: Change Path for Airfoil Single Layer to ...150hn... -->
 
@@
 
+## Effect on runtime
+
 ...
+
+# Results over different Real-World Symbolic Regression Problems
 
@@
 
-Since pre-trained thus far did not demonstrate significant benefits for DAE-GP, this chapter describes two alternative pre-training strategies, Second Generation Pre-Training and  Grow-Initialized Pre-Training, that are also tested for the airfoil dataset.
+Since pre-trained thus far did not demonstrate significant benefits for DAE-GP, this section describes two alternative pre-training strategies, Second Generation Pre-Training and  Grow-Initialized Pre-Training, that are also tested for the airfoil dataset.
 
-One common observation in EC-based optimization algorithms is that the largest improvements in fitness are often occurring in the first generation of the search process.
+One common observation in GP-based optimization algorithms is that the largest improvements in fitness are often occurring in the first generation of the search process.
 Second Generation Pre-Training tries to exploit this observation by shifting the pre-training phase from the start of the algorithm into the second generation with the aim to produce a pre-training Population $\hat{P}$ that more accurately captures the properties of the search problem.

### 2023-01-28 paper.tex
autogit

@@
 While the population diversity steadily decreases over the generations for standard GP, DAE-GP variants strongly increase the population diversity in the first generations and keep them at a high level of over \(0.9\).
-Here it can be observed, that pre-training shows a positive effect on the results by increasing population diversity in comparison to regular DAE-GP.
+Here it can be observed, that pre-training shows a positive effect by increasing population diversity in comparison to regular DAE-GP.
 
@@
 \end{figure}
-
-\hypertarget{airfoil-dataset---1-hidden-layer}{%
-\subsubsection{Airfoil Dataset - 1 Hidden Layer}\label{airfoil-dataset---1-hidden-layer}}
 
@@
 Based on the findings of {[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{3}{]}, I expected that pre-training DAE-GP would lead to a decrease in overall performance.
-Figures \ref{fig:airfoil-fitness-perGen} and \ref{fig:airfoil-finalFitness} show the average best fitness per generation as well as the distribution of final fitness scores after 30 generations.
+Figures \ref{fig:airfoil-fitness-perGen} and \ref{fig:airfoil-finalFitness} shows the average best fitness per generation as well as the distribution of final fitness scores after 30 generations.
 The results are, again, based on \(10\) individual runs per algorithm.
@@
 
-Since pre-trained thus far did not demonstrate significant benefits for DAE-GP, this chapter describes two alternative pre-training strategies, Second Generation Pre-Training and Grow-Initialized Pre-Training, that are also tested for the airfoil dataset.
-
-One common observation in EC-based optimization algorithms is that the largest improvements in fitness are often occurring in the first generation of the search process.
+Since pre-trained thus far did not demonstrate significant benefits for DAE-GP, this section describes two alternative pre-training strategies, Second Generation Pre-Training and Grow-Initialized Pre-Training, that are also tested for the airfoil dataset.
+
+One common observation in GP-based optimization algorithms is that the largest improvements in fitness are often occurring in the first generation of the search process.
 Second Generation Pre-Training tries to exploit this observation by shifting the pre-training phase from the start of the algorithm into the second generation with the aim to produce a pre-training Population \(\hat{P}\) that more accurately captures the properties of the search problem.

### 2023-01-29 paper.Rmd
Sun Jan 29 09:00:04 CET 2023: update-remote.sh

@@
   <center>Pre-Trained Denoising Autoencoders Long Short-Term Memory Networks as probabilistic Models for Estimation of Distribution Genetic Programming</center>
-  <center>Master Thesis - M.Sc. Business Education</center>
   <center>--------------------------------------------------------</center>
@@
   <center>--------------------------------------------------------</center>
-  <center>Master Thesis</center>
+  <center>Master Thesis - M.Sc. Business Education</center>
   <center>FB 03: Chair of Business Administration and Computer Science</center>

### 2023-01-29 paper.tex
Sun Jan 29 09:00:04 CET 2023: update-remote.sh

@@
 \author{}
-\date{\vspace{-2.5em}Date of Submission: 2023-01-28}
+\date{\vspace{-2.5em}Date of Submission: 2023-01-29}
 
@@
 
-\hypertarget{results}{%
-\section{Results}\label{results}}
+\hypertarget{results-airfoil-dataset}{%
+\section{Results Airfoil Dataset}\label{results-airfoil-dataset}}
 
@@
 
-\hypertarget{effect-on-search-performance}{%
-\subsection{Effect on search performance}\label{effect-on-search-performance}}
+\hypertarget{effect-on-solution-quality}{%
+\subsection{Effect on solution quality}\label{effect-on-solution-quality}}
 
@@
 
+\hypertarget{effect-on-runtime}{%
+\subsection{Effect on runtime}\label{effect-on-runtime}}
+
 \ldots{}
+
+\hypertarget{results-over-different-real-world-symbolic-regression-problems}{%
+\section{Results over different Real-World Symbolic Regression Problems}\label{results-over-different-real-world-symbolic-regression-problems}}
 

### 2023-01-29 paper.Rmd
Sun Jan 29 10:00:05 CET 2023: update-remote.sh

@@
 
-```{r altStrategies-Fitness, echo=FALSE, fig.cap="Best Fitness over 30 Generations - Alternative Pre-Trategies - Airfoil", out.width="80%"}
+```{r altStrategies-Fitness, echo=FALSE, fig.cap="Best Fitness over 30 Generations - Alternative Pre-Strategies - Airfoil", out.width="80%"}
 knitr::include_graphics("./img/PT_Strategies_airfoil_2hl_150hn_fullRun/median_fitness_byGens.png")
@@
 
-```{r altStrategies-FitnessDistr, echo=FALSE, fig.cap="Best Fitness after 30 Generations - Alternative Pre-Trategies - Airfoil", out.width="80%"}
+```{r altStrategies-FitnessDistr, echo=FALSE, fig.cap="Best Fitness on Test Set after 30 Generations - Alternative Pre-Trategies - Airfoil", out.width="80%"}
 knitr::include_graphics("./img/PT_Strategies_airfoil_2hl_150hn_fullRun/final_fit_boxplot(test_only).png")

### 2023-01-29 paper.tex
Sun Jan 29 10:00:05 CET 2023: update-remote.sh

@@
 
+--------------------------------------------------------
+
+Student: Roman Höhn
+
+Date of Birth: 1991-04-14
+
+Place of Birth: Wiesbaden, Hesse
+
+Student ID: 2712497
+
+Supervisor: David Wittenberg
+
+--------------------------------------------------------
+
 Master Thesis - M.Sc. Business Education
-
---------------------------------------------------------
-
-Student: Roman Höhn
-
-Date of Birth: 1991-04-14
-
-Place of Birth: Wiesbaden, Hesse
-
-Student ID: 2712497
-
-Supervisor: David Wittenberg
-
---------------------------------------------------------
-
-Master Thesis
 
@@
 
-\caption{Best Fitness over 30 Generations - Alternative Pre-Trategies - Airfoil}\label{fig:altStrategies-Fitness}
+\caption{Best Fitness over 30 Generations - Alternative Pre-Strategies - Airfoil}\label{fig:altStrategies-Fitness}
 \end{figure}
@@
 
-\caption{Best Fitness after 30 Generations - Alternative Pre-Trategies - Airfoil}\label{fig:altStrategies-FitnessDistr}
+\caption{Best Fitness on Test Set after 30 Generations - Alternative Pre-Trategies - Airfoil}\label{fig:altStrategies-FitnessDistr}
 \end{figure}

### 2023-01-29 paper.Rmd
Sun Jan 29 11:00:05 CET 2023: update-remote.sh

@@
 
-# Results Airfoil Dataset
+# Experiments
 
@@
 For each configuration I tested traditional DAE-GP as well as a pre-trained DAE-GP resulting in 12 total sub experiments that were each based on 10 individual runs (total number of runs=$120$).
-To avoid creating biased results through the presence of sampling error inside the pre-training population (see @sampling_err_gp), the population size for the pre-training phase is chosen very high with the size of $\hat{P}$ =20000 where 50% of $\hat{P}$ is used for the training population $\hat{P}_{train}$ and the remaining 50% are used for the hold-out validation population $\hat{P}_{test}$.
+To avoid creating biased results through the presence of sampling error inside the pre-training population (see @sampling_err_gp), the population size for the pre-training phase is chosen very high with the size of $\hat{P}$ =10000 where 50% of $\hat{P}$ is used for the training population $\hat{P}_{train}$ and the remaining 50% are used for the hold-out validation population $\hat{P}_{test}$.
 The training of DAE-LSTM uses a fixed number of 1000 epochs until termination and uses Adam optimization for gradient descent.
@@
 
-...
+Figure \@ref(fig:first-gen-airfoil-byNeurons) summarizes the median reconstruction error for a variable number of hidden neurons.
+As expected, increasing the number of hidden neurons leads to stronger overfitting to the training dataset which results in small training errors but growing testing errors.
+For the experiments run with 50 and 100 hidden neurons, a small benefit of using pre-training can be observed.
+Using only 50 hidden neurons leads to a smaller gap between training and testing error for the pre-trained DAE-LSTM.
+For 100 hidden neurons, regular DAE-GP not only starts to overfit to the training data earlier than pre-trained DAE-GP, also it can be seen that pre-trained DAE-GP starts to overfit at a lower reconstruction error which would result in a higher model performance if training strategies such as early stopping would be employed.
 
-```{r first-gen-airfoil-byNeurons, echo=FALSE, fig.cap="First Generation Median Training Error for variable number of hidden Neurons", out.width="90%"}
+```{r first-gen-airfoil-byNeurons, echo=FALSE, fig.cap="First Generation Median Reconstruction Error for variable number of hidden Neurons", out.width="90%"}
 knitr::include_graphics("./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_neurons_3plots.png")
@@
 
-```{r first-gen-airfoil-byLayers, echo=FALSE, fig.cap="First Generation Median Training Error for variable number of hidden Layers", out.width="90%"}
+```{r first-gen-airfoil-byLayers, echo=FALSE, fig.cap="First Generation Median Reconstruction Error for variable number of hidden Layers", out.width="90%"}
 knitr::include_graphics("./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_layers_3plots.png")

### 2023-01-29 paper.Rmd
Sun Jan 29 12:00:04 CET 2023: update-remote.sh

@@
 \setcounter{page}{1}
-\tableofcontents
+<!-- \tableofcontents -->
 \thispagestyle{empty}
@@
 
-Besides studying the performance for the airfoil dataset, additional real world symbolic regressions experiments used the datasets[@machine_learning_repo] described in table \@ref(tab:full-run-realWorldSymReg-problems).
-
-```{r full-run-realWorldSymReg-problems, echo=FALSE, message=FALSE, warning=FALSE}
-fig.pos="H"
-knitr::kable(
-  read.csv("./tables/full_run_realWorldSymReg_problems.csv"),
-  caption="Real World Symbolic Regression Benchmark Problems"
-)%>% row_spec(0,bold=TRUE ) %>% kableExtra::kable_styling(latex_options = "hold_position")
-```
-
-# Experiments
+# Results for the Airfoil Dataset
 
@@
 For 100 hidden neurons, regular DAE-GP not only starts to overfit to the training data earlier than pre-trained DAE-GP, also it can be seen that pre-trained DAE-GP starts to overfit at a lower reconstruction error which would result in a higher model performance if training strategies such as early stopping would be employed.
+Doubling the number of hidden neurons to 200 results in very strong overfitting behaviour for both DAE-GP variants starting around the 200.th epoch.
 
@@
 
-...
+The reults for the experiment with varying number of hidden layers are summarized in figure \@ref(fig:first-gen-airfoil-byLayers).
+Again, it can be observed that by increasing the dimension of the DAE-LSTM network both algorithms increasingly overfit to the training data.
+While for a single hidden layer overfitting begins at around the 400th.
+epoch, doubling the number of hidden layers to two leads to both algorithms overfitting at around the 200.th epoch.
+In general over all three sub experiments, both algorithms start to overfit after approximately the same number of epochs but pre-training leads to reaching the point of overfitting at a higher reconstruction error which would result in a inferrior model performance if the early stopping training strategy would be used during model training.
 
@@
 
-<!-- Regarding the hidden neurons, the dimension of each hidden layer, I used two different strategies: Regular DAE-GP uses the same strategy as described by earlier work (see [@dae-gp_2022_symreg] or [@dae-gp_2020_rtree]), the number of hidden neurons is dynamically set at each generation to the maximum length of all solutions inside the current training population.
-For pre-trained DAE-GP another strategy had to be used, I used a fixed number of hidden neurons for each run that is set equal to the maximum length of all individuals that are present in the initial pre-training population.
-I decided to take this aproach instead of using a fixed number of hidden neurons for the fact that it simplifies the adjustment of the parameter for later experiments on different problems.
-It should nonetheless be mentioned that this strategy does give pre-training on average more hidden neurons per layer resulting in a more complex network (since the pre-training population is much larger than the regular population, the chances for larger individuals inside it are also higher).
--->
-
 Figure \@ref(fig:airfoil-median-fitness-30gens) shows both the mean and median of the best fitness for each generation on both the test and training set of the airfoil dataset for using 2 hidden layers inside the DAE-LSTM networks.
@@
 
-The first interesting observation from this experiment is, that the pre-trained DAE-GP shows slightly lower mean and median performance than regular DAE-GP for both the training and the testing set in overall fitness.
-Regarding the presence of overfitting, both DAE-GP variants show only a very small gap between the testing and training fitness which indicates a good generalization behaviour.
+The first interesting observation from this experiment is, that the pre-trained DAE-GP shows slightly lower mean RMSE (higher fitness) but a higher median RMSE (lower fitness) than regular DAE-GP for both the training and the testing set in overall fitness.Regarding the presence of overfitting, both DAE-GP variants show only a very small gap between the testing and training fitness which indicates a good generalization behavior.
 
@@
 The distribution for regular DAE-GP shows to be preferable to that of pre-trained DAE-GP on both sets since it shows a lower median RMSE as well as less dispersion of the solutions measured by the standard deviation.
+An interesting observation from studying the distribution of final fitness scores for both algorithms is that for pre-trained DAE-GP the results
 
@@
 
+## Alternative Pre-Training Strategies
+
+Since pre-trained thus far did not demonstrate significant benefits for DAE-GP, this section describes two alternative pre-training strategies, Second Generation Pre-Training and  Grow-Initialized Pre-Training, that are also tested for the airfoil dataset.
+
+One common observation in GP-based optimization algorithms is that the largest improvements in fitness are often occurring in the first generation of the search process.
+Second Generation Pre-Training tries to exploit this observation by shifting the pre-training phase from the start of the algorithm into the second generation with the aim to produce a pre-training Population $\hat{P}$ that more accurately captures the properties of the search problem.
+The basic idea is to run basic DAE-GP for the first generation without any interference, after completing the first generation, the pre-training population $\hat{P}$ is created by sampling the DAE-LSTM model $M_1$.
+The pre-training model $\hat{M}$ is then trained using $\hat{P}$, after finishing DAE-GP continues running and initializes each new DAE-LSTM model $M_g$ for $g\in{2,...,g_{max}}$ using the trainable parameters $\theta_{\hat{M}}$.
+
+The pre-training implementation in previous experiments used ramped half and half initialization [@Koza1993GeneticP] to create $\hat{P}$, this leads to 50% of the population consisting of full grown trees.
+An alternative method of initializing $\hat{P}$ is to use the grow initialization method where solutions are not guaranteed to be full trees.
+This different structural composition of solutions inside the pre-training population might be another way to improve the pre-training strategy by training on a population that more closely resembles the structural properties of futher generations population where a lower percentage of full grown trees is to be expected.
+
+Figure \@ref(fig:altStrategies-Fitness) shows the median best fitness by generations for both alternative pre-training strategies as well as the results for standard DAE-GP and the standard Pre-Training implementation used in earlier experiments.
+Unfortunately, both alternative strategies do show a worse performance in fitness than both DAE-GP and standard pre-trained DAE-GP resulting in lower fitness.
+
+```{r altStrategies-Fitness, echo=FALSE, fig.cap="Best Fitness over 30 Generations - Alternative Pre-Strategies - Airfoil", out.width="80%"}
+knitr::include_graphics("./img/PT_Strategies_airfoil_2hl_150hn_fullRun/median_fitness_byGens.png")
+```
+
+Figure \@ref(fig:altStrategies-FitnessDistr) shows the distribution of best fitness achieved on the test set after running for 30 full generations.
+Both alternative second generation Pre-Training as well as grow-initialized Pre-Trainning show a much higher median RMSE (18.88/17.21) compared to standard pre-training (8.55) and regular DAE-GP (6.99).
+Also both approaches do not demonstrate any benefit on the dispersion of solutions found resulting in similiar standard deviations if compared to standard pre-training.
+
+```{r altStrategies-FitnessDistr, echo=FALSE, fig.cap="Best Fitness on Test Set after 30 Generations - Alternative Pre-Trategies - Airfoil", out.width="80%"}
+knitr::include_graphics("./img/PT_Strategies_airfoil_2hl_150hn_fullRun/final_fit_boxplot(test_only).png")
+```
+
 ## Effect on runtime
@@
 
-<!-- Overview Real World SymReg  -->
+To gather more confidence in the previous results, the pre-training strategy for DAE-GP was tested for three additional real world symbolic regression problems taken from @machine_learning_repo.
+Table \@ref(tab:full-run-realWorldSymReg-problems) briefly summarizes the benchmark problems.
 
-To gather more confidence in the previous results, the pre-training strategy for DAE-GP was tested for three additional real world symbolic regression problems taken from @machine_learning_repo.^[see table \@ref(tab:full-run-realWorldSymReg-problems)]
+```{r full-run-realWorldSymReg-problems, echo=FALSE, message=FALSE, warning=FALSE}
+fig.pos="H"
+knitr::kable(
+  read.csv("./tables/full_run_realWorldSymReg_problems.csv"),
+  caption="Real World Symbolic Regression Benchmark Problems"
+)%>% row_spec(0,bold=TRUE ) %>% kableExtra::kable_styling(latex_options = "hold_position")
+```
+
+<!-- Regarding the hidden neurons, the dimension of each hidden layer, I used two different strategies: Regular DAE-GP uses the same strategy as described by earlier work (see [@dae-gp_2022_symreg] or [@dae-gp_2020_rtree]), the number of hidden neurons is dynamically set at each generation to the maximum length of all solutions inside the current training population.
+For pre-trained DAE-GP another strategy had to be used, I used a fixed number of hidden neurons for each run that is set equal to the maximum length of all individuals that are present in the initial pre-training population.
+I decided to take this aproach instead of using a fixed number of hidden neurons for the fact that it simplifies the adjustment of the parameter for later experiments on different problems.
+It should nonetheless be mentioned that this strategy does give pre-training on average more hidden neurons per layer resulting in a more complex network (since the pre-training population is much larger than the regular population, the chances for larger individuals inside it are also higher).
+-->
 
@@
 
-## Alternative Pre-Training Strategies
-
-Since pre-trained thus far did not demonstrate significant benefits for DAE-GP, this section describes two alternative pre-training strategies, Second Generation Pre-Training and  Grow-Initialized Pre-Training, that are also tested for the airfoil dataset.
-
-One common observation in GP-based optimization algorithms is that the largest improvements in fitness are often occurring in the first generation of the search process.
-Second Generation Pre-Training tries to exploit this observation by shifting the pre-training phase from the start of the algorithm into the second generation with the aim to produce a pre-training Population $\hat{P}$ that more accurately captures the properties of the search problem.
-The basic idea is to run basic DAE-GP for the first generation without any interference, after completing the first generation, the pre-training population $\hat{P}$ is created by sampling the DAE-LSTM model $M_1$.
-The pre-training model $\hat{M}$ is then trained using $\hat{P}$, after finishing DAE-GP continues running and initializes each new DAE-LSTM model $M_g$ for $g\in{2,...,g_{max}}$ using the trainable parameters $\theta_{\hat{M}}$.
-
-The pre-training implementation in previous experiments used ramped half and half initialization [@Koza1993GeneticP] to create $\hat{P}$, this leads to 50% of the population consisting of full grown trees.
-An alternative method of initializing $\hat{P}$ is to use the grow initialization method where solutions are not guaranteed to be full trees.
-This different structural composition of solutions inside the pre-training population might be another way to improve the pre-training strategy by training on a population that more closely resembles the structural properties of futher generations population where a lower percentage of full grown trees is to be expected.
-
-Figure \@ref(fig:altStrategies-Fitness) shows the median best fitness by generations for both alternative pre-training strategies as well as the results for standard DAE-GP and the standard Pre-Training implementation used in earlier experiments.
-Unfortunately, both alternative strategies do show a worse performance in fitness than both DAE-GP and standard pre-trained DAE-GP resulting in lower fitness.
-
-```{r altStrategies-Fitness, echo=FALSE, fig.cap="Best Fitness over 30 Generations - Alternative Pre-Strategies - Airfoil", out.width="80%"}
-knitr::include_graphics("./img/PT_Strategies_airfoil_2hl_150hn_fullRun/median_fitness_byGens.png")
-```
-
-Figure \@ref(fig:altStrategies-FitnessDistr) shows the distribution of best fitness achieved on the test set after running for 30 full generations.
-Both alternative second generation Pre-Training as well as grow-initialized Pre-Trainning show a much higher median RMSE (18.88/17.21) compared to standard pre-training (8.55) and regular DAE-GP (6.99).
-Also both approaches do not demonstrate any benefit on the dispersion of solutions found resulting in similiar standard deviations if compared to standard pre-training.
-
-```{r altStrategies-FitnessDistr, echo=FALSE, fig.cap="Best Fitness on Test Set after 30 Generations - Alternative Pre-Trategies - Airfoil", out.width="80%"}
-knitr::include_graphics("./img/PT_Strategies_airfoil_2hl_150hn_fullRun/final_fit_boxplot(test_only).png")
-```
-
 # Conclusion and Discussion

### 2023-01-29 paper.tex
Sun Jan 29 12:00:04 CET 2023: update-remote.sh

@@
 \setcounter{page}{1}
-\tableofcontents
+
 \thispagestyle{empty}
@@
 
-Besides studying the performance for the airfoil dataset, additional real world symbolic regressions experiments used the datasets{[}\protect\hyperlink{ref-machine_learning_repo}{2}{]} described in table \ref{tab:full-run-realWorldSymReg-problems}.
-
-\begin{table}[!h]
-
-\caption{\label{tab:full-run-realWorldSymReg-problems}Real World Symbolic Regression Benchmark Problems}
-\centering
-\begin{tabular}[t]{l|r|r}
-\hline
-\textbf{Problem} & \textbf{Observations} & \textbf{Features}\\
-\hline
-Airfoil & 1503 & 5\\
-\hline
-Boston\_Housing & 506 & 13\\
-\hline
-Energy\_Cooling & 768 & 8\\
-\hline
-Concrete & 1030 & 8\\
-\hline
-\end{tabular}
-\end{table}
-
-\hypertarget{results-airfoil-dataset}{%
-\section{Results Airfoil Dataset}\label{results-airfoil-dataset}}
+\hypertarget{results-for-the-airfoil-dataset}{%
+\section{Results for the Airfoil Dataset}\label{results-for-the-airfoil-dataset}}
 
@@
 For each configuration I tested traditional DAE-GP as well as a pre-trained DAE-GP resulting in 12 total sub experiments that were each based on 10 individual runs (total number of runs=\(120\)).
-To avoid creating biased results through the presence of sampling error inside the pre-training population (see {[}\protect\hyperlink{ref-sampling_err_gp}{13}{]}), the population size for the pre-training phase is chosen very high with the size of \(\hat{P}\) =20000 where 50\% of \(\hat{P}\) is used for the training population \(\hat{P}_{train}\) and the remaining 50\% are used for the hold-out validation population \(\hat{P}_{test}\).
+To avoid creating biased results through the presence of sampling error inside the pre-training population (see {[}\protect\hyperlink{ref-sampling_err_gp}{13}{]}), the population size for the pre-training phase is chosen very high with the size of \(\hat{P}\) =10000 where 50\% of \(\hat{P}\) is used for the training population \(\hat{P}_{train}\) and the remaining 50\% are used for the hold-out validation population \(\hat{P}_{test}\).
 The training of DAE-LSTM uses a fixed number of 1000 epochs until termination and uses Adam optimization for gradient descent.
@@
 
-\ldots{}
+Figure \ref{fig:first-gen-airfoil-byNeurons} summarizes the median reconstruction error for a variable number of hidden neurons.
+As expected, increasing the number of hidden neurons leads to stronger overfitting to the training dataset which results in small training errors but growing testing errors.
+For the experiments run with 50 and 100 hidden neurons, a small benefit of using pre-training can be observed.
+Using only 50 hidden neurons leads to a smaller gap between training and testing error for the pre-trained DAE-LSTM.
+For 100 hidden neurons, regular DAE-GP not only starts to overfit to the training data earlier than pre-trained DAE-GP, also it can be seen that pre-trained DAE-GP starts to overfit at a lower reconstruction error which would result in a higher model performance if training strategies such as early stopping would be employed.
+Doubling the number of hidden neurons to 200 results in very strong overfitting behaviour for both DAE-GP variants starting around the 200.th epoch.
 
@@
 
-\caption{First Generation Median Training Error for variable number of hidden Neurons}\label{fig:first-gen-airfoil-byNeurons}
-\end{figure}
-
-\ldots{}
+\caption{First Generation Median Reconstruction Error for variable number of hidden Neurons}\label{fig:first-gen-airfoil-byNeurons}
+\end{figure}
+
+The reults for the experiment with varying number of hidden layers are summarized in figure \ref{fig:first-gen-airfoil-byLayers}.
+Again, it can be observed that by increasing the dimension of the DAE-LSTM network both algorithms increasingly overfit to the training data.
+While for a single hidden layer overfitting begins at around the 400th.
+epoch, doubling the number of hidden layers to two leads to both algorithms overfitting at around the 200.th epoch.
+In general over all three sub experiments, both algorithms start to overfit after approximately the same number of epochs but pre-training leads to reaching the point of overfitting at a higher reconstruction error which would result in a inferrior model performance if the early stopping training strategy would be used during model training.
 
@@
 
-\caption{First Generation Median Training Error for variable number of hidden Layers}\label{fig:first-gen-airfoil-byLayers}
+\caption{First Generation Median Reconstruction Error for variable number of hidden Layers}\label{fig:first-gen-airfoil-byLayers}
 \end{figure}
@@
 
+\hypertarget{alternative-pre-training-strategies}{%
+\subsection{Alternative Pre-Training Strategies}\label{alternative-pre-training-strategies}}
+
+Since pre-trained thus far did not demonstrate significant benefits for DAE-GP, this section describes two alternative pre-training strategies, Second Generation Pre-Training and Grow-Initialized Pre-Training, that are also tested for the airfoil dataset.
+
+One common observation in GP-based optimization algorithms is that the largest improvements in fitness are often occurring in the first generation of the search process.
+Second Generation Pre-Training tries to exploit this observation by shifting the pre-training phase from the start of the algorithm into the second generation with the aim to produce a pre-training Population \(\hat{P}\) that more accurately captures the properties of the search problem.
+The basic idea is to run basic DAE-GP for the first generation without any interference, after completing the first generation, the pre-training population \(\hat{P}\) is created by sampling the DAE-LSTM model \(M_1\).
+The pre-training model \(\hat{M}\) is then trained using \(\hat{P}\), after finishing DAE-GP continues running and initializes each new DAE-LSTM model \(M_g\) for \(g\in{2,...,g_{max}}\) using the trainable parameters \(\theta_{\hat{M}}\).
+
+The pre-training implementation in previous experiments used ramped half and half initialization {[}\protect\hyperlink{ref-Koza1993GeneticP}{8}{]} to create \(\hat{P}\), this leads to 50\% of the population consisting of full grown trees.
+An alternative method of initializing \(\hat{P}\) is to use the grow initialization method where solutions are not guaranteed to be full trees.
+This different structural composition of solutions inside the pre-training population might be another way to improve the pre-training strategy by training on a population that more closely resembles the structural properties of futher generations population where a lower percentage of full grown trees is to be expected.
+
+Figure \ref{fig:altStrategies-Fitness} shows the median best fitness by generations for both alternative pre-training strategies as well as the results for standard DAE-GP and the standard Pre-Training implementation used in earlier experiments.
+Unfortunately, both alternative strategies do show a worse performance in fitness than both DAE-GP and standard pre-trained DAE-GP resulting in lower fitness.
+
+\begin{figure}[c]
+
+{\centering \includegraphics[width=0.8\linewidth]{./img/PT_Strategies_airfoil_2hl_150hn_fullRun/median_fitness_byGens}
+
+}
+
+\caption{Best Fitness over 30 Generations - Alternative Pre-Strategies - Airfoil}\label{fig:altStrategies-Fitness}
+\end{figure}
+
+Figure \ref{fig:altStrategies-FitnessDistr} shows the distribution of best fitness achieved on the test set after running for 30 full generations.
+Both alternative second generation Pre-Training as well as grow-initialized Pre-Trainning show a much higher median RMSE (18.88/17.21) compared to standard pre-training (8.55) and regular DAE-GP (6.99).
+Also both approaches do not demonstrate any benefit on the dispersion of solutions found resulting in similiar standard deviations if compared to standard pre-training.
+
+\begin{figure}[c]
+
+{\centering \includegraphics[width=0.8\linewidth]{./img/PT_Strategies_airfoil_2hl_150hn_fullRun/final_fit_boxplot(test_only)}
+
+}
+
+\caption{Best Fitness on Test Set after 30 Generations - Alternative Pre-Trategies - Airfoil}\label{fig:altStrategies-FitnessDistr}
+\end{figure}
+
 \hypertarget{effect-on-runtime}{%
@@
 
-To gather more confidence in the previous results, the pre-training strategy for DAE-GP was tested for three additional real world symbolic regression problems taken from {[}\protect\hyperlink{ref-machine_learning_repo}{2}{]}.\footnote{see table \ref{tab:full-run-realWorldSymReg-problems}}
+To gather more confidence in the previous results, the pre-training strategy for DAE-GP was tested for three additional real world symbolic regression problems taken from {[}\protect\hyperlink{ref-machine_learning_repo}{2}{]}.
+Table \ref{tab:full-run-realWorldSymReg-problems} briefly summarizes the benchmark problems.
+
+\begin{table}[!h]
+
+\caption{\label{tab:full-run-realWorldSymReg-problems}Real World Symbolic Regression Benchmark Problems}
+\centering
+\begin{tabular}[t]{l|r|r}
+\hline
+\textbf{Problem} & \textbf{Observations} & \textbf{Features}\\
+\hline
+Airfoil & 1503 & 5\\
+\hline
+Boston\_Housing & 506 & 13\\
+\hline
+Energy\_Cooling & 768 & 8\\
+\hline
+Concrete & 1030 & 8\\
+\hline
+\end{tabular}
+\end{table}
 
@@
 
-\hypertarget{alternative-pre-training-strategies}{%
-\subsection{Alternative Pre-Training Strategies}\label{alternative-pre-training-strategies}}
-
-Since pre-trained thus far did not demonstrate significant benefits for DAE-GP, this section describes two alternative pre-training strategies, Second Generation Pre-Training and Grow-Initialized Pre-Training, that are also tested for the airfoil dataset.
-
-One common observation in GP-based optimization algorithms is that the largest improvements in fitness are often occurring in the first generation of the search process.
-Second Generation Pre-Training tries to exploit this observation by shifting the pre-training phase from the start of the algorithm into the second generation with the aim to produce a pre-training Population \(\hat{P}\) that more accurately captures the properties of the search problem.
-The basic idea is to run basic DAE-GP for the first generation without any interference, after completing the first generation, the pre-training population \(\hat{P}\) is created by sampling the DAE-LSTM model \(M_1\).
-The pre-training model \(\hat{M}\) is then trained using \(\hat{P}\), after finishing DAE-GP continues running and initializes each new DAE-LSTM model \(M_g\) for \(g\in{2,...,g_{max}}\) using the trainable parameters \(\theta_{\hat{M}}\).
-
-The pre-training implementation in previous experiments used ramped half and half initialization {[}\protect\hyperlink{ref-Koza1993GeneticP}{8}{]} to create \(\hat{P}\), this leads to 50\% of the population consisting of full grown trees.
-An alternative method of initializing \(\hat{P}\) is to use the grow initialization method where solutions are not guaranteed to be full trees.
-This different structural composition of solutions inside the pre-training population might be another way to improve the pre-training strategy by training on a population that more closely resembles the structural properties of futher generations population where a lower percentage of full grown trees is to be expected.
-
-Figure \ref{fig:altStrategies-Fitness} shows the median best fitness by generations for both alternative pre-training strategies as well as the results for standard DAE-GP and the standard Pre-Training implementation used in earlier experiments.
-Unfortunately, both alternative strategies do show a worse performance in fitness than both DAE-GP and standard pre-trained DAE-GP resulting in lower fitness.
-
-\begin{figure}[c]
-
-{\centering \includegraphics[width=0.8\linewidth]{./img/PT_Strategies_airfoil_2hl_150hn_fullRun/median_fitness_byGens}
-
-}
-
-\caption{Best Fitness over 30 Generations - Alternative Pre-Strategies - Airfoil}\label{fig:altStrategies-Fitness}
-\end{figure}
-
-Figure \ref{fig:altStrategies-FitnessDistr} shows the distribution of best fitness achieved on the test set after running for 30 full generations.
-Both alternative second generation Pre-Training as well as grow-initialized Pre-Trainning show a much higher median RMSE (18.88/17.21) compared to standard pre-training (8.55) and regular DAE-GP (6.99).
-Also both approaches do not demonstrate any benefit on the dispersion of solutions found resulting in similiar standard deviations if compared to standard pre-training.
-
-\begin{figure}[c]
-
-{\centering \includegraphics[width=0.8\linewidth]{./img/PT_Strategies_airfoil_2hl_150hn_fullRun/final_fit_boxplot(test_only)}
-
-}
-
-\caption{Best Fitness on Test Set after 30 Generations - Alternative Pre-Trategies - Airfoil}\label{fig:altStrategies-FitnessDistr}
-\end{figure}
-
 \hypertarget{conclusion-and-discussion}{%

### 2023-01-29 paper.Rmd
Sun Jan 29 13:00:05 CET 2023: update-remote.sh

@@
 
+```{r airfoil-median-fitness-30gens, echo=FALSE, fig.cap="Best Fitness over 30 Generations - Airfoil (2HL)", out.width="90%"}
+knitr::include_graphics("./img/airfoil_2hl_150hn_fullRun_30gens_withGP/mean_median_fitness_byGens.png")
+```
+
 Figure \@ref(fig:airfoil-median-fitness-30gens) shows both the mean and median of the best fitness for each generation on both the test and training set of the airfoil dataset for using 2 hidden layers inside the DAE-LSTM networks.
@@
 
-```{r airfoil-median-fitness-30gens, echo=FALSE, fig.cap="Best Fitness over 30 Generations - Airfoil (2HL)", out.width="90%"}
-knitr::include_graphics("./img/airfoil_2hl_150hn_fullRun_30gens_withGP/mean_median_fitness_byGens.png")
-```
-The distribution of final fitness scores for pre-trained and regular DAE-GP are visualized as boxplots in figure  \@ref(fig:airfoil-distribution-fitness-30gens).
-The distribution for regular DAE-GP shows to be preferable to that of pre-trained DAE-GP on both sets since it shows a lower median RMSE as well as less dispersion of the solutions measured by the standard deviation.
-An interesting observation from studying the distribution of final fitness scores for both algorithms is that for pre-trained DAE-GP the results
-
 ```{r airfoil-distribution-fitness-30gens, echo=FALSE, fig.cap="Best Fitness after 30 Generations - Airfoil (2HL)", out.width="80%"}
@@
 ```
+
+The distribution of final fitness scores for pre-trained and regular DAE-GP are visualized as boxplots in figure  \@ref(fig:airfoil-distribution-fitness-30gens).
+The distribution for regular DAE-GP shows to be preferable to that of pre-trained DAE-GP on both sets since it shows a lower median RMSE as well as less dispersion of the solutions measured by the standard deviation.
+An interesting observation from studying the distribution of final fitness scores for both algorithms is that while both algorithms have a very similiar interquartile range (similiar box shape), pre-trained DAE-GP has more extreme outliers for high RMSE which explains the slightly worse performance if looking at median best fitness present in figure \@ref(fig:airfoil-median-fitness-30gens).
 

### 2023-01-29 paper.tex
Sun Jan 29 13:00:05 CET 2023: update-remote.sh

@@
 
+\begin{figure}[c]
+
+{\centering \includegraphics[width=0.9\linewidth]{./img/airfoil_2hl_150hn_fullRun_30gens_withGP/mean_median_fitness_byGens}
+
+}
+
+\caption{Best Fitness over 30 Generations - Airfoil (2HL)}\label{fig:airfoil-median-fitness-30gens}
+\end{figure}
+
 Figure \ref{fig:airfoil-median-fitness-30gens} shows both the mean and median of the best fitness for each generation on both the test and training set of the airfoil dataset for using 2 hidden layers inside the DAE-LSTM networks.
@@
 
-The first interesting observation from this experiment is, that the pre-trained DAE-GP shows slightly lower mean and median performance than regular DAE-GP for both the training and the testing set in overall fitness.
-Regarding the presence of overfitting, both DAE-GP variants show only a very small gap between the testing and training fitness which indicates a good generalization behaviour.
-
-\begin{figure}[c]
-
-{\centering \includegraphics[width=0.9\linewidth]{./img/airfoil_2hl_150hn_fullRun_30gens_withGP/mean_median_fitness_byGens}
-
-}
-
-\caption{Best Fitness over 30 Generations - Airfoil (2HL)}\label{fig:airfoil-median-fitness-30gens}
+The first interesting observation from this experiment is, that the pre-trained DAE-GP shows slightly lower mean RMSE (higher fitness) but a higher median RMSE (lower fitness) than regular DAE-GP for both the training and the testing set in overall fitness.Regarding the presence of overfitting, both DAE-GP variants show only a very small gap between the testing and training fitness which indicates a good generalization behavior.
+
+\begin{figure}[c]
+
+{\centering \includegraphics[width=0.8\linewidth]{./img/airfoil_2hl_150hn_fullRun_30gens/final_fit_boxplot}
+
+}
+
+\caption{Best Fitness after 30 Generations - Airfoil (2HL)}\label{fig:airfoil-distribution-fitness-30gens}
 \end{figure}
@@
 The distribution for regular DAE-GP shows to be preferable to that of pre-trained DAE-GP on both sets since it shows a lower median RMSE as well as less dispersion of the solutions measured by the standard deviation.
-
-\begin{figure}[c]
-
-{\centering \includegraphics[width=0.8\linewidth]{./img/airfoil_2hl_150hn_fullRun_30gens/final_fit_boxplot}
-
-}
-
-\caption{Best Fitness after 30 Generations - Airfoil (2HL)}\label{fig:airfoil-distribution-fitness-30gens}
-\end{figure}
+An interesting observation from studying the distribution of final fitness scores for both algorithms is that while both algorithms have a very similiar interquartile range (similiar box shape), pre-trained DAE-GP has more extreme outliers for high RMSE which explains the slightly worse performance if looking at median best fitness present in figure \ref{fig:airfoil-median-fitness-30gens}.
 

### 2023-01-30 paper.Rmd
Mon Jan 30 21:00:07 CET 2023: update-remote.sh

@@
 
+### Reduced Number of Hidden Layers
+
 Since previous research on DAE-GP used a single hidden layers for the airfoil dataset [@dae-gp_2022_symreg], I repeated the experiment with a single hidden layer.
@@
 
-<!-- TODO: Change Path for Airfoil Single Layer to ...150hn... -->
-
 ```{r airfoil-fitness-perGen, echo=FALSE, fig.cap="Fitness over 30 Generations - Airfoil (1HL)", out.width="80%"}
-knitr::include_graphics("./img/airfoil_1hl_maxIndSize_fullRun_30gens/mean_median_fitness_byGens.png")
+knitr::include_graphics("./img/airfoil_1hl_150hn_30gens/mean_median_fitness_byGens.png")
 ```
@@
 ```{r airfoil-finalFitness, echo=FALSE, fig.cap="Fitness after 30 Generations - Airfoil (1HL)", out.width="80%"}
-knitr::include_graphics("./img/airfoil_1hl_maxIndSize_fullRun_30gens/final_fit_boxplot.png")
+knitr::include_graphics("./img/airfoil_1hl_150hn_fullRun_30gens/final_fit_boxplot.png")
 ```
 
-## Alternative Pre-Training Strategies
+### Alternative Pre-Training Strategies
 
@@
 
-...
+<!-- describe total runtime, including/excluding pre-training time -->
+
+```{r airfoil-runtime-total, echo=FALSE, fig.cap="Median Runtime - Airfoil", out.width="80%"}
+knitr::include_graphics("./img/airfoil_2hl_150hn_fullRun_30gens_TimeBenchmark/median_runtime.png")
+```
+
+```{r airfoil-trainingEpochs, echo=FALSE, fig.cap="Median Number of Training Epochs per Generation - Airfoil", out.width="80%"}
+knitr::include_graphics("./img/airfoil_2hl_150hn_fullRun_30gens_TimeBenchmark/median_training_epochs.png")
+```
+
+```{r airfoil-SamplingTime, echo=FALSE, fig.cap="Median Sampling Time per Generation - Airfoil", out.width="80%"}
+knitr::include_graphics("./img/airfoil_2hl_150hn_fullRun_30gens_TimeBenchmark/median_sampleTime.png")
+```
+
+```{r airfoil-runtime-noSampling-total, echo=FALSE, fig.cap="Median Runtime excluding Time for Sampling - Airfoil", out.width="80%"}
+knitr::include_graphics("./img/airfoil_2hl_150hn_fullRun_30gens_TimeBenchmark/median_runtime_withoutSampling.png")
+```
 

### 2023-01-30 paper.tex
Mon Jan 30 21:00:07 CET 2023: update-remote.sh

@@
 \author{}
-\date{\vspace{-2.5em}Date of Submission: 2023-01-29}
+\date{\vspace{-2.5em}Date of Submission: 2023-01-30}
 
@@
 
-\ldots{}
+\begin{figure}[c]
+
+{\centering \includegraphics[width=0.8\linewidth]{./img/airfoil_2hl_150hn_fullRun_30gens_TimeBenchmark/median_runtime}
+
+}
+
+\caption{Median Runtime - Airfoil}\label{fig:airfoil-runtime-total}
+\end{figure}
+
+\begin{figure}[c]
+
+{\centering \includegraphics[width=0.8\linewidth]{./img/airfoil_2hl_150hn_fullRun_30gens_TimeBenchmark/median_training_epochs}
+
+}
+
+\caption{Median Number of Training Epochs per Generation - Airfoil}\label{fig:airfoil-trainingEpochs}
+\end{figure}
+
+\begin{figure}[c]
+
+{\centering \includegraphics[width=0.8\linewidth]{./img/airfoil_2hl_150hn_fullRun_30gens_TimeBenchmark/median_sampleTime}
+
+}
+
+\caption{Median Sampling Time per Generation - Airfoil}\label{fig:airfoil-SamplingTime}
+\end{figure}
+
+\begin{figure}[c]
+
+{\centering \includegraphics[width=0.8\linewidth]{./img/airfoil_2hl_150hn_fullRun_30gens_TimeBenchmark/median_runtime_withoutSampling}
+
+}
+
+\caption{Median Runtime excluding Time for Sampling - Airfoil}\label{fig:airfoil-runtime-noSampling-total}
+\end{figure}
 

### 2023-01-30 paper.Rmd
autogit

@@
 ```{r airfoil-finalFitness, echo=FALSE, fig.cap="Fitness after 30 Generations - Airfoil (1HL)", out.width="80%"}
-knitr::include_graphics("./img/airfoil_1hl_150hn_fullRun_30gens/final_fit_boxplot.png")
+knitr::include_graphics("./img/airfoil_1hl_150hn_30gens/final_fit_boxplot.png")
 ```
@@
 
-Since pre-trained thus far did not demonstrate significant benefits for DAE-GP, this section describes two alternative pre-training strategies, Second Generation Pre-Training and  Grow-Initialized Pre-Training, that are also tested for the airfoil dataset.
+Since pre-training thus far did not demonstrate significant benefits for the quality of solutions found by DAE-GP, this section describes two alternative pre-training strategies, Second Generation Pre-Training and  Grow-Initialized Pre-Training, that are also tested for the airfoil dataset.
 
@@
 
-<!-- describe total runtime, including/excluding pre-training time -->
+One important argument for using pre-training DAE-GP is that it might be a potential way to reduce the overall run-time of the algorithm by reducing the training time of each generations DAE-LSTM model.
+Especially in the context of optimization problems where solutions have to be found numerous times in a time efficient manner (e.g.
+Vehicle Routing in the Domain of Logistic Services), one major advantage of pre-training is that a pre-trained model can be re-used to optimize many instances of an optimization problem.
 

### 2023-01-30 paper.tex
autogit

@@
 
+\hypertarget{reduced-number-of-hidden-layers}{%
+\subsubsection{Reduced Number of Hidden Layers}\label{reduced-number-of-hidden-layers}}
+
 Since previous research on DAE-GP used a single hidden layers for the airfoil dataset {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]}, I repeated the experiment with a single hidden layer.
@@
 
-{\centering \includegraphics[width=0.8\linewidth]{./img/airfoil_1hl_maxIndSize_fullRun_30gens/mean_median_fitness_byGens}
+{\centering \includegraphics[width=0.8\linewidth]{./img/airfoil_1hl_150hn_30gens/mean_median_fitness_byGens}
 
@@
 
-{\centering \includegraphics[width=0.8\linewidth]{./img/airfoil_1hl_maxIndSize_fullRun_30gens/final_fit_boxplot}
+{\centering \includegraphics[width=0.8\linewidth]{./img/airfoil_1hl_150hn_30gens/final_fit_boxplot}
 
@@
 \hypertarget{alternative-pre-training-strategies}{%
-\subsection{Alternative Pre-Training Strategies}\label{alternative-pre-training-strategies}}
-
-Since pre-trained thus far did not demonstrate significant benefits for DAE-GP, this section describes two alternative pre-training strategies, Second Generation Pre-Training and Grow-Initialized Pre-Training, that are also tested for the airfoil dataset.
+\subsubsection{Alternative Pre-Training Strategies}\label{alternative-pre-training-strategies}}
+
+Since pre-training thus far did not demonstrate significant benefits for the quality of solutions found by DAE-GP, this section describes two alternative pre-training strategies, Second Generation Pre-Training and Grow-Initialized Pre-Training, that are also tested for the airfoil dataset.
 

### 2023-01-30 paper.Rmd
autogit
Mon Jan 30 23:00:10 CET 2023: update-remote.sh

@@
 
-One important argument for using pre-training DAE-GP is that it might be a potential way to reduce the overall run-time of the algorithm by reducing the training time of each generations DAE-LSTM model.
+One important argument for using pre-training with DAE-GP is that it might be a potential way to reduce the overall run-time of the algorithm by reducing the training time of each generation's DAE-LSTM model.
 Especially in the context of optimization problems where solutions have to be found numerous times in a time efficient manner (e.g.
 Vehicle Routing in the Domain of Logistic Services), one major advantage of pre-training is that a pre-trained model can be re-used to optimize many instances of an optimization problem.
+
+Unfortunately, as shown in figure \@ref(fig:airfoil-runtime-total), for the experiment conducted on the airfoil dataset with 2 hidden layers a negative impact of pre-training can be observed for the DAE-GP runtime.
 

### 2023-01-31 paper.Rmd
autogit

@@
 
-Unfortunately, as shown in figure \@ref(fig:airfoil-runtime-total), for the experiment conducted on the airfoil dataset with 2 hidden layers a negative impact of pre-training can be observed for the DAE-GP runtime.
+Unfortunately, as shown in figure \@ref(fig:airfoil-runtime-total), for the experiment conducted on the airfoil dataset with 2 hidden layers a strong negative impact of pre-training can be observed for the median runtime of DAE-GP.
+If the time spend on pre-training is included into the total runtime, the median duration for pre-trained DAE-GP over all 10 runs is $14706.465$ seconds as opposed to regular DAE-GP that has a median runtime of $5076.448.
+Even if the time spend
 

### 2023-01-31 paper.Rmd
Tue Jan 31 10:00:05 CET 2023: update-remote.sh

@@
 
+```{r python_runtime, echo=FALSE}
+REG_TIME=5076.448
+PT_TIME=14706.465
+PT_TIME_NOSAMPLE=12536.889
+
+```
+
 Unfortunately, as shown in figure \@ref(fig:airfoil-runtime-total), for the experiment conducted on the airfoil dataset with 2 hidden layers a strong negative impact of pre-training can be observed for the median runtime of DAE-GP.
-If the time spend on pre-training is included into the total runtime, the median duration for pre-trained DAE-GP over all 10 runs is $14706.465$ seconds as opposed to regular DAE-GP that has a median runtime of $5076.448.
-Even if the time spend
+If the time spend on pre-training is included into the total runtime, the median duration for pre-trained DAE-GP over all 10 runs almost tripples from a median runtime of $5076.448$ for regular DAE-GP to $14706.465$ seconds for the pre-trained algorithm.
+Even in the case where the time spend for pre-training is excluded from the total runtime, pre-training still increases the runtime of DAE-GP to $12536.889$ seconds which
 

### 2023-01-31 paper.Rmd
autogit

--- 2023-01-30 paper.Rmd
@@
 
-Unfortunately, as shown in figure \@ref(fig:airfoil-runtime-total), for the experiment conducted on the airfoil dataset with 2 hidden layers a negative impact of pre-training can be observed for the DAE-GP runtime.
+<!-- TODO: Disclaimer, Runtime is implementation dependant, no GPU use, etc... -->
+
+```{r runtime_vars, echo=FALSE}
+
+# Values are copied from file /eda-gp-2020/experiments/airfoil_2hl_150hn_fullRun_30gens/airfoil_2hl_150hn_fullRun_30gens_TimeBenchmark.ipynb
+
+REG_TIME <- 5076.448
+PT_TIME <- 14706.465
+PT_TIME_NOSAMPLE <- 12536.889
+```
+
+Suprinsingly, as shown in figure \@ref(fig:airfoil-runtime-total), for the experiment conducted on the airfoil dataset with 2 hidden layers a strong negative impact of pre-training can be observed for the median runtime of DAE-GP.
+If the time spend on pre-training is included into the total runtime, the median duration for pre-trained DAE-GP over all 10 runs increases by a factor of `r round(PT_TIME / REG_TIME, 3)` from a median runtime of `r format(REG_TIME, scientific=F)` seconds for regular DAE-GP to `r format(PT_TIME, scientific=F)` seconds for the pre-trained algorithm.
+Even in the case where the time spend for pre-training is excluded from the total runtime, pre-training still increases the runtime of DAE-GP by a factor of `r round(PT_TIME_NOSAMPLE / REG_TIME, 3)` to `r format(PT_TIME_NOSAMPLE, scientific=F)` seconds.
 
@@
 ```
+
+To investigate the strong negative impact of pre-training on DAE-GP runtime, I first studied the influence of pre-training on the median number of training epochs that had to be spend at each generation for the training of the DAE-LSTM models $M_g$.
+Figure \@ref(fig:airfoil-trainingEpochs) shows that pre-training, as expected, strongly reduces the number of training epochs that had to be executed at each generation until the termination condition was reached.
+By reducing the number of training epochs, pre-training should therefore reduce the runtime of DAE-GP.
 

### 2023-01-31 paper.tex
autogit

@@
 \author{}
-\date{\vspace{-2.5em}Date of Submission: 2023-01-30}
+\date{\vspace{-2.5em}Date of Submission: 2023-01-31}
 
@@
 \subsection{Effect on runtime}\label{effect-on-runtime}}
+
+One important argument for using pre-training with DAE-GP is that it might be a potential way to reduce the overall run-time of the algorithm by reducing the training time of each generation's DAE-LSTM model.
+Especially in the context of optimization problems where solutions have to be found numerous times in a time efficient manner (e.g.~Vehicle Routing in the Domain of Logistic Services), one major advantage of pre-training is that a pre-trained model can be re-used to optimize many instances of an optimization problem.
+
+Unfortunately, as shown in figure \ref{fig:airfoil-runtime-total}, for the experiment conducted on the airfoil dataset with 2 hidden layers a strong negative impact of pre-training can be observed for the median runtime of DAE-GP.
+If the time spend on pre-training is included into the total runtime, the median duration for pre-trained DAE-GP over all 10 runs increases by a factor of 2.897 from a median runtime of 5076.448 seconds for regular DAE-GP to 14706.47 seconds for the pre-trained algorithm.
+Even in the case where the time spend for pre-training is excluded from the total runtime, pre-training still increases the runtime of DAE-GP by a factor of 2.47 to 12536.89 seconds.
 

### 2023-01-31 paper.Rmd
autogit

@@
 PT_TIME_NOSAMPLE <- 12536.889
+
+REG_MEDIAN_EPOCHSTRAINED <- 122.0
+PT_MEDIAN_EPOCHSTRAINED <- 53.5
+
+REG_MEDIAN_SAMPLINGTIME <- 29.055806166997172
+PT_MEDIAN_SAMPLINGTIME <- 198.4502993219976
+
 ```
@@
 Figure \@ref(fig:airfoil-trainingEpochs) shows that pre-training, as expected, strongly reduces the number of training epochs that had to be executed at each generation until the termination condition was reached.
-By reducing the number of training epochs, pre-training should therefore reduce the runtime of DAE-GP.
+By reducing the number of training epochs, pre-training should therefore have positive benefit on the runtime of DAE-GP.
 
@@
 ```
+
+If we look at another time consuming operation in DAE-GP, the creation of new individuals by sampling the DAE-LSTM model $M_G$, we see one main reason for the large runtime increase by pre-training DAE-GP.
+Figure \@ref(fig:airfoil-SamplingTime) shows the median sampling time spend during each generation of the search and it clearly shows that pre-training does strongly increase the time spent on model sampling in comparison to regular DAE-GP
 

### 2023-01-31 paper.tex
autogit

@@
 
-Unfortunately, as shown in figure \ref{fig:airfoil-runtime-total}, for the experiment conducted on the airfoil dataset with 2 hidden layers a strong negative impact of pre-training can be observed for the median runtime of DAE-GP.
+Suprinsingly, as shown in figure \ref{fig:airfoil-runtime-total}, for the experiment conducted on the airfoil dataset with 2 hidden layers a strong negative impact of pre-training can be observed for the median runtime of DAE-GP.
 If the time spend on pre-training is included into the total runtime, the median duration for pre-trained DAE-GP over all 10 runs increases by a factor of 2.897 from a median runtime of 5076.448 seconds for regular DAE-GP to 14706.47 seconds for the pre-trained algorithm.
@@
 
+To investigate the strong negative impact of pre-training on DAE-GP runtime, I first studied the influence of pre-training on the median number of training epochs that had to be spend at each generation for the training of the DAE-LSTM models \(M_g\).
+Figure \ref{fig:airfoil-trainingEpochs} shows that pre-training, as expected, strongly reduces the number of training epochs that had to be executed at each generation until the termination condition was reached.
+By reducing the number of training epochs, pre-training should therefore have positive benefit on the runtime of DAE-GP.
+
 \begin{figure}[c]
@@
 \end{figure}
+
+If we look at another time consuming operation in DAE-GP, the creation of new individuals by sampling the DAE-LSTM model \(M_G\), we see one main reason for the large runtime increase by pre-training DAE-GP.
+Figure \ref{fig:airfoil-SamplingTime} shows the median sampling time spend during each generation of the search and it clearly shows that pre-training does strongly increase the time spent on model sampling in comparison to regular DAE-GP
 

### 2023-01-31 paper.Rmd
Tue Jan 31 17:00:08 CET 2023: update-remote.sh

@@
 Figure \@ref(fig:airfoil-trainingEpochs) shows that pre-training, as expected, strongly reduces the number of training epochs that had to be executed at each generation until the termination condition was reached.
-By reducing the number of training epochs, pre-training should therefore have positive benefit on the runtime of DAE-GP.
+By reducing the number of training epochs, pre-training should therefore have a positive benefit on the runtime of DAE-GP.
 
@@
 If we look at another time consuming operation in DAE-GP, the creation of new individuals by sampling the DAE-LSTM model $M_G$, we see one main reason for the large runtime increase by pre-training DAE-GP.
-Figure \@ref(fig:airfoil-SamplingTime) shows the median sampling time spend during each generation of the search and it clearly shows that pre-training does strongly increase the time spent on model sampling in comparison to regular DAE-GP
+Figure \@ref(fig:airfoil-SamplingTime) shows the median sampling time spend during each generation of the search and it clearly demonstrates that pre-training does strongly increase the time spent on model sampling  in comparison to regular DAE-GP.
+The median time spend on sampling new individual for regular DAE-GP increases by a factor of `r round(PT_MEDIAN_SAMPLINGTIME / REG_MEDIAN_SAMPLINGTIME, 3)` from `r format(REG_MEDIAN_SAMPLINGTIME, scientific=F)` to `r format(REG_MEDIAN_SAMPLINGTIME, scientific=F)`
 

### 2023-01-31 paper.Rmd
autogit

@@
 PT_TIME <- 14706.465
-PT_TIME_NOSAMPLE <- 12536.889
+PT_TIME_NOPT <- 12536.889
 
@@
 
+REG_TIME_NOSAMLING <-  3550.276852743067
+PT_TIME_NOSAMPLING  <- 8705.007590330817
+PT_TIME_NOSAMPLING_NOPT  <- 6622.027293765
+
 ```
@@
 If the time spend on pre-training is included into the total runtime, the median duration for pre-trained DAE-GP over all 10 runs increases by a factor of `r round(PT_TIME / REG_TIME, 3)` from a median runtime of `r format(REG_TIME, scientific=F)` seconds for regular DAE-GP to `r format(PT_TIME, scientific=F)` seconds for the pre-trained algorithm.
-Even in the case where the time spend for pre-training is excluded from the total runtime, pre-training still increases the runtime of DAE-GP by a factor of `r round(PT_TIME_NOSAMPLE / REG_TIME, 3)` to `r format(PT_TIME_NOSAMPLE, scientific=F)` seconds.
+Even in the case where the time spend for pre-training is excluded from the total runtime, pre-training still increases the runtime of DAE-GP by a factor of `r round( PT_TIME_NOPT / REG_TIME, 3)` to `r format(PT_TIME_NOPT, scientific=F)` seconds.
 
@@
 Figure \@ref(fig:airfoil-trainingEpochs) shows that pre-training, as expected, strongly reduces the number of training epochs that had to be executed at each generation until the termination condition was reached.
-By reducing the number of training epochs, pre-training should therefore have a positive benefit on the runtime of DAE-GP.
+While regular DAE-GP trains for a median of `r format(REG_MEDIAN_EPOCHSTRAINED, scientific=F)` epochs at each generation, pre-training reduces the number of median training epochs by a factor of `r round(PT_MEDIAN_EPOCHSTRAINED / REG_MEDIAN_EPOCHSTRAINED, 3)` to `r format(PT_MEDIAN_EPOCHSTRAINED, scientific=F)` training epochs.
 
@@
 Figure \@ref(fig:airfoil-SamplingTime) shows the median sampling time spend during each generation of the search and it clearly demonstrates that pre-training does strongly increase the time spent on model sampling  in comparison to regular DAE-GP.
-The median time spend on sampling new individual for regular DAE-GP increases by a factor of `r round(PT_MEDIAN_SAMPLINGTIME / REG_MEDIAN_SAMPLINGTIME, 3)` from `r format(REG_MEDIAN_SAMPLINGTIME, scientific=F)` to `r format(REG_MEDIAN_SAMPLINGTIME, scientific=F)`
+The median time spend on sampling new individual for regular DAE-GP increases by a factor of `r round(PT_MEDIAN_SAMPLINGTIME / REG_MEDIAN_SAMPLINGTIME, 3)` from `r format(REG_MEDIAN_SAMPLINGTIME, scientific=F)` seconds to `r format(PT_MEDIAN_SAMPLINGTIME, scientific=F)` seconds.
 
@@
 
+Suprisingly, even after correcting the runtime for both DAE-GP results by removing the time spend during sampling, my implementation of pre-training for DAE-GP still on average takes more time to finish the search process as shown in figure \@ref(fig:airfoil-runtime-noSampling-total).
+While traditional DAE-GP has a median runtime of `r format(REG_TIME_NOSAMLING, scientific=F)`, s
+
 ```{r airfoil-runtime-noSampling-total, echo=FALSE, fig.cap="Median Runtime excluding Time for Sampling - Airfoil", out.width="80%"}
@@
 ```
+
+<!-- REG_TIME_NOSAMLING <-  3550.276852743067 -->
+<!-- PT_TIME_NOSAMPLING  <- 8705.007590330817 -->
+<!-- PT_TIME_NOSAMPLING_NOPT  <- 6622.027293765 -->
 

### 2023-01-31 paper.tex
autogit

@@
 Figure \ref{fig:airfoil-trainingEpochs} shows that pre-training, as expected, strongly reduces the number of training epochs that had to be executed at each generation until the termination condition was reached.
-By reducing the number of training epochs, pre-training should therefore have positive benefit on the runtime of DAE-GP.
+While regular DAE-GP trains for a median of 122 epochs at each generation, pre-training reduces the number of median training epochs by a factor of 0.439 to 53.5
 
@@
 If we look at another time consuming operation in DAE-GP, the creation of new individuals by sampling the DAE-LSTM model \(M_G\), we see one main reason for the large runtime increase by pre-training DAE-GP.
-Figure \ref{fig:airfoil-SamplingTime} shows the median sampling time spend during each generation of the search and it clearly shows that pre-training does strongly increase the time spent on model sampling in comparison to regular DAE-GP
+Figure \ref{fig:airfoil-SamplingTime} shows the median sampling time spend during each generation of the search and it clearly demonstrates that pre-training does strongly increase the time spent on model sampling in comparison to regular DAE-GP.
+The median time spend on sampling new individual for regular DAE-GP increases by a factor of 6.83 from 29.05581 to 198.4503.
 

### 2023-01-31 paper.Rmd
autogit

@@
 Suprisingly, even after correcting the runtime for both DAE-GP results by removing the time spend during sampling, my implementation of pre-training for DAE-GP still on average takes more time to finish the search process as shown in figure \@ref(fig:airfoil-runtime-noSampling-total).
-While traditional DAE-GP has a median runtime of `r format(REG_TIME_NOSAMLING, scientific=F)`, s
+While traditional DAE-GP has a median runtime without sampling time of `r format(REG_TIME_NOSAMLING, scientific=F)` seconds , the runtime of pre-trained DAE-GP, without time for samling and pre-training, still increases by a factor of `r round(PT_TIME_NOSAMPLING_NOPT / REG_TIME_NOSAMLING, 3)` to `r format(PT_TIME_NOSAMPLING_NOPT, scientific=F)` seconds.
 
@@
 ```
-
-<!-- REG_TIME_NOSAMLING <-  3550.276852743067 -->
-<!-- PT_TIME_NOSAMPLING  <- 8705.007590330817 -->
-<!-- PT_TIME_NOSAMPLING_NOPT  <- 6622.027293765 -->
 

### 2023-01-31 paper.tex
autogit

@@
 Figure \ref{fig:airfoil-trainingEpochs} shows that pre-training, as expected, strongly reduces the number of training epochs that had to be executed at each generation until the termination condition was reached.
-While regular DAE-GP trains for a median of 122 epochs at each generation, pre-training reduces the number of median training epochs by a factor of 0.439 to 53.5
+While regular DAE-GP trains for a median of 122 epochs at each generation, pre-training reduces the number of median training epochs by a factor of 0.439 to 53.5 training epochs.
 
@@
 Figure \ref{fig:airfoil-SamplingTime} shows the median sampling time spend during each generation of the search and it clearly demonstrates that pre-training does strongly increase the time spent on model sampling in comparison to regular DAE-GP.
-The median time spend on sampling new individual for regular DAE-GP increases by a factor of 6.83 from 29.05581 to 198.4503.
+The median time spend on sampling new individual for regular DAE-GP increases by a factor of 6.83 from 29.05581 seconds to 198.4503 seconds.
 
@@
 \end{figure}
+
+Suprisingly, even after correcting the runtime for both DAE-GP results by removing the time spend during sampling, my implementation of pre-training for DAE-GP still on average takes more time to finish the search process as shown in figure \ref{fig:airfoil-runtime-noSampling-total}.
+While traditional DAE-GP has a median runtime without sampling time of 3550.277 seconds , the runtime of pre-trained DAE-GP, without time for samling and pre-training, still increases by a factor of 1.865 to 6622.027 seconds.
 

### 2023-02-01 paper.Rmd
Wed Feb  1 12:00:06 CET 2023: update-remote.sh

@@
 
+```{python read_runtime_vars, echo=FALSE}
+from json import load
+D_AIRFOIL_2hl = load(
+  open("/Users/rmn/github/master_thesis/data/airfoil_2hl_150hn_fullRun_30gens/summary.json", "r", encoding="utf-8")
+  )
+
+D_AIRFOIL_1hl = load(
+  open("/Users/rmn/github/master_thesis/data/airfoil_1hl_150hn_fullRun_30gens/summary.json", "r", encoding="utf-8")
+  )
+
+```
+
+```{r runtime_vars, echo=FALSE}
+
+REG_TRAIN_MED_FF <- py$D_AIRFOIL_2hl$REG_TRAIN_MED_FF
+REG_TEST_MED_FF <- py$D_AIRFOIL_2hl$REG_TEST_MED_FF
+PT_TRAIN_MED_FF <- py$D_AIRFOIL_2hl$PT_TRAIN_MED_FF
+PT_TEST_MED_FF <- py$D_AIRFOIL_2hl$PT_TEST_MED_FF
+GP_TRAIN_MED_FF <- py$D_AIRFOIL_2hl$GP_TRAIN_MED_FF
+GP_TEST_MED_FF <- py$D_AIRFOIL_2hl$GP_TEST_MED_FF
+REG_TRAIN_STDD_FF <- py$D_AIRFOIL_2hl$REG_TRAIN_STDD_FF
+REG_TEST_STDD_FF <- py$D_AIRFOIL_2hl$REG_TEST_STDD_FF
+PT_TRAIN_STDD_FF <- py$D_AIRFOIL_2hl$PT_TRAIN_STDD_FF
+PT_TEST_STDD_FF <- py$D_AIRFOIL_2hl$PT_TEST_STDD_FF
+GP_TRAIN_STDD_FF <- py$D_AIRFOIL_2hl$GP_TRAIN_STDD_FF
+GP_TEST_STDD_FF <- py$D_AIRFOIL_2hl$GP_TEST_STDD_FF
+
+REG_TRAIN_MED_FF_1HL <- py$D_AIRFOIL_1hl$REG_TRAIN_MED_FF
+REG_TEST_MED_FF_1HL <- py$D_AIRFOIL_1hl$REG_TEST_MED_FF
+PT_TRAIN_MED_FF_1HL <- py$D_AIRFOIL_1hl$PT_TRAIN_MED_FF
+PT_TEST_MED_FF_1HL <- py$D_AIRFOIL_1hl$PT_TEST_MED_FF
+GP_TRAIN_MED_FF_1HL <- py$D_AIRFOIL_1hl$GP_TRAIN_MED_FF
+GP_TEST_MED_FF_1HL <- py$D_AIRFOIL_1hl$GP_TEST_MED_FF
+REG_TRAIN_STDD_FF_1HL <- py$D_AIRFOIL_1hl$REG_TRAIN_STDD_FF
+REG_TEST_STDD_FF_1HL <- py$D_AIRFOIL_1hl$REG_TEST_STDD_FF
+PT_TRAIN_STDD_FF_1HL <- py$D_AIRFOIL_1hl$PT_TRAIN_STDD_FF
+PT_TEST_STDD_FF_1HL <- py$D_AIRFOIL_1hl$PT_TEST_STDD_FF
+GP_TRAIN_STDD_FF_1HL <- py$D_AIRFOIL_1hl$GP_TRAIN_STDD_FF
+GP_TEST_STDD_FF_1HL <- py$D_AIRFOIL_1hl$GP_TEST_STDD_FF
+
+```
+
 After closely examining the effect of pre-training on DAE-GPs generalization behaviour another series of experiments is conducted to study how pre-training influences the overall search behaviour of DAE-GP.
@@
 
-The first interesting observation from this experiment is, that the pre-trained DAE-GP shows slightly lower mean RMSE (higher fitness) but a higher median RMSE (lower fitness) than regular DAE-GP for both the training and the testing set in overall fitness.Regarding the presence of overfitting, both DAE-GP variants show only a very small gap between the testing and training fitness which indicates a good generalization behavior.
+The first interesting observation from this experiment is, that the pre-trained DAE-GP shows slightly lower mean RMSE of `r round(, 3)` (higher fitness) but a higher median RMSE (lower fitness) than regular DAE-GP for both the training and the testing set in overall fitness.
+Regarding the presence of overfitting, both DAE-GP variants show only a very small gap between the testing and training fitness which indicates a good generalization behavior.
 
@@
 
+```{python read_runtime_vars, echo=FALSE}
+from json import load
+D_AIRFOIL_TIME = load(
+  open("/Users/rmn/github/master_thesis/data/airfoil_2hl_150hn_fullRun_30gens_TimeBenchmark/summary.json", "r", encoding="utf-8")
+  )
+```
+
 ```{r runtime_vars, echo=FALSE}
 
-# Values are copied from file /eda-gp-2020/experiments/airfoil_2hl_150hn_fullRun_30gens/airfoil_2hl_150hn_fullRun_30gens_TimeBenchmark.ipynb
+REG_TIME <- py$D_AIRFOIL_TIME$REG_TIME
+PT_TIME <- py$D_AIRFOIL_TIME$PT_TIME
+PT_TIME_NOPT <- py$D_AIRFOIL_TIME$PT_TIME_NOPT
 
-REG_TIME <- 5076.448
-PT_TIME <- 14706.465
-PT_TIME_NOPT <- 12536.889
+REG_MEDIAN_EPOCHSTRAINED <- py$D_AIRFOIL_TIME$REG_MEDIAN_EPOCHSTRAINED
+PT_MEDIAN_EPOCHSTRAINED <- py$D_AIRFOIL_TIME$PT_MEDIAN_EPOCHSTRAINED
 
-REG_MEDIAN_EPOCHSTRAINED <- 122.0
-PT_MEDIAN_EPOCHSTRAINED <- 53.5
+REG_MEDIAN_SAMPLINGTIME <- py$D_AIRFOIL_TIME$REG_MEDIAN_SAMPLINGTIME
+PT_MEDIAN_SAMPLINGTIME <- py$D_AIRFOIL_TIME$PT_MEDIAN_SAMPLINGTIME
 
-REG_MEDIAN_SAMPLINGTIME <- 29.055806166997172
-PT_MEDIAN_SAMPLINGTIME <- 198.4502993219976
+REG_TIME_NOSAMLING <-  py$D_AIRFOIL_TIME$REG_TIME_NOSAMLING
+PT_TIME_NOSAMPLING  <- py$D_AIRFOIL_TIME$PT_TIME_NOSAMPLING
+PT_TIME_NOSAMPLING_NOPT  <- py$D_AIRFOIL_TIME$PT_TIME_NOSAMPLING_NOPT
 
-REG_TIME_NOSAMLING <-  3550.276852743067
-PT_TIME_NOSAMPLING  <- 8705.007590330817
-PT_TIME_NOSAMPLING_NOPT  <- 6622.027293765
+REG_PARAMS <- py$D_AIRFOIL_TIME$REG_PARAMS
+PT_PARAMS <- py$D_AIRFOIL_TIME$PT_PARAMS
 
@@
 While traditional DAE-GP has a median runtime without sampling time of `r format(REG_TIME_NOSAMLING, scientific=F)` seconds , the runtime of pre-trained DAE-GP, without time for samling and pre-training, still increases by a factor of `r round(PT_TIME_NOSAMPLING_NOPT / REG_TIME_NOSAMLING, 3)` to `r format(PT_TIME_NOSAMPLING_NOPT, scientific=F)` seconds.
+While investigating other reasons for the differences in runtime, I verified that both DAE-GP variants use the exact same number of trainable parameters for each generations DAE-LSTMS models ( `r format(REG_PRAMS, scientific=F)`).
 

### 2023-02-01 paper.Rmd
Wed Feb  1 14:00:06 CET 2023: update-remote.sh

@@
 
-```{python read_runtime_vars, echo=FALSE}
+```{python read_solQual_vars, echo=FALSE}
 from json import load
 D_AIRFOIL_2hl = load(
-  open("/Users/rmn/github/master_thesis/data/airfoil_2hl_150hn_fullRun_30gens/summary.json", "r", encoding="utf-8")
+  open("/Users/rmn/github/master_thesis/data/airfoil_2hl_150hn_fullRun_30gens_withGP/summary.json", "r", encoding="utf-8")
   )
@@
 
-```{r runtime_vars, echo=FALSE}
+```{r solQual_vars, echo=FALSE}
 
@@
 GP_TEST_STDD_FF <- py$D_AIRFOIL_2hl$GP_TEST_STDD_FF
+
+REG_MED_BESTSIZE <- py$D_AIRFOIL_2hl$REG_MED_BESTSIZE
+PT_MED_BESTSIZE <- py$D_AIRFOIL_2hl$PT_MED_BESTSIZE
+GP_MED_BESTSIZE <- py$D_AIRFOIL_2hl$GP_MED_BESTSIZE
 
@@
 
-The first interesting observation from this experiment is, that the pre-trained DAE-GP shows slightly lower mean RMSE of `r round(, 3)` (higher fitness) but a higher median RMSE (lower fitness) than regular DAE-GP for both the training and the testing set in overall fitness.
+The first interesting observation from this experiment is, that the pre-trained DAE-GP shows slightly lower mean RMSE (higher fitness) but a higher median RMSE (lower fitness) than regular DAE-GP for both the training and the testing set in overall fitness.
 Regarding the presence of overfitting, both DAE-GP variants show only a very small gap between the testing and training fitness which indicates a good generalization behavior.
@@
 The distribution of final fitness scores for pre-trained and regular DAE-GP are visualized as boxplots in figure  \@ref(fig:airfoil-distribution-fitness-30gens).
-The distribution for regular DAE-GP shows to be preferable to that of pre-trained DAE-GP on both sets since it shows a lower median RMSE as well as less dispersion of the solutions measured by the standard deviation.
-An interesting observation from studying the distribution of final fitness scores for both algorithms is that while both algorithms have a very similiar interquartile range (similiar box shape), pre-trained DAE-GP has more extreme outliers for high RMSE which explains the slightly worse performance if looking at median best fitness present in figure \@ref(fig:airfoil-median-fitness-30gens).
+The distribution for regular DAE-GP shows to be preferable to that of pre-trained DAE-GP  since it shows a lower median RMSE of `r round(REG_TEST_MED_FF, 3)` for the test set (`r round(PT_TEST_MED_FF, 3)` for pre-trained DAE-GP) as well as less dispersion of the solutions measured by a standard deviation of `r round(REG_TEST_STDD_FF, 3)` (`r round(PT_TEST_STDD_FF, 3)` for pre-trained DAE-GP).
+For reference, the median final RMSE for the test set of standard GP is at `r round(GP_TEST_MED_FF, 3)` which is larger by a factor of `r round(GP_TEST_MED_FF / PT_TEST_MED_FF, 3)` if compared to pre-trained DAE-GP.
+
+An interesting observation from studying the distribution of final fitness scores for both algorithms is that while both algorithms have a very similiar interquartile range (similar box shape), pre-trained DAE-GP has more extreme outliers for high RMSE which explains the slightly worse performance if looking at median best fitness present in figure \@ref(fig:airfoil-median-fitness-30gens).
 
@@
 Regarding the size of the best solution per generation, the results show that the pre-trained version of algorithm did produce solutions with the same median size as traditional DAE-GP.
+The median size
 
@@
 
-```{r runtime_vars, echo=FALSE}
+```{r runtime_vars, include=FALSE}
 
@@
 While traditional DAE-GP has a median runtime without sampling time of `r format(REG_TIME_NOSAMLING, scientific=F)` seconds , the runtime of pre-trained DAE-GP, without time for samling and pre-training, still increases by a factor of `r round(PT_TIME_NOSAMPLING_NOPT / REG_TIME_NOSAMLING, 3)` to `r format(PT_TIME_NOSAMPLING_NOPT, scientific=F)` seconds.
-While investigating other reasons for the differences in runtime, I verified that both DAE-GP variants use the exact same number of trainable parameters for each generations DAE-LSTMS models ( `r format(REG_PRAMS, scientific=F)`).
+While investigating other reasons for the differences in runtime, I verified that both DAE-GP variants use the exact same number of trainable parameters for each generations DAE-LSTMS models ( `r format(REG_PARAMS, scientific=F)`).
 

### 2023-02-01 paper.tex
Wed Feb  1 14:00:06 CET 2023: update-remote.sh

@@
 \author{}
-\date{\vspace{-2.5em}Date of Submission: 2023-01-31}
+\date{\vspace{-2.5em}Date of Submission: 2023-02-01}
 
@@
 
-The first interesting observation from this experiment is, that the pre-trained DAE-GP shows slightly lower mean RMSE (higher fitness) but a higher median RMSE (lower fitness) than regular DAE-GP for both the training and the testing set in overall fitness.Regarding the presence of overfitting, both DAE-GP variants show only a very small gap between the testing and training fitness which indicates a good generalization behavior.
+The first interesting observation from this experiment is, that the pre-trained DAE-GP shows slightly lower mean RMSE of (higher fitness) but a higher median RMSE (lower fitness) than regular DAE-GP for both the training and the testing set in overall fitness.
+Regarding the presence of overfitting, both DAE-GP variants show only a very small gap between the testing and training fitness which indicates a good generalization behavior.
 
@@
 The distribution of final fitness scores for pre-trained and regular DAE-GP are visualized as boxplots in figure \ref{fig:airfoil-distribution-fitness-30gens}.
-The distribution for regular DAE-GP shows to be preferable to that of pre-trained DAE-GP on both sets since it shows a lower median RMSE as well as less dispersion of the solutions measured by the standard deviation.
+The distribution for regular DAE-GP shows to be preferable to that of pre-trained DAE-GP since it shows a lower median RMSE of 6.991 for the test set (7.68 for pre-trained DAE-GP) as well as less dispersion of the solutions measured by a standard deviation of 11.989 (12.548 for pre-trained DAE-GP).
 An interesting observation from studying the distribution of final fitness scores for both algorithms is that while both algorithms have a very similiar interquartile range (similiar box shape), pre-trained DAE-GP has more extreme outliers for high RMSE which explains the slightly worse performance if looking at median best fitness present in figure \ref{fig:airfoil-median-fitness-30gens}.
@@
 Suprinsingly, as shown in figure \ref{fig:airfoil-runtime-total}, for the experiment conducted on the airfoil dataset with 2 hidden layers a strong negative impact of pre-training can be observed for the median runtime of DAE-GP.
-If the time spend on pre-training is included into the total runtime, the median duration for pre-trained DAE-GP over all 10 runs increases by a factor of 2.897 from a median runtime of 5076.448 seconds for regular DAE-GP to 14706.47 seconds for the pre-trained algorithm.
-Even in the case where the time spend for pre-training is excluded from the total runtime, pre-training still increases the runtime of DAE-GP by a factor of 2.47 to 12536.89 seconds.
+If the time spend on pre-training is included into the total runtime, the median duration for pre-trained DAE-GP over all 10 runs increases by a factor of 3.121 from a median runtime of 4712.303 seconds for regular DAE-GP to 14706.47 seconds for the pre-trained algorithm.
+Even in the case where the time spend for pre-training is excluded from the total runtime, pre-training still increases the runtime of DAE-GP by a factor of 2.66 to 12536.89 seconds.
 
@@
 While traditional DAE-GP has a median runtime without sampling time of 3550.277 seconds , the runtime of pre-trained DAE-GP, without time for samling and pre-training, still increases by a factor of 1.865 to 6622.027 seconds.
+While investigating other reasons for the differences in runtime, I verified that both DAE-GP variants use the exact same number of trainable parameters for each generations DAE-LSTMS models ( NULL).
 

### 2023-02-02 paper.Rmd
Thu Feb  2 08:00:07 CET 2023: update-remote.sh

@@
 GP_MED_BESTSIZE <- py$D_AIRFOIL_2hl$GP_MED_BESTSIZE
+REG_MED_AVGSIZE <- py$D_AIRFOIL_2hl$REG_MED_AVG_SIZE
+PT_MED_AVGSIZE <- py$D_AIRFOIL_2hl$PT_MED_AVG_SIZE
 
@@
 It can be observed that both DAE-GP variants strongly reduce the average solutions size already in the first generations and that the average solution size inside the population constantly stays at this very low level for the rest of the search.
+The median average size over all generations is `r round(REG_MED_AVGSIZE, 3)` for DAE-GP and `r round(PT_MED_AVGSIZE, 3)` for pre-trained DAE-GP.
 Regarding the size of the best solution per generation, the results show that the pre-trained version of algorithm did produce solutions with the same median size as traditional DAE-GP.
-The median size
+The median size of the best solutions found after terminating the search is the same for both algorithms (DAE-GP: `r round(REG_MED_BESTSIZE, 3)`, Pre-Trained DAE-GP: `r round(PT_MED_BESTSIZE, 3)`)
 

### 2023-02-02 paper.Rmd
Thu Feb  2 09:00:06 CET 2023: update-remote.sh

@@
 GP_TEST_MED_FF <- py$D_AIRFOIL_2hl$GP_TEST_MED_FF
+
 REG_TRAIN_STDD_FF <- py$D_AIRFOIL_2hl$REG_TRAIN_STDD_FF
@@
 
-REG_MED_BESTSIZE <- py$D_AIRFOIL_2hl$REG_MED_BESTSIZE
-PT_MED_BESTSIZE <- py$D_AIRFOIL_2hl$PT_MED_BESTSIZE
-GP_MED_BESTSIZE <- py$D_AIRFOIL_2hl$GP_MED_BESTSIZE
+REG_MED_BESTSIZE <- py$D_AIRFOIL_2hl$REG_MED_BEST_SIZE
+PT_MED_BESTSIZE <- py$D_AIRFOIL_2hl$PT_MED_BEST_SIZE
+GP_MED_BESTSIZE <- py$D_AIRFOIL_2hl$GP_MED_BEST_SIZE
+
 REG_MED_AVGSIZE <- py$D_AIRFOIL_2hl$REG_MED_AVG_SIZE
 PT_MED_AVGSIZE <- py$D_AIRFOIL_2hl$PT_MED_AVG_SIZE
+
+REG_MED_LEVDIV <- py$D_AIRFOIL_2hl$REG_MED_LEVDIV
+PT_MED_LEVDIV <- py$D_AIRFOIL_2hl$PT_MED_LEVDIV
+GP_MED_LEVDIV <- py$D_AIRFOIL_2hl$GP_MED_LEVDIV
 
@@
 
-An interesting observation from studying the distribution of final fitness scores for both algorithms is that while both algorithms have a very similiar interquartile range (similar box shape), pre-trained DAE-GP has more extreme outliers for high RMSE which explains the slightly worse performance if looking at median best fitness present in figure \@ref(fig:airfoil-median-fitness-30gens).
+An interesting observation from studying the distribution of final fitness scores for both algorithms is that pre-trained DAE-GP has a smaller interquartile range and more extreme outliers for high RMSE which explains the slightly worse performance if looking at median best fitness present in figure \@ref(fig:airfoil-median-fitness-30gens).
 
-Since previous work by @dae-gp_2022_symreg demonstrated that DAE-GP, for a given number of 10.000 fitness evaluations, is able to produce solutions that are much smaller in tree size that regular GP, I also studied the influence of pre-training on the size of solutions found by DAE-GP.
+Since previous work by @dae-gp_2022_symreg demonstrated that DAE-GP, for a given number of 10.000 fitness evaluations, is able to produce solutions that are much smaller in tree size than regular GP, I also studied the influence of pre-training on the size of solutions found by DAE-GP.
 Figure \@ref(fig:airfoil-size-combined-30gens) shows the median of both the average size of individuals inside the population as well as the size of the currently best performing solution inside the population.
@@
 To conclude the first experiment, figure \@ref(fig:airfoil-popDiversity) shows the median population diversity measured by the normalized Levenshtein edit distance for both DAE-GP algorithms and regular GP as a benchmark.
-While the population diversity steadily decreases over the generations for standard GP, DAE-GP variants strongly increase the population diversity in the first generations and keep them at a high level of over $0.9$.
-Here it can be observed, that pre-training shows a positive effect by increasing population diversity in comparison to regular DAE-GP.
+While the population diversity steadily decreases over the generations for standard GP, DAE-GP variants strongly increase the population diversity in the first generations and keep them at a higher level during the evolutionary search.
+Here it can be observed, that pre-training shows a positive effect by increasing population diversity in comparison to regular DAE-GP (Median over all generations for DAE-GP: `r round(REG_MED_LEVDIV, 3)`, Pre-Trained DAE-GP: `r round(PT_MED_LEVDIV, 3)`, Standard-GP `r round(GP_MED_LEVDIV, 3)`)
 
@@
 Based on the findings of @pmlr-v5-erhan09a, I expected that pre-training DAE-GP would lead to a decrease in overall performance.
-Figures \@ref(fig:airfoil-fitness-perGen) and \@ref(fig:airfoil-finalFitness) shows the average best fitness per generation as well as the distribution of final fitness scores after 30 generations.
+Figures \@ref(fig:airfoil-fitness-perGen) and \@ref(fig:airfoil-finalFitness) shows the mean and median best fitness per generation as well as the distribution of final fitness scores after 30 generations.
 The results are, again, based on $10$ individual runs per algorithm.
-As expected it shows, that for a small DAE-LSTM size pre-training negativley impacts the fitness of solution found by DAE-GG.
+As expected the results show, that for a small DAE-LSTM size pre-training negativley impacts the fitness of solution found by DAE-GP.
+The median final RMSE for the test set of pre-trained DAE-GP increases by a factor of `r round(PT_TEST_MED_FF_1HL, 3)`
 

### 2023-02-02 paper.tex
Thu Feb  2 09:00:06 CET 2023: update-remote.sh

@@
 \author{}
-\date{\vspace{-2.5em}Date of Submission: 2023-02-01}
+\date{\vspace{-2.5em}Date of Submission: 2023-02-02}
 
@@
 
-The first interesting observation from this experiment is, that the pre-trained DAE-GP shows slightly lower mean RMSE of (higher fitness) but a higher median RMSE (lower fitness) than regular DAE-GP for both the training and the testing set in overall fitness.
+The first interesting observation from this experiment is, that the pre-trained DAE-GP shows slightly lower mean RMSE (higher fitness) but a higher median RMSE (lower fitness) than regular DAE-GP for both the training and the testing set in overall fitness.
 Regarding the presence of overfitting, both DAE-GP variants show only a very small gap between the testing and training fitness which indicates a good generalization behavior.
@@
 The distribution for regular DAE-GP shows to be preferable to that of pre-trained DAE-GP since it shows a lower median RMSE of 6.991 for the test set (7.68 for pre-trained DAE-GP) as well as less dispersion of the solutions measured by a standard deviation of 11.989 (12.548 for pre-trained DAE-GP).
-An interesting observation from studying the distribution of final fitness scores for both algorithms is that while both algorithms have a very similiar interquartile range (similiar box shape), pre-trained DAE-GP has more extreme outliers for high RMSE which explains the slightly worse performance if looking at median best fitness present in figure \ref{fig:airfoil-median-fitness-30gens}.
-
-Since previous work by {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]} demonstrated that DAE-GP, for a given number of 10.000 fitness evaluations, is able to produce solutions that are much smaller in tree size that regular GP, I also studied the influence of pre-training on the size of solutions found by DAE-GP.
+For reference, the median final RMSE for the test set of standard GP is at 32.024 which is larger by a factor of 4.17 if compared to pre-trained DAE-GP.
+
+An interesting observation from studying the distribution of final fitness scores for both algorithms is that pre-trained DAE-GP has a smaller interquartile range and more extreme outliers for high RMSE which explains the slightly worse performance if looking at median best fitness present in figure \ref{fig:airfoil-median-fitness-30gens}.
+
+Since previous work by {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]} demonstrated that DAE-GP, for a given number of 10.000 fitness evaluations, is able to produce solutions that are much smaller in tree size than regular GP, I also studied the influence of pre-training on the size of solutions found by DAE-GP.
 Figure \ref{fig:airfoil-size-combined-30gens} shows the median of both the average size of individuals inside the population as well as the size of the currently best performing solution inside the population.
 It can be observed that both DAE-GP variants strongly reduce the average solutions size already in the first generations and that the average solution size inside the population constantly stays at this very low level for the rest of the search.
+The median average size over all generations is 1.536 for DAE-GP and 1.536 for pre-trained DAE-GP.
 Regarding the size of the best solution per generation, the results show that the pre-trained version of algorithm did produce solutions with the same median size as traditional DAE-GP.
+The median size of the best solutions found after terminating the search is the same for both algorithms (DAE-GP: 5, Pre-Trained DAE-GP: 5)
 
@@
 While the population diversity steadily decreases over the generations for standard GP, DAE-GP variants strongly increase the population diversity in the first generations and keep them at a high level of over \(0.9\).
-Here it can be observed, that pre-training shows a positive effect by increasing population diversity in comparison to regular DAE-GP.
+Here it can be observed, that pre-training shows a positive effect by increasing population diversity in comparison to regular DAE-GP (Median over all generations for DAE-GP: 0.861, Pre-Trained DAE-GP: 0.925, Standard-GP 0.705)
 

### 2023-02-02 paper.Rmd
autogit

@@
 As expected the results show, that for a small DAE-LSTM size pre-training negativley impacts the fitness of solution found by DAE-GP.
-The median final RMSE for the test set of pre-trained DAE-GP increases by a factor of `r round(PT_TEST_MED_FF_1HL, 3)`
+The median final RMSE for the test set of pre-trained DAE-GP increases by a factor of `r round(PT_TEST_MED_FF_1HL / REG_TEST_MED_FF_1HL, 3)` to `r round(PT_TEST_MED_FF_1HL , 3)` compared to regular DAE-GP with `r round(REG_TEST_MED_FF_1HL , 3)`.
 
@@
 Second Generation Pre-Training tries to exploit this observation by shifting the pre-training phase from the start of the algorithm into the second generation with the aim to produce a pre-training Population $\hat{P}$ that more accurately captures the properties of the search problem.
-The basic idea is to run basic DAE-GP for the first generation without any interference, after completing the first generation, the pre-training population $\hat{P}$ is created by sampling the DAE-LSTM model $M_1$.
+The basic idea is to run regular DAE-GP for the first generation without any interference, after completing the first generation, the pre-training population $\hat{P}$ is created by sampling the DAE-LSTM model $M_1$.
 The pre-training model $\hat{M}$ is then trained using $\hat{P}$, after finishing DAE-GP continues running and initializes each new DAE-LSTM model $M_g$ for $g\in{2,...,g_{max}}$ using the trainable parameters $\theta_{\hat{M}}$.
@@
 Figure \@ref(fig:altStrategies-FitnessDistr) shows the distribution of best fitness achieved on the test set after running for 30 full generations.
-Both alternative second generation Pre-Training as well as grow-initialized Pre-Trainning show a much higher median RMSE (18.88/17.21) compared to standard pre-training (8.55) and regular DAE-GP (6.99).
+Both alternative second generation Pre-Training as well as grow-initialized Pre-Training show a much higher median RMSE (18.88/17.21) compared to standard pre-training (8.55) and regular DAE-GP (6.99).
 Also both approaches do not demonstrate any benefit on the dispersion of solutions found resulting in similiar standard deviations if compared to standard pre-training.
@@
 
-```{r runtime_vars, include=FALSE}
+```{r runtime_vars, echo=FALSE}
 
@@
 
-REG_PARAMS <- py$D_AIRFOIL_TIME$REG_PARAMS
-PT_PARAMS <- py$D_AIRFOIL_TIME$PT_PARAMS
+REG_PARAMS <- py$D_AIRFOIL_TIME$REG_TOTAL_PARAMS
+PT_PARAMS <- py$D_AIRFOIL_TIME$PT_TOTAL_PARAMS
 
@@
 While traditional DAE-GP has a median runtime without sampling time of `r format(REG_TIME_NOSAMLING, scientific=F)` seconds , the runtime of pre-trained DAE-GP, without time for samling and pre-training, still increases by a factor of `r round(PT_TIME_NOSAMPLING_NOPT / REG_TIME_NOSAMLING, 3)` to `r format(PT_TIME_NOSAMPLING_NOPT, scientific=F)` seconds.
-While investigating other reasons for the differences in runtime, I verified that both DAE-GP variants use the exact same number of trainable parameters for each generations DAE-LSTMS models ( `r format(REG_PARAMS, scientific=F)`).
+While investigating other reasons for the differences in runtime, I verified that both DAE-GP variants use the exact same number of trainable parameters for each generations DAE-LSTMS models (total number of trainable parameters DAE-GP=`r format(REG_PARAMS, scientific=F)`, Pre-trained DAE-GP=`r format(PT_PARAMS, scientific=F)`).
 

### 2023-02-02 paper.tex
autogit

@@
 To conclude the first experiment, figure \ref{fig:airfoil-popDiversity} shows the median population diversity measured by the normalized Levenshtein edit distance for both DAE-GP algorithms and regular GP as a benchmark.
-While the population diversity steadily decreases over the generations for standard GP, DAE-GP variants strongly increase the population diversity in the first generations and keep them at a high level of over \(0.9\).
+While the population diversity steadily decreases over the generations for standard GP, DAE-GP variants strongly increase the population diversity in the first generations and keep them at a higher level during the evolutionary search.
 Here it can be observed, that pre-training shows a positive effect by increasing population diversity in comparison to regular DAE-GP (Median over all generations for DAE-GP: 0.861, Pre-Trained DAE-GP: 0.925, Standard-GP 0.705)
@@
 Based on the findings of {[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{3}{]}, I expected that pre-training DAE-GP would lead to a decrease in overall performance.
-Figures \ref{fig:airfoil-fitness-perGen} and \ref{fig:airfoil-finalFitness} shows the average best fitness per generation as well as the distribution of final fitness scores after 30 generations.
+Figures \ref{fig:airfoil-fitness-perGen} and \ref{fig:airfoil-finalFitness} shows the mean and median best fitness per generation as well as the distribution of final fitness scores after 30 generations.
 The results are, again, based on \(10\) individual runs per algorithm.
-As expected it shows, that for a small DAE-LSTM size pre-training negativley impacts the fitness of solution found by DAE-GG.
+As expected the results show, that for a small DAE-LSTM size pre-training negativley impacts the fitness of solution found by DAE-GP.
+The median final RMSE for the test set of pre-trained DAE-GP increases by a factor of 1.287 to 16.213 compared to regular DAE-GP with 12.594.
 

### 2023-02-02 paper.Rmd
Thu Feb  2 12:00:06 CET 2023: update-remote.sh

@@
 
+The runtime analysis in this chapter is based on experiments that were all executed on an Intel® Xeon® W-2245 Processor without the use of graphical processing unit (GPU) acceleration in the training of DAE-LSTM networks.
+Since the framework that was used is written in the ,comparatively slow, python programming language, the absolute runtime measurements are only of limited
+
 <!-- TODO: Disclaimer, Runtime is implementation dependant, no GPU use, etc... -->

### 2023-02-02 paper.tex
Thu Feb  2 12:00:06 CET 2023: update-remote.sh

@@
 Second Generation Pre-Training tries to exploit this observation by shifting the pre-training phase from the start of the algorithm into the second generation with the aim to produce a pre-training Population \(\hat{P}\) that more accurately captures the properties of the search problem.
-The basic idea is to run basic DAE-GP for the first generation without any interference, after completing the first generation, the pre-training population \(\hat{P}\) is created by sampling the DAE-LSTM model \(M_1\).
+The basic idea is to run regular DAE-GP for the first generation without any interference, after completing the first generation, the pre-training population \(\hat{P}\) is created by sampling the DAE-LSTM model \(M_1\).
 The pre-training model \(\hat{M}\) is then trained using \(\hat{P}\), after finishing DAE-GP continues running and initializes each new DAE-LSTM model \(M_g\) for \(g\in{2,...,g_{max}}\) using the trainable parameters \(\theta_{\hat{M}}\).
@@
 Figure \ref{fig:altStrategies-FitnessDistr} shows the distribution of best fitness achieved on the test set after running for 30 full generations.
-Both alternative second generation Pre-Training as well as grow-initialized Pre-Trainning show a much higher median RMSE (18.88/17.21) compared to standard pre-training (8.55) and regular DAE-GP (6.99).
+Both alternative second generation Pre-Training as well as grow-initialized Pre-Training show a much higher median RMSE (18.88/17.21) compared to standard pre-training (8.55) and regular DAE-GP (6.99).
 Also both approaches do not demonstrate any benefit on the dispersion of solutions found resulting in similiar standard deviations if compared to standard pre-training.
@@
 While traditional DAE-GP has a median runtime without sampling time of 3550.277 seconds , the runtime of pre-trained DAE-GP, without time for samling and pre-training, still increases by a factor of 1.865 to 6622.027 seconds.
-While investigating other reasons for the differences in runtime, I verified that both DAE-GP variants use the exact same number of trainable parameters for each generations DAE-LSTMS models ( NULL).
+While investigating other reasons for the differences in runtime, I verified that both DAE-GP variants use the exact same number of trainable parameters for each generations DAE-LSTMS models (total number of trainable parameters DAE-GP=573473, Pre-trained DAE-GP=573473).
 

### 2023-02-02 paper.Rmd
Thu Feb  2 13:00:07 CET 2023: update-remote.sh

--- 2023-02-02 paper.Rmd
@@
 The key idea is to pre-train an initial DAE-LSTM on a large population of candidate solutions and to use the pre-trained parameters of this initial model in each generation of the search as starting parameters for the current generation DAE-LSTM.
-This thesis studies the influence of using pre-trained DAE-LSTMs in DAE-GP for symbolic regression, especially looking at the influence on overall quality of solutions found by DAE-GP and effects on runtime.
+This thesis studies the influence of using pre-trained DAE-LSTMs in DAE-GP for symbolic regression, especially looking at the influence on overall quality of solutions found by DAE-GP and effects on run-time.
 The aim of this study is to answer the question if a pre-training strategy can be used to improve DAE-GP and to study both positive and negative effects on overall performance.
@@
 One possible way of model building in EDA proposed by @harmless_overfitting_eda is to use denoising autoencoders (DAEs) as generative models to capture complex probability distributions of decision variables.
-DAE, a variation of the autoencoders (AE), are a widely used type of neural networks in the domain of unsupervized machine learning that maps $n$ input variables to $n$ output variables using a hidden representation.
+DAE, a variation of the autoencoders (AE), are a widely used type of neural networks in the domain of unsupervised machine learning that maps $n$ input variables to $n$ output variables using a hidden representation.
 
-AE were introduced by @ae_orig to compress high-dimensional data into lower-dimensions. An AE consists of two different subunits:
+AE were introduced by @ae_orig to compress high-dimensional data into lower-dimensions. An AE consists of two different sub-units:
 
-  * Encoder $g(x)$: Encode input data to a smaller cental layer $h$
+  * Encoder $g(x)$: Encode input data to a smaller central layer $h$
   * Decoder $d(h)$: Decode and output the the encoded data back to its original dimension
 
-The AE is trained to reduce the reconstruction error between input and output data, after the training procedure is finished the network is able to reduce the dimensionality of input data to a compressed representation[@ae_orig].
+The AE is trained to reduce the reconstruction error between input and output data, after the training procedure is finished the network is able to reduce the dimensional of input data to a compressed representation[@ae_orig].
 
@@
 
-Some of the main motivations for using pre-training is the ability to reduce the need for large amounts of training data which can often be unavailable or too expensive to collect as well as an improved performance by either reducing the computational effort that is necessary to train a model to solve a specific task or by improving the models generalisation ability.
+Some of the main motivations for using pre-training is the ability to reduce the need for large amounts of training data which can often be unavailable or too expensive to collect as well as an improved performance by either reducing the computational effort that is necessary to train a model to solve a specific task or by improving the models generalization ability.
 
@@
 
--- describe unsupervized vs supervized pre-training...
+-- describe unsupervised vs supervised pre-training...
 
@@
 Early research on pre-training for DAE by @pmlr-v5-erhan09a comes to the conclusion that besides adding robustness to models the strategy also results in improved generalization and better performing models.
-The authors describe that unsupervised pre-training behaves like a regulizer on DAE networks, the mean test error and it's variance are reduced for DAE networks that are initialized from pre-trained weights if compared to the same architectures that use random initialization.
+The authors describe that unsupervised pre-training behaves like a regularizer on DAE networks, the mean test error and it's variance are reduced for DAE networks that are initialized from pre-trained weights if compared to the same architectures that use random initialization.
 
 Another important finding by @pmlr-v5-erhan09a is that the positive effect of pre-training is dependent on both the depth of the network as well as the size of layers - while increasingly larger networks benefit increasingly more from pre-training the final performance of small architectures tends to be worse with pre-training than with randomized weight initialization.
-The authors find evidence that pre-training DAE is especially usefull for optimizing the parameters of the lower layers of the network.
+The authors find evidence that pre-training DAE is especially useful for optimizing the parameters of the lower layers of the network.
 
@@
 
-One difficulty in the implementation of a pre-training strategy into DAE-GP has been the detemination of the number of hidden neurons in side the DAE-LSTMs hidden layers.
+One difficulty in the implementation of a pre-training strategy into DAE-GP has been the determination of the number of hidden neurons in side the DAE-LSTMs hidden layers.
 The original description of DAE-GP [@dae-gp_2020_rtree] uses a strategy where the number of hidden neurons for each hidden layer is dynamically set per generation to the maximum individual size inside the current population.
@@
 To study generalization behavior this series of experiments is conducted using DAE-GP with only a single generation until the search terminates.
-I disregard the fitness of solutions found and focus soley on the reconstruction error that is produced during the training of each DAE-LSTM.
+I disregard the fitness of solutions found and focus solely on the reconstruction error that is produced during the training of each DAE-LSTM.
 The reconstruction error is measured for two separate populations, a training population $P_{train}$ that is used to train our DAE-LSTM as well as a hold-out validation population $P_{test}$.
@@
 For 100 hidden neurons, regular DAE-GP not only starts to overfit to the training data earlier than pre-trained DAE-GP, also it can be seen that pre-trained DAE-GP starts to overfit at a lower reconstruction error which would result in a higher model performance if training strategies such as early stopping would be employed.
-Doubling the number of hidden neurons to 200 results in very strong overfitting behaviour for both DAE-GP variants starting around the 200.th epoch.
+Doubling the number of hidden neurons to 200 results in very strong overfitting behavior for both DAE-GP variants starting around the 200.th epoch.
 
@@
 
-The reults for the experiment with varying number of hidden layers are summarized in figure \@ref(fig:first-gen-airfoil-byLayers).
+The results for the experiment with varying number of hidden layers are summarized in figure \@ref(fig:first-gen-airfoil-byLayers).
 Again, it can be observed that by increasing the dimension of the DAE-LSTM network both algorithms increasingly overfit to the training data.
@@
 epoch, doubling the number of hidden layers to two leads to both algorithms overfitting at around the 200.th epoch.
-In general over all three sub experiments, both algorithms start to overfit after approximately the same number of epochs but pre-training leads to reaching the point of overfitting at a higher reconstruction error which would result in a inferrior model performance if the early stopping training strategy would be used during model training.
+In general over all three sub experiments, both algorithms start to overfit after approximately the same number of epochs but pre-training leads to reaching the point of overfitting at a higher reconstruction error which would result in a inferior model performance if the early stopping training strategy would be used during model training.
 
@@
 
-After closely examining the effect of pre-training on DAE-GPs generalization behaviour another series of experiments is conducted to study how pre-training influences the overall search behaviour of DAE-GP.
+After closely examining the effect of pre-training on DAE-GPs generalization behavior another series of experiments is conducted to study how pre-training influences the overall search behavior of DAE-GP.
 The airfoil problem was selected as a first benchmark problem to test pre-trained DAE-GP against traditional DAE-GP.
@@
 
-The distribution of final fitness scores for pre-trained and regular DAE-GP are visualized as boxplots in figure  \@ref(fig:airfoil-distribution-fitness-30gens).
+The distribution of final fitness scores for pre-trained and regular DAE-GP are visualized as box plots in figure  \@ref(fig:airfoil-distribution-fitness-30gens).
 The distribution for regular DAE-GP shows to be preferable to that of pre-trained DAE-GP  since it shows a lower median RMSE of `r round(REG_TEST_MED_FF, 3)` for the test set (`r round(PT_TEST_MED_FF, 3)` for pre-trained DAE-GP) as well as less dispersion of the solutions measured by a standard deviation of `r round(REG_TEST_STDD_FF, 3)` (`r round(PT_TEST_STDD_FF, 3)` for pre-trained DAE-GP).
@@
 
-To conclude the first experiment, figure \@ref(fig:airfoil-popDiversity) shows the median population diversity measured by the normalized Levenshtein edit distance for both DAE-GP algorithms and regular GP as a benchmark.
+Another interesting metric to examine in population-based optimization algorithms is the diversity of the current generations population.
+Since individual solutions in DAE-GP are computer programs in the form of parse trees, I selected the normalized Levenshtein edit distance as my primary metric for tracking population diversity.
+Figure \@ref(fig:airfoil-popDiversity) shows the median population diversity for both DAE-GP algorithms and regular GP as a benchmark.
 While the population diversity steadily decreases over the generations for standard GP, DAE-GP variants strongly increase the population diversity in the first generations and keep them at a higher level during the evolutionary search.
@@
 The results are, again, based on $10$ individual runs per algorithm.
-As expected the results show, that for a small DAE-LSTM size pre-training negativley impacts the fitness of solution found by DAE-GP.
+As expected the results show, that for a small DAE-LSTM size pre-training negatively impacts the fitness of solution found by DAE-GP.
 The median final RMSE for the test set of pre-trained DAE-GP increases by a factor of `r round(PT_TEST_MED_FF_1HL / REG_TEST_MED_FF_1HL, 3)` to `r round(PT_TEST_MED_FF_1HL , 3)` compared to regular DAE-GP with `r round(REG_TEST_MED_FF_1HL , 3)`.
@@
 An alternative method of initializing $\hat{P}$ is to use the grow initialization method where solutions are not guaranteed to be full trees.
-This different structural composition of solutions inside the pre-training population might be another way to improve the pre-training strategy by training on a population that more closely resembles the structural properties of futher generations population where a lower percentage of full grown trees is to be expected.
+This different structural composition of solutions inside the pre-training population might be another way to improve the pre-training strategy by training on a population that more closely resembles the structural properties of further generations population where a lower percentage of full grown trees is to be expected.
 
@@
 Both alternative second generation Pre-Training as well as grow-initialized Pre-Training show a much higher median RMSE (18.88/17.21) compared to standard pre-training (8.55) and regular DAE-GP (6.99).
-Also both approaches do not demonstrate any benefit on the dispersion of solutions found resulting in similiar standard deviations if compared to standard pre-training.
+Also both approaches do not demonstrate any benefit on the dispersion of solutions found resulting in similar standard deviations if compared to standard pre-training.
 
@@
 
-<!-- TODO: Disclaimer, Runtime is implementation dependant, no GPU use, etc... -->
+The run-time analysis in this chapter is based on experiments that were all executed on an Intel® Xeon® W-2245 Processor without the use of graphical processing unit (GPU) acceleration in the training of DAE-LSTM networks.
+Since the framework that was used is written in the ,comparatively slow, python programming language, the absolute run-time measurements are only of limited meaningfulness.
+Nonetheless, by analyzing the relative run-time differences between DAE-GP and it's pre-trained alternative, the influence of pre-training on DAE-GP run-time can be explored.
+
+<!-- TODO: Disclaimer, Run-time is implementation dependent, no GPU use, etc... -->
 
@@
 
-Suprinsingly, as shown in figure \@ref(fig:airfoil-runtime-total), for the experiment conducted on the airfoil dataset with 2 hidden layers a strong negative impact of pre-training can be observed for the median runtime of DAE-GP.
-If the time spend on pre-training is included into the total runtime, the median duration for pre-trained DAE-GP over all 10 runs increases by a factor of `r round(PT_TIME / REG_TIME, 3)` from a median runtime of `r format(REG_TIME, scientific=F)` seconds for regular DAE-GP to `r format(PT_TIME, scientific=F)` seconds for the pre-trained algorithm.
-Even in the case where the time spend for pre-training is excluded from the total runtime, pre-training still increases the runtime of DAE-GP by a factor of `r round( PT_TIME_NOPT / REG_TIME, 3)` to `r format(PT_TIME_NOPT, scientific=F)` seconds.
+Surprisingly, as shown in figure \@ref(fig:airfoil-runtime-total), for the experiment conducted on the airfoil dataset with 2 hidden layers a strong negative impact of pre-training can be observed for the median run-time of DAE-GP.
+If the time spend on pre-training is included into the total run-time, the median duration for pre-trained DAE-GP over all 10 runs increases by a factor of `r round(PT_TIME / REG_TIME, 3)` from a median run-time of `r format(REG_TIME, scientific=F)` seconds for regular DAE-GP to `r format(PT_TIME, scientific=F)` seconds for the pre-trained algorithm.
+Even in the case where the time spend for pre-training is excluded from the total run-time, pre-training still increases the run-time of DAE-GP by a factor of `r round( PT_TIME_NOPT / REG_TIME, 3)` to `r format(PT_TIME_NOPT, scientific=F)` seconds.
 
@@
 
-To investigate the strong negative impact of pre-training on DAE-GP runtime, I first studied the influence of pre-training on the median number of training epochs that had to be spend at each generation for the training of the DAE-LSTM models $M_g$.
+To investigate the strong negative impact of pre-training on DAE-GP run-time, I first studied the influence of pre-training on the median number of training epochs that had to be spend at each generation for the training of the DAE-LSTM models $M_g$ (excluding the pre-training DAE-LSTM model $\hat{M}$).
 Figure \@ref(fig:airfoil-trainingEpochs) shows that pre-training, as expected, strongly reduces the number of training epochs that had to be executed at each generation until the termination condition was reached.
@@
 
-If we look at another time consuming operation in DAE-GP, the creation of new individuals by sampling the DAE-LSTM model $M_G$, we see one main reason for the large runtime increase by pre-training DAE-GP.
+If we look at another time consuming operation in DAE-GP, the creation of new individuals by sampling the DAE-LSTM model $M_G$, we see one main reason for the large run-time increase by pre-training DAE-GP.
 Figure \@ref(fig:airfoil-SamplingTime) shows the median sampling time spend during each generation of the search and it clearly demonstrates that pre-training does strongly increase the time spent on model sampling  in comparison to regular DAE-GP.
@@
 
-Suprisingly, even after correcting the runtime for both DAE-GP results by removing the time spend during sampling, my implementation of pre-training for DAE-GP still on average takes more time to finish the search process as shown in figure \@ref(fig:airfoil-runtime-noSampling-total).
-While traditional DAE-GP has a median runtime without sampling time of `r format(REG_TIME_NOSAMLING, scientific=F)` seconds , the runtime of pre-trained DAE-GP, without time for samling and pre-training, still increases by a factor of `r round(PT_TIME_NOSAMPLING_NOPT / REG_TIME_NOSAMLING, 3)` to `r format(PT_TIME_NOSAMPLING_NOPT, scientific=F)` seconds.
-While investigating other reasons for the differences in runtime, I verified that both DAE-GP variants use the exact same number of trainable parameters for each generations DAE-LSTMS models (total number of trainable parameters DAE-GP=`r format(REG_PARAMS, scientific=F)`, Pre-trained DAE-GP=`r format(PT_PARAMS, scientific=F)`).
+Surprisingly, even after correcting the run-time for both DAE-GP results by removing the time spend during sampling, my implementation of pre-training for DAE-GP still on average takes more time to finish the search process as shown in figure \@ref(fig:airfoil-runtime-noSampling-total).
+While traditional DAE-GP has a median run-time without sampling time of `r format(REG_TIME_NOSAMLING, scientific=F)` seconds , the run-time of pre-trained DAE-GP, without time for sampling and pre-training, still increases by a factor of `r round(PT_TIME_NOSAMPLING_NOPT / REG_TIME_NOSAMLING, 3)` to `r format(PT_TIME_NOSAMPLING_NOPT, scientific=F)` seconds.
 
@@
 ```
+While investigating other reasons for the differences in run-time, I verified that both DAE-GP variants use the exact same number of trainable parameters for each generations DAE-LSTMS models (total number of trainable parameters DAE-GP=`r format(REG_PARAMS, scientific=F)`, Pre-trained DAE-GP=`r format(PT_PARAMS, scientific=F)`).
 
@@
 For pre-trained DAE-GP another strategy had to be used, I used a fixed number of hidden neurons for each run that is set equal to the maximum length of all individuals that are present in the initial pre-training population.
-I decided to take this aproach instead of using a fixed number of hidden neurons for the fact that it simplifies the adjustment of the parameter for later experiments on different problems.
+I decided to take this approach instead of using a fixed number of hidden neurons for the fact that it simplifies the adjustment of the parameter for later experiments on different problems.
 It should nonetheless be mentioned that this strategy does give pre-training on average more hidden neurons per layer resulting in a more complex network (since the pre-training population is much larger than the regular population, the chances for larger individuals inside it are also higher).
@@
 
-Table X summarizes the median best fitness achieved after running each algorithm for 10 runs using a population size of 500 individuals for 30 generations.^[P-Values are marked with asteriks (1,2,3) for ($\alpha$)-levels of $0.1$, $0.05$ and $0.01$].
+Table X summarizes the median best fitness achieved after running each algorithm for 10 runs using a population size of 500 individuals for 30 generations.^[P-Values are marked with asterisk (1,2,3) for ($\alpha$)-levels of $0.1$, $0.05$ and $0.01$].
 The median best fitness over all 30 generations are visualized in figure X.
@@
 
-Another interesting metric to examine in EC based metaheuristics is the diversity of the current generations population.
-Since the individual solutions in DAE-GP are computer programs in the form ob parse trees, I selected the normalized Levenshtein edit distance as my primary metric for tracking population diversity.
-figure X shows how the median population diversity develops over the full number of generations for all benchmark problems.
+...
 
@@
 
-One important argument for using pre-training is the reduction of runtime and computational ressources for DAE-GP.
-An important factor in this, is the number of epochs that each DAE-LSTM has to be trained per generation.
-I studied the median number of training epochs that each generations DAE-LSTM $M_g$ (excluding the pre-training DAE-LSTM model $\hat{M}$) was trained until the training was stopped.
-The results are visualized in figure X and summarized in table X
+...
 
@@
 
-[@pmlr-v5-erhan09a]: Pre-Training effect especially usefull for lower-level layers -> Only adapt weights from low layers or embedd the pre-trained model?
+[@pmlr-v5-erhan09a]: Pre-Training effect especially useful for lower-level layers -> Only adapt weights from low layers or embed the pre-trained model?
 

### 2023-02-02 paper.tex
Thu Feb  2 13:00:07 CET 2023: update-remote.sh

@@
 One possible way of model building in EDA proposed by {[}\protect\hyperlink{ref-harmless_overfitting_eda}{11}{]} is to use denoising autoencoders (DAEs) as generative models to capture complex probability distributions of decision variables.
-DAE, a variation of the autoencoders (AE), are a widely used type of neural networks in the domain of unsupervized machine learning that maps \(n\) input variables to \(n\) output variables using a hidden representation.
-
-AE were introduced by {[}\protect\hyperlink{ref-ae_orig}{6}{]} to compress high-dimensional data into lower-dimensions. An AE consists of two different subunits:
+DAE, a variation of the autoencoders (AE), are a widely used type of neural networks in the domain of unsupervised machine learning that maps \(n\) input variables to \(n\) output variables using a hidden representation.
+
+AE were introduced by {[}\protect\hyperlink{ref-ae_orig}{6}{]} to compress high-dimensional data into lower-dimensions. An AE consists of two different sub-units:
 
@@
 \item
-  Encoder \(g(x)\): Encode input data to a smaller cental layer \(h\)
+  Encoder \(g(x)\): Encode input data to a smaller central layer \(h\)
 \item
@@
 
-The AE is trained to reduce the reconstruction error between input and output data, after the training procedure is finished the network is able to reduce the dimensionality of input data to a compressed representation{[}\protect\hyperlink{ref-ae_orig}{6}{]}.
+The AE is trained to reduce the reconstruction error between input and output data, after the training procedure is finished the network is able to reduce the dimensional of input data to a compressed representation{[}\protect\hyperlink{ref-ae_orig}{6}{]}.
 
@@
 
--- describe unsupervized vs supervized pre-training\ldots{}
+-- describe unsupervised vs supervised pre-training\ldots{}
 
@@
 Early research on pre-training for DAE by {[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{3}{]} comes to the conclusion that besides adding robustness to models the strategy also results in improved generalization and better performing models.
-The authors describe that unsupervised pre-training behaves like a regulizer on DAE networks, the mean test error and it's variance are reduced for DAE networks that are initialized from pre-trained weights if compared to the same architectures that use random initialization.
+The authors describe that unsupervised pre-training behaves like a regularizer on DAE networks, the mean test error and it's variance are reduced for DAE networks that are initialized from pre-trained weights if compared to the same architectures that use random initialization.
 
 Another important finding by {[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{3}{]} is that the positive effect of pre-training is dependent on both the depth of the network as well as the size of layers - while increasingly larger networks benefit increasingly more from pre-training the final performance of small architectures tends to be worse with pre-training than with randomized weight initialization.
-The authors find evidence that pre-training DAE is especially usefull for optimizing the parameters of the lower layers of the network.
+The authors find evidence that pre-training DAE is especially useful for optimizing the parameters of the lower layers of the network.
 
@@
 
-One difficulty in the implementation of a pre-training strategy into DAE-GP has been the detemination of the number of hidden neurons in side the DAE-LSTMs hidden layers.
+One difficulty in the implementation of a pre-training strategy into DAE-GP has been the determination of the number of hidden neurons in side the DAE-LSTMs hidden layers.
 The original description of DAE-GP {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]} uses a strategy where the number of hidden neurons for each hidden layer is dynamically set per generation to the maximum individual size inside the current population.
@@
 To study generalization behavior this series of experiments is conducted using DAE-GP with only a single generation until the search terminates.
-I disregard the fitness of solutions found and focus soley on the reconstruction error that is produced during the training of each DAE-LSTM.
+I disregard the fitness of solutions found and focus solely on the reconstruction error that is produced during the training of each DAE-LSTM.
 The reconstruction error is measured for two separate populations, a training population \(P_{train}\) that is used to train our DAE-LSTM as well as a hold-out validation population \(P_{test}\).
@@
 
-The reults for the experiment with varying number of hidden layers are summarized in figure \ref{fig:first-gen-airfoil-byLayers}.
+The results for the experiment with varying number of hidden layers are summarized in figure \ref{fig:first-gen-airfoil-byLayers}.
 Again, it can be observed that by increasing the dimension of the DAE-LSTM network both algorithms increasingly overfit to the training data.
@@
 
-The distribution of final fitness scores for pre-trained and regular DAE-GP are visualized as boxplots in figure \ref{fig:airfoil-distribution-fitness-30gens}.
+The distribution of final fitness scores for pre-trained and regular DAE-GP are visualized as box plots in figure \ref{fig:airfoil-distribution-fitness-30gens}.
 The distribution for regular DAE-GP shows to be preferable to that of pre-trained DAE-GP since it shows a lower median RMSE of 6.991 for the test set (7.68 for pre-trained DAE-GP) as well as less dispersion of the solutions measured by a standard deviation of 11.989 (12.548 for pre-trained DAE-GP).
@@
 
-To conclude the first experiment, figure \ref{fig:airfoil-popDiversity} shows the median population diversity measured by the normalized Levenshtein edit distance for both DAE-GP algorithms and regular GP as a benchmark.
+Another interesting metric to examine in population-based optimization algorithms is the diversity of the current generations population.
+Since individual solutions in DAE-GP are computer programs in the form of parse trees, I selected the normalized Levenshtein edit distance as my primary metric for tracking population diversity.
+Figure \ref{fig:airfoil-popDiversity} shows the median population diversity for both DAE-GP algorithms and regular GP as a benchmark.
 While the population diversity steadily decreases over the generations for standard GP, DAE-GP variants strongly increase the population diversity in the first generations and keep them at a higher level during the evolutionary search.
@@
 The results are, again, based on \(10\) individual runs per algorithm.
-As expected the results show, that for a small DAE-LSTM size pre-training negativley impacts the fitness of solution found by DAE-GP.
+As expected the results show, that for a small DAE-LSTM size pre-training negatively impacts the fitness of solution found by DAE-GP.
 The median final RMSE for the test set of pre-trained DAE-GP increases by a factor of 1.287 to 16.213 compared to regular DAE-GP with 12.594.
@@
 An alternative method of initializing \(\hat{P}\) is to use the grow initialization method where solutions are not guaranteed to be full trees.
-This different structural composition of solutions inside the pre-training population might be another way to improve the pre-training strategy by training on a population that more closely resembles the structural properties of futher generations population where a lower percentage of full grown trees is to be expected.
+This different structural composition of solutions inside the pre-training population might be another way to improve the pre-training strategy by training on a population that more closely resembles the structural properties of further generations population where a lower percentage of full grown trees is to be expected.
 
@@
 
-Suprinsingly, as shown in figure \ref{fig:airfoil-runtime-total}, for the experiment conducted on the airfoil dataset with 2 hidden layers a strong negative impact of pre-training can be observed for the median runtime of DAE-GP.
-If the time spend on pre-training is included into the total runtime, the median duration for pre-trained DAE-GP over all 10 runs increases by a factor of 3.121 from a median runtime of 4712.303 seconds for regular DAE-GP to 14706.47 seconds for the pre-trained algorithm.
-Even in the case where the time spend for pre-training is excluded from the total runtime, pre-training still increases the runtime of DAE-GP by a factor of 2.66 to 12536.89 seconds.
+The run-time analysis in this chapter is based on experiments that were all executed on an Intel® Xeon® W-2245 Processor without the use of graphical processing unit (GPU) acceleration in the training of DAE-LSTM networks.
+Since the framework that was used is written in the ,comparatively slow, python programming language, the absolute run-time measurements are only of limited meaningfulness.
+Nonetheless, by analyzing the relative run-time differences between DAE-GP and it's pre-trained alternative, the influence of pre-training on DAE-GP run-time can be explored.
+
+Surprisingly, as shown in figure \ref{fig:airfoil-runtime-total}, for the experiment conducted on the airfoil dataset with 2 hidden layers a strong negative impact of pre-training can be observed for the median run-time of DAE-GP.
+If the time spend on pre-training is included into the total run-time, the median duration for pre-trained DAE-GP over all 10 runs increases by a factor of 3.121 from a median run-time of 4712.303 seconds for regular DAE-GP to 14706.47 seconds for the pre-trained algorithm.
+Even in the case where the time spend for pre-training is excluded from the total run-time, pre-training still increases the run-time of DAE-GP by a factor of 2.66 to 12536.89 seconds.
 
@@
 
-To investigate the strong negative impact of pre-training on DAE-GP runtime, I first studied the influence of pre-training on the median number of training epochs that had to be spend at each generation for the training of the DAE-LSTM models \(M_g\).
+To investigate the strong negative impact of pre-training on DAE-GP run-time, I first studied the influence of pre-training on the median number of training epochs that had to be spend at each generation for the training of the DAE-LSTM models \(M_g\) (excluding the pre-training DAE-LSTM model \(\hat{M}\)).
 Figure \ref{fig:airfoil-trainingEpochs} shows that pre-training, as expected, strongly reduces the number of training epochs that had to be executed at each generation until the termination condition was reached.
@@
 
-If we look at another time consuming operation in DAE-GP, the creation of new individuals by sampling the DAE-LSTM model \(M_G\), we see one main reason for the large runtime increase by pre-training DAE-GP.
+If we look at another time consuming operation in DAE-GP, the creation of new individuals by sampling the DAE-LSTM model \(M_G\), we see one main reason for the large run-time increase by pre-training DAE-GP.
 Figure \ref{fig:airfoil-SamplingTime} shows the median sampling time spend during each generation of the search and it clearly demonstrates that pre-training does strongly increase the time spent on model sampling in comparison to regular DAE-GP.
@@
 
-Suprisingly, even after correcting the runtime for both DAE-GP results by removing the time spend during sampling, my implementation of pre-training for DAE-GP still on average takes more time to finish the search process as shown in figure \ref{fig:airfoil-runtime-noSampling-total}.
-While traditional DAE-GP has a median runtime without sampling time of 3550.277 seconds , the runtime of pre-trained DAE-GP, without time for samling and pre-training, still increases by a factor of 1.865 to 6622.027 seconds.
-While investigating other reasons for the differences in runtime, I verified that both DAE-GP variants use the exact same number of trainable parameters for each generations DAE-LSTMS models (total number of trainable parameters DAE-GP=573473, Pre-trained DAE-GP=573473).
+Surprisingly, even after correcting the run-time for both DAE-GP results by removing the time spend during sampling, my implementation of pre-training for DAE-GP still on average takes more time to finish the search process as shown in figure \ref{fig:airfoil-runtime-noSampling-total}.
+While traditional DAE-GP has a median run-time without sampling time of 3550.277 seconds , the run-time of pre-trained DAE-GP, without time for sampling and pre-training, still increases by a factor of 1.865 to 6622.027 seconds.
+While investigating other reasons for the differences in run-time, I verified that both DAE-GP variants use the exact same number of trainable parameters for each generations DAE-LSTMS models (total number of trainable parameters DAE-GP=573473, Pre-trained DAE-GP=573473).
 
@@
 
-Table X summarizes the median best fitness achieved after running each algorithm for 10 runs using a population size of 500 individuals for 30 generations.\footnote{P-Values are marked with asteriks (1,2,3) for (\(\alpha\))-levels of \(0.1\), \(0.05\) and \(0.01\)}.
+Table X summarizes the median best fitness achieved after running each algorithm for 10 runs using a population size of 500 individuals for 30 generations.\footnote{P-Values are marked with asterisk (1,2,3) for (\(\alpha\))-levels of \(0.1\), \(0.05\) and \(0.01\)}.
 The median best fitness over all 30 generations are visualized in figure X.
@@
 
-Another interesting metric to examine in EC based metaheuristics is the diversity of the current generations population.
-Since the individual solutions in DAE-GP are computer programs in the form ob parse trees, I selected the normalized Levenshtein edit distance as my primary metric for tracking population diversity.
-figure X shows how the median population diversity develops over the full number of generations for all benchmark problems.
+\ldots{}
 
@@
 
-One important argument for using pre-training is the reduction of runtime and computational ressources for DAE-GP.
-An important factor in this, is the number of epochs that each DAE-LSTM has to be trained per generation.
-I studied the median number of training epochs that each generations DAE-LSTM \(M_g\) (excluding the pre-training DAE-LSTM model \(\hat{M}\)) was trained until the training was stopped.
-The results are visualized in figure X and summarized in table X
+\ldots{}
 
@@
 
-{[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{3}{]}: Pre-Training effect especially usefull for lower-level layers -\textgreater{} Only adapt weights from low layers or embedd the pre-trained model?
+{[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{3}{]}: Pre-Training effect especially useful for lower-level layers -\textgreater{} Only adapt weights from low layers or embed the pre-trained model?
 

### 2023-02-02 data/summarize_data.py

@@ -45,18 +45,18 @@ BASE_TITLE="Symbolic Regression"
 
 
 RESULTS = {
-    "airfoil_1hl" : "/Users/rmn/github/master_thesis/data/airfoil_1hl_maxIndSize_fullRun_30gens",
-    "airfoil_2hl" : "/Users/rmn/github/master_thesis/data/airfoil_2hl_maxIndSize_fullRun_30gens",
-    "bostonHousing": "/Users/rmn/github/master_thesis/data/bostonHousing_2hl_maxIndSize_fullRun_30gens",
-    "energyCooling": "/Users/rmn/github/master_thesis/data/energyCooling_2hl_FullRun_30gens",
-    "concrete": "/Users/rmn/github/master_thesis/data/concrete_2hl_FullRun_30gens"
+    "airfoil_1hl" : "/Users/rmn/github/master_thesis/data/airfoil_1hl_150hn_fullRun_30gens",
+    "airfoil_2hl" : "/Users/rmn/github/master_thesis/data/airfoil_2hl_150hn_fullRun_30gens",
+    "bostonHousing": "/Users/rmn/github/master_thesis/data/bostonHousing_2hl_150hn_fullRun_30gens",
+    "energyCooling": "/Users/rmn/github/master_thesis/data/energyCooling_2hl_150hn_fullRun_30gens",
+    "concrete": "/Users/rmn/github/master_thesis/data/concrete_2hl_150hn_30gens"
 }
 
 RESULTS_2hl = {
-    "Airfoil" : "/Users/rmn/github/master_thesis/data/airfoil_2hl_maxIndSize_fullRun_30gens",
-    "BostonHousing": "/Users/rmn/github/master_thesis/data/bostonHousing_2hl_maxIndSize_fullRun_30gens",
-    "energyCooling": "/Users/rmn/github/master_thesis/data/energyCooling_2hl_FullRun_30gens",
-    "Concrete": "/Users/rmn/github/master_thesis/data/concrete_2hl_FullRun_30gens"
+    "Airfoil" :  "/Users/rmn/github/master_thesis/data/airfoil_2hl_150hn_fullRun_30gens",
+    "BostonHousing": "/Users/rmn/github/master_thesis/data/bostonHousing_2hl_150hn_fullRun_30gens",
+    "energyCooling": "/Users/rmn/github/master_thesis/data/energyCooling_2hl_150hn_fullRun_30gens",
[…]

### 2023-02-02 paper.Rmd
Thu Feb  2 17:00:09 CET 2023: update-remote.sh
Thu Feb  2 18:00:08 CET 2023: update-remote.sh

@@
 While investigating other reasons for the differences in run-time, I verified that both DAE-GP variants use the exact same number of trainable parameters for each generations DAE-LSTMS models (total number of trainable parameters DAE-GP=`r format(REG_PARAMS, scientific=F)`, Pre-trained DAE-GP=`r format(PT_PARAMS, scientific=F)`).
+After profiling a single run of DAE-GP and pre-trained DAE-GP my best explanation for the remaining differences in run-time is an inefficiency in my implementation of pre-training.
+Figure  \@ref(tab:full-run-realWorldSymReg-problems) shows the cumulative time spend for the 20 most expensive function calls.
+Even though it is difficult to identify the exact reasons for the performance hit for pre-trained DAE-GP, it shows two interesting findings: Firstly, both algorithms spend relatively more time during model sampling (calls to `DAE_LSTM.py:208(sample)`) than during model training (calls to `training.py:1303(fit)`).
+Secondly, it shows that internal error handling of the baseline neural network library ´keras´ (calls to `traceback_utils.py:138|50(error_handler)`) takes up a comparatively larger amount of run time for pre-trained DAE-GP than for regular DAE-GP.
+
+```{r airfoil-profiling, echo=FALSE, fig.cap="Cumulative Time consumption by Function Calls (Top 20) - Airfoil", out.width="80%"}
+knitr::include_graphics("./img/Airfoil_RuntimeProfiling/cum_Runtime_HBar.png")
+```
 

### 2023-02-02 paper.tex
Thu Feb  2 18:00:08 CET 2023: update-remote.sh

@@
 While traditional DAE-GP has a median run-time without sampling time of 3550.277 seconds , the run-time of pre-trained DAE-GP, without time for sampling and pre-training, still increases by a factor of 1.865 to 6622.027 seconds.
+
+\begin{figure}[c]
+
+{\centering \includegraphics[width=0.8\linewidth]{./img/airfoil_2hl_150hn_fullRun_30gens_TimeBenchmark/median_runtime_withoutSampling}
+
+}
+
+\caption{Median Runtime excluding Time for Sampling - Airfoil}\label{fig:airfoil-runtime-noSampling-total}
+\end{figure}
+
 While investigating other reasons for the differences in run-time, I verified that both DAE-GP variants use the exact same number of trainable parameters for each generations DAE-LSTMS models (total number of trainable parameters DAE-GP=573473, Pre-trained DAE-GP=573473).
-
-\begin{figure}[c]
-
-{\centering \includegraphics[width=0.8\linewidth]{./img/airfoil_2hl_150hn_fullRun_30gens_TimeBenchmark/median_runtime_withoutSampling}
-
-}
-
-\caption{Median Runtime excluding Time for Sampling - Airfoil}\label{fig:airfoil-runtime-noSampling-total}
+After profiling a single run of DAE-GP and pre-trained DAE-GP my best explanation for the remaining differences in run-time is an inefficiency in my implementation of pre-training.
+Figure \ref{tab:full-run-realWorldSymReg-problems} shows the cumulative time spend for the 20 most expensive function calls.
+Even though it is difficult to identify the exact reasons for the performance hit for pre-trained DAE-GP, it shows two interesting findings: Firstly, both algorithms spend relatively more time during model sampling (calls to \texttt{DAE\_LSTM.py:208(sample)}) than during model training (calls to \texttt{training.py:1303(fit)}).
+Secondly, it shows that internal error handling of the baseline neural network library ´keras´ (calls to \texttt{traceback\_utils.py:138\textbar{}50(error\_handler)}) takes up a comparatively larger amount of run time for pre-trained DAE-GP than for regular DAE-GP.
+
+\begin{figure}[c]
+
+{\centering \includegraphics[width=0.8\linewidth]{./img/Airfoil_RuntimeProfiling/cum_Runtime_HBar}
+
+}
+
+\caption{Cumulative Time consumption by Function Calls (Top 20) - Airfoil}\label{fig:airfoil-profiling}
 \end{figure}

### 2023-02-02 paper.Rmd
autogit

@@
 indent: false
-fontsize: 11pt
+fontsize: 12pt
 bibliography: ./ref/ref.bib
@@
 The distribution of final fitness scores for pre-trained and regular DAE-GP are visualized as box plots in figure  \@ref(fig:airfoil-distribution-fitness-30gens).
-The distribution for regular DAE-GP shows to be preferable to that of pre-trained DAE-GP  since it shows a lower median RMSE of `r round(REG_TEST_MED_FF, 3)` for the test set (`r round(PT_TEST_MED_FF, 3)` for pre-trained DAE-GP) as well as less dispersion of the solutions measured by a standard deviation of `r round(REG_TEST_STDD_FF, 3)` (`r round(PT_TEST_STDD_FF, 3)` for pre-trained DAE-GP).
+The distribution for regular DAE-GP shows to be preferable to that of pre-trained DAE-GP  since it has a lower median RMSE of `r round(REG_TEST_MED_FF, 3)` for the test set (`r round(PT_TEST_MED_FF, 3)` for pre-trained DAE-GP) as well as less dispersion of the solutions measured by a standard deviation of `r round(REG_TEST_STDD_FF, 3)` (`r round(PT_TEST_STDD_FF, 3)` for pre-trained DAE-GP).
 For reference, the median final RMSE for the test set of standard GP is at `r round(GP_TEST_MED_FF, 3)` which is larger by a factor of `r round(GP_TEST_MED_FF / PT_TEST_MED_FF, 3)` if compared to pre-trained DAE-GP.
@@
 The run-time analysis in this chapter is based on experiments that were all executed on an Intel® Xeon® W-2245 Processor without the use of graphical processing unit (GPU) acceleration in the training of DAE-LSTM networks.
-Since the framework that was used is written in the ,comparatively slow, python programming language, the absolute run-time measurements are only of limited meaningfulness.
+Since the framework that was used is written in the, comparatively slow, python programming language, the absolute run-time measurements are only of limited meaningfulness.
 Nonetheless, by analyzing the relative run-time differences between DAE-GP and it's pre-trained alternative, the influence of pre-training on DAE-GP run-time can be explored.
@@
 After profiling a single run of DAE-GP and pre-trained DAE-GP my best explanation for the remaining differences in run-time is an inefficiency in my implementation of pre-training.
-Figure  \@ref(tab:full-run-realWorldSymReg-problems) shows the cumulative time spend for the 20 most expensive function calls.
+Figure  \@ref(fig:airfoil-profiling) shows the cumulative time spend for the 20 most expensive function calls.
 Even though it is difficult to identify the exact reasons for the performance hit for pre-trained DAE-GP, it shows two interesting findings: Firstly, both algorithms spend relatively more time during model sampling (calls to `DAE_LSTM.py:208(sample)`) than during model training (calls to `training.py:1303(fit)`).
-Secondly, it shows that internal error handling of the baseline neural network library ´keras´ (calls to `traceback_utils.py:138|50(error_handler)`) takes up a comparatively larger amount of run time for pre-trained DAE-GP than for regular DAE-GP.
+Secondly, it shows that internal error handling of the baseline neural network library `keras` (calls to `traceback_utils.py:138(error_handler)` and `traceback_utils.py:59(error_handler)`) takes up a comparatively larger amount of run time for pre-trained DAE-GP than for regular DAE-GP.
 
@@
 
-To gather more confidence in the previous results, the pre-training strategy for DAE-GP was tested for three additional real world symbolic regression problems taken from @machine_learning_repo.
+To gather more confidence in the previous results, the pre-training strategy for DAE-GP was applied for three additional real world symbolic regression problems taken from @machine_learning_repo and tested for statistically significant differences in the underlying distributions.
 Table \@ref(tab:full-run-realWorldSymReg-problems) briefly summarizes the benchmark problems.

### 2023-02-02 paper.tex
autogit

@@
 \documentclass[
-  11pt,
+  12pt,
 ]{article}
@@
 The distribution of final fitness scores for pre-trained and regular DAE-GP are visualized as box plots in figure \ref{fig:airfoil-distribution-fitness-30gens}.
-The distribution for regular DAE-GP shows to be preferable to that of pre-trained DAE-GP since it shows a lower median RMSE of 6.991 for the test set (7.68 for pre-trained DAE-GP) as well as less dispersion of the solutions measured by a standard deviation of 11.989 (12.548 for pre-trained DAE-GP).
+The distribution for regular DAE-GP shows to be preferable to that of pre-trained DAE-GP since it has a lower median RMSE of 6.991 for the test set (7.68 for pre-trained DAE-GP) as well as less dispersion of the solutions measured by a standard deviation of 11.989 (12.548 for pre-trained DAE-GP).
 For reference, the median final RMSE for the test set of standard GP is at 32.024 which is larger by a factor of 4.17 if compared to pre-trained DAE-GP.
@@
 After profiling a single run of DAE-GP and pre-trained DAE-GP my best explanation for the remaining differences in run-time is an inefficiency in my implementation of pre-training.
-Figure \ref{tab:full-run-realWorldSymReg-problems} shows the cumulative time spend for the 20 most expensive function calls.
+Figure \ref{fig:airfoil-profiling} shows the cumulative time spend for the 20 most expensive function calls.
 Even though it is difficult to identify the exact reasons for the performance hit for pre-trained DAE-GP, it shows two interesting findings: Firstly, both algorithms spend relatively more time during model sampling (calls to \texttt{DAE\_LSTM.py:208(sample)}) than during model training (calls to \texttt{training.py:1303(fit)}).
-Secondly, it shows that internal error handling of the baseline neural network library ´keras´ (calls to \texttt{traceback\_utils.py:138\textbar{}50(error\_handler)}) takes up a comparatively larger amount of run time for pre-trained DAE-GP than for regular DAE-GP.
+Secondly, it shows that internal error handling of the baseline neural network library \texttt{keras} (calls to \texttt{traceback\_utils.py:138(error\_handler)} and \texttt{traceback\_utils.py:59(error\_handler)}) takes up a comparatively larger amount of run time for pre-trained DAE-GP than for regular DAE-GP.
 
@@
 
-To gather more confidence in the previous results, the pre-training strategy for DAE-GP was tested for three additional real world symbolic regression problems taken from {[}\protect\hyperlink{ref-machine_learning_repo}{2}{]}.
+To gather more confidence in the previous results, the pre-training strategy for DAE-GP was applied for three additional real world symbolic regression problems taken from {[}\protect\hyperlink{ref-machine_learning_repo}{2}{]} and tested for statistically significant differences in the underlying distributions.
 Table \ref{tab:full-run-realWorldSymReg-problems} briefly summarizes the benchmark problems.

### 2023-02-02 paper.Rmd
Thu Feb  2 22:00:05 CET 2023: update-remote.sh

@@
 
+## Benefits of using Pre-Training
+
+The experiments conducted in the scope of this master thesis did find that the implemented pre-training strategy resulted in an increased population diversity measured by the normalized levenshtein edit distance.
+Increasing the population diversity can be an important mechanism in controlling the search behavior of population based metaheuristics and is especially useful for increasing the exploration of the search space of possible solutions [@QUELLE?]
+
+## Drawbacks of using Pre-Training
+
 The experiments conducted in this thesis did not find any significant benefits in using a pre-training strategy for using DAE-GP to solve real world symbolic regression problems.

### 2023-02-02 paper.Rmd
Thu Feb  2 23:00:06 CET 2023: update-remote.sh

@@
 The objective of the airfoil problem is to find a function that accurately predicts the output variable $y$ by taking in a subset of the input variables $x1,x2,x3,x4,x5$.
-The function set used by all DAE-GP variations for the airfoil problem is summarized in table 2, the terminal set consists of the 5 input variables $x1,x2,x3,x4,x5$ and ephemeral random integers in the range of $[{}-5,..,5]$ [@dae-gp_2022_symreg].
+The function set used by all DAE-GP variations for the airfoil problem is summarized in \@ref(tab:airfoil-fullRun-2hl-maxIndSize-params), the terminal set consists of the 5 input variables $x1,x2,x3,x4,x5$ and the ephemeral integer constants listed in \@ref(tab:airfoil-fullRun-2hl-maxIndSize-params)[@dae-gp_2022_symreg].
 
@@
 
-To gather a deeper understanding about the effect of pre-training DAE-LSTM in DAE-GP a series of experiments was conducted using the airfoil dataset for symbolic regression while using and different parameter configurations for the number of hidden layers as well as the number of hidden neurons per hidden layer.
+To gather a deeper understanding about the effect of pre-training DAE-LSTM in DAE-GP a series of experiments was conducted using the airfoil dataset for symbolic regression while using different parameter configurations for the number of hidden layers as well as the number of hidden neurons per hidden layer.
 
@@
 
-In general I expect that with a growing number of model parameters (either by adding more hidden layers or more hidden neurons per layer) the DAE-LSTM will be more prone to overfit to the training population $P_{train}$ resulting in a small reconstruction error for the training population and a large one for the validation population $P_{test}$.
+In general I expect that with a growing number of trainable parameters (either by adding more hidden layers or more hidden neurons per layer) the DAE-LSTM will be more prone to overfit to the training population $P_{train}$ resulting in a small reconstruction error for the training population and a large one for the validation population $P_{test}$.
 The reason for this effect is that a larger network trained over an extended period of time (without the use of strategies like early stopping), has much more potential to learn noise from the training dataset than a smaller network, which is more likely to result in worsening performance for previously unseen cases [@weigend1994overfitting].
@@
 ```
+... describe differences in distribution, interquartile range blablabla...
 
@@
 The experiments conducted in the scope of this master thesis did find that the implemented pre-training strategy resulted in an increased population diversity measured by the normalized levenshtein edit distance.
-Increasing the population diversity can be an important mechanism in controlling the search behavior of population based metaheuristics and is especially useful for increasing the exploration of the search space of possible solutions [@QUELLE?]
+Increasing the population diversity can be an important mechanism in controlling the search behavior of population based metaheuristics and is especially useful for increasing the exploration of the search space of possible solutions !QUELLE!.
+
+Another significant benefit of using pre-training in DAE-GP is that it is highly efficient in decreasing the number of training epochs that are necessary to train each generations DAE-LSTM model.
+Even though pre-training increased the median totalrun-time for the tested benchmark problem in combination with the chosen DAE-LSTM dimensions, reducing the number of training epochs might still be a possible way to lower the total
 

### 2023-02-03 paper.Rmd
Fri Feb  3 09:00:08 CET 2023: update-remote.sh

@@
 
-# bash /Users/rmn/masterThesis/eda-gp-2020/experiments/airfoil_1hl_maxIndSize_fullRun_30gens/re_evaluate_airfoil_1hl_maxIndSize_30gen.sh
-
-# bash /Users/rmn/masterThesis/eda-gp-2020/experiments/airfoil_2hl_maxIndSize_fullRun_30gens/re_evaluate_airfoil_2hl_maxIndSize_30gen.sh
-
 # energy (cooling)
@@
 # boston housing
-bash /Users/rmn/masterThesis/eda-gp-2020/experiments/bostonHousing_2hl_maxIndSize_fullRun_30gens/re_evaluate_bostonHousing_2hl_maxIndSize_fullRun_30gens.sh
+bash /Users/rmn/masterThesis/eda-gp-2020/experiments/bostonHousing_2hl_150hn_fullRun_30gens/re_evaluate_bostonHousing_2hl_150hn_fullRun_30gens.sh
 
 # concrete
-bash /Users/rmn/masterThesis/eda-gp-2020/experiments/concrete_2hl_maxIndSize_fullRun_30gens/re_evaluate_concrete_2hl_maxIndSize_fullRun_30gens.sh
+bash /Users/rmn/masterThesis/eda-gp-2020/experiments/concrete_2hl_150hn_fullRun_30gens/re_evaluate_concrete_2hl_150hn_fullRun_30gens.sh
 
@@
 
+<!-- add test results for statistical significance! -->
+
 The experiments conducted in the scope of this master thesis did find that the implemented pre-training strategy resulted in an increased population diversity measured by the normalized levenshtein edit distance.
@@
 Another significant benefit of using pre-training in DAE-GP is that it is highly efficient in decreasing the number of training epochs that are necessary to train each generations DAE-LSTM model.
-Even though pre-training increased the median totalrun-time for the tested benchmark problem in combination with the chosen DAE-LSTM dimensions, reducing the number of training epochs might still be a possible way to lower the total
+Even though pre-training increased the median total run-time for the tested benchmark problem in combination with the chosen DAE-LSTM dimensions, reducing the number of training epochs might be a possible way to lower the runtime of DAE-GP if further performance optimizations to the source code, especially during model sampling, can be achieved.
 
@@
 
-The experiments conducted in this thesis did not find any significant benefits in using a pre-training strategy for using DAE-GP to solve real world symbolic regression problems.
+The experiments conducted in this thesis did not find any significant benefits on both solution quality as well as run-time for solving pre-training strategy for using DAE-GP to solve real world symbolic regression problems.
 

### 2023-02-03 paper.Rmd
Fri Feb  3 10:00:08 CET 2023: update-remote.sh

@@
 bash /Users/rmn/masterThesis/eda-gp-2020/experiments/airfoil_1hl_150hn_fullRun_30gens/re_evaluate_airfoil_1hl_150hn_30gen.sh
+
+bash /Users/rmn/masterThesis/eda-gp-2020/experiments/airfoil_2hl_maxIndSize_fullRun_30gens/re_evaluate_airfoil_2hl_maxIndSize_30gen.sh
 
@@
 For a pre-training implementation, this strategy can not be easily adapted since it leads to a changing number of neurons at each generation resulting in different dimensions of the DAE-LSTM.
-To allow the sharing of the state of the pre-trained parameters $\theta_{\hat{M}}$ from the pre-trained model $\hat{M}$ to each $M_g$ for $g\in{1,...,g_{max}}$  I set the number of hidden neurons per hidden layer for all pre-trained runs statically to the maximum individual size inside the pre-training population $\hat{P}_{train}$
+To allow the sharing of the state of the pre-trained parameters $\theta_{\hat{M}}$ from the pre-trained model $\hat{M}$ to each $M_g$ for $g\in{1,...,g_{max}}$ the number of hidden neurons per hidden layer for all DAE-GP runs, pre-trained as well as regular, is statically set to a predefined value.
 
@@
 knitr::include_graphics("./img/airfoil_2hl_150hn_fullRun_30gens_withGP/median_popDiversity_byGens_levOnly.png")
+```
+### Dynamic Adjustment of Hidden Neurons
+
+To ensure a fair comparison between pre-trained and regular DAE-GP, the prior experiments used a static number of hidden neurons for each hidden Layer inside the DAE-LSTM networks.
+For comparison, figure \@ref(fig:airfoil-maxIndSize-params) shows the total number of trainable parameters for each generation's DAE-LSTM of another experiment where regular DAE-GP was tested with dynamic adjustment of the number of hidden neurons as described in [@dae-gp_2020_rtree] and [@dae-gp_2022_symreg].
+Again, the experiments was based on $10$ individual runs and otherwise used the same hyperparameters as the previous one.
+It clearly shows that DAE-GP can strongly reduce the number of trainable parameters by adjusting the depth of it's hidden layers to the current generations population which will further reduce the amount of training necessary and likely reduce the run-time of DAE-GP.
+
+```{r airfoil-maxIndSize-params, echo=FALSE, fig.cap="Median Number of trainable Parameters over 30 Generations - Airfoil (2HL)", out.width="80%"}
+knitr::include_graphics("./img/airfoil_2hl_maxIndSize_fullRun_30gens/trainableParams.png")
+```
+
+As expected, this reduction of hidden neurons by dynamic adjustment, does also result in a reduction of the fitness of solutions found.
+Figure \@ref(fig:airfoil-maxIndSize-fitness) shows that pre-trained DAE-GP does achieve a higher median fitness than regular DAE-GP with dynmaic adjustment of hidden neurons.
+
+```{r airfoil-maxIndSize-fitness, echo=FALSE, fig.cap="Median Fitness over 30 Generations - Airfoil (2HL), dynamic adjustment of regular DAE-GP", out.width="80%"}
+knitr::include_graphics("./img/airfoil_2hl_maxIndSize_fullRun_30gens/median_fitness_byGens.png")
 ```
@@
 
-## Drawbacks of using Pre-Training
+## Disadvantages of using Pre-Training
 
-The experiments conducted in this thesis did not find any significant benefits on both solution quality as well as run-time for solving pre-training strategy for using DAE-GP to solve real world symbolic regression problems.
+The experiments conducted in this thesis did not find any significant benefits on both solution quality as well as run-time for solving real world symbolic regression problems with pre-trained DAE-GP in comparison to regular DAE-GP.
+
+One major drawback of using pre-training with DAE-GP
 

### 2023-02-03 paper.Rmd
autogit

@@
 As expected, this reduction of hidden neurons by dynamic adjustment, does also result in a reduction of the fitness of solutions found.
-Figure \@ref(fig:airfoil-maxIndSize-fitness) shows that pre-trained DAE-GP does achieve a higher median fitness than regular DAE-GP with dynmaic adjustment of hidden neurons.
+Figure \@ref(fig:airfoil-maxIndSize-fitness) shows that pre-trained DAE-GP, with it's more complex DAE-LSTMs over the full 30 generations, does achieve a higher median fitness than regular DAE-GP with dynamic adjustment of hidden neurons.
+Pre-Training does therefore introduce a new Trade-Off decision into the adjustment of DAE-GP hyperparameters where the number of hidden neurons needs to be carefully selected since it positivly impacts solution quality but also increases the computational cost of DAE-LSTM training.
 
-```{r airfoil-maxIndSize-fitness, echo=FALSE, fig.cap="Median Fitness over 30 Generations - Airfoil (2HL), dynamic adjustment of regular DAE-GP", out.width="80%"}
+```{r airfoil-maxIndSize-fitness, echo=FALSE, fig.cap="Median Best Fitness over 30 Generations - Airfoil (2HL) with dynamic adjustment of regular DAE-GP", out.width="80%"}
 knitr::include_graphics("./img/airfoil_2hl_maxIndSize_fullRun_30gens/median_fitness_byGens.png")

### 2023-02-03 paper.tex
autogit

@@
 \author{}
-\date{\vspace{-2.5em}Date of Submission: 2023-02-02}
+\date{\vspace{-2.5em}Date of Submission: 2023-02-03}
 
@@
 For a pre-training implementation, this strategy can not be easily adapted since it leads to a changing number of neurons at each generation resulting in different dimensions of the DAE-LSTM.
-To allow the sharing of the state of the pre-trained parameters \(\theta_{\hat{M}}\) from the pre-trained model \(\hat{M}\) to each \(M_g\) for \(g\in{1,...,g_{max}}\) I set the number of hidden neurons per hidden layer for all pre-trained runs statically to the maximum individual size inside the pre-training population \(\hat{P}_{train}\)
+To allow the sharing of the state of the pre-trained parameters \(\theta_{\hat{M}}\) from the pre-trained model \(\hat{M}\) to each \(M_g\) for \(g\in{1,...,g_{max}}\) the number of hidden neurons per hidden layer for all DAE-GP runs, pre-trained as well as regular, is statically set to a predefined value.
 
@@
 The objective of the airfoil problem is to find a function that accurately predicts the output variable \(y\) by taking in a subset of the input variables \(x1,x2,x3,x4,x5\).
-The function set used by all DAE-GP variations for the airfoil problem is summarized in table 2, the terminal set consists of the 5 input variables \(x1,x2,x3,x4,x5\) and ephemeral random integers in the range of \([{}-5,..,5]\) {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]}.
+The function set used by all DAE-GP variations for the airfoil problem is summarized in \ref{tab:airfoil-fullRun-2hl-maxIndSize-params}, the terminal set consists of the 5 input variables \(x1,x2,x3,x4,x5\) and the ephemeral integer constants listed in \ref{tab:airfoil-fullRun-2hl-maxIndSize-params}{[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]}.
 
@@
 
-In general I expect that with a growing number of model parameters (either by adding more hidden layers or more hidden neurons per layer) the DAE-LSTM will be more prone to overfit to the training population \(P_{train}\) resulting in a small reconstruction error for the training population and a large one for the validation population \(P_{test}\).
+In general I expect that with a growing number of trainable parameters (either by adding more hidden layers or more hidden neurons per layer) the DAE-LSTM will be more prone to overfit to the training population \(P_{train}\) resulting in a small reconstruction error for the training population and a large one for the validation population \(P_{test}\).
 The reason for this effect is that a larger network trained over an extended period of time (without the use of strategies like early stopping), has much more potential to learn noise from the training dataset than a smaller network, which is more likely to result in worsening performance for previously unseen cases {[}\protect\hyperlink{ref-weigend1994overfitting}{15}{]}.
@@
 
+\hypertarget{dynamic-adjustment-of-hidden-neurons}{%
+\subsubsection{Dynamic Adjustment of Hidden Neurons}\label{dynamic-adjustment-of-hidden-neurons}}
+
+To ensure a fair comparison between pre-trained and regular DAE-GP, the prior experiments used a static number of hidden neurons for each hidden Layer inside the DAE-LSTM networks.
+For comparison, figure \ref{fig:airfoil-maxIndSize-params} shows the total number of trainable parameters for each generation's DAE-LSTM of another experiment where regular DAE-GP was tested with dynamic adjustment of the number of hidden neurons as described in {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]} and {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]}.
+Again, the experiments was based on \(10\) individual runs and otherwise used the same hyperparameters as the previous one.
+It clearly shows that DAE-GP can strongly reduce the number of trainable parameters by adjusting the depth of it's hidden layers to the current generations population which will further reduce the amount of training necessary and likely reduce the run-time of DAE-GP.
+
+\begin{figure}[c]
+
+{\centering \includegraphics[width=0.8\linewidth]{./img/airfoil_2hl_maxIndSize_fullRun_30gens/trainableParams}
+
+}
+
+\caption{Median Number of trainable Parameters over 30 Generations - Airfoil (2HL)}\label{fig:airfoil-maxIndSize-params}
+\end{figure}
+
+As expected, this reduction of hidden neurons by dynamic adjustment, does also result in a reduction of the fitness of solutions found.
+Figure \ref{fig:airfoil-maxIndSize-fitness} shows that pre-trained DAE-GP, with it's more complex DAE-LSTMs over the full 30 generations, does achieve a higher median fitness than regular DAE-GP with dynamic adjustment of hidden neurons.
+Pre-Training does therefore introduce a new Trade-Off decision into the adjustment of DAE-GP hyperparameters where the number of hidden neurons needs to be carefully selected since it positivly impacts solution quality but also increases the computational cost of DAE-LSTM training.
+
+\begin{figure}[c]
+
+{\centering \includegraphics[width=0.8\linewidth]{./img/airfoil_2hl_maxIndSize_fullRun_30gens/median_fitness_byGens}
+
+}
+
+\caption{Median Best Fitness over 30 Generations - Airfoil (2HL) with dynamic adjustment of regular DAE-GP}\label{fig:airfoil-maxIndSize-fitness}
+\end{figure}
+
 \hypertarget{reduced-number-of-hidden-layers}{%
@@
 \end{figure}
+
+\ldots{} describe differences in distribution, interquartile range blablabla\ldots{}
 
@@
 
-The experiments conducted in this thesis did not find any significant benefits in using a pre-training strategy for using DAE-GP to solve real world symbolic regression problems.
+\hypertarget{benefits-of-using-pre-training}{%
+\subsection{Benefits of using Pre-Training}\label{benefits-of-using-pre-training}}
+
+The experiments conducted in the scope of this master thesis did find that the implemented pre-training strategy resulted in an increased population diversity measured by the normalized levenshtein edit distance.
+Increasing the population diversity can be an important mechanism in controlling the search behavior of population based metaheuristics and is especially useful for increasing the exploration of the search space of possible solutions !QUELLE!.
+
+Another significant benefit of using pre-training in DAE-GP is that it is highly efficient in decreasing the number of training epochs that are necessary to train each generations DAE-LSTM model.
+Even though pre-training increased the median total run-time for the tested benchmark problem in combination with the chosen DAE-LSTM dimensions, reducing the number of training epochs might be a possible way to lower the runtime of DAE-GP if further performance optimizations to the source code, especially during model sampling, can be achieved.
+
+\hypertarget{disadvantages-of-using-pre-training}{%
+\subsection{Disadvantages of using Pre-Training}\label{disadvantages-of-using-pre-training}}
+
+The experiments conducted in this thesis did not find any significant benefits on both solution quality as well as run-time for solving real world symbolic regression problems with pre-trained DAE-GP in comparison to regular DAE-GP.
+
+One major drawback of using pre-training with DAE-GP
 

### 2023-02-03 paper.Rmd
autogit
Fri Feb  3 12:00:08 CET 2023: update-remote.sh

@@
 EDAs use this probabilistic model to sample new offspring solutions inside an evolutionary search to replace crossover and/or mutation operators.
+
+### EDA-GP
+
+... insert some smart stuff here...
 
@@
 
-One major drawback of using pre-training with DAE-GP
+As detailed in the subsection [Dynamic Adjustment of Hidden Neurons], one major drawback of using pre-training with DAE-GP is the loss of the ability to dynamically adjust the number of hidden neurons to the maximum size of individuals inside the current generations population.
+This dynamic mechanism does reduce the computational resources necessary for training DAE-LSTM networks which is one of the most time consuming operations besides sampling new individuals.
+
+Another disadvantage of pre-training in DAE-GP is that in order for it to show
 
@@
 
-<!-- * der Erhöhung der Populationsdiversität (mögliche Strategie zur Kontrolle des Explore/Exploit Verhaltens von DAE-GP?) -->
-
 <!-- * der Reduktion der Trainingsepochen pro Generation (Reduktion der Rechenleistung, insbesondere bei komplexen DAE-LSTM nützlich?) -->
@@
 <!-- ## Nachteile -->
-
-<!-- * DAE-GP verliert durch Einsatz von Pre-Training die Möglichkeit die Anzahl künstlicher Neuronen pro verstecktem Layer anzupassen -->
 

### 2023-02-03 paper.tex
Fri Feb  3 12:00:08 CET 2023: update-remote.sh

@@
 
+\hypertarget{eda-gp}{%
+\subsubsection{EDA-GP}\label{eda-gp}}
+
+\ldots{} insert some smart stuff here\ldots{}
+
 \hypertarget{denoising-autoencoders}{%
@@
 
-One major drawback of using pre-training with DAE-GP
+As detailed in the subsection {[}\protect\hyperlink{ref-Dynamic}{\textbf{Dynamic?}} Adjustment of Hidden Neurons{]}, one major drawback of using pre-training with DAE-GP is the loss of the ability to dynamically adjust the number of hidden neurons to the maximum size of individuals inside the current generations population.
+This dynamic mechanism does reduce the computational resources necessary for training DAE-LSTM networks which is one of the most time consuming operations besides sampling new individuals.
+
+Another disadvantage
 

### 2023-02-03 paper.Rmd
Fri Feb  3 13:00:09 CET 2023: update-remote.sh

@@
 
-```{r airfoil-median-fitness-30gens, echo=FALSE, fig.cap="Best Fitness over 30 Generations - Airfoil (2HL)", out.width="90%"}
+```{r airfoil-median-fitness-30gens, echo=FALSE, fig.cap="Best Fitness over 30 Generations - Airfoil", out.width="90%"}
 knitr::include_graphics("./img/airfoil_2hl_150hn_fullRun_30gens_withGP/mean_median_fitness_byGens.png")
@@
 
-```{r airfoil-distribution-fitness-30gens, echo=FALSE, fig.cap="Best Fitness after 30 Generations - Airfoil (2HL)", out.width="80%"}
+```{r airfoil-distribution-fitness-30gens, echo=FALSE, fig.cap="Best Fitness after 30 Generations - Airfoil", out.width="80%"}
 knitr::include_graphics("./img/airfoil_2hl_150hn_fullRun_30gens/final_fit_boxplot.png")
@@
 
-```{r airfoil-size-combined-30gens, echo=FALSE, fig.cap="Median Solution Size over 30 Generations - Airfoil (2HL)", out.width="80%"}
+```{r airfoil-size-combined-30gens, echo=FALSE, fig.cap="Median Solution Size over 30 Generations - Airfoil", out.width="80%"}
 knitr::include_graphics("./img/airfoil_2hl_150hn_fullRun_30gens/size_combined_median.png")
@@
 
-```{r airfoil-popDiversity, echo=FALSE, fig.cap="Median Population Diversity over 30 Generations - Airfoil (2HL)", out.width="80%"}
+```{r airfoil-popDiversity, echo=FALSE, fig.cap="Median Population Diversity over 30 Generations - Airfoil", out.width="80%"}
 knitr::include_graphics("./img/airfoil_2hl_150hn_fullRun_30gens_withGP/median_popDiversity_byGens_levOnly.png")
@@
 
-```{r airfoil-maxIndSize-params, echo=FALSE, fig.cap="Median Number of trainable Parameters over 30 Generations - Airfoil (2HL)", out.width="80%"}
+```{r airfoil-maxIndSize-params, echo=FALSE, fig.cap="Median Number of trainable Parameters over 30 Generations - Airfoil - Dynamic adjustment of regular DAE-GP", out.width="80%"}
 knitr::include_graphics("./img/airfoil_2hl_maxIndSize_fullRun_30gens/trainableParams.png")
@@
 
-```{r airfoil-maxIndSize-fitness, echo=FALSE, fig.cap="Median Best Fitness over 30 Generations - Airfoil (2HL) with dynamic adjustment of regular DAE-GP", out.width="80%"}
+```{r airfoil-maxIndSize-fitness, echo=FALSE, fig.cap="Median Best Fitness over 30 Generations - Airfoil - Dynamic adjustment of regular DAE-GP", out.width="80%"}
 knitr::include_graphics("./img/airfoil_2hl_maxIndSize_fullRun_30gens/median_fitness_byGens.png")
@@
 
-```{r airfoil-fitness-perGen, echo=FALSE, fig.cap="Fitness over 30 Generations - Airfoil (1HL)", out.width="80%"}
+```{r airfoil-fitness-perGen, echo=FALSE, fig.cap="Fitness over 30 Generations - Airfoil - Single Hidden Layer", out.width="80%"}
 knitr::include_graphics("./img/airfoil_1hl_150hn_30gens/mean_median_fitness_byGens.png")
@@
 
-```{r airfoil-finalFitness, echo=FALSE, fig.cap="Fitness after 30 Generations - Airfoil (1HL)", out.width="80%"}
+```{r airfoil-finalFitness, echo=FALSE, fig.cap="Fitness after 30 Generations - Airfoil - Single Hidden Layer", out.width="80%"}
 knitr::include_graphics("./img/airfoil_1hl_150hn_30gens/final_fit_boxplot.png")
@@
 
-```{r altStrategies-Fitness, echo=FALSE, fig.cap="Best Fitness over 30 Generations - Alternative Pre-Strategies - Airfoil", out.width="80%"}
+```{r altStrategies-Fitness, echo=FALSE, fig.cap="Best Fitness over 30 Generations - Airfoil - Alternative Pre-Strategies ", out.width="80%"}
 knitr::include_graphics("./img/PT_Strategies_airfoil_2hl_150hn_fullRun/median_fitness_byGens.png")
@@
 
-```{r altStrategies-FitnessDistr, echo=FALSE, fig.cap="Best Fitness on Test Set after 30 Generations - Alternative Pre-Trategies - Airfoil", out.width="80%"}
+```{r altStrategies-FitnessDistr, echo=FALSE, fig.cap="Best Fitness on Test Set after 30 Generations - Airfoil - Alternative Pre-Trategies", out.width="80%"}
 knitr::include_graphics("./img/PT_Strategies_airfoil_2hl_150hn_fullRun/final_fit_boxplot(test_only).png")
@@
 
-## Effect on runtime
+## Effect on Run-Time
 
@@
 ```
-
-<!-- Regarding the hidden neurons, the dimension of each hidden layer, I used two different strategies: Regular DAE-GP uses the same strategy as described by earlier work (see [@dae-gp_2022_symreg] or [@dae-gp_2020_rtree]), the number of hidden neurons is dynamically set at each generation to the maximum length of all solutions inside the current training population.
-For pre-trained DAE-GP another strategy had to be used, I used a fixed number of hidden neurons for each run that is set equal to the maximum length of all individuals that are present in the initial pre-training population.
-I decided to take this approach instead of using a fixed number of hidden neurons for the fact that it simplifies the adjustment of the parameter for later experiments on different problems.
-It should nonetheless be mentioned that this strategy does give pre-training on average more hidden neurons per layer resulting in a more complex network (since the pre-training population is much larger than the regular population, the chances for larger individuals inside it are also higher).
--->
 
@@
 Another significant benefit of using pre-training in DAE-GP is that it is highly efficient in decreasing the number of training epochs that are necessary to train each generations DAE-LSTM model.
-Even though pre-training increased the median total run-time for the tested benchmark problem in combination with the chosen DAE-LSTM dimensions, reducing the number of training epochs might be a possible way to lower the runtime of DAE-GP if further performance optimizations to the source code, especially during model sampling, can be achieved.
+Even though pre-training increased the median total run-time for the tested benchmark problem in combination with the chosen DAE-LSTM dimensions, reducing the number of training epochs might be a possible way to lower the run time of DAE-GP if further performance optimizations to the source code, especially during model sampling, can be achieved.
 
@@
 
-Another disadvantage of pre-training in DAE-GP is that in order for it to show
+Another disadvantage of using pre-training in DAE-GP is that, in my experiments, it required DAE-LSTM networks to use two hidden layers (see subsection [Reduced Number of Hidden Layers]).
+For a single hidden layer, pre-training did show a negative impact on the median final fitness of solutions found as it was suggested by [@pmlr-v9-erhan10a].
+Since current research on DAE-GP for real-world symbolic regression relies on a single hidden layer (see [@dae-gp_2022_symreg]), pre-training does come with additional computational expenses for using deeper DAE-LSTMs.
 
-<!-- ## Vorteile -->
+## Limitations and open Questions
 
-<!-- Der Einsatz von Pre-Training dient: -->
-
-<!-- * der Reduktion der Trainingsepochen pro Generation (Reduktion der Rechenleistung, insbesondere bei komplexen DAE-LSTM nützlich?) -->
-
-<!-- ## Nachteile -->
-
-<!-- * Erhöhung der Rechenzeit/Ressourcen: DAE-LSTMs müssen ausreichend dimensioniert sein (2+ Hidden Layer), aktuelle Publikationen nutzen nur einen einzelnen Hidden Layer^[@dae-gp_2022_symreg] ^[@daegp_explore_exploit].
--->
-
-<!-- * Keine signifikanten Vorteile für die Qualität der gefundenen Lösungen -->
-
-# Limitations and open Questions
-
-## Further Questions
+One major limitation for the experiments conducted during the reseach phase of this thesis has been the large computational effort for experimenting with DAE-GP.
+Since single runs for DAE-GP have a run-time of many hours, it was not possible for me to extend the experiments to larger population sizes or more generations.
+Also I did not perform an extended optimization of hyperparameters (e.g by using a grid search strategy) which might yield configurations that increase the benefits of pre-training.
+As described in subsection [Effect on Run-Time], sampling new solutions has been a large contributor to the large performance hit on run-time by using pre-training.
+It could be a worthwhile endevour to explore new strategies, that
 

### 2023-02-03 paper.tex
Fri Feb  3 13:00:09 CET 2023: update-remote.sh

--- 2023-02-03 paper.tex
@@
 
-Denoising Autoencoder Genetic Programming (DAE-GP) is a novel variation of an genetic programming based Estimation of Distribution Algorithm (EDA-GP) that uses a denoising autoencoders long short-term memory network (DAE-LSTMs) as a probabilistic model to sample new candidate solutions {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]}.
+Denoising Autoencoder Genetic Programming (DAE-GP) is a novel variation of an genetic programming based Estimation of Distribution Algorithm (EDA-GP) that uses a denoising autoencoders long short-term memory network (DAE-LSTMs) as a probabilistic model to sample new candidate solutions {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{19}{]}.
 
@@
 In DAE-GP, DAE-LSTMs are trained with a subset of high-fitness solutions selected from a parent population with the aim to capture their promising qualities.
-The resulting model is then used to sample new offspring solutions by propagating partially mutated solutions from the parent population through the DAE-LSTMS {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]}.
-In previous work DAE-GP has been shown to outperform GP for both a generalized version of the royal tree problem {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]} as well as for a real-world symbolic regression problem {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]}.
-
-The DAE-GP algorithm first described by {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]} trains a DAE-LSTMs for each generation \(g\) of the search from scratch.
-{[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]} and {[}\protect\hyperlink{ref-daegp_explore_exploit}{16}{]} suggests the incorporation of a pre-training strategy into the evolutionary search as a possible way of improving the overall performance of the DAE-GP algorithm.
+The resulting model is then used to sample new offspring solutions by propagating partially mutated solutions from the parent population through the DAE-LSTMS {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{19}{]}.
+In previous work DAE-GP has been shown to outperform GP for both a generalized version of the royal tree problem {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{19}{]} as well as for a real-world symbolic regression problem {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{18}{]}.
+
+The DAE-GP algorithm first described by {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{19}{]} trains a DAE-LSTMs for each generation \(g\) of the search from scratch.
+{[}\protect\hyperlink{ref-dae-gp_2022_symreg}{18}{]} and {[}\protect\hyperlink{ref-daegp_explore_exploit}{17}{]} suggests the incorporation of a pre-training strategy into the evolutionary search as a possible way of improving the overall performance of the DAE-GP algorithm.
 The key idea is to pre-train an initial DAE-LSTM on a large population of candidate solutions and to use the pre-trained parameters of this initial model in each generation of the search as starting parameters for the current generation DAE-LSTM.
@@
 
-DAE-GP is an EDA-GP algorithm that uses DAE-LSTM networks as a probabilistic model to sample new offspring solutions {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]}.
+DAE-GP is an EDA-GP algorithm that uses DAE-LSTM networks as a probabilistic model to sample new offspring solutions {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{19}{]}.
 
@@
 Another important concept in EC is the distinction between genotypes and phenotypes of solutions, the genotype contains the information that is necessary to construct the phenotype, the outer appearance of a particular solution on which we measure the overall quality of solutions.
-The representation of a solution is therefore defined by the mapping of genotypes to phenotypes {[}\protect\hyperlink{ref-design_of_modern_heuristics}{12}{]}.
+The representation of a solution is therefore defined by the mapping of genotypes to phenotypes {[}\protect\hyperlink{ref-design_of_modern_heuristics}{13}{]}.
 Genetic operators, such as mutation or recombination are usually applied to the genotype of solutions.
@@
 
-GP follows the same basic evolutionary principle of EC but searches for more general, hierarchical computer programs of dynamically varying size and shape {[}\protect\hyperlink{ref-Koza1993GeneticP}{8}{]}.
-The computer programs that are at the center of the evolutionary search in GP are commonly represented by tree structures at the level of their phenotype {[}\protect\hyperlink{ref-design_of_modern_heuristics}{12}{]}.
-Since GP searches for high fitness computer programs that produce a desired output for some input, it can be applied to various different problem domains such as symbolic regression, automatic programming, or evolving game-playing strategies {[}\protect\hyperlink{ref-Koza1993GeneticP}{8}{]}.
+GP follows the same basic evolutionary principle of EC but searches for more general, hierarchical computer programs of dynamically varying size and shape {[}\protect\hyperlink{ref-Koza1993GeneticP}{9}{]}.
+The computer programs that are at the center of the evolutionary search in GP are commonly represented by tree structures at the level of their phenotype {[}\protect\hyperlink{ref-design_of_modern_heuristics}{13}{]}.
+Since GP searches for high fitness computer programs that produce a desired output for some input, it can be applied to various different problem domains such as symbolic regression, automatic programming, or evolving game-playing strategies {[}\protect\hyperlink{ref-Koza1993GeneticP}{9}{]}.
 An important quality of GP is the ability to search for solutions of variable length and structure.
 GP is an especially useful meta heuristic for problems where no a priori knowledge about the final form of good solutions is available.
-GPs ability to optimize solutions for their structure as well as for their parameters led to it being one of the most prevalent methods used for symbolic regression {[}\protect\hyperlink{ref-10.1007ux2f978-3-540-24621-3_22}{10}{]}.
+GPs ability to optimize solutions for their structure as well as for their parameters led to it being one of the most prevalent methods used for symbolic regression {[}\protect\hyperlink{ref-10.1007ux2f978-3-540-24621-3_22}{11}{]}.
 
@@
 
-The aim of Estimation of distribution algorithms (EDA) is to replace the standard variation operators used in GA by building probabilistic models that can capture complex dependencies between different decision variables of an optimization problem {[}\protect\hyperlink{ref-design_of_modern_heuristics}{12}{]}.
+The aim of Estimation of distribution algorithms (EDA) is to replace the standard variation operators used in GA by building probabilistic models that can capture complex dependencies between different decision variables of an optimization problem {[}\protect\hyperlink{ref-design_of_modern_heuristics}{13}{]}.
 EDAs use this probabilistic model to sample new offspring solutions inside an evolutionary search to replace crossover and/or mutation operators.
+
+\hypertarget{eda-gp}{%
+\subsubsection{EDA-GP}\label{eda-gp}}
+
+\ldots{} insert some smart stuff here\ldots{}
 
@@
 
-One possible way of model building in EDA proposed by {[}\protect\hyperlink{ref-harmless_overfitting_eda}{11}{]} is to use denoising autoencoders (DAEs) as generative models to capture complex probability distributions of decision variables.
+One possible way of model building in EDA proposed by {[}\protect\hyperlink{ref-harmless_overfitting_eda}{12}{]} is to use denoising autoencoders (DAEs) as generative models to capture complex probability distributions of decision variables.
 DAE, a variation of the autoencoders (AE), are a widely used type of neural networks in the domain of unsupervised machine learning that maps \(n\) input variables to \(n\) output variables using a hidden representation.
 
-AE were introduced by {[}\protect\hyperlink{ref-ae_orig}{6}{]} to compress high-dimensional data into lower-dimensions. An AE consists of two different sub-units:
+AE were introduced by {[}\protect\hyperlink{ref-ae_orig}{7}{]} to compress high-dimensional data into lower-dimensions. An AE consists of two different sub-units:
 
@@
 
-The AE is trained to reduce the reconstruction error between input and output data, after the training procedure is finished the network is able to reduce the dimensional of input data to a compressed representation{[}\protect\hyperlink{ref-ae_orig}{6}{]}.
-
-DAE was first introduced by {[}\protect\hyperlink{ref-dae_orig2008}{14}{]} as an improved AE with the ability to learn new representations of data that is especially robust to partially corrupted input data.
+The AE is trained to reduce the reconstruction error between input and output data, after the training procedure is finished the network is able to reduce the dimensional of input data to a compressed representation{[}\protect\hyperlink{ref-ae_orig}{7}{]}.
+
+DAE was first introduced by {[}\protect\hyperlink{ref-dae_orig2008}{15}{]} as an improved AE with the ability to learn new representations of data that is especially robust to partially corrupted input data.
 DAE modifies the AE by using partially corrupted input data for the AE and training it to reconstruct the uncorrupted, original version of the input data.
 
-Since the hidden representation of DAE captures the dependency structure of the input data it can therefore also be used to generate new solutions in the context of GAs {[}\protect\hyperlink{ref-harmless_overfitting_eda}{11}{]}.
-
-DAE-GP builds upon the concept of using DAEs in EDAs described by {[}\protect\hyperlink{ref-harmless_overfitting_eda}{11}{]} and transfers the concept to the domain of GP.
-The mutation and crossover operators of standard GP are replaced by sampling new solutions from a probabilistic model that is build by training a DAEs long short-term memory (LSTM) network on a subset of high fitness solutions from the current population {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]}.
-
-LSTMs are a variant of artificial neural networks first introduced by {[}\protect\hyperlink{ref-lstm_orig}{7}{]} that can store learned information over en extended time period while avoiding the problem of vanishing and/or exploding gradients.
-Since DAE-GP encodes candidate solutions as linear strings in prefix notation, the DAE in DAE-GP uses LSTMs for both encoding and decoding where the total amount of time steps \(T\) is equal to sum of the length of the input solution and the output solution {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]}.
+Since the hidden representation of DAE captures the dependency structure of the input data it can therefore also be used to generate new solutions in the context of GAs {[}\protect\hyperlink{ref-harmless_overfitting_eda}{12}{]}.
+
+DAE-GP builds upon the concept of using DAEs in EDAs described by {[}\protect\hyperlink{ref-harmless_overfitting_eda}{12}{]} and transfers the concept to the domain of GP.
+The mutation and crossover operators of standard GP are replaced by sampling new solutions from a probabilistic model that is build by training a DAEs long short-term memory (LSTM) network on a subset of high fitness solutions from the current population {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{19}{]}.
+
+LSTMs are a variant of artificial neural networks first introduced by {[}\protect\hyperlink{ref-lstm_orig}{8}{]} that can store learned information over en extended time period while avoiding the problem of vanishing and/or exploding gradients.
+Since DAE-GP encodes candidate solutions as linear strings in prefix notation, the DAE in DAE-GP uses LSTMs for both encoding and decoding where the total amount of time steps \(T\) is equal to sum of the length of the input solution and the output solution {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{19}{]}.
 
@@
 Pre-Training describes the concept of initially training an artificial neural network on a large dataset before using it to solve a more specific task.
-It is a commonly used strategy in deep architectures that has been shown to improve both the optimization process itself as well as the generalization behavior if compared to the standard approach of using randomly initialized parameters {[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{3}{]}.
-
-The strategy of pre-training artificial neural networks is based on the idea of transfer learning that aims at retaining previously learned knowledge from one task to re-use it for another task {[}\protect\hyperlink{ref-HAN2021225}{5}{]} {[}\protect\hyperlink{ref-survey_transfer_learning}{9}{]}.
+It is a commonly used strategy in deep architectures that has been shown to improve both the optimization process itself as well as the generalization behavior if compared to the standard approach of using randomly initialized parameters {[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{4}{]}.
+
+The strategy of pre-training artificial neural networks is based on the idea of transfer learning that aims at retaining previously learned knowledge from one task to re-use it for another task {[}\protect\hyperlink{ref-HAN2021225}{6}{]} {[}\protect\hyperlink{ref-survey_transfer_learning}{10}{]}.
 Transfer learning traditionally follows a two phase approach:
@@
 
-where source task and target task are usually similar but may differ in their feature space and the distribution of training data {[}\protect\hyperlink{ref-survey_transfer_learning}{9}{]}.
+where source task and target task are usually similar but may differ in their feature space and the distribution of training data {[}\protect\hyperlink{ref-survey_transfer_learning}{10}{]}.
 
@@
 Instead of initializing and optimizing each model from scratch as done in previous work (e.g.
-{[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]} {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]}), a separate DAE-LSTM network \(\hat{M}\) will be trained on a large population of randomly initialized solutions.\\
+{[}\protect\hyperlink{ref-dae-gp_2020_rtree}{19}{]} {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{18}{]}), a separate DAE-LSTM network \(\hat{M}\) will be trained on a large population of randomly initialized solutions.\\
 The trainable parameters \(\theta_{\hat{M}}\) obtained after finishing the training procedure of \(\hat{M}\) are then used as the starting parameters for each following DAE-LSTM model \(M_g\) for \(g\in{1,...,g_{max}}\).
@@
 
-Early research on pre-training for DAE by {[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{3}{]} comes to the conclusion that besides adding robustness to models the strategy also results in improved generalization and better performing models.
+Early research on pre-training for DAE by {[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{4}{]} comes to the conclusion that besides adding robustness to models the strategy also results in improved generalization and better performing models.
 The authors describe that unsupervised pre-training behaves like a regularizer on DAE networks, the mean test error and it's variance are reduced for DAE networks that are initialized from pre-trained weights if compared to the same architectures that use random initialization.
 
-Another important finding by {[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{3}{]} is that the positive effect of pre-training is dependent on both the depth of the network as well as the size of layers - while increasingly larger networks benefit increasingly more from pre-training the final performance of small architectures tends to be worse with pre-training than with randomized weight initialization.
+Another important finding by {[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{4}{]} is that the positive effect of pre-training is dependent on both the depth of the network as well as the size of layers - while increasingly larger networks benefit increasingly more from pre-training the final performance of small architectures tends to be worse with pre-training than with randomized weight initialization.
 The authors find evidence that pre-training DAE is especially useful for optimizing the parameters of the lower layers of the network.
@@
 Another motivation for introducing pre-training into DAE-GP is the prevalence of sampling error for small GP populations sizes.
-{[}\protect\hyperlink{ref-sampling_err_gp}{13}{]} finds that sampling error, non-systematic errors that are caused by observing only a small subset of the statistical population, is a severe problem in the domain of GP since the initial population may not be a representative sample of all possible solutions.
-{[}\protect\hyperlink{ref-sampling_err_gp}{13}{]} also introduces a method for calculating optimal population sizes to minimize the presence of sampling error.
+{[}\protect\hyperlink{ref-sampling_err_gp}{14}{]} finds that sampling error, non-systematic errors that are caused by observing only a small subset of the statistical population, is a severe problem in the domain of GP since the initial population may not be a representative sample of all possible solutions.
+{[}\protect\hyperlink{ref-sampling_err_gp}{14}{]} also introduces a method for calculating optimal population sizes to minimize the presence of sampling error.
 Since DAE-LSTM of DAE-GP learns the properties of it's training population and reuses the acquired knowledge in the sampling procedure, using the parameters obtained from training \(\hat{M}\) on a sufficiently large training population might reduce the need for large population sizes in the following generations of the search if \(\hat{M}\) already implicitly captured the properties of a representative sample of solutions.
@@
 
-The DAE-GP algorithm, first described by {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]}, is summarized as a flowchart in figure \ref{fig:daegp-flowchart}.
+The DAE-GP algorithm, first described by {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{19}{]}, is summarized as a flowchart in figure \ref{fig:daegp-flowchart}.
 After setting the generation counter \(g\) to \(0\) and creating the initial population of solutions \(P_0\), DAE-GP enters the main loop that checks if a termination criteria is satisfied (i.e.~maximum number of generations or fitness evaluations is reached).
@@
 One difficulty in the implementation of a pre-training strategy into DAE-GP has been the determination of the number of hidden neurons in side the DAE-LSTMs hidden layers.
-The original description of DAE-GP {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]} uses a strategy where the number of hidden neurons for each hidden layer is dynamically set per generation to the maximum individual size inside the current population.
+The original description of DAE-GP {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{19}{]} uses a strategy where the number of hidden neurons for each hidden layer is dynamically set per generation to the maximum individual size inside the current population.
 For a pre-training implementation, this strategy can not be easily adapted since it leads to a changing number of neurons at each generation resulting in different dimensions of the DAE-LSTM.
@@
 \hline
+Hidden\_Layer & 2\\
+\hline
 Selection\_Operator & Binary Tournament Selection\\
@@
 
-The framework that was used to conduct all experiments of this thesis was provided by the supervisor of this thesis David Wittenberg and uses the python programming language in conjunction with the baseline libraries \texttt{Keras} {[}\protect\hyperlink{ref-chollet2015keras}{1}{]} for deep neural networks and \texttt{deap} {[}\protect\hyperlink{ref-DEAP_JMLR2012}{4}{]} for evolutionary computation.
+The framework that was used to conduct all experiments of this thesis was provided by the supervisor of this thesis David Wittenberg and uses the python programming language in conjunction with the baseline libraries \texttt{Keras} {[}\protect\hyperlink{ref-chollet2015keras}{1}{]} for deep neural networks and \texttt{deap} {[}\protect\hyperlink{ref-DEAP_JMLR2012}{5}{]} for evolutionary computation.
 
@@
 To test pre-training in DAE-GP this thesis focuses on the domain of real-world symbolic regression problems.
-Symbolic Regression problems have been one of the first GP applications {[}\protect\hyperlink{ref-Koza1993GeneticP}{8}{]} and are an actively studied and highly relevant research area.
-The goal in symbolic regression is to find a mathematical model for a given set of data points {[}\protect\hyperlink{ref-10.1007ux2f978-3-540-24621-3_22}{10}{]}, in real-world symbolic regression these data points are sourced from real-world observations which in contrast to synthetic symbolic regression problem are more likely to contain random noise and bias.
+Symbolic Regression problems have been one of the first GP applications {[}\protect\hyperlink{ref-Koza1993GeneticP}{9}{]} and are an actively studied and highly relevant research area.
+The goal in symbolic regression is to find a mathematical model for a given set of data points {[}\protect\hyperlink{ref-10.1007ux2f978-3-540-24621-3_22}{11}{]}, in real-world symbolic regression these data points are sourced from real-world observations which in contrast to synthetic symbolic regression problem are more likely to contain random noise and bias.
 Another important challenge in solving real-world symbolic regression problems is the ability for a given model to generalize, we want the final model to show high accuracy in predicting outcomes for previously unseen cases.
@@
 The objective of the airfoil problem is to find a function that accurately predicts the output variable \(y\) by taking in a subset of the input variables \(x1,x2,x3,x4,x5\).
-The function set used by all DAE-GP variations for the airfoil problem is summarized in \ref{tab:airfoil-fullRun-2hl-maxIndSize-params}, the terminal set consists of the 5 input variables \(x1,x2,x3,x4,x5\) and the ephemeral integer constants listed in \ref{tab:airfoil-fullRun-2hl-maxIndSize-params}{[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]}.
+The function set used by all DAE-GP variations for the airfoil problem is summarized in \ref{tab:airfoil-fullRun-2hl-maxIndSize-params}, the terminal set consists of the 5 input variables \(x1,x2,x3,x4,x5\) and the ephemeral integer constants listed in \ref{tab:airfoil-fullRun-2hl-maxIndSize-params}{[}\protect\hyperlink{ref-dae-gp_2022_symreg}{18}{]}.
 
@@
 For each configuration I tested traditional DAE-GP as well as a pre-trained DAE-GP resulting in 12 total sub experiments that were each based on 10 individual runs (total number of runs=\(120\)).
-To avoid creating biased results through the presence of sampling error inside the pre-training population (see {[}\protect\hyperlink{ref-sampling_err_gp}{13}{]}), the population size for the pre-training phase is chosen very high with the size of \(\hat{P}\) =10000 where 50\% of \(\hat{P}\) is used for the training population \(\hat{P}_{train}\) and the remaining 50\% are used for the hold-out validation population \(\hat{P}_{test}\).
+To avoid creating biased results through the presence of sampling error inside the pre-training population (see {[}\protect\hyperlink{ref-sampling_err_gp}{14}{]}), the population size for the pre-training phase is chosen very high with the size of \(\hat{P}\) =10000 where 50\% of \(\hat{P}\) is used for the training population \(\hat{P}_{train}\) and the remaining 50\% are used for the hold-out validation population \(\hat{P}_{test}\).
 The training of DAE-LSTM uses a fixed number of 1000 epochs until termination and uses Adam optimization for gradient descent.
@@
 In general I expect that with a growing number of trainable parameters (either by adding more hidden layers or more hidden neurons per layer) the DAE-LSTM will be more prone to overfit to the training population \(P_{train}\) resulting in a small reconstruction error for the training population and a large one for the validation population \(P_{test}\).
-The reason for this effect is that a larger network trained over an extended period of time (without the use of strategies like early stopping), has much more potential to learn noise from the training dataset than a smaller network, which is more likely to result in worsening performance for previously unseen cases {[}\protect\hyperlink{ref-weigend1994overfitting}{15}{]}.
-
-Based on the review of {[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{3}{]} I also expect that pre-training will have an insignificant or even negative influence on small DAE-LSTM instances while improving the networks generalization ability with growing size.
+The reason for this effect is that a larger network trained over an extended period of time (without the use of strategies like early stopping), has much more potential to learn noise from the training dataset than a smaller network, which is more likely to result in worsening performance for previously unseen cases {[}\protect\hyperlink{ref-weigend1994overfitting}{16}{]}.
+
+Based on the review of {[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{4}{]} I also expect that pre-training will have an insignificant or even negative influence on small DAE-LSTM instances while improving the networks generalization ability with growing size.
 
@@
 
-\caption{Best Fitness over 30 Generations - Airfoil (2HL)}\label{fig:airfoil-median-fitness-30gens}
+\caption{Best Fitness over 30 Generations - Airfoil}\label{fig:airfoil-median-fitness-30gens}
 \end{figure}
@@
 
-\caption{Best Fitness after 30 Generations - Airfoil (2HL)}\label{fig:airfoil-distribution-fitness-30gens}
+\caption{Best Fitness after 30 Generations - Airfoil}\label{fig:airfoil-distribution-fitness-30gens}
 \end{figure}
@@
 
-Since previous work by {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]} demonstrated that DAE-GP, for a given number of 10.000 fitness evaluations, is able to produce solutions that are much smaller in tree size than regular GP, I also studied the influence of pre-training on the size of solutions found by DAE-GP.
+Since previous work by {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{18}{]} demonstrated that DAE-GP, for a given number of 10.000 fitness evaluations, is able to produce solutions that are much smaller in tree size than regular GP, I also studied the influence of pre-training on the size of solutions found by DAE-GP.
 Figure \ref{fig:airfoil-size-combined-30gens} shows the median of both the average size of individuals inside the population as well as the size of the currently best performing solution inside the population.
@@
 
-\caption{Median Solution Size over 30 Generations - Airfoil (2HL)}\label{fig:airfoil-size-combined-30gens}
+\caption{Median Solution Size over 30 Generations - Airfoil}\label{fig:airfoil-size-combined-30gens}
 \end{figure}
@@
 
-\caption{Median Population Diversity over 30 Generations - Airfoil (2HL)}\label{fig:airfoil-popDiversity}
+\caption{Median Population Diversity over 30 Generations - Airfoil}\label{fig:airfoil-popDiversity}
 \end{figure}
@@
 To ensure a fair comparison between pre-trained and regular DAE-GP, the prior experiments used a static number of hidden neurons for each hidden Layer inside the DAE-LSTM networks.
-For comparison, figure \ref{fig:airfoil-maxIndSize-params} shows the total number of trainable parameters for each generation's DAE-LSTM of another experiment where regular DAE-GP was tested with dynamic adjustment of the number of hidden neurons as described in {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]} and {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]}.
+For comparison, figure \ref{fig:airfoil-maxIndSize-params} shows the total number of trainable parameters for each generation's DAE-LSTM of another experiment where regular DAE-GP was tested with dynamic adjustment of the number of hidden neurons as described in {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{19}{]} and {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{18}{]}.
 Again, the experiments was based on \(10\) individual runs and otherwise used the same hyperparameters as the previous one.
@@
 
-\caption{Median Number of trainable Parameters over 30 Generations - Airfoil (2HL)}\label{fig:airfoil-maxIndSize-params}
+\caption{Median Number of trainable Parameters over 30 Generations - Airfoil - Dynamic adjustment of regular DAE-GP}\label{fig:airfoil-maxIndSize-params}
 \end{figure}
@@
 
-\caption{Median Best Fitness over 30 Generations - Airfoil (2HL) with dynamic adjustment of regular DAE-GP}\label{fig:airfoil-maxIndSize-fitness}
+\caption{Median Best Fitness over 30 Generations - Airfoil - Dynamic adjustment of regular DAE-GP}\label{fig:airfoil-maxIndSize-fitness}
 \end{figure}
@@
 
-Since previous research on DAE-GP used a single hidden layers for the airfoil dataset {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]}, I repeated the experiment with a single hidden layer.
-Based on the findings of {[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{3}{]}, I expected that pre-training DAE-GP would lead to a decrease in overall performance.
+Since previous research on DAE-GP used a single hidden layers for the airfoil dataset {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{18}{]}, I repeated the experiment with a single hidden layer.
+Based on the findings of {[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{4}{]}, I expected that pre-training DAE-GP would lead to a decrease in overall performance.
 Figures \ref{fig:airfoil-fitness-perGen} and \ref{fig:airfoil-finalFitness} shows the mean and median best fitness per generation as well as the distribution of final fitness scores after 30 generations.
@@
 
-\caption{Fitness over 30 Generations - Airfoil (1HL)}\label{fig:airfoil-fitness-perGen}
+\caption{Fitness over 30 Generations - Airfoil - Single Hidden Layer}\label{fig:airfoil-fitness-perGen}
 \end{figure}
@@
 
-\caption{Fitness after 30 Generations - Airfoil (1HL)}\label{fig:airfoil-finalFitness}
+\caption{Fitness after 30 Generations - Airfoil - Single Hidden Layer}\label{fig:airfoil-finalFitness}
 \end{figure}
@@
 
-The pre-training implementation in previous experiments used ramped half and half initialization {[}\protect\hyperlink{ref-Koza1993GeneticP}{8}{]} to create \(\hat{P}\), this leads to 50\% of the population consisting of full grown trees.
+The pre-training implementation in previous experiments used ramped half and half initialization {[}\protect\hyperlink{ref-Koza1993GeneticP}{9}{]} to create \(\hat{P}\), this leads to 50\% of the population consisting of full grown trees.
 An alternative method of initializing \(\hat{P}\) is to use the grow initialization method where solutions are not guaranteed to be full trees.
@@
 
-\caption{Best Fitness over 30 Generations - Alternative Pre-Strategies - Airfoil}\label{fig:altStrategies-Fitness}
+\caption{Best Fitness over 30 Generations - Airfoil - Alternative Pre-Strategies }\label{fig:altStrategies-Fitness}
 \end{figure}
@@
 
-\caption{Best Fitness on Test Set after 30 Generations - Alternative Pre-Trategies - Airfoil}\label{fig:altStrategies-FitnessDistr}
+\caption{Best Fitness on Test Set after 30 Generations - Airfoil - Alternative Pre-Trategies}\label{fig:altStrategies-FitnessDistr}
 \end{figure}
@@
 
-One major drawback of using pre-training with DAE-GP
-
-\hypertarget{limitations-and-open-questions}{%
-\section{Limitations and open Questions}\label{limitations-and-open-questions}}
+As detailed in the subsection \protect\hyperlink{dynamic-adjustment-of-hidden-neurons}{Dynamic Adjustment of Hidden Neurons}, one major drawback of using pre-training with DAE-GP is the loss of the ability to dynamically adjust the number of hidden neurons to the maximum size of individuals inside the current generations population.
+This dynamic mechanism does reduce the computational resources necessary for training DAE-LSTM networks which is one of the most time consuming operations besides sampling new individuals.
+
+Another disadvantage of using pre-training in DAE-GP is that, in my experiments, it required DAE-LSTM networks to use two hidden layers (see subsection \protect\hyperlink{reduced-number-of-hidden-layers}{Reduced Number of Hidden Layers}).
+For a single hidden layer, pre-training did show a negative impact on the median final fitness of solutions found as it was suggested by {[}\protect\hyperlink{ref-pmlr-v9-erhan10a}{3}{]}.
+Since current research on DAE-GP for real-world symbolic regression relies on a single hidden layer (see {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{18}{]}), pre-training does come with additional computational expenses for using deeper DAE-LSTMs.
+
+\hypertarget{limitations}{%
+\subsection{Limitations}\label{limitations}}
+
+\ldots{}
 
@@
 
-{[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{3}{]}: Pre-Training effect especially useful for lower-level layers -\textgreater{} Only adapt weights from low layers or embed the pre-trained model?
+{[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{4}{]}: Pre-Training effect especially useful for lower-level layers -\textgreater{} Only adapt weights from low layers or embed the pre-trained model?
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-pmlr-v9-erhan10a}{}}%
+\CSLLeftMargin{{[}3{]} }%
+\CSLRightInline{Dumitru Erhan, Aaron Courville, Yoshua Bengio, and Pascal Vincent.
+2010.
+Why does unsupervised pre-training help deep learning?
+In \emph{Proceedings of the thirteenth international conference on artificial intelligence and statistics} (Proceedings of machine learning research), PMLR, Chia Laguna Resort, Sardinia, Italy, 201--208.
+Retrieved from \url{https://proceedings.mlr.press/v9/erhan10a.html}}
+
 \leavevmode\vadjust pre{\hypertarget{ref-pmlr-v5-erhan09a}{}}%
-\CSLLeftMargin{{[}3{]} }%
+\CSLLeftMargin{{[}4{]} }%
 \CSLRightInline{Dumitru Erhan, Pierre-Antoine Manzagol, Yoshua Bengio, Samy Bengio, and Pascal Vincent.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-DEAP_JMLR2012}{}}%
-\CSLLeftMargin{{[}4{]} }%
+\CSLLeftMargin{{[}5{]} }%
 \CSLRightInline{Félix-Antoine Fortin, François-Michel De Rainville, Marc-André Gardner, Marc Parizeau, and Christian Gagné.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-HAN2021225}{}}%
-\CSLLeftMargin{{[}5{]} }%
+\CSLLeftMargin{{[}6{]} }%
 \CSLRightInline{Xu Han, Zhengyan Zhang, Ning Ding, Yuxian Gu, Xiao Liu, Yuqi Huo, Jiezhong Qiu, Yuan Yao, Ao Zhang, Liang Zhang, Wentao Han, Minlie Huang, Qin Jin, Yanyan Lan, Yang Liu, Zhiyuan Liu, Zhiwu Lu, Xipeng Qiu, Ruihua Song, Jie Tang, Ji-Rong Wen, Jinhui Yuan, Wayne Xin Zhao, and Jun Zhu.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ae_orig}{}}%
-\CSLLeftMargin{{[}6{]} }%
+\CSLLeftMargin{{[}7{]} }%
 \CSLRightInline{G.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-lstm_orig}{}}%
-\CSLLeftMargin{{[}7{]} }%
+\CSLLeftMargin{{[}8{]} }%
 \CSLRightInline{Sepp Hochreiter and Jürgen Schmidhuber.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Koza1993GeneticP}{}}%
-\CSLLeftMargin{{[}8{]} }%
+\CSLLeftMargin{{[}9{]} }%
 \CSLRightInline{John R. Koza. 1993. Genetic programming - on the programming of computers by means of natural selection. In \emph{Complex adaptive systems}.}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-survey_transfer_learning}{}}%
-\CSLLeftMargin{{[}9{]} }%
+\CSLLeftMargin{{[}10{]} }%
 \CSLRightInline{Sinno Jialin Pan and Qiang Yang.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-10.1007ux2f978-3-540-24621-3_22}{}}%
-\CSLLeftMargin{{[}10{]} }%
+\CSLLeftMargin{{[}11{]} }%
 \CSLRightInline{Grégory Paris, Denis Robilliard, and Cyril Fonlupt.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-harmless_overfitting_eda}{}}%
-\CSLLeftMargin{{[}11{]} }%
+\CSLLeftMargin{{[}12{]} }%
 \CSLRightInline{Malte Probst and Franz Rothlauf.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-design_of_modern_heuristics}{}}%
-\CSLLeftMargin{{[}12{]} }%
+\CSLLeftMargin{{[}13{]} }%
 \CSLRightInline{Franz Rothlauf. 2011. \emph{Design of modern heuristics: Principles and application}. DOI:https://doi.org/\href{https://doi.org/10.1007/978-3-540-72962-4}{10.1007/978-3-540-72962-4}}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-sampling_err_gp}{}}%
-\CSLLeftMargin{{[}13{]} }%
+\CSLLeftMargin{{[}14{]} }%
 \CSLRightInline{Dirk Schweim, David Wittenberg, and Franz Rothlauf.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-dae_orig2008}{}}%
-\CSLLeftMargin{{[}14{]} }%
+\CSLLeftMargin{{[}15{]} }%
 \CSLRightInline{Pascal Vincent, Hugo Larochelle, Y.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-weigend1994overfitting}{}}%
-\CSLLeftMargin{{[}15{]} }%
+\CSLLeftMargin{{[}16{]} }%
 \CSLRightInline{Andreas Weigend. 1994. On overfitting and the effective number of hidden units. In \emph{Proceedings of the 1993 connectionist models summer school}, 335--342.}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-daegp_explore_exploit}{}}%
-\CSLLeftMargin{{[}16{]} }%
+\CSLLeftMargin{{[}17{]} }%
 \CSLRightInline{David Wittenberg.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-dae-gp_2022_symreg}{}}%
-\CSLLeftMargin{{[}17{]} }%
+\CSLLeftMargin{{[}18{]} }%
 \CSLRightInline{David Wittenberg and Franz Rothlauf.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-dae-gp_2020_rtree}{}}%
-\CSLLeftMargin{{[}18{]} }%
+\CSLLeftMargin{{[}19{]} }%
 \CSLRightInline{David Wittenberg, Franz Rothlauf, and Dirk Schweim.

### 2023-02-03 paper.Rmd
autogit

@@
 As described in subsection [Effect on Run-Time], sampling new solutions has been a large contributor to the large performance hit on run-time by using pre-training.
-It could be a worthwhile endevour to explore new strategies, that
+It could be a worthwhile endevour to explore the main factors that increase sampling time for pre-trained DAE-GP to more efficiently exploit the benefits of pre-training like reducing the number of training epochs.
 
-[@pmlr-v5-erhan09a]: Pre-Training effect especially useful for lower-level layers -> Only adapt weights from low layers or embed the pre-trained model?
+Another limitation of this research is the software implementation used for all experiments.
+As briefly touched in the subsection [Effect on Run-Time], I found some evidence that my implementation of pre-training suffered from increased run-time costs for internal error handling in the used `keras` library for deep neural networks.
+By optimizing the source code or even re-implementing it in a more performance-focused programming language (such as `C++` or `Rust`), large improvements to the pre-training strategy might be achievable.
 
-Different pre-training strategies?
+Another possibility to improve the benefits of pre-training for DAE-GP is based on the findings of @pmlr-v5-erhan09a.
+Pre-Training has shown to be especially useful for the lower layers of DAEs so it could be worthwhile to further experiment with only adapting the weights for a subset of the hidden layers inside the DAE-LSTM networks to either reduce run-time or to improve on the quality of solutions found during the evolutionary search.
 

### 2023-02-03 paper.tex
autogit

@@
 
-\hypertarget{effect-on-runtime}{%
-\subsection{Effect on runtime}\label{effect-on-runtime}}
+\hypertarget{effect-on-run-time}{%
+\subsection{Effect on Run-Time}\label{effect-on-run-time}}
 
@@
 
-\hypertarget{limitations}{%
-\subsection{Limitations}\label{limitations}}
-
-\ldots{}
-
-\hypertarget{further-questions}{%
-\subsection{Further Questions}\label{further-questions}}
-
-{[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{4}{]}: Pre-Training effect especially useful for lower-level layers -\textgreater{} Only adapt weights from low layers or embed the pre-trained model?
-
-Different pre-training strategies?
+\hypertarget{limitations-and-open-questions}{%
+\subsection{Limitations and open Questions}\label{limitations-and-open-questions}}
+
+One major limitation for the experiments conducted during the reseach phase of this thesis has been the large computational effort for experimenting with DAE-GP.
+Since single runs for DAE-GP have a run-time of many hours, it was not possible for me to extend the experiments to larger population sizes or more generations.
+Also I did not perform an extended optimization of hyperparameters (e.g by using a grid search strategy) which might yield configurations that increase the benefits of pre-training.
+As described in subsection \protect\hyperlink{effect-on-run-time}{Effect on Run-Time}, sampling new solutions has been a large contributor to the large performance hit on run-time by using pre-training.
+It could be a worthwhile endevour to explore the main factors that increase sampling time for pre-trained DAE-GP to more efficiently exploit the benefits of pre-training like reducing the number of training epochs.
+
+Another limitation of this research is the software implementation used for all experiments.
+As briefly touched in the subsection \protect\hyperlink{effect-on-run-time}{Effect on Run-Time}, I found some evidence that my implementation of pre-training suffered from increased run-time costs for internal error handling in the used \texttt{keras} library for deep neural networks.
+By optimizing the source code or even re-implementing it in a more performance-focused programming language (such as \texttt{C++} or \texttt{Rust}), large improvements to the pre-training strategy might be achievable.
+
+Another possibility to improve the benefits of pre-training for DAE-GP is based on the findings of {[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{4}{]}.
+Pre-Training has shown to be especially useful for the lower layers of DAEs so it could be worthwhile to further experiment with only adapting the weights for a subset of the hidden layers inside the DAE-LSTM networks to either reduce run-time or to improve on the quality of solutions found during the evolutionary search.
 

### 2023-02-03 paper.Rmd
Fri Feb  3 20:00:06 CET 2023: update-remote.sh

@@
 csl: csl/association-for-computing-machinery.csl
+urlcolor: blue
+linkcolor: blue
 #csl: csl/harvard-cite-them-right.csl
@@
 The objective of the airfoil problem is to find a function that accurately predicts the output variable $y$ by taking in a subset of the input variables $x1,x2,x3,x4,x5$.
-The function set used by all DAE-GP variations for the airfoil problem is summarized in \@ref(tab:airfoil-fullRun-2hl-maxIndSize-params), the terminal set consists of the 5 input variables $x1,x2,x3,x4,x5$ and the ephemeral integer constants listed in \@ref(tab:airfoil-fullRun-2hl-maxIndSize-params)[@dae-gp_2022_symreg].
+The function set used by all DAE-GP variations for symbolic regression is summarized in \@ref(tab:airfoil-fullRun-2hl-maxIndSize-params), the terminal set consists of the the input variables $x1,x2,x3,x4,x5$ and the ephemeral integer constants listed in \@ref(tab:airfoil-fullRun-2hl-maxIndSize-params)[@dae-gp_2022_symreg].
 
@@
 
-## Effect on Generalisation
+## Influence on Generalisation
 
@@
 
-## Effect on solution quality
+## Influence on solution quality
 
@@
 
-## Effect on Run-Time
+## Influence on Run-Time
 
@@
 Another significant benefit of using pre-training in DAE-GP is that it is highly efficient in decreasing the number of training epochs that are necessary to train each generations DAE-LSTM model.
-Even though pre-training increased the median total run-time for the tested benchmark problem in combination with the chosen DAE-LSTM dimensions, reducing the number of training epochs might be a possible way to lower the run time of DAE-GP if further performance optimizations to the source code, especially during model sampling, can be achieved.
+Even though pre-training increased the median total run-time for the tested benchmark problem in combination with the chosen DAE-LSTM dimensions, reducing the number of training epochs might be a possible way to lower the run time of DAE-GP if further performance optimizations, especially during model sampling, can be achieved.
 
@@
 Another limitation of this research is the software implementation used for all experiments.
-As briefly touched in the subsection [Effect on Run-Time], I found some evidence that my implementation of pre-training suffered from increased run-time costs for internal error handling in the used `keras` library for deep neural networks.
+As briefly touched in the subsection [Influence on Run-Time], I found some evidence that my implementation of pre-training suffered from increased run-time costs for internal error handling in the used `keras` library for deep neural networks.
 By optimizing the source code or even re-implementing it in a more performance-focused programming language (such as `C++` or `Rust`), large improvements to the pre-training strategy might be achievable.

### 2023-02-03 paper.tex
Fri Feb  3 20:00:06 CET 2023: update-remote.sh

@@
 \PassOptionsToPackage{hyphens}{url}
+\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
 %
@@
 \hypersetup{
-  hidelinks,
+  colorlinks=true,
+  linkcolor={blue},
+  filecolor={Maroon},
+  citecolor={Blue},
+  urlcolor={blue},
   pdfcreator={LaTeX via pandoc}}
@@
 {
+\hypersetup{linkcolor=}
 \setcounter{tocdepth}{3}
@@
 The objective of the airfoil problem is to find a function that accurately predicts the output variable \(y\) by taking in a subset of the input variables \(x1,x2,x3,x4,x5\).
-The function set used by all DAE-GP variations for the airfoil problem is summarized in \ref{tab:airfoil-fullRun-2hl-maxIndSize-params}, the terminal set consists of the 5 input variables \(x1,x2,x3,x4,x5\) and the ephemeral integer constants listed in \ref{tab:airfoil-fullRun-2hl-maxIndSize-params}{[}\protect\hyperlink{ref-dae-gp_2022_symreg}{18}{]}.
+The function set used by all DAE-GP variations for symbolic regression is summarized in \ref{tab:airfoil-fullRun-2hl-maxIndSize-params}, the terminal set consists of the the input variables \(x1,x2,x3,x4,x5\) and the ephemeral integer constants listed in \ref{tab:airfoil-fullRun-2hl-maxIndSize-params}{[}\protect\hyperlink{ref-dae-gp_2022_symreg}{18}{]}.
 
@@
 
-\hypertarget{effect-on-generalisation}{%
-\subsection{Effect on Generalisation}\label{effect-on-generalisation}}
+\hypertarget{influence-on-generalisation}{%
+\subsection{Influence on Generalisation}\label{influence-on-generalisation}}
 
@@
 
-\hypertarget{effect-on-solution-quality}{%
-\subsection{Effect on solution quality}\label{effect-on-solution-quality}}
+\hypertarget{influence-on-solution-quality}{%
+\subsection{Influence on solution quality}\label{influence-on-solution-quality}}
 
@@
 
-\hypertarget{effect-on-run-time}{%
-\subsection{Effect on Run-Time}\label{effect-on-run-time}}
+\hypertarget{influence-on-run-time}{%
+\subsection{Influence on Run-Time}\label{influence-on-run-time}}
 
@@
 Also I did not perform an extended optimization of hyperparameters (e.g by using a grid search strategy) which might yield configurations that increase the benefits of pre-training.
-As described in subsection \protect\hyperlink{effect-on-run-time}{Effect on Run-Time}, sampling new solutions has been a large contributor to the large performance hit on run-time by using pre-training.
+As described in subsection {[}Effect on Run-Time{]}, sampling new solutions has been a large contributor to the large performance hit on run-time by using pre-training.
 It could be a worthwhile endevour to explore the main factors that increase sampling time for pre-trained DAE-GP to more efficiently exploit the benefits of pre-training like reducing the number of training epochs.
@@
 Another limitation of this research is the software implementation used for all experiments.
-As briefly touched in the subsection \protect\hyperlink{effect-on-run-time}{Effect on Run-Time}, I found some evidence that my implementation of pre-training suffered from increased run-time costs for internal error handling in the used \texttt{keras} library for deep neural networks.
+As briefly touched in the subsection \protect\hyperlink{influence-on-run-time}{Influence on Run-Time}, I found some evidence that my implementation of pre-training suffered from increased run-time costs for internal error handling in the used \texttt{keras} library for deep neural networks.
 By optimizing the source code or even re-implementing it in a more performance-focused programming language (such as \texttt{C++} or \texttt{Rust}), large improvements to the pre-training strategy might be achievable.

### 2023-02-03 paper.Rmd
autogit

@@
 indent: false
-fontsize: 12pt
+fontsize: 11pt
 bibliography: ./ref/ref.bib
@@
 linkcolor: blue
+toccolor: blue
 #csl: csl/harvard-cite-them-right.csl
@@
 For comparison, figure \@ref(fig:airfoil-maxIndSize-params) shows the total number of trainable parameters for each generation's DAE-LSTM of another experiment where regular DAE-GP was tested with dynamic adjustment of the number of hidden neurons as described in [@dae-gp_2020_rtree] and [@dae-gp_2022_symreg].
-Again, the experiments was based on $10$ individual runs and otherwise used the same hyperparameters as the previous one.
-It clearly shows that DAE-GP can strongly reduce the number of trainable parameters by adjusting the depth of it's hidden layers to the current generations population which will further reduce the amount of training necessary and likely reduce the run-time of DAE-GP.
+Again, the experiments was based on $10$ individual runs and otherwise used the same hyperparameters as listed in table \@ref(tab:airfoil-fullRun-2hl-maxIndSize-params).
+It clearly shows that DAE-GP can strongly reduce the number of trainable parameters by adjusting the number of hidden neurons per hidden layers to the current generations population which largley reduces the amount of training necessary and likely reduce the run-time of DAE-GP.
 
@@
 ```
-... describe differences in distribution, interquartile range blablabla...
+
+<!-- An interesting observation from studying the distribution of final fitness scores for both algorithms is that pre-trained DAE-GP has a smaller interquartile range and more extreme outliers for high RMSE which explains the slightly worse performance if looking at median best fitness present in figure \@ref(fig:airfoil-median-fitness-30gens).
+-->
+
+Interestingly, figure \@ref(fig:airfoil-finalFitness) shows that the distribution of final fitness scores for pre-trained DAE-GP has a larger interquartile range than regular DAE-GP.
+This is opposed to the distributions of final fitness score that was measured for the same problem run with two hidden layers (see figure \@ref(fig:airfoil-distribution-fitness-30gens)).
+Otherwise, pre-trained DAE-GP again has more outliers for high RMSE than regular DAE-GP as seen in the previous experiment with 2 hidden layers.
 
@@
 
-```{r altStrategies-Fitness, echo=FALSE, fig.cap="Best Fitness over 30 Generations - Airfoil - Alternative Pre-Strategies ", out.width="80%"}
+```{r altStrategies-Fitness, echo=FALSE, fig.cap="Best Fitness over 30 Generations - Airfoil - Alternative Pre-Training Strategies ", out.width="80%"}
 knitr::include_graphics("./img/PT_Strategies_airfoil_2hl_150hn_fullRun/median_fitness_byGens.png")
@@
 
-```{r altStrategies-FitnessDistr, echo=FALSE, fig.cap="Best Fitness on Test Set after 30 Generations - Airfoil - Alternative Pre-Trategies", out.width="80%"}
+```{r altStrategies-FitnessDistr, echo=FALSE, fig.cap="Best Fitness on Test Set after 30 Generations - Airfoil - Alternative Pre-Training Strategies", out.width="80%"}
 knitr::include_graphics("./img/PT_Strategies_airfoil_2hl_150hn_fullRun/final_fit_boxplot(test_only).png")

### 2023-02-03 paper.tex
autogit

@@
 \documentclass[
-  12pt,
+  11pt,
 ]{article}
@@
 {
-\hypersetup{linkcolor=}
+\hypersetup{linkcolor=blue}
 \setcounter{tocdepth}{3}
@@
 For comparison, figure \ref{fig:airfoil-maxIndSize-params} shows the total number of trainable parameters for each generation's DAE-LSTM of another experiment where regular DAE-GP was tested with dynamic adjustment of the number of hidden neurons as described in {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{19}{]} and {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{18}{]}.
-Again, the experiments was based on \(10\) individual runs and otherwise used the same hyperparameters as the previous one.
+Again, the experiments was based on \(10\) individual runs and otherwise used the same hyperparameters as listed in table \ref{tab:airfoil-fullRun-2hl-maxIndSize-params}.
 It clearly shows that DAE-GP can strongly reduce the number of trainable parameters by adjusting the depth of it's hidden layers to the current generations population which will further reduce the amount of training necessary and likely reduce the run-time of DAE-GP.
@@
 
-\ldots{} describe differences in distribution, interquartile range blablabla\ldots{}
+Interestingly, figure \ref{fig:airfoil-finalFitness} shows that the distribution of final fitness scores for pre-trained DAE-GP has a larger interquartile range than regular DAE-GP.
+This is opposed to the distributions of final fitness score that was measured for the same problem run with two hidden layers (see figure \ref{fig:airfoil-distribution-fitness-30gens}).
+Otherwise, pre-trained DAE-GP again has more outliers for high RMSE than regular DAE-GP as seen in the previous experiment with 2 hidden layers.
 
@@
 
-\caption{Best Fitness over 30 Generations - Airfoil - Alternative Pre-Strategies }\label{fig:altStrategies-Fitness}
+\caption{Best Fitness over 30 Generations - Airfoil - Alternative Pre-Training Strategies }\label{fig:altStrategies-Fitness}
 \end{figure}
@@
 
-\caption{Best Fitness on Test Set after 30 Generations - Airfoil - Alternative Pre-Trategies}\label{fig:altStrategies-FitnessDistr}
+\caption{Best Fitness on Test Set after 30 Generations - Airfoil - Alternative Pre-Training Strategies}\label{fig:altStrategies-FitnessDistr}
 \end{figure}

### 2023-02-04 paper.Rmd
Sat Feb  4 17:00:05 CET 2023: update-remote.sh

@@
 <!--- airfoil dataset -->
-The main experiments conducted in this thesis uses the NASA Airfoil Self-Noise Data Set which is part of the UCI machine learning repository [@machine_learning_repo].
-The dataset consists of 5 input variables and 1 output variable that are listed in table \@ref(tab:airfoil-dataset-description).
+The main experiments conducted in this thesis uses the NASA Airfoil Self-Noise Data Set [@AIRFOIL_DATASET] that consists of 5 input variables and 1 output variable that are listed in table \@ref(tab:airfoil-dataset-description).
 
@@
 
-To gather more confidence in the previous results, the pre-training strategy for DAE-GP was applied for three additional real world symbolic regression problems taken from @machine_learning_repo and tested for statistically significant differences in the underlying distributions.
+To gather more confidence in the previous results, the pre-training strategy for DAE-GP was applied for three additional real world symbolic regression problems and tested for statistically significant differences in the underlying distributions.
 Table \@ref(tab:full-run-realWorldSymReg-problems) briefly summarizes the benchmark problems.
@@
   read.csv("./tables/full_run_realWorldSymReg_problems.csv"),
-  caption="Real World Symbolic Regression Benchmark Problems"
+  caption="Real World Symbolic Regression Benchmark Problems",
+  format = "markdown"
 )%>% row_spec(0,bold=TRUE ) %>% kableExtra::kable_styling(latex_options = "hold_position")

### 2023-02-04 paper.tex
Sat Feb  4 17:00:05 CET 2023: update-remote.sh

@@
 \author{}
-\date{\vspace{-2.5em}Date of Submission: 2023-02-03}
+\date{\vspace{-2.5em}Date of Submission: 2023-02-04}
 
@@
 
-Denoising Autoencoder Genetic Programming (DAE-GP) is a novel variation of an genetic programming based Estimation of Distribution Algorithm (EDA-GP) that uses a denoising autoencoders long short-term memory network (DAE-LSTMs) as a probabilistic model to sample new candidate solutions {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{19}{]}.
+Denoising Autoencoder Genetic Programming (DAE-GP) is a novel variation of an genetic programming based Estimation of Distribution Algorithm (EDA-GP) that uses a denoising autoencoders long short-term memory network (DAE-LSTMs) as a probabilistic model to sample new candidate solutions {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{21}{]}.
 
@@
 In DAE-GP, DAE-LSTMs are trained with a subset of high-fitness solutions selected from a parent population with the aim to capture their promising qualities.
-The resulting model is then used to sample new offspring solutions by propagating partially mutated solutions from the parent population through the DAE-LSTMS {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{19}{]}.
-In previous work DAE-GP has been shown to outperform GP for both a generalized version of the royal tree problem {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{19}{]} as well as for a real-world symbolic regression problem {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{18}{]}.
-
-The DAE-GP algorithm first described by {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{19}{]} trains a DAE-LSTMs for each generation \(g\) of the search from scratch.
-{[}\protect\hyperlink{ref-dae-gp_2022_symreg}{18}{]} and {[}\protect\hyperlink{ref-daegp_explore_exploit}{17}{]} suggests the incorporation of a pre-training strategy into the evolutionary search as a possible way of improving the overall performance of the DAE-GP algorithm.
+The resulting model is then used to sample new offspring solutions by propagating partially mutated solutions from the parent population through the DAE-LSTMS {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{21}{]}.
+In previous work DAE-GP has been shown to outperform GP for both a generalized version of the royal tree problem {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{21}{]} as well as for a real-world symbolic regression problem {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{20}{]}.
+
+The DAE-GP algorithm first described by {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{21}{]} trains a DAE-LSTMs for each generation \(g\) of the search from scratch.
+{[}\protect\hyperlink{ref-dae-gp_2022_symreg}{20}{]} and {[}\protect\hyperlink{ref-daegp_explore_exploit}{19}{]} suggests the incorporation of a pre-training strategy into the evolutionary search as a possible way of improving the overall performance of the DAE-GP algorithm.
 The key idea is to pre-train an initial DAE-LSTM on a large population of candidate solutions and to use the pre-trained parameters of this initial model in each generation of the search as starting parameters for the current generation DAE-LSTM.
@@
 
-DAE-GP is an EDA-GP algorithm that uses DAE-LSTM networks as a probabilistic model to sample new offspring solutions {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{19}{]}.
+DAE-GP is an EDA-GP algorithm that uses DAE-LSTM networks as a probabilistic model to sample new offspring solutions {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{21}{]}.
 
@@
 Another important concept in EC is the distinction between genotypes and phenotypes of solutions, the genotype contains the information that is necessary to construct the phenotype, the outer appearance of a particular solution on which we measure the overall quality of solutions.
-The representation of a solution is therefore defined by the mapping of genotypes to phenotypes {[}\protect\hyperlink{ref-design_of_modern_heuristics}{13}{]}.
+The representation of a solution is therefore defined by the mapping of genotypes to phenotypes {[}\protect\hyperlink{ref-design_of_modern_heuristics}{14}{]}.
 Genetic operators, such as mutation or recombination are usually applied to the genotype of solutions.
@@
 
-GP follows the same basic evolutionary principle of EC but searches for more general, hierarchical computer programs of dynamically varying size and shape {[}\protect\hyperlink{ref-Koza1993GeneticP}{9}{]}.
-The computer programs that are at the center of the evolutionary search in GP are commonly represented by tree structures at the level of their phenotype {[}\protect\hyperlink{ref-design_of_modern_heuristics}{13}{]}.
-Since GP searches for high fitness computer programs that produce a desired output for some input, it can be applied to various different problem domains such as symbolic regression, automatic programming, or evolving game-playing strategies {[}\protect\hyperlink{ref-Koza1993GeneticP}{9}{]}.
+GP follows the same basic evolutionary principle of EC but searches for more general, hierarchical computer programs of dynamically varying size and shape {[}\protect\hyperlink{ref-Koza1993GeneticP}{10}{]}.
+The computer programs that are at the center of the evolutionary search in GP are commonly represented by tree structures at the level of their phenotype {[}\protect\hyperlink{ref-design_of_modern_heuristics}{14}{]}.
+Since GP searches for high fitness computer programs that produce a desired output for some input, it can be applied to various different problem domains such as symbolic regression, automatic programming, or evolving game-playing strategies {[}\protect\hyperlink{ref-Koza1993GeneticP}{10}{]}.
 An important quality of GP is the ability to search for solutions of variable length and structure.
 GP is an especially useful meta heuristic for problems where no a priori knowledge about the final form of good solutions is available.
-GPs ability to optimize solutions for their structure as well as for their parameters led to it being one of the most prevalent methods used for symbolic regression {[}\protect\hyperlink{ref-10.1007ux2f978-3-540-24621-3_22}{11}{]}.
+GPs ability to optimize solutions for their structure as well as for their parameters led to it being one of the most prevalent methods used for symbolic regression {[}\protect\hyperlink{ref-10.1007ux2f978-3-540-24621-3_22}{12}{]}.
 
@@
 
-The aim of Estimation of distribution algorithms (EDA) is to replace the standard variation operators used in GA by building probabilistic models that can capture complex dependencies between different decision variables of an optimization problem {[}\protect\hyperlink{ref-design_of_modern_heuristics}{13}{]}.
+The aim of Estimation of distribution algorithms (EDA) is to replace the standard variation operators used in GA by building probabilistic models that can capture complex dependencies between different decision variables of an optimization problem {[}\protect\hyperlink{ref-design_of_modern_heuristics}{14}{]}.
 EDAs use this probabilistic model to sample new offspring solutions inside an evolutionary search to replace crossover and/or mutation operators.
@@
 
-One possible way of model building in EDA proposed by {[}\protect\hyperlink{ref-harmless_overfitting_eda}{12}{]} is to use denoising autoencoders (DAEs) as generative models to capture complex probability distributions of decision variables.
+One possible way of model building in EDA proposed by {[}\protect\hyperlink{ref-harmless_overfitting_eda}{13}{]} is to use denoising autoencoders (DAEs) as generative models to capture complex probability distributions of decision variables.
 DAE, a variation of the autoencoders (AE), are a widely used type of neural networks in the domain of unsupervised machine learning that maps \(n\) input variables to \(n\) output variables using a hidden representation.
 
-AE were introduced by {[}\protect\hyperlink{ref-ae_orig}{7}{]} to compress high-dimensional data into lower-dimensions. An AE consists of two different sub-units:
+AE were introduced by {[}\protect\hyperlink{ref-ae_orig}{8}{]} to compress high-dimensional data into lower-dimensions. An AE consists of two different sub-units:
 
@@
 
-The AE is trained to reduce the reconstruction error between input and output data, after the training procedure is finished the network is able to reduce the dimensional of input data to a compressed representation{[}\protect\hyperlink{ref-ae_orig}{7}{]}.
-
-DAE was first introduced by {[}\protect\hyperlink{ref-dae_orig2008}{15}{]} as an improved AE with the ability to learn new representations of data that is especially robust to partially corrupted input data.
+The AE is trained to reduce the reconstruction error between input and output data, after the training procedure is finished the network is able to reduce the dimensional of input data to a compressed representation{[}\protect\hyperlink{ref-ae_orig}{8}{]}.
+
+DAE was first introduced by {[}\protect\hyperlink{ref-dae_orig2008}{17}{]} as an improved AE with the ability to learn new representations of data that is especially robust to partially corrupted input data.
 DAE modifies the AE by using partially corrupted input data for the AE and training it to reconstruct the uncorrupted, original version of the input data.
 
-Since the hidden representation of DAE captures the dependency structure of the input data it can therefore also be used to generate new solutions in the context of GAs {[}\protect\hyperlink{ref-harmless_overfitting_eda}{12}{]}.
-
-DAE-GP builds upon the concept of using DAEs in EDAs described by {[}\protect\hyperlink{ref-harmless_overfitting_eda}{12}{]} and transfers the concept to the domain of GP.
-The mutation and crossover operators of standard GP are replaced by sampling new solutions from a probabilistic model that is build by training a DAEs long short-term memory (LSTM) network on a subset of high fitness solutions from the current population {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{19}{]}.
-
-LSTMs are a variant of artificial neural networks first introduced by {[}\protect\hyperlink{ref-lstm_orig}{8}{]} that can store learned information over en extended time period while avoiding the problem of vanishing and/or exploding gradients.
-Since DAE-GP encodes candidate solutions as linear strings in prefix notation, the DAE in DAE-GP uses LSTMs for both encoding and decoding where the total amount of time steps \(T\) is equal to sum of the length of the input solution and the output solution {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{19}{]}.
+Since the hidden representation of DAE captures the dependency structure of the input data it can therefore also be used to generate new solutions in the context of GAs {[}\protect\hyperlink{ref-harmless_overfitting_eda}{13}{]}.
+
+DAE-GP builds upon the concept of using DAEs in EDAs described by {[}\protect\hyperlink{ref-harmless_overfitting_eda}{13}{]} and transfers the concept to the domain of GP.
+The mutation and crossover operators of standard GP are replaced by sampling new solutions from a probabilistic model that is build by training a DAEs long short-term memory (LSTM) network on a subset of high fitness solutions from the current population {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{21}{]}.
+
+LSTMs are a variant of artificial neural networks first introduced by {[}\protect\hyperlink{ref-lstm_orig}{9}{]} that can store learned information over en extended time period while avoiding the problem of vanishing and/or exploding gradients.
+Since DAE-GP encodes candidate solutions as linear strings in prefix notation, the DAE in DAE-GP uses LSTMs for both encoding and decoding where the total amount of time steps \(T\) is equal to sum of the length of the input solution and the output solution {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{21}{]}.
 
@@
 
-The strategy of pre-training artificial neural networks is based on the idea of transfer learning that aims at retaining previously learned knowledge from one task to re-use it for another task {[}\protect\hyperlink{ref-HAN2021225}{6}{]} {[}\protect\hyperlink{ref-survey_transfer_learning}{10}{]}.
+The strategy of pre-training artificial neural networks is based on the idea of transfer learning that aims at retaining previously learned knowledge from one task to re-use it for another task {[}\protect\hyperlink{ref-HAN2021225}{6}{]} {[}\protect\hyperlink{ref-survey_transfer_learning}{11}{]}.
 Transfer learning traditionally follows a two phase approach:
@@
 
-where source task and target task are usually similar but may differ in their feature space and the distribution of training data {[}\protect\hyperlink{ref-survey_transfer_learning}{10}{]}.
+where source task and target task are usually similar but may differ in their feature space and the distribution of training data {[}\protect\hyperlink{ref-survey_transfer_learning}{11}{]}.
 
@@
 Instead of initializing and optimizing each model from scratch as done in previous work (e.g.
-{[}\protect\hyperlink{ref-dae-gp_2020_rtree}{19}{]} {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{18}{]}), a separate DAE-LSTM network \(\hat{M}\) will be trained on a large population of randomly initialized solutions.\\
+{[}\protect\hyperlink{ref-dae-gp_2020_rtree}{21}{]} {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{20}{]}), a separate DAE-LSTM network \(\hat{M}\) will be trained on a large population of randomly initialized solutions.\\
 The trainable parameters \(\theta_{\hat{M}}\) obtained after finishing the training procedure of \(\hat{M}\) are then used as the starting parameters for each following DAE-LSTM model \(M_g\) for \(g\in{1,...,g_{max}}\).
@@
 Another motivation for introducing pre-training into DAE-GP is the prevalence of sampling error for small GP populations sizes.
-{[}\protect\hyperlink{ref-sampling_err_gp}{14}{]} finds that sampling error, non-systematic errors that are caused by observing only a small subset of the statistical population, is a severe problem in the domain of GP since the initial population may not be a representative sample of all possible solutions.
-{[}\protect\hyperlink{ref-sampling_err_gp}{14}{]} also introduces a method for calculating optimal population sizes to minimize the presence of sampling error.
+{[}\protect\hyperlink{ref-sampling_err_gp}{15}{]} finds that sampling error, non-systematic errors that are caused by observing only a small subset of the statistical population, is a severe problem in the domain of GP since the initial population may not be a representative sample of all possible solutions.
+{[}\protect\hyperlink{ref-sampling_err_gp}{15}{]} also introduces a method for calculating optimal population sizes to minimize the presence of sampling error.
 Since DAE-LSTM of DAE-GP learns the properties of it's training population and reuses the acquired knowledge in the sampling procedure, using the parameters obtained from training \(\hat{M}\) on a sufficiently large training population might reduce the need for large population sizes in the following generations of the search if \(\hat{M}\) already implicitly captured the properties of a representative sample of solutions.
@@
 
-The DAE-GP algorithm, first described by {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{19}{]}, is summarized as a flowchart in figure \ref{fig:daegp-flowchart}.
+The DAE-GP algorithm, first described by {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{21}{]}, is summarized as a flowchart in figure \ref{fig:daegp-flowchart}.
 After setting the generation counter \(g\) to \(0\) and creating the initial population of solutions \(P_0\), DAE-GP enters the main loop that checks if a termination criteria is satisfied (i.e.~maximum number of generations or fitness evaluations is reached).
@@
 One difficulty in the implementation of a pre-training strategy into DAE-GP has been the determination of the number of hidden neurons in side the DAE-LSTMs hidden layers.
-The original description of DAE-GP {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{19}{]} uses a strategy where the number of hidden neurons for each hidden layer is dynamically set per generation to the maximum individual size inside the current population.
+The original description of DAE-GP {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{21}{]} uses a strategy where the number of hidden neurons for each hidden layer is dynamically set per generation to the maximum individual size inside the current population.
 For a pre-training implementation, this strategy can not be easily adapted since it leads to a changing number of neurons at each generation resulting in different dimensions of the DAE-LSTM.
@@
 
-The framework that was used to conduct all experiments of this thesis was provided by the supervisor of this thesis David Wittenberg and uses the python programming language in conjunction with the baseline libraries \texttt{Keras} {[}\protect\hyperlink{ref-chollet2015keras}{1}{]} for deep neural networks and \texttt{deap} {[}\protect\hyperlink{ref-DEAP_JMLR2012}{5}{]} for evolutionary computation.
+The framework that was used to conduct all experiments of this thesis was provided by the supervisor of this thesis David Wittenberg and uses the python programming language in conjunction with the baseline libraries \texttt{Keras} {[}\protect\hyperlink{ref-chollet2015keras}{2}{]} for deep neural networks and \texttt{deap} {[}\protect\hyperlink{ref-DEAP_JMLR2012}{5}{]} for evolutionary computation.
 
@@
 To test pre-training in DAE-GP this thesis focuses on the domain of real-world symbolic regression problems.
-Symbolic Regression problems have been one of the first GP applications {[}\protect\hyperlink{ref-Koza1993GeneticP}{9}{]} and are an actively studied and highly relevant research area.
-The goal in symbolic regression is to find a mathematical model for a given set of data points {[}\protect\hyperlink{ref-10.1007ux2f978-3-540-24621-3_22}{11}{]}, in real-world symbolic regression these data points are sourced from real-world observations which in contrast to synthetic symbolic regression problem are more likely to contain random noise and bias.
+Symbolic Regression problems have been one of the first GP applications {[}\protect\hyperlink{ref-Koza1993GeneticP}{10}{]} and are an actively studied and highly relevant research area.
+The goal in symbolic regression is to find a mathematical model for a given set of data points {[}\protect\hyperlink{ref-10.1007ux2f978-3-540-24621-3_22}{12}{]}, in real-world symbolic regression these data points are sourced from real-world observations which in contrast to synthetic symbolic regression problem are more likely to contain random noise and bias.
 Another important challenge in solving real-world symbolic regression problems is the ability for a given model to generalize, we want the final model to show high accuracy in predicting outcomes for previously unseen cases.
 
-The main experiments conducted in this thesis uses the NASA Airfoil Self-Noise Data Set which is part of the UCI machine learning repository {[}\protect\hyperlink{ref-machine_learning_repo}{2}{]}.
-The dataset consists of 5 input variables and 1 output variable that are listed in table \ref{tab:airfoil-dataset-description}.
+The main experiments conducted in this thesis uses the NASA Airfoil Self-Noise Data Set {[}\protect\hyperlink{ref-AIRFOIL_DATASET}{1}{]} that consists of 5 input variables and 1 output variable that are listed in table \ref{tab:airfoil-dataset-description}.
 
@@
 The objective of the airfoil problem is to find a function that accurately predicts the output variable \(y\) by taking in a subset of the input variables \(x1,x2,x3,x4,x5\).
-The function set used by all DAE-GP variations for symbolic regression is summarized in \ref{tab:airfoil-fullRun-2hl-maxIndSize-params}, the terminal set consists of the the input variables \(x1,x2,x3,x4,x5\) and the ephemeral integer constants listed in \ref{tab:airfoil-fullRun-2hl-maxIndSize-params}{[}\protect\hyperlink{ref-dae-gp_2022_symreg}{18}{]}.
+The function set used by all DAE-GP variations for symbolic regression is summarized in \ref{tab:airfoil-fullRun-2hl-maxIndSize-params}, the terminal set consists of the the input variables \(x1,x2,x3,x4,x5\) and the ephemeral integer constants listed in \ref{tab:airfoil-fullRun-2hl-maxIndSize-params}{[}\protect\hyperlink{ref-dae-gp_2022_symreg}{20}{]}.
 
@@
 For each configuration I tested traditional DAE-GP as well as a pre-trained DAE-GP resulting in 12 total sub experiments that were each based on 10 individual runs (total number of runs=\(120\)).
-To avoid creating biased results through the presence of sampling error inside the pre-training population (see {[}\protect\hyperlink{ref-sampling_err_gp}{14}{]}), the population size for the pre-training phase is chosen very high with the size of \(\hat{P}\) =10000 where 50\% of \(\hat{P}\) is used for the training population \(\hat{P}_{train}\) and the remaining 50\% are used for the hold-out validation population \(\hat{P}_{test}\).
+To avoid creating biased results through the presence of sampling error inside the pre-training population (see {[}\protect\hyperlink{ref-sampling_err_gp}{15}{]}), the population size for the pre-training phase is chosen very high with the size of \(\hat{P}\) =10000 where 50\% of \(\hat{P}\) is used for the training population \(\hat{P}_{train}\) and the remaining 50\% are used for the hold-out validation population \(\hat{P}_{test}\).
 The training of DAE-LSTM uses a fixed number of 1000 epochs until termination and uses Adam optimization for gradient descent.
@@
 In general I expect that with a growing number of trainable parameters (either by adding more hidden layers or more hidden neurons per layer) the DAE-LSTM will be more prone to overfit to the training population \(P_{train}\) resulting in a small reconstruction error for the training population and a large one for the validation population \(P_{test}\).
-The reason for this effect is that a larger network trained over an extended period of time (without the use of strategies like early stopping), has much more potential to learn noise from the training dataset than a smaller network, which is more likely to result in worsening performance for previously unseen cases {[}\protect\hyperlink{ref-weigend1994overfitting}{16}{]}.
+The reason for this effect is that a larger network trained over an extended period of time (without the use of strategies like early stopping), has much more potential to learn noise from the training dataset than a smaller network, which is more likely to result in worsening performance for previously unseen cases {[}\protect\hyperlink{ref-weigend1994overfitting}{18}{]}.
 
@@
 
-Since previous work by {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{18}{]} demonstrated that DAE-GP, for a given number of 10.000 fitness evaluations, is able to produce solutions that are much smaller in tree size than regular GP, I also studied the influence of pre-training on the size of solutions found by DAE-GP.
+Since previous work by {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{20}{]} demonstrated that DAE-GP, for a given number of 10.000 fitness evaluations, is able to produce solutions that are much smaller in tree size than regular GP, I also studied the influence of pre-training on the size of solutions found by DAE-GP.
 Figure \ref{fig:airfoil-size-combined-30gens} shows the median of both the average size of individuals inside the population as well as the size of the currently best performing solution inside the population.
@@
 To ensure a fair comparison between pre-trained and regular DAE-GP, the prior experiments used a static number of hidden neurons for each hidden Layer inside the DAE-LSTM networks.
-For comparison, figure \ref{fig:airfoil-maxIndSize-params} shows the total number of trainable parameters for each generation's DAE-LSTM of another experiment where regular DAE-GP was tested with dynamic adjustment of the number of hidden neurons as described in {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{19}{]} and {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{18}{]}.
+For comparison, figure \ref{fig:airfoil-maxIndSize-params} shows the total number of trainable parameters for each generation's DAE-LSTM of another experiment where regular DAE-GP was tested with dynamic adjustment of the number of hidden neurons as described in {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{21}{]} and {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{20}{]}.
 Again, the experiments was based on \(10\) individual runs and otherwise used the same hyperparameters as listed in table \ref{tab:airfoil-fullRun-2hl-maxIndSize-params}.
-It clearly shows that DAE-GP can strongly reduce the number of trainable parameters by adjusting the depth of it's hidden layers to the current generations population which will further reduce the amount of training necessary and likely reduce the run-time of DAE-GP.
+It clearly shows that DAE-GP can strongly reduce the number of trainable parameters by adjusting the number of hidden neurons per hidden layers to the current generations population which largley reduces the amount of training necessary and likely reduce the run-time of DAE-GP.
 
@@
 
-Since previous research on DAE-GP used a single hidden layers for the airfoil dataset {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{18}{]}, I repeated the experiment with a single hidden layer.
+Since previous research on DAE-GP used a single hidden layers for the airfoil dataset {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{20}{]}, I repeated the experiment with a single hidden layer.
 Based on the findings of {[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{4}{]}, I expected that pre-training DAE-GP would lead to a decrease in overall performance.
@@
 
-The pre-training implementation in previous experiments used ramped half and half initialization {[}\protect\hyperlink{ref-Koza1993GeneticP}{9}{]} to create \(\hat{P}\), this leads to 50\% of the population consisting of full grown trees.
+The pre-training implementation in previous experiments used ramped half and half initialization {[}\protect\hyperlink{ref-Koza1993GeneticP}{10}{]} to create \(\hat{P}\), this leads to 50\% of the population consisting of full grown trees.
 An alternative method of initializing \(\hat{P}\) is to use the grow initialization method where solutions are not guaranteed to be full trees.
@@
 
-To gather more confidence in the previous results, the pre-training strategy for DAE-GP was applied for three additional real world symbolic regression problems taken from {[}\protect\hyperlink{ref-machine_learning_repo}{2}{]} and tested for statistically significant differences in the underlying distributions.
+To gather more confidence in the previous results, the pre-training strategy for DAE-GP was applied for three additional real world symbolic regression problems and tested for statistically significant differences in the underlying distributions.
 Table \ref{tab:full-run-realWorldSymReg-problems} briefly summarizes the benchmark problems.
 
-\begin{table}[!h]
-
-\caption{\label{tab:full-run-realWorldSymReg-problems}Real World Symbolic Regression Benchmark Problems}
-\centering
-\begin{tabular}[t]{l|r|r}
-\hline
-\textbf{Problem} & \textbf{Observations} & \textbf{Features}\\
-\hline
-Airfoil & 1503 & 5\\
-\hline
-Boston\_Housing & 506 & 13\\
-\hline
-Energy\_Cooling & 768 & 8\\
-\hline
-Concrete & 1030 & 8\\
-\hline
-\end{tabular}
-\end{table}
+\begin{longtable}[]{@{}lrrl@{}}
+\caption{\label{tab:full-run-realWorldSymReg-problems}Real World Symbolic Regression Benchmark Problems}\tabularnewline
+\toprule\noalign{}
+Problem & Observations & Features & Source \\
+\midrule\noalign{}
+\endfirsthead
+\toprule\noalign{}
+Problem & Observations & Features & Source \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+Airfoil & 1503 & 5 & {[}\protect\hyperlink{ref-AIRFOIL_DATASET}{1}{]} \\
+Boston\_Housing & 506 & 13 & {[}\protect\hyperlink{ref-BOSTON_HOUSING_DATASET}{7}{]} \\
+Energy\_Cooling & 768 & 8 & {[}\protect\hyperlink{ref-ENERGY_DATASET}{16}{]} \\
+Concrete & 1030 & 8 & {[}\protect\hyperlink{ref-CONCRETE_DATASET}{22}{]} \\
+\end{longtable}
 
@@
 For a single hidden layer, pre-training did show a negative impact on the median final fitness of solutions found as it was suggested by {[}\protect\hyperlink{ref-pmlr-v9-erhan10a}{3}{]}.
-Since current research on DAE-GP for real-world symbolic regression relies on a single hidden layer (see {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{18}{]}), pre-training does come with additional computational expenses for using deeper DAE-LSTMs.
+Since current research on DAE-GP for real-world symbolic regression relies on a single hidden layer (see {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{20}{]}), pre-training does come with additional computational expenses for using deeper DAE-LSTMs.
 
@@
 \begin{CSLReferences}{0}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-AIRFOIL_DATASET}{}}%
+\CSLLeftMargin{{[}1{]} }%
+\CSLRightInline{Thomas F. Brooks, Dennis S. Pope, and Michael A. Marcolini. 1989. Airfoil self-noise and prediction. }
+
 \leavevmode\vadjust pre{\hypertarget{ref-chollet2015keras}{}}%
-\CSLLeftMargin{{[}1{]} }%
+\CSLLeftMargin{{[}2{]} }%
 \CSLRightInline{Francois Chollet and others. 2015. Keras. Retrieved from \url{https://github.com/fchollet/keras}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-machine_learning_repo}{}}%
-\CSLLeftMargin{{[}2{]} }%
-\CSLRightInline{Dheeru Dua and Casey Graff. 2017. {UCI} machine learning repository. Retrieved from \url{http://archive.ics.uci.edu/ml}}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-BOSTON_HOUSING_DATASET}{}}%
+\CSLLeftMargin{{[}7{]} }%
+\CSLRightInline{David Harrison and Daniel Rubinfeld.
+1978.
+Hedonic housing prices and the demand for clean air.
+\emph{Journal of Environmental Economics and Management} 5, (March 1978), 81--102.
+DOI:https://doi.org/\href{https://doi.org/10.1016/0095-0696(78)90006-2}{10.1016/0095-0696(78)90006-2}}
+
 \leavevmode\vadjust pre{\hypertarget{ref-ae_orig}{}}%
-\CSLLeftMargin{{[}7{]} }%
+\CSLLeftMargin{{[}8{]} }%
 \CSLRightInline{G.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-lstm_orig}{}}%
-\CSLLeftMargin{{[}8{]} }%
+\CSLLeftMargin{{[}9{]} }%
 \CSLRightInline{Sepp Hochreiter and Jürgen Schmidhuber.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Koza1993GeneticP}{}}%
-\CSLLeftMargin{{[}9{]} }%
+\CSLLeftMargin{{[}10{]} }%
 \CSLRightInline{John R. Koza. 1993. Genetic programming - on the programming of computers by means of natural selection. In \emph{Complex adaptive systems}.}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-survey_transfer_learning}{}}%
-\CSLLeftMargin{{[}10{]} }%
+\CSLLeftMargin{{[}11{]} }%
 \CSLRightInline{Sinno Jialin Pan and Qiang Yang.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-10.1007ux2f978-3-540-24621-3_22}{}}%
-\CSLLeftMargin{{[}11{]} }%
+\CSLLeftMargin{{[}12{]} }%
 \CSLRightInline{Grégory Paris, Denis Robilliard, and Cyril Fonlupt.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-harmless_overfitting_eda}{}}%
-\CSLLeftMargin{{[}12{]} }%
+\CSLLeftMargin{{[}13{]} }%
 \CSLRightInline{Malte Probst and Franz Rothlauf.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-design_of_modern_heuristics}{}}%
-\CSLLeftMargin{{[}13{]} }%
+\CSLLeftMargin{{[}14{]} }%
 \CSLRightInline{Franz Rothlauf. 2011. \emph{Design of modern heuristics: Principles and application}. DOI:https://doi.org/\href{https://doi.org/10.1007/978-3-540-72962-4}{10.1007/978-3-540-72962-4}}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-sampling_err_gp}{}}%
-\CSLLeftMargin{{[}14{]} }%
+\CSLLeftMargin{{[}15{]} }%
 \CSLRightInline{Dirk Schweim, David Wittenberg, and Franz Rothlauf.
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-ENERGY_DATASET}{}}%
+\CSLLeftMargin{{[}16{]} }%
+\CSLRightInline{Athanasios Tsanas and Angeliki Xifara.
+2012.
+Accurate quantitative estimation of energy performance of residential buildings using statistical machine learning tools.
+\emph{Energy and Buildings} 49, (June 2012), 560--567.
+DOI:https://doi.org/\href{https://doi.org/10.1016/j.enbuild.2012.03.003}{10.1016/j.enbuild.2012.03.003}}
+
 \leavevmode\vadjust pre{\hypertarget{ref-dae_orig2008}{}}%
-\CSLLeftMargin{{[}15{]} }%
+\CSLLeftMargin{{[}17{]} }%
 \CSLRightInline{Pascal Vincent, Hugo Larochelle, Y.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-weigend1994overfitting}{}}%
-\CSLLeftMargin{{[}16{]} }%
+\CSLLeftMargin{{[}18{]} }%
 \CSLRightInline{Andreas Weigend. 1994. On overfitting and the effective number of hidden units. In \emph{Proceedings of the 1993 connectionist models summer school}, 335--342.}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-daegp_explore_exploit}{}}%
-\CSLLeftMargin{{[}17{]} }%
+\CSLLeftMargin{{[}19{]} }%
 \CSLRightInline{David Wittenberg.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-dae-gp_2022_symreg}{}}%
-\CSLLeftMargin{{[}18{]} }%
+\CSLLeftMargin{{[}20{]} }%
 \CSLRightInline{David Wittenberg and Franz Rothlauf.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-dae-gp_2020_rtree}{}}%
-\CSLLeftMargin{{[}19{]} }%
+\CSLLeftMargin{{[}21{]} }%
 \CSLRightInline{David Wittenberg, Franz Rothlauf, and Dirk Schweim.
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-CONCRETE_DATASET}{}}%
+\CSLLeftMargin{{[}22{]} }%
+\CSLRightInline{I-Cheng Yeh.
+1998.
+Modeling of strength of high-performance concrete using artificial neural networks.'' cement and concrete research, 28(12), 1797-1808.
+\emph{Cement and Concrete Research} 28, (December 1998), 1797--1808.
+DOI:https://doi.org/\href{https://doi.org/10.1016/S0008-8846(98)00165-3}{10.1016/S0008-8846(98)00165-3}}
+
 \end{CSLReferences}

### 2023-02-04 data/summarize_data.py

@@ -48,15 +48,15 @@ RESULTS = {
     "airfoil_1hl" : "/Users/rmn/github/master_thesis/data/airfoil_1hl_150hn_fullRun_30gens",
     "airfoil_2hl" : "/Users/rmn/github/master_thesis/data/airfoil_2hl_150hn_fullRun_30gens",
     "bostonHousing": "/Users/rmn/github/master_thesis/data/bostonHousing_2hl_150hn_fullRun_30gens",
-    "energyCooling": "/Users/rmn/github/master_thesis/data/energyCooling_2hl_150hn_fullRun_30gens",
-    "concrete": "/Users/rmn/github/master_thesis/data/concrete_2hl_150hn_30gens"
+    "energyCooling": "/Users/rmn/github/master_thesis/data/energyCooling_2hl_150hn_FullRun_30gens",
+    "concrete": "/Users/rmn/github/master_thesis/data/concrete_2hl_150hn_FullRun_30gens"
 }
 
 RESULTS_2hl = {
     "Airfoil" :  "/Users/rmn/github/master_thesis/data/airfoil_2hl_150hn_fullRun_30gens",
     "BostonHousing": "/Users/rmn/github/master_thesis/data/bostonHousing_2hl_150hn_fullRun_30gens",
-    "energyCooling": "/Users/rmn/github/master_thesis/data/energyCooling_2hl_150hn_fullRun_30gens",
-    "Concrete": "/Users/rmn/github/master_thesis/data/concrete_2hl_150hn_30gens"
+    "energyCooling": "/Users/rmn/github/master_thesis/data/energyCooling_2hl_150hn_FullRun_30gens",
+    "Concrete": "/Users/rmn/github/master_thesis/data/concrete_2hl_150hn_FullRun_30gens"
 }
 
 

### 2023-02-04 data/vargha_delahni_A.py

@@ -1,45 +1,0 @@
-# code taken from https://gist.github.com/jacksonpradolima/f9b19d65b7f16603c837024d5f8c8a65
-
-import itertools as it
-
-from bisect import bisect_left
-from typing import List
-import scipy.stats as ss
-
-from pandas import Categorical
-
-
-def VD_A(treatment: List[float], control: List[float]):
-    """
-    Computes Vargha and Delaney A index
-    A. Vargha and H. D. Delaney.
-    A critique and improvement of the CL common language
-    effect size statistics of McGraw and Wong.
-    Journal of Educational and Behavioral Statistics, 25(2):101-132, 2000
-    The formula to compute A has been transformed to minimize accuracy errors
-    See: http://mtorchiano.wordpress.com/2014/05/19/effect-size-of-r-precision/
-    :param treatment: a numeric list
-    :param control: another numeric list
-    :returns the value estimate and the magnitude
-    """
-    m = len(treatment)
-    n = len(control)
-
-    if m != n:
-        raise ValueError("Data d and f must have the same length")
-
-    r = ss.rankdata(treatment + control)
-    r1 = sum(r[0:m])
-
-    # Compute the measure
-    # A = (r1/m - (m+1)/2)/n # formula (14) in Vargha and Delaney, 2000
-    A = (2 * r1 - m * (m + 1)) / (2 * n * m)  # equivalent formula to avoid accuracy errors
-
-    levels = [0.147, 0.33, 0.474]  # effect sizes from Hess and Kromrey, 2004
-    magnitude = ["negligible", "small", "medium", "large"]
-    scaled_A = (A - 0.5) * 2
-
-    magnitude = magnitude[bisect_left(levels, abs(scaled_A))]
-    estimate = A
-
-    return f"{estimate}({magnitude})"

### 2023-02-04 paper.Rmd
Sat Feb  4 19:00:06 CET 2023: update-remote.sh

@@
 The objective of the airfoil problem is to find a function that accurately predicts the output variable $y$ by taking in a subset of the input variables $x1,x2,x3,x4,x5$.
-The function set used by all DAE-GP variations for symbolic regression is summarized in \@ref(tab:airfoil-fullRun-2hl-maxIndSize-params), the terminal set consists of the the input variables $x1,x2,x3,x4,x5$ and the ephemeral integer constants listed in \@ref(tab:airfoil-fullRun-2hl-maxIndSize-params)[@dae-gp_2022_symreg].
+The function set used by all DAE-GP variations for symbolic regression is summarized in \@ref(tab:airfoil-fullRun-2hl-maxIndSize-params), the terminal set consists of the the input variables $x1,x2,x3,x4,x5$ and the ephemeral integer constants listed in table \@ref(tab:airfoil-fullRun-2hl-maxIndSize-params)[@dae-gp_2022_symreg].
 
@@
 ```
+To test for statistical differences between the distributions of results achieved by either pre-trained or regular DAE-GP, I selected the  
+nonparametric Mann-Whitney-U Test ([@mwu1], [@mwu2]) to test if the distribution .... Additionaly the effect strength of pre-training is computed by using Cliffs Delta[@Cliff1993DominanceSO].
 
-Table X summarizes the median best fitness achieved after running each algorithm for 10 runs using a population size of 500 individuals for 30 generations.^[P-Values are marked with asterisk (1,2,3) for ($\alpha$)-levels of $0.1$, $0.05$ and $0.01$].
-The median best fitness over all 30 generations are visualized in figure X.
+## Fitness
+
+The results for the median best fitness over the evolutionary search are visualized in figure \@ref(fig:symReg-Fitness).
+
+```{r symReg-Fitness, echo=FALSE, fig.cap="Fitness over 30 Generations - Summary - Real World Symbolic Regression", out.width="80%"}
+knitr::include_graphics("~/masterThesis/master_thesis/img/symbolicRegressionSummarized/median_fitness_byGens.png")
+```
+
+Table \@ref(tab:full-run-realWorldSymReg-fitness) summarizes the results for the median best fitness achieved after 30 generations and shows the results of computing the statistical tests.
+Higher fitness values (lower RMSE) are printed bold and p-values are marked with asterisk (1,2,3) for $\alpha$ levels ($0.1$, $0.05$, $0.01$).
 
@@
 df <- read.csv(
-  "/Users/rmn/github/master_thesis/data/summary_table_final_fit_mean.csv",
-  # "/Users/rmn/github/master_thesis/data/summary_table_final_fit_median.csv",  -> Use this path to summarize by median
+  #"/Users/rmn/github/master_thesis/data/summary_table_final_fit_mean.csv",
+  "/Users/rmn/github/master_thesis/data/summary_table_final_fit_median.csv",
   sep=",",
@@
  col.names = col_names,
- align = "c",
- caption = "Median Best Fitness - Symbolic Regression"
+ align = "l",
+ caption = "Median Best Fitness after 30 generations - Summary - Real World Symbolic Regression"
 ) %>% row_spec(0,bold=TRUE ) %>% kableExtra::kable_styling(latex_options = "hold_position")
@@
 
-![Median Best Fitness - Symbolic Regression](~/masterThesis/master_thesis/img/symbolicRegressionSummarized/median_fitness_byGens.png){height=80%}
+## Solution Size
 
-## Solution Size in Symbolic Regression
+Next I studied the size of solutions found during DAE-GP.
+Figure \@ref(fig:symReg-Size) shows the median size of the best solution per generation aggregated over all benchmark problems for symbolic regression.
 
-Next I studied the size of solutions based on using a pre-training strategy. Figure X shows the median size of the best solutions aggregated of all benchmark problems for symbolic regression
+```{r symReg-Size, echo=FALSE, fig.cap="Size of the best Solution over 30 Generations - Summary - Real World Symbolic Regression", out.width="80%"}
+knitr::include_graphics("~/masterThesis/master_thesis/img/symbolicRegressionSummarized/median_bestSol_size_byGens.png")
+```
 
-![Median Solution Size - Symbolic Regression](~/masterThesis/master_thesis/img/symbolicRegressionSummarized/median_bestSol_size_byGens.png){height=80%}
-
-Table X summarizes the results of the statistical tests.
+Table \@ref(tab:full-run-realWorldSymReg-best-solution-size) shows the median size of the best solution after 30 generations and results for computing the statistical tests.
 
@@
  align = "l",
- caption = "Median Solution Size - Symbolic Regression"
+ caption = "Median Size of the best solution after 30 generations - Summary - Real World Symbolic Regression"
 ) %>% row_spec(0,bold=FALSE) %>% kableExtra::kable_styling(latex_options = "hold_position", font_size = 7)
@@
  col.names = col_names,
- align = "c",
+ align = "l",
  caption = "Median Population Diversity over Generations - Symbolic Regression"

### 2023-02-04 paper.tex
Sat Feb  4 19:00:06 CET 2023: update-remote.sh

@@
 
-Denoising Autoencoder Genetic Programming (DAE-GP) is a novel variation of an genetic programming based Estimation of Distribution Algorithm (EDA-GP) that uses a denoising autoencoders long short-term memory network (DAE-LSTMs) as a probabilistic model to sample new candidate solutions {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{21}{]}.
+Denoising Autoencoder Genetic Programming (DAE-GP) is a novel variation of an genetic programming based Estimation of Distribution Algorithm (EDA-GP) that uses a denoising autoencoders long short-term memory network (DAE-LSTMs) as a probabilistic model to sample new candidate solutions {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{24}{]}.
 
@@
 In DAE-GP, DAE-LSTMs are trained with a subset of high-fitness solutions selected from a parent population with the aim to capture their promising qualities.
-The resulting model is then used to sample new offspring solutions by propagating partially mutated solutions from the parent population through the DAE-LSTMS {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{21}{]}.
-In previous work DAE-GP has been shown to outperform GP for both a generalized version of the royal tree problem {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{21}{]} as well as for a real-world symbolic regression problem {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{20}{]}.
-
-The DAE-GP algorithm first described by {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{21}{]} trains a DAE-LSTMs for each generation \(g\) of the search from scratch.
-{[}\protect\hyperlink{ref-dae-gp_2022_symreg}{20}{]} and {[}\protect\hyperlink{ref-daegp_explore_exploit}{19}{]} suggests the incorporation of a pre-training strategy into the evolutionary search as a possible way of improving the overall performance of the DAE-GP algorithm.
+The resulting model is then used to sample new offspring solutions by propagating partially mutated solutions from the parent population through the DAE-LSTMS {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{24}{]}.
+In previous work DAE-GP has been shown to outperform GP for both a generalized version of the royal tree problem {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{24}{]} as well as for a real-world symbolic regression problem {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{23}{]}.
+
+The DAE-GP algorithm first described by {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{24}{]} trains a DAE-LSTMs for each generation \(g\) of the search from scratch.
+{[}\protect\hyperlink{ref-dae-gp_2022_symreg}{23}{]} and {[}\protect\hyperlink{ref-daegp_explore_exploit}{22}{]} suggests the incorporation of a pre-training strategy into the evolutionary search as a possible way of improving the overall performance of the DAE-GP algorithm.
 The key idea is to pre-train an initial DAE-LSTM on a large population of candidate solutions and to use the pre-trained parameters of this initial model in each generation of the search as starting parameters for the current generation DAE-LSTM.
@@
 
-DAE-GP is an EDA-GP algorithm that uses DAE-LSTM networks as a probabilistic model to sample new offspring solutions {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{21}{]}.
+DAE-GP is an EDA-GP algorithm that uses DAE-LSTM networks as a probabilistic model to sample new offspring solutions {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{24}{]}.
 
@@
 Another important concept in EC is the distinction between genotypes and phenotypes of solutions, the genotype contains the information that is necessary to construct the phenotype, the outer appearance of a particular solution on which we measure the overall quality of solutions.
-The representation of a solution is therefore defined by the mapping of genotypes to phenotypes {[}\protect\hyperlink{ref-design_of_modern_heuristics}{14}{]}.
+The representation of a solution is therefore defined by the mapping of genotypes to phenotypes {[}\protect\hyperlink{ref-design_of_modern_heuristics}{16}{]}.
 Genetic operators, such as mutation or recombination are usually applied to the genotype of solutions.
@@
 
-GP follows the same basic evolutionary principle of EC but searches for more general, hierarchical computer programs of dynamically varying size and shape {[}\protect\hyperlink{ref-Koza1993GeneticP}{10}{]}.
-The computer programs that are at the center of the evolutionary search in GP are commonly represented by tree structures at the level of their phenotype {[}\protect\hyperlink{ref-design_of_modern_heuristics}{14}{]}.
-Since GP searches for high fitness computer programs that produce a desired output for some input, it can be applied to various different problem domains such as symbolic regression, automatic programming, or evolving game-playing strategies {[}\protect\hyperlink{ref-Koza1993GeneticP}{10}{]}.
+GP follows the same basic evolutionary principle of EC but searches for more general, hierarchical computer programs of dynamically varying size and shape {[}\protect\hyperlink{ref-Koza1993GeneticP}{11}{]}.
+The computer programs that are at the center of the evolutionary search in GP are commonly represented by tree structures at the level of their phenotype {[}\protect\hyperlink{ref-design_of_modern_heuristics}{16}{]}.
+Since GP searches for high fitness computer programs that produce a desired output for some input, it can be applied to various different problem domains such as symbolic regression, automatic programming, or evolving game-playing strategies {[}\protect\hyperlink{ref-Koza1993GeneticP}{11}{]}.
 An important quality of GP is the ability to search for solutions of variable length and structure.
 GP is an especially useful meta heuristic for problems where no a priori knowledge about the final form of good solutions is available.
-GPs ability to optimize solutions for their structure as well as for their parameters led to it being one of the most prevalent methods used for symbolic regression {[}\protect\hyperlink{ref-10.1007ux2f978-3-540-24621-3_22}{12}{]}.
+GPs ability to optimize solutions for their structure as well as for their parameters led to it being one of the most prevalent methods used for symbolic regression {[}\protect\hyperlink{ref-10.1007ux2f978-3-540-24621-3_22}{14}{]}.
 
@@
 
-The aim of Estimation of distribution algorithms (EDA) is to replace the standard variation operators used in GA by building probabilistic models that can capture complex dependencies between different decision variables of an optimization problem {[}\protect\hyperlink{ref-design_of_modern_heuristics}{14}{]}.
+The aim of Estimation of distribution algorithms (EDA) is to replace the standard variation operators used in GA by building probabilistic models that can capture complex dependencies between different decision variables of an optimization problem {[}\protect\hyperlink{ref-design_of_modern_heuristics}{16}{]}.
 EDAs use this probabilistic model to sample new offspring solutions inside an evolutionary search to replace crossover and/or mutation operators.
@@
 
-One possible way of model building in EDA proposed by {[}\protect\hyperlink{ref-harmless_overfitting_eda}{13}{]} is to use denoising autoencoders (DAEs) as generative models to capture complex probability distributions of decision variables.
+One possible way of model building in EDA proposed by {[}\protect\hyperlink{ref-harmless_overfitting_eda}{15}{]} is to use denoising autoencoders (DAEs) as generative models to capture complex probability distributions of decision variables.
 DAE, a variation of the autoencoders (AE), are a widely used type of neural networks in the domain of unsupervised machine learning that maps \(n\) input variables to \(n\) output variables using a hidden representation.
 
-AE were introduced by {[}\protect\hyperlink{ref-ae_orig}{8}{]} to compress high-dimensional data into lower-dimensions. An AE consists of two different sub-units:
+AE were introduced by {[}\protect\hyperlink{ref-ae_orig}{9}{]} to compress high-dimensional data into lower-dimensions. An AE consists of two different sub-units:
 
@@
 
-The AE is trained to reduce the reconstruction error between input and output data, after the training procedure is finished the network is able to reduce the dimensional of input data to a compressed representation{[}\protect\hyperlink{ref-ae_orig}{8}{]}.
-
-DAE was first introduced by {[}\protect\hyperlink{ref-dae_orig2008}{17}{]} as an improved AE with the ability to learn new representations of data that is especially robust to partially corrupted input data.
+The AE is trained to reduce the reconstruction error between input and output data, after the training procedure is finished the network is able to reduce the dimensional of input data to a compressed representation{[}\protect\hyperlink{ref-ae_orig}{9}{]}.
+
+DAE was first introduced by {[}\protect\hyperlink{ref-dae_orig2008}{19}{]} as an improved AE with the ability to learn new representations of data that is especially robust to partially corrupted input data.
 DAE modifies the AE by using partially corrupted input data for the AE and training it to reconstruct the uncorrupted, original version of the input data.
 
-Since the hidden representation of DAE captures the dependency structure of the input data it can therefore also be used to generate new solutions in the context of GAs {[}\protect\hyperlink{ref-harmless_overfitting_eda}{13}{]}.
-
-DAE-GP builds upon the concept of using DAEs in EDAs described by {[}\protect\hyperlink{ref-harmless_overfitting_eda}{13}{]} and transfers the concept to the domain of GP.
-The mutation and crossover operators of standard GP are replaced by sampling new solutions from a probabilistic model that is build by training a DAEs long short-term memory (LSTM) network on a subset of high fitness solutions from the current population {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{21}{]}.
-
-LSTMs are a variant of artificial neural networks first introduced by {[}\protect\hyperlink{ref-lstm_orig}{9}{]} that can store learned information over en extended time period while avoiding the problem of vanishing and/or exploding gradients.
-Since DAE-GP encodes candidate solutions as linear strings in prefix notation, the DAE in DAE-GP uses LSTMs for both encoding and decoding where the total amount of time steps \(T\) is equal to sum of the length of the input solution and the output solution {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{21}{]}.
+Since the hidden representation of DAE captures the dependency structure of the input data it can therefore also be used to generate new solutions in the context of GAs {[}\protect\hyperlink{ref-harmless_overfitting_eda}{15}{]}.
+
+DAE-GP builds upon the concept of using DAEs in EDAs described by {[}\protect\hyperlink{ref-harmless_overfitting_eda}{15}{]} and transfers the concept to the domain of GP.
+The mutation and crossover operators of standard GP are replaced by sampling new solutions from a probabilistic model that is build by training a DAEs long short-term memory (LSTM) network on a subset of high fitness solutions from the current population {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{24}{]}.
+
+LSTMs are a variant of artificial neural networks first introduced by {[}\protect\hyperlink{ref-lstm_orig}{10}{]} that can store learned information over en extended time period while avoiding the problem of vanishing and/or exploding gradients.
+Since DAE-GP encodes candidate solutions as linear strings in prefix notation, the DAE in DAE-GP uses LSTMs for both encoding and decoding where the total amount of time steps \(T\) is equal to sum of the length of the input solution and the output solution {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{24}{]}.
 
@@
 Pre-Training describes the concept of initially training an artificial neural network on a large dataset before using it to solve a more specific task.
-It is a commonly used strategy in deep architectures that has been shown to improve both the optimization process itself as well as the generalization behavior if compared to the standard approach of using randomly initialized parameters {[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{4}{]}.
-
-The strategy of pre-training artificial neural networks is based on the idea of transfer learning that aims at retaining previously learned knowledge from one task to re-use it for another task {[}\protect\hyperlink{ref-HAN2021225}{6}{]} {[}\protect\hyperlink{ref-survey_transfer_learning}{11}{]}.
+It is a commonly used strategy in deep architectures that has been shown to improve both the optimization process itself as well as the generalization behavior if compared to the standard approach of using randomly initialized parameters {[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{5}{]}.
+
+The strategy of pre-training artificial neural networks is based on the idea of transfer learning that aims at retaining previously learned knowledge from one task to re-use it for another task {[}\protect\hyperlink{ref-HAN2021225}{7}{]} {[}\protect\hyperlink{ref-survey_transfer_learning}{13}{]}.
 Transfer learning traditionally follows a two phase approach:
@@
 
-where source task and target task are usually similar but may differ in their feature space and the distribution of training data {[}\protect\hyperlink{ref-survey_transfer_learning}{11}{]}.
+where source task and target task are usually similar but may differ in their feature space and the distribution of training data {[}\protect\hyperlink{ref-survey_transfer_learning}{13}{]}.
 
@@
 Instead of initializing and optimizing each model from scratch as done in previous work (e.g.
-{[}\protect\hyperlink{ref-dae-gp_2020_rtree}{21}{]} {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{20}{]}), a separate DAE-LSTM network \(\hat{M}\) will be trained on a large population of randomly initialized solutions.\\
+{[}\protect\hyperlink{ref-dae-gp_2020_rtree}{24}{]} {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{23}{]}), a separate DAE-LSTM network \(\hat{M}\) will be trained on a large population of randomly initialized solutions.\\
 The trainable parameters \(\theta_{\hat{M}}\) obtained after finishing the training procedure of \(\hat{M}\) are then used as the starting parameters for each following DAE-LSTM model \(M_g\) for \(g\in{1,...,g_{max}}\).
@@
 
-Early research on pre-training for DAE by {[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{4}{]} comes to the conclusion that besides adding robustness to models the strategy also results in improved generalization and better performing models.
+Early research on pre-training for DAE by {[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{5}{]} comes to the conclusion that besides adding robustness to models the strategy also results in improved generalization and better performing models.
 The authors describe that unsupervised pre-training behaves like a regularizer on DAE networks, the mean test error and it's variance are reduced for DAE networks that are initialized from pre-trained weights if compared to the same architectures that use random initialization.
 
-Another important finding by {[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{4}{]} is that the positive effect of pre-training is dependent on both the depth of the network as well as the size of layers - while increasingly larger networks benefit increasingly more from pre-training the final performance of small architectures tends to be worse with pre-training than with randomized weight initialization.
+Another important finding by {[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{5}{]} is that the positive effect of pre-training is dependent on both the depth of the network as well as the size of layers - while increasingly larger networks benefit increasingly more from pre-training the final performance of small architectures tends to be worse with pre-training than with randomized weight initialization.
 The authors find evidence that pre-training DAE is especially useful for optimizing the parameters of the lower layers of the network.
@@
 Another motivation for introducing pre-training into DAE-GP is the prevalence of sampling error for small GP populations sizes.
-{[}\protect\hyperlink{ref-sampling_err_gp}{15}{]} finds that sampling error, non-systematic errors that are caused by observing only a small subset of the statistical population, is a severe problem in the domain of GP since the initial population may not be a representative sample of all possible solutions.
-{[}\protect\hyperlink{ref-sampling_err_gp}{15}{]} also introduces a method for calculating optimal population sizes to minimize the presence of sampling error.
+{[}\protect\hyperlink{ref-sampling_err_gp}{17}{]} finds that sampling error, non-systematic errors that are caused by observing only a small subset of the statistical population, is a severe problem in the domain of GP since the initial population may not be a representative sample of all possible solutions.
+{[}\protect\hyperlink{ref-sampling_err_gp}{17}{]} also introduces a method for calculating optimal population sizes to minimize the presence of sampling error.
 Since DAE-LSTM of DAE-GP learns the properties of it's training population and reuses the acquired knowledge in the sampling procedure, using the parameters obtained from training \(\hat{M}\) on a sufficiently large training population might reduce the need for large population sizes in the following generations of the search if \(\hat{M}\) already implicitly captured the properties of a representative sample of solutions.
@@
 
-The DAE-GP algorithm, first described by {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{21}{]}, is summarized as a flowchart in figure \ref{fig:daegp-flowchart}.
+The DAE-GP algorithm, first described by {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{24}{]}, is summarized as a flowchart in figure \ref{fig:daegp-flowchart}.
 After setting the generation counter \(g\) to \(0\) and creating the initial population of solutions \(P_0\), DAE-GP enters the main loop that checks if a termination criteria is satisfied (i.e.~maximum number of generations or fitness evaluations is reached).
@@
 One difficulty in the implementation of a pre-training strategy into DAE-GP has been the determination of the number of hidden neurons in side the DAE-LSTMs hidden layers.
-The original description of DAE-GP {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{21}{]} uses a strategy where the number of hidden neurons for each hidden layer is dynamically set per generation to the maximum individual size inside the current population.
+The original description of DAE-GP {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{24}{]} uses a strategy where the number of hidden neurons for each hidden layer is dynamically set per generation to the maximum individual size inside the current population.
 For a pre-training implementation, this strategy can not be easily adapted since it leads to a changing number of neurons at each generation resulting in different dimensions of the DAE-LSTM.
@@
 
-The framework that was used to conduct all experiments of this thesis was provided by the supervisor of this thesis David Wittenberg and uses the python programming language in conjunction with the baseline libraries \texttt{Keras} {[}\protect\hyperlink{ref-chollet2015keras}{2}{]} for deep neural networks and \texttt{deap} {[}\protect\hyperlink{ref-DEAP_JMLR2012}{5}{]} for evolutionary computation.
+The framework that was used to conduct all experiments of this thesis was provided by the supervisor of this thesis David Wittenberg and uses the python programming language in conjunction with the baseline libraries \texttt{Keras} {[}\protect\hyperlink{ref-chollet2015keras}{2}{]} for deep neural networks and \texttt{deap} {[}\protect\hyperlink{ref-DEAP_JMLR2012}{6}{]} for evolutionary computation.
 
@@
 To test pre-training in DAE-GP this thesis focuses on the domain of real-world symbolic regression problems.
-Symbolic Regression problems have been one of the first GP applications {[}\protect\hyperlink{ref-Koza1993GeneticP}{10}{]} and are an actively studied and highly relevant research area.
-The goal in symbolic regression is to find a mathematical model for a given set of data points {[}\protect\hyperlink{ref-10.1007ux2f978-3-540-24621-3_22}{12}{]}, in real-world symbolic regression these data points are sourced from real-world observations which in contrast to synthetic symbolic regression problem are more likely to contain random noise and bias.
+Symbolic Regression problems have been one of the first GP applications {[}\protect\hyperlink{ref-Koza1993GeneticP}{11}{]} and are an actively studied and highly relevant research area.
+The goal in symbolic regression is to find a mathematical model for a given set of data points {[}\protect\hyperlink{ref-10.1007ux2f978-3-540-24621-3_22}{14}{]}, in real-world symbolic regression these data points are sourced from real-world observations which in contrast to synthetic symbolic regression problem are more likely to contain random noise and bias.
 Another important challenge in solving real-world symbolic regression problems is the ability for a given model to generalize, we want the final model to show high accuracy in predicting outcomes for previously unseen cases.
@@
 The objective of the airfoil problem is to find a function that accurately predicts the output variable \(y\) by taking in a subset of the input variables \(x1,x2,x3,x4,x5\).
-The function set used by all DAE-GP variations for symbolic regression is summarized in \ref{tab:airfoil-fullRun-2hl-maxIndSize-params}, the terminal set consists of the the input variables \(x1,x2,x3,x4,x5\) and the ephemeral integer constants listed in \ref{tab:airfoil-fullRun-2hl-maxIndSize-params}{[}\protect\hyperlink{ref-dae-gp_2022_symreg}{20}{]}.
+The function set used by all DAE-GP variations for symbolic regression is summarized in \ref{tab:airfoil-fullRun-2hl-maxIndSize-params}, the terminal set consists of the the input variables \(x1,x2,x3,x4,x5\) and the ephemeral integer constants listed in table \ref{tab:airfoil-fullRun-2hl-maxIndSize-params}{[}\protect\hyperlink{ref-dae-gp_2022_symreg}{23}{]}.
 
@@
 For each configuration I tested traditional DAE-GP as well as a pre-trained DAE-GP resulting in 12 total sub experiments that were each based on 10 individual runs (total number of runs=\(120\)).
-To avoid creating biased results through the presence of sampling error inside the pre-training population (see {[}\protect\hyperlink{ref-sampling_err_gp}{15}{]}), the population size for the pre-training phase is chosen very high with the size of \(\hat{P}\) =10000 where 50\% of \(\hat{P}\) is used for the training population \(\hat{P}_{train}\) and the remaining 50\% are used for the hold-out validation population \(\hat{P}_{test}\).
+To avoid creating biased results through the presence of sampling error inside the pre-training population (see {[}\protect\hyperlink{ref-sampling_err_gp}{17}{]}), the population size for the pre-training phase is chosen very high with the size of \(\hat{P}\) =10000 where 50\% of \(\hat{P}\) is used for the training population \(\hat{P}_{train}\) and the remaining 50\% are used for the hold-out validation population \(\hat{P}_{test}\).
 The training of DAE-LSTM uses a fixed number of 1000 epochs until termination and uses Adam optimization for gradient descent.
@@
 In general I expect that with a growing number of trainable parameters (either by adding more hidden layers or more hidden neurons per layer) the DAE-LSTM will be more prone to overfit to the training population \(P_{train}\) resulting in a small reconstruction error for the training population and a large one for the validation population \(P_{test}\).
-The reason for this effect is that a larger network trained over an extended period of time (without the use of strategies like early stopping), has much more potential to learn noise from the training dataset than a smaller network, which is more likely to result in worsening performance for previously unseen cases {[}\protect\hyperlink{ref-weigend1994overfitting}{18}{]}.
-
-Based on the review of {[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{4}{]} I also expect that pre-training will have an insignificant or even negative influence on small DAE-LSTM instances while improving the networks generalization ability with growing size.
+The reason for this effect is that a larger network trained over an extended period of time (without the use of strategies like early stopping), has much more potential to learn noise from the training dataset than a smaller network, which is more likely to result in worsening performance for previously unseen cases {[}\protect\hyperlink{ref-weigend1994overfitting}{20}{]}.
+
+Based on the review of {[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{5}{]} I also expect that pre-training will have an insignificant or even negative influence on small DAE-LSTM instances while improving the networks generalization ability with growing size.
 
@@
 
-Since previous work by {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{20}{]} demonstrated that DAE-GP, for a given number of 10.000 fitness evaluations, is able to produce solutions that are much smaller in tree size than regular GP, I also studied the influence of pre-training on the size of solutions found by DAE-GP.
+Since previous work by {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{23}{]} demonstrated that DAE-GP, for a given number of 10.000 fitness evaluations, is able to produce solutions that are much smaller in tree size than regular GP, I also studied the influence of pre-training on the size of solutions found by DAE-GP.
 Figure \ref{fig:airfoil-size-combined-30gens} shows the median of both the average size of individuals inside the population as well as the size of the currently best performing solution inside the population.
@@
 To ensure a fair comparison between pre-trained and regular DAE-GP, the prior experiments used a static number of hidden neurons for each hidden Layer inside the DAE-LSTM networks.
-For comparison, figure \ref{fig:airfoil-maxIndSize-params} shows the total number of trainable parameters for each generation's DAE-LSTM of another experiment where regular DAE-GP was tested with dynamic adjustment of the number of hidden neurons as described in {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{21}{]} and {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{20}{]}.
+For comparison, figure \ref{fig:airfoil-maxIndSize-params} shows the total number of trainable parameters for each generation's DAE-LSTM of another experiment where regular DAE-GP was tested with dynamic adjustment of the number of hidden neurons as described in {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{24}{]} and {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{23}{]}.
 Again, the experiments was based on \(10\) individual runs and otherwise used the same hyperparameters as listed in table \ref{tab:airfoil-fullRun-2hl-maxIndSize-params}.
@@
 
-Since previous research on DAE-GP used a single hidden layers for the airfoil dataset {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{20}{]}, I repeated the experiment with a single hidden layer.
-Based on the findings of {[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{4}{]}, I expected that pre-training DAE-GP would lead to a decrease in overall performance.
+Since previous research on DAE-GP used a single hidden layers for the airfoil dataset {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{23}{]}, I repeated the experiment with a single hidden layer.
+Based on the findings of {[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{5}{]}, I expected that pre-training DAE-GP would lead to a decrease in overall performance.
 Figures \ref{fig:airfoil-fitness-perGen} and \ref{fig:airfoil-finalFitness} shows the mean and median best fitness per generation as well as the distribution of final fitness scores after 30 generations.
@@
 
-The pre-training implementation in previous experiments used ramped half and half initialization {[}\protect\hyperlink{ref-Koza1993GeneticP}{10}{]} to create \(\hat{P}\), this leads to 50\% of the population consisting of full grown trees.
+The pre-training implementation in previous experiments used ramped half and half initialization {[}\protect\hyperlink{ref-Koza1993GeneticP}{11}{]} to create \(\hat{P}\), this leads to 50\% of the population consisting of full grown trees.
 An alternative method of initializing \(\hat{P}\) is to use the grow initialization method where solutions are not guaranteed to be full trees.
@@
 Airfoil & 1503 & 5 & {[}\protect\hyperlink{ref-AIRFOIL_DATASET}{1}{]} \\
-Boston\_Housing & 506 & 13 & {[}\protect\hyperlink{ref-BOSTON_HOUSING_DATASET}{7}{]} \\
-Energy\_Cooling & 768 & 8 & {[}\protect\hyperlink{ref-ENERGY_DATASET}{16}{]} \\
-Concrete & 1030 & 8 & {[}\protect\hyperlink{ref-CONCRETE_DATASET}{22}{]} \\
+Boston\_Housing & 506 & 13 & {[}\protect\hyperlink{ref-BOSTON_HOUSING_DATASET}{8}{]} \\
+Energy\_Cooling & 768 & 8 & {[}\protect\hyperlink{ref-ENERGY_DATASET}{18}{]} \\
+Concrete & 1030 & 8 & {[}\protect\hyperlink{ref-CONCRETE_DATASET}{25}{]} \\
 \end{longtable}
 
-Table X summarizes the median best fitness achieved after running each algorithm for 10 runs using a population size of 500 individuals for 30 generations.\footnote{P-Values are marked with asterisk (1,2,3) for (\(\alpha\))-levels of \(0.1\), \(0.05\) and \(0.01\)}.
-The median best fitness over all 30 generations are visualized in figure X.
+\hypertarget{fitness}{%
+\subsection{Fitness}\label{fitness}}
+
+The results for the median best fitness over the evolutionary search are visualized in figure \ref{fig:symReg-Fitness}.
+
+\begin{figure}[c]
+
+{\centering \includegraphics[width=0.8\linewidth]{img/symbolicRegressionSummarized/median_fitness_byGens}
+
+}
+
+\caption{Fitness over 30 Generations - Summary - Real World Symbolic Regression}\label{fig:symReg-Fitness}
+\end{figure}
+
+Table \ref{tab:full-run-realWorldSymReg-fitness} summarizes the results for the median best fitness achieved after 30 generations and shows the results of computing both the Mann-Whitney-U Test ({[}\protect\hyperlink{ref-mwu1}{21}{]}, {[}\protect\hyperlink{ref-mwu2}{12}{]}) as well as the effect strength measured by Cliffs Delta{[}\protect\hyperlink{ref-Cliff1993DominanceSO}{3}{]} over the distribution of final fitness scores.
+Higher fitness values (lower RMSE) are printed bold and p-values are marked with asterisk (1,2,3) for \(\alpha\) levels (\(0.1\), \(0.05\), \(0.01\)).
 
 \begin{longtable}[]{@{}
-  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1932}}
-  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1705}}
-  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.0795}}
-  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1477}}
-  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1477}}
-  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1023}}
-  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1591}}@{}}
-\caption{\label{tab:full-run-realWorldSymReg-fitness}Median Best Fitness - Symbolic Regression}\tabularnewline
-\toprule\noalign{}
-\begin{minipage}[b]{\linewidth}\centering
-Problem
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
-Hidden-Layers
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
-Set
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
-DAE-GP
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
-Pre-Trained
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
-P-Value
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
-Cliffs-Delta
-\end{minipage} \\
-\midrule\noalign{}
-\endfirsthead
-\toprule\noalign{}
-\begin{minipage}[b]{\linewidth}\centering
-Problem
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
-Hidden-Layers
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
-Set
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
-DAE-GP
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
-Pre-Trained
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
-P-Value
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
-Cliffs-Delta
-\end{minipage} \\
-\midrule\noalign{}
-\endhead
-\bottomrule\noalign{}
-\endlastfoot
-Airfoil & 1 & Train & \textbf{33.3674} & 35.3454 & 0.02** & 0.64 \\
-& 1 & Test & \textbf{33.7271} & 35.612 & 0.14 & 0.40 \\
-Airfoil & 2 & Train & 17.4762 & \textbf{14.7233} & 0.31 & -0.28 \\
-& 2 & Test & 17.5299 & \textbf{14.7603} & 0.57 & -0.16 \\
-Boston\_Housing & 2 & Train & 8.1922 & \textbf{8.0763} & 0.29 & -0.29 \\
-& 2 & Test & 8.0501 & \textbf{7.9692} & 0.71 & -0.11 \\
-Energy(Cooling) & 2 & Train & \textbf{4.5271} & 4.7782 & 0.02** & 0.63 \\
-& 2 & Test & \textbf{4.67} & 4.9196 & 0.29 & 0.29 \\
-Concrete & 2 & Train & 17.0052 & \textbf{16.9085} & 0.97 & 0.02 \\
-& 2 & Test & \textbf{16.5845} & 16.9298 & 0.52 & 0.18 \\
-\end{longtable}
-
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth,height=0.8\textheight]{~/masterThesis/master_thesis/img/symbolicRegressionSummarized/median_fitness_byGens.png}
-\caption{Median Best Fitness - Symbolic Regression}
-\end{figure}
-
-\hypertarget{solution-size-in-symbolic-regression}{%
-\subsection{Solution Size in Symbolic Regression}\label{solution-size-in-symbolic-regression}}
-
-Next I studied the size of solutions based on using a pre-training strategy. Figure X shows the median size of the best solutions aggregated of all benchmark problems for symbolic regression
-
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth,height=0.8\textheight]{~/masterThesis/master_thesis/img/symbolicRegressionSummarized/median_bestSol_size_byGens.png}
-\caption{Median Solution Size - Symbolic Regression}
-\end{figure}
-
-Table X summarizes the results of the statistical tests.
-
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2254}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1549}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1549}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1690}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1127}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1831}}@{}}
-\caption{\label{tab:full-run-realWorldSymReg-best-solution-size}Median Solution Size - Symbolic Regression}\tabularnewline
+  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1975}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1728}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.0741}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1481}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1481}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.0988}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1605}}@{}}
+\caption{\label{tab:full-run-realWorldSymReg-fitness}Median Best Fitness after 30 generations - Summary - Real World Symbolic Regression}\tabularnewline
 \toprule\noalign{}
@@
 \end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Hid.Layers
+Hidden-Layers
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Set
 \end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
@@
 \endlastfoot
-Airfoil & 1 & 12 & \textbf{10.6} & 0.46 & -0.20 \\
-Airfoil & 2 & 17.4762 & \textbf{14.7233} & 0.31 & -0.28 \\
-Boston\_Housing & 2 & 8.1922 & \textbf{8.0763} & 0.29 & -0.29 \\
-Energy(Cooling) & 2 & \textbf{4.5271} & 4.7782 & 0.02** & 0.63 \\
-Concrete & 2 & 17.0052 & \textbf{16.9085} & 0.97 & 0.02 \\
+Airfoil & 1 & Train & \textbf{12.6611} & 16.2778 & 0.57 & 0.16 \\
+& 1 & Test & \textbf{12.5944} & 16.2131 & 0.62 & 0.14 \\
+Airfoil & 2 & Train & \textbf{7.0644} & 7.7109 & 0.62 & 0.14 \\
+& 2 & Test & \textbf{6.991} & 7.6799 & 0.91 & 0.04 \\
+Boston\_Housing & 2 & Train & \textbf{8.0217} & 8.0543 & 0.85 & 0.06 \\
+& 2 & Test & \textbf{7.9472} & 7.9683 & 0.79 & 0.08 \\
+Energy(Cooling) & 2 & Train & \textbf{4.449} & 4.5217 & 1.00 & -0.03 \\
+& 2 & Test & 4.6258 & \textbf{4.5708} & 1.00 & -0.03 \\
+Concrete & 2 & Train & \textbf{17.1299} & 17.9175 & 0.51 & 0.24 \\
+& 2 & Test & \textbf{17.1585} & 17.6584 & 0.68 & 0.16 \\
+\end{longtable}
+
+\hypertarget{solution-size}{%
+\subsection{Solution Size}\label{solution-size}}
+
+Next I studied the size of solutions found during DAE-GP.
+Figure \ref{fig:symReg-Size} shows the median size of the best solution per generation aggregated over all benchmark problems for symbolic regression.
+
+\begin{figure}[c]
+
+{\centering \includegraphics[width=0.8\linewidth]{img/symbolicRegressionSummarized/median_bestSol_size_byGens}
+
+}
+
+\caption{Size of the best Solution over 30 Generations - Summary - Real World Symbolic Regression}\label{fig:symReg-Size}
+\end{figure}
+
+Table \ref{tab:full-run-realWorldSymReg-best-solution-size} shows the median size of the best solution after 30 generations and results for computing the statistical tests.
+
+\begin{longtable}[]{@{}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2353}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1618}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1176}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1765}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1176}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1912}}@{}}
+\caption{\label{tab:full-run-realWorldSymReg-best-solution-size}Median Size of the best solution after 30 generations - Summary - Real World Symbolic Regression}\tabularnewline
+\toprule\noalign{}
+\begin{minipage}[b]{\linewidth}\raggedright
+Problem
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Hid.Layers
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+DAE-GP
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Pre-Trained
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+P-Value
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Cliffs-Delta
+\end{minipage} \\
+\midrule\noalign{}
+\endfirsthead
+\toprule\noalign{}
+\begin{minipage}[b]{\linewidth}\raggedright
+Problem
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Hid.Layers
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+DAE-GP
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Pre-Trained
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+P-Value
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Cliffs-Delta
+\end{minipage} \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+Airfoil & 1 & \textbf{7.8} & 8.4 & 0.44 & 0.20 \\
+Airfoil & 2 & 13.2 & \textbf{8.6} & 0.97 & -0.02 \\
+Boston\_Housing & 2 & 4.6 & \textbf{3.8} & 0.37 & -0.20 \\
+Energy(Cooling) & 2 & 7.8 & \textbf{3} & 0.18 & -0.40 \\
+Concrete & 2 & 9.2 & \textbf{5} & 0.09* & -0.56 \\
 \end{longtable}
@@
 \endlastfoot
-Airfoil & 1 & \textbf{0.34} & 0.49 & 0.02** & 0.36 \\
-Airfoil & 2 & \textbf{0.91} & 0.93 & 0.00*** & 0.88 \\
-Boston\_Housing & 2 & \textbf{0.94} & 0.95 & 0.00*** & 0.82 \\
-Energy(Cooling) & 2 & \textbf{0.93} & 0.94 & 0.00*** & 0.80 \\
-Concrete & 2 & \textbf{0.93} & 0.94 & 0.00*** & 0.88 \\
+Airfoil & 1 & \textbf{0.86} & 0.93 & 0.00*** & 0.93 \\
+Airfoil & 1 & \textbf{0.86} & 0.93 & 0.00*** & 0.93 \\
+Boston\_Housing & 1 & \textbf{0.88} & 0.95 & 0.00*** & 0.92 \\
+Energy(Cooling) & 1 & \textbf{0.87} & 0.94 & 0.00*** & 0.94 \\
+Concrete & 1 & \textbf{0.87} & 0.94 & 0.00*** & 0.94 \\
 \end{longtable}
@@
 \begin{longtable}[]{@{}
-  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2222}}
-  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}
-  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1111}}
-  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1806}}
-  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1250}}
-  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1944}}@{}}
+  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2329}}
+  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1644}}
+  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1096}}
+  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1781}}
+  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1233}}
+  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1918}}@{}}
 \caption{\label{tab:full-run-realWorldSymReg-epochsPerGen}Median Number of Training Epochs per Generation - Symbolic Regression}\tabularnewline
@@
 \endlastfoot
-Airfoil & 1 & 190.93 & \textbf{60} & 0.00*** & -0.94 \\
-Airfoil & 2 & 151.87 & \textbf{64.95} & 0.00*** & -0.88 \\
-Boston\_Housing & 2 & 182.84 & \textbf{67.27} & 0.00*** & -0.88 \\
-Energy\_Cooling & 2 & 169.39 & \textbf{69.87} & 0.00*** & -0.90 \\
-concrete & 2 & 171.69 & \textbf{64.9} & 0.00*** & -0.92 \\
+Airfoil & 1 & 143.49 & \textbf{64.66} & 0.00*** & -0.88 \\
+Airfoil & 2 & 143.83 & \textbf{61.22} & 0.00*** & -0.88 \\
+Boston\_Housing & 2 & 141.8 & \textbf{65.36} & 0.00*** & -0.87 \\
+Energy(Cooling) & 2 & 144.24 & \textbf{63.5} & 0.00*** & -0.93 \\
+Concrete & 2 & 139.2 & \textbf{64.21} & 0.00*** & -0.87 \\
 \end{longtable}
@@
 Another disadvantage of using pre-training in DAE-GP is that, in my experiments, it required DAE-LSTM networks to use two hidden layers (see subsection \protect\hyperlink{reduced-number-of-hidden-layers}{Reduced Number of Hidden Layers}).
-For a single hidden layer, pre-training did show a negative impact on the median final fitness of solutions found as it was suggested by {[}\protect\hyperlink{ref-pmlr-v9-erhan10a}{3}{]}.
-Since current research on DAE-GP for real-world symbolic regression relies on a single hidden layer (see {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{20}{]}), pre-training does come with additional computational expenses for using deeper DAE-LSTMs.
+For a single hidden layer, pre-training did show a negative impact on the median final fitness of solutions found as it was suggested by {[}\protect\hyperlink{ref-pmlr-v9-erhan10a}{4}{]}.
+Since current research on DAE-GP for real-world symbolic regression relies on a single hidden layer (see {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{23}{]}), pre-training does come with additional computational expenses for using deeper DAE-LSTMs.
 
@@
 
-Another possibility to improve the benefits of pre-training for DAE-GP is based on the findings of {[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{4}{]}.
+Another possibility to improve the benefits of pre-training for DAE-GP is based on the findings of {[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{5}{]}.
 Pre-Training has shown to be especially useful for the lower layers of DAEs so it could be worthwhile to further experiment with only adapting the weights for a subset of the hidden layers inside the DAE-LSTM networks to either reduce run-time or to improve on the quality of solutions found during the evolutionary search.
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Cliff1993DominanceSO}{}}%
+\CSLLeftMargin{{[}3{]} }%
+\CSLRightInline{Norman Cliff. 1993. Dominance statistics: Ordinal analyses to answer ordinal questions. \emph{Psychological Bulletin} 114, (1993), 494--509.}
+
 \leavevmode\vadjust pre{\hypertarget{ref-pmlr-v9-erhan10a}{}}%
-\CSLLeftMargin{{[}3{]} }%
+\CSLLeftMargin{{[}4{]} }%
 \CSLRightInline{Dumitru Erhan, Aaron Courville, Yoshua Bengio, and Pascal Vincent.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-pmlr-v5-erhan09a}{}}%
-\CSLLeftMargin{{[}4{]} }%
+\CSLLeftMargin{{[}5{]} }%
 \CSLRightInline{Dumitru Erhan, Pierre-Antoine Manzagol, Yoshua Bengio, Samy Bengio, and Pascal Vincent.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-DEAP_JMLR2012}{}}%
-\CSLLeftMargin{{[}5{]} }%
+\CSLLeftMargin{{[}6{]} }%
 \CSLRightInline{Félix-Antoine Fortin, François-Michel De Rainville, Marc-André Gardner, Marc Parizeau, and Christian Gagné.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-HAN2021225}{}}%
-\CSLLeftMargin{{[}6{]} }%
+\CSLLeftMargin{{[}7{]} }%
 \CSLRightInline{Xu Han, Zhengyan Zhang, Ning Ding, Yuxian Gu, Xiao Liu, Yuqi Huo, Jiezhong Qiu, Yuan Yao, Ao Zhang, Liang Zhang, Wentao Han, Minlie Huang, Qin Jin, Yanyan Lan, Yang Liu, Zhiyuan Liu, Zhiwu Lu, Xipeng Qiu, Ruihua Song, Jie Tang, Ji-Rong Wen, Jinhui Yuan, Wayne Xin Zhao, and Jun Zhu.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-BOSTON_HOUSING_DATASET}{}}%
-\CSLLeftMargin{{[}7{]} }%
+\CSLLeftMargin{{[}8{]} }%
 \CSLRightInline{David Harrison and Daniel Rubinfeld.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ae_orig}{}}%
-\CSLLeftMargin{{[}8{]} }%
+\CSLLeftMargin{{[}9{]} }%
 \CSLRightInline{G.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-lstm_orig}{}}%
-\CSLLeftMargin{{[}9{]} }%
+\CSLLeftMargin{{[}10{]} }%
 \CSLRightInline{Sepp Hochreiter and Jürgen Schmidhuber.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Koza1993GeneticP}{}}%
-\CSLLeftMargin{{[}10{]} }%
+\CSLLeftMargin{{[}11{]} }%
 \CSLRightInline{John R. Koza. 1993. Genetic programming - on the programming of computers by means of natural selection. In \emph{Complex adaptive systems}.}
 
+\leavevmode\vadjust pre{\hypertarget{ref-mwu2}{}}%
+\CSLLeftMargin{{[}12{]} }%
+\CSLRightInline{Henry B.
+Mann and Douglas R.
+Whitney.
+1947.
+On a test of whether one of two random variables is stochastically larger than the other.
+\emph{Annals of Mathematical Statistics} 18, (1947), 50--60.}
+
 \leavevmode\vadjust pre{\hypertarget{ref-survey_transfer_learning}{}}%
-\CSLLeftMargin{{[}11{]} }%
+\CSLLeftMargin{{[}13{]} }%
 \CSLRightInline{Sinno Jialin Pan and Qiang Yang.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-10.1007ux2f978-3-540-24621-3_22}{}}%
-\CSLLeftMargin{{[}12{]} }%
+\CSLLeftMargin{{[}14{]} }%
 \CSLRightInline{Grégory Paris, Denis Robilliard, and Cyril Fonlupt.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-harmless_overfitting_eda}{}}%
-\CSLLeftMargin{{[}13{]} }%
+\CSLLeftMargin{{[}15{]} }%
 \CSLRightInline{Malte Probst and Franz Rothlauf.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-design_of_modern_heuristics}{}}%
-\CSLLeftMargin{{[}14{]} }%
+\CSLLeftMargin{{[}16{]} }%
 \CSLRightInline{Franz Rothlauf. 2011. \emph{Design of modern heuristics: Principles and application}. DOI:https://doi.org/\href{https://doi.org/10.1007/978-3-540-72962-4}{10.1007/978-3-540-72962-4}}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-sampling_err_gp}{}}%
-\CSLLeftMargin{{[}15{]} }%
+\CSLLeftMargin{{[}17{]} }%
 \CSLRightInline{Dirk Schweim, David Wittenberg, and Franz Rothlauf.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ENERGY_DATASET}{}}%
-\CSLLeftMargin{{[}16{]} }%
+\CSLLeftMargin{{[}18{]} }%
 \CSLRightInline{Athanasios Tsanas and Angeliki Xifara.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-dae_orig2008}{}}%
-\CSLLeftMargin{{[}17{]} }%
+\CSLLeftMargin{{[}19{]} }%
 \CSLRightInline{Pascal Vincent, Hugo Larochelle, Y.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-weigend1994overfitting}{}}%
-\CSLLeftMargin{{[}18{]} }%
+\CSLLeftMargin{{[}20{]} }%
 \CSLRightInline{Andreas Weigend. 1994. On overfitting and the effective number of hidden units. In \emph{Proceedings of the 1993 connectionist models summer school}, 335--342.}
 
+\leavevmode\vadjust pre{\hypertarget{ref-mwu1}{}}%
+\CSLLeftMargin{{[}21{]} }%
+\CSLRightInline{Frank. Wilcoxon. 1945. Individual comparisons by ranking methods. \emph{Biometrics} 1, (1945), 196--202.}
+
 \leavevmode\vadjust pre{\hypertarget{ref-daegp_explore_exploit}{}}%
-\CSLLeftMargin{{[}19{]} }%
+\CSLLeftMargin{{[}22{]} }%
 \CSLRightInline{David Wittenberg.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-dae-gp_2022_symreg}{}}%
-\CSLLeftMargin{{[}20{]} }%
+\CSLLeftMargin{{[}23{]} }%
 \CSLRightInline{David Wittenberg and Franz Rothlauf.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-dae-gp_2020_rtree}{}}%
-\CSLLeftMargin{{[}21{]} }%
+\CSLLeftMargin{{[}24{]} }%
 \CSLRightInline{David Wittenberg, Franz Rothlauf, and Dirk Schweim.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-CONCRETE_DATASET}{}}%
-\CSLLeftMargin{{[}22{]} }%
+\CSLLeftMargin{{[}25{]} }%
 \CSLRightInline{I-Cheng Yeh.

### 2023-02-04 data/summarize_data.py

@@ -437,7 +437,7 @@ fig.savefig(f"{IMG_PATH}/median_levDistance_byGen.png")
 PT_STRATEGIES_PATH = "/Users/rmn/github/master_thesis/data/PT_Strategies_airfoil_2hl_150hn_fullRun"
 
 FILES = {
-    "Norm._Levenshtein_Distance" : "lev_div_overGens.json",
+    "Norm._Levenshtein_Distance" : "lev_div_ovrGens.json",
     "Size_best_Solution" : "size_best_solution.json",
     "Fitness_lastGen" : "fitness_per_gen.json",
     "Training_Epochs" : "epochs_trained_perGen.json"

### 2023-02-04 paper.Rmd
Sat Feb  4 22:00:06 CET 2023: update-remote.sh

@@
 csl: csl/association-for-computing-machinery.csl
+# csl: csl/harvard-cite-them-right.csl
+# csl: csl/springer-science-reviews.csl
 urlcolor: blue
@@
 toccolor: blue
-#csl: csl/harvard-cite-them-right.csl
-#csl: csl/springer-science-reviews.csl
+github-repo: https://github.com/roman91DE/master_thesis
 include-before: '`\thispagestyle{empty} \newpage{}`{=latex}'
@@
   read.csv("./tables/airfoil_fullRun_2hl_maxIndSize_params", sep = "?"),
-  caption = "DAE-GP - Hyperparameter"
+  caption = "DAE-GP - Hyperparameter",
+  format = "markdown"
 )%>% row_spec(0,bold=TRUE )
@@
 Another interesting metric to examine in population-based optimization algorithms is the diversity of the current generations population.
-Since individual solutions in DAE-GP are computer programs in the form of parse trees, I selected the normalized Levenshtein edit distance as my primary metric for tracking population diversity.
+Since individual solutions in DAE-GP are computer programs in the form of parse trees, I selected the normalized Levenshtein edit distance [@NormLevDistance] as my primary metric for tracking population diversity.
 Figure \@ref(fig:airfoil-popDiversity) shows the median population diversity for both DAE-GP algorithms and regular GP as a benchmark.
@@
 
-To gather more confidence in the previous results, the pre-training strategy for DAE-GP was applied for three additional real world symbolic regression problems and tested for statistically significant differences in the underlying distributions.
+To gather more confidence in the previous results, I applied the pre-training strategy for DAE-GP to three additional real world symbolic regression problems.
 Table \@ref(tab:full-run-realWorldSymReg-problems) briefly summarizes the benchmark problems.
@@
 ```
-To test for statistical differences between the distributions of results achieved by either pre-trained or regular DAE-GP, I selected the  
-nonparametric Mann-Whitney-U Test ([@mwu1], [@mwu2]) to test if the distribution .... Additionaly the effect strength of pre-training is computed by using Cliffs Delta[@Cliff1993DominanceSO].
+To test for statistical differences of the results achieved by either pre-trained or regular DAE-GP, I selected the nonparametric Mann-Whitney-U Test ([@mwu1], [@mwu2]) to test for differences in the distribution underlying the samples sourced from my experiments.
+Additionaly, I included the effect strength of pre-training as measured by Cliffs Delta [@Cliff1993DominanceSO] into the statistical analysis.
+Better median values (e.g.
+higher fitness) are printed bold and p-values are marked with asterisk (1,2,3) for $\alpha$ levels ($0.1$, $0.05$, $0.01$).
 
-## Fitness
+Note that the results for the real world symbolic regression problems inside this chapter, except for the airfoil problem (see subsection [Influence on Run-Time]), have not been analyzed for their total run-time.
+This is due to the fact, that experiments were run on different computers architectures which resulted in strongly biased results.
+
+## Influence on Fitness
 
@@
 
-```{r symReg-Fitness, echo=FALSE, fig.cap="Fitness over 30 Generations - Summary - Real World Symbolic Regression", out.width="80%"}
+```{r symReg-Fitness, echo=FALSE, fig.cap="Fitness over 30 Generations - Real World Symbolic Regression", out.width="90%"}
 knitr::include_graphics("~/masterThesis/master_thesis/img/symbolicRegressionSummarized/median_fitness_byGens.png")
@@
 Table \@ref(tab:full-run-realWorldSymReg-fitness) summarizes the results for the median best fitness achieved after 30 generations and shows the results of computing the statistical tests.
-Higher fitness values (lower RMSE) are printed bold and p-values are marked with asterisk (1,2,3) for $\alpha$ levels ($0.1$, $0.05$, $0.01$).
 
@@
  align = "l",
- caption = "Median Best Fitness after 30 generations - Summary - Real World Symbolic Regression"
+ caption = "Median Best Fitness after 30 generations - Real World Symbolic Regression"
 ) %>% row_spec(0,bold=TRUE ) %>% kableExtra::kable_styling(latex_options = "hold_position")
@@
 
-## Solution Size
+## Influence on Solution Size
 
@@
 
-```{r symReg-Size, echo=FALSE, fig.cap="Size of the best Solution over 30 Generations - Summary - Real World Symbolic Regression", out.width="80%"}
+```{r symReg-Size, echo=FALSE, fig.cap="Size of the best Solution over 30 Generations - Real World Symbolic Regression", out.width="90%"}
 knitr::include_graphics("~/masterThesis/master_thesis/img/symbolicRegressionSummarized/median_bestSol_size_byGens.png")
@@
  align = "l",
- caption = "Median Size of the best solution after 30 generations - Summary - Real World Symbolic Regression"
+ caption = "Median Size of the best solution after 30 generations - Real World Symbolic Regression"
 ) %>% row_spec(0,bold=FALSE) %>% kableExtra::kable_styling(latex_options = "hold_position", font_size = 7)
@@
 ```
-## Population Diversity
+## Influence on Population Diversity
 
-...
+Figure \@ref(fig:symReg-LevDiv) shows the median levenshtein edit distance as a metric for population diversity over the evolutionary search.
 
-![Median Population Diversity over Generations](~/masterThesis/master_thesis/img/symbolicRegressionSummarized/median_levDistance_byGen.png){height=80%}
+```{r symReg-LevDiv, echo=FALSE, fig.cap="Population Diversity over 30 Generations - Real World Symbolic Regression", out.width="90%"}
+knitr::include_graphics("~/masterThesis/master_thesis/img/symbolicRegressionSummarized/median_levDistance_byGen.png")
+```
 
-Testing for statistical differences in the underlying distributions yields the P-Values listed in table X.
+The results of the statistical analysis is summarized in table \@ref(tab:full-run-realWorldSymReg-popDiversity).
 
@@
 for(i in 1:nrow(df)){
-  if(df[i, "DAE.GP"] < df[i, "Pre.Trained_DAE.GP"]){
+  if(df[i, "DAE.GP"] > df[i, "Pre.Trained_DAE.GP"]){
     df_bold[i, "DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(round(df[i, "DAE.GP"],2)))
@@
     }
-  else if (df[i, "DAE.GP"] > df[i, "Pre.Trained_DAE.GP"]) {
+  else if (df[i, "DAE.GP"] < df[i, "Pre.Trained_DAE.GP"]) {
     df_bold[i, "Pre.Trained_DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(round(df[i, "Pre.Trained_DAE.GP"],2)))
@@
  align = "l",
- caption = "Median Population Diversity over Generations - Symbolic Regression"
+ caption = "Median Population Diversity over 30 Generations - Symbolic Regression"
 ) %>% 
@@
 
-## Training epochs per Generation
+## Influence on the number of Training epochs per Generation
 
-...
+Figure \@ref(fig:symReg-EpochsPerGen) shows the median number of epochs spend on training each generations DAE-LSTM network until the termination criterion was satisfied.
 
-![Median Number of Training Epochs per Generation - Symbolic Regression](~/masterThesis/master_thesis/img/symbolicRegressionSummarized/median_epochsTrained.png){height=80%}
+```{r symReg-EpochsPerGen, echo=FALSE, fig.cap="Training Epochs over 30 Generations - Real World Symbolic Regression", out.width="90%"}
+knitr::include_graphics("~/masterThesis/master_thesis/img/symbolicRegressionSummarized/median_epochsTrained.png")
+```
+
+The results of the statistical analysis is summarized in table \@ref(tab:full-run-realWorldSymReg-epochsPerGen).
 
@@
  col.names = col_names,
- align = "c",
+ align = "l",
  caption = "Median Number of Training Epochs per Generation - Symbolic Regression"

### 2023-02-04 paper.tex
Sat Feb  4 22:00:06 CET 2023: update-remote.sh

@@
 
-Denoising Autoencoder Genetic Programming (DAE-GP) is a novel variation of an genetic programming based Estimation of Distribution Algorithm (EDA-GP) that uses a denoising autoencoders long short-term memory network (DAE-LSTMs) as a probabilistic model to sample new candidate solutions {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{24}{]}.
+Denoising Autoencoder Genetic Programming (DAE-GP) is a novel variation of an genetic programming based Estimation of Distribution Algorithm (EDA-GP) that uses a denoising autoencoders long short-term memory network (DAE-LSTMs) as a probabilistic model to sample new candidate solutions {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{25}{]}.
 
@@
 In DAE-GP, DAE-LSTMs are trained with a subset of high-fitness solutions selected from a parent population with the aim to capture their promising qualities.
-The resulting model is then used to sample new offspring solutions by propagating partially mutated solutions from the parent population through the DAE-LSTMS {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{24}{]}.
-In previous work DAE-GP has been shown to outperform GP for both a generalized version of the royal tree problem {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{24}{]} as well as for a real-world symbolic regression problem {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{23}{]}.
-
-The DAE-GP algorithm first described by {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{24}{]} trains a DAE-LSTMs for each generation \(g\) of the search from scratch.
-{[}\protect\hyperlink{ref-dae-gp_2022_symreg}{23}{]} and {[}\protect\hyperlink{ref-daegp_explore_exploit}{22}{]} suggests the incorporation of a pre-training strategy into the evolutionary search as a possible way of improving the overall performance of the DAE-GP algorithm.
+The resulting model is then used to sample new offspring solutions by propagating partially mutated solutions from the parent population through the DAE-LSTMS {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{25}{]}.
+In previous work DAE-GP has been shown to outperform GP for both a generalized version of the royal tree problem {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{25}{]} as well as for a real-world symbolic regression problem {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{24}{]}.
+
+The DAE-GP algorithm first described by {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{25}{]} trains a DAE-LSTMs for each generation \(g\) of the search from scratch.
+{[}\protect\hyperlink{ref-dae-gp_2022_symreg}{24}{]} and {[}\protect\hyperlink{ref-daegp_explore_exploit}{23}{]} suggests the incorporation of a pre-training strategy into the evolutionary search as a possible way of improving the overall performance of the DAE-GP algorithm.
 The key idea is to pre-train an initial DAE-LSTM on a large population of candidate solutions and to use the pre-trained parameters of this initial model in each generation of the search as starting parameters for the current generation DAE-LSTM.
@@
 
-DAE-GP is an EDA-GP algorithm that uses DAE-LSTM networks as a probabilistic model to sample new offspring solutions {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{24}{]}.
+DAE-GP is an EDA-GP algorithm that uses DAE-LSTM networks as a probabilistic model to sample new offspring solutions {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{25}{]}.
 
@@
 Another important concept in EC is the distinction between genotypes and phenotypes of solutions, the genotype contains the information that is necessary to construct the phenotype, the outer appearance of a particular solution on which we measure the overall quality of solutions.
-The representation of a solution is therefore defined by the mapping of genotypes to phenotypes {[}\protect\hyperlink{ref-design_of_modern_heuristics}{16}{]}.
+The representation of a solution is therefore defined by the mapping of genotypes to phenotypes {[}\protect\hyperlink{ref-design_of_modern_heuristics}{17}{]}.
 Genetic operators, such as mutation or recombination are usually applied to the genotype of solutions.
@@
 GP follows the same basic evolutionary principle of EC but searches for more general, hierarchical computer programs of dynamically varying size and shape {[}\protect\hyperlink{ref-Koza1993GeneticP}{11}{]}.
-The computer programs that are at the center of the evolutionary search in GP are commonly represented by tree structures at the level of their phenotype {[}\protect\hyperlink{ref-design_of_modern_heuristics}{16}{]}.
+The computer programs that are at the center of the evolutionary search in GP are commonly represented by tree structures at the level of their phenotype {[}\protect\hyperlink{ref-design_of_modern_heuristics}{17}{]}.
 Since GP searches for high fitness computer programs that produce a desired output for some input, it can be applied to various different problem domains such as symbolic regression, automatic programming, or evolving game-playing strategies {[}\protect\hyperlink{ref-Koza1993GeneticP}{11}{]}.
@@
 GP is an especially useful meta heuristic for problems where no a priori knowledge about the final form of good solutions is available.
-GPs ability to optimize solutions for their structure as well as for their parameters led to it being one of the most prevalent methods used for symbolic regression {[}\protect\hyperlink{ref-10.1007ux2f978-3-540-24621-3_22}{14}{]}.
+GPs ability to optimize solutions for their structure as well as for their parameters led to it being one of the most prevalent methods used for symbolic regression {[}\protect\hyperlink{ref-10.1007ux2f978-3-540-24621-3_22}{15}{]}.
 
@@
 
-The aim of Estimation of distribution algorithms (EDA) is to replace the standard variation operators used in GA by building probabilistic models that can capture complex dependencies between different decision variables of an optimization problem {[}\protect\hyperlink{ref-design_of_modern_heuristics}{16}{]}.
+The aim of Estimation of distribution algorithms (EDA) is to replace the standard variation operators used in GA by building probabilistic models that can capture complex dependencies between different decision variables of an optimization problem {[}\protect\hyperlink{ref-design_of_modern_heuristics}{17}{]}.
 EDAs use this probabilistic model to sample new offspring solutions inside an evolutionary search to replace crossover and/or mutation operators.
@@
 
-One possible way of model building in EDA proposed by {[}\protect\hyperlink{ref-harmless_overfitting_eda}{15}{]} is to use denoising autoencoders (DAEs) as generative models to capture complex probability distributions of decision variables.
+One possible way of model building in EDA proposed by {[}\protect\hyperlink{ref-harmless_overfitting_eda}{16}{]} is to use denoising autoencoders (DAEs) as generative models to capture complex probability distributions of decision variables.
 DAE, a variation of the autoencoders (AE), are a widely used type of neural networks in the domain of unsupervised machine learning that maps \(n\) input variables to \(n\) output variables using a hidden representation.
@@
 
-DAE was first introduced by {[}\protect\hyperlink{ref-dae_orig2008}{19}{]} as an improved AE with the ability to learn new representations of data that is especially robust to partially corrupted input data.
+DAE was first introduced by {[}\protect\hyperlink{ref-dae_orig2008}{20}{]} as an improved AE with the ability to learn new representations of data that is especially robust to partially corrupted input data.
 DAE modifies the AE by using partially corrupted input data for the AE and training it to reconstruct the uncorrupted, original version of the input data.
 
-Since the hidden representation of DAE captures the dependency structure of the input data it can therefore also be used to generate new solutions in the context of GAs {[}\protect\hyperlink{ref-harmless_overfitting_eda}{15}{]}.
-
-DAE-GP builds upon the concept of using DAEs in EDAs described by {[}\protect\hyperlink{ref-harmless_overfitting_eda}{15}{]} and transfers the concept to the domain of GP.
-The mutation and crossover operators of standard GP are replaced by sampling new solutions from a probabilistic model that is build by training a DAEs long short-term memory (LSTM) network on a subset of high fitness solutions from the current population {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{24}{]}.
+Since the hidden representation of DAE captures the dependency structure of the input data it can therefore also be used to generate new solutions in the context of GAs {[}\protect\hyperlink{ref-harmless_overfitting_eda}{16}{]}.
+
+DAE-GP builds upon the concept of using DAEs in EDAs described by {[}\protect\hyperlink{ref-harmless_overfitting_eda}{16}{]} and transfers the concept to the domain of GP.
+The mutation and crossover operators of standard GP are replaced by sampling new solutions from a probabilistic model that is build by training a DAEs long short-term memory (LSTM) network on a subset of high fitness solutions from the current population {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{25}{]}.
 
 LSTMs are a variant of artificial neural networks first introduced by {[}\protect\hyperlink{ref-lstm_orig}{10}{]} that can store learned information over en extended time period while avoiding the problem of vanishing and/or exploding gradients.
-Since DAE-GP encodes candidate solutions as linear strings in prefix notation, the DAE in DAE-GP uses LSTMs for both encoding and decoding where the total amount of time steps \(T\) is equal to sum of the length of the input solution and the output solution {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{24}{]}.
+Since DAE-GP encodes candidate solutions as linear strings in prefix notation, the DAE in DAE-GP uses LSTMs for both encoding and decoding where the total amount of time steps \(T\) is equal to sum of the length of the input solution and the output solution {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{25}{]}.
 
@@
 
-The strategy of pre-training artificial neural networks is based on the idea of transfer learning that aims at retaining previously learned knowledge from one task to re-use it for another task {[}\protect\hyperlink{ref-HAN2021225}{7}{]} {[}\protect\hyperlink{ref-survey_transfer_learning}{13}{]}.
+The strategy of pre-training artificial neural networks is based on the idea of transfer learning that aims at retaining previously learned knowledge from one task to re-use it for another task {[}\protect\hyperlink{ref-HAN2021225}{7}{]} {[}\protect\hyperlink{ref-survey_transfer_learning}{14}{]}.
 Transfer learning traditionally follows a two phase approach:
@@
 
-where source task and target task are usually similar but may differ in their feature space and the distribution of training data {[}\protect\hyperlink{ref-survey_transfer_learning}{13}{]}.
+where source task and target task are usually similar but may differ in their feature space and the distribution of training data {[}\protect\hyperlink{ref-survey_transfer_learning}{14}{]}.
 
@@
 Instead of initializing and optimizing each model from scratch as done in previous work (e.g.
-{[}\protect\hyperlink{ref-dae-gp_2020_rtree}{24}{]} {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{23}{]}), a separate DAE-LSTM network \(\hat{M}\) will be trained on a large population of randomly initialized solutions.\\
+{[}\protect\hyperlink{ref-dae-gp_2020_rtree}{25}{]} {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{24}{]}), a separate DAE-LSTM network \(\hat{M}\) will be trained on a large population of randomly initialized solutions.\\
 The trainable parameters \(\theta_{\hat{M}}\) obtained after finishing the training procedure of \(\hat{M}\) are then used as the starting parameters for each following DAE-LSTM model \(M_g\) for \(g\in{1,...,g_{max}}\).
@@
 Another motivation for introducing pre-training into DAE-GP is the prevalence of sampling error for small GP populations sizes.
-{[}\protect\hyperlink{ref-sampling_err_gp}{17}{]} finds that sampling error, non-systematic errors that are caused by observing only a small subset of the statistical population, is a severe problem in the domain of GP since the initial population may not be a representative sample of all possible solutions.
-{[}\protect\hyperlink{ref-sampling_err_gp}{17}{]} also introduces a method for calculating optimal population sizes to minimize the presence of sampling error.
+{[}\protect\hyperlink{ref-sampling_err_gp}{18}{]} finds that sampling error, non-systematic errors that are caused by observing only a small subset of the statistical population, is a severe problem in the domain of GP since the initial population may not be a representative sample of all possible solutions.
+{[}\protect\hyperlink{ref-sampling_err_gp}{18}{]} also introduces a method for calculating optimal population sizes to minimize the presence of sampling error.
 Since DAE-LSTM of DAE-GP learns the properties of it's training population and reuses the acquired knowledge in the sampling procedure, using the parameters obtained from training \(\hat{M}\) on a sufficiently large training population might reduce the need for large population sizes in the following generations of the search if \(\hat{M}\) already implicitly captured the properties of a representative sample of solutions.
@@
 
-The DAE-GP algorithm, first described by {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{24}{]}, is summarized as a flowchart in figure \ref{fig:daegp-flowchart}.
+The DAE-GP algorithm, first described by {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{25}{]}, is summarized as a flowchart in figure \ref{fig:daegp-flowchart}.
 After setting the generation counter \(g\) to \(0\) and creating the initial population of solutions \(P_0\), DAE-GP enters the main loop that checks if a termination criteria is satisfied (i.e.~maximum number of generations or fitness evaluations is reached).
@@
 One difficulty in the implementation of a pre-training strategy into DAE-GP has been the determination of the number of hidden neurons in side the DAE-LSTMs hidden layers.
-The original description of DAE-GP {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{24}{]} uses a strategy where the number of hidden neurons for each hidden layer is dynamically set per generation to the maximum individual size inside the current population.
+The original description of DAE-GP {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{25}{]} uses a strategy where the number of hidden neurons for each hidden layer is dynamically set per generation to the maximum individual size inside the current population.
 For a pre-training implementation, this strategy can not be easily adapted since it leads to a changing number of neurons at each generation resulting in different dimensions of the DAE-LSTM.
@@
 
-\begin{table}
-
-\caption{\label{tab:airfoil-fullRun-2hl-maxIndSize-params}DAE-GP - Hyperparameter}
-\centering
-\begin{tabular}[t]{l|l}
-\hline
-\textbf{Hyperparameter} & \textbf{Value}\\
-\hline
-Population\_Size & 500\\
-\hline
-Generations & 30\\
-\hline
-Fitness\_Metric & RMSE\\
-\hline
-Training\_Mode & Convergence\\
-\hline
-Sampling\_Steps & 1\\
-\hline
-Hidden\_Neurons & 150\\
-\hline
-Hidden\_Layer & 2\\
-\hline
-Selection\_Operator & Binary Tournament Selection\\
-\hline
-Corruption\_Operator & Levenshtein Edit\\
-\hline
-Denoising\_Edit\_Probability(training/sampling) & 0.05/0.95\\
-\hline
-Function\_Set & \{+ , - , * , aq\}\\
-\hline
-Ephemeral\_Constants & [-5,..,5]\\
-\hline
-Pre-Training\_Population\_Size & 10000\\
-\hline
-Pre-Training\_Train/Test\_Split & 50\%\\
-\hline
-Pre-Training\_Training\_Mode & Early Stopping\\
-\hline
-\end{tabular}
-\end{table}
+\begin{longtable}[]{@{}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.5287}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.4713}}@{}}
+\caption{\label{tab:airfoil-fullRun-2hl-maxIndSize-params}DAE-GP - Hyperparameter}\tabularnewline
+\toprule\noalign{}
+\begin{minipage}[b]{\linewidth}\raggedright
+Hyperparameter
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Value
+\end{minipage} \\
+\midrule\noalign{}
+\endfirsthead
+\toprule\noalign{}
+\begin{minipage}[b]{\linewidth}\raggedright
+Hyperparameter
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Value
+\end{minipage} \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+Population\_Size & 500 \\
+Generations & 30 \\
+Fitness\_Metric & RMSE \\
+Training\_Mode & Convergence({[}\protect\hyperlink{ref-daegp_explore_exploit}{23}{]}) \\
+Reconstruction\_Error\_Metric & Multiclass\_Cross\_Entropy \\
+Learning\_Rate & 0.001 \\
+Batch\_Size & 0.1 \\
+Max\_Tree\_Depth & 17({[}\protect\hyperlink{ref-DEAP_JMLR2012}{6}{]}) \\
+loss function & \\
+Sampling\_Steps & 1 \\
+Hidden\_Neurons & 150 \\
+Hidden\_Layer & 2 \\
+Selection\_Operator & Binary Tournament Selection \\
+Corruption\_Operator & Levenshtein Edit({[}\protect\hyperlink{ref-daegp_explore_exploit}{23}{]}) \\
+Denoising\_Edit\_Probability(training/sampling) & 0.05/0.95 \\
+Function\_Set & \{+ , - , * , aq\} \\
+Ephemeral\_Constants & {[}-5,..,5{]} \\
+Pre-Training\_Population\_Size & 10000 \\
+Pre-Training\_Train/Test\_Split & 50\% \\
+Pre-Training\_Training\_Mode & Early Stopping \\
+\end{longtable}
 
@@
 Symbolic Regression problems have been one of the first GP applications {[}\protect\hyperlink{ref-Koza1993GeneticP}{11}{]} and are an actively studied and highly relevant research area.
-The goal in symbolic regression is to find a mathematical model for a given set of data points {[}\protect\hyperlink{ref-10.1007ux2f978-3-540-24621-3_22}{14}{]}, in real-world symbolic regression these data points are sourced from real-world observations which in contrast to synthetic symbolic regression problem are more likely to contain random noise and bias.
+The goal in symbolic regression is to find a mathematical model for a given set of data points {[}\protect\hyperlink{ref-10.1007ux2f978-3-540-24621-3_22}{15}{]}, in real-world symbolic regression these data points are sourced from real-world observations which in contrast to synthetic symbolic regression problem are more likely to contain random noise and bias.
 Another important challenge in solving real-world symbolic regression problems is the ability for a given model to generalize, we want the final model to show high accuracy in predicting outcomes for previously unseen cases.
@@
 The objective of the airfoil problem is to find a function that accurately predicts the output variable \(y\) by taking in a subset of the input variables \(x1,x2,x3,x4,x5\).
-The function set used by all DAE-GP variations for symbolic regression is summarized in \ref{tab:airfoil-fullRun-2hl-maxIndSize-params}, the terminal set consists of the the input variables \(x1,x2,x3,x4,x5\) and the ephemeral integer constants listed in table \ref{tab:airfoil-fullRun-2hl-maxIndSize-params}{[}\protect\hyperlink{ref-dae-gp_2022_symreg}{23}{]}.
+The function set used by all DAE-GP variations for symbolic regression is summarized in \ref{tab:airfoil-fullRun-2hl-maxIndSize-params}, the terminal set consists of the the input variables \(x1,x2,x3,x4,x5\) and the ephemeral integer constants listed in table \ref{tab:airfoil-fullRun-2hl-maxIndSize-params}{[}\protect\hyperlink{ref-dae-gp_2022_symreg}{24}{]}.
 
@@
 For each configuration I tested traditional DAE-GP as well as a pre-trained DAE-GP resulting in 12 total sub experiments that were each based on 10 individual runs (total number of runs=\(120\)).
-To avoid creating biased results through the presence of sampling error inside the pre-training population (see {[}\protect\hyperlink{ref-sampling_err_gp}{17}{]}), the population size for the pre-training phase is chosen very high with the size of \(\hat{P}\) =10000 where 50\% of \(\hat{P}\) is used for the training population \(\hat{P}_{train}\) and the remaining 50\% are used for the hold-out validation population \(\hat{P}_{test}\).
+To avoid creating biased results through the presence of sampling error inside the pre-training population (see {[}\protect\hyperlink{ref-sampling_err_gp}{18}{]}), the population size for the pre-training phase is chosen very high with the size of \(\hat{P}\) =10000 where 50\% of \(\hat{P}\) is used for the training population \(\hat{P}_{train}\) and the remaining 50\% are used for the hold-out validation population \(\hat{P}_{test}\).
 The training of DAE-LSTM uses a fixed number of 1000 epochs until termination and uses Adam optimization for gradient descent.
@@
 In general I expect that with a growing number of trainable parameters (either by adding more hidden layers or more hidden neurons per layer) the DAE-LSTM will be more prone to overfit to the training population \(P_{train}\) resulting in a small reconstruction error for the training population and a large one for the validation population \(P_{test}\).
-The reason for this effect is that a larger network trained over an extended period of time (without the use of strategies like early stopping), has much more potential to learn noise from the training dataset than a smaller network, which is more likely to result in worsening performance for previously unseen cases {[}\protect\hyperlink{ref-weigend1994overfitting}{20}{]}.
+The reason for this effect is that a larger network trained over an extended period of time (without the use of strategies like early stopping), has much more potential to learn noise from the training dataset than a smaller network, which is more likely to result in worsening performance for previously unseen cases {[}\protect\hyperlink{ref-weigend1994overfitting}{21}{]}.
 
@@
 
-Since previous work by {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{23}{]} demonstrated that DAE-GP, for a given number of 10.000 fitness evaluations, is able to produce solutions that are much smaller in tree size than regular GP, I also studied the influence of pre-training on the size of solutions found by DAE-GP.
+Since previous work by {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{24}{]} demonstrated that DAE-GP, for a given number of 10.000 fitness evaluations, is able to produce solutions that are much smaller in tree size than regular GP, I also studied the influence of pre-training on the size of solutions found by DAE-GP.
 Figure \ref{fig:airfoil-size-combined-30gens} shows the median of both the average size of individuals inside the population as well as the size of the currently best performing solution inside the population.
@@
 Another interesting metric to examine in population-based optimization algorithms is the diversity of the current generations population.
-Since individual solutions in DAE-GP are computer programs in the form of parse trees, I selected the normalized Levenshtein edit distance as my primary metric for tracking population diversity.
+Since individual solutions in DAE-GP are computer programs in the form of parse trees, I selected the normalized Levenshtein edit distance {[}\protect\hyperlink{ref-NormLevDistance}{12}{]} as my primary metric for tracking population diversity.
 Figure \ref{fig:airfoil-popDiversity} shows the median population diversity for both DAE-GP algorithms and regular GP as a benchmark.
@@
 To ensure a fair comparison between pre-trained and regular DAE-GP, the prior experiments used a static number of hidden neurons for each hidden Layer inside the DAE-LSTM networks.
-For comparison, figure \ref{fig:airfoil-maxIndSize-params} shows the total number of trainable parameters for each generation's DAE-LSTM of another experiment where regular DAE-GP was tested with dynamic adjustment of the number of hidden neurons as described in {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{24}{]} and {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{23}{]}.
+For comparison, figure \ref{fig:airfoil-maxIndSize-params} shows the total number of trainable parameters for each generation's DAE-LSTM of another experiment where regular DAE-GP was tested with dynamic adjustment of the number of hidden neurons as described in {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{25}{]} and {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{24}{]}.
 Again, the experiments was based on \(10\) individual runs and otherwise used the same hyperparameters as listed in table \ref{tab:airfoil-fullRun-2hl-maxIndSize-params}.
@@
 
-Since previous research on DAE-GP used a single hidden layers for the airfoil dataset {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{23}{]}, I repeated the experiment with a single hidden layer.
+Since previous research on DAE-GP used a single hidden layers for the airfoil dataset {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{24}{]}, I repeated the experiment with a single hidden layer.
 Based on the findings of {[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{5}{]}, I expected that pre-training DAE-GP would lead to a decrease in overall performance.
@@
 
-To gather more confidence in the previous results, the pre-training strategy for DAE-GP was applied for three additional real world symbolic regression problems and tested for statistically significant differences in the underlying distributions.
+To gather more confidence in the previous results, I applied the pre-training strategy for DAE-GP to three additional real world symbolic regression problems.
 Table \ref{tab:full-run-realWorldSymReg-problems} briefly summarizes the benchmark problems.
@@
 Boston\_Housing & 506 & 13 & {[}\protect\hyperlink{ref-BOSTON_HOUSING_DATASET}{8}{]} \\
-Energy\_Cooling & 768 & 8 & {[}\protect\hyperlink{ref-ENERGY_DATASET}{18}{]} \\
-Concrete & 1030 & 8 & {[}\protect\hyperlink{ref-CONCRETE_DATASET}{25}{]} \\
+Energy\_Cooling & 768 & 8 & {[}\protect\hyperlink{ref-ENERGY_DATASET}{19}{]} \\
+Concrete & 1030 & 8 & {[}\protect\hyperlink{ref-CONCRETE_DATASET}{26}{]} \\
 \end{longtable}
 
-\hypertarget{fitness}{%
-\subsection{Fitness}\label{fitness}}
+To test for statistical differences of the results achieved by either pre-trained or regular DAE-GP, I selected the nonparametric Mann-Whitney-U Test ({[}\protect\hyperlink{ref-mwu1}{22}{]}, {[}\protect\hyperlink{ref-mwu2}{13}{]}) to test for differences in the distribution underlying the samples sourced from my experiments.
+Additionaly, I included the effect strength of pre-training as measured by Cliffs Delta {[}\protect\hyperlink{ref-Cliff1993DominanceSO}{3}{]} into the statistical analysis.
+Better median values (e.g.~higher fitness) are printed bold and p-values are marked with asterisk (1,2,3) for \(\alpha\) levels (\(0.1\), \(0.05\), \(0.01\)).
+
+Note that the results for the real world symbolic regression problems inside this chapter, except for the airfoil problem (see subsection \protect\hyperlink{influence-on-run-time}{Influence on Run-Time}), have not been analyzed for their total run-time.
+This is due to the fact, that experiments were run on different computers architectures which resulted in strongly biased results.
+
+\hypertarget{influence-on-fitness}{%
+\subsection{Influence on Fitness}\label{influence-on-fitness}}
 
@@
 
-{\centering \includegraphics[width=0.8\linewidth]{img/symbolicRegressionSummarized/median_fitness_byGens}
-
-}
-
-\caption{Fitness over 30 Generations - Summary - Real World Symbolic Regression}\label{fig:symReg-Fitness}
-\end{figure}
-
-Table \ref{tab:full-run-realWorldSymReg-fitness} summarizes the results for the median best fitness achieved after 30 generations and shows the results of computing both the Mann-Whitney-U Test ({[}\protect\hyperlink{ref-mwu1}{21}{]}, {[}\protect\hyperlink{ref-mwu2}{12}{]}) as well as the effect strength measured by Cliffs Delta{[}\protect\hyperlink{ref-Cliff1993DominanceSO}{3}{]} over the distribution of final fitness scores.
-Higher fitness values (lower RMSE) are printed bold and p-values are marked with asterisk (1,2,3) for \(\alpha\) levels (\(0.1\), \(0.05\), \(0.01\)).
+{\centering \includegraphics[width=0.9\linewidth]{img/symbolicRegressionSummarized/median_fitness_byGens}
+
+}
+
+\caption{Fitness over 30 Generations - Real World Symbolic Regression}\label{fig:symReg-Fitness}
+\end{figure}
+
+Table \ref{tab:full-run-realWorldSymReg-fitness} summarizes the results for the median best fitness achieved after 30 generations and shows the results of computing the statistical tests.
 
@@
   >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1605}}@{}}
-\caption{\label{tab:full-run-realWorldSymReg-fitness}Median Best Fitness after 30 generations - Summary - Real World Symbolic Regression}\tabularnewline
+\caption{\label{tab:full-run-realWorldSymReg-fitness}Median Best Fitness after 30 generations - Real World Symbolic Regression}\tabularnewline
 \toprule\noalign{}
@@
 
-\hypertarget{solution-size}{%
-\subsection{Solution Size}\label{solution-size}}
+\hypertarget{influence-on-solution-size}{%
+\subsection{Influence on Solution Size}\label{influence-on-solution-size}}
 
@@
 
-{\centering \includegraphics[width=0.8\linewidth]{img/symbolicRegressionSummarized/median_bestSol_size_byGens}
-
-}
-
-\caption{Size of the best Solution over 30 Generations - Summary - Real World Symbolic Regression}\label{fig:symReg-Size}
+{\centering \includegraphics[width=0.9\linewidth]{img/symbolicRegressionSummarized/median_bestSol_size_byGens}
+
+}
+
+\caption{Size of the best Solution over 30 Generations - Real World Symbolic Regression}\label{fig:symReg-Size}
 \end{figure}
@@
 \begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2353}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1618}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1176}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1765}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1176}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1912}}@{}}
-\caption{\label{tab:full-run-realWorldSymReg-best-solution-size}Median Size of the best solution after 30 generations - Summary - Real World Symbolic Regression}\tabularnewline
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2388}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1642}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1045}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1791}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1194}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1940}}@{}}
+\caption{\label{tab:full-run-realWorldSymReg-best-solution-size}Median Size of the best solution after 30 generations - Real World Symbolic Regression}\tabularnewline
 \toprule\noalign{}
@@
 \endlastfoot
-Airfoil & 1 & \textbf{7.8} & 8.4 & 0.44 & 0.20 \\
-Airfoil & 2 & 13.2 & \textbf{8.6} & 0.97 & -0.02 \\
-Boston\_Housing & 2 & 4.6 & \textbf{3.8} & 0.37 & -0.20 \\
-Energy(Cooling) & 2 & 7.8 & \textbf{3} & 0.18 & -0.40 \\
-Concrete & 2 & 9.2 & \textbf{5} & 0.09* & -0.56 \\
+Airfoil & 1 & \textbf{5} & 7 & 0.44 & 0.20 \\
+Airfoil & 2 & 5 & 5 & 0.97 & -0.02 \\
+Boston\_Housing & 2 & 3 & 3 & 0.37 & -0.20 \\
+Energy(Cooling) & 2 & 3 & 3 & 0.18 & -0.40 \\
+Concrete & 2 & 7 & \textbf{5} & 0.09* & -0.56 \\
 \end{longtable}
 
-\hypertarget{population-diversity}{%
-\subsection{Population Diversity}\label{population-diversity}}
-
-\ldots{}
-
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth,height=0.8\textheight]{~/masterThesis/master_thesis/img/symbolicRegressionSummarized/median_levDistance_byGen.png}
-\caption{Median Population Diversity over Generations}
-\end{figure}
-
-Testing for statistical differences in the underlying distributions yields the P-Values listed in table X.
+\hypertarget{influence-on-population-diversity}{%
+\subsection{Influence on Population Diversity}\label{influence-on-population-diversity}}
+
+Figure \ref{fig:symReg-LevDiv} shows the median levenshtein edit distance as a metric for population diversity over the evolutionary search.
+
+\begin{figure}[c]
+
+{\centering \includegraphics[width=0.9\linewidth]{img/symbolicRegressionSummarized/median_levDistance_byGen}
+
+}
+
+\caption{Population Diversity over 30 Generations - Real World Symbolic Regression}\label{fig:symReg-LevDiv}
+\end{figure}
+
+The results of the statistical analysis is summarized in table \ref{tab:full-run-realWorldSymReg-popDiversity}.
 
 \begin{longtable}[]{@{}
-  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2267}}
-  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1600}}
-  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1333}}
-  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1733}}
-  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1200}}
-  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1867}}@{}}
-\caption{\label{tab:full-run-realWorldSymReg-popDiversity}Median Population Diversity over Generations - Symbolic Regression}\tabularnewline
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2388}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1642}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1045}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1791}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1194}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1940}}@{}}
+\caption{\label{tab:full-run-realWorldSymReg-popDiversity}Median Population Diversity over 30 Generations - Symbolic Regression}\tabularnewline
 \toprule\noalign{}
-\begin{minipage}[b]{\linewidth}\centering
+\begin{minipage}[b]{\linewidth}\raggedright
 Problem
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
 Hid.Layers
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
 DAE-GP
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
 Pre-Trained
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
 P-Value
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
 Cliffs-Delta
@@
 \toprule\noalign{}
-\begin{minipage}[b]{\linewidth}\centering
+\begin{minipage}[b]{\linewidth}\raggedright
 Problem
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
 Hid.Layers
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
 DAE-GP
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
 Pre-Trained
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
 P-Value
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
 Cliffs-Delta
@@
 \endlastfoot
-Airfoil & 1 & \textbf{0.86} & 0.93 & 0.00*** & 0.93 \\
-Airfoil & 1 & \textbf{0.86} & 0.93 & 0.00*** & 0.93 \\
-Boston\_Housing & 1 & \textbf{0.88} & 0.95 & 0.00*** & 0.92 \\
-Energy(Cooling) & 1 & \textbf{0.87} & 0.94 & 0.00*** & 0.94 \\
-Concrete & 1 & \textbf{0.87} & 0.94 & 0.00*** & 0.94 \\
+Airfoil & 1 & 0.86 & \textbf{0.93} & 0.00*** & 0.93 \\
+Airfoil & 1 & 0.86 & \textbf{0.93} & 0.00*** & 0.93 \\
+Boston\_Housing & 1 & 0.88 & \textbf{0.95} & 0.00*** & 0.92 \\
+Energy(Cooling) & 1 & 0.87 & \textbf{0.94} & 0.00*** & 0.94 \\
+Concrete & 1 & 0.87 & \textbf{0.94} & 0.00*** & 0.94 \\
 \end{longtable}
 
-\hypertarget{training-epochs-per-generation}{%
-\subsection{Training epochs per Generation}\label{training-epochs-per-generation}}
-
-\ldots{}
-
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth,height=0.8\textheight]{~/masterThesis/master_thesis/img/symbolicRegressionSummarized/median_epochsTrained.png}
-\caption{Median Number of Training Epochs per Generation - Symbolic Regression}
-\end{figure}
+\hypertarget{influence-on-the-number-of-training-epochs-per-generation}{%
+\subsection{Influence on the number of Training epochs per Generation}\label{influence-on-the-number-of-training-epochs-per-generation}}
+
+Figure \ref{fig:symReg-EpochsPerGen} shows the median number of epochs spend on training each generations DAE-LSTM network until the termination criterion was satisfied.
+
+\begin{figure}[c]
+
+{\centering \includegraphics[width=0.9\linewidth]{img/symbolicRegressionSummarized/median_epochsTrained}
+
+}
+
+\caption{Training Epochs over 30 Generations - Real World Symbolic Regression}\label{fig:symReg-EpochsPerGen}
+\end{figure}
+
+The results of the statistical analysis is summarized in table \ref{tab:full-run-realWorldSymReg-epochsPerGen}.
 
 \begin{longtable}[]{@{}
-  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2329}}
-  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1644}}
-  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1096}}
-  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1781}}
-  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1233}}
-  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1918}}@{}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2388}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1642}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1045}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1791}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1194}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1940}}@{}}
 \caption{\label{tab:full-run-realWorldSymReg-epochsPerGen}Median Number of Training Epochs per Generation - Symbolic Regression}\tabularnewline
 \toprule\noalign{}
-\begin{minipage}[b]{\linewidth}\centering
+\begin{minipage}[b]{\linewidth}\raggedright
 Problem
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
 Hid.Layers
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
 DAE-GP
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
 Pre-Trained
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
 P-Value
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
 Cliffs-Delta
@@
 For a single hidden layer, pre-training did show a negative impact on the median final fitness of solutions found as it was suggested by {[}\protect\hyperlink{ref-pmlr-v9-erhan10a}{4}{]}.
-Since current research on DAE-GP for real-world symbolic regression relies on a single hidden layer (see {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{23}{]}), pre-training does come with additional computational expenses for using deeper DAE-LSTMs.
+Since current research on DAE-GP for real-world symbolic regression relies on a single hidden layer (see {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{24}{]}), pre-training does come with additional computational expenses for using deeper DAE-LSTMs.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-NormLevDistance}{}}%
+\CSLLeftMargin{{[}12{]} }%
+\CSLRightInline{Joseph B.
+Kruskal.
+1983.
+An overview of sequence comparison: Time warps, string edits, and macromolecules.
+\emph{SIAM Rev.} 25, 2 (April 1983), 201--237.
+DOI:https://doi.org/\href{https://doi.org/10.1137/1025045}{10.1137/1025045}}
+
 \leavevmode\vadjust pre{\hypertarget{ref-mwu2}{}}%
-\CSLLeftMargin{{[}12{]} }%
+\CSLLeftMargin{{[}13{]} }%
 \CSLRightInline{Henry B.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-survey_transfer_learning}{}}%
-\CSLLeftMargin{{[}13{]} }%
+\CSLLeftMargin{{[}14{]} }%
 \CSLRightInline{Sinno Jialin Pan and Qiang Yang.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-10.1007ux2f978-3-540-24621-3_22}{}}%
-\CSLLeftMargin{{[}14{]} }%
+\CSLLeftMargin{{[}15{]} }%
 \CSLRightInline{Grégory Paris, Denis Robilliard, and Cyril Fonlupt.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-harmless_overfitting_eda}{}}%
-\CSLLeftMargin{{[}15{]} }%
+\CSLLeftMargin{{[}16{]} }%
 \CSLRightInline{Malte Probst and Franz Rothlauf.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-design_of_modern_heuristics}{}}%
-\CSLLeftMargin{{[}16{]} }%
+\CSLLeftMargin{{[}17{]} }%
 \CSLRightInline{Franz Rothlauf. 2011. \emph{Design of modern heuristics: Principles and application}. DOI:https://doi.org/\href{https://doi.org/10.1007/978-3-540-72962-4}{10.1007/978-3-540-72962-4}}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-sampling_err_gp}{}}%
-\CSLLeftMargin{{[}17{]} }%
+\CSLLeftMargin{{[}18{]} }%
 \CSLRightInline{Dirk Schweim, David Wittenberg, and Franz Rothlauf.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ENERGY_DATASET}{}}%
-\CSLLeftMargin{{[}18{]} }%
+\CSLLeftMargin{{[}19{]} }%
 \CSLRightInline{Athanasios Tsanas and Angeliki Xifara.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-dae_orig2008}{}}%
-\CSLLeftMargin{{[}19{]} }%
+\CSLLeftMargin{{[}20{]} }%
 \CSLRightInline{Pascal Vincent, Hugo Larochelle, Y.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-weigend1994overfitting}{}}%
-\CSLLeftMargin{{[}20{]} }%
+\CSLLeftMargin{{[}21{]} }%
 \CSLRightInline{Andreas Weigend. 1994. On overfitting and the effective number of hidden units. In \emph{Proceedings of the 1993 connectionist models summer school}, 335--342.}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-mwu1}{}}%
-\CSLLeftMargin{{[}21{]} }%
+\CSLLeftMargin{{[}22{]} }%
 \CSLRightInline{Frank. Wilcoxon. 1945. Individual comparisons by ranking methods. \emph{Biometrics} 1, (1945), 196--202.}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-daegp_explore_exploit}{}}%
-\CSLLeftMargin{{[}22{]} }%
+\CSLLeftMargin{{[}23{]} }%
 \CSLRightInline{David Wittenberg.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-dae-gp_2022_symreg}{}}%
-\CSLLeftMargin{{[}23{]} }%
+\CSLLeftMargin{{[}24{]} }%
 \CSLRightInline{David Wittenberg and Franz Rothlauf.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-dae-gp_2020_rtree}{}}%
-\CSLLeftMargin{{[}24{]} }%
+\CSLLeftMargin{{[}25{]} }%
 \CSLRightInline{David Wittenberg, Franz Rothlauf, and Dirk Schweim.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-CONCRETE_DATASET}{}}%
-\CSLLeftMargin{{[}25{]} }%
+\CSLLeftMargin{{[}26{]} }%
 \CSLRightInline{I-Cheng Yeh.

### 2023-02-05 paper.Rmd
Sun Feb  5 13:00:05 CET 2023: update-remote.sh

@@
 
-If not stated otherwise, all experiments in this thesis use the hyperparameters listed in table \@ref(tab:airfoil-fullRun-2hl-maxIndSize-params).
+If not stated otherwise, all experiments in this thesis use the GP hyperparameters listed in table \@ref(tab:baseParams-GP) and the DAE-LSTM specific hyperparameters listed in table \@ref(tab:baseParams-DAEGP).
 
-```{r airfoil-fullRun-2hl-maxIndSize-params, echo=FALSE, message=FALSE, warning=FALSE}
+```{r baseParams-GP, echo=FALSE, message=FALSE, warning=FALSE}
 fig.pos="H"
 knitr::kable(
-  read.csv("./tables/airfoil_fullRun_2hl_maxIndSize_params", sep = "?"),
-  caption = "DAE-GP - Hyperparameter",
+  read.csv("./tables/baseParams_GP", sep = "?"),
+  caption = "GP - Hyperparameter",
+  format = "markdown"
+)%>% row_spec(0,bold=TRUE )
+```
+
+```{r baseParams-DAEGP, echo=FALSE, message=FALSE, warning=FALSE}
+fig.pos="H"
+knitr::kable(
+  read.csv("./tables/baseParams_DAEGP", sep = "?"),
+  caption = "DAE-LSTM - Hyperparameter",
   format = "markdown"
@@
 The objective of the airfoil problem is to find a function that accurately predicts the output variable $y$ by taking in a subset of the input variables $x1,x2,x3,x4,x5$.
-The function set used by all DAE-GP variations for symbolic regression is summarized in \@ref(tab:airfoil-fullRun-2hl-maxIndSize-params), the terminal set consists of the the input variables $x1,x2,x3,x4,x5$ and the ephemeral integer constants listed in table \@ref(tab:airfoil-fullRun-2hl-maxIndSize-params)[@dae-gp_2022_symreg].
+The function set used by all DAE-GP variations for symbolic regression is summarized in tables \@ref(tab:baseParams-GP) and , the terminal set consists of the the input variables $x1,x2,x3,x4,x5$ and the ephemeral integer constants listed in table \@ref(tab:baseParams-GP) [@dae-gp_2022_symreg].
 
@@
 For comparison, figure \@ref(fig:airfoil-maxIndSize-params) shows the total number of trainable parameters for each generation's DAE-LSTM of another experiment where regular DAE-GP was tested with dynamic adjustment of the number of hidden neurons as described in [@dae-gp_2020_rtree] and [@dae-gp_2022_symreg].
-Again, the experiments was based on $10$ individual runs and otherwise used the same hyperparameters as listed in table \@ref(tab:airfoil-fullRun-2hl-maxIndSize-params).
+Again, the experiments was based on $10$ individual runs and otherwise used the same hyperparameters as listed in tables \@ref(tab:baseParams-GP) and \@ref(tab:baseParams-DAEGP).
 It clearly shows that DAE-GP can strongly reduce the number of trainable parameters by adjusting the number of hidden neurons per hidden layers to the current generations population which largley reduces the amount of training necessary and likely reduce the run-time of DAE-GP.

### 2023-02-05 paper.tex
Sun Feb  5 13:00:05 CET 2023: update-remote.sh

@@
 \author{}
-\date{\vspace{-2.5em}Date of Submission: 2023-02-04}
+\date{\vspace{-2.5em}Date of Submission: 2023-02-05}
 
@@
 
-If not stated otherwise, all experiments in this thesis use the hyperparameters listed in table \ref{tab:airfoil-fullRun-2hl-maxIndSize-params}.
+If not stated otherwise, all experiments in this thesis use the GP hyperparameters listed in table \ref{tab:baseParams-GP} and the DAE-LSTM specific hyperparameters listed in table \ref{tab:baseParams-DAEGP}.
 
 \begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.5287}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.4713}}@{}}
-\caption{\label{tab:airfoil-fullRun-2hl-maxIndSize-params}DAE-GP - Hyperparameter}\tabularnewline
+  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.4225}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.5775}}@{}}
+\caption{\label{tab:baseParams-GP}GP - Hyperparameter}\tabularnewline
 \toprule\noalign{}
@@
 \endlastfoot
-Population\_Size & 500 \\
+Population Size & 500 \\
 Generations & 30 \\
-Fitness\_Metric & RMSE \\
-Training\_Mode & Convergence({[}\protect\hyperlink{ref-daegp_explore_exploit}{23}{]}) \\
-Reconstruction\_Error\_Metric & Multiclass\_Cross\_Entropy \\
-Learning\_Rate & 0.001 \\
-Batch\_Size & 0.1 \\
-Max\_Tree\_Depth & 17({[}\protect\hyperlink{ref-DEAP_JMLR2012}{6}{]}) \\
-loss function & \\
-Sampling\_Steps & 1 \\
-Hidden\_Neurons & 150 \\
-Hidden\_Layer & 2 \\
-Selection\_Operator & Binary Tournament Selection \\
-Corruption\_Operator & Levenshtein Edit({[}\protect\hyperlink{ref-daegp_explore_exploit}{23}{]}) \\
-Denoising\_Edit\_Probability(training/sampling) & 0.05/0.95 \\
-Function\_Set & \{+ , - , * , aq\} \\
-Ephemeral\_Constants & {[}-5,..,5{]} \\
-Pre-Training\_Population\_Size & 10000 \\
-Pre-Training\_Train/Test\_Split & 50\% \\
-Pre-Training\_Training\_Mode & Early Stopping \\
+Fitness Metric & Root Mean Squared Error (RMSE) \\
+Train/Test Split & 50:50 \\
+Maximum Tree Depth & 17({[}\protect\hyperlink{ref-DEAP_JMLR2012}{6}{]}) \\
+Initialization Algorithm & Ramped Half and Half ({[}\protect\hyperlink{ref-Koza1993GeneticP}{11}{]}) \\
+Mix/Max Initialization-Depth & 2/6 \\
+Selection Operator & Binary Tournament Selection \\
+Function Set & \{+ , - , * , aq\} \\
+Ephemeral Constants & {[}-5,..,5{]} \\
+Pre-Training Population Size & 10000 \\
+Pre-Training Train/Test Split & 50:50 \\
+\end{longtable}
+
+\begin{longtable}[]{@{}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.5176}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.4824}}@{}}
+\caption{\label{tab:baseParams-DAEGP}DAE-LSTM - Hyperparameter}\tabularnewline
+\toprule\noalign{}
+\begin{minipage}[b]{\linewidth}\raggedright
+Hyperparameter
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Value
+\end{minipage} \\
+\midrule\noalign{}
+\endfirsthead
+\toprule\noalign{}
+\begin{minipage}[b]{\linewidth}\raggedright
+Hyperparameter
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Value
+\end{minipage} \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+Stopping Condition & Convergence({[}\protect\hyperlink{ref-daegp_explore_exploit}{23}{]}) \\
+Reconstruction Error Metric & Multiclass Cross Entropy \\
+Learning Rate & 0.001 \\
+Batch Size & 0.1 \\
+Sampling Steps & 1 \\
+Hidden Layer & 2 \\
+Hidden Neurons & 150 \\
+Corruption Operator & Levenshtein Edit({[}\protect\hyperlink{ref-daegp_explore_exploit}{23}{]}) \\
+Denoising Edit Probability (Model Building) & 0.05 \\
+Denoising Edit Probability (Model Sampling) & 0.95 \\
+Pre-Training Stopping Condition & Early Stopping \\
 \end{longtable}
@@
 The objective of the airfoil problem is to find a function that accurately predicts the output variable \(y\) by taking in a subset of the input variables \(x1,x2,x3,x4,x5\).
-The function set used by all DAE-GP variations for symbolic regression is summarized in \ref{tab:airfoil-fullRun-2hl-maxIndSize-params}, the terminal set consists of the the input variables \(x1,x2,x3,x4,x5\) and the ephemeral integer constants listed in table \ref{tab:airfoil-fullRun-2hl-maxIndSize-params}{[}\protect\hyperlink{ref-dae-gp_2022_symreg}{24}{]}.
+The function set used by all DAE-GP variations for symbolic regression is summarized in tables \ref{tab:baseParams-GP} and , the terminal set consists of the the input variables \(x1,x2,x3,x4,x5\) and the ephemeral integer constants listed in table \ref{tab:baseParams-GP} {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{24}{]}.
 
@@
 For comparison, figure \ref{fig:airfoil-maxIndSize-params} shows the total number of trainable parameters for each generation's DAE-LSTM of another experiment where regular DAE-GP was tested with dynamic adjustment of the number of hidden neurons as described in {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{25}{]} and {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{24}{]}.
-Again, the experiments was based on \(10\) individual runs and otherwise used the same hyperparameters as listed in table \ref{tab:airfoil-fullRun-2hl-maxIndSize-params}.
+Again, the experiments was based on \(10\) individual runs and otherwise used the same hyperparameters as listed in tables \ref{tab:baseParams-GP} and \ref{tab:baseParams-DAEGP}.
 It clearly shows that DAE-GP can strongly reduce the number of trainable parameters by adjusting the number of hidden neurons per hidden layers to the current generations population which largley reduces the amount of training necessary and likely reduce the run-time of DAE-GP.

### 2023-02-05 paper.Rmd
autogit

@@
 Also I did not perform an extended optimization of hyperparameters (e.g by using a grid search strategy) which might yield configurations that increase the benefits of pre-training.
-As described in subsection [Effect on Run-Time], sampling new solutions has been a large contributor to the large performance hit on run-time by using pre-training.
+As described in subsection [Influence on Run-Time], sampling new solutions has been a large contributor to the large performance hit on run-time by using pre-training.
 It could be a worthwhile endevour to explore the main factors that increase sampling time for pre-trained DAE-GP to more efficiently exploit the benefits of pre-training like reducing the number of training epochs.

### 2023-02-05 paper.tex
autogit

@@
 
-Denoising Autoencoder Genetic Programming (DAE-GP) is a novel variation of an genetic programming based Estimation of Distribution Algorithm (EDA-GP) that uses a denoising autoencoders long short-term memory network (DAE-LSTMs) as a probabilistic model to sample new candidate solutions {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{25}{]}.
+Denoising Autoencoder Genetic Programming (DAE-GP) is a novel variation of an genetic programming based Estimation of Distribution Algorithm (EDA-GP) that uses a denoising autoencoders long short-term memory network (DAE-LSTMs) as a probabilistic model to sample new candidate solutions {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{26}{]}.
 
@@
 In DAE-GP, DAE-LSTMs are trained with a subset of high-fitness solutions selected from a parent population with the aim to capture their promising qualities.
-The resulting model is then used to sample new offspring solutions by propagating partially mutated solutions from the parent population through the DAE-LSTMS {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{25}{]}.
-In previous work DAE-GP has been shown to outperform GP for both a generalized version of the royal tree problem {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{25}{]} as well as for a real-world symbolic regression problem {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{24}{]}.
-
-The DAE-GP algorithm first described by {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{25}{]} trains a DAE-LSTMs for each generation \(g\) of the search from scratch.
-{[}\protect\hyperlink{ref-dae-gp_2022_symreg}{24}{]} and {[}\protect\hyperlink{ref-daegp_explore_exploit}{23}{]} suggests the incorporation of a pre-training strategy into the evolutionary search as a possible way of improving the overall performance of the DAE-GP algorithm.
+The resulting model is then used to sample new offspring solutions by propagating partially mutated solutions from the parent population through the DAE-LSTMS {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{26}{]}.
+In previous work DAE-GP has been shown to outperform GP for both a generalized version of the royal tree problem {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{26}{]} as well as for a real-world symbolic regression problem {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{25}{]}.
+
+The DAE-GP algorithm first described by {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{26}{]} trains a DAE-LSTMs for each generation \(g\) of the search from scratch.
+{[}\protect\hyperlink{ref-dae-gp_2022_symreg}{25}{]} and {[}\protect\hyperlink{ref-daegp_explore_exploit}{24}{]} suggests the incorporation of a pre-training strategy into the evolutionary search as a possible way of improving the overall performance of the DAE-GP algorithm.
 The key idea is to pre-train an initial DAE-LSTM on a large population of candidate solutions and to use the pre-trained parameters of this initial model in each generation of the search as starting parameters for the current generation DAE-LSTM.
@@
 
-DAE-GP is an EDA-GP algorithm that uses DAE-LSTM networks as a probabilistic model to sample new offspring solutions {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{25}{]}.
+DAE-GP is an EDA-GP algorithm that uses DAE-LSTM networks as a probabilistic model to sample new offspring solutions {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{26}{]}.
 
@@
 Another important concept in EC is the distinction between genotypes and phenotypes of solutions, the genotype contains the information that is necessary to construct the phenotype, the outer appearance of a particular solution on which we measure the overall quality of solutions.
-The representation of a solution is therefore defined by the mapping of genotypes to phenotypes {[}\protect\hyperlink{ref-design_of_modern_heuristics}{17}{]}.
+The representation of a solution is therefore defined by the mapping of genotypes to phenotypes {[}\protect\hyperlink{ref-design_of_modern_heuristics}{18}{]}.
 Genetic operators, such as mutation or recombination are usually applied to the genotype of solutions.
@@
 
-GP follows the same basic evolutionary principle of EC but searches for more general, hierarchical computer programs of dynamically varying size and shape {[}\protect\hyperlink{ref-Koza1993GeneticP}{11}{]}.
-The computer programs that are at the center of the evolutionary search in GP are commonly represented by tree structures at the level of their phenotype {[}\protect\hyperlink{ref-design_of_modern_heuristics}{17}{]}.
-Since GP searches for high fitness computer programs that produce a desired output for some input, it can be applied to various different problem domains such as symbolic regression, automatic programming, or evolving game-playing strategies {[}\protect\hyperlink{ref-Koza1993GeneticP}{11}{]}.
+GP follows the same basic evolutionary principle of EC but searches for more general, hierarchical computer programs of dynamically varying size and shape {[}\protect\hyperlink{ref-Koza1993GeneticP}{12}{]}.
+The computer programs that are at the center of the evolutionary search in GP are commonly represented by tree structures at the level of their phenotype {[}\protect\hyperlink{ref-design_of_modern_heuristics}{18}{]}.
+Since GP searches for high fitness computer programs that produce a desired output for some input, it can be applied to various different problem domains such as symbolic regression, automatic programming, or evolving game-playing strategies {[}\protect\hyperlink{ref-Koza1993GeneticP}{12}{]}.
 An important quality of GP is the ability to search for solutions of variable length and structure.
 GP is an especially useful meta heuristic for problems where no a priori knowledge about the final form of good solutions is available.
-GPs ability to optimize solutions for their structure as well as for their parameters led to it being one of the most prevalent methods used for symbolic regression {[}\protect\hyperlink{ref-10.1007ux2f978-3-540-24621-3_22}{15}{]}.
+GPs ability to optimize solutions for their structure as well as for their parameters led to it being one of the most prevalent methods used for symbolic regression {[}\protect\hyperlink{ref-10.1007ux2f978-3-540-24621-3_22}{16}{]}.
 
@@
 
-The aim of Estimation of distribution algorithms (EDA) is to replace the standard variation operators used in GA by building probabilistic models that can capture complex dependencies between different decision variables of an optimization problem {[}\protect\hyperlink{ref-design_of_modern_heuristics}{17}{]}.
+The aim of Estimation of distribution algorithms (EDA) is to replace the standard variation operators used in GA by building probabilistic models that can capture complex dependencies between different decision variables of an optimization problem {[}\protect\hyperlink{ref-design_of_modern_heuristics}{18}{]}.
 EDAs use this probabilistic model to sample new offspring solutions inside an evolutionary search to replace crossover and/or mutation operators.
@@
 
-One possible way of model building in EDA proposed by {[}\protect\hyperlink{ref-harmless_overfitting_eda}{16}{]} is to use denoising autoencoders (DAEs) as generative models to capture complex probability distributions of decision variables.
+One possible way of model building in EDA proposed by {[}\protect\hyperlink{ref-harmless_overfitting_eda}{17}{]} is to use denoising autoencoders (DAEs) as generative models to capture complex probability distributions of decision variables.
 DAE, a variation of the autoencoders (AE), are a widely used type of neural networks in the domain of unsupervised machine learning that maps \(n\) input variables to \(n\) output variables using a hidden representation.
@@
 
-DAE was first introduced by {[}\protect\hyperlink{ref-dae_orig2008}{20}{]} as an improved AE with the ability to learn new representations of data that is especially robust to partially corrupted input data.
+DAE was first introduced by {[}\protect\hyperlink{ref-dae_orig2008}{21}{]} as an improved AE with the ability to learn new representations of data that is especially robust to partially corrupted input data.
 DAE modifies the AE by using partially corrupted input data for the AE and training it to reconstruct the uncorrupted, original version of the input data.
 
-Since the hidden representation of DAE captures the dependency structure of the input data it can therefore also be used to generate new solutions in the context of GAs {[}\protect\hyperlink{ref-harmless_overfitting_eda}{16}{]}.
-
-DAE-GP builds upon the concept of using DAEs in EDAs described by {[}\protect\hyperlink{ref-harmless_overfitting_eda}{16}{]} and transfers the concept to the domain of GP.
-The mutation and crossover operators of standard GP are replaced by sampling new solutions from a probabilistic model that is build by training a DAEs long short-term memory (LSTM) network on a subset of high fitness solutions from the current population {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{25}{]}.
+Since the hidden representation of DAE captures the dependency structure of the input data it can therefore also be used to generate new solutions in the context of GAs {[}\protect\hyperlink{ref-harmless_overfitting_eda}{17}{]}.
+
+DAE-GP builds upon the concept of using DAEs in EDAs described by {[}\protect\hyperlink{ref-harmless_overfitting_eda}{17}{]} and transfers the concept to the domain of GP.
+The mutation and crossover operators of standard GP are replaced by sampling new solutions from a probabilistic model that is build by training a DAEs long short-term memory (LSTM) network on a subset of high fitness solutions from the current population {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{26}{]}.
 
 LSTMs are a variant of artificial neural networks first introduced by {[}\protect\hyperlink{ref-lstm_orig}{10}{]} that can store learned information over en extended time period while avoiding the problem of vanishing and/or exploding gradients.
-Since DAE-GP encodes candidate solutions as linear strings in prefix notation, the DAE in DAE-GP uses LSTMs for both encoding and decoding where the total amount of time steps \(T\) is equal to sum of the length of the input solution and the output solution {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{25}{]}.
+Since DAE-GP encodes candidate solutions as linear strings in prefix notation, the DAE in DAE-GP uses LSTMs for both encoding and decoding where the total amount of time steps \(T\) is equal to sum of the length of the input solution and the output solution {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{26}{]}.
 
@@
 
-The strategy of pre-training artificial neural networks is based on the idea of transfer learning that aims at retaining previously learned knowledge from one task to re-use it for another task {[}\protect\hyperlink{ref-HAN2021225}{7}{]} {[}\protect\hyperlink{ref-survey_transfer_learning}{14}{]}.
+The strategy of pre-training artificial neural networks is based on the idea of transfer learning that aims at retaining previously learned knowledge from one task to re-use it for another task {[}\protect\hyperlink{ref-HAN2021225}{7}{]} {[}\protect\hyperlink{ref-survey_transfer_learning}{15}{]}.
 Transfer learning traditionally follows a two phase approach:
@@
 
-where source task and target task are usually similar but may differ in their feature space and the distribution of training data {[}\protect\hyperlink{ref-survey_transfer_learning}{14}{]}.
+where source task and target task are usually similar but may differ in their feature space and the distribution of training data {[}\protect\hyperlink{ref-survey_transfer_learning}{15}{]}.
 
@@
 Instead of initializing and optimizing each model from scratch as done in previous work (e.g.
-{[}\protect\hyperlink{ref-dae-gp_2020_rtree}{25}{]} {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{24}{]}), a separate DAE-LSTM network \(\hat{M}\) will be trained on a large population of randomly initialized solutions.\\
+{[}\protect\hyperlink{ref-dae-gp_2020_rtree}{26}{]} {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{25}{]}), a separate DAE-LSTM network \(\hat{M}\) will be trained on a large population of randomly initialized solutions.\\
 The trainable parameters \(\theta_{\hat{M}}\) obtained after finishing the training procedure of \(\hat{M}\) are then used as the starting parameters for each following DAE-LSTM model \(M_g\) for \(g\in{1,...,g_{max}}\).
@@
 Another motivation for introducing pre-training into DAE-GP is the prevalence of sampling error for small GP populations sizes.
-{[}\protect\hyperlink{ref-sampling_err_gp}{18}{]} finds that sampling error, non-systematic errors that are caused by observing only a small subset of the statistical population, is a severe problem in the domain of GP since the initial population may not be a representative sample of all possible solutions.
-{[}\protect\hyperlink{ref-sampling_err_gp}{18}{]} also introduces a method for calculating optimal population sizes to minimize the presence of sampling error.
+{[}\protect\hyperlink{ref-sampling_err_gp}{19}{]} finds that sampling error, non-systematic errors that are caused by observing only a small subset of the statistical population, is a severe problem in the domain of GP since the initial population may not be a representative sample of all possible solutions.
+{[}\protect\hyperlink{ref-sampling_err_gp}{19}{]} also introduces a method for calculating optimal population sizes to minimize the presence of sampling error.
 Since DAE-LSTM of DAE-GP learns the properties of it's training population and reuses the acquired knowledge in the sampling procedure, using the parameters obtained from training \(\hat{M}\) on a sufficiently large training population might reduce the need for large population sizes in the following generations of the search if \(\hat{M}\) already implicitly captured the properties of a representative sample of solutions.
@@
 
-The DAE-GP algorithm, first described by {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{25}{]}, is summarized as a flowchart in figure \ref{fig:daegp-flowchart}.
+The DAE-GP algorithm, first described by {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{26}{]}, is summarized as a flowchart in figure \ref{fig:daegp-flowchart}.
 After setting the generation counter \(g\) to \(0\) and creating the initial population of solutions \(P_0\), DAE-GP enters the main loop that checks if a termination criteria is satisfied (i.e.~maximum number of generations or fitness evaluations is reached).
@@
 One difficulty in the implementation of a pre-training strategy into DAE-GP has been the determination of the number of hidden neurons in side the DAE-LSTMs hidden layers.
-The original description of DAE-GP {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{25}{]} uses a strategy where the number of hidden neurons for each hidden layer is dynamically set per generation to the maximum individual size inside the current population.
+The original description of DAE-GP {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{26}{]} uses a strategy where the number of hidden neurons for each hidden layer is dynamically set per generation to the maximum individual size inside the current population.
 For a pre-training implementation, this strategy can not be easily adapted since it leads to a changing number of neurons at each generation resulting in different dimensions of the DAE-LSTM.
@@
 Train/Test Split & 50:50 \\
-Maximum Tree Depth & 17({[}\protect\hyperlink{ref-DEAP_JMLR2012}{6}{]}) \\
-Initialization Algorithm & Ramped Half and Half ({[}\protect\hyperlink{ref-Koza1993GeneticP}{11}{]}) \\
+Maximum Tree Depth & 17 {[}\protect\hyperlink{ref-DEAP_JMLR2012}{6}{]} \\
+Initialization Algorithm & Ramped Half and Half {[}\protect\hyperlink{ref-Koza1993GeneticP}{12}{]} \\
 Mix/Max Initialization-Depth & 2/6 \\
@@
 \begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.5176}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.4824}}@{}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.5116}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.4884}}@{}}
 \caption{\label{tab:baseParams-DAEGP}DAE-LSTM - Hyperparameter}\tabularnewline
@@
 \endlastfoot
-Stopping Condition & Convergence({[}\protect\hyperlink{ref-daegp_explore_exploit}{23}{]}) \\
+Stopping Condition & Convergence {[}\protect\hyperlink{ref-daegp_explore_exploit}{24}{]} \\
 Reconstruction Error Metric & Multiclass Cross Entropy \\
+Gradient Descent Algorithm & Adam Optimization {[}\protect\hyperlink{ref-Kingma2014AdamAM}{11}{]} \\
 Learning Rate & 0.001 \\
@@
 Hidden Neurons & 150 \\
-Corruption Operator & Levenshtein Edit({[}\protect\hyperlink{ref-daegp_explore_exploit}{23}{]}) \\
+Corruption Operator & Levenshtein Edit {[}\protect\hyperlink{ref-daegp_explore_exploit}{24}{]} \\
 Denoising Edit Probability (Model Building) & 0.05 \\
@@
 To test pre-training in DAE-GP this thesis focuses on the domain of real-world symbolic regression problems.
-Symbolic Regression problems have been one of the first GP applications {[}\protect\hyperlink{ref-Koza1993GeneticP}{11}{]} and are an actively studied and highly relevant research area.
-The goal in symbolic regression is to find a mathematical model for a given set of data points {[}\protect\hyperlink{ref-10.1007ux2f978-3-540-24621-3_22}{15}{]}, in real-world symbolic regression these data points are sourced from real-world observations which in contrast to synthetic symbolic regression problem are more likely to contain random noise and bias.
+Symbolic Regression problems have been one of the first GP applications {[}\protect\hyperlink{ref-Koza1993GeneticP}{12}{]} and are an actively studied and highly relevant research area.
+The goal in symbolic regression is to find a mathematical model for a given set of data points {[}\protect\hyperlink{ref-10.1007ux2f978-3-540-24621-3_22}{16}{]}, in real-world symbolic regression these data points are sourced from real-world observations which in contrast to synthetic symbolic regression problem are more likely to contain random noise and bias.
 Another important challenge in solving real-world symbolic regression problems is the ability for a given model to generalize, we want the final model to show high accuracy in predicting outcomes for previously unseen cases.
@@
 The objective of the airfoil problem is to find a function that accurately predicts the output variable \(y\) by taking in a subset of the input variables \(x1,x2,x3,x4,x5\).
-The function set used by all DAE-GP variations for symbolic regression is summarized in tables \ref{tab:baseParams-GP} and , the terminal set consists of the the input variables \(x1,x2,x3,x4,x5\) and the ephemeral integer constants listed in table \ref{tab:baseParams-GP} {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{24}{]}.
+The function set used by all DAE-GP variations for symbolic regression is summarized in tables \ref{tab:baseParams-GP} and , the terminal set consists of the the input variables \(x1,x2,x3,x4,x5\) and the ephemeral integer constants listed in table \ref{tab:baseParams-GP} {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{25}{]}.
 
@@
 For each configuration I tested traditional DAE-GP as well as a pre-trained DAE-GP resulting in 12 total sub experiments that were each based on 10 individual runs (total number of runs=\(120\)).
-To avoid creating biased results through the presence of sampling error inside the pre-training population (see {[}\protect\hyperlink{ref-sampling_err_gp}{18}{]}), the population size for the pre-training phase is chosen very high with the size of \(\hat{P}\) =10000 where 50\% of \(\hat{P}\) is used for the training population \(\hat{P}_{train}\) and the remaining 50\% are used for the hold-out validation population \(\hat{P}_{test}\).
+To avoid creating biased results through the presence of sampling error inside the pre-training population (see {[}\protect\hyperlink{ref-sampling_err_gp}{19}{]}), the population size for the pre-training phase is chosen very high with the size of \(\hat{P}\) =10000 where 50\% of \(\hat{P}\) is used for the training population \(\hat{P}_{train}\) and the remaining 50\% are used for the hold-out validation population \(\hat{P}_{test}\).
 The training of DAE-LSTM uses a fixed number of 1000 epochs until termination and uses Adam optimization for gradient descent.
@@
 In general I expect that with a growing number of trainable parameters (either by adding more hidden layers or more hidden neurons per layer) the DAE-LSTM will be more prone to overfit to the training population \(P_{train}\) resulting in a small reconstruction error for the training population and a large one for the validation population \(P_{test}\).
-The reason for this effect is that a larger network trained over an extended period of time (without the use of strategies like early stopping), has much more potential to learn noise from the training dataset than a smaller network, which is more likely to result in worsening performance for previously unseen cases {[}\protect\hyperlink{ref-weigend1994overfitting}{21}{]}.
+The reason for this effect is that a larger network trained over an extended period of time (without the use of strategies like early stopping), has much more potential to learn noise from the training dataset than a smaller network, which is more likely to result in worsening performance for previously unseen cases {[}\protect\hyperlink{ref-weigend1994overfitting}{22}{]}.
 
@@
 
-Since previous work by {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{24}{]} demonstrated that DAE-GP, for a given number of 10.000 fitness evaluations, is able to produce solutions that are much smaller in tree size than regular GP, I also studied the influence of pre-training on the size of solutions found by DAE-GP.
+Since previous work by {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{25}{]} demonstrated that DAE-GP, for a given number of 10.000 fitness evaluations, is able to produce solutions that are much smaller in tree size than regular GP, I also studied the influence of pre-training on the size of solutions found by DAE-GP.
 Figure \ref{fig:airfoil-size-combined-30gens} shows the median of both the average size of individuals inside the population as well as the size of the currently best performing solution inside the population.
@@
 Another interesting metric to examine in population-based optimization algorithms is the diversity of the current generations population.
-Since individual solutions in DAE-GP are computer programs in the form of parse trees, I selected the normalized Levenshtein edit distance {[}\protect\hyperlink{ref-NormLevDistance}{12}{]} as my primary metric for tracking population diversity.
+Since individual solutions in DAE-GP are computer programs in the form of parse trees, I selected the normalized Levenshtein edit distance {[}\protect\hyperlink{ref-NormLevDistance}{13}{]} as my primary metric for tracking population diversity.
 Figure \ref{fig:airfoil-popDiversity} shows the median population diversity for both DAE-GP algorithms and regular GP as a benchmark.
@@
 To ensure a fair comparison between pre-trained and regular DAE-GP, the prior experiments used a static number of hidden neurons for each hidden Layer inside the DAE-LSTM networks.
-For comparison, figure \ref{fig:airfoil-maxIndSize-params} shows the total number of trainable parameters for each generation's DAE-LSTM of another experiment where regular DAE-GP was tested with dynamic adjustment of the number of hidden neurons as described in {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{25}{]} and {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{24}{]}.
+For comparison, figure \ref{fig:airfoil-maxIndSize-params} shows the total number of trainable parameters for each generation's DAE-LSTM of another experiment where regular DAE-GP was tested with dynamic adjustment of the number of hidden neurons as described in {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{26}{]} and {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{25}{]}.
 Again, the experiments was based on \(10\) individual runs and otherwise used the same hyperparameters as listed in tables \ref{tab:baseParams-GP} and \ref{tab:baseParams-DAEGP}.
@@
 
-Since previous research on DAE-GP used a single hidden layers for the airfoil dataset {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{24}{]}, I repeated the experiment with a single hidden layer.
+Since previous research on DAE-GP used a single hidden layers for the airfoil dataset {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{25}{]}, I repeated the experiment with a single hidden layer.
 Based on the findings of {[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{5}{]}, I expected that pre-training DAE-GP would lead to a decrease in overall performance.
@@
 
-The pre-training implementation in previous experiments used ramped half and half initialization {[}\protect\hyperlink{ref-Koza1993GeneticP}{11}{]} to create \(\hat{P}\), this leads to 50\% of the population consisting of full grown trees.
+The pre-training implementation in previous experiments used ramped half and half initialization {[}\protect\hyperlink{ref-Koza1993GeneticP}{12}{]} to create \(\hat{P}\), this leads to 50\% of the population consisting of full grown trees.
 An alternative method of initializing \(\hat{P}\) is to use the grow initialization method where solutions are not guaranteed to be full trees.
@@
 Boston\_Housing & 506 & 13 & {[}\protect\hyperlink{ref-BOSTON_HOUSING_DATASET}{8}{]} \\
-Energy\_Cooling & 768 & 8 & {[}\protect\hyperlink{ref-ENERGY_DATASET}{19}{]} \\
-Concrete & 1030 & 8 & {[}\protect\hyperlink{ref-CONCRETE_DATASET}{26}{]} \\
+Energy\_Cooling & 768 & 8 & {[}\protect\hyperlink{ref-ENERGY_DATASET}{20}{]} \\
+Concrete & 1030 & 8 & {[}\protect\hyperlink{ref-CONCRETE_DATASET}{27}{]} \\
 \end{longtable}
 
-To test for statistical differences of the results achieved by either pre-trained or regular DAE-GP, I selected the nonparametric Mann-Whitney-U Test ({[}\protect\hyperlink{ref-mwu1}{22}{]}, {[}\protect\hyperlink{ref-mwu2}{13}{]}) to test for differences in the distribution underlying the samples sourced from my experiments.
+To test for statistical differences of the results achieved by either pre-trained or regular DAE-GP, I selected the nonparametric Mann-Whitney-U Test ({[}\protect\hyperlink{ref-mwu1}{23}{]}, {[}\protect\hyperlink{ref-mwu2}{14}{]}) to test for differences in the distribution underlying the samples sourced from my experiments.
 Additionaly, I included the effect strength of pre-training as measured by Cliffs Delta {[}\protect\hyperlink{ref-Cliff1993DominanceSO}{3}{]} into the statistical analysis.
@@
 & 2 & Test & \textbf{7.9472} & 7.9683 & 0.79 & 0.08 \\
-Energy(Cooling) & 2 & Train & \textbf{4.449} & 4.5217 & 1.00 & -0.03 \\
-& 2 & Test & 4.6258 & \textbf{4.5708} & 1.00 & -0.03 \\
+Energy(Cooling) & 2 & Train & \textbf{4.449} & 4.4627 & 0.82 & 0.07 \\
+& 2 & Test & \textbf{4.6258} & 4.6297 & 0.96 & -0.03 \\
 Concrete & 2 & Train & \textbf{17.1299} & 17.9175 & 0.51 & 0.24 \\
@@
 Boston\_Housing & 2 & 3 & 3 & 0.37 & -0.20 \\
-Energy(Cooling) & 2 & 3 & 3 & 0.18 & -0.40 \\
+Energy(Cooling) & 2 & 3 & 3 & 0.16 & -0.33 \\
 Concrete & 2 & 7 & \textbf{5} & 0.09* & -0.56 \\
@@
 Boston\_Housing & 2 & 141.8 & \textbf{65.36} & 0.00*** & -0.87 \\
-Energy(Cooling) & 2 & 144.24 & \textbf{63.5} & 0.00*** & -0.93 \\
+Energy(Cooling) & 2 & 144.24 & \textbf{66.81} & 0.00*** & -0.93 \\
 Concrete & 2 & 139.2 & \textbf{64.21} & 0.00*** & -0.87 \\
@@
 For a single hidden layer, pre-training did show a negative impact on the median final fitness of solutions found as it was suggested by {[}\protect\hyperlink{ref-pmlr-v9-erhan10a}{4}{]}.
-Since current research on DAE-GP for real-world symbolic regression relies on a single hidden layer (see {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{24}{]}), pre-training does come with additional computational expenses for using deeper DAE-LSTMs.
+Since current research on DAE-GP for real-world symbolic regression relies on a single hidden layer (see {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{25}{]}), pre-training does come with additional computational expenses for using deeper DAE-LSTMs.
 
@@
 Also I did not perform an extended optimization of hyperparameters (e.g by using a grid search strategy) which might yield configurations that increase the benefits of pre-training.
-As described in subsection {[}Effect on Run-Time{]}, sampling new solutions has been a large contributor to the large performance hit on run-time by using pre-training.
+As described in subsection \protect\hyperlink{influence-on-run-time}{Influence on Run-Time}, sampling new solutions has been a large contributor to the large performance hit on run-time by using pre-training.
 It could be a worthwhile endevour to explore the main factors that increase sampling time for pre-trained DAE-GP to more efficiently exploit the benefits of pre-training like reducing the number of training epochs.
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Kingma2014AdamAM}{}}%
+\CSLLeftMargin{{[}11{]} }%
+\CSLRightInline{Diederik P. Kingma and Jimmy Ba. 2014. Adam: A method for stochastic optimization. \emph{CoRR} abs/1412.6980, (2014).}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Koza1993GeneticP}{}}%
-\CSLLeftMargin{{[}11{]} }%
+\CSLLeftMargin{{[}12{]} }%
 \CSLRightInline{John R. Koza. 1993. Genetic programming - on the programming of computers by means of natural selection. In \emph{Complex adaptive systems}.}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-NormLevDistance}{}}%
-\CSLLeftMargin{{[}12{]} }%
+\CSLLeftMargin{{[}13{]} }%
 \CSLRightInline{Joseph B.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-mwu2}{}}%
-\CSLLeftMargin{{[}13{]} }%
+\CSLLeftMargin{{[}14{]} }%
 \CSLRightInline{Henry B.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-survey_transfer_learning}{}}%
-\CSLLeftMargin{{[}14{]} }%
+\CSLLeftMargin{{[}15{]} }%
 \CSLRightInline{Sinno Jialin Pan and Qiang Yang.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-10.1007ux2f978-3-540-24621-3_22}{}}%
-\CSLLeftMargin{{[}15{]} }%
+\CSLLeftMargin{{[}16{]} }%
 \CSLRightInline{Grégory Paris, Denis Robilliard, and Cyril Fonlupt.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-harmless_overfitting_eda}{}}%
-\CSLLeftMargin{{[}16{]} }%
+\CSLLeftMargin{{[}17{]} }%
 \CSLRightInline{Malte Probst and Franz Rothlauf.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-design_of_modern_heuristics}{}}%
-\CSLLeftMargin{{[}17{]} }%
+\CSLLeftMargin{{[}18{]} }%
 \CSLRightInline{Franz Rothlauf. 2011. \emph{Design of modern heuristics: Principles and application}. DOI:https://doi.org/\href{https://doi.org/10.1007/978-3-540-72962-4}{10.1007/978-3-540-72962-4}}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-sampling_err_gp}{}}%
-\CSLLeftMargin{{[}18{]} }%
+\CSLLeftMargin{{[}19{]} }%
 \CSLRightInline{Dirk Schweim, David Wittenberg, and Franz Rothlauf.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ENERGY_DATASET}{}}%
-\CSLLeftMargin{{[}19{]} }%
+\CSLLeftMargin{{[}20{]} }%
 \CSLRightInline{Athanasios Tsanas and Angeliki Xifara.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-dae_orig2008}{}}%
-\CSLLeftMargin{{[}20{]} }%
+\CSLLeftMargin{{[}21{]} }%
 \CSLRightInline{Pascal Vincent, Hugo Larochelle, Y.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-weigend1994overfitting}{}}%
-\CSLLeftMargin{{[}21{]} }%
+\CSLLeftMargin{{[}22{]} }%
 \CSLRightInline{Andreas Weigend. 1994. On overfitting and the effective number of hidden units. In \emph{Proceedings of the 1993 connectionist models summer school}, 335--342.}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-mwu1}{}}%
-\CSLLeftMargin{{[}22{]} }%
+\CSLLeftMargin{{[}23{]} }%
 \CSLRightInline{Frank. Wilcoxon. 1945. Individual comparisons by ranking methods. \emph{Biometrics} 1, (1945), 196--202.}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-daegp_explore_exploit}{}}%
-\CSLLeftMargin{{[}23{]} }%
+\CSLLeftMargin{{[}24{]} }%
 \CSLRightInline{David Wittenberg.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-dae-gp_2022_symreg}{}}%
-\CSLLeftMargin{{[}24{]} }%
+\CSLLeftMargin{{[}25{]} }%
 \CSLRightInline{David Wittenberg and Franz Rothlauf.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-dae-gp_2020_rtree}{}}%
-\CSLLeftMargin{{[}25{]} }%
+\CSLLeftMargin{{[}26{]} }%
 \CSLRightInline{David Wittenberg, Franz Rothlauf, and Dirk Schweim.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-CONCRETE_DATASET}{}}%
-\CSLLeftMargin{{[}26{]} }%
+\CSLLeftMargin{{[}27{]} }%
 \CSLRightInline{I-Cheng Yeh.

### 2023-02-06 paper.tex
Mon Feb  6 09:00:07 CET 2023: update-remote.sh

@@
 
-Denoising Autoencoder Genetic Programming (DAE-GP) is a novel variation of an genetic programming based Estimation of Distribution Algorithm (EDA-GP) that uses a denoising autoencoders long short-term memory network (DAE-LSTMs) as a probabilistic model to sample new candidate solutions {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{26}{]}.
+Denoising Autoencoder Genetic Programming (DAE-GP) is a novel variation of an genetic programming based Estimation of Distribution Algorithm (EDA-GP) that uses a denoising autoencoders long short-term memory network (DAE-LSTMs) as a probabilistic model to sample new candidate solutions {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{27}{]}.
 
@@
 In DAE-GP, DAE-LSTMs are trained with a subset of high-fitness solutions selected from a parent population with the aim to capture their promising qualities.
-The resulting model is then used to sample new offspring solutions by propagating partially mutated solutions from the parent population through the DAE-LSTMS {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{26}{]}.
-In previous work DAE-GP has been shown to outperform GP for both a generalized version of the royal tree problem {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{26}{]} as well as for a real-world symbolic regression problem {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{25}{]}.
-
-The DAE-GP algorithm first described by {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{26}{]} trains a DAE-LSTMs for each generation \(g\) of the search from scratch.
-{[}\protect\hyperlink{ref-dae-gp_2022_symreg}{25}{]} and {[}\protect\hyperlink{ref-daegp_explore_exploit}{24}{]} suggests the incorporation of a pre-training strategy into the evolutionary search as a possible way of improving the overall performance of the DAE-GP algorithm.
+The resulting model is then used to sample new offspring solutions by propagating partially mutated solutions from the parent population through the DAE-LSTMS {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{27}{]}.
+In previous work DAE-GP has been shown to outperform GP for both a generalized version of the royal tree problem {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{27}{]} as well as for a real-world symbolic regression problem {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{26}{]}.
+
+The DAE-GP algorithm first described by {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{27}{]} trains a DAE-LSTMs for each generation \(g\) of the search from scratch.
+{[}\protect\hyperlink{ref-dae-gp_2022_symreg}{26}{]} and {[}\protect\hyperlink{ref-daegp_explore_exploit}{25}{]} suggests the incorporation of a pre-training strategy into the evolutionary search as a possible way of improving the overall performance of the DAE-GP algorithm.
 The key idea is to pre-train an initial DAE-LSTM on a large population of candidate solutions and to use the pre-trained parameters of this initial model in each generation of the search as starting parameters for the current generation DAE-LSTM.
@@
 
-DAE-GP is an EDA-GP algorithm that uses DAE-LSTM networks as a probabilistic model to sample new offspring solutions {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{26}{]}.
+DAE-GP is an EDA-GP algorithm that uses DAE-LSTM networks as a probabilistic model to sample new offspring solutions {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{27}{]}.
 
@@
 Another important concept in EC is the distinction between genotypes and phenotypes of solutions, the genotype contains the information that is necessary to construct the phenotype, the outer appearance of a particular solution on which we measure the overall quality of solutions.
-The representation of a solution is therefore defined by the mapping of genotypes to phenotypes {[}\protect\hyperlink{ref-design_of_modern_heuristics}{18}{]}.
+The representation of a solution is therefore defined by the mapping of genotypes to phenotypes {[}\protect\hyperlink{ref-design_of_modern_heuristics}{19}{]}.
 Genetic operators, such as mutation or recombination are usually applied to the genotype of solutions.
@@
 GP follows the same basic evolutionary principle of EC but searches for more general, hierarchical computer programs of dynamically varying size and shape {[}\protect\hyperlink{ref-Koza1993GeneticP}{12}{]}.
-The computer programs that are at the center of the evolutionary search in GP are commonly represented by tree structures at the level of their phenotype {[}\protect\hyperlink{ref-design_of_modern_heuristics}{18}{]}.
+The computer programs that are at the center of the evolutionary search in GP are commonly represented by tree structures at the level of their phenotype {[}\protect\hyperlink{ref-design_of_modern_heuristics}{19}{]}.
 Since GP searches for high fitness computer programs that produce a desired output for some input, it can be applied to various different problem domains such as symbolic regression, automatic programming, or evolving game-playing strategies {[}\protect\hyperlink{ref-Koza1993GeneticP}{12}{]}.
@@
 GP is an especially useful meta heuristic for problems where no a priori knowledge about the final form of good solutions is available.
-GPs ability to optimize solutions for their structure as well as for their parameters led to it being one of the most prevalent methods used for symbolic regression {[}\protect\hyperlink{ref-10.1007ux2f978-3-540-24621-3_22}{16}{]}.
+GPs ability to optimize solutions for their structure as well as for their parameters led to it being one of the most prevalent methods used for symbolic regression {[}\protect\hyperlink{ref-10.1007ux2f978-3-540-24621-3_22}{17}{]}.
 
@@
 
-The aim of Estimation of distribution algorithms (EDA) is to replace the standard variation operators used in GA by building probabilistic models that can capture complex dependencies between different decision variables of an optimization problem {[}\protect\hyperlink{ref-design_of_modern_heuristics}{18}{]}.
+The aim of Estimation of distribution algorithms (EDA) is to replace the standard variation operators used in GA by building probabilistic models that can capture complex dependencies between different decision variables of an optimization problem {[}\protect\hyperlink{ref-design_of_modern_heuristics}{19}{]}.
 EDAs use this probabilistic model to sample new offspring solutions inside an evolutionary search to replace crossover and/or mutation operators.
@@
 
-One possible way of model building in EDA proposed by {[}\protect\hyperlink{ref-harmless_overfitting_eda}{17}{]} is to use denoising autoencoders (DAEs) as generative models to capture complex probability distributions of decision variables.
+One possible way of model building in EDA proposed by {[}\protect\hyperlink{ref-harmless_overfitting_eda}{18}{]} is to use denoising autoencoders (DAEs) as generative models to capture complex probability distributions of decision variables.
 DAE, a variation of the autoencoders (AE), are a widely used type of neural networks in the domain of unsupervised machine learning that maps \(n\) input variables to \(n\) output variables using a hidden representation.
@@
 
-DAE was first introduced by {[}\protect\hyperlink{ref-dae_orig2008}{21}{]} as an improved AE with the ability to learn new representations of data that is especially robust to partially corrupted input data.
+DAE was first introduced by {[}\protect\hyperlink{ref-dae_orig2008}{22}{]} as an improved AE with the ability to learn new representations of data that is especially robust to partially corrupted input data.
 DAE modifies the AE by using partially corrupted input data for the AE and training it to reconstruct the uncorrupted, original version of the input data.
 
-Since the hidden representation of DAE captures the dependency structure of the input data it can therefore also be used to generate new solutions in the context of GAs {[}\protect\hyperlink{ref-harmless_overfitting_eda}{17}{]}.
-
-DAE-GP builds upon the concept of using DAEs in EDAs described by {[}\protect\hyperlink{ref-harmless_overfitting_eda}{17}{]} and transfers the concept to the domain of GP.
-The mutation and crossover operators of standard GP are replaced by sampling new solutions from a probabilistic model that is build by training a DAEs long short-term memory (LSTM) network on a subset of high fitness solutions from the current population {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{26}{]}.
+Since the hidden representation of DAE captures the dependency structure of the input data it can therefore also be used to generate new solutions in the context of GAs {[}\protect\hyperlink{ref-harmless_overfitting_eda}{18}{]}.
+
+DAE-GP builds upon the concept of using DAEs in EDAs described by {[}\protect\hyperlink{ref-harmless_overfitting_eda}{18}{]} and transfers the concept to the domain of GP.
+The mutation and crossover operators of standard GP are replaced by sampling new solutions from a probabilistic model that is build by training a DAEs long short-term memory (LSTM) network on a subset of high fitness solutions from the current population {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{27}{]}.
 
 LSTMs are a variant of artificial neural networks first introduced by {[}\protect\hyperlink{ref-lstm_orig}{10}{]} that can store learned information over en extended time period while avoiding the problem of vanishing and/or exploding gradients.
-Since DAE-GP encodes candidate solutions as linear strings in prefix notation, the DAE in DAE-GP uses LSTMs for both encoding and decoding where the total amount of time steps \(T\) is equal to sum of the length of the input solution and the output solution {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{26}{]}.
+Since DAE-GP encodes candidate solutions as linear strings in prefix notation, the DAE in DAE-GP uses LSTMs for both encoding and decoding where the total amount of time steps \(T\) is equal to sum of the length of the input solution and the output solution {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{27}{]}.
 
@@
 
-The strategy of pre-training artificial neural networks is based on the idea of transfer learning that aims at retaining previously learned knowledge from one task to re-use it for another task {[}\protect\hyperlink{ref-HAN2021225}{7}{]} {[}\protect\hyperlink{ref-survey_transfer_learning}{15}{]}.
+The strategy of pre-training artificial neural networks is based on the idea of transfer learning that aims at retaining previously learned knowledge from one task to re-use it for another task {[}\protect\hyperlink{ref-HAN2021225}{7}{]} {[}\protect\hyperlink{ref-survey_transfer_learning}{16}{]}.
 Transfer learning traditionally follows a two phase approach:
@@
 
-where source task and target task are usually similar but may differ in their feature space and the distribution of training data {[}\protect\hyperlink{ref-survey_transfer_learning}{15}{]}.
+where source task and target task are usually similar but may differ in their feature space and the distribution of training data {[}\protect\hyperlink{ref-survey_transfer_learning}{16}{]}.
 
@@
 Instead of initializing and optimizing each model from scratch as done in previous work (e.g.
-{[}\protect\hyperlink{ref-dae-gp_2020_rtree}{26}{]} {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{25}{]}), a separate DAE-LSTM network \(\hat{M}\) will be trained on a large population of randomly initialized solutions.\\
+{[}\protect\hyperlink{ref-dae-gp_2020_rtree}{27}{]} {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{26}{]}), a separate DAE-LSTM network \(\hat{M}\) will be trained on a large population of randomly initialized solutions.\\
 The trainable parameters \(\theta_{\hat{M}}\) obtained after finishing the training procedure of \(\hat{M}\) are then used as the starting parameters for each following DAE-LSTM model \(M_g\) for \(g\in{1,...,g_{max}}\).
@@
 Another motivation for introducing pre-training into DAE-GP is the prevalence of sampling error for small GP populations sizes.
-{[}\protect\hyperlink{ref-sampling_err_gp}{19}{]} finds that sampling error, non-systematic errors that are caused by observing only a small subset of the statistical population, is a severe problem in the domain of GP since the initial population may not be a representative sample of all possible solutions.
-{[}\protect\hyperlink{ref-sampling_err_gp}{19}{]} also introduces a method for calculating optimal population sizes to minimize the presence of sampling error.
+{[}\protect\hyperlink{ref-sampling_err_gp}{20}{]} finds that sampling error, non-systematic errors that are caused by observing only a small subset of the statistical population, is a severe problem in the domain of GP since the initial population may not be a representative sample of all possible solutions.
+{[}\protect\hyperlink{ref-sampling_err_gp}{20}{]} also introduces a method for calculating optimal population sizes to minimize the presence of sampling error.
 Since DAE-LSTM of DAE-GP learns the properties of it's training population and reuses the acquired knowledge in the sampling procedure, using the parameters obtained from training \(\hat{M}\) on a sufficiently large training population might reduce the need for large population sizes in the following generations of the search if \(\hat{M}\) already implicitly captured the properties of a representative sample of solutions.
@@
 
-The DAE-GP algorithm, first described by {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{26}{]}, is summarized as a flowchart in figure \ref{fig:daegp-flowchart}.
+The DAE-GP algorithm, first described by {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{27}{]}, is summarized as a flowchart in figure \ref{fig:daegp-flowchart}.
 After setting the generation counter \(g\) to \(0\) and creating the initial population of solutions \(P_0\), DAE-GP enters the main loop that checks if a termination criteria is satisfied (i.e.~maximum number of generations or fitness evaluations is reached).
@@
 One difficulty in the implementation of a pre-training strategy into DAE-GP has been the determination of the number of hidden neurons in side the DAE-LSTMs hidden layers.
-The original description of DAE-GP {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{26}{]} uses a strategy where the number of hidden neurons for each hidden layer is dynamically set per generation to the maximum individual size inside the current population.
+The original description of DAE-GP {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{27}{]} uses a strategy where the number of hidden neurons for each hidden layer is dynamically set per generation to the maximum individual size inside the current population.
 For a pre-training implementation, this strategy can not be easily adapted since it leads to a changing number of neurons at each generation resulting in different dimensions of the DAE-LSTM.
@@
 Train/Test Split & 50:50 \\
-Maximum Tree Depth & 17 {[}\protect\hyperlink{ref-DEAP_JMLR2012}{6}{]} \\
+Maximum Tree Depth & 17 {[}\protect\hyperlink{ref-Koza1993GeneticP}{12}{]} \\
 Initialization Algorithm & Ramped Half and Half {[}\protect\hyperlink{ref-Koza1993GeneticP}{12}{]} \\
@@
 Selection Operator & Binary Tournament Selection \\
-Function Set & \{+ , - , * , aq\} \\
+Function Set & \{+ , - , * , AQ {[}\protect\hyperlink{ref-AQ}{15}{]}\} \\
 Ephemeral Constants & {[}-5,..,5{]} \\
@@
 \endlastfoot
-Stopping Condition & Convergence {[}\protect\hyperlink{ref-daegp_explore_exploit}{24}{]} \\
+Stopping Condition & Convergence {[}\protect\hyperlink{ref-daegp_explore_exploit}{25}{]} \\
 Reconstruction Error Metric & Multiclass Cross Entropy \\
@@
 Hidden Neurons & 150 \\
-Corruption Operator & Levenshtein Edit {[}\protect\hyperlink{ref-daegp_explore_exploit}{24}{]} \\
+Corruption Operator & Levenshtein Edit {[}\protect\hyperlink{ref-daegp_explore_exploit}{25}{]} \\
 Denoising Edit Probability (Model Building) & 0.05 \\
@@
 Symbolic Regression problems have been one of the first GP applications {[}\protect\hyperlink{ref-Koza1993GeneticP}{12}{]} and are an actively studied and highly relevant research area.
-The goal in symbolic regression is to find a mathematical model for a given set of data points {[}\protect\hyperlink{ref-10.1007ux2f978-3-540-24621-3_22}{16}{]}, in real-world symbolic regression these data points are sourced from real-world observations which in contrast to synthetic symbolic regression problem are more likely to contain random noise and bias.
+The goal in symbolic regression is to find a mathematical model for a given set of data points {[}\protect\hyperlink{ref-10.1007ux2f978-3-540-24621-3_22}{17}{]}, in real-world symbolic regression these data points are sourced from real-world observations which in contrast to synthetic symbolic regression problem are more likely to contain random noise and bias.
 Another important challenge in solving real-world symbolic regression problems is the ability for a given model to generalize, we want the final model to show high accuracy in predicting outcomes for previously unseen cases.
@@
 The objective of the airfoil problem is to find a function that accurately predicts the output variable \(y\) by taking in a subset of the input variables \(x1,x2,x3,x4,x5\).
-The function set used by all DAE-GP variations for symbolic regression is summarized in tables \ref{tab:baseParams-GP} and , the terminal set consists of the the input variables \(x1,x2,x3,x4,x5\) and the ephemeral integer constants listed in table \ref{tab:baseParams-GP} {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{25}{]}.
+The function set used by all DAE-GP variations for symbolic regression is summarized in tables \ref{tab:baseParams-GP} and , the terminal set consists of the the input variables \(x1,x2,x3,x4,x5\) and the ephemeral integer constants listed in table \ref{tab:baseParams-GP} {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{26}{]}.
 
@@
 For each configuration I tested traditional DAE-GP as well as a pre-trained DAE-GP resulting in 12 total sub experiments that were each based on 10 individual runs (total number of runs=\(120\)).
-To avoid creating biased results through the presence of sampling error inside the pre-training population (see {[}\protect\hyperlink{ref-sampling_err_gp}{19}{]}), the population size for the pre-training phase is chosen very high with the size of \(\hat{P}\) =10000 where 50\% of \(\hat{P}\) is used for the training population \(\hat{P}_{train}\) and the remaining 50\% are used for the hold-out validation population \(\hat{P}_{test}\).
+To avoid creating biased results through the presence of sampling error inside the pre-training population (see {[}\protect\hyperlink{ref-sampling_err_gp}{20}{]}), the population size for the pre-training phase is chosen very high with the size of \(\hat{P}\) =10000 where 50\% of \(\hat{P}\) is used for the training population \(\hat{P}_{train}\) and the remaining 50\% are used for the hold-out validation population \(\hat{P}_{test}\).
 The training of DAE-LSTM uses a fixed number of 1000 epochs until termination and uses Adam optimization for gradient descent.
@@
 In general I expect that with a growing number of trainable parameters (either by adding more hidden layers or more hidden neurons per layer) the DAE-LSTM will be more prone to overfit to the training population \(P_{train}\) resulting in a small reconstruction error for the training population and a large one for the validation population \(P_{test}\).
-The reason for this effect is that a larger network trained over an extended period of time (without the use of strategies like early stopping), has much more potential to learn noise from the training dataset than a smaller network, which is more likely to result in worsening performance for previously unseen cases {[}\protect\hyperlink{ref-weigend1994overfitting}{22}{]}.
+The reason for this effect is that a larger network trained over an extended period of time (without the use of strategies like early stopping), has much more potential to learn noise from the training dataset than a smaller network, which is more likely to result in worsening performance for previously unseen cases {[}\protect\hyperlink{ref-weigend1994overfitting}{23}{]}.
 
@@
 
-Since previous work by {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{25}{]} demonstrated that DAE-GP, for a given number of 10.000 fitness evaluations, is able to produce solutions that are much smaller in tree size than regular GP, I also studied the influence of pre-training on the size of solutions found by DAE-GP.
+Since previous work by {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{26}{]} demonstrated that DAE-GP, for a given number of 10.000 fitness evaluations, is able to produce solutions that are much smaller in tree size than regular GP, I also studied the influence of pre-training on the size of solutions found by DAE-GP.
 Figure \ref{fig:airfoil-size-combined-30gens} shows the median of both the average size of individuals inside the population as well as the size of the currently best performing solution inside the population.
@@
 To ensure a fair comparison between pre-trained and regular DAE-GP, the prior experiments used a static number of hidden neurons for each hidden Layer inside the DAE-LSTM networks.
-For comparison, figure \ref{fig:airfoil-maxIndSize-params} shows the total number of trainable parameters for each generation's DAE-LSTM of another experiment where regular DAE-GP was tested with dynamic adjustment of the number of hidden neurons as described in {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{26}{]} and {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{25}{]}.
+For comparison, figure \ref{fig:airfoil-maxIndSize-params} shows the total number of trainable parameters for each generation's DAE-LSTM of another experiment where regular DAE-GP was tested with dynamic adjustment of the number of hidden neurons as described in {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{27}{]} and {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{26}{]}.
 Again, the experiments was based on \(10\) individual runs and otherwise used the same hyperparameters as listed in tables \ref{tab:baseParams-GP} and \ref{tab:baseParams-DAEGP}.
@@
 
-Since previous research on DAE-GP used a single hidden layers for the airfoil dataset {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{25}{]}, I repeated the experiment with a single hidden layer.
+Since previous research on DAE-GP used a single hidden layers for the airfoil dataset {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{26}{]}, I repeated the experiment with a single hidden layer.
 Based on the findings of {[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{5}{]}, I expected that pre-training DAE-GP would lead to a decrease in overall performance.
@@
 Boston\_Housing & 506 & 13 & {[}\protect\hyperlink{ref-BOSTON_HOUSING_DATASET}{8}{]} \\
-Energy\_Cooling & 768 & 8 & {[}\protect\hyperlink{ref-ENERGY_DATASET}{20}{]} \\
-Concrete & 1030 & 8 & {[}\protect\hyperlink{ref-CONCRETE_DATASET}{27}{]} \\
+Energy\_Cooling & 768 & 8 & {[}\protect\hyperlink{ref-ENERGY_DATASET}{21}{]} \\
+Concrete & 1030 & 8 & {[}\protect\hyperlink{ref-CONCRETE_DATASET}{28}{]} \\
 \end{longtable}
 
-To test for statistical differences of the results achieved by either pre-trained or regular DAE-GP, I selected the nonparametric Mann-Whitney-U Test ({[}\protect\hyperlink{ref-mwu1}{23}{]}, {[}\protect\hyperlink{ref-mwu2}{14}{]}) to test for differences in the distribution underlying the samples sourced from my experiments.
+To test for statistical differences of the results achieved by either pre-trained or regular DAE-GP, I selected the nonparametric Mann-Whitney-U Test ({[}\protect\hyperlink{ref-mwu1}{24}{]}, {[}\protect\hyperlink{ref-mwu2}{14}{]}) to test for differences in the distribution underlying the samples sourced from my experiments.
 Additionaly, I included the effect strength of pre-training as measured by Cliffs Delta {[}\protect\hyperlink{ref-Cliff1993DominanceSO}{3}{]} into the statistical analysis.
@@
 For a single hidden layer, pre-training did show a negative impact on the median final fitness of solutions found as it was suggested by {[}\protect\hyperlink{ref-pmlr-v9-erhan10a}{4}{]}.
-Since current research on DAE-GP for real-world symbolic regression relies on a single hidden layer (see {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{25}{]}), pre-training does come with additional computational expenses for using deeper DAE-LSTMs.
+Since current research on DAE-GP for real-world symbolic regression relies on a single hidden layer (see {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{26}{]}), pre-training does come with additional computational expenses for using deeper DAE-LSTMs.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-AQ}{}}%
+\CSLLeftMargin{{[}15{]} }%
+\CSLRightInline{Ji Ni, Russ H.
+Drieberg, and Peter I.
+Rockett.
+2013.
+The use of an analytic quotient operator in genetic programming.
+\emph{IEEE Transactions on Evolutionary Computation} 17, 1 (February 2013), 146--152.
+DOI:https://doi.org/\href{https://doi.org/doi:10.1109/TEVC.2012.2195319}{doi:10.1109/TEVC.2012.2195319}}
+
 \leavevmode\vadjust pre{\hypertarget{ref-survey_transfer_learning}{}}%
-\CSLLeftMargin{{[}15{]} }%
+\CSLLeftMargin{{[}16{]} }%
 \CSLRightInline{Sinno Jialin Pan and Qiang Yang.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-10.1007ux2f978-3-540-24621-3_22}{}}%
-\CSLLeftMargin{{[}16{]} }%
+\CSLLeftMargin{{[}17{]} }%
 \CSLRightInline{Grégory Paris, Denis Robilliard, and Cyril Fonlupt.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-harmless_overfitting_eda}{}}%
-\CSLLeftMargin{{[}17{]} }%
+\CSLLeftMargin{{[}18{]} }%
 \CSLRightInline{Malte Probst and Franz Rothlauf.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-design_of_modern_heuristics}{}}%
-\CSLLeftMargin{{[}18{]} }%
+\CSLLeftMargin{{[}19{]} }%
 \CSLRightInline{Franz Rothlauf. 2011. \emph{Design of modern heuristics: Principles and application}. DOI:https://doi.org/\href{https://doi.org/10.1007/978-3-540-72962-4}{10.1007/978-3-540-72962-4}}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-sampling_err_gp}{}}%
-\CSLLeftMargin{{[}19{]} }%
+\CSLLeftMargin{{[}20{]} }%
 \CSLRightInline{Dirk Schweim, David Wittenberg, and Franz Rothlauf.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ENERGY_DATASET}{}}%
-\CSLLeftMargin{{[}20{]} }%
+\CSLLeftMargin{{[}21{]} }%
 \CSLRightInline{Athanasios Tsanas and Angeliki Xifara.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-dae_orig2008}{}}%
-\CSLLeftMargin{{[}21{]} }%
+\CSLLeftMargin{{[}22{]} }%
 \CSLRightInline{Pascal Vincent, Hugo Larochelle, Y.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-weigend1994overfitting}{}}%
-\CSLLeftMargin{{[}22{]} }%
+\CSLLeftMargin{{[}23{]} }%
 \CSLRightInline{Andreas Weigend. 1994. On overfitting and the effective number of hidden units. In \emph{Proceedings of the 1993 connectionist models summer school}, 335--342.}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-mwu1}{}}%
-\CSLLeftMargin{{[}23{]} }%
+\CSLLeftMargin{{[}24{]} }%
 \CSLRightInline{Frank. Wilcoxon. 1945. Individual comparisons by ranking methods. \emph{Biometrics} 1, (1945), 196--202.}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-daegp_explore_exploit}{}}%
-\CSLLeftMargin{{[}24{]} }%
+\CSLLeftMargin{{[}25{]} }%
 \CSLRightInline{David Wittenberg.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-dae-gp_2022_symreg}{}}%
-\CSLLeftMargin{{[}25{]} }%
+\CSLLeftMargin{{[}26{]} }%
 \CSLRightInline{David Wittenberg and Franz Rothlauf.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-dae-gp_2020_rtree}{}}%
-\CSLLeftMargin{{[}26{]} }%
+\CSLLeftMargin{{[}27{]} }%
 \CSLRightInline{David Wittenberg, Franz Rothlauf, and Dirk Schweim.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-CONCRETE_DATASET}{}}%
-\CSLLeftMargin{{[}27{]} }%
+\CSLLeftMargin{{[}28{]} }%
 \CSLRightInline{I-Cheng Yeh.

### 2023-02-06 data/summarize_data.py

@@ -157,6 +157,8 @@ for problem,path in RESULTS.items():
 with open("/Users/rmn/github/master_thesis/data/summary_table_final_fit_median.csv", "w", encoding="utf-8") as f:
     f.write(csv_string)
 
+# over gens
+# ---
 
 fig, ((ul, ur), (dl, dr)) = plt.subplots(nrows=2, ncols=2, sharex=True, sharey=False, dpi=DPI, layout="constrained")
 fig.set_size_inches(14,12)
@@ -176,6 +178,62 @@ for (problem,path), ax in zip(RESULTS_2hl.items(), (ul, ur, dl, dr)):
 
 ur.legend(fontsize=MID)
 fig.savefig(f"{IMG_PATH}/median_fitness_byGens.png")
+
+# fitness boxplot
+# ----
+
+def last_fits(arr):
+    ret = []
+    for run in arr:
+        ret.append(run[-1])
+    return ret
+
+
+
+fig, ((ul, ur), (dl, dr)) = plt.subplots(nrows=2, ncols=2, sharex=False, sharey=False, dpi=DPI, layout="constrained")
+fig.set_size_inches(14,12)
+# fig.suptitle(f"{BASE_TITLE} - Median Best Fitness", fontsize=BIG)
+# fig.supxlabel("Generations", fontsize=MID)
+fig.supylabel("RMSE", fontsize=MID)
+
+for (problem,path), ax in zip(RESULTS_2hl.items(), (ul, ur, dl, dr)):
+
+    d = json.load(open(join(path, FILE_FULL_FITNESS_DATA),"r",encoding="utf-8"))
+    ax.set_title(problem, fontsize=MID)
+
+    LABELS = ["DAE-GP(Train)", "DAE-GP(Test)", "Pre-Trained(Train)", "Pre-Trained(Test)"]
+    X = [
+        last_fits(d["DAE-GP (train)"]),
+        last_fits(d["DAE-GP (test)"]),
+        last_fits(d["Pre-Trained (train)"]),
+        last_fits(d["Pre-Trained (test)"])
+    ]
+
+    # std_dev = np.std(X, 1)
+    # medians = np.median(X,1)
+
+    # fig.suptitle(f"{BASE_TITLE} - Best Fitness after 30 gens", fontsize=BIG)
+
+    bp_dict = ax.boxplot(
+        x=X,
+        labels=LABELS,
+
+    )
+    ax.grid()
+
[…]

### 2023-02-06 paper.Rmd
autogit

@@
 The results for the median best fitness over the evolutionary search are visualized in figure \@ref(fig:symReg-Fitness).
+Figure \@ref(fig:symReg-Fitness-BoxPlots) shows the distribution of final fitness values on both the test and training set as boxplots.
 
@@
 knitr::include_graphics("~/masterThesis/master_thesis/img/symbolicRegressionSummarized/median_fitness_byGens.png")
+```
+
+```{r symReg-Fitness-BoxPlots, echo=FALSE, fig.cap="Fitness after 30 Generations - Real World Symbolic Regression", out.width="90%"}
+knitr::include_graphics("~/masterThesis/master_thesis/img/symbolicRegressionSummarized/final_fitness_boxplot.png")
 ```

### 2023-02-06 paper.tex
autogit

@@
 \author{}
-\date{\vspace{-2.5em}Date of Submission: 2023-02-05}
+\date{\vspace{-2.5em}Date of Submission: 2023-02-06}
 
@@
 The results for the median best fitness over the evolutionary search are visualized in figure \ref{fig:symReg-Fitness}.
+Figure \ref{fig:symReg-Fitness-BoxPlots} shows the distribution of final fitness values on both the test and training set as boxplots.
 
@@
 \caption{Fitness over 30 Generations - Real World Symbolic Regression}\label{fig:symReg-Fitness}
+\end{figure}
+
+\begin{figure}[c]
+
+{\centering \includegraphics[width=0.9\linewidth]{img/symbolicRegressionSummarized/final_fitness_boxplot}
+
+}
+
+\caption{Fitness after 30 Generations - Real World Symbolic Regression}\label{fig:symReg-Fitness-BoxPlots}
 \end{figure}
@@
 & 2 & Test & \textbf{7.9472} & 7.9683 & 0.79 & 0.08 \\
-Energy(Cooling) & 2 & Train & \textbf{4.449} & 4.4627 & 0.82 & 0.07 \\
-& 2 & Test & \textbf{4.6258} & 4.6297 & 0.96 & -0.03 \\
-Concrete & 2 & Train & \textbf{17.1299} & 17.9175 & 0.51 & 0.24 \\
-& 2 & Test & \textbf{17.1585} & 17.6584 & 0.68 & 0.16 \\
+Energy(Cooling) & 2 & Train & \textbf{4.449} & 4.4668 & 0.54 & 0.17 \\
+& 2 & Test & \textbf{4.6258} & 4.6429 & 0.76 & 0.09 \\
+Concrete & 2 & Train & 17.1299 & \textbf{16.9825} & 0.91 & -0.04 \\
+& 2 & Test & 17.1585 & \textbf{16.7413} & 0.62 & -0.14 \\
 \end{longtable}
@@
 Boston\_Housing & 2 & 3 & 3 & 0.37 & -0.20 \\
-Energy(Cooling) & 2 & 3 & 3 & 0.16 & -0.33 \\
-Concrete & 2 & 7 & \textbf{5} & 0.09* & -0.56 \\
+Energy(Cooling) & 2 & 3 & 3 & 0.21 & -0.28 \\
+Concrete & 2 & 7 & \textbf{5} & 0.21 & -0.33 \\
 \end{longtable}
@@
 Energy(Cooling) & 1 & 0.87 & \textbf{0.94} & 0.00*** & 0.94 \\
-Concrete & 1 & 0.87 & \textbf{0.94} & 0.00*** & 0.94 \\
+Concrete & 1 & 0.87 & \textbf{0.94} & 0.00*** & 0.93 \\
 \end{longtable}
@@
 Boston\_Housing & 2 & 141.8 & \textbf{65.36} & 0.00*** & -0.87 \\
-Energy(Cooling) & 2 & 144.24 & \textbf{66.81} & 0.00*** & -0.93 \\
-Concrete & 2 & 139.2 & \textbf{64.21} & 0.00*** & -0.87 \\
+Energy(Cooling) & 2 & 144.24 & \textbf{66.7} & 0.00*** & -0.93 \\
+Concrete & 2 & 139.2 & \textbf{65.15} & 0.00*** & -0.87 \\
 \end{longtable}

### 2023-02-06 paper.Rmd
Mon Feb  6 12:00:07 CET 2023: update-remote.sh
Mon Feb  6 13:00:08 CET 2023: update-remote.sh

@@
 Again, the experiments was based on $10$ individual runs and otherwise used the same hyperparameters as listed in tables \@ref(tab:baseParams-GP) and \@ref(tab:baseParams-DAEGP).
-It clearly shows that DAE-GP can strongly reduce the number of trainable parameters by adjusting the number of hidden neurons per hidden layers to the current generations population which largley reduces the amount of training necessary and likely reduce the run-time of DAE-GP.
+It clearly shows that DAE-GP can strongly reduce the number of trainable parameters by adjusting the number of hidden neurons per hidden layers to the current generations population which largely reduces the amount of training necessary and likely reduce the run-time of DAE-GP.
 
@@
 Figure \@ref(fig:airfoil-maxIndSize-fitness) shows that pre-trained DAE-GP, with it's more complex DAE-LSTMs over the full 30 generations, does achieve a higher median fitness than regular DAE-GP with dynamic adjustment of hidden neurons.
-Pre-Training does therefore introduce a new Trade-Off decision into the adjustment of DAE-GP hyperparameters where the number of hidden neurons needs to be carefully selected since it positivly impacts solution quality but also increases the computational cost of DAE-LSTM training.
+Pre-Training does therefore introduce a new Trade-Off decision into the adjustment of DAE-GP hyperparameters where the number of hidden neurons needs to be carefully selected since it positively impacts solution quality but also increases the computational cost of DAE-LSTM training.
 
@@
 To test for statistical differences of the results achieved by either pre-trained or regular DAE-GP, I selected the nonparametric Mann-Whitney-U Test ([@mwu1], [@mwu2]) to test for differences in the distribution underlying the samples sourced from my experiments.
-Additionaly, I included the effect strength of pre-training as measured by Cliffs Delta [@Cliff1993DominanceSO] into the statistical analysis.
+Additionally, I included the effect strength of pre-training as measured by Cliffs Delta [@Cliff1993DominanceSO] into the statistical analysis.
 Better median values (e.g.
@@
 The results for the median best fitness over the evolutionary search are visualized in figure \@ref(fig:symReg-Fitness).
-Figure \@ref(fig:symReg-Fitness-BoxPlots) shows the distribution of final fitness values on both the test and training set as boxplots.
+Figure \@ref(fig:symReg-Fitness-BoxPlots) shows the distribution of final fitness values on both the test and training set as box plots.
 
@@
 The experiments conducted in the scope of this master thesis did find that the implemented pre-training strategy resulted in an increased population diversity measured by the normalized levenshtein edit distance.
-Increasing the population diversity can be an important mechanism in controlling the search behavior of population based metaheuristics and is especially useful for increasing the exploration of the search space of possible solutions !QUELLE!.
+Increasing the population diversity can be an important mechanism in controlling the search behavior of population based meta heuristics and is especially useful for increasing the exploration of the search space of possible solutions !QUELLE!.
 
@@
 
-One major limitation for the experiments conducted during the reseach phase of this thesis has been the large computational effort for experimenting with DAE-GP.
+One major limitation for the experiments conducted during the research phase of this thesis has been the large computational effort for experimenting with DAE-GP.
 Since single runs for DAE-GP have a run-time of many hours, it was not possible for me to extend the experiments to larger population sizes or more generations.
@@
 
-# References {.unnumbered #I}
+# Nomenclature {.unnumbered #I}
+
+## Symbols {.unnumbered #I}
+
+```{r Symbols, echo=FALSE, message=FALSE, warning=FALSE}
+fig.pos="H"
+knitr::kable(
+  read.csv("./tables/nomenclature.csv", sep = "?"),
+  caption = "List of Mathematical Symbols",
+  format = "markdown"
+)%>% row_spec(0,bold=TRUE )
+```
+## Abbreviations {.unnumbered #II}
+
+```{r abbreviations, echo=FALSE, message=FALSE, warning=FALSE}
+fig.pos="H"
+knitr::kable(
+  read.csv("./tables/abbreviations.csv", sep = "?"),
+  caption = "List of Abbreviations",
+  format = "markdown"
+)%>% row_spec(0,bold=TRUE )
+```
+
+\newpage
+
+# References {.unnumbered #II}
 

### 2023-02-06 paper.tex
Mon Feb  6 13:00:08 CET 2023: update-remote.sh

@@
 Again, the experiments was based on \(10\) individual runs and otherwise used the same hyperparameters as listed in tables \ref{tab:baseParams-GP} and \ref{tab:baseParams-DAEGP}.
-It clearly shows that DAE-GP can strongly reduce the number of trainable parameters by adjusting the number of hidden neurons per hidden layers to the current generations population which largley reduces the amount of training necessary and likely reduce the run-time of DAE-GP.
+It clearly shows that DAE-GP can strongly reduce the number of trainable parameters by adjusting the number of hidden neurons per hidden layers to the current generations population which largely reduces the amount of training necessary and likely reduce the run-time of DAE-GP.
 
@@
 Figure \ref{fig:airfoil-maxIndSize-fitness} shows that pre-trained DAE-GP, with it's more complex DAE-LSTMs over the full 30 generations, does achieve a higher median fitness than regular DAE-GP with dynamic adjustment of hidden neurons.
-Pre-Training does therefore introduce a new Trade-Off decision into the adjustment of DAE-GP hyperparameters where the number of hidden neurons needs to be carefully selected since it positivly impacts solution quality but also increases the computational cost of DAE-LSTM training.
+Pre-Training does therefore introduce a new Trade-Off decision into the adjustment of DAE-GP hyperparameters where the number of hidden neurons needs to be carefully selected since it positively impacts solution quality but also increases the computational cost of DAE-LSTM training.
 
@@
 To test for statistical differences of the results achieved by either pre-trained or regular DAE-GP, I selected the nonparametric Mann-Whitney-U Test ({[}\protect\hyperlink{ref-mwu1}{24}{]}, {[}\protect\hyperlink{ref-mwu2}{14}{]}) to test for differences in the distribution underlying the samples sourced from my experiments.
-Additionaly, I included the effect strength of pre-training as measured by Cliffs Delta {[}\protect\hyperlink{ref-Cliff1993DominanceSO}{3}{]} into the statistical analysis.
+Additionally, I included the effect strength of pre-training as measured by Cliffs Delta {[}\protect\hyperlink{ref-Cliff1993DominanceSO}{3}{]} into the statistical analysis.
 Better median values (e.g.~higher fitness) are printed bold and p-values are marked with asterisk (1,2,3) for \(\alpha\) levels (\(0.1\), \(0.05\), \(0.01\)).
@@
 The results for the median best fitness over the evolutionary search are visualized in figure \ref{fig:symReg-Fitness}.
-Figure \ref{fig:symReg-Fitness-BoxPlots} shows the distribution of final fitness values on both the test and training set as boxplots.
+Figure \ref{fig:symReg-Fitness-BoxPlots} shows the distribution of final fitness values on both the test and training set as box plots.
 
@@
 \hypertarget{I}{%
-\section*{References}\label{I}}
+\section*{Nomenclature}\label{I}}
+\addcontentsline{toc}{section}{Nomenclature}
+
+\hypertarget{symbols}{%
+\subsection{Symbols}\label{symbols}}
+
+\begin{longtable}[]{@{}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.2045}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.7955}}@{}}
+\caption{\label{tab:Symbols}List of Mathematical Symbols}\tabularnewline
+\toprule\noalign{}
+\begin{minipage}[b]{\linewidth}\raggedright
+Symbol
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Meaning
+\end{minipage} \\
+\midrule\noalign{}
+\endfirsthead
+\toprule\noalign{}
+\begin{minipage}[b]{\linewidth}\raggedright
+Symbol
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Meaning
+\end{minipage} \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+\(g\) & Current generation \\
+\(g_{max}\) & Maximum number of generations \\
+\(M_g\) & DAE-LSTM model for generation \(g\) \\
+\(\hat{M}\) & DAE-LSTM model for Pre-Training \\
+\(\theta_{M}\) & Trainable parameters of DAE-LSTM model \(M\) \\
+\(P_g\) & Population of generation \(g\) \\
+\(\hat{P}\) & Pre-Training Population \\
+\(\hat{P}_{train}\) & Pre-Training Population used for training DAE-LSTM model \(\hat{M}\) \\
+\(\hat{P}_{test}\) & Pre-Training Population used for validating DAE-LSTM model \(\hat{M}\) \\
+\(X_{g}\) & Selected Training Population for DAE-LSTM model \(M\) in generation \(g\) \\
+\end{longtable}
+
+\hypertarget{abbreviations}{%
+\subsection{Abbreviations}\label{abbreviations}}
+
+\begin{longtable}[]{@{}ll@{}}
+\caption{\label{tab:abbreviations}List of Abbreviations}\tabularnewline
+\toprule\noalign{}
+Abbreviation & Meaning \\
+\midrule\noalign{}
+\endfirsthead
+\toprule\noalign{}
+Abbreviation & Meaning \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+DAE-GP & Denoising Autoencoder Genetic Programming \\
+GP & Genetic Programming \\
+EDA & Estimation of Distribution Algorithm \\
+EC & Evolutionary Computation \\
+DAE & Denoising Autoencoder \\
+DAE-LSTM & Denoising Autoencoder Long Short Term Memory Network \\
+RMSE & Root Mean Squared Error \\
+AQ & Analytic Quotient \\
+\end{longtable}
+
+\newpage
+
+\hypertarget{II}{%
+\section*{References}\label{II}}
 \addcontentsline{toc}{section}{References}

### 2023-02-06 paper.Rmd
autogit

@@
 
-# Abstract
-
+# Abstract {.unnumbered #I}
 ...
@@
 
-# Nomenclature {.unnumbered #I}
+# Nomenclature {.unnumbered #II}
 
-## Symbols {.unnumbered #I}
+## Symbols {.unnumbered #IIa}
 
@@
 ```
-## Abbreviations {.unnumbered #II}
+## Abbreviations {.unnumbered #IIb}
 
@@
 
-# References {.unnumbered #II}
+# References {.unnumbered #III}
 

### 2023-02-06 paper.Rmd
Mon Feb  6 17:00:09 CET 2023: update-remote.sh

@@
 
-The framework that was used to conduct all experiments of this thesis was provided by the supervisor of this thesis David Wittenberg and uses the python programming language in conjunction with the baseline libraries `Keras` [@chollet2015keras] for deep neural networks and `deap` [@DEAP_JMLR2012] for evolutionary computation.
+The framework that was used to run all experiments of this thesis was provided by the supervisor of this thesis David Wittenberg and uses the python programming language [@python3] in conjunction with the baseline libraries `Keras` [@chollet2015keras] for deep neural networks and `deap` [@DEAP_JMLR2012] for evolutionary computation.
+For data visualizations I used the python package `matplotlib` [@matplotlib], statistical analysis and evaluations were conducted using primarily the datscience libraries `numpy` [@numpy], `pandas` [@pandas], and
 
@@
 While investigating other reasons for the differences in run-time, I verified that both DAE-GP variants use the exact same number of trainable parameters for each generations DAE-LSTMS models (total number of trainable parameters DAE-GP=`r format(REG_PARAMS, scientific=F)`, Pre-trained DAE-GP=`r format(PT_PARAMS, scientific=F)`).
-After profiling a single run of DAE-GP and pre-trained DAE-GP my best explanation for the remaining differences in run-time is an inefficiency in my implementation of pre-training.
+After profiling a single run of DAE-GP and pre-trained DAE-GP using the python `cProfiler` module, my best explanation for the remaining differences in run-time is an inefficiency in my implementation of pre-training.
 Figure  \@ref(fig:airfoil-profiling) shows the cumulative time spend for the 20 most expensive function calls.
@@
 # shorten column names
-col_names <- c("Problem", "Hid.Layers", "DAE-GP", "Pre-Trained", "P-Value", "Cliffs-Delta")
+col_names <- c("Problem", "Hidden Layers", "DAE-GP", "Pre-Trained", "P-Value", "Cliffs-Delta")
 
@@
 
-<!--
-  EIDESSTAATLICHE ERKLAERUNG
-# II - Statutory Declaration {.unnumbered #II}
+\newpage
 
-![Statutory Declaration](./private/erklaerung.pdf)\
--->
+# II - Statutory Declaration {.unnumbered #IV}
+
+...hier einfügen...

### 2023-02-06 paper.tex
autogit
Mon Feb  6 17:00:09 CET 2023: update-remote.sh

@@
 
-\hypertarget{abstract}{%
-\section{Abstract}\label{abstract}}
+\hypertarget{I}{%
+\section*{Abstract}\label{I}}
+\addcontentsline{toc}{section}{Abstract}
 
@@
 
-Denoising Autoencoder Genetic Programming (DAE-GP) is a novel variation of an genetic programming based Estimation of Distribution Algorithm (EDA-GP) that uses a denoising autoencoders long short-term memory network (DAE-LSTMs) as a probabilistic model to sample new candidate solutions {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{27}{]}.
+Denoising Autoencoder Genetic Programming (DAE-GP) is a novel variation of an genetic programming based Estimation of Distribution Algorithm (EDA-GP) that uses a denoising autoencoders long short-term memory network (DAE-LSTMs) as a probabilistic model to sample new candidate solutions {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{28}{]}.
 
@@
 In DAE-GP, DAE-LSTMs are trained with a subset of high-fitness solutions selected from a parent population with the aim to capture their promising qualities.
-The resulting model is then used to sample new offspring solutions by propagating partially mutated solutions from the parent population through the DAE-LSTMS {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{27}{]}.
-In previous work DAE-GP has been shown to outperform GP for both a generalized version of the royal tree problem {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{27}{]} as well as for a real-world symbolic regression problem {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{26}{]}.
-
-The DAE-GP algorithm first described by {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{27}{]} trains a DAE-LSTMs for each generation \(g\) of the search from scratch.
-{[}\protect\hyperlink{ref-dae-gp_2022_symreg}{26}{]} and {[}\protect\hyperlink{ref-daegp_explore_exploit}{25}{]} suggests the incorporation of a pre-training strategy into the evolutionary search as a possible way of improving the overall performance of the DAE-GP algorithm.
+The resulting model is then used to sample new offspring solutions by propagating partially mutated solutions from the parent population through the DAE-LSTMS {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{28}{]}.
+In previous work DAE-GP has been shown to outperform GP for both a generalized version of the royal tree problem {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{28}{]} as well as for a real-world symbolic regression problem {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{27}{]}.
+
+The DAE-GP algorithm first described by {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{28}{]} trains a DAE-LSTMs for each generation \(g\) of the search from scratch.
+{[}\protect\hyperlink{ref-dae-gp_2022_symreg}{27}{]} and {[}\protect\hyperlink{ref-daegp_explore_exploit}{26}{]} suggests the incorporation of a pre-training strategy into the evolutionary search as a possible way of improving the overall performance of the DAE-GP algorithm.
 The key idea is to pre-train an initial DAE-LSTM on a large population of candidate solutions and to use the pre-trained parameters of this initial model in each generation of the search as starting parameters for the current generation DAE-LSTM.
@@
 
-DAE-GP is an EDA-GP algorithm that uses DAE-LSTM networks as a probabilistic model to sample new offspring solutions {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{27}{]}.
+DAE-GP is an EDA-GP algorithm that uses DAE-LSTM networks as a probabilistic model to sample new offspring solutions {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{28}{]}.
 
@@
 
-DAE was first introduced by {[}\protect\hyperlink{ref-dae_orig2008}{22}{]} as an improved AE with the ability to learn new representations of data that is especially robust to partially corrupted input data.
+DAE was first introduced by {[}\protect\hyperlink{ref-dae_orig2008}{23}{]} as an improved AE with the ability to learn new representations of data that is especially robust to partially corrupted input data.
 DAE modifies the AE by using partially corrupted input data for the AE and training it to reconstruct the uncorrupted, original version of the input data.
@@
 DAE-GP builds upon the concept of using DAEs in EDAs described by {[}\protect\hyperlink{ref-harmless_overfitting_eda}{18}{]} and transfers the concept to the domain of GP.
-The mutation and crossover operators of standard GP are replaced by sampling new solutions from a probabilistic model that is build by training a DAEs long short-term memory (LSTM) network on a subset of high fitness solutions from the current population {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{27}{]}.
+The mutation and crossover operators of standard GP are replaced by sampling new solutions from a probabilistic model that is build by training a DAEs long short-term memory (LSTM) network on a subset of high fitness solutions from the current population {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{28}{]}.
 
 LSTMs are a variant of artificial neural networks first introduced by {[}\protect\hyperlink{ref-lstm_orig}{10}{]} that can store learned information over en extended time period while avoiding the problem of vanishing and/or exploding gradients.
-Since DAE-GP encodes candidate solutions as linear strings in prefix notation, the DAE in DAE-GP uses LSTMs for both encoding and decoding where the total amount of time steps \(T\) is equal to sum of the length of the input solution and the output solution {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{27}{]}.
+Since DAE-GP encodes candidate solutions as linear strings in prefix notation, the DAE in DAE-GP uses LSTMs for both encoding and decoding where the total amount of time steps \(T\) is equal to sum of the length of the input solution and the output solution {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{28}{]}.
 
@@
 Instead of initializing and optimizing each model from scratch as done in previous work (e.g.
-{[}\protect\hyperlink{ref-dae-gp_2020_rtree}{27}{]} {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{26}{]}), a separate DAE-LSTM network \(\hat{M}\) will be trained on a large population of randomly initialized solutions.\\
+{[}\protect\hyperlink{ref-dae-gp_2020_rtree}{28}{]} {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{27}{]}), a separate DAE-LSTM network \(\hat{M}\) will be trained on a large population of randomly initialized solutions.\\
 The trainable parameters \(\theta_{\hat{M}}\) obtained after finishing the training procedure of \(\hat{M}\) are then used as the starting parameters for each following DAE-LSTM model \(M_g\) for \(g\in{1,...,g_{max}}\).
@@
 
-The DAE-GP algorithm, first described by {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{27}{]}, is summarized as a flowchart in figure \ref{fig:daegp-flowchart}.
+The DAE-GP algorithm, first described by {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{28}{]}, is summarized as a flowchart in figure \ref{fig:daegp-flowchart}.
 After setting the generation counter \(g\) to \(0\) and creating the initial population of solutions \(P_0\), DAE-GP enters the main loop that checks if a termination criteria is satisfied (i.e.~maximum number of generations or fitness evaluations is reached).
@@
 One difficulty in the implementation of a pre-training strategy into DAE-GP has been the determination of the number of hidden neurons in side the DAE-LSTMs hidden layers.
-The original description of DAE-GP {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{27}{]} uses a strategy where the number of hidden neurons for each hidden layer is dynamically set per generation to the maximum individual size inside the current population.
+The original description of DAE-GP {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{28}{]} uses a strategy where the number of hidden neurons for each hidden layer is dynamically set per generation to the maximum individual size inside the current population.
 For a pre-training implementation, this strategy can not be easily adapted since it leads to a changing number of neurons at each generation resulting in different dimensions of the DAE-LSTM.
@@
 \endlastfoot
-Stopping Condition & Convergence {[}\protect\hyperlink{ref-daegp_explore_exploit}{25}{]} \\
+Stopping Condition & Convergence {[}\protect\hyperlink{ref-daegp_explore_exploit}{26}{]} \\
 Reconstruction Error Metric & Multiclass Cross Entropy \\
@@
 Hidden Neurons & 150 \\
-Corruption Operator & Levenshtein Edit {[}\protect\hyperlink{ref-daegp_explore_exploit}{25}{]} \\
+Corruption Operator & Levenshtein Edit {[}\protect\hyperlink{ref-daegp_explore_exploit}{26}{]} \\
 Denoising Edit Probability (Model Building) & 0.05 \\
@@
 The objective of the airfoil problem is to find a function that accurately predicts the output variable \(y\) by taking in a subset of the input variables \(x1,x2,x3,x4,x5\).
-The function set used by all DAE-GP variations for symbolic regression is summarized in tables \ref{tab:baseParams-GP} and , the terminal set consists of the the input variables \(x1,x2,x3,x4,x5\) and the ephemeral integer constants listed in table \ref{tab:baseParams-GP} {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{26}{]}.
+The function set used by all DAE-GP variations for symbolic regression is summarized in tables \ref{tab:baseParams-GP} and , the terminal set consists of the the input variables \(x1,x2,x3,x4,x5\) and the ephemeral integer constants listed in table \ref{tab:baseParams-GP} {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{27}{]}.
 
@@
 In general I expect that with a growing number of trainable parameters (either by adding more hidden layers or more hidden neurons per layer) the DAE-LSTM will be more prone to overfit to the training population \(P_{train}\) resulting in a small reconstruction error for the training population and a large one for the validation population \(P_{test}\).
-The reason for this effect is that a larger network trained over an extended period of time (without the use of strategies like early stopping), has much more potential to learn noise from the training dataset than a smaller network, which is more likely to result in worsening performance for previously unseen cases {[}\protect\hyperlink{ref-weigend1994overfitting}{23}{]}.
+The reason for this effect is that a larger network trained over an extended period of time (without the use of strategies like early stopping), has much more potential to learn noise from the training dataset than a smaller network, which is more likely to result in worsening performance for previously unseen cases {[}\protect\hyperlink{ref-weigend1994overfitting}{24}{]}.
 
@@
 
-Since previous work by {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{26}{]} demonstrated that DAE-GP, for a given number of 10.000 fitness evaluations, is able to produce solutions that are much smaller in tree size than regular GP, I also studied the influence of pre-training on the size of solutions found by DAE-GP.
+Since previous work by {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{27}{]} demonstrated that DAE-GP, for a given number of 10.000 fitness evaluations, is able to produce solutions that are much smaller in tree size than regular GP, I also studied the influence of pre-training on the size of solutions found by DAE-GP.
 Figure \ref{fig:airfoil-size-combined-30gens} shows the median of both the average size of individuals inside the population as well as the size of the currently best performing solution inside the population.
@@
 To ensure a fair comparison between pre-trained and regular DAE-GP, the prior experiments used a static number of hidden neurons for each hidden Layer inside the DAE-LSTM networks.
-For comparison, figure \ref{fig:airfoil-maxIndSize-params} shows the total number of trainable parameters for each generation's DAE-LSTM of another experiment where regular DAE-GP was tested with dynamic adjustment of the number of hidden neurons as described in {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{27}{]} and {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{26}{]}.
+For comparison, figure \ref{fig:airfoil-maxIndSize-params} shows the total number of trainable parameters for each generation's DAE-LSTM of another experiment where regular DAE-GP was tested with dynamic adjustment of the number of hidden neurons as described in {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{28}{]} and {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{27}{]}.
 Again, the experiments was based on \(10\) individual runs and otherwise used the same hyperparameters as listed in tables \ref{tab:baseParams-GP} and \ref{tab:baseParams-DAEGP}.
@@
 
-Since previous research on DAE-GP used a single hidden layers for the airfoil dataset {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{26}{]}, I repeated the experiment with a single hidden layer.
+Since previous research on DAE-GP used a single hidden layers for the airfoil dataset {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{27}{]}, I repeated the experiment with a single hidden layer.
 Based on the findings of {[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{5}{]}, I expected that pre-training DAE-GP would lead to a decrease in overall performance.
@@
 While investigating other reasons for the differences in run-time, I verified that both DAE-GP variants use the exact same number of trainable parameters for each generations DAE-LSTMS models (total number of trainable parameters DAE-GP=573473, Pre-trained DAE-GP=573473).
-After profiling a single run of DAE-GP and pre-trained DAE-GP my best explanation for the remaining differences in run-time is an inefficiency in my implementation of pre-training.
+After profiling a single run of DAE-GP and pre-trained DAE-GP using the python \texttt{cProfiler} module {[}\protect\hyperlink{ref-python3}{22}{]}, my best explanation for the remaining differences in run-time is an inefficiency in my implementation of pre-training.
 Figure \ref{fig:airfoil-profiling} shows the cumulative time spend for the 20 most expensive function calls.
@@
 Energy\_Cooling & 768 & 8 & {[}\protect\hyperlink{ref-ENERGY_DATASET}{21}{]} \\
-Concrete & 1030 & 8 & {[}\protect\hyperlink{ref-CONCRETE_DATASET}{28}{]} \\
+Concrete & 1030 & 8 & {[}\protect\hyperlink{ref-CONCRETE_DATASET}{29}{]} \\
 \end{longtable}
 
-To test for statistical differences of the results achieved by either pre-trained or regular DAE-GP, I selected the nonparametric Mann-Whitney-U Test ({[}\protect\hyperlink{ref-mwu1}{24}{]}, {[}\protect\hyperlink{ref-mwu2}{14}{]}) to test for differences in the distribution underlying the samples sourced from my experiments.
+To test for statistical differences of the results achieved by either pre-trained or regular DAE-GP, I selected the nonparametric Mann-Whitney-U Test ({[}\protect\hyperlink{ref-mwu1}{25}{]}, {[}\protect\hyperlink{ref-mwu2}{14}{]}) to test for differences in the distribution underlying the samples sourced from my experiments.
 Additionally, I included the effect strength of pre-training as measured by Cliffs Delta {[}\protect\hyperlink{ref-Cliff1993DominanceSO}{3}{]} into the statistical analysis.
@@
 \begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2388}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1642}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1045}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1791}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1194}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1940}}@{}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2286}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2000}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1000}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1714}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1143}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1857}}@{}}
 \caption{\label{tab:full-run-realWorldSymReg-best-solution-size}Median Size of the best solution after 30 generations - Real World Symbolic Regression}\tabularnewline
@@
 \end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Hid.Layers
+Hidden Layers
 \end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
@@
 \begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2388}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1642}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1045}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1791}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1194}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1940}}@{}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2286}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2000}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1000}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1714}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1143}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1857}}@{}}
 \caption{\label{tab:full-run-realWorldSymReg-popDiversity}Median Population Diversity over 30 Generations - Symbolic Regression}\tabularnewline
@@
 \begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2388}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1642}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1045}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1791}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1194}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1940}}@{}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2286}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2000}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1000}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1714}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1143}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1857}}@{}}
 \caption{\label{tab:full-run-realWorldSymReg-epochsPerGen}Median Number of Training Epochs per Generation - Symbolic Regression}\tabularnewline
@@
 For a single hidden layer, pre-training did show a negative impact on the median final fitness of solutions found as it was suggested by {[}\protect\hyperlink{ref-pmlr-v9-erhan10a}{4}{]}.
-Since current research on DAE-GP for real-world symbolic regression relies on a single hidden layer (see {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{26}{]}), pre-training does come with additional computational expenses for using deeper DAE-LSTMs.
+Since current research on DAE-GP for real-world symbolic regression relies on a single hidden layer (see {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{27}{]}), pre-training does come with additional computational expenses for using deeper DAE-LSTMs.
 
@@
 
-\hypertarget{I}{%
-\section*{Nomenclature}\label{I}}
+\hypertarget{II}{%
+\section*{Nomenclature}\label{II}}
 \addcontentsline{toc}{section}{Nomenclature}
 
-\hypertarget{symbols}{%
-\subsection{Symbols}\label{symbols}}
+\hypertarget{IIa}{%
+\subsection*{Symbols}\label{IIa}}
+\addcontentsline{toc}{subsection}{Symbols}
 
 \begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.2045}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.7955}}@{}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.1957}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.8043}}@{}}
 \caption{\label{tab:Symbols}List of Mathematical Symbols}\tabularnewline
@@
 \endlastfoot
-\(g\) & Current generation \\
+\(g_{n}\) & \(n\).th generation \\
 \(g_{max}\) & Maximum number of generations \\
@@
 \(\hat{P}_{test}\) & Pre-Training Population used for validating DAE-LSTM model \(\hat{M}\) \\
-\(X_{g}\) & Selected Training Population for DAE-LSTM model \(M\) in generation \(g\) \\
+\(X_{g}\) & Selected Training Population for DAE-LSTM model \(M_{g}\) in generation \(g\) \\
 \end{longtable}
 
-\hypertarget{abbreviations}{%
-\subsection{Abbreviations}\label{abbreviations}}
+\hypertarget{IIb}{%
+\subsection*{Abbreviations}\label{IIb}}
+\addcontentsline{toc}{subsection}{Abbreviations}
 
@@
 
-\hypertarget{II}{%
-\section*{References}\label{II}}
+\hypertarget{III}{%
+\section*{References}\label{III}}
 \addcontentsline{toc}{section}{References}
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-python3}{}}%
+\CSLLeftMargin{{[}22{]} }%
+\CSLRightInline{Guido Van Rossum and Fred L. Drake. 2009. \emph{Python 3 reference manual}. CreateSpace, Scotts Valley, CA.}
+
 \leavevmode\vadjust pre{\hypertarget{ref-dae_orig2008}{}}%
-\CSLLeftMargin{{[}22{]} }%
+\CSLLeftMargin{{[}23{]} }%
 \CSLRightInline{Pascal Vincent, Hugo Larochelle, Y.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-weigend1994overfitting}{}}%
-\CSLLeftMargin{{[}23{]} }%
+\CSLLeftMargin{{[}24{]} }%
 \CSLRightInline{Andreas Weigend. 1994. On overfitting and the effective number of hidden units. In \emph{Proceedings of the 1993 connectionist models summer school}, 335--342.}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-mwu1}{}}%
-\CSLLeftMargin{{[}24{]} }%
+\CSLLeftMargin{{[}25{]} }%
 \CSLRightInline{Frank. Wilcoxon. 1945. Individual comparisons by ranking methods. \emph{Biometrics} 1, (1945), 196--202.}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-daegp_explore_exploit}{}}%
-\CSLLeftMargin{{[}25{]} }%
+\CSLLeftMargin{{[}26{]} }%
 \CSLRightInline{David Wittenberg.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-dae-gp_2022_symreg}{}}%
-\CSLLeftMargin{{[}26{]} }%
+\CSLLeftMargin{{[}27{]} }%
 \CSLRightInline{David Wittenberg and Franz Rothlauf.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-dae-gp_2020_rtree}{}}%
-\CSLLeftMargin{{[}27{]} }%
+\CSLLeftMargin{{[}28{]} }%
 \CSLRightInline{David Wittenberg, Franz Rothlauf, and Dirk Schweim.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-CONCRETE_DATASET}{}}%
-\CSLLeftMargin{{[}28{]} }%
+\CSLLeftMargin{{[}29{]} }%
 \CSLRightInline{I-Cheng Yeh.
@@
 
+\newpage
+
+\hypertarget{IV}{%
+\section*{II - Statutory Declaration}\label{IV}}
+\addcontentsline{toc}{section}{II - Statutory Declaration}
+
+\ldots hier einfügen\ldots{}
+
 \end{document}

### 2023-02-06 paper.Rmd
Mon Feb  6 18:02:44 CET 2023: update-remote.sh

@@
 The framework that was used to run all experiments of this thesis was provided by the supervisor of this thesis David Wittenberg and uses the python programming language [@python3] in conjunction with the baseline libraries `Keras` [@chollet2015keras] for deep neural networks and `deap` [@DEAP_JMLR2012] for evolutionary computation.
-For data visualizations I used the python package `matplotlib` [@matplotlib], statistical analysis and evaluations were conducted using primarily the datscience libraries `numpy` [@numpy], `pandas` [@pandas], and
+For data visualizations I used the python package `matplotlib` [@matplotlib], statistical analysis and evaluations were conducted using primarily the data science libraries `scipy` [@scipy], `numpy` [@numpy], `pandas` [@pandas], and  `IPython` [@ipython].
 
@@
 Regarding the presence of overfitting, both DAE-GP variants show only a very small gap between the testing and training fitness which indicates a good generalization behavior.
+Another interesting observation from figure \@ref(fig:airfoil-median-fitness-30gens) is that pre-trained DAE-GP is faster at improving fitness than regular DAE-GP, this is especially true if only looking at the first 10 generations
 
@@
 
-# II - Statutory Declaration {.unnumbered #IV}
+# Statutory Declaration {.unnumbered #IV}
 

### 2023-02-06 paper.tex
Mon Feb  6 18:02:44 CET 2023: update-remote.sh

@@
 
-Denoising Autoencoder Genetic Programming (DAE-GP) is a novel variation of an genetic programming based Estimation of Distribution Algorithm (EDA-GP) that uses a denoising autoencoders long short-term memory network (DAE-LSTMs) as a probabilistic model to sample new candidate solutions {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{28}{]}.
+Denoising Autoencoder Genetic Programming (DAE-GP) is a novel variation of an genetic programming based Estimation of Distribution Algorithm (EDA-GP) that uses a denoising autoencoders long short-term memory network (DAE-LSTMs) as a probabilistic model to sample new candidate solutions {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{33}{]}.
 
@@
 In DAE-GP, DAE-LSTMs are trained with a subset of high-fitness solutions selected from a parent population with the aim to capture their promising qualities.
-The resulting model is then used to sample new offspring solutions by propagating partially mutated solutions from the parent population through the DAE-LSTMS {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{28}{]}.
-In previous work DAE-GP has been shown to outperform GP for both a generalized version of the royal tree problem {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{28}{]} as well as for a real-world symbolic regression problem {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{27}{]}.
-
-The DAE-GP algorithm first described by {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{28}{]} trains a DAE-LSTMs for each generation \(g\) of the search from scratch.
-{[}\protect\hyperlink{ref-dae-gp_2022_symreg}{27}{]} and {[}\protect\hyperlink{ref-daegp_explore_exploit}{26}{]} suggests the incorporation of a pre-training strategy into the evolutionary search as a possible way of improving the overall performance of the DAE-GP algorithm.
+The resulting model is then used to sample new offspring solutions by propagating partially mutated solutions from the parent population through the DAE-LSTMS {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{33}{]}.
+In previous work DAE-GP has been shown to outperform GP for both a generalized version of the royal tree problem {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{33}{]} as well as for a real-world symbolic regression problem {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{32}{]}.
+
+The DAE-GP algorithm first described by {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{33}{]} trains a DAE-LSTMs for each generation \(g\) of the search from scratch.
+{[}\protect\hyperlink{ref-dae-gp_2022_symreg}{32}{]} and {[}\protect\hyperlink{ref-daegp_explore_exploit}{31}{]} suggests the incorporation of a pre-training strategy into the evolutionary search as a possible way of improving the overall performance of the DAE-GP algorithm.
 The key idea is to pre-train an initial DAE-LSTM on a large population of candidate solutions and to use the pre-trained parameters of this initial model in each generation of the search as starting parameters for the current generation DAE-LSTM.
@@
 
-DAE-GP is an EDA-GP algorithm that uses DAE-LSTM networks as a probabilistic model to sample new offspring solutions {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{28}{]}.
+DAE-GP is an EDA-GP algorithm that uses DAE-LSTM networks as a probabilistic model to sample new offspring solutions {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{33}{]}.
 
@@
 Another important concept in EC is the distinction between genotypes and phenotypes of solutions, the genotype contains the information that is necessary to construct the phenotype, the outer appearance of a particular solution on which we measure the overall quality of solutions.
-The representation of a solution is therefore defined by the mapping of genotypes to phenotypes {[}\protect\hyperlink{ref-design_of_modern_heuristics}{19}{]}.
+The representation of a solution is therefore defined by the mapping of genotypes to phenotypes {[}\protect\hyperlink{ref-design_of_modern_heuristics}{22}{]}.
 Genetic operators, such as mutation or recombination are usually applied to the genotype of solutions.
@@
 
-GP follows the same basic evolutionary principle of EC but searches for more general, hierarchical computer programs of dynamically varying size and shape {[}\protect\hyperlink{ref-Koza1993GeneticP}{12}{]}.
-The computer programs that are at the center of the evolutionary search in GP are commonly represented by tree structures at the level of their phenotype {[}\protect\hyperlink{ref-design_of_modern_heuristics}{19}{]}.
-Since GP searches for high fitness computer programs that produce a desired output for some input, it can be applied to various different problem domains such as symbolic regression, automatic programming, or evolving game-playing strategies {[}\protect\hyperlink{ref-Koza1993GeneticP}{12}{]}.
+GP follows the same basic evolutionary principle of EC but searches for more general, hierarchical computer programs of dynamically varying size and shape {[}\protect\hyperlink{ref-Koza1993GeneticP}{14}{]}.
+The computer programs that are at the center of the evolutionary search in GP are commonly represented by tree structures at the level of their phenotype {[}\protect\hyperlink{ref-design_of_modern_heuristics}{22}{]}.
+Since GP searches for high fitness computer programs that produce a desired output for some input, it can be applied to various different problem domains such as symbolic regression, automatic programming, or evolving game-playing strategies {[}\protect\hyperlink{ref-Koza1993GeneticP}{14}{]}.
 An important quality of GP is the ability to search for solutions of variable length and structure.
 GP is an especially useful meta heuristic for problems where no a priori knowledge about the final form of good solutions is available.
-GPs ability to optimize solutions for their structure as well as for their parameters led to it being one of the most prevalent methods used for symbolic regression {[}\protect\hyperlink{ref-10.1007ux2f978-3-540-24621-3_22}{17}{]}.
+GPs ability to optimize solutions for their structure as well as for their parameters led to it being one of the most prevalent methods used for symbolic regression {[}\protect\hyperlink{ref-10.1007ux2f978-3-540-24621-3_22}{19}{]}.
 
@@
 
-The aim of Estimation of distribution algorithms (EDA) is to replace the standard variation operators used in GA by building probabilistic models that can capture complex dependencies between different decision variables of an optimization problem {[}\protect\hyperlink{ref-design_of_modern_heuristics}{19}{]}.
+The aim of Estimation of distribution algorithms (EDA) is to replace the standard variation operators used in GA by building probabilistic models that can capture complex dependencies between different decision variables of an optimization problem {[}\protect\hyperlink{ref-design_of_modern_heuristics}{22}{]}.
 EDAs use this probabilistic model to sample new offspring solutions inside an evolutionary search to replace crossover and/or mutation operators.
@@
 
-One possible way of model building in EDA proposed by {[}\protect\hyperlink{ref-harmless_overfitting_eda}{18}{]} is to use denoising autoencoders (DAEs) as generative models to capture complex probability distributions of decision variables.
+One possible way of model building in EDA proposed by {[}\protect\hyperlink{ref-harmless_overfitting_eda}{21}{]} is to use denoising autoencoders (DAEs) as generative models to capture complex probability distributions of decision variables.
 DAE, a variation of the autoencoders (AE), are a widely used type of neural networks in the domain of unsupervised machine learning that maps \(n\) input variables to \(n\) output variables using a hidden representation.
 
-AE were introduced by {[}\protect\hyperlink{ref-ae_orig}{9}{]} to compress high-dimensional data into lower-dimensions. An AE consists of two different sub-units:
+AE were introduced by {[}\protect\hyperlink{ref-ae_orig}{10}{]} to compress high-dimensional data into lower-dimensions. An AE consists of two different sub-units:
 
@@
 
-The AE is trained to reduce the reconstruction error between input and output data, after the training procedure is finished the network is able to reduce the dimensional of input data to a compressed representation{[}\protect\hyperlink{ref-ae_orig}{9}{]}.
-
-DAE was first introduced by {[}\protect\hyperlink{ref-dae_orig2008}{23}{]} as an improved AE with the ability to learn new representations of data that is especially robust to partially corrupted input data.
+The AE is trained to reduce the reconstruction error between input and output data, after the training procedure is finished the network is able to reduce the dimensional of input data to a compressed representation{[}\protect\hyperlink{ref-ae_orig}{10}{]}.
+
+DAE was first introduced by {[}\protect\hyperlink{ref-dae_orig2008}{27}{]} as an improved AE with the ability to learn new representations of data that is especially robust to partially corrupted input data.
 DAE modifies the AE by using partially corrupted input data for the AE and training it to reconstruct the uncorrupted, original version of the input data.
 
-Since the hidden representation of DAE captures the dependency structure of the input data it can therefore also be used to generate new solutions in the context of GAs {[}\protect\hyperlink{ref-harmless_overfitting_eda}{18}{]}.
-
-DAE-GP builds upon the concept of using DAEs in EDAs described by {[}\protect\hyperlink{ref-harmless_overfitting_eda}{18}{]} and transfers the concept to the domain of GP.
-The mutation and crossover operators of standard GP are replaced by sampling new solutions from a probabilistic model that is build by training a DAEs long short-term memory (LSTM) network on a subset of high fitness solutions from the current population {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{28}{]}.
-
-LSTMs are a variant of artificial neural networks first introduced by {[}\protect\hyperlink{ref-lstm_orig}{10}{]} that can store learned information over en extended time period while avoiding the problem of vanishing and/or exploding gradients.
-Since DAE-GP encodes candidate solutions as linear strings in prefix notation, the DAE in DAE-GP uses LSTMs for both encoding and decoding where the total amount of time steps \(T\) is equal to sum of the length of the input solution and the output solution {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{28}{]}.
+Since the hidden representation of DAE captures the dependency structure of the input data it can therefore also be used to generate new solutions in the context of GAs {[}\protect\hyperlink{ref-harmless_overfitting_eda}{21}{]}.
+
+DAE-GP builds upon the concept of using DAEs in EDAs described by {[}\protect\hyperlink{ref-harmless_overfitting_eda}{21}{]} and transfers the concept to the domain of GP.
+The mutation and crossover operators of standard GP are replaced by sampling new solutions from a probabilistic model that is build by training a DAEs long short-term memory (LSTM) network on a subset of high fitness solutions from the current population {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{33}{]}.
+
+LSTMs are a variant of artificial neural networks first introduced by {[}\protect\hyperlink{ref-lstm_orig}{11}{]} that can store learned information over en extended time period while avoiding the problem of vanishing and/or exploding gradients.
+Since DAE-GP encodes candidate solutions as linear strings in prefix notation, the DAE in DAE-GP uses LSTMs for both encoding and decoding where the total amount of time steps \(T\) is equal to sum of the length of the input solution and the output solution {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{33}{]}.
 
@@
 
-The strategy of pre-training artificial neural networks is based on the idea of transfer learning that aims at retaining previously learned knowledge from one task to re-use it for another task {[}\protect\hyperlink{ref-HAN2021225}{7}{]} {[}\protect\hyperlink{ref-survey_transfer_learning}{16}{]}.
+The strategy of pre-training artificial neural networks is based on the idea of transfer learning that aims at retaining previously learned knowledge from one task to re-use it for another task {[}\protect\hyperlink{ref-HAN2021225}{7}{]} {[}\protect\hyperlink{ref-survey_transfer_learning}{18}{]}.
 Transfer learning traditionally follows a two phase approach:
@@
 
-where source task and target task are usually similar but may differ in their feature space and the distribution of training data {[}\protect\hyperlink{ref-survey_transfer_learning}{16}{]}.
+where source task and target task are usually similar but may differ in their feature space and the distribution of training data {[}\protect\hyperlink{ref-survey_transfer_learning}{18}{]}.
 
@@
 Instead of initializing and optimizing each model from scratch as done in previous work (e.g.
-{[}\protect\hyperlink{ref-dae-gp_2020_rtree}{28}{]} {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{27}{]}), a separate DAE-LSTM network \(\hat{M}\) will be trained on a large population of randomly initialized solutions.\\
+{[}\protect\hyperlink{ref-dae-gp_2020_rtree}{33}{]} {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{32}{]}), a separate DAE-LSTM network \(\hat{M}\) will be trained on a large population of randomly initialized solutions.\\
 The trainable parameters \(\theta_{\hat{M}}\) obtained after finishing the training procedure of \(\hat{M}\) are then used as the starting parameters for each following DAE-LSTM model \(M_g\) for \(g\in{1,...,g_{max}}\).
@@
 Another motivation for introducing pre-training into DAE-GP is the prevalence of sampling error for small GP populations sizes.
-{[}\protect\hyperlink{ref-sampling_err_gp}{20}{]} finds that sampling error, non-systematic errors that are caused by observing only a small subset of the statistical population, is a severe problem in the domain of GP since the initial population may not be a representative sample of all possible solutions.
-{[}\protect\hyperlink{ref-sampling_err_gp}{20}{]} also introduces a method for calculating optimal population sizes to minimize the presence of sampling error.
+{[}\protect\hyperlink{ref-sampling_err_gp}{23}{]} finds that sampling error, non-systematic errors that are caused by observing only a small subset of the statistical population, is a severe problem in the domain of GP since the initial population may not be a representative sample of all possible solutions.
+{[}\protect\hyperlink{ref-sampling_err_gp}{23}{]} also introduces a method for calculating optimal population sizes to minimize the presence of sampling error.
 Since DAE-LSTM of DAE-GP learns the properties of it's training population and reuses the acquired knowledge in the sampling procedure, using the parameters obtained from training \(\hat{M}\) on a sufficiently large training population might reduce the need for large population sizes in the following generations of the search if \(\hat{M}\) already implicitly captured the properties of a representative sample of solutions.
@@
 
-The DAE-GP algorithm, first described by {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{28}{]}, is summarized as a flowchart in figure \ref{fig:daegp-flowchart}.
+The DAE-GP algorithm, first described by {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{33}{]}, is summarized as a flowchart in figure \ref{fig:daegp-flowchart}.
 After setting the generation counter \(g\) to \(0\) and creating the initial population of solutions \(P_0\), DAE-GP enters the main loop that checks if a termination criteria is satisfied (i.e.~maximum number of generations or fitness evaluations is reached).
@@
 One difficulty in the implementation of a pre-training strategy into DAE-GP has been the determination of the number of hidden neurons in side the DAE-LSTMs hidden layers.
-The original description of DAE-GP {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{28}{]} uses a strategy where the number of hidden neurons for each hidden layer is dynamically set per generation to the maximum individual size inside the current population.
+The original description of DAE-GP {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{33}{]} uses a strategy where the number of hidden neurons for each hidden layer is dynamically set per generation to the maximum individual size inside the current population.
 For a pre-training implementation, this strategy can not be easily adapted since it leads to a changing number of neurons at each generation resulting in different dimensions of the DAE-LSTM.
@@
 Train/Test Split & 50:50 \\
-Maximum Tree Depth & 17 {[}\protect\hyperlink{ref-Koza1993GeneticP}{12}{]} \\
-Initialization Algorithm & Ramped Half and Half {[}\protect\hyperlink{ref-Koza1993GeneticP}{12}{]} \\
+Maximum Tree Depth & 17 {[}\protect\hyperlink{ref-Koza1993GeneticP}{14}{]} \\
+Initialization Algorithm & Ramped Half and Half {[}\protect\hyperlink{ref-Koza1993GeneticP}{14}{]} \\
 Mix/Max Initialization-Depth & 2/6 \\
 Selection Operator & Binary Tournament Selection \\
-Function Set & \{+ , - , * , AQ {[}\protect\hyperlink{ref-AQ}{15}{]}\} \\
+Function Set & \{+ , - , * , AQ {[}\protect\hyperlink{ref-AQ}{17}{]}\} \\
 Ephemeral Constants & {[}-5,..,5{]} \\
@@
 \endlastfoot
-Stopping Condition & Convergence {[}\protect\hyperlink{ref-daegp_explore_exploit}{26}{]} \\
+Stopping Condition & Convergence {[}\protect\hyperlink{ref-daegp_explore_exploit}{31}{]} \\
 Reconstruction Error Metric & Multiclass Cross Entropy \\
-Gradient Descent Algorithm & Adam Optimization {[}\protect\hyperlink{ref-Kingma2014AdamAM}{11}{]} \\
+Gradient Descent Algorithm & Adam Optimization {[}\protect\hyperlink{ref-Kingma2014AdamAM}{13}{]} \\
 Learning Rate & 0.001 \\
@@
 Hidden Neurons & 150 \\
-Corruption Operator & Levenshtein Edit {[}\protect\hyperlink{ref-daegp_explore_exploit}{26}{]} \\
+Corruption Operator & Levenshtein Edit {[}\protect\hyperlink{ref-daegp_explore_exploit}{31}{]} \\
 Denoising Edit Probability (Model Building) & 0.05 \\
@@
 
-The framework that was used to conduct all experiments of this thesis was provided by the supervisor of this thesis David Wittenberg and uses the python programming language in conjunction with the baseline libraries \texttt{Keras} {[}\protect\hyperlink{ref-chollet2015keras}{2}{]} for deep neural networks and \texttt{deap} {[}\protect\hyperlink{ref-DEAP_JMLR2012}{6}{]} for evolutionary computation.
+The framework that was used to run all experiments of this thesis was provided by the supervisor of this thesis David Wittenberg and uses the python programming language {[}\protect\hyperlink{ref-python3}{26}{]} in conjunction with the baseline libraries \texttt{Keras} {[}\protect\hyperlink{ref-chollet2015keras}{2}{]} for deep neural networks and \texttt{deap} {[}\protect\hyperlink{ref-DEAP_JMLR2012}{6}{]} for evolutionary computation.
+For data visualizations I used the python package \texttt{matplotlib} {[}\protect\hyperlink{ref-matplotlib}{12}{]}, statistical analysis and evaluations were conducted using primarily the data science libraries \texttt{scipy} {[}\protect\hyperlink{ref-scipy}{28}{]}, \texttt{numpy} {[}\protect\hyperlink{ref-numpy}{8}{]}, \texttt{pandas} {[}\protect\hyperlink{ref-pandas}{24}{]}, and \texttt{IPython} {[}\protect\hyperlink{ref-ipython}{20}{]}.
 
@@
 To test pre-training in DAE-GP this thesis focuses on the domain of real-world symbolic regression problems.
-Symbolic Regression problems have been one of the first GP applications {[}\protect\hyperlink{ref-Koza1993GeneticP}{12}{]} and are an actively studied and highly relevant research area.
-The goal in symbolic regression is to find a mathematical model for a given set of data points {[}\protect\hyperlink{ref-10.1007ux2f978-3-540-24621-3_22}{17}{]}, in real-world symbolic regression these data points are sourced from real-world observations which in contrast to synthetic symbolic regression problem are more likely to contain random noise and bias.
+Symbolic Regression problems have been one of the first GP applications {[}\protect\hyperlink{ref-Koza1993GeneticP}{14}{]} and are an actively studied and highly relevant research area.
+The goal in symbolic regression is to find a mathematical model for a given set of data points {[}\protect\hyperlink{ref-10.1007ux2f978-3-540-24621-3_22}{19}{]}, in real-world symbolic regression these data points are sourced from real-world observations which in contrast to synthetic symbolic regression problem are more likely to contain random noise and bias.
 Another important challenge in solving real-world symbolic regression problems is the ability for a given model to generalize, we want the final model to show high accuracy in predicting outcomes for previously unseen cases.
@@
 The objective of the airfoil problem is to find a function that accurately predicts the output variable \(y\) by taking in a subset of the input variables \(x1,x2,x3,x4,x5\).
-The function set used by all DAE-GP variations for symbolic regression is summarized in tables \ref{tab:baseParams-GP} and , the terminal set consists of the the input variables \(x1,x2,x3,x4,x5\) and the ephemeral integer constants listed in table \ref{tab:baseParams-GP} {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{27}{]}.
+The function set used by all DAE-GP variations for symbolic regression is summarized in tables \ref{tab:baseParams-GP} and , the terminal set consists of the the input variables \(x1,x2,x3,x4,x5\) and the ephemeral integer constants listed in table \ref{tab:baseParams-GP} {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{32}{]}.
 
@@
 For each configuration I tested traditional DAE-GP as well as a pre-trained DAE-GP resulting in 12 total sub experiments that were each based on 10 individual runs (total number of runs=\(120\)).
-To avoid creating biased results through the presence of sampling error inside the pre-training population (see {[}\protect\hyperlink{ref-sampling_err_gp}{20}{]}), the population size for the pre-training phase is chosen very high with the size of \(\hat{P}\) =10000 where 50\% of \(\hat{P}\) is used for the training population \(\hat{P}_{train}\) and the remaining 50\% are used for the hold-out validation population \(\hat{P}_{test}\).
+To avoid creating biased results through the presence of sampling error inside the pre-training population (see {[}\protect\hyperlink{ref-sampling_err_gp}{23}{]}), the population size for the pre-training phase is chosen very high with the size of \(\hat{P}\) =10000 where 50\% of \(\hat{P}\) is used for the training population \(\hat{P}_{train}\) and the remaining 50\% are used for the hold-out validation population \(\hat{P}_{test}\).
 The training of DAE-LSTM uses a fixed number of 1000 epochs until termination and uses Adam optimization for gradient descent.
@@
 In general I expect that with a growing number of trainable parameters (either by adding more hidden layers or more hidden neurons per layer) the DAE-LSTM will be more prone to overfit to the training population \(P_{train}\) resulting in a small reconstruction error for the training population and a large one for the validation population \(P_{test}\).
-The reason for this effect is that a larger network trained over an extended period of time (without the use of strategies like early stopping), has much more potential to learn noise from the training dataset than a smaller network, which is more likely to result in worsening performance for previously unseen cases {[}\protect\hyperlink{ref-weigend1994overfitting}{24}{]}.
+The reason for this effect is that a larger network trained over an extended period of time (without the use of strategies like early stopping), has much more potential to learn noise from the training dataset than a smaller network, which is more likely to result in worsening performance for previously unseen cases {[}\protect\hyperlink{ref-weigend1994overfitting}{29}{]}.
 
@@
 
-Since previous work by {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{27}{]} demonstrated that DAE-GP, for a given number of 10.000 fitness evaluations, is able to produce solutions that are much smaller in tree size than regular GP, I also studied the influence of pre-training on the size of solutions found by DAE-GP.
+Since previous work by {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{32}{]} demonstrated that DAE-GP, for a given number of 10.000 fitness evaluations, is able to produce solutions that are much smaller in tree size than regular GP, I also studied the influence of pre-training on the size of solutions found by DAE-GP.
 Figure \ref{fig:airfoil-size-combined-30gens} shows the median of both the average size of individuals inside the population as well as the size of the currently best performing solution inside the population.
@@
 Another interesting metric to examine in population-based optimization algorithms is the diversity of the current generations population.
-Since individual solutions in DAE-GP are computer programs in the form of parse trees, I selected the normalized Levenshtein edit distance {[}\protect\hyperlink{ref-NormLevDistance}{13}{]} as my primary metric for tracking population diversity.
+Since individual solutions in DAE-GP are computer programs in the form of parse trees, I selected the normalized Levenshtein edit distance {[}\protect\hyperlink{ref-NormLevDistance}{15}{]} as my primary metric for tracking population diversity.
 Figure \ref{fig:airfoil-popDiversity} shows the median population diversity for both DAE-GP algorithms and regular GP as a benchmark.
@@
 To ensure a fair comparison between pre-trained and regular DAE-GP, the prior experiments used a static number of hidden neurons for each hidden Layer inside the DAE-LSTM networks.
-For comparison, figure \ref{fig:airfoil-maxIndSize-params} shows the total number of trainable parameters for each generation's DAE-LSTM of another experiment where regular DAE-GP was tested with dynamic adjustment of the number of hidden neurons as described in {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{28}{]} and {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{27}{]}.
+For comparison, figure \ref{fig:airfoil-maxIndSize-params} shows the total number of trainable parameters for each generation's DAE-LSTM of another experiment where regular DAE-GP was tested with dynamic adjustment of the number of hidden neurons as described in {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{33}{]} and {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{32}{]}.
 Again, the experiments was based on \(10\) individual runs and otherwise used the same hyperparameters as listed in tables \ref{tab:baseParams-GP} and \ref{tab:baseParams-DAEGP}.
@@
 
-Since previous research on DAE-GP used a single hidden layers for the airfoil dataset {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{27}{]}, I repeated the experiment with a single hidden layer.
+Since previous research on DAE-GP used a single hidden layers for the airfoil dataset {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{32}{]}, I repeated the experiment with a single hidden layer.
 Based on the findings of {[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{5}{]}, I expected that pre-training DAE-GP would lead to a decrease in overall performance.
@@
 
-The pre-training implementation in previous experiments used ramped half and half initialization {[}\protect\hyperlink{ref-Koza1993GeneticP}{12}{]} to create \(\hat{P}\), this leads to 50\% of the population consisting of full grown trees.
+The pre-training implementation in previous experiments used ramped half and half initialization {[}\protect\hyperlink{ref-Koza1993GeneticP}{14}{]} to create \(\hat{P}\), this leads to 50\% of the population consisting of full grown trees.
 An alternative method of initializing \(\hat{P}\) is to use the grow initialization method where solutions are not guaranteed to be full trees.
@@
 While investigating other reasons for the differences in run-time, I verified that both DAE-GP variants use the exact same number of trainable parameters for each generations DAE-LSTMS models (total number of trainable parameters DAE-GP=573473, Pre-trained DAE-GP=573473).
-After profiling a single run of DAE-GP and pre-trained DAE-GP using the python \texttt{cProfiler} module {[}\protect\hyperlink{ref-python3}{22}{]}, my best explanation for the remaining differences in run-time is an inefficiency in my implementation of pre-training.
+After profiling a single run of DAE-GP and pre-trained DAE-GP using the python \texttt{cProfiler} module, my best explanation for the remaining differences in run-time is an inefficiency in my implementation of pre-training.
 Figure \ref{fig:airfoil-profiling} shows the cumulative time spend for the 20 most expensive function calls.
@@
 Airfoil & 1503 & 5 & {[}\protect\hyperlink{ref-AIRFOIL_DATASET}{1}{]} \\
-Boston\_Housing & 506 & 13 & {[}\protect\hyperlink{ref-BOSTON_HOUSING_DATASET}{8}{]} \\
-Energy\_Cooling & 768 & 8 & {[}\protect\hyperlink{ref-ENERGY_DATASET}{21}{]} \\
-Concrete & 1030 & 8 & {[}\protect\hyperlink{ref-CONCRETE_DATASET}{29}{]} \\
+Boston\_Housing & 506 & 13 & {[}\protect\hyperlink{ref-BOSTON_HOUSING_DATASET}{9}{]} \\
+Energy\_Cooling & 768 & 8 & {[}\protect\hyperlink{ref-ENERGY_DATASET}{25}{]} \\
+Concrete & 1030 & 8 & {[}\protect\hyperlink{ref-CONCRETE_DATASET}{34}{]} \\
 \end{longtable}
 
-To test for statistical differences of the results achieved by either pre-trained or regular DAE-GP, I selected the nonparametric Mann-Whitney-U Test ({[}\protect\hyperlink{ref-mwu1}{25}{]}, {[}\protect\hyperlink{ref-mwu2}{14}{]}) to test for differences in the distribution underlying the samples sourced from my experiments.
+To test for statistical differences of the results achieved by either pre-trained or regular DAE-GP, I selected the nonparametric Mann-Whitney-U Test ({[}\protect\hyperlink{ref-mwu1}{30}{]}, {[}\protect\hyperlink{ref-mwu2}{16}{]}) to test for differences in the distribution underlying the samples sourced from my experiments.
 Additionally, I included the effect strength of pre-training as measured by Cliffs Delta {[}\protect\hyperlink{ref-Cliff1993DominanceSO}{3}{]} into the statistical analysis.
@@
 For a single hidden layer, pre-training did show a negative impact on the median final fitness of solutions found as it was suggested by {[}\protect\hyperlink{ref-pmlr-v9-erhan10a}{4}{]}.
-Since current research on DAE-GP for real-world symbolic regression relies on a single hidden layer (see {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{27}{]}), pre-training does come with additional computational expenses for using deeper DAE-LSTMs.
+Since current research on DAE-GP for real-world symbolic regression relies on a single hidden layer (see {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{32}{]}), pre-training does come with additional computational expenses for using deeper DAE-LSTMs.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-numpy}{}}%
+\CSLLeftMargin{{[}8{]} }%
+\CSLRightInline{Charles R.
+Harris, K.
+Jarrod Millman, Stéfan J.
+van der Walt, Ralf Gommers, Pauli Virtanen, David Cournapeau, Eric Wieser, Julian Taylor, Sebastian Berg, Nathaniel J.
+Smith, Robert Kern, Matti Picus, Stephan Hoyer, Marten H.
+van Kerkwijk, Matthew Brett, Allan Haldane, Jaime Fernández del Río, Mark Wiebe, Pearu Peterson, Pierre Gérard-Marchant, Kevin Sheppard, Tyler Reddy, Warren Weckesser, Hameer Abbasi, Christoph Gohlke, and Travis E.
+Oliphant.
+2020.
+Array programming with {NumPy}.
+\emph{Nature} 585, 7825 (September 2020), 357--362.
+DOI:https://doi.org/\href{https://doi.org/10.1038/s41586-020-2649-2}{10.1038/s41586-020-2649-2}}
+
 \leavevmode\vadjust pre{\hypertarget{ref-BOSTON_HOUSING_DATASET}{}}%
-\CSLLeftMargin{{[}8{]} }%
+\CSLLeftMargin{{[}9{]} }%
 \CSLRightInline{David Harrison and Daniel Rubinfeld.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ae_orig}{}}%
-\CSLLeftMargin{{[}9{]} }%
+\CSLLeftMargin{{[}10{]} }%
 \CSLRightInline{G.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-lstm_orig}{}}%
-\CSLLeftMargin{{[}10{]} }%
+\CSLLeftMargin{{[}11{]} }%
 \CSLRightInline{Sepp Hochreiter and Jürgen Schmidhuber.
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-matplotlib}{}}%
+\CSLLeftMargin{{[}12{]} }%
+\CSLRightInline{J.
+D.
+Hunter.
+2007.
+Matplotlib: A 2D graphics environment.
+\emph{Computing in Science \& Engineering} 9, 3 (2007), 90--95.
+DOI:https://doi.org/\href{https://doi.org/10.1109/MCSE.2007.55}{10.1109/MCSE.2007.55}}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Kingma2014AdamAM}{}}%
-\CSLLeftMargin{{[}11{]} }%
+\CSLLeftMargin{{[}13{]} }%
 \CSLRightInline{Diederik P. Kingma and Jimmy Ba. 2014. Adam: A method for stochastic optimization. \emph{CoRR} abs/1412.6980, (2014).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Koza1993GeneticP}{}}%
-\CSLLeftMargin{{[}12{]} }%
+\CSLLeftMargin{{[}14{]} }%
 \CSLRightInline{John R. Koza. 1993. Genetic programming - on the programming of computers by means of natural selection. In \emph{Complex adaptive systems}.}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-NormLevDistance}{}}%
-\CSLLeftMargin{{[}13{]} }%
+\CSLLeftMargin{{[}15{]} }%
 \CSLRightInline{Joseph B.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-mwu2}{}}%
-\CSLLeftMargin{{[}14{]} }%
+\CSLLeftMargin{{[}16{]} }%
 \CSLRightInline{Henry B.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-AQ}{}}%
-\CSLLeftMargin{{[}15{]} }%
+\CSLLeftMargin{{[}17{]} }%
 \CSLRightInline{Ji Ni, Russ H.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-survey_transfer_learning}{}}%
-\CSLLeftMargin{{[}16{]} }%
+\CSLLeftMargin{{[}18{]} }%
 \CSLRightInline{Sinno Jialin Pan and Qiang Yang.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-10.1007ux2f978-3-540-24621-3_22}{}}%
-\CSLLeftMargin{{[}17{]} }%
+\CSLLeftMargin{{[}19{]} }%
 \CSLRightInline{Grégory Paris, Denis Robilliard, and Cyril Fonlupt.
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-ipython}{}}%
+\CSLLeftMargin{{[}20{]} }%
+\CSLRightInline{Fernando Pérez and Brian E.
+Granger.
+2007.
+{IP}ython: A system for interactive scientific computing.
+\emph{Computing in Science and Engineering} 9, 3 (May 2007), 21--29.
+DOI:https://doi.org/\href{https://doi.org/10.1109/MCSE.2007.53}{10.1109/MCSE.2007.53}}
+
 \leavevmode\vadjust pre{\hypertarget{ref-harmless_overfitting_eda}{}}%
-\CSLLeftMargin{{[}18{]} }%
+\CSLLeftMargin{{[}21{]} }%
 \CSLRightInline{Malte Probst and Franz Rothlauf.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-design_of_modern_heuristics}{}}%
-\CSLLeftMargin{{[}19{]} }%
+\CSLLeftMargin{{[}22{]} }%
 \CSLRightInline{Franz Rothlauf. 2011. \emph{Design of modern heuristics: Principles and application}. DOI:https://doi.org/\href{https://doi.org/10.1007/978-3-540-72962-4}{10.1007/978-3-540-72962-4}}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-sampling_err_gp}{}}%
-\CSLLeftMargin{{[}20{]} }%
+\CSLLeftMargin{{[}23{]} }%
 \CSLRightInline{Dirk Schweim, David Wittenberg, and Franz Rothlauf.
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-pandas}{}}%
+\CSLLeftMargin{{[}24{]} }%
+\CSLRightInline{The pandas development team. 2020. Pandas-dev/pandas: pandas. DOI:https://doi.org/\href{https://doi.org/10.5281/zenodo.3509134}{10.5281/zenodo.3509134}}
+
 \leavevmode\vadjust pre{\hypertarget{ref-ENERGY_DATASET}{}}%
-\CSLLeftMargin{{[}21{]} }%
+\CSLLeftMargin{{[}25{]} }%
 \CSLRightInline{Athanasios Tsanas and Angeliki Xifara.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-python3}{}}%
-\CSLLeftMargin{{[}22{]} }%
+\CSLLeftMargin{{[}26{]} }%
 \CSLRightInline{Guido Van Rossum and Fred L. Drake. 2009. \emph{Python 3 reference manual}. CreateSpace, Scotts Valley, CA.}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-dae_orig2008}{}}%
-\CSLLeftMargin{{[}23{]} }%
+\CSLLeftMargin{{[}27{]} }%
 \CSLRightInline{Pascal Vincent, Hugo Larochelle, Y.
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-scipy}{}}%
+\CSLLeftMargin{{[}28{]} }%
+\CSLRightInline{Pauli Virtanen, Ralf Gommers, Travis E.
+Oliphant, Matt Haberland, Tyler Reddy, David Cournapeau, Evgeni Burovski, Pearu Peterson, Warren Weckesser, Jonathan Bright, Stéfan J.
+van der Walt, Matthew Brett, Joshua Wilson, K.
+Jarrod Millman, Nikolay Mayorov, Andrew R.
+J.
+Nelson, Eric Jones, Robert Kern, Eric Larson, C J Carey, İlhan Polat, Yu Feng, Eric W.
+Moore, Jake VanderPlas, Denis Laxalde, Josef Perktold, Robert Cimrman, Ian Henriksen, E.
+A.
+Quintero, Charles R.
+Harris, Anne M.
+Archibald, Antônio H.
+Ribeiro, Fabian Pedregosa, Paul van Mulbregt, and SciPy 1.0 Contributors.
+2020.
+{{SciPy} 1.0: Fundamental Algorithms for Scientific Computing in Python}.
+\emph{Nature Methods} 17, (2020), 261--272.
+DOI:https://doi.org/\href{https://doi.org/10.1038/s41592-019-0686-2}{10.1038/s41592-019-0686-2}}
+
 \leavevmode\vadjust pre{\hypertarget{ref-weigend1994overfitting}{}}%
-\CSLLeftMargin{{[}24{]} }%
+\CSLLeftMargin{{[}29{]} }%
 \CSLRightInline{Andreas Weigend. 1994. On overfitting and the effective number of hidden units. In \emph{Proceedings of the 1993 connectionist models summer school}, 335--342.}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-mwu1}{}}%
-\CSLLeftMargin{{[}25{]} }%
+\CSLLeftMargin{{[}30{]} }%
 \CSLRightInline{Frank. Wilcoxon. 1945. Individual comparisons by ranking methods. \emph{Biometrics} 1, (1945), 196--202.}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-daegp_explore_exploit}{}}%
-\CSLLeftMargin{{[}26{]} }%
+\CSLLeftMargin{{[}31{]} }%
 \CSLRightInline{David Wittenberg.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-dae-gp_2022_symreg}{}}%
-\CSLLeftMargin{{[}27{]} }%
+\CSLLeftMargin{{[}32{]} }%
 \CSLRightInline{David Wittenberg and Franz Rothlauf.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-dae-gp_2020_rtree}{}}%
-\CSLLeftMargin{{[}28{]} }%
+\CSLLeftMargin{{[}33{]} }%
 \CSLRightInline{David Wittenberg, Franz Rothlauf, and Dirk Schweim.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-CONCRETE_DATASET}{}}%
-\CSLLeftMargin{{[}29{]} }%
+\CSLLeftMargin{{[}34{]} }%
 \CSLRightInline{I-Cheng Yeh.
@@
 \hypertarget{IV}{%
-\section*{II - Statutory Declaration}\label{IV}}
-\addcontentsline{toc}{section}{II - Statutory Declaration}
+\section*{Statutory Declaration}\label{IV}}
+\addcontentsline{toc}{section}{Statutory Declaration}
 

### 2023-02-06 paper.Rmd
Mon Feb  6 19:00:07 CET 2023: update-remote.sh

@@
 
+REG_TRAIN_MED_10F <- py$D_AIRFOIL_2hl$REG_TRAIN_MED_10F
+REG_TEST_MED_10F <- py$D_AIRFOIL_2hl$REG_TEST_MED_10F
+PT_TRAIN_MED_10F <- py$D_AIRFOIL_2hl$PT_TRAIN_MED_10F
+PT_TEST_MED_10F <- py$D_AIRFOIL_2hl$PT_TEST_MED_10F
+GP_TRAIN_MED_10F <- py$D_AIRFOIL_2hl$GP_TRAIN_MED_10F
+GP_TEST_MED_10F <- py$D_AIRFOIL_2hl$GP_TEST_MED_10F
+
 REG_TRAIN_STDD_FF <- py$D_AIRFOIL_2hl$REG_TRAIN_STDD_FF
@@
 Regarding the presence of overfitting, both DAE-GP variants show only a very small gap between the testing and training fitness which indicates a good generalization behavior.
-Another interesting observation from figure \@ref(fig:airfoil-median-fitness-30gens) is that pre-trained DAE-GP is faster at improving fitness than regular DAE-GP, this is especially true if only looking at the first 10 generations
+Another interesting observation from figure \@ref(fig:airfoil-median-fitness-30gens) is that pre-trained DAE-GP is faster at improving fitness than regular DAE-GP, this is especially true if only looking at the first 10 generations.
+The median best RMSE on the test set after 10 generations for pre-trained DAE-GP is `r round(PT_TEST_MED_10F, 3)` which is smaller by a factor of `r round( PT_TEST_MED_10F / REG_TEST_MED_10F, 3)` than regular DAE-GP with a median best RMSE of `r round( REG_TEST_MED_10F, 3)`
 
@@
 Figure \@ref(fig:symReg-LevDiv) shows the median levenshtein edit distance as a metric for population diversity over the evolutionary search.
+It clearly shows that pre-training does increase the population diversity for DAE-GP consistently for all tested benchmark problems.
 

### 2023-02-06 paper.tex
Mon Feb  6 19:00:07 CET 2023: update-remote.sh

@@
 Regarding the presence of overfitting, both DAE-GP variants show only a very small gap between the testing and training fitness which indicates a good generalization behavior.
+Another interesting observation from figure \ref{fig:airfoil-median-fitness-30gens} is that pre-trained DAE-GP is faster at improving fitness than regular DAE-GP, this is especially true if only looking at the first 10 generations.
+The median best RMSE on the test set after 10 generations for pre-trained DAE-GP is 17.093 which is smaller by a factor of 0.558 than regular DAE-GP with a median best RMSE of 30.658
 

### 2023-02-08 paper.Rmd
autogit

@@
 
-<!-- TODO: Disclaimer, Run-time is implementation dependent, no GPU use, etc... -->
-
 ```{python read_runtime_vars, echo=FALSE}
@@
 knitr::include_graphics("./img/airfoil_2hl_150hn_fullRun_30gens_TimeBenchmark/median_runtime.png")
+```
+Figure \@ref(fig:airfoil-runtime-boxplot) shows the distribution of the total runtime for both DAE-GP variants.
+It shows that Pre-Training, besides largely increasing the median runtime, also leads to a higher dispersion which is shown by both a larger interquartile range as well as by a a standard deviation that increases by a factor of `r round(782.68 / 635.04, 3)` from 635,04 for regular DAE-GP to 782,68 for pre-trained DAE-GP.
+
+```{r airfoil-runtime-boxplot, echo=FALSE, fig.cap="Total Runtime Boxplot - Airfoil", out.width="80%"}
+knitr::include_graphics("./img/airfoil_2hl_150hn_fullRun_30gens_TimeBenchmark/total_runtime_distribution_boxplot.png")
 ```

### 2023-02-08 paper.tex
autogit

@@
 \author{}
-\date{\vspace{-2.5em}Date of Submission: 2023-02-06}
+\date{\vspace{-2.5em}Date of Submission: 2023-02-08}
 
@@
 
+Figure \ref{fig:airfoil-runtime-boxplot} shows the distribution of the total runtime for both DAE-GP variants.
+It shows that Pre-Training, besides largely increasing the median runtime, also leads to a higher dispersion which is shown by both a larger interquartile range as well as by a a standard deviation that increases by a factor of 1.232 from 635,04 for regular DAE-GP to 782,68 for pre-trained DAE-GP.
+
+\begin{figure}[c]
+
+{\centering \includegraphics[width=0.8\linewidth]{./img/airfoil_2hl_150hn_fullRun_30gens_TimeBenchmark/total_runtime_distribution_boxplot}
+
+}
+
+\caption{Total Runtime Boxplot - Airfoil}\label{fig:airfoil-runtime-boxplot}
+\end{figure}
+
 To investigate the strong negative impact of pre-training on DAE-GP run-time, I first studied the influence of pre-training on the median number of training epochs that had to be spend at each generation for the training of the DAE-LSTM models \(M_g\) (excluding the pre-training DAE-LSTM model \(\hat{M}\)).
@@
 Figure \ref{fig:symReg-LevDiv} shows the median levenshtein edit distance as a metric for population diversity over the evolutionary search.
+It clearly shows that pre-training does increase the population diversity for DAE-GP consistently for all tested benchmark problems.
 

### 2023-02-08 paper.Rmd
autogit

@@
 During model building, DAE-GP selects a subset $X_g$ of the current population that is used as training data for the DAE-LSTM model $M_g$ which is trained to learn the properties of $X_g$.
-The trained model $M_g$ is then used for model sampling, partially mutated solutions from $P_G$ are propagated through $M_g$ to create a new population $P_{g+1}$.
+The trained model $M_g$ is then used for model sampling, partially mutated solutions from $P_g$ are propagated through $M_g$ to create a new population $P_{g+1}$.
 This process is repeated until the termination criteria is met and DAE-GP returns the highest fitness solutions.
@@
 
-If we look at another time consuming operation in DAE-GP, the creation of new individuals by sampling the DAE-LSTM model $M_G$, we see one main reason for the large run-time increase by pre-training DAE-GP.
+If we look at another time consuming operation in DAE-GP, the creation of new individuals by sampling the DAE-LSTM model $M_g$, we see one main reason for the large run-time increase by pre-training DAE-GP.
 Figure \@ref(fig:airfoil-SamplingTime) shows the median sampling time spend during each generation of the search and it clearly demonstrates that pre-training does strongly increase the time spent on model sampling  in comparison to regular DAE-GP.

### 2023-02-08 paper.tex
autogit

@@
 Airfoil & 1 & 0.86 & \textbf{0.93} & 0.00*** & 0.93 \\
-Airfoil & 1 & 0.86 & \textbf{0.93} & 0.00*** & 0.93 \\
-Boston\_Housing & 1 & 0.88 & \textbf{0.95} & 0.00*** & 0.92 \\
-Energy(Cooling) & 1 & 0.87 & \textbf{0.94} & 0.00*** & 0.94 \\
-Concrete & 1 & 0.87 & \textbf{0.94} & 0.00*** & 0.93 \\
+Airfoil & 2 & 0.86 & \textbf{0.93} & 0.00*** & 0.93 \\
+Boston\_Housing & 2 & 0.88 & \textbf{0.95} & 0.00*** & 0.92 \\
+Energy(Cooling) & 2 & 0.87 & \textbf{0.94} & 0.00*** & 0.94 \\
+Concrete & 2 & 0.87 & \textbf{0.94} & 0.00*** & 0.93 \\
 \end{longtable}

### 2023-02-08 template.Rmd
autogit

---
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Statutory Declaration

I herewith declare that I have composed the present thesis myself and without use of any other than the
cited sources and aids. Sentences or parts of sentences quoted literally are marked as such; other references
with regard to the statement and scope are indicated by full details of the publications concerned. The thesis
in the same or similar form has not been submitted to any examination body and has not been published.
This thesis was not yet, even in part, used in another examination or as a course performance

Place, Date:  
Signature:

## Ehrenwörtliche Erklärung

Ich versichere hiermit, dass ich die vorliegende Arbeit selbständig und ohne Benutzung anderer als der
angegebenen Quellen und Hilfsmittel verfasst habe. Wörtlich übernommene Sätze oder Satzteile sind als
Zitat belegt, andere Anlehnungen, hinsichtlich Aussage und Umfang, unter Quellenangabe kenntlich
gemacht. Die Arbeit hat in gleicher oder ähnlicher Form noch keiner Prüfungsbehörde vorgelegen und ist
nicht veröffentlicht. Sie wurde nicht, auch nicht auszugsweise, für eine andere Prüfungs- oder
Studienleistung verwendet.

Ort, Datum:  
Unterschrift:

### 2023-02-08 paper.tex
Wed Feb  8 19:08:36 CET 2023: update-remote.sh

@@
 During model building, DAE-GP selects a subset \(X_g\) of the current population that is used as training data for the DAE-LSTM model \(M_g\) which is trained to learn the properties of \(X_g\).
-The trained model \(M_g\) is then used for model sampling, partially mutated solutions from \(P_G\) are propagated through \(M_g\) to create a new population \(P_{g+1}\).
+The trained model \(M_g\) is then used for model sampling, partially mutated solutions from \(P_g\) are propagated through \(M_g\) to create a new population \(P_{g+1}\).
 This process is repeated until the termination criteria is met and DAE-GP returns the highest fitness solutions.
@@
 
-If we look at another time consuming operation in DAE-GP, the creation of new individuals by sampling the DAE-LSTM model \(M_G\), we see one main reason for the large run-time increase by pre-training DAE-GP.
+If we look at another time consuming operation in DAE-GP, the creation of new individuals by sampling the DAE-LSTM model \(M_g\), we see one main reason for the large run-time increase by pre-training DAE-GP.
 Figure \ref{fig:airfoil-SamplingTime} shows the median sampling time spend during each generation of the search and it clearly demonstrates that pre-training does strongly increase the time spent on model sampling in comparison to regular DAE-GP.
@@
 
-\ldots hier einfügen\ldots{}
+\begin{center}\includegraphics{private/statutory_declaration} \end{center}
 

### 2023-02-08 paper.Rmd
Wed Feb  8 13:00:08 CET 2023: update-remote.sh
Wed Feb  8 19:08:36 CET 2023: update-remote.sh
Wed Feb  8 20:00:06 CET 2023: update-remote.sh

@@
 # energy (cooling)
-bash /Users/rmn/masterThesis/eda-gp-2020/experiments/energyCooling_2hl_maxIndSize_fullRun_30gens/re_evaluate_energyCooling_2hl_maxIndSize_fullRun_30gens.sh
+bash /Users/rmn/masterThesis/eda-gp-2020/experiments/energyCooling_2hl_150hn_fullRun_30gens/re_evaluate_energyCooling_2hl_150hn_fullRun_30gens.sh
 
@@
 
-...hier einfügen...
+```{r statutory_declaration, echo=FALSE}
+knitr::include_graphics("~/masterThesis/master_thesis/private/statutory_declaration.pdf")
+```

### 2023-02-08 data/summarize_data.py

@@ -48,14 +48,14 @@ RESULTS = {
     "airfoil_1hl" : "/Users/rmn/github/master_thesis/data/airfoil_1hl_150hn_fullRun_30gens",
     "airfoil_2hl" : "/Users/rmn/github/master_thesis/data/airfoil_2hl_150hn_fullRun_30gens",
     "bostonHousing": "/Users/rmn/github/master_thesis/data/bostonHousing_2hl_150hn_fullRun_30gens",
-    "energyCooling": "/Users/rmn/github/master_thesis/data/energyCooling_2hl_150hn_FullRun_30gens",
+    "energyCooling": "/Users/rmn/masterThesis/master_thesis/data/energyCooling_2hl_150hn_FullRun_30gens",
     "concrete": "/Users/rmn/github/master_thesis/data/concrete_2hl_150hn_FullRun_30gens"
 }
 
 RESULTS_2hl = {
     "Airfoil" :  "/Users/rmn/github/master_thesis/data/airfoil_2hl_150hn_fullRun_30gens",
     "BostonHousing": "/Users/rmn/github/master_thesis/data/bostonHousing_2hl_150hn_fullRun_30gens",
-    "energyCooling": "/Users/rmn/github/master_thesis/data/energyCooling_2hl_150hn_FullRun_30gens",
+    "energyCooling": "/Users/rmn/masterThesis/master_thesis/data/energyCooling_2hl_150hn_FullRun_30gens",
     "Concrete": "/Users/rmn/github/master_thesis/data/concrete_2hl_150hn_FullRun_30gens"
 }
 

### 2023-02-08 paper.Rmd
Wed Feb  8 21:00:07 CET 2023: update-remote.sh

@@
 Note that the results for the real world symbolic regression problems inside this chapter, except for the airfoil problem (see subsection [Influence on Run-Time]), have not been analyzed for their total run-time.
-This is due to the fact, that experiments were run on different computers architectures which resulted in strongly biased results.
+This is due to the fact, that experiments were run on different computers architectures which would result in biased results.
 
 ## Influence on Fitness
-
-The results for the median best fitness over the evolutionary search are visualized in figure \@ref(fig:symReg-Fitness).
-Figure \@ref(fig:symReg-Fitness-BoxPlots) shows the distribution of final fitness values on both the test and training set as box plots.
 
@@
 
-```{r symReg-Fitness-BoxPlots, echo=FALSE, fig.cap="Fitness after 30 Generations - Real World Symbolic Regression", out.width="90%"}
-knitr::include_graphics("~/masterThesis/master_thesis/img/symbolicRegressionSummarized/final_fitness_boxplot.png")
-```
-
-Table \@ref(tab:full-run-realWorldSymReg-fitness) summarizes the results for the median best fitness achieved after 30 generations and shows the results of computing the statistical tests.
+The results for the median best fitness over the evolutionary search are visualized in figure \@ref(fig:symReg-Fitness).
+In general, pre-training and regular DAE-GP both are able to improve on the best fitness found over the evolutionary search.
+Looking at figure \@ref(fig:symReg-Fitness), it appears that no significant differences for the best fitness between both algorithms exist.
+Regular DAE-GP achieves a slightly lower RMSE on the test set for Airfoil, Boston Housing and the Energy Cooling dataset while pre-trained DAE-GP is slightly better for the Concrete dataset.
 
@@
 # Create the data frame from csv
-df <- read.csv(
-  #"/Users/rmn/github/master_thesis/data/summary_table_final_fit_mean.csv",
+df_symreg_fitness <- read.csv(
   "/Users/rmn/github/master_thesis/data/summary_table_final_fit_median.csv",
@@
 # Create a new dataframe with the modified values of the two columns,
-df_bold <- df
+df_bold <- df_symreg_fitness
 
-for(i in 1:nrow(df)){
-  if(df[i, "DAE.GP"] < df[i, "Pre.Trained_DAE.GP"]){
-    df_bold[i, "DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(round(df[i, "DAE.GP"],4)))
-    df_bold[i, "Pre.Trained_DAE.GP"] <- as.character(round(df[i, "Pre.Trained_DAE.GP"],4))
+for(i in 1:nrow(df_symreg_fitness)){
+  if(df_symreg_fitness[i, "DAE.GP"] < df_symreg_fitness[i, "Pre.Trained_DAE.GP"]){
+    df_bold[i, "DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(round(df_symreg_fitness[i, "DAE.GP"],4)))
+    df_bold[i, "Pre.Trained_DAE.GP"] <- as.character(round(df_symreg_fitness[i, "Pre.Trained_DAE.GP"],4))
     }
-  else if (df[i, "DAE.GP"] > df[i, "Pre.Trained_DAE.GP"]) {
-    df_bold[i, "Pre.Trained_DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(round(df[i, "Pre.Trained_DAE.GP"],4)))
-  df_bold[i, "DAE.GP"] <- as.character(round(df[i, "DAE.GP"],4))
+  else if (df_symreg_fitness[i, "DAE.GP"] > df_symreg_fitness[i, "Pre.Trained_DAE.GP"]) {
+    df_bold[i, "Pre.Trained_DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(round(df_symreg_fitness[i, "Pre.Trained_DAE.GP"],4)))
+  df_bold[i, "DAE.GP"] <- as.character(round(df_symreg_fitness[i, "DAE.GP"],4))
   }
@@
 
+```
+
+Table \@ref(tab:full-run-realWorldSymReg-fitness) summarizes the results for the median best fitness achieved after 30 generations and shows the results of computing the statistical tests.
+As expected after reviewing figure \@ref(fig:symReg-Fitness), none of the differences in median best fitness are statistically significant and pre-training has only a very small, negligible effect strength for Cliff's Delta ranging between `r min(df_symreg_fitness$Cliffs_Delta)` and `r max(df_symreg_fitness$Cliffs_Delta)` [@Vargha2000ACA].
+
+Figure \@ref(fig:symReg-Fitness-BoxPlots) shows the distribution of final fitness values on both the test and training set as box plots.
+It shows that depending on the symbolic regression problem, the dispersion of fitness values is different for both DAE-GP variants.
+For example, a comparativley large span is visible for the airfoil problem while fitness scores for Energy Cooling are much more concentrated.
+The distribution are nontheless similar for both algorithms which does not suggest any meaningful differences between using pre-training and the distribution of final fitness scores.
+
+```{r symReg-Fitness-BoxPlots, echo=FALSE, fig.cap="Fitness after 30 Generations - Real World Symbolic Regression", out.width="90%"}
+knitr::include_graphics("~/masterThesis/master_thesis/img/symbolicRegressionSummarized/final_fitness_boxplot.png")
 ```

### 2023-02-08 paper.tex
Wed Feb  8 21:00:07 CET 2023: update-remote.sh

@@
 
-Denoising Autoencoder Genetic Programming (DAE-GP) is a novel variation of an genetic programming based Estimation of Distribution Algorithm (EDA-GP) that uses a denoising autoencoders long short-term memory network (DAE-LSTMs) as a probabilistic model to sample new candidate solutions {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{33}{]}.
+Denoising Autoencoder Genetic Programming (DAE-GP) is a novel variation of an genetic programming based Estimation of Distribution Algorithm (EDA-GP) that uses a denoising autoencoders long short-term memory network (DAE-LSTMs) as a probabilistic model to sample new candidate solutions {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{34}{]}.
 
@@
 In DAE-GP, DAE-LSTMs are trained with a subset of high-fitness solutions selected from a parent population with the aim to capture their promising qualities.
-The resulting model is then used to sample new offspring solutions by propagating partially mutated solutions from the parent population through the DAE-LSTMS {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{33}{]}.
-In previous work DAE-GP has been shown to outperform GP for both a generalized version of the royal tree problem {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{33}{]} as well as for a real-world symbolic regression problem {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{32}{]}.
-
-The DAE-GP algorithm first described by {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{33}{]} trains a DAE-LSTMs for each generation \(g\) of the search from scratch.
-{[}\protect\hyperlink{ref-dae-gp_2022_symreg}{32}{]} and {[}\protect\hyperlink{ref-daegp_explore_exploit}{31}{]} suggests the incorporation of a pre-training strategy into the evolutionary search as a possible way of improving the overall performance of the DAE-GP algorithm.
+The resulting model is then used to sample new offspring solutions by propagating partially mutated solutions from the parent population through the DAE-LSTMS {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{34}{]}.
+In previous work DAE-GP has been shown to outperform GP for both a generalized version of the royal tree problem {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{34}{]} as well as for a real-world symbolic regression problem {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{33}{]}.
+
+The DAE-GP algorithm first described by {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{34}{]} trains a DAE-LSTMs for each generation \(g\) of the search from scratch.
+{[}\protect\hyperlink{ref-dae-gp_2022_symreg}{33}{]} and {[}\protect\hyperlink{ref-daegp_explore_exploit}{32}{]} suggests the incorporation of a pre-training strategy into the evolutionary search as a possible way of improving the overall performance of the DAE-GP algorithm.
 The key idea is to pre-train an initial DAE-LSTM on a large population of candidate solutions and to use the pre-trained parameters of this initial model in each generation of the search as starting parameters for the current generation DAE-LSTM.
@@
 
-DAE-GP is an EDA-GP algorithm that uses DAE-LSTM networks as a probabilistic model to sample new offspring solutions {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{33}{]}.
+DAE-GP is an EDA-GP algorithm that uses DAE-LSTM networks as a probabilistic model to sample new offspring solutions {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{34}{]}.
 
@@
 
-DAE was first introduced by {[}\protect\hyperlink{ref-dae_orig2008}{27}{]} as an improved AE with the ability to learn new representations of data that is especially robust to partially corrupted input data.
+DAE was first introduced by {[}\protect\hyperlink{ref-dae_orig2008}{28}{]} as an improved AE with the ability to learn new representations of data that is especially robust to partially corrupted input data.
 DAE modifies the AE by using partially corrupted input data for the AE and training it to reconstruct the uncorrupted, original version of the input data.
@@
 DAE-GP builds upon the concept of using DAEs in EDAs described by {[}\protect\hyperlink{ref-harmless_overfitting_eda}{21}{]} and transfers the concept to the domain of GP.
-The mutation and crossover operators of standard GP are replaced by sampling new solutions from a probabilistic model that is build by training a DAEs long short-term memory (LSTM) network on a subset of high fitness solutions from the current population {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{33}{]}.
+The mutation and crossover operators of standard GP are replaced by sampling new solutions from a probabilistic model that is build by training a DAEs long short-term memory (LSTM) network on a subset of high fitness solutions from the current population {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{34}{]}.
 
 LSTMs are a variant of artificial neural networks first introduced by {[}\protect\hyperlink{ref-lstm_orig}{11}{]} that can store learned information over en extended time period while avoiding the problem of vanishing and/or exploding gradients.
-Since DAE-GP encodes candidate solutions as linear strings in prefix notation, the DAE in DAE-GP uses LSTMs for both encoding and decoding where the total amount of time steps \(T\) is equal to sum of the length of the input solution and the output solution {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{33}{]}.
+Since DAE-GP encodes candidate solutions as linear strings in prefix notation, the DAE in DAE-GP uses LSTMs for both encoding and decoding where the total amount of time steps \(T\) is equal to sum of the length of the input solution and the output solution {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{34}{]}.
 
@@
 Instead of initializing and optimizing each model from scratch as done in previous work (e.g.
-{[}\protect\hyperlink{ref-dae-gp_2020_rtree}{33}{]} {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{32}{]}), a separate DAE-LSTM network \(\hat{M}\) will be trained on a large population of randomly initialized solutions.\\
+{[}\protect\hyperlink{ref-dae-gp_2020_rtree}{34}{]} {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{33}{]}), a separate DAE-LSTM network \(\hat{M}\) will be trained on a large population of randomly initialized solutions.\\
 The trainable parameters \(\theta_{\hat{M}}\) obtained after finishing the training procedure of \(\hat{M}\) are then used as the starting parameters for each following DAE-LSTM model \(M_g\) for \(g\in{1,...,g_{max}}\).
@@
 
-The DAE-GP algorithm, first described by {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{33}{]}, is summarized as a flowchart in figure \ref{fig:daegp-flowchart}.
+The DAE-GP algorithm, first described by {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{34}{]}, is summarized as a flowchart in figure \ref{fig:daegp-flowchart}.
 After setting the generation counter \(g\) to \(0\) and creating the initial population of solutions \(P_0\), DAE-GP enters the main loop that checks if a termination criteria is satisfied (i.e.~maximum number of generations or fitness evaluations is reached).
@@
 One difficulty in the implementation of a pre-training strategy into DAE-GP has been the determination of the number of hidden neurons in side the DAE-LSTMs hidden layers.
-The original description of DAE-GP {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{33}{]} uses a strategy where the number of hidden neurons for each hidden layer is dynamically set per generation to the maximum individual size inside the current population.
+The original description of DAE-GP {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{34}{]} uses a strategy where the number of hidden neurons for each hidden layer is dynamically set per generation to the maximum individual size inside the current population.
 For a pre-training implementation, this strategy can not be easily adapted since it leads to a changing number of neurons at each generation resulting in different dimensions of the DAE-LSTM.
@@
 \endlastfoot
-Stopping Condition & Convergence {[}\protect\hyperlink{ref-daegp_explore_exploit}{31}{]} \\
+Stopping Condition & Convergence {[}\protect\hyperlink{ref-daegp_explore_exploit}{32}{]} \\
 Reconstruction Error Metric & Multiclass Cross Entropy \\
@@
 Hidden Neurons & 150 \\
-Corruption Operator & Levenshtein Edit {[}\protect\hyperlink{ref-daegp_explore_exploit}{31}{]} \\
+Corruption Operator & Levenshtein Edit {[}\protect\hyperlink{ref-daegp_explore_exploit}{32}{]} \\
 Denoising Edit Probability (Model Building) & 0.05 \\
@@
 The framework that was used to run all experiments of this thesis was provided by the supervisor of this thesis David Wittenberg and uses the python programming language {[}\protect\hyperlink{ref-python3}{26}{]} in conjunction with the baseline libraries \texttt{Keras} {[}\protect\hyperlink{ref-chollet2015keras}{2}{]} for deep neural networks and \texttt{deap} {[}\protect\hyperlink{ref-DEAP_JMLR2012}{6}{]} for evolutionary computation.
-For data visualizations I used the python package \texttt{matplotlib} {[}\protect\hyperlink{ref-matplotlib}{12}{]}, statistical analysis and evaluations were conducted using primarily the data science libraries \texttt{scipy} {[}\protect\hyperlink{ref-scipy}{28}{]}, \texttt{numpy} {[}\protect\hyperlink{ref-numpy}{8}{]}, \texttt{pandas} {[}\protect\hyperlink{ref-pandas}{24}{]}, and \texttt{IPython} {[}\protect\hyperlink{ref-ipython}{20}{]}.
+For data visualizations I used the python package \texttt{matplotlib} {[}\protect\hyperlink{ref-matplotlib}{12}{]}, statistical analysis and evaluations were conducted using primarily the data science libraries \texttt{scipy} {[}\protect\hyperlink{ref-scipy}{29}{]}, \texttt{numpy} {[}\protect\hyperlink{ref-numpy}{8}{]}, \texttt{pandas} {[}\protect\hyperlink{ref-pandas}{24}{]}, and \texttt{IPython} {[}\protect\hyperlink{ref-ipython}{20}{]}.
 
@@
 The objective of the airfoil problem is to find a function that accurately predicts the output variable \(y\) by taking in a subset of the input variables \(x1,x2,x3,x4,x5\).
-The function set used by all DAE-GP variations for symbolic regression is summarized in tables \ref{tab:baseParams-GP} and , the terminal set consists of the the input variables \(x1,x2,x3,x4,x5\) and the ephemeral integer constants listed in table \ref{tab:baseParams-GP} {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{32}{]}.
+The function set used by all DAE-GP variations for symbolic regression is summarized in tables \ref{tab:baseParams-GP} and , the terminal set consists of the the input variables \(x1,x2,x3,x4,x5\) and the ephemeral integer constants listed in table \ref{tab:baseParams-GP} {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{33}{]}.
 
@@
 In general I expect that with a growing number of trainable parameters (either by adding more hidden layers or more hidden neurons per layer) the DAE-LSTM will be more prone to overfit to the training population \(P_{train}\) resulting in a small reconstruction error for the training population and a large one for the validation population \(P_{test}\).
-The reason for this effect is that a larger network trained over an extended period of time (without the use of strategies like early stopping), has much more potential to learn noise from the training dataset than a smaller network, which is more likely to result in worsening performance for previously unseen cases {[}\protect\hyperlink{ref-weigend1994overfitting}{29}{]}.
+The reason for this effect is that a larger network trained over an extended period of time (without the use of strategies like early stopping), has much more potential to learn noise from the training dataset than a smaller network, which is more likely to result in worsening performance for previously unseen cases {[}\protect\hyperlink{ref-weigend1994overfitting}{30}{]}.
 
@@
 
-Since previous work by {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{32}{]} demonstrated that DAE-GP, for a given number of 10.000 fitness evaluations, is able to produce solutions that are much smaller in tree size than regular GP, I also studied the influence of pre-training on the size of solutions found by DAE-GP.
+Since previous work by {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{33}{]} demonstrated that DAE-GP, for a given number of 10.000 fitness evaluations, is able to produce solutions that are much smaller in tree size than regular GP, I also studied the influence of pre-training on the size of solutions found by DAE-GP.
 Figure \ref{fig:airfoil-size-combined-30gens} shows the median of both the average size of individuals inside the population as well as the size of the currently best performing solution inside the population.
@@
 To ensure a fair comparison between pre-trained and regular DAE-GP, the prior experiments used a static number of hidden neurons for each hidden Layer inside the DAE-LSTM networks.
-For comparison, figure \ref{fig:airfoil-maxIndSize-params} shows the total number of trainable parameters for each generation's DAE-LSTM of another experiment where regular DAE-GP was tested with dynamic adjustment of the number of hidden neurons as described in {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{33}{]} and {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{32}{]}.
+For comparison, figure \ref{fig:airfoil-maxIndSize-params} shows the total number of trainable parameters for each generation's DAE-LSTM of another experiment where regular DAE-GP was tested with dynamic adjustment of the number of hidden neurons as described in {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{34}{]} and {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{33}{]}.
 Again, the experiments was based on \(10\) individual runs and otherwise used the same hyperparameters as listed in tables \ref{tab:baseParams-GP} and \ref{tab:baseParams-DAEGP}.
@@
 
-Since previous research on DAE-GP used a single hidden layers for the airfoil dataset {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{32}{]}, I repeated the experiment with a single hidden layer.
+Since previous research on DAE-GP used a single hidden layers for the airfoil dataset {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{33}{]}, I repeated the experiment with a single hidden layer.
 Based on the findings of {[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{5}{]}, I expected that pre-training DAE-GP would lead to a decrease in overall performance.
@@
 Energy\_Cooling & 768 & 8 & {[}\protect\hyperlink{ref-ENERGY_DATASET}{25}{]} \\
-Concrete & 1030 & 8 & {[}\protect\hyperlink{ref-CONCRETE_DATASET}{34}{]} \\
+Concrete & 1030 & 8 & {[}\protect\hyperlink{ref-CONCRETE_DATASET}{35}{]} \\
 \end{longtable}
 
-To test for statistical differences of the results achieved by either pre-trained or regular DAE-GP, I selected the nonparametric Mann-Whitney-U Test ({[}\protect\hyperlink{ref-mwu1}{30}{]}, {[}\protect\hyperlink{ref-mwu2}{16}{]}) to test for differences in the distribution underlying the samples sourced from my experiments.
+To test for statistical differences of the results achieved by either pre-trained or regular DAE-GP, I selected the nonparametric Mann-Whitney-U Test ({[}\protect\hyperlink{ref-mwu1}{31}{]}, {[}\protect\hyperlink{ref-mwu2}{16}{]}) to test for differences in the distribution underlying the samples sourced from my experiments.
 Additionally, I included the effect strength of pre-training as measured by Cliffs Delta {[}\protect\hyperlink{ref-Cliff1993DominanceSO}{3}{]} into the statistical analysis.
@@
 Note that the results for the real world symbolic regression problems inside this chapter, except for the airfoil problem (see subsection \protect\hyperlink{influence-on-run-time}{Influence on Run-Time}), have not been analyzed for their total run-time.
-This is due to the fact, that experiments were run on different computers architectures which resulted in strongly biased results.
+This is due to the fact, that experiments were run on different computers architectures which would result in biased results.
 
@@
 
+\begin{figure}[c]
+
+{\centering \includegraphics[width=0.9\linewidth]{img/symbolicRegressionSummarized/median_fitness_byGens}
+
+}
+
+\caption{Fitness over 30 Generations - Real World Symbolic Regression}\label{fig:symReg-Fitness}
+\end{figure}
+
 The results for the median best fitness over the evolutionary search are visualized in figure \ref{fig:symReg-Fitness}.
-Figure \ref{fig:symReg-Fitness-BoxPlots} shows the distribution of final fitness values on both the test and training set as box plots.
-
-\begin{figure}[c]
-
-{\centering \includegraphics[width=0.9\linewidth]{img/symbolicRegressionSummarized/median_fitness_byGens}
-
-}
-
-\caption{Fitness over 30 Generations - Real World Symbolic Regression}\label{fig:symReg-Fitness}
-\end{figure}
-
-\begin{figure}[c]
-
-{\centering \includegraphics[width=0.9\linewidth]{img/symbolicRegressionSummarized/final_fitness_boxplot}
-
-}
-
-\caption{Fitness after 30 Generations - Real World Symbolic Regression}\label{fig:symReg-Fitness-BoxPlots}
-\end{figure}
-
-Table \ref{tab:full-run-realWorldSymReg-fitness} summarizes the results for the median best fitness achieved after 30 generations and shows the results of computing the statistical tests.
+In general, pre-training and regular DAE-GP both are able to improve on the best fitness found over the evolutionary search.
+Looking at figure \ref{fig:symReg-Fitness}, it appears that no significant differences for the best fitness between both algorithms exist.
+Regular DAE-GP achieves a slightly lower RMSE on the test set for Airfoil, Boston Housing and the Energy Cooling dataset while pre-trained DAE-GP is slightly better for the Concrete dataset.
 
@@
 
+Table \ref{tab:full-run-realWorldSymReg-fitness} summarizes the results for the median best fitness achieved after 30 generations and shows the results of computing the statistical tests.
+As expected after reviewing figure \ref{fig:symReg-Fitness}, none of the differences in median best fitness are statistically significant and pre-training has only a very small, negligible effect strength for Cliff's Delta ranging between -0.04 and 0.17 {[}\protect\hyperlink{ref-Vargha2000ACA}{27}{]}.
+
+Figure \ref{fig:symReg-Fitness-BoxPlots} shows the distribution of final fitness values on both the test and training set as box plots.
+It shows that depending on the symbolic regression problem, the dispersion of fitness values is different for both DAE-GP variants.
+For example, a comparativley large span is visible for the airfoil problem while fitness scores for Energy Cooling are much more concentrated.
+The distribution are nontheless similar for both algorithms which does not suggest any meaningful differences between using pre-training and the distribution of final fitness scores.
+
+\begin{figure}[c]
+
+{\centering \includegraphics[width=0.9\linewidth]{img/symbolicRegressionSummarized/final_fitness_boxplot}
+
+}
+
+\caption{Fitness after 30 Generations - Real World Symbolic Regression}\label{fig:symReg-Fitness-BoxPlots}
+\end{figure}
+
 \hypertarget{influence-on-solution-size}{%
@@
 For a single hidden layer, pre-training did show a negative impact on the median final fitness of solutions found as it was suggested by {[}\protect\hyperlink{ref-pmlr-v9-erhan10a}{4}{]}.
-Since current research on DAE-GP for real-world symbolic regression relies on a single hidden layer (see {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{32}{]}), pre-training does come with additional computational expenses for using deeper DAE-LSTMs.
+Since current research on DAE-GP for real-world symbolic regression relies on a single hidden layer (see {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{33}{]}), pre-training does come with additional computational expenses for using deeper DAE-LSTMs.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Vargha2000ACA}{}}%
+\CSLLeftMargin{{[}27{]} }%
+\CSLRightInline{András Vargha and Harold D Delaney.
+2000.
+A critique and improvement of the CL common language effect size statistics of McGraw and wong.
+\emph{Journal of Educational and Behavioral Statistics} 25, (2000), 101--132.}
+
 \leavevmode\vadjust pre{\hypertarget{ref-dae_orig2008}{}}%
-\CSLLeftMargin{{[}27{]} }%
+\CSLLeftMargin{{[}28{]} }%
 \CSLRightInline{Pascal Vincent, Hugo Larochelle, Y.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-scipy}{}}%
-\CSLLeftMargin{{[}28{]} }%
+\CSLLeftMargin{{[}29{]} }%
 \CSLRightInline{Pauli Virtanen, Ralf Gommers, Travis E.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-weigend1994overfitting}{}}%
-\CSLLeftMargin{{[}29{]} }%
+\CSLLeftMargin{{[}30{]} }%
 \CSLRightInline{Andreas Weigend. 1994. On overfitting and the effective number of hidden units. In \emph{Proceedings of the 1993 connectionist models summer school}, 335--342.}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-mwu1}{}}%
-\CSLLeftMargin{{[}30{]} }%
+\CSLLeftMargin{{[}31{]} }%
 \CSLRightInline{Frank. Wilcoxon. 1945. Individual comparisons by ranking methods. \emph{Biometrics} 1, (1945), 196--202.}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-daegp_explore_exploit}{}}%
-\CSLLeftMargin{{[}31{]} }%
+\CSLLeftMargin{{[}32{]} }%
 \CSLRightInline{David Wittenberg.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-dae-gp_2022_symreg}{}}%
-\CSLLeftMargin{{[}32{]} }%
+\CSLLeftMargin{{[}33{]} }%
 \CSLRightInline{David Wittenberg and Franz Rothlauf.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-dae-gp_2020_rtree}{}}%
-\CSLLeftMargin{{[}33{]} }%
+\CSLLeftMargin{{[}34{]} }%
 \CSLRightInline{David Wittenberg, Franz Rothlauf, and Dirk Schweim.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-CONCRETE_DATASET}{}}%
-\CSLLeftMargin{{[}34{]} }%
+\CSLLeftMargin{{[}35{]} }%
 \CSLRightInline{I-Cheng Yeh.

### 2023-02-08 paper.Rmd
autogit

@@
 higher fitness) are printed bold and p-values are marked with asterisk (1,2,3) for $\alpha$ levels ($0.1$, $0.05$, $0.01$).
+Note that I included the results for the airfoil problem with a single hidden layer into the statistical evaluation which is not visualized during this chapter (see subsection [Reduced Number of Hidden Layers]).
 
-Note that the results for the real world symbolic regression problems inside this chapter, except for the airfoil problem (see subsection [Influence on Run-Time]), have not been analyzed for their total run-time.
+Also note that the results for the real world symbolic regression problems inside this chapter, except for the airfoil problem (see subsection [Influence on Run-Time]), have not been analyzed for their total run-time.
 This is due to the fact, that experiments were run on different computers architectures which would result in biased results.
@@
 Table \@ref(tab:full-run-realWorldSymReg-fitness) summarizes the results for the median best fitness achieved after 30 generations and shows the results of computing the statistical tests.
+
 As expected after reviewing figure \@ref(fig:symReg-Fitness), none of the differences in median best fitness are statistically significant and pre-training has only a very small, negligible effect strength for Cliff's Delta ranging between `r min(df_symreg_fitness$Cliffs_Delta)` and `r max(df_symreg_fitness$Cliffs_Delta)` [@Vargha2000ACA].
@@
 Figure \@ref(fig:symReg-Size) shows the median size of the best solution per generation aggregated over all benchmark problems for symbolic regression.
+Similiar to the previous results regarding the fitness of the best found solution, no significant differences in the size of the best found solution are visible between both DAE-GP variants.
 
@@
 ```
-
-Table \@ref(tab:full-run-realWorldSymReg-best-solution-size) shows the median size of the best solution after 30 generations and results for computing the statistical tests.
 
@@
 # Create the data frame from csv
-df <- read.csv(
+df_size <- read.csv(
   "/Users/rmn/github/master_thesis/data/summary_table_best_size_median.csv",
@@
 # Create a new dataframe with the modified values of the two columns,
-df_bold <- df
+df_bold <- df_size
 
-for(i in 1:nrow(df)){
-  if(df[i, "DAE.GP"] < df[i, "Pre.Trained_DAE.GP"]){
-    df_bold[i, "DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(round(df[i, "DAE.GP"],4)))
-    df_bold[i, "Pre.Trained_DAE.GP"] <- as.character(round(df[i, "Pre.Trained_DAE.GP"],4))
+for(i in 1:nrow(df_size)){
+  if(df_size[i, "DAE.GP"] < df_size[i, "Pre.Trained_DAE.GP"]){
+    df_bold[i, "DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(round(df_size[i, "DAE.GP"],4)))
+    df_bold[i, "Pre.Trained_DAE.GP"] <- as.character(round(df_size[i, "Pre.Trained_DAE.GP"],4))
     }
-  else if (df[i, "DAE.GP"] > df[i, "Pre.Trained_DAE.GP"]) {
-    df_bold[i, "Pre.Trained_DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(round(df[i, "Pre.Trained_DAE.GP"],4)))
-  df_bold[i, "DAE.GP"] <- as.character(round(df[i, "DAE.GP"],4))
+  else if (df_size[i, "DAE.GP"] > df_size[i, "Pre.Trained_DAE.GP"]) {
+    df_bold[i, "Pre.Trained_DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(round(df_size[i, "Pre.Trained_DAE.GP"],4)))
+  df_bold[i, "DAE.GP"] <- as.character(round(df_size[i, "DAE.GP"],4))
   }
@@
 ```
+Table \@ref(tab:full-run-realWorldSymReg-best-solution-size) shows the median size of the best solution after 30 generations and results for computing the statistical tests.
+For Airfoil, Boston Housing and Energy Cooling datasets, the median size of the best soultion are equal, only for the concrete dataset a smaller median size for pre-trained DAE-GP was achieved (pre-trained DAE-GP also showed a higher median best fitness for the concrete problem).
+The results of the Mann-Whitney U Test again show, that differences in the underlying distributions are not statistically significant which also shows in low effect strenghts of using pre-training on the size of the best found solutions (Cliff's Delta ranges from`r min(df_size$Cliffs_Delta)` to `r max(df_size$Cliffs_Delta)`).
+
 ## Influence on Population Diversity

### 2023-02-08 paper.tex
autogit

@@
 Better median values (e.g.~higher fitness) are printed bold and p-values are marked with asterisk (1,2,3) for \(\alpha\) levels (\(0.1\), \(0.05\), \(0.01\)).
-
-Note that the results for the real world symbolic regression problems inside this chapter, except for the airfoil problem (see subsection \protect\hyperlink{influence-on-run-time}{Influence on Run-Time}), have not been analyzed for their total run-time.
+Note that I included the results for the airfoil problem with a single hidden layer into the statistical evaluation which is not visualized during this chapter (see subsection \protect\hyperlink{reduced-number-of-hidden-layers}{Reduced Number of Hidden Layers}).
+
+Also note that the results for the real world symbolic regression problems inside this chapter, except for the airfoil problem (see subsection \protect\hyperlink{influence-on-run-time}{Influence on Run-Time}), have not been analyzed for their total run-time.
 This is due to the fact, that experiments were run on different computers architectures which would result in biased results.
@@
 Table \ref{tab:full-run-realWorldSymReg-fitness} summarizes the results for the median best fitness achieved after 30 generations and shows the results of computing the statistical tests.
+
 As expected after reviewing figure \ref{fig:symReg-Fitness}, none of the differences in median best fitness are statistically significant and pre-training has only a very small, negligible effect strength for Cliff's Delta ranging between -0.04 and 0.17 {[}\protect\hyperlink{ref-Vargha2000ACA}{27}{]}.
@@
 Figure \ref{fig:symReg-Size} shows the median size of the best solution per generation aggregated over all benchmark problems for symbolic regression.
+Similiar to the previous results regarding the fitness of the best found solution, no significant differences in the size of the best found solution are visible between both DAE-GP variants.
 
@@
 \end{figure}
-
-Table \ref{tab:full-run-realWorldSymReg-best-solution-size} shows the median size of the best solution after 30 generations and results for computing the statistical tests.
 
@@
 \end{longtable}
+
+Table \ref{tab:full-run-realWorldSymReg-best-solution-size} shows the median size of the best solution after 30 generations and results for computing the statistical tests.
+For Airfoil, Boston Housing and Energy Cooling datasets, the median size of the best soultion are equal, only for the concrete dataset a smaller median size for pre-trained DAE-GP was achieved (pre-trained DAE-GP also showed a higher median best fitness for the concrete problem).
+The results of the Mann-Whitney U Test again show, that differences in the underlying distributions are not statistically significant which also shows in low effect strenghts of using pre-training on the size of the best found solutions (Cliff's Delta ranges from-0.02 to 0.20).
 

### 2023-02-09 paper.Rmd
Thu Feb  9 14:00:07 CET 2023: update-remote.sh

@@
 <!-- LaTeX Setup -->
+
 \setcounter{page}{1}
+
 <!-- \tableofcontents -->
+
+```{=tex}
 \thispagestyle{empty}
@@
 \newpage
-
+```
 <!--Markdown Text -->
 
-# Abstract {.unnumbered #I}
+# Abstract {#I .unnumbered}
+
 ...
@@
 <!--- Describe experiments... -->
+
 ...
@@
 
-AE were introduced by @ae_orig to compress high-dimensional data into lower-dimensions. An AE consists of two different sub-units:
+AE were introduced by @ae_orig to compress high-dimensional data into lower-dimensions.
+An AE consists of two different sub-units:
 
-  * Encoder $g(x)$: Encode input data to a smaller central layer $h$
-  * Decoder $d(h)$: Decode and output the the encoded data back to its original dimension
+-   Encoder $g(x)$: Encode input data to a smaller central layer $h$
+-   Decoder $d(h)$: Decode and output the the encoded data back to its original dimension
 
@@
 Instead of initializing and optimizing each model from scratch as done in previous work (e.g.
-[@dae-gp_2020_rtree] [@dae-gp_2022_symreg]), a separate DAE-LSTM network $\hat{M}$ will be trained on a large population of randomly initialized solutions.
+[@dae-gp_2020_rtree] [@dae-gp_2022_symreg]), a separate DAE-LSTM network $\hat{M}$ will be trained on a large population of randomly initialized solutions.\
+
 The trainable parameters $\theta_{\hat{M}}$ obtained after finishing the training procedure of $\hat{M}$ are then used as the starting parameters for each following DAE-LSTM model $M_g$ for $g\in{1,...,g_{max}}$.
@@
 <!--- Symbolic regression -->
+
 To test pre-training in DAE-GP this thesis focuses on the domain of real-world symbolic regression problems.
@@
 <!--- airfoil dataset -->
+
 The main experiments conducted in this thesis uses the NASA Airfoil Self-Noise Data Set [@AIRFOIL_DATASET] that consists of 5 input variables and 1 output variable that are listed in table \@ref(tab:airfoil-dataset-description).
@@
 
-* Variable number of hidden neurons (50, 100, 200) with a single hidden layer
-* Fixed number of hidden neurons (100) per hidden layer with variable number of hidden layers (1, 2, 3)
+-   Variable number of hidden neurons (50, 100, 200) with a single hidden layer
+-   Fixed number of hidden neurons (100) per hidden layer with variable number of hidden layers (1, 2, 3)
 
@@
 ```
+
 ### Dynamic Adjustment of Hidden Neurons
@@
 
-<!-- An interesting observation from studying the distribution of final fitness scores for both algorithms is that pre-trained DAE-GP has a smaller interquartile range and more extreme outliers for high RMSE which explains the slightly worse performance if looking at median best fitness present in figure \@ref(fig:airfoil-median-fitness-30gens).
+<!-- An interesting observation from studying the distribution of final fitness scores for both algorithms is that pre-trained DAE-GP has a smaller interquartile range and more extreme outliers for high RMSE which explains the slightly worse performance if looking at median best fitness present in figure @ref(fig:airfoil-median-fitness-30gens).
 -->
@@
 ```
+
 Figure \@ref(fig:airfoil-runtime-boxplot) shows the distribution of the total runtime for both DAE-GP variants.
@@
 ```
+
 While investigating other reasons for the differences in run-time, I verified that both DAE-GP variants use the exact same number of trainable parameters for each generations DAE-LSTMS models (total number of trainable parameters DAE-GP=`r format(REG_PARAMS, scientific=F)`, Pre-trained DAE-GP=`r format(PT_PARAMS, scientific=F)`).
@@
 ```
+
 To test for statistical differences of the results achieved by either pre-trained or regular DAE-GP, I selected the nonparametric Mann-Whitney-U Test ([@mwu1], [@mwu2]) to test for differences in the distribution underlying the samples sourced from my experiments.
 Additionally, I included the effect strength of pre-training as measured by Cliffs Delta [@Cliff1993DominanceSO] into the statistical analysis.
-Better median values (e.g.
-higher fitness) are printed bold and p-values are marked with asterisk (1,2,3) for $\alpha$ levels ($0.1$, $0.05$, $0.01$).
+Better median values (e.g. higher fitness) are printed bold and p-values are marked with asterisk (1,2,3) for $\alpha$ levels ($0.1$, $0.05$, $0.01$).
 Note that I included the results for the airfoil problem with a single hidden layer into the statistical evaluation which is not visualized during this chapter (see subsection [Reduced Number of Hidden Layers]).
@@
 ```
+
 Table \@ref(tab:full-run-realWorldSymReg-best-solution-size) shows the median size of the best solution after 30 generations and results for computing the statistical tests.
@@
 
-The results of the statistical analysis is summarized in table \@ref(tab:full-run-realWorldSymReg-popDiversity).
-
 ```{r full-run-realWorldSymReg-popDiversity, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
@@
 # Create the data frame from csv
-df <- read.csv(
+df_popDiv <- read.csv(
   "/Users/rmn/github/master_thesis/data/summary_table_median_LevDistance.csv",
@@
 # Create a new dataframe with the modified values of the two columns,
-df_bold <- df
+df_bold <- df_popDiv
 
-for(i in 1:nrow(df)){
-  if(df[i, "DAE.GP"] > df[i, "Pre.Trained_DAE.GP"]){
-    df_bold[i, "DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(round(df[i, "DAE.GP"],2)))
-    df_bold[i, "Pre.Trained_DAE.GP"] <- as.character(round(df[i, "Pre.Trained_DAE.GP"],2))
+for(i in 1:nrow(df_popDiv)){
+  if(df_popDiv[i, "DAE.GP"] > df_popDiv[i, "Pre.Trained_DAE.GP"]){
+    df_bold[i, "DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(round(df_popDiv[i, "DAE.GP"],2)))
+    df_bold[i, "Pre.Trained_DAE.GP"] <- as.character(round(df_popDiv[i, "Pre.Trained_DAE.GP"],2))
     }
-  else if (df[i, "DAE.GP"] < df[i, "Pre.Trained_DAE.GP"]) {
-    df_bold[i, "Pre.Trained_DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(round(df[i, "Pre.Trained_DAE.GP"],2)))
-  df_bold[i, "DAE.GP"] <- as.character(round(df[i, "DAE.GP"],2))
+  else if (df_popDiv[i, "DAE.GP"] < df_popDiv[i, "Pre.Trained_DAE.GP"]) {
+    df_bold[i, "Pre.Trained_DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(round(df_popDiv[i, "Pre.Trained_DAE.GP"],2)))
+  df_bold[i, "DAE.GP"] <- as.character(round(df_popDiv[i, "DAE.GP"],2))
   }
@@
 ```
+
+The results of the statistical analysis is summarized in table \@ref(tab:full-run-realWorldSymReg-popDiversity).
+For all sub experiments, computing the Mann-Whitney U Test returns P-Values that are smaller than 0.01 which indicates a highly statistical significant difference in the distributions.
+Looking at the effect strength of pre-training on population diversity, it shows that pre-training does have a strong positive effect (Cliffs Delta ranges from \`r \`)in improving the levenshtein distance between individual solutions inside the population $P_g$
 
@@
 The experiments conducted in the scope of this master thesis did find that the implemented pre-training strategy resulted in an increased population diversity measured by the normalized levenshtein edit distance.
-Increasing the population diversity can be an important mechanism in controlling the search behavior of population based meta heuristics and is especially useful for increasing the exploration of the search space of possible solutions !QUELLE!.
+Increasing the population diversity can be an important mechanism in controlling the search behavior of population based meta heuristics and is especially useful for increasing the exploration of the search space of possible solutions !Q
+UELLE!.
 
@@
 
-# Nomenclature {.unnumbered #II}
+# Nomenclature {#II .unnumbered}
 
-## Symbols {.unnumbered #IIa}
+## Symbols {#IIa .unnumbered}
 
@@
 ```
-## Abbreviations {.unnumbered #IIb}
+
+## Abbreviations {#IIb .unnumbered}
 
@@
 
-# References {.unnumbered #III}
+# References {#III .unnumbered}
 
@@
 
-# Statutory Declaration {.unnumbered #IV}
+# Statutory Declaration {#IV .unnumbered}
 

### 2023-02-09 paper.Rmd
Thu Feb  9 15:00:07 CET 2023: update-remote.sh

@@
 
-<!--- Gang der Untersuchung -->
-
-<!--- Describe experiments... -->
-
-...
-
 # Theoretical Foundations
@@
 EDAs use this probabilistic model to sample new offspring solutions inside an evolutionary search to replace crossover and/or mutation operators.
-
-### EDA-GP
-
-... insert some smart stuff here...
+Even though a majority of EDA algorithms are designed to work with fixed length string representations as used in classical GAs, EDA algorithms have been researched an successfully applied to the domain of GP in various different research streams.
 
@@
 It shows that depending on the symbolic regression problem, the dispersion of fitness values is different for both DAE-GP variants.
-For example, a comparativley large span is visible for the airfoil problem while fitness scores for Energy Cooling are much more concentrated.
-The distribution are nontheless similar for both algorithms which does not suggest any meaningful differences between using pre-training and the distribution of final fitness scores.
+For example, a comparatively large span is visible for the airfoil problem while fitness scores for Energy Cooling are much more concentrated.
+The distribution are nonetheless similar for both algorithms which does not suggest any meaningful differences between using pre-training and the distribution of final fitness scores.
 
@@
 Figure \@ref(fig:symReg-Size) shows the median size of the best solution per generation aggregated over all benchmark problems for symbolic regression.
-Similiar to the previous results regarding the fitness of the best found solution, no significant differences in the size of the best found solution are visible between both DAE-GP variants.
+Similar to the previous results regarding the fitness of the best found solution, no significant differences in the size of the best found solution are visible between both DAE-GP variants.
 
@@
 Table \@ref(tab:full-run-realWorldSymReg-best-solution-size) shows the median size of the best solution after 30 generations and results for computing the statistical tests.
-For Airfoil, Boston Housing and Energy Cooling datasets, the median size of the best soultion are equal, only for the concrete dataset a smaller median size for pre-trained DAE-GP was achieved (pre-trained DAE-GP also showed a higher median best fitness for the concrete problem).
-The results of the Mann-Whitney U Test again show, that differences in the underlying distributions are not statistically significant which also shows in low effect strenghts of using pre-training on the size of the best found solutions (Cliff's Delta ranges from`r min(df_size$Cliffs_Delta)` to `r max(df_size$Cliffs_Delta)`).
+For Airfoil, Boston Housing and Energy Cooling data-sets, the median size of the best solution are equal, only for the concrete dataset a smaller median size for pre-trained DAE-GP was achieved (pre-trained DAE-GP also showed a higher median best fitness for the concrete problem).
+The results of the Mann-Whitney-U Test again show, that differences in the underlying distributions are not statistically significant which also shows in low effect strengths of using pre-training on the size of the best found solutions (Cliff's Delta ranges from`r min(df_size$Cliffs_Delta)` to `r max(df_size$Cliffs_Delta)`).
 
@@
 The results of the statistical analysis is summarized in table \@ref(tab:full-run-realWorldSymReg-popDiversity).
-For all sub experiments, computing the Mann-Whitney U Test returns P-Values that are smaller than 0.01 which indicates a highly statistical significant difference in the distributions.
-Looking at the effect strength of pre-training on population diversity, it shows that pre-training does have a strong positive effect (Cliffs Delta ranges from \`r \`)in improving the levenshtein distance between individual solutions inside the population $P_g$
+For all sub experiments, computing the Mann-Whitney-U Test returns P-Values that are smaller than $0.01$ which indicates a highly statistically significant difference in the underlying distributions.
+Looking at the effect strength of pre-training, it shows that pre-training does have a strong positive effect (Cliff's Delta ranges from `r min(df_popDiv$Cliffs_Delta)` to `r max(df_popDiv$Cliffs_Delta)`) on population diversity inside the population.
+Pre-Training did increase the median normalized levenshtein distance over all generations and sub-experiments by a factor of `r round( median(df_popDiv$Pre.Trained_DAE.GP) / median(df_popDiv$DAE.GP) , 3)` from `r round(median(df_popDiv$DAE.GP) , 3)` to `r round(median(df_popDiv$Pre.Trained_DAE.GP) , 3)`.
 
@@
 
-Figure \@ref(fig:symReg-EpochsPerGen) shows the median number of epochs spend on training each generations DAE-LSTM network until the termination criterion was satisfied.
+Finally, figure \@ref(fig:symReg-EpochsPerGen) shows the median number of epochs spend on training each generations DAE-LSTM network $M_g$ until the termination criterion was satisfied.
+Pre-Training shows to be very effective at reducing the number of training epochs in comparison to regular DAE-GP.
 
@@
 ```
-
-The results of the statistical analysis is summarized in table \@ref(tab:full-run-realWorldSymReg-epochsPerGen).
 
@@
 # Create the data frame from csv
-df <- read.csv(
+df_epochs <- read.csv(
   "/Users/rmn/github/master_thesis/data/summary_table_mean_epochsTrained.csv",
@@
 # Create a new dataframe with the modified values of the two columns,
-df_bold <- df
+df_bold <- df_epochs
 
-for(i in 1:nrow(df)){
-  if(df[i, "DAE.GP"] < df[i, "Pre.Trained_DAE.GP"]){
-    df_bold[i, "DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(round(df[i, "DAE.GP"],2)))
-    df_bold[i, "Pre.Trained_DAE.GP"] <- as.character(round(df[i, "Pre.Trained_DAE.GP"],2))
+for(i in 1:nrow(df_epochs)){
+  if(df_epochs[i, "DAE.GP"] < df_epochs[i, "Pre.Trained_DAE.GP"]){
+    df_bold[i, "DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(round(df_epochs[i, "DAE.GP"],2)))
+    df_bold[i, "Pre.Trained_DAE.GP"] <- as.character(round(df_epochs[i, "Pre.Trained_DAE.GP"],2))
     }
-  else if (df[i, "DAE.GP"] > df[i, "Pre.Trained_DAE.GP"]) {
-    df_bold[i, "Pre.Trained_DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(round(df[i, "Pre.Trained_DAE.GP"],2)))
-  df_bold[i, "DAE.GP"] <- as.character(round(df[i, "DAE.GP"],2))
+  else if (df_epochs[i, "DAE.GP"] > df_epochs[i, "Pre.Trained_DAE.GP"]) {
+    df_bold[i, "Pre.Trained_DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(round(df_epochs[i, "Pre.Trained_DAE.GP"],2)))
+  df_bold[i, "DAE.GP"] <- as.character(round(df_epochs[i, "DAE.GP"],2))
   }
@@
 
+Looking at the results of the statistical analysis summarized in table \@ref(tab:full-run-realWorldSymReg-epochsPerGen), computing the Mann-Whitney-U test does consistently return P-Values below $0.01$ for all experiments which proves the statistical significant differences in distributions.
+Analyzing the effect strength of pre-training on the number of training epochs shows values for Cliff's Delta in the range of `r min(df_epochs$Cliffs_Delta)` and `r max(df_epochs$Cliffs_Delta)` which indicates a very strong negative effect on the number of training epochs necessary at each generation of the search.
+Computing the median number of training epochs over all generations for all benchmark problems shows a total decrease by a factor of `r round( (median(df_epochs$Pre.Trained_DAE.GP) / median(df_epochs$DAE.GP)) , 2)` from `r round(median(df_epochs$DAE.GP), 2)` epochs for regular DAE-GP to `r round(median(df_epochs$Pre.Trained_DAE.GP), 2)` epochs for pre-trained DAE-GP.
+
 # Conclusion and Discussion
@@
 
-The experiments conducted in the scope of this master thesis did find that the implemented pre-training strategy resulted in an increased population diversity measured by the normalized levenshtein edit distance.
-Increasing the population diversity can be an important mechanism in controlling the search behavior of population based meta heuristics and is especially useful for increasing the exploration of the search space of possible solutions !Q
-UELLE!.
+The experiments conducted in the scope of this master thesis did find that the implemented pre-training strategy resulted in a highly statistically significant increase in population diversity measured by the normalized levenshtein edit distance (see subsection [Influence on Population Diversity]).
+Increasing the population diversity can be an important mechanism in controlling the search behavior of evolutionary algorithms and is especially important for the global exploration of the search space and the avoidance of premature convergence of the search [@Sudholt2020].
 
-Another significant benefit of using pre-training in DAE-GP is that it is highly efficient in decreasing the number of training epochs that are necessary to train each generations DAE-LSTM model.
-Even though pre-training increased the median total run-time for the tested benchmark problem in combination with the chosen DAE-LSTM dimensions, reducing the number of training epochs might be a possible way to lower the run time of DAE-GP if further performance optimizations, especially during model sampling, can be achieved.
+Another significant benefit of using pre-training in DAE-GP is that it is highly efficient in decreasing the number of training epochs that are necessary to train each generations DAE-LSTM model (see subsection [Influence on Population Diversity]).
+Even though pre-training increased the median total run-time for the tested benchmark problem in combination with the chosen DAE-LSTM dimensions, reducing the number of training epochs might still be a possible way to lower the run time of DAE-GP if further performance optimizations, especially during model sampling, can be achieved.
 

### 2023-02-09 paper.tex
Thu Feb  9 15:00:07 CET 2023: update-remote.sh

@@
 \author{}
-\date{\vspace{-2.5em}Date of Submission: 2023-02-08}
+\date{\vspace{-2.5em}Date of Submission: 2023-02-09}
 
@@
 
-AE were introduced by {[}\protect\hyperlink{ref-ae_orig}{10}{]} to compress high-dimensional data into lower-dimensions. An AE consists of two different sub-units:
+AE were introduced by {[}\protect\hyperlink{ref-ae_orig}{10}{]} to compress high-dimensional data into lower-dimensions.
+An AE consists of two different sub-units:
 
@@
 {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{34}{]} {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{33}{]}), a separate DAE-LSTM network \(\hat{M}\) will be trained on a large population of randomly initialized solutions.\\
+
 The trainable parameters \(\theta_{\hat{M}}\) obtained after finishing the training procedure of \(\hat{M}\) are then used as the starting parameters for each following DAE-LSTM model \(M_g\) for \(g\in{1,...,g_{max}}\).
@@
 For Airfoil, Boston Housing and Energy Cooling datasets, the median size of the best soultion are equal, only for the concrete dataset a smaller median size for pre-trained DAE-GP was achieved (pre-trained DAE-GP also showed a higher median best fitness for the concrete problem).
-The results of the Mann-Whitney U Test again show, that differences in the underlying distributions are not statistically significant which also shows in low effect strenghts of using pre-training on the size of the best found solutions (Cliff's Delta ranges from-0.02 to 0.20).
+The results of the Mann-Whitney-U Test again show, that differences in the underlying distributions are not statistically significant which also shows in low effect strenghts of using pre-training on the size of the best found solutions (Cliff's Delta ranges from-0.02 to 0.20).
 
@@
 \end{figure}
-
-The results of the statistical analysis is summarized in table \ref{tab:full-run-realWorldSymReg-popDiversity}.
 
@@
 
+The results of the statistical analysis is summarized in table \ref{tab:full-run-realWorldSymReg-popDiversity}.
+For all sub experiments, computing the Mann-Whitney-U Test returns P-Values that are smaller than \(0.01\) which indicates a highly statistically significant difference in the underlying distributions.
+Looking at the effect strength of pre-training, it shows that pre-training does have a strong positive effect (Cliff's Delta ranges from 0.92 to 0.94) on population diversity inside the population.
+Pre-Training did increase the median normalized levenshtein distance over all generations and sub-experiments by a factor of 1.077 from 0.869 to 0.935.
+
 \hypertarget{influence-on-the-number-of-training-epochs-per-generation}{%
@@
 
-Figure \ref{fig:symReg-EpochsPerGen} shows the median number of epochs spend on training each generations DAE-LSTM network until the termination criterion was satisfied.
+Finally, figure \ref{fig:symReg-EpochsPerGen} shows the median number of epochs spend on training each generations DAE-LSTM network \(M_g\) until the termination criterion was satisfied.
+Pre-Training shows to be very effective at reducing the number of training epochs in comparison to regular DAE-GP.
 
@@
 \end{figure}
-
-The results of the statistical analysis is summarized in table \ref{tab:full-run-realWorldSymReg-epochsPerGen}.
 
@@
 
+Looking at the results of the statistical analysis summarized in table \ref{tab:full-run-realWorldSymReg-epochsPerGen}, computing the Mann-Whitney-U test does consistently return P-Values below \(0.01\) for all experiments which proves the statistical significant differences in distributions.
+Analyzing the effect strength of pre-training on the number of training epochs shows values for Cliff's Delta in the range of -0.87 and -0.93 which indicates a very strong negative effect on the number of training epochs necessary at each generation of the search.
+Computing the median number of training epochs over all generations for all benchmark problems shows a total decrease by a factor of 0.45 from 143.49 epochs for regular DAE-GP to 65.15 epochs for pre-trained DAE-GP.
+
 \hypertarget{conclusion-and-discussion}{%

### 2023-02-09 paper.Rmd
autogit

@@
 
-...
+Denoising Autoencoder Genetic Programming (DAE-GP) is an Estimation of Distribution Algorithm in the domain of Genetic Programming that uses Denoising Autoencoders Long Short-Term Memory Networks (DAE-LSTM) as probabilistic models for sampling new populations of solutions.
+This thesis investigates the possible benefits and downsides of using pre-training for the DAE-LSTM networks of DAE-GP for four real world symbolic regression problems.
+The Experiments conducted did show that pre-training can drastically reduce the number of training epochs that are necessary at each generation of the DAE-GP search while also increasing the diversity of the population.
+Unfortunately pre-training did not show any improvements for both final fitness and the final size of solutions while largley increasing the total run-time for DAE-GP.
+
+# Keywords {.unlisted .unnumbered}
+
+Genetic Programming, Estimation of Distribution Algorithms, Denoising Autoencoder Genetic Programming, Pre-Training, Long Short-Term Memory Networks, Symbolic Regression
+
+\newpage
 
@@
 EDAs use this probabilistic model to sample new offspring solutions inside an evolutionary search to replace crossover and/or mutation operators.
-Even though a majority of EDA algorithms are designed to work with fixed length string representations as used in classical GAs, EDA algorithms have been researched an successfully applied to the domain of GP in various different research streams.
+Even though a majority of EDA algorithms are designed to work with fixed length string representations as used in classical GAs, EDA algorithms have been researched an successfully applied to the domain of GP in various different research streams [@kim_probMod_GP].
 
@@
 Some of the main motivations for using pre-training is the ability to reduce the need for large amounts of training data which can often be unavailable or too expensive to collect as well as an improved performance by either reducing the computational effort that is necessary to train a model to solve a specific task or by improving the models generalization ability.
-
-...describe different pre-training strategies, e.g. few shot, perpetual, re-using...
-
--- describe unsupervised vs supervised pre-training...
 

### 2023-02-09 paper.tex
autogit

@@
 
-\ldots{}
+Denoising Autoencoder Genetic Programming (DAE-GP) is an Estimation of Distribution Algorithm in the domain of Genetic Programming that uses Denoising Autoencoders Long Short-Term Memory Networks (DAE-LSTM) as probabilistic models for sampling new populations of solutions.
+This thesis investigates the possible benefits and downsides of using pre-training for the DAE-LSTM networks of DAE-GP for four real world symbolic regression problems.
+The Experiments conducted did show that pre-training can drastically reduce the number of training epochs that are necessary at each generation of the DAE-GP search while also increasing the diversity of the population.
+Unfortunately pre-training did not show any improvements for both final fitness and the final size of solutions while largley increasing the total run-time for DAE-GP.
+
+\hypertarget{keywords}{%
+\section*{Keywords}\label{keywords}}
+
+Genetic Programming, Estimation of Distribution Algorithms, Denoising Autoencoder Genetic Programming, Pre-Training, Long Short-Term Memory Networks, Symbolic Regression
+
+\newpage
 
@@
 
-Denoising Autoencoder Genetic Programming (DAE-GP) is a novel variation of an genetic programming based Estimation of Distribution Algorithm (EDA-GP) that uses a denoising autoencoders long short-term memory network (DAE-LSTMs) as a probabilistic model to sample new candidate solutions {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{34}{]}.
+Denoising Autoencoder Genetic Programming (DAE-GP) is a novel variation of an genetic programming based Estimation of Distribution Algorithm (EDA-GP) that uses a denoising autoencoders long short-term memory network (DAE-LSTMs) as a probabilistic model to sample new candidate solutions {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{36}{]}.
 
@@
 In DAE-GP, DAE-LSTMs are trained with a subset of high-fitness solutions selected from a parent population with the aim to capture their promising qualities.
-The resulting model is then used to sample new offspring solutions by propagating partially mutated solutions from the parent population through the DAE-LSTMS {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{34}{]}.
-In previous work DAE-GP has been shown to outperform GP for both a generalized version of the royal tree problem {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{34}{]} as well as for a real-world symbolic regression problem {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{33}{]}.
-
-The DAE-GP algorithm first described by {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{34}{]} trains a DAE-LSTMs for each generation \(g\) of the search from scratch.
-{[}\protect\hyperlink{ref-dae-gp_2022_symreg}{33}{]} and {[}\protect\hyperlink{ref-daegp_explore_exploit}{32}{]} suggests the incorporation of a pre-training strategy into the evolutionary search as a possible way of improving the overall performance of the DAE-GP algorithm.
+The resulting model is then used to sample new offspring solutions by propagating partially mutated solutions from the parent population through the DAE-LSTMS {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{36}{]}.
+In previous work DAE-GP has been shown to outperform GP for both a generalized version of the royal tree problem {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{36}{]} as well as for a real-world symbolic regression problem {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{35}{]}.
+
+The DAE-GP algorithm first described by {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{36}{]} trains a DAE-LSTMs for each generation \(g\) of the search from scratch.
+{[}\protect\hyperlink{ref-dae-gp_2022_symreg}{35}{]} and {[}\protect\hyperlink{ref-daegp_explore_exploit}{34}{]} suggests the incorporation of a pre-training strategy into the evolutionary search as a possible way of improving the overall performance of the DAE-GP algorithm.
 The key idea is to pre-train an initial DAE-LSTM on a large population of candidate solutions and to use the pre-trained parameters of this initial model in each generation of the search as starting parameters for the current generation DAE-LSTM.
@@
 
-\ldots{}
-
 \hypertarget{theoretical-foundations}{%
@@
 
-DAE-GP is an EDA-GP algorithm that uses DAE-LSTM networks as a probabilistic model to sample new offspring solutions {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{34}{]}.
+DAE-GP is an EDA-GP algorithm that uses DAE-LSTM networks as a probabilistic model to sample new offspring solutions {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{36}{]}.
 
@@
 Another important concept in EC is the distinction between genotypes and phenotypes of solutions, the genotype contains the information that is necessary to construct the phenotype, the outer appearance of a particular solution on which we measure the overall quality of solutions.
-The representation of a solution is therefore defined by the mapping of genotypes to phenotypes {[}\protect\hyperlink{ref-design_of_modern_heuristics}{22}{]}.
+The representation of a solution is therefore defined by the mapping of genotypes to phenotypes {[}\protect\hyperlink{ref-design_of_modern_heuristics}{23}{]}.
 Genetic operators, such as mutation or recombination are usually applied to the genotype of solutions.
@@
 
-GP follows the same basic evolutionary principle of EC but searches for more general, hierarchical computer programs of dynamically varying size and shape {[}\protect\hyperlink{ref-Koza1993GeneticP}{14}{]}.
-The computer programs that are at the center of the evolutionary search in GP are commonly represented by tree structures at the level of their phenotype {[}\protect\hyperlink{ref-design_of_modern_heuristics}{22}{]}.
-Since GP searches for high fitness computer programs that produce a desired output for some input, it can be applied to various different problem domains such as symbolic regression, automatic programming, or evolving game-playing strategies {[}\protect\hyperlink{ref-Koza1993GeneticP}{14}{]}.
+GP follows the same basic evolutionary principle of EC but searches for more general, hierarchical computer programs of dynamically varying size and shape {[}\protect\hyperlink{ref-Koza1993GeneticP}{15}{]}.
+The computer programs that are at the center of the evolutionary search in GP are commonly represented by tree structures at the level of their phenotype {[}\protect\hyperlink{ref-design_of_modern_heuristics}{23}{]}.
+Since GP searches for high fitness computer programs that produce a desired output for some input, it can be applied to various different problem domains such as symbolic regression, automatic programming, or evolving game-playing strategies {[}\protect\hyperlink{ref-Koza1993GeneticP}{15}{]}.
 An important quality of GP is the ability to search for solutions of variable length and structure.
 GP is an especially useful meta heuristic for problems where no a priori knowledge about the final form of good solutions is available.
-GPs ability to optimize solutions for their structure as well as for their parameters led to it being one of the most prevalent methods used for symbolic regression {[}\protect\hyperlink{ref-10.1007ux2f978-3-540-24621-3_22}{19}{]}.
+GPs ability to optimize solutions for their structure as well as for their parameters led to it being one of the most prevalent methods used for symbolic regression {[}\protect\hyperlink{ref-10.1007ux2f978-3-540-24621-3_22}{20}{]}.
 
@@
 
-The aim of Estimation of distribution algorithms (EDA) is to replace the standard variation operators used in GA by building probabilistic models that can capture complex dependencies between different decision variables of an optimization problem {[}\protect\hyperlink{ref-design_of_modern_heuristics}{22}{]}.
+The aim of Estimation of distribution algorithms (EDA) is to replace the standard variation operators used in GA by building probabilistic models that can capture complex dependencies between different decision variables of an optimization problem {[}\protect\hyperlink{ref-design_of_modern_heuristics}{23}{]}.
 EDAs use this probabilistic model to sample new offspring solutions inside an evolutionary search to replace crossover and/or mutation operators.
-
-\hypertarget{eda-gp}{%
-\subsubsection{EDA-GP}\label{eda-gp}}
-
-\ldots{} insert some smart stuff here\ldots{}
+Even though a majority of EDA algorithms are designed to work with fixed length string representations as used in classical GAs, EDA algorithms have been researched an successfully applied to the domain of GP in various different research streams {[}\protect\hyperlink{ref-kim_probMod_GP}{13}{]}.
 
@@
 
-One possible way of model building in EDA proposed by {[}\protect\hyperlink{ref-harmless_overfitting_eda}{21}{]} is to use denoising autoencoders (DAEs) as generative models to capture complex probability distributions of decision variables.
+One possible way of model building in EDA proposed by {[}\protect\hyperlink{ref-harmless_overfitting_eda}{22}{]} is to use denoising autoencoders (DAEs) as generative models to capture complex probability distributions of decision variables.
 DAE, a variation of the autoencoders (AE), are a widely used type of neural networks in the domain of unsupervised machine learning that maps \(n\) input variables to \(n\) output variables using a hidden representation.
@@
 
-DAE was first introduced by {[}\protect\hyperlink{ref-dae_orig2008}{28}{]} as an improved AE with the ability to learn new representations of data that is especially robust to partially corrupted input data.
+DAE was first introduced by {[}\protect\hyperlink{ref-dae_orig2008}{30}{]} as an improved AE with the ability to learn new representations of data that is especially robust to partially corrupted input data.
 DAE modifies the AE by using partially corrupted input data for the AE and training it to reconstruct the uncorrupted, original version of the input data.
 
-Since the hidden representation of DAE captures the dependency structure of the input data it can therefore also be used to generate new solutions in the context of GAs {[}\protect\hyperlink{ref-harmless_overfitting_eda}{21}{]}.
-
-DAE-GP builds upon the concept of using DAEs in EDAs described by {[}\protect\hyperlink{ref-harmless_overfitting_eda}{21}{]} and transfers the concept to the domain of GP.
-The mutation and crossover operators of standard GP are replaced by sampling new solutions from a probabilistic model that is build by training a DAEs long short-term memory (LSTM) network on a subset of high fitness solutions from the current population {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{34}{]}.
+Since the hidden representation of DAE captures the dependency structure of the input data it can therefore also be used to generate new solutions in the context of GAs {[}\protect\hyperlink{ref-harmless_overfitting_eda}{22}{]}.
+
+DAE-GP builds upon the concept of using DAEs in EDAs described by {[}\protect\hyperlink{ref-harmless_overfitting_eda}{22}{]} and transfers the concept to the domain of GP.
+The mutation and crossover operators of standard GP are replaced by sampling new solutions from a probabilistic model that is build by training a DAEs long short-term memory (LSTM) network on a subset of high fitness solutions from the current population {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{36}{]}.
 
 LSTMs are a variant of artificial neural networks first introduced by {[}\protect\hyperlink{ref-lstm_orig}{11}{]} that can store learned information over en extended time period while avoiding the problem of vanishing and/or exploding gradients.
-Since DAE-GP encodes candidate solutions as linear strings in prefix notation, the DAE in DAE-GP uses LSTMs for both encoding and decoding where the total amount of time steps \(T\) is equal to sum of the length of the input solution and the output solution {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{34}{]}.
+Since DAE-GP encodes candidate solutions as linear strings in prefix notation, the DAE in DAE-GP uses LSTMs for both encoding and decoding where the total amount of time steps \(T\) is equal to sum of the length of the input solution and the output solution {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{36}{]}.
 
@@
 
-The strategy of pre-training artificial neural networks is based on the idea of transfer learning that aims at retaining previously learned knowledge from one task to re-use it for another task {[}\protect\hyperlink{ref-HAN2021225}{7}{]} {[}\protect\hyperlink{ref-survey_transfer_learning}{18}{]}.
+The strategy of pre-training artificial neural networks is based on the idea of transfer learning that aims at retaining previously learned knowledge from one task to re-use it for another task {[}\protect\hyperlink{ref-HAN2021225}{7}{]} {[}\protect\hyperlink{ref-survey_transfer_learning}{19}{]}.
 Transfer learning traditionally follows a two phase approach:
@@
 
-where source task and target task are usually similar but may differ in their feature space and the distribution of training data {[}\protect\hyperlink{ref-survey_transfer_learning}{18}{]}.
+where source task and target task are usually similar but may differ in their feature space and the distribution of training data {[}\protect\hyperlink{ref-survey_transfer_learning}{19}{]}.
 
 Some of the main motivations for using pre-training is the ability to reduce the need for large amounts of training data which can often be unavailable or too expensive to collect as well as an improved performance by either reducing the computational effort that is necessary to train a model to solve a specific task or by improving the models generalization ability.
-
-\ldots describe different pre-training strategies, e.g.~few shot, perpetual, re-using\ldots{}
-
--- describe unsupervised vs supervised pre-training\ldots{}
 
@@
 Instead of initializing and optimizing each model from scratch as done in previous work (e.g.
-{[}\protect\hyperlink{ref-dae-gp_2020_rtree}{34}{]} {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{33}{]}), a separate DAE-LSTM network \(\hat{M}\) will be trained on a large population of randomly initialized solutions.\\
+{[}\protect\hyperlink{ref-dae-gp_2020_rtree}{36}{]} {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{35}{]}), a separate DAE-LSTM network \(\hat{M}\) will be trained on a large population of randomly initialized solutions.\\
 
@@
 Another motivation for introducing pre-training into DAE-GP is the prevalence of sampling error for small GP populations sizes.
-{[}\protect\hyperlink{ref-sampling_err_gp}{23}{]} finds that sampling error, non-systematic errors that are caused by observing only a small subset of the statistical population, is a severe problem in the domain of GP since the initial population may not be a representative sample of all possible solutions.
-{[}\protect\hyperlink{ref-sampling_err_gp}{23}{]} also introduces a method for calculating optimal population sizes to minimize the presence of sampling error.
+{[}\protect\hyperlink{ref-sampling_err_gp}{24}{]} finds that sampling error, non-systematic errors that are caused by observing only a small subset of the statistical population, is a severe problem in the domain of GP since the initial population may not be a representative sample of all possible solutions.
+{[}\protect\hyperlink{ref-sampling_err_gp}{24}{]} also introduces a method for calculating optimal population sizes to minimize the presence of sampling error.
 Since DAE-LSTM of DAE-GP learns the properties of it's training population and reuses the acquired knowledge in the sampling procedure, using the parameters obtained from training \(\hat{M}\) on a sufficiently large training population might reduce the need for large population sizes in the following generations of the search if \(\hat{M}\) already implicitly captured the properties of a representative sample of solutions.
@@
 
-The DAE-GP algorithm, first described by {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{34}{]}, is summarized as a flowchart in figure \ref{fig:daegp-flowchart}.
+The DAE-GP algorithm, first described by {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{36}{]}, is summarized as a flowchart in figure \ref{fig:daegp-flowchart}.
 After setting the generation counter \(g\) to \(0\) and creating the initial population of solutions \(P_0\), DAE-GP enters the main loop that checks if a termination criteria is satisfied (i.e.~maximum number of generations or fitness evaluations is reached).
@@
 One difficulty in the implementation of a pre-training strategy into DAE-GP has been the determination of the number of hidden neurons in side the DAE-LSTMs hidden layers.
-The original description of DAE-GP {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{34}{]} uses a strategy where the number of hidden neurons for each hidden layer is dynamically set per generation to the maximum individual size inside the current population.
+The original description of DAE-GP {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{36}{]} uses a strategy where the number of hidden neurons for each hidden layer is dynamically set per generation to the maximum individual size inside the current population.
 For a pre-training implementation, this strategy can not be easily adapted since it leads to a changing number of neurons at each generation resulting in different dimensions of the DAE-LSTM.
@@
 Train/Test Split & 50:50 \\
-Maximum Tree Depth & 17 {[}\protect\hyperlink{ref-Koza1993GeneticP}{14}{]} \\
-Initialization Algorithm & Ramped Half and Half {[}\protect\hyperlink{ref-Koza1993GeneticP}{14}{]} \\
+Maximum Tree Depth & 17 {[}\protect\hyperlink{ref-Koza1993GeneticP}{15}{]} \\
+Initialization Algorithm & Ramped Half and Half {[}\protect\hyperlink{ref-Koza1993GeneticP}{15}{]} \\
 Mix/Max Initialization-Depth & 2/6 \\
 Selection Operator & Binary Tournament Selection \\
-Function Set & \{+ , - , * , AQ {[}\protect\hyperlink{ref-AQ}{17}{]}\} \\
+Function Set & \{+ , - , * , AQ {[}\protect\hyperlink{ref-AQ}{18}{]}\} \\
 Ephemeral Constants & {[}-5,..,5{]} \\
@@
 \endlastfoot
-Stopping Condition & Convergence {[}\protect\hyperlink{ref-daegp_explore_exploit}{32}{]} \\
+Stopping Condition & Convergence {[}\protect\hyperlink{ref-daegp_explore_exploit}{34}{]} \\
 Reconstruction Error Metric & Multiclass Cross Entropy \\
-Gradient Descent Algorithm & Adam Optimization {[}\protect\hyperlink{ref-Kingma2014AdamAM}{13}{]} \\
+Gradient Descent Algorithm & Adam Optimization {[}\protect\hyperlink{ref-Kingma2014AdamAM}{14}{]} \\
 Learning Rate & 0.001 \\
@@
 Hidden Neurons & 150 \\
-Corruption Operator & Levenshtein Edit {[}\protect\hyperlink{ref-daegp_explore_exploit}{32}{]} \\
+Corruption Operator & Levenshtein Edit {[}\protect\hyperlink{ref-daegp_explore_exploit}{34}{]} \\
 Denoising Edit Probability (Model Building) & 0.05 \\
@@
 
-The framework that was used to run all experiments of this thesis was provided by the supervisor of this thesis David Wittenberg and uses the python programming language {[}\protect\hyperlink{ref-python3}{26}{]} in conjunction with the baseline libraries \texttt{Keras} {[}\protect\hyperlink{ref-chollet2015keras}{2}{]} for deep neural networks and \texttt{deap} {[}\protect\hyperlink{ref-DEAP_JMLR2012}{6}{]} for evolutionary computation.
-For data visualizations I used the python package \texttt{matplotlib} {[}\protect\hyperlink{ref-matplotlib}{12}{]}, statistical analysis and evaluations were conducted using primarily the data science libraries \texttt{scipy} {[}\protect\hyperlink{ref-scipy}{29}{]}, \texttt{numpy} {[}\protect\hyperlink{ref-numpy}{8}{]}, \texttt{pandas} {[}\protect\hyperlink{ref-pandas}{24}{]}, and \texttt{IPython} {[}\protect\hyperlink{ref-ipython}{20}{]}.
+The framework that was used to run all experiments of this thesis was provided by the supervisor of this thesis David Wittenberg and uses the python programming language {[}\protect\hyperlink{ref-python3}{28}{]} in conjunction with the baseline libraries \texttt{Keras} {[}\protect\hyperlink{ref-chollet2015keras}{2}{]} for deep neural networks and \texttt{deap} {[}\protect\hyperlink{ref-DEAP_JMLR2012}{6}{]} for evolutionary computation.
+For data visualizations I used the python package \texttt{matplotlib} {[}\protect\hyperlink{ref-matplotlib}{12}{]}, statistical analysis and evaluations were conducted using primarily the data science libraries \texttt{scipy} {[}\protect\hyperlink{ref-scipy}{31}{]}, \texttt{numpy} {[}\protect\hyperlink{ref-numpy}{8}{]}, \texttt{pandas} {[}\protect\hyperlink{ref-pandas}{26}{]}, and \texttt{IPython} {[}\protect\hyperlink{ref-ipython}{21}{]}.
 
@@
 To test pre-training in DAE-GP this thesis focuses on the domain of real-world symbolic regression problems.
-Symbolic Regression problems have been one of the first GP applications {[}\protect\hyperlink{ref-Koza1993GeneticP}{14}{]} and are an actively studied and highly relevant research area.
-The goal in symbolic regression is to find a mathematical model for a given set of data points {[}\protect\hyperlink{ref-10.1007ux2f978-3-540-24621-3_22}{19}{]}, in real-world symbolic regression these data points are sourced from real-world observations which in contrast to synthetic symbolic regression problem are more likely to contain random noise and bias.
+Symbolic Regression problems have been one of the first GP applications {[}\protect\hyperlink{ref-Koza1993GeneticP}{15}{]} and are an actively studied and highly relevant research area.
+The goal in symbolic regression is to find a mathematical model for a given set of data points {[}\protect\hyperlink{ref-10.1007ux2f978-3-540-24621-3_22}{20}{]}, in real-world symbolic regression these data points are sourced from real-world observations which in contrast to synthetic symbolic regression problem are more likely to contain random noise and bias.
 Another important challenge in solving real-world symbolic regression problems is the ability for a given model to generalize, we want the final model to show high accuracy in predicting outcomes for previously unseen cases.
@@
 The objective of the airfoil problem is to find a function that accurately predicts the output variable \(y\) by taking in a subset of the input variables \(x1,x2,x3,x4,x5\).
-The function set used by all DAE-GP variations for symbolic regression is summarized in tables \ref{tab:baseParams-GP} and , the terminal set consists of the the input variables \(x1,x2,x3,x4,x5\) and the ephemeral integer constants listed in table \ref{tab:baseParams-GP} {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{33}{]}.
+The function set used by all DAE-GP variations for symbolic regression is summarized in tables \ref{tab:baseParams-GP} and , the terminal set consists of the the input variables \(x1,x2,x3,x4,x5\) and the ephemeral integer constants listed in table \ref{tab:baseParams-GP} {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{35}{]}.
 
@@
 For each configuration I tested traditional DAE-GP as well as a pre-trained DAE-GP resulting in 12 total sub experiments that were each based on 10 individual runs (total number of runs=\(120\)).
-To avoid creating biased results through the presence of sampling error inside the pre-training population (see {[}\protect\hyperlink{ref-sampling_err_gp}{23}{]}), the population size for the pre-training phase is chosen very high with the size of \(\hat{P}\) =10000 where 50\% of \(\hat{P}\) is used for the training population \(\hat{P}_{train}\) and the remaining 50\% are used for the hold-out validation population \(\hat{P}_{test}\).
+To avoid creating biased results through the presence of sampling error inside the pre-training population (see {[}\protect\hyperlink{ref-sampling_err_gp}{24}{]}), the population size for the pre-training phase is chosen very high with the size of \(\hat{P}\) =10000 where 50\% of \(\hat{P}\) is used for the training population \(\hat{P}_{train}\) and the remaining 50\% are used for the hold-out validation population \(\hat{P}_{test}\).
 The training of DAE-LSTM uses a fixed number of 1000 epochs until termination and uses Adam optimization for gradient descent.
@@
 In general I expect that with a growing number of trainable parameters (either by adding more hidden layers or more hidden neurons per layer) the DAE-LSTM will be more prone to overfit to the training population \(P_{train}\) resulting in a small reconstruction error for the training population and a large one for the validation population \(P_{test}\).
-The reason for this effect is that a larger network trained over an extended period of time (without the use of strategies like early stopping), has much more potential to learn noise from the training dataset than a smaller network, which is more likely to result in worsening performance for previously unseen cases {[}\protect\hyperlink{ref-weigend1994overfitting}{30}{]}.
+The reason for this effect is that a larger network trained over an extended period of time (without the use of strategies like early stopping), has much more potential to learn noise from the training dataset than a smaller network, which is more likely to result in worsening performance for previously unseen cases {[}\protect\hyperlink{ref-weigend1994overfitting}{32}{]}.
 
@@
 
-Since previous work by {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{33}{]} demonstrated that DAE-GP, for a given number of 10.000 fitness evaluations, is able to produce solutions that are much smaller in tree size than regular GP, I also studied the influence of pre-training on the size of solutions found by DAE-GP.
+Since previous work by {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{35}{]} demonstrated that DAE-GP, for a given number of 10.000 fitness evaluations, is able to produce solutions that are much smaller in tree size than regular GP, I also studied the influence of pre-training on the size of solutions found by DAE-GP.
 Figure \ref{fig:airfoil-size-combined-30gens} shows the median of both the average size of individuals inside the population as well as the size of the currently best performing solution inside the population.
@@
 Another interesting metric to examine in population-based optimization algorithms is the diversity of the current generations population.
-Since individual solutions in DAE-GP are computer programs in the form of parse trees, I selected the normalized Levenshtein edit distance {[}\protect\hyperlink{ref-NormLevDistance}{15}{]} as my primary metric for tracking population diversity.
+Since individual solutions in DAE-GP are computer programs in the form of parse trees, I selected the normalized Levenshtein edit distance {[}\protect\hyperlink{ref-NormLevDistance}{16}{]} as my primary metric for tracking population diversity.
 Figure \ref{fig:airfoil-popDiversity} shows the median population diversity for both DAE-GP algorithms and regular GP as a benchmark.
@@
 To ensure a fair comparison between pre-trained and regular DAE-GP, the prior experiments used a static number of hidden neurons for each hidden Layer inside the DAE-LSTM networks.
-For comparison, figure \ref{fig:airfoil-maxIndSize-params} shows the total number of trainable parameters for each generation's DAE-LSTM of another experiment where regular DAE-GP was tested with dynamic adjustment of the number of hidden neurons as described in {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{34}{]} and {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{33}{]}.
+For comparison, figure \ref{fig:airfoil-maxIndSize-params} shows the total number of trainable parameters for each generation's DAE-LSTM of another experiment where regular DAE-GP was tested with dynamic adjustment of the number of hidden neurons as described in {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{36}{]} and {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{35}{]}.
 Again, the experiments was based on \(10\) individual runs and otherwise used the same hyperparameters as listed in tables \ref{tab:baseParams-GP} and \ref{tab:baseParams-DAEGP}.
@@
 
-Since previous research on DAE-GP used a single hidden layers for the airfoil dataset {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{33}{]}, I repeated the experiment with a single hidden layer.
+Since previous research on DAE-GP used a single hidden layers for the airfoil dataset {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{35}{]}, I repeated the experiment with a single hidden layer.
 Based on the findings of {[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{5}{]}, I expected that pre-training DAE-GP would lead to a decrease in overall performance.
@@
 
-The pre-training implementation in previous experiments used ramped half and half initialization {[}\protect\hyperlink{ref-Koza1993GeneticP}{14}{]} to create \(\hat{P}\), this leads to 50\% of the population consisting of full grown trees.
+The pre-training implementation in previous experiments used ramped half and half initialization {[}\protect\hyperlink{ref-Koza1993GeneticP}{15}{]} to create \(\hat{P}\), this leads to 50\% of the population consisting of full grown trees.
 An alternative method of initializing \(\hat{P}\) is to use the grow initialization method where solutions are not guaranteed to be full trees.
@@
 Boston\_Housing & 506 & 13 & {[}\protect\hyperlink{ref-BOSTON_HOUSING_DATASET}{9}{]} \\
-Energy\_Cooling & 768 & 8 & {[}\protect\hyperlink{ref-ENERGY_DATASET}{25}{]} \\
-Concrete & 1030 & 8 & {[}\protect\hyperlink{ref-CONCRETE_DATASET}{35}{]} \\
+Energy\_Cooling & 768 & 8 & {[}\protect\hyperlink{ref-ENERGY_DATASET}{27}{]} \\
+Concrete & 1030 & 8 & {[}\protect\hyperlink{ref-CONCRETE_DATASET}{37}{]} \\
 \end{longtable}
 
-To test for statistical differences of the results achieved by either pre-trained or regular DAE-GP, I selected the nonparametric Mann-Whitney-U Test ({[}\protect\hyperlink{ref-mwu1}{31}{]}, {[}\protect\hyperlink{ref-mwu2}{16}{]}) to test for differences in the distribution underlying the samples sourced from my experiments.
+To test for statistical differences of the results achieved by either pre-trained or regular DAE-GP, I selected the nonparametric Mann-Whitney-U Test ({[}\protect\hyperlink{ref-mwu1}{33}{]}, {[}\protect\hyperlink{ref-mwu2}{17}{]}) to test for differences in the distribution underlying the samples sourced from my experiments.
 Additionally, I included the effect strength of pre-training as measured by Cliffs Delta {[}\protect\hyperlink{ref-Cliff1993DominanceSO}{3}{]} into the statistical analysis.
@@
 
-As expected after reviewing figure \ref{fig:symReg-Fitness}, none of the differences in median best fitness are statistically significant and pre-training has only a very small, negligible effect strength for Cliff's Delta ranging between -0.04 and 0.17 {[}\protect\hyperlink{ref-Vargha2000ACA}{27}{]}.
+As expected after reviewing figure \ref{fig:symReg-Fitness}, none of the differences in median best fitness are statistically significant and pre-training has only a very small, negligible effect strength for Cliff's Delta ranging between -0.04 and 0.17 {[}\protect\hyperlink{ref-Vargha2000ACA}{29}{]}.
 
@@
 Figure \ref{fig:symReg-Size} shows the median size of the best solution per generation aggregated over all benchmark problems for symbolic regression.
-Similiar to the previous results regarding the fitness of the best found solution, no significant differences in the size of the best found solution are visible between both DAE-GP variants.
+Similar to the previous results regarding the fitness of the best found solution, no significant differences in the size of the best found solution are visible between both DAE-GP variants.
 
@@
 Table \ref{tab:full-run-realWorldSymReg-best-solution-size} shows the median size of the best solution after 30 generations and results for computing the statistical tests.
-For Airfoil, Boston Housing and Energy Cooling datasets, the median size of the best soultion are equal, only for the concrete dataset a smaller median size for pre-trained DAE-GP was achieved (pre-trained DAE-GP also showed a higher median best fitness for the concrete problem).
-The results of the Mann-Whitney-U Test again show, that differences in the underlying distributions are not statistically significant which also shows in low effect strenghts of using pre-training on the size of the best found solutions (Cliff's Delta ranges from-0.02 to 0.20).
+For Airfoil, Boston Housing and Energy Cooling data-sets, the median size of the best solution are equal, only for the concrete dataset a smaller median size for pre-trained DAE-GP was achieved (pre-trained DAE-GP also showed a higher median best fitness for the concrete problem).
+The results of the Mann-Whitney-U Test again show, that differences in the underlying distributions are not statistically significant which also shows in low effect strengths of using pre-training on the size of the best found solutions (Cliff's Delta ranges from-0.02 to 0.20).
 
@@
 
-The experiments conducted in the scope of this master thesis did find that the implemented pre-training strategy resulted in an increased population diversity measured by the normalized levenshtein edit distance.
-Increasing the population diversity can be an important mechanism in controlling the search behavior of population based meta heuristics and is especially useful for increasing the exploration of the search space of possible solutions !Q
-UELLE!.
-
-Another significant benefit of using pre-training in DAE-GP is that it is highly efficient in decreasing the number of training epochs that are necessary to train each generations DAE-LSTM model.
-Even though pre-training increased the median total run-time for the tested benchmark problem in combination with the chosen DAE-LSTM dimensions, reducing the number of training epochs might be a possible way to lower the run time of DAE-GP if further performance optimizations, especially during model sampling, can be achieved.
+The experiments conducted in the scope of this master thesis did find that the implemented pre-training strategy resulted in a highly statistically significant increase in population diversity measured by the normalized levenshtein edit distance (see subsection \protect\hyperlink{influence-on-population-diversity}{Influence on Population Diversity}).
+Increasing the population diversity can be an important mechanism in controlling the search behavior of evolutionary algorithms and is especially important for the global exploration of the search space and the avoidance of premature convergence of the search {[}\protect\hyperlink{ref-Sudholt2020}{25}{]}.
+
+Another significant benefit of using pre-training in DAE-GP is that it is highly efficient in decreasing the number of training epochs that are necessary to train each generations DAE-LSTM model (see subsection \protect\hyperlink{influence-on-population-diversity}{Influence on Population Diversity}).
+Even though pre-training increased the median total run-time for the tested benchmark problem in combination with the chosen DAE-LSTM dimensions, reducing the number of training epochs might still be a possible way to lower the run time of DAE-GP if further performance optimizations, especially during model sampling, can be achieved.
 
@@
 For a single hidden layer, pre-training did show a negative impact on the median final fitness of solutions found as it was suggested by {[}\protect\hyperlink{ref-pmlr-v9-erhan10a}{4}{]}.
-Since current research on DAE-GP for real-world symbolic regression relies on a single hidden layer (see {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{33}{]}), pre-training does come with additional computational expenses for using deeper DAE-LSTMs.
+Since current research on DAE-GP for real-world symbolic regression relies on a single hidden layer (see {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{35}{]}), pre-training does come with additional computational expenses for using deeper DAE-LSTMs.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-kim_probMod_GP}{}}%
+\CSLLeftMargin{{[}13{]} }%
+\CSLRightInline{Kangil Kim, Yin Shan, Nguyen Hoai, and Robert McKay.
+2014.
+Probabilistic model building in genetic programming: A critical review.
+\emph{Genetic Programming and Evolvable Machines} 15, (June 2014).
+DOI:https://doi.org/\href{https://doi.org/10.1007/s10710-013-9205-x}{10.1007/s10710-013-9205-x}}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Kingma2014AdamAM}{}}%
-\CSLLeftMargin{{[}13{]} }%
+\CSLLeftMargin{{[}14{]} }%
 \CSLRightInline{Diederik P. Kingma and Jimmy Ba. 2014. Adam: A method for stochastic optimization. \emph{CoRR} abs/1412.6980, (2014).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Koza1993GeneticP}{}}%
-\CSLLeftMargin{{[}14{]} }%
+\CSLLeftMargin{{[}15{]} }%
 \CSLRightInline{John R. Koza. 1993. Genetic programming - on the programming of computers by means of natural selection. In \emph{Complex adaptive systems}.}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-NormLevDistance}{}}%
-\CSLLeftMargin{{[}15{]} }%
+\CSLLeftMargin{{[}16{]} }%
 \CSLRightInline{Joseph B.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-mwu2}{}}%
-\CSLLeftMargin{{[}16{]} }%
+\CSLLeftMargin{{[}17{]} }%
 \CSLRightInline{Henry B.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-AQ}{}}%
-\CSLLeftMargin{{[}17{]} }%
+\CSLLeftMargin{{[}18{]} }%
 \CSLRightInline{Ji Ni, Russ H.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-survey_transfer_learning}{}}%
-\CSLLeftMargin{{[}18{]} }%
+\CSLLeftMargin{{[}19{]} }%
 \CSLRightInline{Sinno Jialin Pan and Qiang Yang.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-10.1007ux2f978-3-540-24621-3_22}{}}%
-\CSLLeftMargin{{[}19{]} }%
+\CSLLeftMargin{{[}20{]} }%
 \CSLRightInline{Grégory Paris, Denis Robilliard, and Cyril Fonlupt.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ipython}{}}%
-\CSLLeftMargin{{[}20{]} }%
+\CSLLeftMargin{{[}21{]} }%
 \CSLRightInline{Fernando Pérez and Brian E.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-harmless_overfitting_eda}{}}%
-\CSLLeftMargin{{[}21{]} }%
+\CSLLeftMargin{{[}22{]} }%
 \CSLRightInline{Malte Probst and Franz Rothlauf.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-design_of_modern_heuristics}{}}%
-\CSLLeftMargin{{[}22{]} }%
+\CSLLeftMargin{{[}23{]} }%
 \CSLRightInline{Franz Rothlauf. 2011. \emph{Design of modern heuristics: Principles and application}. DOI:https://doi.org/\href{https://doi.org/10.1007/978-3-540-72962-4}{10.1007/978-3-540-72962-4}}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-sampling_err_gp}{}}%
-\CSLLeftMargin{{[}23{]} }%
+\CSLLeftMargin{{[}24{]} }%
 \CSLRightInline{Dirk Schweim, David Wittenberg, and Franz Rothlauf.
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Sudholt2020}{}}%
+\CSLLeftMargin{{[}25{]} }%
+\CSLRightInline{Dirk Sudholt.
+2020.
+The benefits of population diversity in evolutionary algorithms: A survey of rigorous runtime analyses.
+In \emph{Theory of evolutionary computation: Recent developments in discrete optimization}, Benjamin Doerr and Frank Neumann (eds.).
+Springer International Publishing, Cham, 359--404.
+DOI:https://doi.org/\href{https://doi.org/10.1007/978-3-030-29414-4_8}{10.1007/978-3-030-29414-4\_8}}
+
 \leavevmode\vadjust pre{\hypertarget{ref-pandas}{}}%
-\CSLLeftMargin{{[}24{]} }%
+\CSLLeftMargin{{[}26{]} }%
 \CSLRightInline{The pandas development team. 2020. Pandas-dev/pandas: pandas. DOI:https://doi.org/\href{https://doi.org/10.5281/zenodo.3509134}{10.5281/zenodo.3509134}}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ENERGY_DATASET}{}}%
-\CSLLeftMargin{{[}25{]} }%
+\CSLLeftMargin{{[}27{]} }%
 \CSLRightInline{Athanasios Tsanas and Angeliki Xifara.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-python3}{}}%
-\CSLLeftMargin{{[}26{]} }%
+\CSLLeftMargin{{[}28{]} }%
 \CSLRightInline{Guido Van Rossum and Fred L. Drake. 2009. \emph{Python 3 reference manual}. CreateSpace, Scotts Valley, CA.}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Vargha2000ACA}{}}%
-\CSLLeftMargin{{[}27{]} }%
+\CSLLeftMargin{{[}29{]} }%
 \CSLRightInline{András Vargha and Harold D Delaney.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-dae_orig2008}{}}%
-\CSLLeftMargin{{[}28{]} }%
+\CSLLeftMargin{{[}30{]} }%
 \CSLRightInline{Pascal Vincent, Hugo Larochelle, Y.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-scipy}{}}%
-\CSLLeftMargin{{[}29{]} }%
+\CSLLeftMargin{{[}31{]} }%
 \CSLRightInline{Pauli Virtanen, Ralf Gommers, Travis E.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-weigend1994overfitting}{}}%
-\CSLLeftMargin{{[}30{]} }%
+\CSLLeftMargin{{[}32{]} }%
 \CSLRightInline{Andreas Weigend. 1994. On overfitting and the effective number of hidden units. In \emph{Proceedings of the 1993 connectionist models summer school}, 335--342.}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-mwu1}{}}%
-\CSLLeftMargin{{[}31{]} }%
+\CSLLeftMargin{{[}33{]} }%
 \CSLRightInline{Frank. Wilcoxon. 1945. Individual comparisons by ranking methods. \emph{Biometrics} 1, (1945), 196--202.}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-daegp_explore_exploit}{}}%
-\CSLLeftMargin{{[}32{]} }%
+\CSLLeftMargin{{[}34{]} }%
 \CSLRightInline{David Wittenberg.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-dae-gp_2022_symreg}{}}%
-\CSLLeftMargin{{[}33{]} }%
+\CSLLeftMargin{{[}35{]} }%
 \CSLRightInline{David Wittenberg and Franz Rothlauf.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-dae-gp_2020_rtree}{}}%
-\CSLLeftMargin{{[}34{]} }%
+\CSLLeftMargin{{[}36{]} }%
 \CSLRightInline{David Wittenberg, Franz Rothlauf, and Dirk Schweim.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-CONCRETE_DATASET}{}}%
-\CSLLeftMargin{{[}35{]} }%
+\CSLLeftMargin{{[}37{]} }%
 \CSLRightInline{I-Cheng Yeh.

### 2023-02-09 paper.Rmd
Thu Feb  9 16:00:07 CET 2023: update-remote.sh

@@
 Also note that the results for the real world symbolic regression problems inside this chapter, except for the airfoil problem (see subsection [Influence on Run-Time]), have not been analyzed for their total run-time.
-This is due to the fact, that experiments were run on different computers architectures which would result in biased results.
+This is due to the fact, that experiments were run on different computer architectures which would result in biased results.
 

### 2023-02-10 paper.Rmd
autogit

@@
 The Experiments conducted did show that pre-training can drastically reduce the number of training epochs that are necessary at each generation of the DAE-GP search while also increasing the diversity of the population.
-Unfortunately pre-training did not show any improvements for both final fitness and the final size of solutions while largley increasing the total run-time for DAE-GP.
+Unfortunately pre-training did not show any improvements for both final fitness and the final size of solutions while largely increasing the total run-time for DAE-GP.
 
@@
 The basic idea is to run regular DAE-GP for the first generation without any interference, after completing the first generation, the pre-training population $\hat{P}$ is created by sampling the DAE-LSTM model $M_1$.
-The pre-training model $\hat{M}$ is then trained using $\hat{P}$, after finishing DAE-GP continues running and initializes each new DAE-LSTM model $M_g$ for $g\in{2,...,g_{max}}$ using the trainable parameters $\theta_{\hat{M}}$.
+The pre-training model $\hat{M}$ is then trained using $\hat{P}$, after finishing, DAE-GP continues running and initializes each new DAE-LSTM model $M_g$ for $g\in{2,...,g_{max}}$ using the trainable parameters $\theta_{\hat{M}}$.
 
-The pre-training implementation in previous experiments used ramped half and half initialization [@Koza1993GeneticP] to create $\hat{P}$, this leads to 50% of the population consisting of full grown trees.
-An alternative method of initializing $\hat{P}$ is to use the grow initialization method where solutions are not guaranteed to be full trees.
+The second alternative pre-training strategy is based on the choice of an alternative initialization algorithm to create $\hat{P}$.
+The pre-training implementation in previous experiments used `ramped half and half` initialization [@Koza1993GeneticP] to create $\hat{P}$.
+Using ramped half and half leads to 50% of the population being initialized by the `full` method, where terminals are only inserted at the maximum allowed depth of a tree resulting in full trees.
+An alternative method of initializing $\hat{P}$ is to use the `grow` initialization method where solutions are not guaranteed to be full trees since every leaf above the maximum depth of a tree can also be a terminal expression.
 This different structural composition of solutions inside the pre-training population might be another way to improve the pre-training strategy by training on a population that more closely resembles the structural properties of further generations population where a lower percentage of full grown trees is to be expected.
@@
 Figure \@ref(fig:altStrategies-Fitness) shows the median best fitness by generations for both alternative pre-training strategies as well as the results for standard DAE-GP and the standard Pre-Training implementation used in earlier experiments.
-Unfortunately, both alternative strategies do show a worse performance in fitness than both DAE-GP and standard pre-trained DAE-GP resulting in lower fitness.
+Unfortunately, both alternative strategies demonstrate an inferior performance in fitness compared to both regular DAE-GP and standard pre-trained DAE-GP resulting in lower fitness.
 
@@
 
-Figure \@ref(fig:altStrategies-FitnessDistr) shows the distribution of best fitness achieved on the test set after running for 30 full generations.
-Both alternative second generation Pre-Training as well as grow-initialized Pre-Training show a much higher median RMSE (18.88/17.21) compared to standard pre-training (8.55) and regular DAE-GP (6.99).
+Figure \@ref(fig:altStrategies-FitnessDistr) shows the distribution of best fitness scores achieved on the test set after running for 30 full generations.
+Both alternatives, second generation Pre-Training as well as grow-initialized Pre-Training, show a much higher median RMSE (18.88/17.21) compared to standard pre-training (8.55) and regular DAE-GP (6.99).
 Also both approaches do not demonstrate any benefit on the dispersion of solutions found resulting in similar standard deviations if compared to standard pre-training.

### 2023-02-10 paper.tex
Thu Feb  9 16:00:07 CET 2023: update-remote.sh
autogit

@@
 \author{}
-\date{\vspace{-2.5em}Date of Submission: 2023-02-09}
+\date{\vspace{-2.5em}Date of Submission: 2023-02-10}
 
@@
 The basic idea is to run regular DAE-GP for the first generation without any interference, after completing the first generation, the pre-training population \(\hat{P}\) is created by sampling the DAE-LSTM model \(M_1\).
-The pre-training model \(\hat{M}\) is then trained using \(\hat{P}\), after finishing DAE-GP continues running and initializes each new DAE-LSTM model \(M_g\) for \(g\in{2,...,g_{max}}\) using the trainable parameters \(\theta_{\hat{M}}\).
-
-The pre-training implementation in previous experiments used ramped half and half initialization {[}\protect\hyperlink{ref-Koza1993GeneticP}{15}{]} to create \(\hat{P}\), this leads to 50\% of the population consisting of full grown trees.
-An alternative method of initializing \(\hat{P}\) is to use the grow initialization method where solutions are not guaranteed to be full trees.
+The pre-training model \(\hat{M}\) is then trained using \(\hat{P}\), after finishing, DAE-GP continues running and initializes each new DAE-LSTM model \(M_g\) for \(g\in{2,...,g_{max}}\) using the trainable parameters \(\theta_{\hat{M}}\).
+
+The second alternative pre-training strategy is based on the choice of an alternative initialization algorithm to create \(\hat{P}\).
+The pre-training implementation in previous experiments used \texttt{ramped\ half\ and\ half} initialization {[}\protect\hyperlink{ref-Koza1993GeneticP}{15}{]} to create \(\hat{P}\).
+Using ramped half and half leads to 50\% of the population being initialized by the \texttt{full} method, where terminals are only inserted at the maximum allowed depth of a tree resulting in full trees.
+An alternative method of initializing \(\hat{P}\) is to use the \texttt{grow} initialization method where solutions are not guaranteed to be full trees since every leaf above the maximum depth of a tree can also be a terminal expression.
 This different structural composition of solutions inside the pre-training population might be another way to improve the pre-training strategy by training on a population that more closely resembles the structural properties of further generations population where a lower percentage of full grown trees is to be expected.
@@
 Figure \ref{fig:altStrategies-Fitness} shows the median best fitness by generations for both alternative pre-training strategies as well as the results for standard DAE-GP and the standard Pre-Training implementation used in earlier experiments.
-Unfortunately, both alternative strategies do show a worse performance in fitness than both DAE-GP and standard pre-trained DAE-GP resulting in lower fitness.
+Unfortunately, both alternative strategies demonstrate an inferior performance in fitness compared to both regular DAE-GP and standard pre-trained DAE-GP resulting in lower fitness.
 
@@
 
-Figure \ref{fig:altStrategies-FitnessDistr} shows the distribution of best fitness achieved on the test set after running for 30 full generations.
-Both alternative second generation Pre-Training as well as grow-initialized Pre-Training show a much higher median RMSE (18.88/17.21) compared to standard pre-training (8.55) and regular DAE-GP (6.99).
+Figure \ref{fig:altStrategies-FitnessDistr} shows the distribution of best fitness scores achieved on the test set after running for 30 full generations.
+Both alternatives, second generation Pre-Training as well as grow-initialized Pre-Training, show a much higher median RMSE (18.88/17.21) compared to standard pre-training (8.55) and regular DAE-GP (6.99).
 Also both approaches do not demonstrate any benefit on the dispersion of solutions found resulting in similar standard deviations if compared to standard pre-training.
@@
 Also note that the results for the real world symbolic regression problems inside this chapter, except for the airfoil problem (see subsection \protect\hyperlink{influence-on-run-time}{Influence on Run-Time}), have not been analyzed for their total run-time.
-This is due to the fact, that experiments were run on different computers architectures which would result in biased results.
+This is due to the fact, that experiments were run on different computer architectures which would result in biased results.
 

### 2023-02-10 paper.Rmd
autogit

@@
 
-As expected after reviewing figure \@ref(fig:symReg-Fitness), none of the differences in median best fitness are statistically significant and pre-training has only a very small, negligible effect strength for Cliff's Delta ranging between `r min(df_symreg_fitness$Cliffs_Delta)` and `r max(df_symreg_fitness$Cliffs_Delta)` [@Vargha2000ACA].
+As expected after reviewing figure \@ref(fig:symReg-Fitness), none of the differences in median best fitness are statistically significant and pre-training has only a very small, negligible effect strength for Cliff's Delta ranging between `r min(as.numeric(df_symreg_fitness$Cliffs_Delta))` and `r max(as.numeric(df_symreg_fitness$Cliffs_Delta))` [@Vargha2000ACA].
 
@@
 For Airfoil, Boston Housing and Energy Cooling data-sets, the median size of the best solution are equal, only for the concrete dataset a smaller median size for pre-trained DAE-GP was achieved (pre-trained DAE-GP also showed a higher median best fitness for the concrete problem).
-The results of the Mann-Whitney-U Test again show, that differences in the underlying distributions are not statistically significant which also shows in low effect strengths of using pre-training on the size of the best found solutions (Cliff's Delta ranges from`r min(df_size$Cliffs_Delta)` to `r max(df_size$Cliffs_Delta)`).
+The results of the Mann-Whitney-U Test again show, that differences in the underlying distributions are not statistically significant which also shows in low effect strengths of using pre-training on the size of the best found solutions (Cliff's Delta ranges from`r min(as.numeric(df_size$Cliffs_Delta))` to `r max(as.numeric(df_size$Cliffs_Delta))`).
 
@@
 For all sub experiments, computing the Mann-Whitney-U Test returns P-Values that are smaller than $0.01$ which indicates a highly statistically significant difference in the underlying distributions.
-Looking at the effect strength of pre-training, it shows that pre-training does have a strong positive effect (Cliff's Delta ranges from `r min(df_popDiv$Cliffs_Delta)` to `r max(df_popDiv$Cliffs_Delta)`) on population diversity inside the population.
+Looking at the effect strength of pre-training, it shows that pre-training does have a strong positive effect (Cliff's Delta ranges from `r min(as.numeric(df_popDiv$Cliffs_Delta))` to `r max(as.numeric(df_popDiv$Cliffs_Delta))`) on population diversity inside the population.
 Pre-Training did increase the median normalized levenshtein distance over all generations and sub-experiments by a factor of `r round( median(df_popDiv$Pre.Trained_DAE.GP) / median(df_popDiv$DAE.GP) , 3)` from `r round(median(df_popDiv$DAE.GP) , 3)` to `r round(median(df_popDiv$Pre.Trained_DAE.GP) , 3)`.
@@
 Looking at the results of the statistical analysis summarized in table \@ref(tab:full-run-realWorldSymReg-epochsPerGen), computing the Mann-Whitney-U test does consistently return P-Values below $0.01$ for all experiments which proves the statistical significant differences in distributions.
-Analyzing the effect strength of pre-training on the number of training epochs shows values for Cliff's Delta in the range of `r min(df_epochs$Cliffs_Delta)` and `r max(df_epochs$Cliffs_Delta)` which indicates a very strong negative effect on the number of training epochs necessary at each generation of the search.
+Analyzing the effect strength of pre-training on the number of training epochs shows values for Cliff's Delta in the range of `r min(as.numeric(df_epochs$Cliffs_Delta))` and `r max(as.numeric(df_epochs$Cliffs_Delta))` which indicates a very strong negative effect on the number of training epochs necessary at each generation of the search.
 Computing the median number of training epochs over all generations for all benchmark problems shows a total decrease by a factor of `r round( (median(df_epochs$Pre.Trained_DAE.GP) / median(df_epochs$DAE.GP)) , 2)` from `r round(median(df_epochs$DAE.GP), 2)` epochs for regular DAE-GP to `r round(median(df_epochs$Pre.Trained_DAE.GP), 2)` epochs for pre-trained DAE-GP.

### 2023-02-10 paper.tex
autogit

@@
 Initialization Algorithm & Ramped Half and Half {[}\protect\hyperlink{ref-Koza1993GeneticP}{15}{]} \\
-Mix/Max Initialization-Depth & 2/6 \\
+Min/Max Initialization-Depth & 2/6 \\
 Selection Operator & Binary Tournament Selection \\
@@
 
-As expected after reviewing figure \ref{fig:symReg-Fitness}, none of the differences in median best fitness are statistically significant and pre-training has only a very small, negligible effect strength for Cliff's Delta ranging between -0.04 and 0.17 {[}\protect\hyperlink{ref-Vargha2000ACA}{29}{]}.
+As expected after reviewing figure \ref{fig:symReg-Fitness}, none of the differences in median best fitness are statistically significant and pre-training has only a very small, negligible effect strength for Cliff's Delta ranging between -0.14 and 0.17 {[}\protect\hyperlink{ref-Vargha2000ACA}{29}{]}.
 
@@
 For Airfoil, Boston Housing and Energy Cooling data-sets, the median size of the best solution are equal, only for the concrete dataset a smaller median size for pre-trained DAE-GP was achieved (pre-trained DAE-GP also showed a higher median best fitness for the concrete problem).
-The results of the Mann-Whitney-U Test again show, that differences in the underlying distributions are not statistically significant which also shows in low effect strengths of using pre-training on the size of the best found solutions (Cliff's Delta ranges from-0.02 to 0.20).
+The results of the Mann-Whitney-U Test again show, that differences in the underlying distributions are not statistically significant which also shows in low effect strengths of using pre-training on the size of the best found solutions (Cliff's Delta ranges from-0.33 to 0.2).
 
@@
 Looking at the results of the statistical analysis summarized in table \ref{tab:full-run-realWorldSymReg-epochsPerGen}, computing the Mann-Whitney-U test does consistently return P-Values below \(0.01\) for all experiments which proves the statistical significant differences in distributions.
-Analyzing the effect strength of pre-training on the number of training epochs shows values for Cliff's Delta in the range of -0.87 and -0.93 which indicates a very strong negative effect on the number of training epochs necessary at each generation of the search.
+Analyzing the effect strength of pre-training on the number of training epochs shows values for Cliff's Delta in the range of -0.93 and -0.87 which indicates a very strong negative effect on the number of training epochs necessary at each generation of the search.
 Computing the median number of training epochs over all generations for all benchmark problems shows a total decrease by a factor of 0.45 from 143.49 epochs for regular DAE-GP to 65.15 epochs for pre-trained DAE-GP.

### 2023-02-11 paper.Rmd
Sat Feb 11 12:00:05 CET 2023: update-remote.sh

@@
 This thesis investigates the possible benefits and downsides of using pre-training for the DAE-LSTM networks of DAE-GP for four real world symbolic regression problems.
-The Experiments conducted did show that pre-training can drastically reduce the number of training epochs that are necessary at each generation of the DAE-GP search while also increasing the diversity of the population.
+The experiments conducted did show that pre-training can drastically reduce the number of training epochs that are necessary at each generation of the DAE-GP search while also increasing the diversity of the population.
 Unfortunately pre-training did not show any improvements for both final fitness and the final size of solutions while largely increasing the total run-time for DAE-GP.
@@
 The DAE-GP algorithm first described by [@dae-gp_2020_rtree] trains a DAE-LSTMs for each generation $g$ of the search from scratch.
-@dae-gp_2022_symreg and @daegp_explore_exploit suggests the incorporation of a pre-training strategy into the evolutionary search as a possible way of improving the overall performance of the DAE-GP algorithm.
+@dae-gp_2022_symreg and @daegp_explore_exploit suggest the incorporation of a pre-training strategy into the evolutionary search as a possible way of improving the overall performance of the DAE-GP algorithm.
 The key idea is to pre-train an initial DAE-LSTM on a large population of candidate solutions and to use the pre-trained parameters of this initial model in each generation of the search as starting parameters for the current generation DAE-LSTM.
@@
 
-*This section describes the relevant concepts that are necessary for the understanding and classification of DAE-GP as well as the concept of pre-training in artificial neural networks*
+*This section describes the relevant concepts that are necessary for the understanding and classification of DAE-GP as well as the concept of pre-training in artificial neural networks.*
 
@@
 In EC, the quality of a solution is commonly measured as fitness and the time steps of the search are called generations.
-Another important concept in EC is the distinction between genotypes and phenotypes of solutions, the genotype contains the information that is necessary to construct the phenotype, the outer appearance of a particular solution on which we measure the overall quality of solutions.
+Another important concept in EC is the distinction between genotypes and phenotypes of solutions.
+The genotype contains the information that is necessary to construct the phenotype, the outer appearance of a particular solution on which we measure the overall quality of solutions.
 The representation of a solution is therefore defined by the mapping of genotypes to phenotypes [@design_of_modern_heuristics].
-Genetic operators, such as mutation or recombination are usually applied to the genotype of solutions.
+Genetic operators, such as mutation or recombination, are usually applied to the genotype of solutions.
 
@@
 An important quality of GP is the ability to search for solutions of variable length and structure.
-GP is an especially useful meta heuristic for problems where no a priori knowledge about the final form of good solutions is available.
-GPs ability to optimize solutions for their structure as well as for their parameters led to it being one of the most prevalent methods used for symbolic regression [@10.1007/978-3-540-24621-3_22].
+GP is an especially useful meta-heuristic for problems where no a priori knowledge about the final form of good solutions is available.
+GPs ability to optimize solutions for both their structure as well as for their parameters led to it being one of the most prevalent methods used for symbolic regression [@10.1007/978-3-540-24621-3_22].
 
@@
 
-The aim of Estimation of distribution algorithms (EDA) is to replace the standard variation operators used in GA by building probabilistic models that can capture complex dependencies between different decision variables of an optimization problem [@design_of_modern_heuristics].
+The aim of estimation of distribution algorithms (EDA) is to replace the standard variation operators used in EC by building probabilistic models that can capture complex dependencies between different decision variables of an optimization problem [@design_of_modern_heuristics].
 EDAs use this probabilistic model to sample new offspring solutions inside an evolutionary search to replace crossover and/or mutation operators.
-Even though a majority of EDA algorithms are designed to work with fixed length string representations as used in classical GAs, EDA algorithms have been researched an successfully applied to the domain of GP in various different research streams [@kim_probMod_GP].
+Even though a majority of EDA algorithms are designed to work with fixed length string representations as used in classical GAs, EDA algorithms have been researched and successfully applied to the domain of GP in various different research streams [@kim_probMod_GP].
 
@@
 
-One possible way of model building in EDA proposed by @harmless_overfitting_eda is to use denoising autoencoders (DAEs) as generative models to capture complex probability distributions of decision variables.
-DAE, a variation of the autoencoders (AE), are a widely used type of neural networks in the domain of unsupervised machine learning that maps $n$ input variables to $n$ output variables using a hidden representation.
+One possible way of model building in EDA proposed by @harmless_overfitting_eda is to use denoising autoencoders (DAE) as generative models to capture complex probability distributions of decision variables.
+DAE, a variation of the autoencoder (AE), is a widely used type of neural networks in the domain of unsupervised machine learning that maps $n$ input variables to $n$ output variables using a hidden representation.
 
@@
 -   Encoder $g(x)$: Encode input data to a smaller central layer $h$
--   Decoder $d(h)$: Decode and output the the encoded data back to its original dimension
+-   Decoder $d(h)$: Decode and output the encoded data back to its original dimension
 
-The AE is trained to reduce the reconstruction error between input and output data, after the training procedure is finished the network is able to reduce the dimensional of input data to a compressed representation[@ae_orig].
+The AE is trained to reduce the reconstruction error between input and output data.
+After the training procedure is finished the network is able to reduce the dimensionality of input data to a compressed representation[@ae_orig].
 
@@
 
-LSTMs are a variant of artificial neural networks first introduced by @lstm_orig that can store learned information over en extended time period while avoiding the problem of vanishing and/or exploding gradients.
-Since DAE-GP encodes candidate solutions as linear strings in prefix notation, the DAE in DAE-GP uses LSTMs for both encoding and decoding where the total amount of time steps $T$ is equal to sum of the length of the input solution and the output solution [@dae-gp_2020_rtree].
+LSTMs are a variant of artificial neural networks first introduced by @lstm_orig that can store learned information over en extended period of time while avoiding the problem of vanishing and/or exploding gradients.
+Since DAE-GP encodes candidate solutions as linear strings in prefix notation, the DAE in DAE-GP uses LSTMs for both encoding and decoding where the total amount of time steps $T$ is equal to the sum of the length of the input solution and the output solution [@dae-gp_2020_rtree].
 

### 2023-02-11 paper.Rmd
autogit

@@
 
-Some of the main motivations for using pre-training is the ability to reduce the need for large amounts of training data which can often be unavailable or too expensive to collect as well as an improved performance by either reducing the computational effort that is necessary to train a model to solve a specific task or by improving the models generalization ability.
+One of the main benefits of using pre-training is the ability to reduce the need for large amounts of training data which can often be unavailable or too expensive to collect.
+Additionally, pre-training can lead to an improved performance by either reducing the computational effort that is necessary to train a model to solve a specific task or by improving the models generalization ability.
 
@@
 
-The pre-training strategy used in this thesis is applied to the DAE-LSTM model $M_g$ that are used in each DAE-GP generation $g$ for $g\in{1,...,g_{max}}$.
+The pre-training strategy used in this thesis is applied to the DAE-LSTM models $M_g$ that are used in each DAE-GP generation $g$ for $g\in{1,...,g_{max}}$.
 Instead of initializing and optimizing each model from scratch as done in previous work (e.g.
@@
 
-The probably most obvious motivation for using a pre-training strategy in DAE-GP is based on the fact, that initializing $M_g$ with pre-trained weights $\theta_{\hat{M}}$ is likely to reduce the amount of training epochs that are necessary until the point of training error convergence is reached.
+The probably most obvious motivation for using a pre-training strategy in DAE-GP is based on the fact that initializing $M_g$ with pre-trained weights $\theta_{\hat{M}}$ is likely to reduce the amount of training epochs that are necessary until the point of training error convergence is reached.
 
 Early research on pre-training for DAE by @pmlr-v5-erhan09a comes to the conclusion that besides adding robustness to models the strategy also results in improved generalization and better performing models.
-The authors describe that unsupervised pre-training behaves like a regularizer on DAE networks, the mean test error and it's variance are reduced for DAE networks that are initialized from pre-trained weights if compared to the same architectures that use random initialization.
+The authors describe that unsupervised pre-training behaves like a regularizer on DAE networks.
+The mean test error and its variance are reduced for DAE networks that are initialized from pre-trained weights if compared to the same architectures that use random initialization.
 
@@
 [@sampling_err_gp] also introduces a method for calculating optimal population sizes to minimize the presence of sampling error.
-Since DAE-LSTM of DAE-GP learns the properties of it's training population and reuses the acquired knowledge in the sampling procedure, using the parameters obtained from training $\hat{M}$ on a sufficiently large training population might reduce the need for large population sizes in the following generations of the search if $\hat{M}$ already implicitly captured the properties of a representative sample of solutions.
+Since DAE-LSTM of DAE-GP learns the properties of its training population and reuses the acquired knowledge in the sampling procedure, using the parameters obtained from training $\hat{M}$ on a sufficiently large training population might reduce the need for large population sizes in the following generations of the search if $\hat{M}$ already implicitly captured the properties of a representative sample of solutions.
 If this mechanism can be applied successfully it would benefit the performance of DAE-GP by increasing the population diversity (resulting in better solution quality) as well as by reducing the need for large population sizes which require computational resources.
@@
 
-One difficulty in the implementation of a pre-training strategy into DAE-GP has been the determination of the number of hidden neurons in side the DAE-LSTMs hidden layers.
+One difficulty in the implementation of a pre-training strategy into DAE-GP has been the determination of the number of hidden neurons inside the DAE-LSTMs hidden layers.
 The original description of DAE-GP [@dae-gp_2020_rtree] uses a strategy where the number of hidden neurons for each hidden layer is dynamically set per generation to the maximum individual size inside the current population.
@@
 Symbolic Regression problems have been one of the first GP applications [@Koza1993GeneticP] and are an actively studied and highly relevant research area.
-The goal in symbolic regression is to find a mathematical model for a given set of data points [@10.1007/978-3-540-24621-3_22], in real-world symbolic regression these data points are sourced from real-world observations which in contrast to synthetic symbolic regression problem are more likely to contain random noise and bias.
-Another important challenge in solving real-world symbolic regression problems is the ability for a given model to generalize, we want the final model to show high accuracy in predicting outcomes for previously unseen cases.
+The goal in symbolic regression is to find a mathematical model for a given set of data points [@10.1007/978-3-540-24621-3_22].
+In real-world symbolic regression these data points are sourced from real-world observations which in contrast to synthetic symbolic regression problem are more likely to contain random noise and bias.
+An important challenge in solving real-world symbolic regression problems is the ability for a given model to generalize, we want the final model to show high accuracy in predicting outcomes for previously unseen cases.
 
@@
 The objective of the airfoil problem is to find a function that accurately predicts the output variable $y$ by taking in a subset of the input variables $x1,x2,x3,x4,x5$.
-The function set used by all DAE-GP variations for symbolic regression is summarized in tables \@ref(tab:baseParams-GP) and , the terminal set consists of the the input variables $x1,x2,x3,x4,x5$ and the ephemeral integer constants listed in table \@ref(tab:baseParams-GP) [@dae-gp_2022_symreg].
+The function set used by all DAE-GP variations for symbolic regression is summarized in tables \@ref(tab:baseParams-GP).
+The terminal set consists of the the input variables $x1,x2,x3,x4,x5$ and the ephemeral integer constants listed in table \@ref(tab:baseParams-GP) [@dae-gp_2022_symreg].
 

### 2023-02-13 paper.Rmd
Mon Feb 13 10:00:08 CET 2023: update-remote.sh

@@
 As expected, this reduction of hidden neurons by dynamic adjustment, does also result in a reduction of the fitness of solutions found.
-Figure \@ref(fig:airfoil-maxIndSize-fitness) shows that pre-trained DAE-GP, with it's more complex DAE-LSTMs over the full 30 generations, does achieve a higher median fitness than regular DAE-GP with dynamic adjustment of hidden neurons.
+Figure \@ref(fig:airfoil-maxIndSize-fitness) shows that pre-trained DAE-GP, with its more complex DAE-LSTMs over the full 30 generations, does achieve a higher median fitness than regular DAE-GP with dynamic adjustment of hidden neurons.
 Pre-Training does therefore introduce a new Trade-Off decision into the adjustment of DAE-GP hyperparameters where the number of hidden neurons needs to be carefully selected since it positively impacts solution quality but also increases the computational cost of DAE-LSTM training.
@@
 Since the framework that was used is written in the, comparatively slow, python programming language, the absolute run-time measurements are only of limited meaningfulness.
-Nonetheless, by analyzing the relative run-time differences between DAE-GP and it's pre-trained alternative, the influence of pre-training on DAE-GP run-time can be explored.
+Nonetheless, by analyzing the relative run-time differences between DAE-GP and its pre-trained alternative, the influence of pre-training on DAE-GP run-time can be explored.
 
@@
 Better median values (e.g. higher fitness) are printed bold and p-values are marked with asterisk (1,2,3) for $\alpha$ levels ($0.1$, $0.05$, $0.01$).
-Note that I included the results for the airfoil problem with a single hidden layer into the statistical evaluation which is not visualized during this chapter (see subsection [Reduced Number of Hidden Layers]).
+Note that I included the results for the airfoil problem with a single hidden layer into the statistical evaluation which is not visualized during this chapter (see subsection [Reduced Number of Hidden Layers](#reduced-number-of-hidden-layers-1)).
 
-Also note that the results for the real world symbolic regression problems inside this chapter, except for the airfoil problem (see subsection [Influence on Run-Time]), have not been analyzed for their total run-time.
+Also note that the results for the real world symbolic regression problems inside this chapter, except for the airfoil problem (see subsection [Influence on Run-Time](#influence-on-run-time-1)), have not been analyzed for their total run-time.
 This is due to the fact, that experiments were run on different computer architectures which would result in biased results.
@@
 
-The experiments conducted in the scope of this master thesis did find that the implemented pre-training strategy resulted in a highly statistically significant increase in population diversity measured by the normalized levenshtein edit distance (see subsection [Influence on Population Diversity]).
+The experiments conducted in the scope of this master thesis did find that the implemented pre-training strategy resulted in a highly statistically significant increase in population diversity measured by the normalized levenshtein edit distance (see subsection [Influence on Population Diversity](#influence-on-population-diversity-1)).
 Increasing the population diversity can be an important mechanism in controlling the search behavior of evolutionary algorithms and is especially important for the global exploration of the search space and the avoidance of premature convergence of the search [@Sudholt2020].
 
-Another significant benefit of using pre-training in DAE-GP is that it is highly efficient in decreasing the number of training epochs that are necessary to train each generations DAE-LSTM model (see subsection [Influence on Population Diversity]).
+Another significant benefit of using pre-training in DAE-GP is that it is highly efficient in decreasing the number of training epochs that are necessary to train each generations DAE-LSTM model (see subsection [Influence on Population Diversity](#influence-on-population-diversity-1)).
 Even though pre-training increased the median total run-time for the tested benchmark problem in combination with the chosen DAE-LSTM dimensions, reducing the number of training epochs might still be a possible way to lower the run time of DAE-GP if further performance optimizations, especially during model sampling, can be achieved.
@@
 
-As detailed in the subsection [Dynamic Adjustment of Hidden Neurons], one major drawback of using pre-training with DAE-GP is the loss of the ability to dynamically adjust the number of hidden neurons to the maximum size of individuals inside the current generations population.
+As detailed in the subsection [Dynamic Adjustment of Hidden Neurons](#dynamic-adjustment-of-hidden-neurons-1), one major drawback of using pre-training with DAE-GP is the loss of the ability to dynamically adjust the number of hidden neurons to the maximum size of individuals inside the current generations population.
 This dynamic mechanism does reduce the computational resources necessary for training DAE-LSTM networks which is one of the most time consuming operations besides sampling new individuals.
 
-Another disadvantage of using pre-training in DAE-GP is that, in my experiments, it required DAE-LSTM networks to use two hidden layers (see subsection [Reduced Number of Hidden Layers]).
+Another disadvantage of using pre-training in DAE-GP is that, in my experiments, it required DAE-LSTM networks to use two hidden layers (see subsection [Reduced Number of Hidden Layers](#reduced-number-of-hidden-layers-1)).
 For a single hidden layer, pre-training did show a negative impact on the median final fitness of solutions found as it was suggested by [@pmlr-v9-erhan10a].
@@
 Also I did not perform an extended optimization of hyperparameters (e.g by using a grid search strategy) which might yield configurations that increase the benefits of pre-training.
-As described in subsection [Influence on Run-Time], sampling new solutions has been a large contributor to the large performance hit on run-time by using pre-training.
+As described in subsection [Influence on Run-Time](#influence-on-run-time-1), sampling new solutions has been a large contributor to the large performance hit on run-time by using pre-training.
 It could be a worthwhile endevour to explore the main factors that increase sampling time for pre-trained DAE-GP to more efficiently exploit the benefits of pre-training like reducing the number of training epochs.
@@
 Another limitation of this research is the software implementation used for all experiments.
-As briefly touched in the subsection [Influence on Run-Time], I found some evidence that my implementation of pre-training suffered from increased run-time costs for internal error handling in the used `keras` library for deep neural networks.
+As briefly touched in the subsection [Influence on Run-Time](#influence-on-run-time-1), I found some evidence that my implementation of pre-training suffered from increased run-time costs for internal error handling in the used `keras` library for deep neural networks.
 By optimizing the source code or even re-implementing it in a more performance-focused programming language (such as `C++` or `Rust`), large improvements to the pre-training strategy might be achievable.

### 2023-02-13 paper.tex
Sat Feb 11 10:00:05 CET 2023: update-remote.sh
Mon Feb 13 10:00:08 CET 2023: update-remote.sh

@@
 \author{}
-\date{\vspace{-2.5em}Date of Submission: 2023-02-10}
+\date{\vspace{-2.5em}Date of Submission: 2023-02-13}
 
@@
 The DAE-GP algorithm first described by {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{36}{]} trains a DAE-LSTMs for each generation \(g\) of the search from scratch.
-{[}\protect\hyperlink{ref-dae-gp_2022_symreg}{35}{]} and {[}\protect\hyperlink{ref-daegp_explore_exploit}{34}{]} suggests the incorporation of a pre-training strategy into the evolutionary search as a possible way of improving the overall performance of the DAE-GP algorithm.
+{[}\protect\hyperlink{ref-dae-gp_2022_symreg}{35}{]} and {[}\protect\hyperlink{ref-daegp_explore_exploit}{34}{]} suggest the incorporation of a pre-training strategy into the evolutionary search as a possible way of improving the overall performance of the DAE-GP algorithm.
 The key idea is to pre-train an initial DAE-LSTM on a large population of candidate solutions and to use the pre-trained parameters of this initial model in each generation of the search as starting parameters for the current generation DAE-LSTM.
@@
 
-\emph{This section describes the relevant concepts that are necessary for the understanding and classification of DAE-GP as well as the concept of pre-training in artificial neural networks}
+\emph{This section describes the relevant concepts that are necessary for the understanding and classification of DAE-GP as well as the concept of pre-training in artificial neural networks.}
 
@@
 In EC, the quality of a solution is commonly measured as fitness and the time steps of the search are called generations.
-Another important concept in EC is the distinction between genotypes and phenotypes of solutions, the genotype contains the information that is necessary to construct the phenotype, the outer appearance of a particular solution on which we measure the overall quality of solutions.
+Another important concept in EC is the distinction between genotypes and phenotypes of solutions.
+The genotype contains the information that is necessary to construct the phenotype, the outer appearance of a particular solution on which we measure the overall quality of solutions.
 The representation of a solution is therefore defined by the mapping of genotypes to phenotypes {[}\protect\hyperlink{ref-design_of_modern_heuristics}{23}{]}.
-Genetic operators, such as mutation or recombination are usually applied to the genotype of solutions.
+Genetic operators, such as mutation or recombination, are usually applied to the genotype of solutions.
 
@@
 An important quality of GP is the ability to search for solutions of variable length and structure.
-GP is an especially useful meta heuristic for problems where no a priori knowledge about the final form of good solutions is available.
-GPs ability to optimize solutions for their structure as well as for their parameters led to it being one of the most prevalent methods used for symbolic regression {[}\protect\hyperlink{ref-10.1007ux2f978-3-540-24621-3_22}{20}{]}.
+GP is an especially useful meta-heuristic for problems where no a priori knowledge about the final form of good solutions is available.
+GPs ability to optimize solutions for both their structure as well as for their parameters led to it being one of the most prevalent methods used for symbolic regression {[}\protect\hyperlink{ref-10.1007ux2f978-3-540-24621-3_22}{20}{]}.
 
@@
 
-The aim of Estimation of distribution algorithms (EDA) is to replace the standard variation operators used in GA by building probabilistic models that can capture complex dependencies between different decision variables of an optimization problem {[}\protect\hyperlink{ref-design_of_modern_heuristics}{23}{]}.
+The aim of estimation of distribution algorithms (EDA) is to replace the standard variation operators used in EC by building probabilistic models that can capture complex dependencies between different decision variables of an optimization problem {[}\protect\hyperlink{ref-design_of_modern_heuristics}{23}{]}.
 EDAs use this probabilistic model to sample new offspring solutions inside an evolutionary search to replace crossover and/or mutation operators.
-Even though a majority of EDA algorithms are designed to work with fixed length string representations as used in classical GAs, EDA algorithms have been researched an successfully applied to the domain of GP in various different research streams {[}\protect\hyperlink{ref-kim_probMod_GP}{13}{]}.
+Even though a majority of EDA algorithms are designed to work with fixed length string representations as used in classical GAs, EDA algorithms have been researched and successfully applied to the domain of GP in various different research streams {[}\protect\hyperlink{ref-kim_probMod_GP}{13}{]}.
 
@@
 
-One possible way of model building in EDA proposed by {[}\protect\hyperlink{ref-harmless_overfitting_eda}{22}{]} is to use denoising autoencoders (DAEs) as generative models to capture complex probability distributions of decision variables.
-DAE, a variation of the autoencoders (AE), are a widely used type of neural networks in the domain of unsupervised machine learning that maps \(n\) input variables to \(n\) output variables using a hidden representation.
+One possible way of model building in EDA proposed by {[}\protect\hyperlink{ref-harmless_overfitting_eda}{22}{]} is to use denoising autoencoders (DAE) as generative models to capture complex probability distributions of decision variables.
+DAE, a variation of the autoencoder (AE), is a widely used type of neural networks in the domain of unsupervised machine learning that maps \(n\) input variables to \(n\) output variables using a hidden representation.
 
@@
 \item
-  Decoder \(d(h)\): Decode and output the the encoded data back to its original dimension
+  Decoder \(d(h)\): Decode and output the encoded data back to its original dimension
 \end{itemize}
 
-The AE is trained to reduce the reconstruction error between input and output data, after the training procedure is finished the network is able to reduce the dimensional of input data to a compressed representation{[}\protect\hyperlink{ref-ae_orig}{10}{]}.
+The AE is trained to reduce the reconstruction error between input and output data.
+After the training procedure is finished the network is able to reduce the dimensionality of input data to a compressed representation{[}\protect\hyperlink{ref-ae_orig}{10}{]}.
 
@@
 
-LSTMs are a variant of artificial neural networks first introduced by {[}\protect\hyperlink{ref-lstm_orig}{11}{]} that can store learned information over en extended time period while avoiding the problem of vanishing and/or exploding gradients.
-Since DAE-GP encodes candidate solutions as linear strings in prefix notation, the DAE in DAE-GP uses LSTMs for both encoding and decoding where the total amount of time steps \(T\) is equal to sum of the length of the input solution and the output solution {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{36}{]}.
+LSTMs are a variant of artificial neural networks first introduced by {[}\protect\hyperlink{ref-lstm_orig}{11}{]} that can store learned information over en extended period of time while avoiding the problem of vanishing and/or exploding gradients.
+Since DAE-GP encodes candidate solutions as linear strings in prefix notation, the DAE in DAE-GP uses LSTMs for both encoding and decoding where the total amount of time steps \(T\) is equal to the sum of the length of the input solution and the output solution {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{36}{]}.
 
@@
 
-The pre-training strategy used in this thesis is applied to the DAE-LSTM model \(M_g\) that are used in each DAE-GP generation \(g\) for \(g\in{1,...,g_{max}}\).
+The pre-training strategy used in this thesis is applied to the DAE-LSTM models \(M_g\) that are used in each DAE-GP generation \(g\) for \(g\in{1,...,g_{max}}\).
 Instead of initializing and optimizing each model from scratch as done in previous work (e.g.
@@
 
-The probably most obvious motivation for using a pre-training strategy in DAE-GP is based on the fact, that initializing \(M_g\) with pre-trained weights \(\theta_{\hat{M}}\) is likely to reduce the amount of training epochs that are necessary until the point of training error convergence is reached.
+The probably most obvious motivation for using a pre-training strategy in DAE-GP is based on the fact that initializing \(M_g\) with pre-trained weights \(\theta_{\hat{M}}\) is likely to reduce the amount of training epochs that are necessary until the point of training error convergence is reached.
 
 Early research on pre-training for DAE by {[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{5}{]} comes to the conclusion that besides adding robustness to models the strategy also results in improved generalization and better performing models.
-The authors describe that unsupervised pre-training behaves like a regularizer on DAE networks, the mean test error and it's variance are reduced for DAE networks that are initialized from pre-trained weights if compared to the same architectures that use random initialization.
+The authors describe that unsupervised pre-training behaves like a regularizer on DAE networks.
+The mean test error and its variance are reduced for DAE networks that are initialized from pre-trained weights if compared to the same architectures that use random initialization.
 
@@
 {[}\protect\hyperlink{ref-sampling_err_gp}{24}{]} also introduces a method for calculating optimal population sizes to minimize the presence of sampling error.
-Since DAE-LSTM of DAE-GP learns the properties of it's training population and reuses the acquired knowledge in the sampling procedure, using the parameters obtained from training \(\hat{M}\) on a sufficiently large training population might reduce the need for large population sizes in the following generations of the search if \(\hat{M}\) already implicitly captured the properties of a representative sample of solutions.
+Since DAE-LSTM of DAE-GP learns the properties of its training population and reuses the acquired knowledge in the sampling procedure, using the parameters obtained from training \(\hat{M}\) on a sufficiently large training population might reduce the need for large population sizes in the following generations of the search if \(\hat{M}\) already implicitly captured the properties of a representative sample of solutions.
 If this mechanism can be applied successfully it would benefit the performance of DAE-GP by increasing the population diversity (resulting in better solution quality) as well as by reducing the need for large population sizes which require computational resources.
@@
 
-One difficulty in the implementation of a pre-training strategy into DAE-GP has been the determination of the number of hidden neurons in side the DAE-LSTMs hidden layers.
+One difficulty in the implementation of a pre-training strategy into DAE-GP has been the determination of the number of hidden neurons inside the DAE-LSTMs hidden layers.
 The original description of DAE-GP {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{36}{]} uses a strategy where the number of hidden neurons for each hidden layer is dynamically set per generation to the maximum individual size inside the current population.
@@
 Symbolic Regression problems have been one of the first GP applications {[}\protect\hyperlink{ref-Koza1993GeneticP}{15}{]} and are an actively studied and highly relevant research area.
-The goal in symbolic regression is to find a mathematical model for a given set of data points {[}\protect\hyperlink{ref-10.1007ux2f978-3-540-24621-3_22}{20}{]}, in real-world symbolic regression these data points are sourced from real-world observations which in contrast to synthetic symbolic regression problem are more likely to contain random noise and bias.
-Another important challenge in solving real-world symbolic regression problems is the ability for a given model to generalize, we want the final model to show high accuracy in predicting outcomes for previously unseen cases.
+The goal in symbolic regression is to find a mathematical model for a given set of data points {[}\protect\hyperlink{ref-10.1007ux2f978-3-540-24621-3_22}{20}{]}.
+In real-world symbolic regression these data points are sourced from real-world observations which in contrast to synthetic symbolic regression problem are more likely to contain random noise and bias.
+An important challenge in solving real-world symbolic regression problems is the ability for a given model to generalize, we want the final model to show high accuracy in predicting outcomes for previously unseen cases.
 
@@
 The objective of the airfoil problem is to find a function that accurately predicts the output variable \(y\) by taking in a subset of the input variables \(x1,x2,x3,x4,x5\).
-The function set used by all DAE-GP variations for symbolic regression is summarized in tables \ref{tab:baseParams-GP} and , the terminal set consists of the the input variables \(x1,x2,x3,x4,x5\) and the ephemeral integer constants listed in table \ref{tab:baseParams-GP} {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{35}{]}.
+The function set used by all DAE-GP variations for symbolic regression is summarized in tables \ref{tab:baseParams-GP}.
+The terminal set consists of the the input variables \(x1,x2,x3,x4,x5\) and the ephemeral integer constants listed in table \ref{tab:baseParams-GP} {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{35}{]}.
 
@@
 As expected, this reduction of hidden neurons by dynamic adjustment, does also result in a reduction of the fitness of solutions found.
-Figure \ref{fig:airfoil-maxIndSize-fitness} shows that pre-trained DAE-GP, with it's more complex DAE-LSTMs over the full 30 generations, does achieve a higher median fitness than regular DAE-GP with dynamic adjustment of hidden neurons.
+Figure \ref{fig:airfoil-maxIndSize-fitness} shows that pre-trained DAE-GP, with its more complex DAE-LSTMs over the full 30 generations, does achieve a higher median fitness than regular DAE-GP with dynamic adjustment of hidden neurons.
 Pre-Training does therefore introduce a new Trade-Off decision into the adjustment of DAE-GP hyperparameters where the number of hidden neurons needs to be carefully selected since it positively impacts solution quality but also increases the computational cost of DAE-LSTM training.
@@
 Better median values (e.g.~higher fitness) are printed bold and p-values are marked with asterisk (1,2,3) for \(\alpha\) levels (\(0.1\), \(0.05\), \(0.01\)).
-Note that I included the results for the airfoil problem with a single hidden layer into the statistical evaluation which is not visualized during this chapter (see subsection \protect\hyperlink{reduced-number-of-hidden-layers}{Reduced Number of Hidden Layers}).
-
-Also note that the results for the real world symbolic regression problems inside this chapter, except for the airfoil problem (see subsection \protect\hyperlink{influence-on-run-time}{Influence on Run-Time}), have not been analyzed for their total run-time.
+Note that I included the results for the airfoil problem with a single hidden layer into the statistical evaluation which is not visualized during this chapter (see subsection \protect\hyperlink{reduced-number-of-hidden-layers-1}{Reduced Number of Hidden Layers}).
+
+Also note that the results for the real world symbolic regression problems inside this chapter, except for the airfoil problem (see subsection \protect\hyperlink{influence-on-run-time-1}{Influence on Run-Time}), have not been analyzed for their total run-time.
 This is due to the fact, that experiments were run on different computer architectures which would result in biased results.
@@
 
-The experiments conducted in the scope of this master thesis did find that the implemented pre-training strategy resulted in a highly statistically significant increase in population diversity measured by the normalized levenshtein edit distance (see subsection \protect\hyperlink{influence-on-population-diversity}{Influence on Population Diversity}).
+The experiments conducted in the scope of this master thesis did find that the implemented pre-training strategy resulted in a highly statistically significant increase in population diversity measured by the normalized levenshtein edit distance (see subsection \protect\hyperlink{influence-on-population-diversity-1}{Influence on Population Diversity}).
 Increasing the population diversity can be an important mechanism in controlling the search behavior of evolutionary algorithms and is especially important for the global exploration of the search space and the avoidance of premature convergence of the search {[}\protect\hyperlink{ref-Sudholt2020}{25}{]}.
 
-Another significant benefit of using pre-training in DAE-GP is that it is highly efficient in decreasing the number of training epochs that are necessary to train each generations DAE-LSTM model (see subsection \protect\hyperlink{influence-on-population-diversity}{Influence on Population Diversity}).
+Another significant benefit of using pre-training in DAE-GP is that it is highly efficient in decreasing the number of training epochs that are necessary to train each generations DAE-LSTM model (see subsection \protect\hyperlink{influence-on-population-diversity-1}{Influence on Population Diversity}).
 Even though pre-training increased the median total run-time for the tested benchmark problem in combination with the chosen DAE-LSTM dimensions, reducing the number of training epochs might still be a possible way to lower the run time of DAE-GP if further performance optimizations, especially during model sampling, can be achieved.
@@
 
-As detailed in the subsection \protect\hyperlink{dynamic-adjustment-of-hidden-neurons}{Dynamic Adjustment of Hidden Neurons}, one major drawback of using pre-training with DAE-GP is the loss of the ability to dynamically adjust the number of hidden neurons to the maximum size of individuals inside the current generations population.
+As detailed in the subsection \protect\hyperlink{dynamic-adjustment-of-hidden-neurons-1}{Dynamic Adjustment of Hidden Neurons}, one major drawback of using pre-training with DAE-GP is the loss of the ability to dynamically adjust the number of hidden neurons to the maximum size of individuals inside the current generations population.
 This dynamic mechanism does reduce the computational resources necessary for training DAE-LSTM networks which is one of the most time consuming operations besides sampling new individuals.
 
-Another disadvantage of using pre-training in DAE-GP is that, in my experiments, it required DAE-LSTM networks to use two hidden layers (see subsection \protect\hyperlink{reduced-number-of-hidden-layers}{Reduced Number of Hidden Layers}).
+Another disadvantage of using pre-training in DAE-GP is that, in my experiments, it required DAE-LSTM networks to use two hidden layers (see subsection \protect\hyperlink{reduced-number-of-hidden-layers-1}{Reduced Number of Hidden Layers}).
 For a single hidden layer, pre-training did show a negative impact on the median final fitness of solutions found as it was suggested by {[}\protect\hyperlink{ref-pmlr-v9-erhan10a}{4}{]}.
@@
 Also I did not perform an extended optimization of hyperparameters (e.g by using a grid search strategy) which might yield configurations that increase the benefits of pre-training.
-As described in subsection \protect\hyperlink{influence-on-run-time}{Influence on Run-Time}, sampling new solutions has been a large contributor to the large performance hit on run-time by using pre-training.
+As described in subsection \protect\hyperlink{influence-on-run-time-1}{Influence on Run-Time}, sampling new solutions has been a large contributor to the large performance hit on run-time by using pre-training.
 It could be a worthwhile endevour to explore the main factors that increase sampling time for pre-trained DAE-GP to more efficiently exploit the benefits of pre-training like reducing the number of training epochs.
@@
 Another limitation of this research is the software implementation used for all experiments.
-As briefly touched in the subsection \protect\hyperlink{influence-on-run-time}{Influence on Run-Time}, I found some evidence that my implementation of pre-training suffered from increased run-time costs for internal error handling in the used \texttt{keras} library for deep neural networks.
+As briefly touched in the subsection \protect\hyperlink{influence-on-run-time-1}{Influence on Run-Time}, I found some evidence that my implementation of pre-training suffered from increased run-time costs for internal error handling in the used \texttt{keras} library for deep neural networks.
 By optimizing the source code or even re-implementing it in a more performance-focused programming language (such as \texttt{C++} or \texttt{Rust}), large improvements to the pre-training strategy might be achievable.

### 2023-02-14 paper.Rmd
Tue Feb 14 15:00:08 CET 2023: update-remote.sh

@@
 
-To gather a deeper understanding about the effect of pre-training DAE-LSTM in DAE-GP a series of experiments was conducted using the airfoil dataset for symbolic regression while using different parameter configurations for the number of hidden layers as well as the number of hidden neurons per hidden layer.
+To gain a deeper understanding about the effect of pre-training the DAE-LSTM networks in DAE-GP, a series of experiments was conducted using the airfoil dataset for symbolic regression while using different parameter configurations for the number of hidden layers as well as the number of hidden neurons per hidden layer.
 
-To study generalization behavior this series of experiments is conducted using DAE-GP with only a single generation until the search terminates.
-I disregard the fitness of solutions found and focus solely on the reconstruction error that is produced during the training of each DAE-LSTM.
+To study the generalization behavior of the DAE-LSTM model, this series of experiments is conducted using DAE-GP with only a single generation until the search terminates.
+The fitness of solutions found was disregarded to focus solely on the reconstruction error that is produced during the training of each DAE-LSTM.
 The reconstruction error is measured for two separate populations, a training population $P_{train}$ that is used to train our DAE-LSTM as well as a hold-out validation population $P_{test}$.
@@
 
-For each configuration I tested traditional DAE-GP as well as a pre-trained DAE-GP resulting in 12 total sub experiments that were each based on 10 individual runs (total number of runs=$120$).
+For each configuration, traditional DAE-GP as well as a pre-trained DAE-GP was tested resulting in 12 total sub experiments that were each based on 10 individual runs (total number of runs=$120$).
 To avoid creating biased results through the presence of sampling error inside the pre-training population (see @sampling_err_gp), the population size for the pre-training phase is chosen very high with the size of $\hat{P}$ =10000 where 50% of $\hat{P}$ is used for the training population $\hat{P}_{train}$ and the remaining 50% are used for the hold-out validation population $\hat{P}_{test}$.
-The training of DAE-LSTM uses a fixed number of 1000 epochs until termination and uses Adam optimization for gradient descent.
-The reason for using a high amount of 1000 fixed training epochs is to deliberately force the DAE-LSTM to overfit to the training data.
+The training of each DAE-LSTM stopped after a fixed number of 1000 epochs.
+The reason for using a high amount of 1000 fixed training epochs was to deliberately force the DAE-LSTM to overfit to the training data.
 
-In general I expect that with a growing number of trainable parameters (either by adding more hidden layers or more hidden neurons per layer) the DAE-LSTM will be more prone to overfit to the training population $P_{train}$ resulting in a small reconstruction error for the training population and a large one for the validation population $P_{test}$.
+In general, it is expected that with a growing number of trainable parameters (either by adding more hidden layers or more hidden neurons per layer) the DAE-LSTM will be more likely to overfit to the training population $P_{train}$ which will result in a small reconstruction error for the training population and a large one for the validation population $P_{test}$.
 The reason for this effect is that a larger network trained over an extended period of time (without the use of strategies like early stopping), has much more potential to learn noise from the training dataset than a smaller network, which is more likely to result in worsening performance for previously unseen cases [@weigend1994overfitting].
 
-Based on the review of [@pmlr-v5-erhan09a] I also expect that pre-training will have an insignificant or even negative influence on small DAE-LSTM instances while improving the networks generalization ability with growing size.
+Based on the review of [@pmlr-v5-erhan09a] it is also expected that pre-training will have an insignificant or even negative influence on small DAE-LSTM instances while improving the networks generalization ability with growing size.
 
@@
 
-The results for the experiment with varying number of hidden layers are summarized in figure \@ref(fig:first-gen-airfoil-byLayers).
+The results for the experiment with a varying number of hidden layers are summarized in figure \@ref(fig:first-gen-airfoil-byLayers).
 Again, it can be observed that by increasing the dimension of the DAE-LSTM network both algorithms increasingly overfit to the training data.
@@
 
-After closely examining the effect of pre-training on DAE-GPs generalization behavior another series of experiments is conducted to study how pre-training influences the overall search behavior of DAE-GP.
-The airfoil problem was selected as a first benchmark problem to test pre-trained DAE-GP against traditional DAE-GP.
+After closely examining the effect of pre-training on DAE-GPs generalization behavior, another series of experiments is conducted to study how pre-training influences the overall search behavior of DAE-GP.
 
@@
 
-Figure \@ref(fig:airfoil-median-fitness-30gens) shows both the mean and median of the best fitness for each generation on both the test and training set of the airfoil dataset for using 2 hidden layers inside the DAE-LSTM networks.
+Figure \@ref(fig:airfoil-median-fitness-30gens) shows both the mean and median the best fitness by generations for the airfoil dataset.
 The results are aggregated for $10$ individuals runs per algorithm and also include regular GP as a benchmark.
@@
 The first interesting observation from this experiment is, that the pre-trained DAE-GP shows slightly lower mean RMSE (higher fitness) but a higher median RMSE (lower fitness) than regular DAE-GP for both the training and the testing set in overall fitness.
-Regarding the presence of overfitting, both DAE-GP variants show only a very small gap between the testing and training fitness which indicates a good generalization behavior.
-Another interesting observation from figure \@ref(fig:airfoil-median-fitness-30gens) is that pre-trained DAE-GP is faster at improving fitness than regular DAE-GP, this is especially true if only looking at the first 10 generations.
+Regarding the presence of overfitting, both DAE-GP variants show only a very small gap between the testing and training fitness which indicates a good generalization behavior.\
+
+Another interesting observation from figure \@ref(fig:airfoil-median-fitness-30gens) is that pre-trained DAE-GP is faster at improving fitness than regular DAE-GP.
 The median best RMSE on the test set after 10 generations for pre-trained DAE-GP is `r round(PT_TEST_MED_10F, 3)` which is smaller by a factor of `r round( PT_TEST_MED_10F / REG_TEST_MED_10F, 3)` than regular DAE-GP with a median best RMSE of `r round( REG_TEST_MED_10F, 3)`
@@
 
-Since previous work by @dae-gp_2022_symreg demonstrated that DAE-GP, for a given number of 10.000 fitness evaluations, is able to produce solutions that are much smaller in tree size than regular GP, I also studied the influence of pre-training on the size of solutions found by DAE-GP.
+Since previous work by @dae-gp_2022_symreg demonstrated that DAE-GP, for a given number of 10.000 fitness evaluations, is able to produce solutions that are much smaller in tree size than regular GP, the experiment also analyzed the influence of pre-training on the size of solutions found by DAE-GP.
 Figure \@ref(fig:airfoil-size-combined-30gens) shows the median of both the average size of individuals inside the population as well as the size of the currently best performing solution inside the population.

### 2023-02-14 paper.Rmd
autogit

@@
 Since previous work by @dae-gp_2022_symreg demonstrated that DAE-GP, for a given number of 10.000 fitness evaluations, is able to produce solutions that are much smaller in tree size than regular GP, the experiment also analyzed the influence of pre-training on the size of solutions found by DAE-GP.
+
 Figure \@ref(fig:airfoil-size-combined-30gens) shows the median of both the average size of individuals inside the population as well as the size of the currently best performing solution inside the population.
-It can be observed that both DAE-GP variants strongly reduce the average solutions size already in the first generations and that the average solution size inside the population constantly stays at this very low level for the rest of the search.
+It can be observed that both DAE-GP variants strongly reduce the average solution size already in the first generations and that the average solution size inside the population constantly stays at this very low level for the rest of the search.
 The median average size over all generations is `r round(REG_MED_AVGSIZE, 3)` for DAE-GP and `r round(PT_MED_AVGSIZE, 3)` for pre-trained DAE-GP.
@@
 Another interesting metric to examine in population-based optimization algorithms is the diversity of the current generations population.
-Since individual solutions in DAE-GP are computer programs in the form of parse trees, I selected the normalized Levenshtein edit distance [@NormLevDistance] as my primary metric for tracking population diversity.
+Since individual solutions in DAE-GP are computer programs in the form of parse trees, the normalized Levenshtein edit distance [@NormLevDistance] was selected as a metric for tracking population diversity.
 Figure \@ref(fig:airfoil-popDiversity) shows the median population diversity for both DAE-GP algorithms and regular GP as a benchmark.
 While the population diversity steadily decreases over the generations for standard GP, DAE-GP variants strongly increase the population diversity in the first generations and keep them at a higher level during the evolutionary search.
-Here it can be observed, that pre-training shows a positive effect by increasing population diversity in comparison to regular DAE-GP (Median over all generations for DAE-GP: `r round(REG_MED_LEVDIV, 3)`, Pre-Trained DAE-GP: `r round(PT_MED_LEVDIV, 3)`, Standard-GP `r round(GP_MED_LEVDIV, 3)`)
+Here it can be observed, that pre-training shows a positive effect by increasing population diversity in comparison to regular DAE-GP (Median over all generations for DAE-GP: `r round(REG_MED_LEVDIV, 3)`, Pre-Trained DAE-GP: `r round(PT_MED_LEVDIV, 3)`, Standard-GP `r round(GP_MED_LEVDIV, 3)`).
 
@@
 To ensure a fair comparison between pre-trained and regular DAE-GP, the prior experiments used a static number of hidden neurons for each hidden Layer inside the DAE-LSTM networks.
-For comparison, figure \@ref(fig:airfoil-maxIndSize-params) shows the total number of trainable parameters for each generation's DAE-LSTM of another experiment where regular DAE-GP was tested with dynamic adjustment of the number of hidden neurons as described in [@dae-gp_2020_rtree] and [@dae-gp_2022_symreg].
-Again, the experiments was based on $10$ individual runs and otherwise used the same hyperparameters as listed in tables \@ref(tab:baseParams-GP) and \@ref(tab:baseParams-DAEGP).
+For comparison, figure \@ref(fig:airfoil-maxIndSize-params) shows the total number of trainable parameters for each generations DAE-LSTM of another experiment where regular DAE-GP was tested with dynamic adjustment of the number of hidden neurons as described in [@dae-gp_2020_rtree] and [@dae-gp_2022_symreg].
+Again, the experiment was based on $10$ individual runs and otherwise used the same hyperparameters as listed in tables \@ref(tab:baseParams-GP) and \@ref(tab:baseParams-DAEGP).
 It clearly shows that DAE-GP can strongly reduce the number of trainable parameters by adjusting the number of hidden neurons per hidden layers to the current generations population which largely reduces the amount of training necessary and likely reduce the run-time of DAE-GP.
@@
 
-As expected, this reduction of hidden neurons by dynamic adjustment, does also result in a reduction of the fitness of solutions found.
+As expected, the reduction of hidden neurons by dynamic adjustment does also result in a reduction of the fitness of solutions found.
 Figure \@ref(fig:airfoil-maxIndSize-fitness) shows that pre-trained DAE-GP, with its more complex DAE-LSTMs over the full 30 generations, does achieve a higher median fitness than regular DAE-GP with dynamic adjustment of hidden neurons.
@@
 
-Since previous research on DAE-GP used a single hidden layers for the airfoil dataset [@dae-gp_2022_symreg], I repeated the experiment with a single hidden layer.
-Based on the findings of @pmlr-v5-erhan09a, I expected that pre-training DAE-GP would lead to a decrease in overall performance.
-Figures \@ref(fig:airfoil-fitness-perGen) and \@ref(fig:airfoil-finalFitness) shows the mean and median best fitness per generation as well as the distribution of final fitness scores after 30 generations.
+Since previous research on DAE-GP used a single hidden layers for the airfoil dataset [@dae-gp_2022_symreg], the experiment was repeated with a single hidden layer.
+Based on the findings of @pmlr-v5-erhan09a, it is to be expected that pre-training DAE-GP with a single hidden layer inside the DAE-LSTM networks leads to a decrease in overall performance.
+Figures \@ref(fig:airfoil-fitness-perGen) and \@ref(fig:airfoil-finalFitness) show the mean and median best fitness per generation as well as the distribution of final fitness scores after 30 generations.
 The results are, again, based on $10$ individual runs per algorithm.
@@
 Interestingly, figure \@ref(fig:airfoil-finalFitness) shows that the distribution of final fitness scores for pre-trained DAE-GP has a larger interquartile range than regular DAE-GP.
-This is opposed to the distributions of final fitness score that was measured for the same problem run with two hidden layers (see figure \@ref(fig:airfoil-distribution-fitness-30gens)).
-Otherwise, pre-trained DAE-GP again has more outliers for high RMSE than regular DAE-GP as seen in the previous experiment with 2 hidden layers.
+This is opposed to the distributions of final fitness scores that was measured for the same problem run with two hidden layers (see figure \@ref(fig:airfoil-distribution-fitness-30gens)).
+Otherwise, pre-trained DAE-GP again has more outliers for high RMSE than regular DAE-GP as seen in the previous experiment with two hidden layers.
 
@@
 The pre-training implementation in previous experiments used `ramped half and half` initialization [@Koza1993GeneticP] to create $\hat{P}$.
-Using ramped half and half leads to 50% of the population being initialized by the `full` method, where terminals are only inserted at the maximum allowed depth of a tree resulting in full trees.
-An alternative method of initializing $\hat{P}$ is to use the `grow` initialization method where solutions are not guaranteed to be full trees since every leaf above the maximum depth of a tree can also be a terminal expression.
+Using `ramped half and half` leads to 50% of the population being initialized by the `full` method, where terminals are only inserted at the maximum allowed depth of a tree resulting in full trees.
+An alternative method of initializing $\hat{P}$ is to use the `grow` initialization method where solutions are not guaranteed to be full trees since every leaf below the maximum depth of a tree can also be a terminal expression.
 This different structural composition of solutions inside the pre-training population might be another way to improve the pre-training strategy by training on a population that more closely resembles the structural properties of further generations population where a lower percentage of full grown trees is to be expected.
@@
 
-Surprisingly, as shown in figure \@ref(fig:airfoil-runtime-total), for the experiment conducted on the airfoil dataset with 2 hidden layers a strong negative impact of pre-training can be observed for the median run-time of DAE-GP.
+Surprisingly, as shown in figure \@ref(fig:airfoil-runtime-total), for the experiment conducted on the airfoil dataset with two hidden layers, a strong negative impact of pre-training can be observed for the median run-time of DAE-GP.
 If the time spend on pre-training is included into the total run-time, the median duration for pre-trained DAE-GP over all 10 runs increases by a factor of `r round(PT_TIME / REG_TIME, 3)` from a median run-time of `r format(REG_TIME, scientific=F)` seconds for regular DAE-GP to `r format(PT_TIME, scientific=F)` seconds for the pre-trained algorithm.
@@
 
-To investigate the strong negative impact of pre-training on DAE-GP run-time, I first studied the influence of pre-training on the median number of training epochs that had to be spend at each generation for the training of the DAE-LSTM models $M_g$ (excluding the pre-training DAE-LSTM model $\hat{M}$).
-Figure \@ref(fig:airfoil-trainingEpochs) shows that pre-training, as expected, strongly reduces the number of training epochs that had to be executed at each generation until the termination condition was reached.
-While regular DAE-GP trains for a median of `r format(REG_MEDIAN_EPOCHSTRAINED, scientific=F)` epochs at each generation, pre-training reduces the number of median training epochs by a factor of `r round(PT_MEDIAN_EPOCHSTRAINED / REG_MEDIAN_EPOCHSTRAINED, 3)` to `r format(PT_MEDIAN_EPOCHSTRAINED, scientific=F)` training epochs.
+To investigate the strong negative impact of pre-training on DAE-GP run-time, the influence of pre-training on the median number of training epochs that had to be spend at each generation for the training of the DAE-LSTM models $M_g$ (excluding the pre-training DAE-LSTM model $\hat{M}$) was studied.
+Figure \@ref(fig:airfoil-trainingEpochs) shows that pre-training, as expected, strongly reduces the number of training epochs that had to be executed at each generation until the termination condition was satisfied.
+While regular DAE-GP trained for a median of `r format(REG_MEDIAN_EPOCHSTRAINED, scientific=F)` epochs at each generation, pre-training reduces the number of median training epochs by a factor of `r round(PT_MEDIAN_EPOCHSTRAINED / REG_MEDIAN_EPOCHSTRAINED, 3)` to `r format(PT_MEDIAN_EPOCHSTRAINED, scientific=F)` training epochs.
 
@@
 
-While investigating other reasons for the differences in run-time, I verified that both DAE-GP variants use the exact same number of trainable parameters for each generations DAE-LSTMS models (total number of trainable parameters DAE-GP=`r format(REG_PARAMS, scientific=F)`, Pre-trained DAE-GP=`r format(PT_PARAMS, scientific=F)`).
+While investigating other reasons for the differences in run-time, I verified that both DAE-GP variants use the exact same number of trainable parameters for each generation's DAE-LSTM model $M_g$ (total number of trainable parameters DAE-GP=`r format(REG_PARAMS, scientific=F)`, Pre-trained DAE-GP=`r format(PT_PARAMS, scientific=F)`).
 After profiling a single run of DAE-GP and pre-trained DAE-GP using the python `cProfiler` module, my best explanation for the remaining differences in run-time is an inefficiency in my implementation of pre-training.

### 2023-02-14 paper.Rmd
autogit

@@
 The framework that was used to run all experiments of this thesis was provided by the supervisor of this thesis David Wittenberg and uses the python programming language [@python3] in conjunction with the baseline libraries `Keras` [@chollet2015keras] for deep neural networks and `deap` [@DEAP_JMLR2012] for evolutionary computation.
-For data visualizations I used the python package `matplotlib` [@matplotlib], statistical analysis and evaluations were conducted using primarily the data science libraries `scipy` [@scipy], `numpy` [@numpy], `pandas` [@pandas], and `IPython` [@ipython].
+For data visualizations the python package `matplotlib` [@matplotlib] was used, statistical analysis and evaluations were conducted using primarily the data science libraries `scipy` [@scipy], `numpy` [@numpy] and `pandas` [@pandas].
 
@@
 
-While investigating other reasons for the differences in run-time, I verified that both DAE-GP variants use the exact same number of trainable parameters for each generation's DAE-LSTM model $M_g$ (total number of trainable parameters DAE-GP=`r format(REG_PARAMS, scientific=F)`, Pre-trained DAE-GP=`r format(PT_PARAMS, scientific=F)`).
+While investigating other reasons for the differences in run-time, it was verified that both DAE-GP variants use the exact same number of trainable parameters for each generation's DAE-LSTM model $M_g$ (total number of trainable parameters DAE-GP=`r format(REG_PARAMS, scientific=F)`, Pre-trained DAE-GP=`r format(PT_PARAMS, scientific=F)`).
 After profiling a single run of DAE-GP and pre-trained DAE-GP using the python `cProfiler` module, my best explanation for the remaining differences in run-time is an inefficiency in my implementation of pre-training.
@@
 
-To gather more confidence in the previous results, I applied the pre-training strategy for DAE-GP to three additional real world symbolic regression problems.
+To gather more confidence in the previous results, the pre-training strategy for DAE-GP was applied to three additional real world symbolic regression problems.
 Table \@ref(tab:full-run-realWorldSymReg-problems) briefly summarizes the benchmark problems.
@@
 
-To test for statistical differences of the results achieved by either pre-trained or regular DAE-GP, I selected the nonparametric Mann-Whitney-U Test ([@mwu1], [@mwu2]) to test for differences in the distribution underlying the samples sourced from my experiments.
-Additionally, I included the effect strength of pre-training as measured by Cliffs Delta [@Cliff1993DominanceSO] into the statistical analysis.
+To test for statistical differences of the results achieved by either pre-trained or regular DAE-GP, the nonparametric Mann-Whitney-U Test ([@mwu1], [@mwu2]) was selected to test for differences in the underlying distribution.
+Additionally, the effect strength of pre-training as measured by Cliffs Delta [@Cliff1993DominanceSO] was included into the statistical analysis.
 Better median values (e.g. higher fitness) are printed bold and p-values are marked with asterisk (1,2,3) for $\alpha$ levels ($0.1$, $0.05$, $0.01$).
-Note that I included the results for the airfoil problem with a single hidden layer into the statistical evaluation which is not visualized during this chapter (see subsection [Reduced Number of Hidden Layers](#reduced-number-of-hidden-layers-1)).
+Note that the results for the airfoil problem with a single hidden layer were also included into the statistical evaluation which is not visualized during this chapter (see subsection [Reduced Number of Hidden Layers](#reduced-number-of-hidden-layers-1)).
 
@@
 
-Next I studied the size of solutions found during DAE-GP.
+Next, the size of solutions found during DAE-GP was analyzed.
 Figure \@ref(fig:symReg-Size) shows the median size of the best solution per generation aggregated over all benchmark problems for symbolic regression.
@@
 One major limitation for the experiments conducted during the research phase of this thesis has been the large computational effort for experimenting with DAE-GP.
-Since single runs for DAE-GP have a run-time of many hours, it was not possible for me to extend the experiments to larger population sizes or more generations.
-Also I did not perform an extended optimization of hyperparameters (e.g by using a grid search strategy) which might yield configurations that increase the benefits of pre-training.
+Since single runs for DAE-GP have a run-time of many hours, it was not possible to extend the experiments to larger population sizes or more generations.
+Also no extended optimization of hyperparameters (e.g by using a grid search strategy) have been performed which might yield configurations that increase the benefits of pre-training.
 As described in subsection [Influence on Run-Time](#influence-on-run-time-1), sampling new solutions has been a large contributor to the large performance hit on run-time by using pre-training.
@@
 Another limitation of this research is the software implementation used for all experiments.
-As briefly touched in the subsection [Influence on Run-Time](#influence-on-run-time-1), I found some evidence that my implementation of pre-training suffered from increased run-time costs for internal error handling in the used `keras` library for deep neural networks.
+As briefly touched in the subsection [Influence on Run-Time](#influence-on-run-time-1), some evidence was found that my implementation of pre-training suffered from increased run-time costs for internal error handling in the used `keras` library for deep neural networks.
 By optimizing the source code or even re-implementing it in a more performance-focused programming language (such as `C++` or `Rust`), large improvements to the pre-training strategy might be achievable.

### 2023-02-14 paper.tex
Tue Feb 14 13:00:08 CET 2023: update-remote.sh
Tue Feb 14 20:00:07 CET 2023: update-remote.sh

@@
 \author{}
-\date{\vspace{-2.5em}Date of Submission: 2023-02-13}
+\date{\vspace{-2.5em}Date of Submission: 2023-02-14}
 
@@
 
-Denoising Autoencoder Genetic Programming (DAE-GP) is a novel variation of an genetic programming based Estimation of Distribution Algorithm (EDA-GP) that uses a denoising autoencoders long short-term memory network (DAE-LSTMs) as a probabilistic model to sample new candidate solutions {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{36}{]}.
+Denoising Autoencoder Genetic Programming (DAE-GP) is a novel variation of an genetic programming based Estimation of Distribution Algorithm (EDA-GP) that uses a denoising autoencoders long short-term memory network (DAE-LSTMs) as a probabilistic model to sample new candidate solutions {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{35}{]}.
 
@@
 In DAE-GP, DAE-LSTMs are trained with a subset of high-fitness solutions selected from a parent population with the aim to capture their promising qualities.
-The resulting model is then used to sample new offspring solutions by propagating partially mutated solutions from the parent population through the DAE-LSTMS {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{36}{]}.
-In previous work DAE-GP has been shown to outperform GP for both a generalized version of the royal tree problem {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{36}{]} as well as for a real-world symbolic regression problem {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{35}{]}.
-
-The DAE-GP algorithm first described by {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{36}{]} trains a DAE-LSTMs for each generation \(g\) of the search from scratch.
-{[}\protect\hyperlink{ref-dae-gp_2022_symreg}{35}{]} and {[}\protect\hyperlink{ref-daegp_explore_exploit}{34}{]} suggest the incorporation of a pre-training strategy into the evolutionary search as a possible way of improving the overall performance of the DAE-GP algorithm.
+The resulting model is then used to sample new offspring solutions by propagating partially mutated solutions from the parent population through the DAE-LSTMS {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{35}{]}.
+In previous work DAE-GP has been shown to outperform GP for both a generalized version of the royal tree problem {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{35}{]} as well as for a real-world symbolic regression problem {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{34}{]}.
+
+The DAE-GP algorithm first described by {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{35}{]} trains a DAE-LSTMs for each generation \(g\) of the search from scratch.
+{[}\protect\hyperlink{ref-dae-gp_2022_symreg}{34}{]} and {[}\protect\hyperlink{ref-daegp_explore_exploit}{33}{]} suggest the incorporation of a pre-training strategy into the evolutionary search as a possible way of improving the overall performance of the DAE-GP algorithm.
 The key idea is to pre-train an initial DAE-LSTM on a large population of candidate solutions and to use the pre-trained parameters of this initial model in each generation of the search as starting parameters for the current generation DAE-LSTM.
@@
 
-DAE-GP is an EDA-GP algorithm that uses DAE-LSTM networks as a probabilistic model to sample new offspring solutions {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{36}{]}.
+DAE-GP is an EDA-GP algorithm that uses DAE-LSTM networks as a probabilistic model to sample new offspring solutions {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{35}{]}.
 
@@
 The genotype contains the information that is necessary to construct the phenotype, the outer appearance of a particular solution on which we measure the overall quality of solutions.
-The representation of a solution is therefore defined by the mapping of genotypes to phenotypes {[}\protect\hyperlink{ref-design_of_modern_heuristics}{23}{]}.
+The representation of a solution is therefore defined by the mapping of genotypes to phenotypes {[}\protect\hyperlink{ref-design_of_modern_heuristics}{22}{]}.
 Genetic operators, such as mutation or recombination, are usually applied to the genotype of solutions.
@@
 GP follows the same basic evolutionary principle of EC but searches for more general, hierarchical computer programs of dynamically varying size and shape {[}\protect\hyperlink{ref-Koza1993GeneticP}{15}{]}.
-The computer programs that are at the center of the evolutionary search in GP are commonly represented by tree structures at the level of their phenotype {[}\protect\hyperlink{ref-design_of_modern_heuristics}{23}{]}.
+The computer programs that are at the center of the evolutionary search in GP are commonly represented by tree structures at the level of their phenotype {[}\protect\hyperlink{ref-design_of_modern_heuristics}{22}{]}.
 Since GP searches for high fitness computer programs that produce a desired output for some input, it can be applied to various different problem domains such as symbolic regression, automatic programming, or evolving game-playing strategies {[}\protect\hyperlink{ref-Koza1993GeneticP}{15}{]}.
@@
 
-The aim of estimation of distribution algorithms (EDA) is to replace the standard variation operators used in EC by building probabilistic models that can capture complex dependencies between different decision variables of an optimization problem {[}\protect\hyperlink{ref-design_of_modern_heuristics}{23}{]}.
+The aim of estimation of distribution algorithms (EDA) is to replace the standard variation operators used in EC by building probabilistic models that can capture complex dependencies between different decision variables of an optimization problem {[}\protect\hyperlink{ref-design_of_modern_heuristics}{22}{]}.
 EDAs use this probabilistic model to sample new offspring solutions inside an evolutionary search to replace crossover and/or mutation operators.
@@
 
-One possible way of model building in EDA proposed by {[}\protect\hyperlink{ref-harmless_overfitting_eda}{22}{]} is to use denoising autoencoders (DAE) as generative models to capture complex probability distributions of decision variables.
+One possible way of model building in EDA proposed by {[}\protect\hyperlink{ref-harmless_overfitting_eda}{21}{]} is to use denoising autoencoders (DAE) as generative models to capture complex probability distributions of decision variables.
 DAE, a variation of the autoencoder (AE), is a widely used type of neural networks in the domain of unsupervised machine learning that maps \(n\) input variables to \(n\) output variables using a hidden representation.
@@
 
-DAE was first introduced by {[}\protect\hyperlink{ref-dae_orig2008}{30}{]} as an improved AE with the ability to learn new representations of data that is especially robust to partially corrupted input data.
+DAE was first introduced by {[}\protect\hyperlink{ref-dae_orig2008}{29}{]} as an improved AE with the ability to learn new representations of data that is especially robust to partially corrupted input data.
 DAE modifies the AE by using partially corrupted input data for the AE and training it to reconstruct the uncorrupted, original version of the input data.
 
-Since the hidden representation of DAE captures the dependency structure of the input data it can therefore also be used to generate new solutions in the context of GAs {[}\protect\hyperlink{ref-harmless_overfitting_eda}{22}{]}.
-
-DAE-GP builds upon the concept of using DAEs in EDAs described by {[}\protect\hyperlink{ref-harmless_overfitting_eda}{22}{]} and transfers the concept to the domain of GP.
-The mutation and crossover operators of standard GP are replaced by sampling new solutions from a probabilistic model that is build by training a DAEs long short-term memory (LSTM) network on a subset of high fitness solutions from the current population {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{36}{]}.
+Since the hidden representation of DAE captures the dependency structure of the input data it can therefore also be used to generate new solutions in the context of GAs {[}\protect\hyperlink{ref-harmless_overfitting_eda}{21}{]}.
+
+DAE-GP builds upon the concept of using DAEs in EDAs described by {[}\protect\hyperlink{ref-harmless_overfitting_eda}{21}{]} and transfers the concept to the domain of GP.
+The mutation and crossover operators of standard GP are replaced by sampling new solutions from a probabilistic model that is build by training a DAEs long short-term memory (LSTM) network on a subset of high fitness solutions from the current population {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{35}{]}.
 
 LSTMs are a variant of artificial neural networks first introduced by {[}\protect\hyperlink{ref-lstm_orig}{11}{]} that can store learned information over en extended period of time while avoiding the problem of vanishing and/or exploding gradients.
-Since DAE-GP encodes candidate solutions as linear strings in prefix notation, the DAE in DAE-GP uses LSTMs for both encoding and decoding where the total amount of time steps \(T\) is equal to the sum of the length of the input solution and the output solution {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{36}{]}.
+Since DAE-GP encodes candidate solutions as linear strings in prefix notation, the DAE in DAE-GP uses LSTMs for both encoding and decoding where the total amount of time steps \(T\) is equal to the sum of the length of the input solution and the output solution {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{35}{]}.
 
@@
 Instead of initializing and optimizing each model from scratch as done in previous work (e.g.
-{[}\protect\hyperlink{ref-dae-gp_2020_rtree}{36}{]} {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{35}{]}), a separate DAE-LSTM network \(\hat{M}\) will be trained on a large population of randomly initialized solutions.\\
+{[}\protect\hyperlink{ref-dae-gp_2020_rtree}{35}{]} {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{34}{]}), a separate DAE-LSTM network \(\hat{M}\) will be trained on a large population of randomly initialized solutions.\\
 
@@
 Another motivation for introducing pre-training into DAE-GP is the prevalence of sampling error for small GP populations sizes.
-{[}\protect\hyperlink{ref-sampling_err_gp}{24}{]} finds that sampling error, non-systematic errors that are caused by observing only a small subset of the statistical population, is a severe problem in the domain of GP since the initial population may not be a representative sample of all possible solutions.
-{[}\protect\hyperlink{ref-sampling_err_gp}{24}{]} also introduces a method for calculating optimal population sizes to minimize the presence of sampling error.
+{[}\protect\hyperlink{ref-sampling_err_gp}{23}{]} finds that sampling error, non-systematic errors that are caused by observing only a small subset of the statistical population, is a severe problem in the domain of GP since the initial population may not be a representative sample of all possible solutions.
+{[}\protect\hyperlink{ref-sampling_err_gp}{23}{]} also introduces a method for calculating optimal population sizes to minimize the presence of sampling error.
 Since DAE-LSTM of DAE-GP learns the properties of its training population and reuses the acquired knowledge in the sampling procedure, using the parameters obtained from training \(\hat{M}\) on a sufficiently large training population might reduce the need for large population sizes in the following generations of the search if \(\hat{M}\) already implicitly captured the properties of a representative sample of solutions.
@@
 
-The DAE-GP algorithm, first described by {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{36}{]}, is summarized as a flowchart in figure \ref{fig:daegp-flowchart}.
+The DAE-GP algorithm, first described by {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{35}{]}, is summarized as a flowchart in figure \ref{fig:daegp-flowchart}.
 After setting the generation counter \(g\) to \(0\) and creating the initial population of solutions \(P_0\), DAE-GP enters the main loop that checks if a termination criteria is satisfied (i.e.~maximum number of generations or fitness evaluations is reached).
@@
 One difficulty in the implementation of a pre-training strategy into DAE-GP has been the determination of the number of hidden neurons inside the DAE-LSTMs hidden layers.
-The original description of DAE-GP {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{36}{]} uses a strategy where the number of hidden neurons for each hidden layer is dynamically set per generation to the maximum individual size inside the current population.
+The original description of DAE-GP {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{35}{]} uses a strategy where the number of hidden neurons for each hidden layer is dynamically set per generation to the maximum individual size inside the current population.
 For a pre-training implementation, this strategy can not be easily adapted since it leads to a changing number of neurons at each generation resulting in different dimensions of the DAE-LSTM.
@@
 \endlastfoot
-Stopping Condition & Convergence {[}\protect\hyperlink{ref-daegp_explore_exploit}{34}{]} \\
+Stopping Condition & Convergence {[}\protect\hyperlink{ref-daegp_explore_exploit}{33}{]} \\
 Reconstruction Error Metric & Multiclass Cross Entropy \\
@@
 Hidden Neurons & 150 \\
-Corruption Operator & Levenshtein Edit {[}\protect\hyperlink{ref-daegp_explore_exploit}{34}{]} \\
+Corruption Operator & Levenshtein Edit {[}\protect\hyperlink{ref-daegp_explore_exploit}{33}{]} \\
 Denoising Edit Probability (Model Building) & 0.05 \\
@@
 
-The framework that was used to run all experiments of this thesis was provided by the supervisor of this thesis David Wittenberg and uses the python programming language {[}\protect\hyperlink{ref-python3}{28}{]} in conjunction with the baseline libraries \texttt{Keras} {[}\protect\hyperlink{ref-chollet2015keras}{2}{]} for deep neural networks and \texttt{deap} {[}\protect\hyperlink{ref-DEAP_JMLR2012}{6}{]} for evolutionary computation.
-For data visualizations I used the python package \texttt{matplotlib} {[}\protect\hyperlink{ref-matplotlib}{12}{]}, statistical analysis and evaluations were conducted using primarily the data science libraries \texttt{scipy} {[}\protect\hyperlink{ref-scipy}{31}{]}, \texttt{numpy} {[}\protect\hyperlink{ref-numpy}{8}{]}, \texttt{pandas} {[}\protect\hyperlink{ref-pandas}{26}{]}, and \texttt{IPython} {[}\protect\hyperlink{ref-ipython}{21}{]}.
+The framework that was used to run all experiments of this thesis was provided by the supervisor of this thesis David Wittenberg and uses the python programming language {[}\protect\hyperlink{ref-python3}{27}{]} in conjunction with the baseline libraries \texttt{Keras} {[}\protect\hyperlink{ref-chollet2015keras}{2}{]} for deep neural networks and \texttt{deap} {[}\protect\hyperlink{ref-DEAP_JMLR2012}{6}{]} for evolutionary computation.
+For data visualizations the python package \texttt{matplotlib} {[}\protect\hyperlink{ref-matplotlib}{12}{]} was used, statistical analysis and evaluations were conducted using primarily the data science libraries \texttt{scipy} {[}\protect\hyperlink{ref-scipy}{30}{]}, \texttt{numpy} {[}\protect\hyperlink{ref-numpy}{8}{]} and \texttt{pandas} {[}\protect\hyperlink{ref-pandas}{25}{]}.
 
@@
 The function set used by all DAE-GP variations for symbolic regression is summarized in tables \ref{tab:baseParams-GP}.
-The terminal set consists of the the input variables \(x1,x2,x3,x4,x5\) and the ephemeral integer constants listed in table \ref{tab:baseParams-GP} {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{35}{]}.
+The terminal set consists of the the input variables \(x1,x2,x3,x4,x5\) and the ephemeral integer constants listed in table \ref{tab:baseParams-GP} {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{34}{]}.
 
@@
 
-To gather a deeper understanding about the effect of pre-training DAE-LSTM in DAE-GP a series of experiments was conducted using the airfoil dataset for symbolic regression while using different parameter configurations for the number of hidden layers as well as the number of hidden neurons per hidden layer.
-
-To study generalization behavior this series of experiments is conducted using DAE-GP with only a single generation until the search terminates.
-I disregard the fitness of solutions found and focus solely on the reconstruction error that is produced during the training of each DAE-LSTM.
+To gain a deeper understanding about the effect of pre-training the DAE-LSTM networks in DAE-GP, a series of experiments was conducted using the airfoil dataset for symbolic regression while using different parameter configurations for the number of hidden layers as well as the number of hidden neurons per hidden layer.
+
+To study the generalization behavior of the DAE-LSTM model, this series of experiments is conducted using DAE-GP with only a single generation until the search terminates.
+The fitness of solutions found was disregarded to focus solely on the reconstruction error that is produced during the training of each DAE-LSTM.
 The reconstruction error is measured for two separate populations, a training population \(P_{train}\) that is used to train our DAE-LSTM as well as a hold-out validation population \(P_{test}\).
@@
 
-For each configuration I tested traditional DAE-GP as well as a pre-trained DAE-GP resulting in 12 total sub experiments that were each based on 10 individual runs (total number of runs=\(120\)).
-To avoid creating biased results through the presence of sampling error inside the pre-training population (see {[}\protect\hyperlink{ref-sampling_err_gp}{24}{]}), the population size for the pre-training phase is chosen very high with the size of \(\hat{P}\) =10000 where 50\% of \(\hat{P}\) is used for the training population \(\hat{P}_{train}\) and the remaining 50\% are used for the hold-out validation population \(\hat{P}_{test}\).
-The training of DAE-LSTM uses a fixed number of 1000 epochs until termination and uses Adam optimization for gradient descent.
-The reason for using a high amount of 1000 fixed training epochs is to deliberately force the DAE-LSTM to overfit to the training data.
-
-In general I expect that with a growing number of trainable parameters (either by adding more hidden layers or more hidden neurons per layer) the DAE-LSTM will be more prone to overfit to the training population \(P_{train}\) resulting in a small reconstruction error for the training population and a large one for the validation population \(P_{test}\).
-The reason for this effect is that a larger network trained over an extended period of time (without the use of strategies like early stopping), has much more potential to learn noise from the training dataset than a smaller network, which is more likely to result in worsening performance for previously unseen cases {[}\protect\hyperlink{ref-weigend1994overfitting}{32}{]}.
-
-Based on the review of {[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{5}{]} I also expect that pre-training will have an insignificant or even negative influence on small DAE-LSTM instances while improving the networks generalization ability with growing size.
+For each configuration, traditional DAE-GP as well as a pre-trained DAE-GP was tested resulting in 12 total sub experiments that were each based on 10 individual runs (total number of runs=\(120\)).
+To avoid creating biased results through the presence of sampling error inside the pre-training population (see {[}\protect\hyperlink{ref-sampling_err_gp}{23}{]}), the population size for the pre-training phase is chosen very high with the size of \(\hat{P}\) =10000 where 50\% of \(\hat{P}\) is used for the training population \(\hat{P}_{train}\) and the remaining 50\% are used for the hold-out validation population \(\hat{P}_{test}\).
+The training of each DAE-LSTM stopped after a fixed number of 1000 epochs.
+The reason for using a high amount of 1000 fixed training epochs was to deliberately force the DAE-LSTM to overfit to the training data.
+
+In general, it is expected that with a growing number of trainable parameters (either by adding more hidden layers or more hidden neurons per layer) the DAE-LSTM will be more likely to overfit to the training population \(P_{train}\) which will result in a small reconstruction error for the training population and a large one for the validation population \(P_{test}\).
+The reason for this effect is that a larger network trained over an extended period of time (without the use of strategies like early stopping), has much more potential to learn noise from the training dataset than a smaller network, which is more likely to result in worsening performance for previously unseen cases {[}\protect\hyperlink{ref-weigend1994overfitting}{31}{]}.
+
+Based on the review of {[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{5}{]} it is also expected that pre-training will have an insignificant or even negative influence on small DAE-LSTM instances while improving the networks generalization ability with growing size.
 
@@
 
-The results for the experiment with varying number of hidden layers are summarized in figure \ref{fig:first-gen-airfoil-byLayers}.
+The results for the experiment with a varying number of hidden layers are summarized in figure \ref{fig:first-gen-airfoil-byLayers}.
 Again, it can be observed that by increasing the dimension of the DAE-LSTM network both algorithms increasingly overfit to the training data.
@@
 
-Figure \ref{fig:airfoil-median-fitness-30gens} shows both the mean and median of the best fitness for each generation on both the test and training set of the airfoil dataset for using 2 hidden layers inside the DAE-LSTM networks.
+Figure \ref{fig:airfoil-median-fitness-30gens} shows both the mean and median the best fitness by generations for the airfoil dataset.
 The results are aggregated for \(10\) individuals runs per algorithm and also include regular GP as a benchmark.
@@
 The first interesting observation from this experiment is, that the pre-trained DAE-GP shows slightly lower mean RMSE (higher fitness) but a higher median RMSE (lower fitness) than regular DAE-GP for both the training and the testing set in overall fitness.
-Regarding the presence of overfitting, both DAE-GP variants show only a very small gap between the testing and training fitness which indicates a good generalization behavior.
-Another interesting observation from figure \ref{fig:airfoil-median-fitness-30gens} is that pre-trained DAE-GP is faster at improving fitness than regular DAE-GP, this is especially true if only looking at the first 10 generations.
+Regarding the presence of overfitting, both DAE-GP variants show only a very small gap between the testing and training fitness which indicates a good generalization behavior.\\
+
+Another interesting observation from figure \ref{fig:airfoil-median-fitness-30gens} is that pre-trained DAE-GP is faster at improving fitness than regular DAE-GP.
 The median best RMSE on the test set after 10 generations for pre-trained DAE-GP is 17.093 which is smaller by a factor of 0.558 than regular DAE-GP with a median best RMSE of 30.658
@@
 
-Since previous work by {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{35}{]} demonstrated that DAE-GP, for a given number of 10.000 fitness evaluations, is able to produce solutions that are much smaller in tree size than regular GP, I also studied the influence of pre-training on the size of solutions found by DAE-GP.
+Since previous work by {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{34}{]} demonstrated that DAE-GP, for a given number of 10.000 fitness evaluations, is able to produce solutions that are much smaller in tree size than regular GP, the experiment also analyzed the influence of pre-training on the size of solutions found by DAE-GP.
+
 Figure \ref{fig:airfoil-size-combined-30gens} shows the median of both the average size of individuals inside the population as well as the size of the currently best performing solution inside the population.
-It can be observed that both DAE-GP variants strongly reduce the average solutions size already in the first generations and that the average solution size inside the population constantly stays at this very low level for the rest of the search.
+It can be observed that both DAE-GP variants strongly reduce the average solution size already in the first generations and that the average solution size inside the population constantly stays at this very low level for the rest of the search.
 The median average size over all generations is 1.536 for DAE-GP and 1.536 for pre-trained DAE-GP.
@@
 Another interesting metric to examine in population-based optimization algorithms is the diversity of the current generations population.
-Since individual solutions in DAE-GP are computer programs in the form of parse trees, I selected the normalized Levenshtein edit distance {[}\protect\hyperlink{ref-NormLevDistance}{16}{]} as my primary metric for tracking population diversity.
+Since individual solutions in DAE-GP are computer programs in the form of parse trees, the normalized Levenshtein edit distance {[}\protect\hyperlink{ref-NormLevDistance}{16}{]} was selected as a metric for tracking population diversity.
 Figure \ref{fig:airfoil-popDiversity} shows the median population diversity for both DAE-GP algorithms and regular GP as a benchmark.
 While the population diversity steadily decreases over the generations for standard GP, DAE-GP variants strongly increase the population diversity in the first generations and keep them at a higher level during the evolutionary search.
-Here it can be observed, that pre-training shows a positive effect by increasing population diversity in comparison to regular DAE-GP (Median over all generations for DAE-GP: 0.861, Pre-Trained DAE-GP: 0.925, Standard-GP 0.705)
+Here it can be observed, that pre-training shows a positive effect by increasing population diversity in comparison to regular DAE-GP (Median over all generations for DAE-GP: 0.861, Pre-Trained DAE-GP: 0.925, Standard-GP 0.705).
 
@@
 To ensure a fair comparison between pre-trained and regular DAE-GP, the prior experiments used a static number of hidden neurons for each hidden Layer inside the DAE-LSTM networks.
-For comparison, figure \ref{fig:airfoil-maxIndSize-params} shows the total number of trainable parameters for each generation's DAE-LSTM of another experiment where regular DAE-GP was tested with dynamic adjustment of the number of hidden neurons as described in {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{36}{]} and {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{35}{]}.
-Again, the experiments was based on \(10\) individual runs and otherwise used the same hyperparameters as listed in tables \ref{tab:baseParams-GP} and \ref{tab:baseParams-DAEGP}.
+For comparison, figure \ref{fig:airfoil-maxIndSize-params} shows the total number of trainable parameters for each generations DAE-LSTM of another experiment where regular DAE-GP was tested with dynamic adjustment of the number of hidden neurons as described in {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{35}{]} and {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{34}{]}.
+Again, the experiment was based on \(10\) individual runs and otherwise used the same hyperparameters as listed in tables \ref{tab:baseParams-GP} and \ref{tab:baseParams-DAEGP}.
 It clearly shows that DAE-GP can strongly reduce the number of trainable parameters by adjusting the number of hidden neurons per hidden layers to the current generations population which largely reduces the amount of training necessary and likely reduce the run-time of DAE-GP.
@@
 
-As expected, this reduction of hidden neurons by dynamic adjustment, does also result in a reduction of the fitness of solutions found.
+As expected, the reduction of hidden neurons by dynamic adjustment does also result in a reduction of the fitness of solutions found.
 Figure \ref{fig:airfoil-maxIndSize-fitness} shows that pre-trained DAE-GP, with its more complex DAE-LSTMs over the full 30 generations, does achieve a higher median fitness than regular DAE-GP with dynamic adjustment of hidden neurons.
@@
 
-Since previous research on DAE-GP used a single hidden layers for the airfoil dataset {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{35}{]}, I repeated the experiment with a single hidden layer.
-Based on the findings of {[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{5}{]}, I expected that pre-training DAE-GP would lead to a decrease in overall performance.
-Figures \ref{fig:airfoil-fitness-perGen} and \ref{fig:airfoil-finalFitness} shows the mean and median best fitness per generation as well as the distribution of final fitness scores after 30 generations.
+Since previous research on DAE-GP used a single hidden layers for the airfoil dataset {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{34}{]}, the experiment was repeated with a single hidden layer.
+Based on the findings of {[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{5}{]}, it is to be expected that pre-training DAE-GP with a single hidden layer inside the DAE-LSTM networks leads to a decrease in overall performance.
+Figures \ref{fig:airfoil-fitness-perGen} and \ref{fig:airfoil-finalFitness} show the mean and median best fitness per generation as well as the distribution of final fitness scores after 30 generations.
 The results are, again, based on \(10\) individual runs per algorithm.
@@
 Interestingly, figure \ref{fig:airfoil-finalFitness} shows that the distribution of final fitness scores for pre-trained DAE-GP has a larger interquartile range than regular DAE-GP.
-This is opposed to the distributions of final fitness score that was measured for the same problem run with two hidden layers (see figure \ref{fig:airfoil-distribution-fitness-30gens}).
-Otherwise, pre-trained DAE-GP again has more outliers for high RMSE than regular DAE-GP as seen in the previous experiment with 2 hidden layers.
+This is opposed to the distributions of final fitness scores that was measured for the same problem run with two hidden layers (see figure \ref{fig:airfoil-distribution-fitness-30gens}).
+Otherwise, pre-trained DAE-GP again has more outliers for high RMSE than regular DAE-GP as seen in the previous experiment with two hidden layers.
 
@@
 The pre-training implementation in previous experiments used \texttt{ramped\ half\ and\ half} initialization {[}\protect\hyperlink{ref-Koza1993GeneticP}{15}{]} to create \(\hat{P}\).
-Using ramped half and half leads to 50\% of the population being initialized by the \texttt{full} method, where terminals are only inserted at the maximum allowed depth of a tree resulting in full trees.
-An alternative method of initializing \(\hat{P}\) is to use the \texttt{grow} initialization method where solutions are not guaranteed to be full trees since every leaf above the maximum depth of a tree can also be a terminal expression.
+Using \texttt{ramped\ half\ and\ half} leads to 50\% of the population being initialized by the \texttt{full} method, where terminals are only inserted at the maximum allowed depth of a tree resulting in full trees.
+An alternative method of initializing \(\hat{P}\) is to use the \texttt{grow} initialization method where solutions are not guaranteed to be full trees since every leaf below the maximum depth of a tree can also be a terminal expression.
 This different structural composition of solutions inside the pre-training population might be another way to improve the pre-training strategy by training on a population that more closely resembles the structural properties of further generations population where a lower percentage of full grown trees is to be expected.
@@
 
-Surprisingly, as shown in figure \ref{fig:airfoil-runtime-total}, for the experiment conducted on the airfoil dataset with 2 hidden layers a strong negative impact of pre-training can be observed for the median run-time of DAE-GP.
+Surprisingly, as shown in figure \ref{fig:airfoil-runtime-total}, for the experiment conducted on the airfoil dataset with two hidden layers, a strong negative impact of pre-training can be observed for the median run-time of DAE-GP.
 If the time spend on pre-training is included into the total run-time, the median duration for pre-trained DAE-GP over all 10 runs increases by a factor of 3.121 from a median run-time of 4712.303 seconds for regular DAE-GP to 14706.47 seconds for the pre-trained algorithm.
@@
 
-To investigate the strong negative impact of pre-training on DAE-GP run-time, I first studied the influence of pre-training on the median number of training epochs that had to be spend at each generation for the training of the DAE-LSTM models \(M_g\) (excluding the pre-training DAE-LSTM model \(\hat{M}\)).
-Figure \ref{fig:airfoil-trainingEpochs} shows that pre-training, as expected, strongly reduces the number of training epochs that had to be executed at each generation until the termination condition was reached.
-While regular DAE-GP trains for a median of 122 epochs at each generation, pre-training reduces the number of median training epochs by a factor of 0.439 to 53.5 training epochs.
+To investigate the strong negative impact of pre-training on DAE-GP run-time, the influence of pre-training on the median number of training epochs that had to be spend at each generation for the training of the DAE-LSTM models \(M_g\) (excluding the pre-training DAE-LSTM model \(\hat{M}\)) was studied.
+Figure \ref{fig:airfoil-trainingEpochs} shows that pre-training, as expected, strongly reduces the number of training epochs that had to be executed at each generation until the termination condition was satisfied.
+While regular DAE-GP trained for a median of 122 epochs at each generation, pre-training reduces the number of median training epochs by a factor of 0.439 to 53.5 training epochs.
 
@@
 
-While investigating other reasons for the differences in run-time, I verified that both DAE-GP variants use the exact same number of trainable parameters for each generations DAE-LSTMS models (total number of trainable parameters DAE-GP=573473, Pre-trained DAE-GP=573473).
+While investigating other reasons for the differences in run-time, it was verified that both DAE-GP variants use the exact same number of trainable parameters for each generation's DAE-LSTM model \(M_g\) (total number of trainable parameters DAE-GP=573473, Pre-trained DAE-GP=573473).
 After profiling a single run of DAE-GP and pre-trained DAE-GP using the python \texttt{cProfiler} module, my best explanation for the remaining differences in run-time is an inefficiency in my implementation of pre-training.
@@
 
-To gather more confidence in the previous results, I applied the pre-training strategy for DAE-GP to three additional real world symbolic regression problems.
+To gather more confidence in the previous results, the pre-training strategy for DAE-GP was applied to three additional real world symbolic regression problems.
 Table \ref{tab:full-run-realWorldSymReg-problems} briefly summarizes the benchmark problems.
@@
 Boston\_Housing & 506 & 13 & {[}\protect\hyperlink{ref-BOSTON_HOUSING_DATASET}{9}{]} \\
-Energy\_Cooling & 768 & 8 & {[}\protect\hyperlink{ref-ENERGY_DATASET}{27}{]} \\
-Concrete & 1030 & 8 & {[}\protect\hyperlink{ref-CONCRETE_DATASET}{37}{]} \\
+Energy\_Cooling & 768 & 8 & {[}\protect\hyperlink{ref-ENERGY_DATASET}{26}{]} \\
+Concrete & 1030 & 8 & {[}\protect\hyperlink{ref-CONCRETE_DATASET}{36}{]} \\
 \end{longtable}
 
-To test for statistical differences of the results achieved by either pre-trained or regular DAE-GP, I selected the nonparametric Mann-Whitney-U Test ({[}\protect\hyperlink{ref-mwu1}{33}{]}, {[}\protect\hyperlink{ref-mwu2}{17}{]}) to test for differences in the distribution underlying the samples sourced from my experiments.
-Additionally, I included the effect strength of pre-training as measured by Cliffs Delta {[}\protect\hyperlink{ref-Cliff1993DominanceSO}{3}{]} into the statistical analysis.
+To test for statistical differences of the results achieved by either pre-trained or regular DAE-GP, the nonparametric Mann-Whitney-U Test ({[}\protect\hyperlink{ref-mwu1}{32}{]}, {[}\protect\hyperlink{ref-mwu2}{17}{]}) was selected to test for differences in the underlying distribution.
+Additionally, the effect strength of pre-training as measured by Cliffs Delta {[}\protect\hyperlink{ref-Cliff1993DominanceSO}{3}{]} was included into the statistical analysis.
 Better median values (e.g.~higher fitness) are printed bold and p-values are marked with asterisk (1,2,3) for \(\alpha\) levels (\(0.1\), \(0.05\), \(0.01\)).
-Note that I included the results for the airfoil problem with a single hidden layer into the statistical evaluation which is not visualized during this chapter (see subsection \protect\hyperlink{reduced-number-of-hidden-layers-1}{Reduced Number of Hidden Layers}).
+Note that the results for the airfoil problem with a single hidden layer were also included into the statistical evaluation which is not visualized during this chapter (see subsection \protect\hyperlink{reduced-number-of-hidden-layers-1}{Reduced Number of Hidden Layers}).
 
@@
 
-As expected after reviewing figure \ref{fig:symReg-Fitness}, none of the differences in median best fitness are statistically significant and pre-training has only a very small, negligible effect strength for Cliff's Delta ranging between -0.14 and 0.17 {[}\protect\hyperlink{ref-Vargha2000ACA}{29}{]}.
+As expected after reviewing figure \ref{fig:symReg-Fitness}, none of the differences in median best fitness are statistically significant and pre-training has only a very small, negligible effect strength for Cliff's Delta ranging between -0.14 and 0.17 {[}\protect\hyperlink{ref-Vargha2000ACA}{28}{]}.
 
@@
 
-Next I studied the size of solutions found during DAE-GP.
+Next, the size of solutions found during DAE-GP was analyzed.
 Figure \ref{fig:symReg-Size} shows the median size of the best solution per generation aggregated over all benchmark problems for symbolic regression.
@@
 The experiments conducted in the scope of this master thesis did find that the implemented pre-training strategy resulted in a highly statistically significant increase in population diversity measured by the normalized levenshtein edit distance (see subsection \protect\hyperlink{influence-on-population-diversity-1}{Influence on Population Diversity}).
-Increasing the population diversity can be an important mechanism in controlling the search behavior of evolutionary algorithms and is especially important for the global exploration of the search space and the avoidance of premature convergence of the search {[}\protect\hyperlink{ref-Sudholt2020}{25}{]}.
+Increasing the population diversity can be an important mechanism in controlling the search behavior of evolutionary algorithms and is especially important for the global exploration of the search space and the avoidance of premature convergence of the search {[}\protect\hyperlink{ref-Sudholt2020}{24}{]}.
 
@@
 For a single hidden layer, pre-training did show a negative impact on the median final fitness of solutions found as it was suggested by {[}\protect\hyperlink{ref-pmlr-v9-erhan10a}{4}{]}.
-Since current research on DAE-GP for real-world symbolic regression relies on a single hidden layer (see {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{35}{]}), pre-training does come with additional computational expenses for using deeper DAE-LSTMs.
+Since current research on DAE-GP for real-world symbolic regression relies on a single hidden layer (see {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{34}{]}), pre-training does come with additional computational expenses for using deeper DAE-LSTMs.
 
@@
 One major limitation for the experiments conducted during the research phase of this thesis has been the large computational effort for experimenting with DAE-GP.
-Since single runs for DAE-GP have a run-time of many hours, it was not possible for me to extend the experiments to larger population sizes or more generations.
-Also I did not perform an extended optimization of hyperparameters (e.g by using a grid search strategy) which might yield configurations that increase the benefits of pre-training.
+Since single runs for DAE-GP have a run-time of many hours, it was not possible to extend the experiments to larger population sizes or more generations.
+Also no extended optimization of hyperparameters (e.g by using a grid search strategy) have been performed which might yield configurations that increase the benefits of pre-training.
 As described in subsection \protect\hyperlink{influence-on-run-time-1}{Influence on Run-Time}, sampling new solutions has been a large contributor to the large performance hit on run-time by using pre-training.
@@
 Another limitation of this research is the software implementation used for all experiments.
-As briefly touched in the subsection \protect\hyperlink{influence-on-run-time-1}{Influence on Run-Time}, I found some evidence that my implementation of pre-training suffered from increased run-time costs for internal error handling in the used \texttt{keras} library for deep neural networks.
+As briefly touched in the subsection \protect\hyperlink{influence-on-run-time-1}{Influence on Run-Time}, some evidence was found that my implementation of pre-training suffered from increased run-time costs for internal error handling in the used \texttt{keras} library for deep neural networks.
 By optimizing the source code or even re-implementing it in a more performance-focused programming language (such as \texttt{C++} or \texttt{Rust}), large improvements to the pre-training strategy might be achievable.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-ipython}{}}%
+\leavevmode\vadjust pre{\hypertarget{ref-harmless_overfitting_eda}{}}%
 \CSLLeftMargin{{[}21{]} }%
-\CSLRightInline{Fernando Pérez and Brian E.
-Granger.
-2007.
-{IP}ython: A system for interactive scientific computing.
-\emph{Computing in Science and Engineering} 9, 3 (May 2007), 21--29.
-DOI:https://doi.org/\href{https://doi.org/10.1109/MCSE.2007.53}{10.1109/MCSE.2007.53}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-harmless_overfitting_eda}{}}%
-\CSLLeftMargin{{[}22{]} }%
 \CSLRightInline{Malte Probst and Franz Rothlauf.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-design_of_modern_heuristics}{}}%
+\CSLLeftMargin{{[}22{]} }%
+\CSLRightInline{Franz Rothlauf. 2011. \emph{Design of modern heuristics: Principles and application}. DOI:https://doi.org/\href{https://doi.org/10.1007/978-3-540-72962-4}{10.1007/978-3-540-72962-4}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-sampling_err_gp}{}}%
 \CSLLeftMargin{{[}23{]} }%
-\CSLRightInline{Franz Rothlauf. 2011. \emph{Design of modern heuristics: Principles and application}. DOI:https://doi.org/\href{https://doi.org/10.1007/978-3-540-72962-4}{10.1007/978-3-540-72962-4}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-sampling_err_gp}{}}%
-\CSLLeftMargin{{[}24{]} }%
 \CSLRightInline{Dirk Schweim, David Wittenberg, and Franz Rothlauf.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Sudholt2020}{}}%
-\CSLLeftMargin{{[}25{]} }%
+\CSLLeftMargin{{[}24{]} }%
 \CSLRightInline{Dirk Sudholt.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-pandas}{}}%
+\CSLLeftMargin{{[}25{]} }%
+\CSLRightInline{The pandas development team. 2020. Pandas-dev/pandas: pandas. DOI:https://doi.org/\href{https://doi.org/10.5281/zenodo.3509134}{10.5281/zenodo.3509134}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-ENERGY_DATASET}{}}%
 \CSLLeftMargin{{[}26{]} }%
-\CSLRightInline{The pandas development team. 2020. Pandas-dev/pandas: pandas. DOI:https://doi.org/\href{https://doi.org/10.5281/zenodo.3509134}{10.5281/zenodo.3509134}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-ENERGY_DATASET}{}}%
-\CSLLeftMargin{{[}27{]} }%
 \CSLRightInline{Athanasios Tsanas and Angeliki Xifara.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-python3}{}}%
+\CSLLeftMargin{{[}27{]} }%
+\CSLRightInline{Guido Van Rossum and Fred L. Drake. 2009. \emph{Python 3 reference manual}. CreateSpace, Scotts Valley, CA.}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Vargha2000ACA}{}}%
 \CSLLeftMargin{{[}28{]} }%
-\CSLRightInline{Guido Van Rossum and Fred L. Drake. 2009. \emph{Python 3 reference manual}. CreateSpace, Scotts Valley, CA.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Vargha2000ACA}{}}%
-\CSLLeftMargin{{[}29{]} }%
 \CSLRightInline{András Vargha and Harold D Delaney.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-dae_orig2008}{}}%
-\CSLLeftMargin{{[}30{]} }%
+\CSLLeftMargin{{[}29{]} }%
 \CSLRightInline{Pascal Vincent, Hugo Larochelle, Y.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-scipy}{}}%
-\CSLLeftMargin{{[}31{]} }%
+\CSLLeftMargin{{[}30{]} }%
 \CSLRightInline{Pauli Virtanen, Ralf Gommers, Travis E.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-weigend1994overfitting}{}}%
+\CSLLeftMargin{{[}31{]} }%
+\CSLRightInline{Andreas Weigend. 1994. On overfitting and the effective number of hidden units. In \emph{Proceedings of the 1993 connectionist models summer school}, 335--342.}
+
+\leavevmode\vadjust pre{\hypertarget{ref-mwu1}{}}%
 \CSLLeftMargin{{[}32{]} }%
-\CSLRightInline{Andreas Weigend. 1994. On overfitting and the effective number of hidden units. In \emph{Proceedings of the 1993 connectionist models summer school}, 335--342.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-mwu1}{}}%
+\CSLRightInline{Frank. Wilcoxon. 1945. Individual comparisons by ranking methods. \emph{Biometrics} 1, (1945), 196--202.}
+
+\leavevmode\vadjust pre{\hypertarget{ref-daegp_explore_exploit}{}}%
 \CSLLeftMargin{{[}33{]} }%
-\CSLRightInline{Frank. Wilcoxon. 1945. Individual comparisons by ranking methods. \emph{Biometrics} 1, (1945), 196--202.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-daegp_explore_exploit}{}}%
-\CSLLeftMargin{{[}34{]} }%
 \CSLRightInline{David Wittenberg.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-dae-gp_2022_symreg}{}}%
-\CSLLeftMargin{{[}35{]} }%
+\CSLLeftMargin{{[}34{]} }%
 \CSLRightInline{David Wittenberg and Franz Rothlauf.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-dae-gp_2020_rtree}{}}%
-\CSLLeftMargin{{[}36{]} }%
+\CSLLeftMargin{{[}35{]} }%
 \CSLRightInline{David Wittenberg, Franz Rothlauf, and Dirk Schweim.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-CONCRETE_DATASET}{}}%
-\CSLLeftMargin{{[}37{]} }%
+\CSLLeftMargin{{[}36{]} }%
 \CSLRightInline{I-Cheng Yeh.

### 2023-02-14 paper.Rmd
autogit

@@
 If the time spend on pre-training is included into the total run-time, the median duration for pre-trained DAE-GP over all 10 runs increases by a factor of `r round(PT_TIME / REG_TIME, 3)` from a median run-time of `r format(REG_TIME, scientific=F)` seconds for regular DAE-GP to `r format(PT_TIME, scientific=F)` seconds for the pre-trained algorithm.
-Even in the case where the time spend for pre-training is excluded from the total run-time, pre-training still increases the run-time of DAE-GP by a factor of `r round( PT_TIME_NOPT / REG_TIME, 3)` to `r format(PT_TIME_NOPT, scientific=F)` seconds.
+Even when the time spend for pre-training is excluded from the total run-time, pre-training still increases the run-time of DAE-GP by a factor of `r round( PT_TIME_NOPT / REG_TIME, 3)` to `r format(PT_TIME_NOPT, scientific=F)` seconds.
 
@@
 Surprisingly, even after correcting the run-time for both DAE-GP results by removing the time spend during sampling, my implementation of pre-training for DAE-GP still on average takes more time to finish the search process as shown in figure \@ref(fig:airfoil-runtime-noSampling-total).
-While traditional DAE-GP has a median run-time without sampling time of `r format(REG_TIME_NOSAMLING, scientific=F)` seconds , the run-time of pre-trained DAE-GP, without time for sampling and pre-training, still increases by a factor of `r round(PT_TIME_NOSAMPLING_NOPT / REG_TIME_NOSAMLING, 3)` to `r format(PT_TIME_NOSAMPLING_NOPT, scientific=F)` seconds.
+While traditional DAE-GP has a median corrected run-time of `r format(REG_TIME_NOSAMLING, scientific=F)` seconds , the corrected run-time of pre-trained DAE-GP still increases by a factor of `r round(PT_TIME_NOSAMPLING_NOPT / REG_TIME_NOSAMLING, 3)` to `r format(PT_TIME_NOSAMPLING_NOPT, scientific=F)` seconds.
 
@@
 
-As expected after reviewing figure \@ref(fig:symReg-Fitness), none of the differences in median best fitness are statistically significant and pre-training has only a very small, negligible effect strength for Cliff's Delta ranging between `r min(as.numeric(df_symreg_fitness$Cliffs_Delta))` and `r max(as.numeric(df_symreg_fitness$Cliffs_Delta))` [@Vargha2000ACA].
+As expected after reviewing figure \@ref(fig:symReg-Fitness), none of the differences in median best fitness are statistically significant and pre-training only has a small, negligible effect strength with Cliff's Delta ranging between `r min(as.numeric(df_symreg_fitness$Cliffs_Delta))` and `r max(as.numeric(df_symreg_fitness$Cliffs_Delta))` [@Vargha2000ACA].
 
-Figure \@ref(fig:symReg-Fitness-BoxPlots) shows the distribution of final fitness values on both the test and training set as box plots.
+Figure \@ref(fig:symReg-Fitness-BoxPlots) shows the distribution of final fitness values on both the test and the training set as box plots.
 It shows that depending on the symbolic regression problem, the dispersion of fitness values is different for both DAE-GP variants.
@@
 For Airfoil, Boston Housing and Energy Cooling data-sets, the median size of the best solution are equal, only for the concrete dataset a smaller median size for pre-trained DAE-GP was achieved (pre-trained DAE-GP also showed a higher median best fitness for the concrete problem).
-The results of the Mann-Whitney-U Test again show, that differences in the underlying distributions are not statistically significant which also shows in low effect strengths of using pre-training on the size of the best found solutions (Cliff's Delta ranges from`r min(as.numeric(df_size$Cliffs_Delta))` to `r max(as.numeric(df_size$Cliffs_Delta))`).
+The results of the Mann-Whitney-U Test show, that differences in the underlying distributions are not statistically significant, which also shows in low effect strengths of pre-training on the size of the best found solutions (Cliff's Delta ranges from`r min(as.numeric(df_size$Cliffs_Delta))` to `r max(as.numeric(df_size$Cliffs_Delta))`).
 
@@
 
-The results of the statistical analysis is summarized in table \@ref(tab:full-run-realWorldSymReg-popDiversity).
-For all sub experiments, computing the Mann-Whitney-U Test returns P-Values that are smaller than $0.01$ which indicates a highly statistically significant difference in the underlying distributions.
+The results of the statistical analysis are summarized in table \@ref(tab:full-run-realWorldSymReg-popDiversity).
+For all sub experiments, computing the Mann-Whitney-U Test returns P-Values smaller than $0.01$ which indicates a highly statistically significant difference in the underlying distributions.
 Looking at the effect strength of pre-training, it shows that pre-training does have a strong positive effect (Cliff's Delta ranges from `r min(as.numeric(df_popDiv$Cliffs_Delta))` to `r max(as.numeric(df_popDiv$Cliffs_Delta))`) on population diversity inside the population.
@@
 
-Looking at the results of the statistical analysis summarized in table \@ref(tab:full-run-realWorldSymReg-epochsPerGen), computing the Mann-Whitney-U test does consistently return P-Values below $0.01$ for all experiments which proves the statistical significant differences in distributions.
-Analyzing the effect strength of pre-training on the number of training epochs shows values for Cliff's Delta in the range of `r min(as.numeric(df_epochs$Cliffs_Delta))` and `r max(as.numeric(df_epochs$Cliffs_Delta))` which indicates a very strong negative effect on the number of training epochs necessary at each generation of the search.
+Looking at the results of the statistical analysis summarized in table \@ref(tab:full-run-realWorldSymReg-epochsPerGen), computing the Mann-Whitney-U test does consistently return P-Values below $0.01$ for all experiments.
+Analyzing the effect strength of pre-training on the number of training epochs shows values for Cliff's Delta in the range of `r min(as.numeric(df_epochs$Cliffs_Delta))` and `r max(as.numeric(df_epochs$Cliffs_Delta))` which indicates a strong negative effect strength.
+This evidence suggests that pre-training can substantially decrease the required number of training epochs, indicating a statistically significant impact.
+
 Computing the median number of training epochs over all generations for all benchmark problems shows a total decrease by a factor of `r round( (median(df_epochs$Pre.Trained_DAE.GP) / median(df_epochs$DAE.GP)) , 2)` from `r round(median(df_epochs$DAE.GP), 2)` epochs for regular DAE-GP to `r round(median(df_epochs$Pre.Trained_DAE.GP), 2)` epochs for pre-trained DAE-GP.
@@
 
-Another significant benefit of using pre-training in DAE-GP is that it is highly efficient in decreasing the number of training epochs that are necessary to train each generations DAE-LSTM model (see subsection [Influence on Population Diversity](#influence-on-population-diversity-1)).
+Another significant benefit of using pre-training in DAE-GP is that it is highly efficient in decreasing the number of training epochs that are necessary to train each generation´s DAE-LSTM model (see subsection [Influence on Population Diversity](#influence-on-population-diversity-1)).
 Even though pre-training increased the median total run-time for the tested benchmark problem in combination with the chosen DAE-LSTM dimensions, reducing the number of training epochs might still be a possible way to lower the run time of DAE-GP if further performance optimizations, especially during model sampling, can be achieved.
@@
 
-The experiments conducted in this thesis did not find any significant benefits on both solution quality as well as run-time for solving real world symbolic regression problems with pre-trained DAE-GP in comparison to regular DAE-GP.
+The experiments conducted in this thesis did not find any significant benefits, in terms of solution quality or run-time, when using pre-trained DAE-GP compared to regular DAE-GP for solving real-world symbolic regression problems.
 
@@
 As described in subsection [Influence on Run-Time](#influence-on-run-time-1), sampling new solutions has been a large contributor to the large performance hit on run-time by using pre-training.
-It could be a worthwhile endevour to explore the main factors that increase sampling time for pre-trained DAE-GP to more efficiently exploit the benefits of pre-training like reducing the number of training epochs.
+It could be a worthwhile endevour to explore the main factors that increase sampling time for pre-trained DAE-GP to more efficiently exploit the benefits of pre-training such as the reducing the number of training epochs.
 
@@
 Another possibility to improve the benefits of pre-training for DAE-GP is based on the findings of @pmlr-v5-erhan09a.
-Pre-Training has shown to be especially useful for the lower layers of DAEs so it could be worthwhile to further experiment with only adapting the weights for a subset of the hidden layers inside the DAE-LSTM networks to either reduce run-time or to improve on the quality of solutions found during the evolutionary search.
+Pre-Training has shown to be especially useful for the lower layers of DAEs therefore it could be worthwhile to further experiment with only adapting the weights for a subset of the hidden layers inside the DAE-LSTM networks to either reduce run-time or to improve on the quality of solutions found during the evolutionary search.
 

### 2023-02-14 paper.tex
autogit

@@
 If the time spend on pre-training is included into the total run-time, the median duration for pre-trained DAE-GP over all 10 runs increases by a factor of 3.121 from a median run-time of 4712.303 seconds for regular DAE-GP to 14706.47 seconds for the pre-trained algorithm.
-Even in the case where the time spend for pre-training is excluded from the total run-time, pre-training still increases the run-time of DAE-GP by a factor of 2.66 to 12536.89 seconds.
+Even when the time spend for pre-training is excluded from the total run-time, pre-training still increases the run-time of DAE-GP by a factor of 2.66 to 12536.89 seconds.
 
@@
 Surprisingly, even after correcting the run-time for both DAE-GP results by removing the time spend during sampling, my implementation of pre-training for DAE-GP still on average takes more time to finish the search process as shown in figure \ref{fig:airfoil-runtime-noSampling-total}.
-While traditional DAE-GP has a median run-time without sampling time of 3550.277 seconds , the run-time of pre-trained DAE-GP, without time for sampling and pre-training, still increases by a factor of 1.865 to 6622.027 seconds.
+While traditional DAE-GP has a median corrected run-time of 3550.277 seconds , the corrected run-time of pre-trained DAE-GP still increases by a factor of 1.865 to 6622.027 seconds.
 
@@
 
-As expected after reviewing figure \ref{fig:symReg-Fitness}, none of the differences in median best fitness are statistically significant and pre-training has only a very small, negligible effect strength for Cliff's Delta ranging between -0.14 and 0.17 {[}\protect\hyperlink{ref-Vargha2000ACA}{28}{]}.
-
-Figure \ref{fig:symReg-Fitness-BoxPlots} shows the distribution of final fitness values on both the test and training set as box plots.
+As expected after reviewing figure \ref{fig:symReg-Fitness}, none of the differences in median best fitness are statistically significant and pre-training only has a small, negligible effect strength with Cliff's Delta ranging between -0.14 and 0.17 {[}\protect\hyperlink{ref-Vargha2000ACA}{28}{]}.
+
+Figure \ref{fig:symReg-Fitness-BoxPlots} shows the distribution of final fitness values on both the test and the training set as box plots.
 It shows that depending on the symbolic regression problem, the dispersion of fitness values is different for both DAE-GP variants.
@@
 For Airfoil, Boston Housing and Energy Cooling data-sets, the median size of the best solution are equal, only for the concrete dataset a smaller median size for pre-trained DAE-GP was achieved (pre-trained DAE-GP also showed a higher median best fitness for the concrete problem).
-The results of the Mann-Whitney-U Test again show, that differences in the underlying distributions are not statistically significant which also shows in low effect strengths of using pre-training on the size of the best found solutions (Cliff's Delta ranges from-0.33 to 0.2).
+The results of the Mann-Whitney-U Test show, that differences in the underlying distributions are not statistically significant, which also shows in low effect strengths of pre-training on the size of the best found solutions (Cliff's Delta ranges from-0.33 to 0.2).
 
@@
 
-The results of the statistical analysis is summarized in table \ref{tab:full-run-realWorldSymReg-popDiversity}.
-For all sub experiments, computing the Mann-Whitney-U Test returns P-Values that are smaller than \(0.01\) which indicates a highly statistically significant difference in the underlying distributions.
+The results of the statistical analysis are summarized in table \ref{tab:full-run-realWorldSymReg-popDiversity}.
+For all sub experiments, computing the Mann-Whitney-U Test returns P-Values smaller than \(0.01\) which indicates a highly statistically significant difference in the underlying distributions.
 Looking at the effect strength of pre-training, it shows that pre-training does have a strong positive effect (Cliff's Delta ranges from 0.92 to 0.94) on population diversity inside the population.
@@
 
-Looking at the results of the statistical analysis summarized in table \ref{tab:full-run-realWorldSymReg-epochsPerGen}, computing the Mann-Whitney-U test does consistently return P-Values below \(0.01\) for all experiments which proves the statistical significant differences in distributions.
-Analyzing the effect strength of pre-training on the number of training epochs shows values for Cliff's Delta in the range of -0.93 and -0.87 which indicates a very strong negative effect on the number of training epochs necessary at each generation of the search.
+Looking at the results of the statistical analysis summarized in table \ref{tab:full-run-realWorldSymReg-epochsPerGen}, computing the Mann-Whitney-U test does consistently return P-Values below \(0.01\) for all experiments.
+Analyzing the effect strength of pre-training on the number of training epochs shows values for Cliff's Delta in the range of -0.93 and -0.87 which indicates a strong negative effect strength.
+This evidence suggests that pre-training can substantially decrease the required number of training epochs, indicating a statistically significant impact.
+
 Computing the median number of training epochs over all generations for all benchmark problems shows a total decrease by a factor of 0.45 from 143.49 epochs for regular DAE-GP to 65.15 epochs for pre-trained DAE-GP.
@@
 
-Another significant benefit of using pre-training in DAE-GP is that it is highly efficient in decreasing the number of training epochs that are necessary to train each generations DAE-LSTM model (see subsection \protect\hyperlink{influence-on-population-diversity-1}{Influence on Population Diversity}).
+Another significant benefit of using pre-training in DAE-GP is that it is highly efficient in decreasing the number of training epochs that are necessary to train each generation´s DAE-LSTM model (see subsection \protect\hyperlink{influence-on-population-diversity-1}{Influence on Population Diversity}).
 Even though pre-training increased the median total run-time for the tested benchmark problem in combination with the chosen DAE-LSTM dimensions, reducing the number of training epochs might still be a possible way to lower the run time of DAE-GP if further performance optimizations, especially during model sampling, can be achieved.
@@
 As described in subsection \protect\hyperlink{influence-on-run-time-1}{Influence on Run-Time}, sampling new solutions has been a large contributor to the large performance hit on run-time by using pre-training.
-It could be a worthwhile endevour to explore the main factors that increase sampling time for pre-trained DAE-GP to more efficiently exploit the benefits of pre-training like reducing the number of training epochs.
+It could be a worthwhile endevour to explore the main factors that increase sampling time for pre-trained DAE-GP to more efficiently exploit the benefits of pre-training such as the reducing the number of training epochs.
 
@@
 Another possibility to improve the benefits of pre-training for DAE-GP is based on the findings of {[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{5}{]}.
-Pre-Training has shown to be especially useful for the lower layers of DAEs so it could be worthwhile to further experiment with only adapting the weights for a subset of the hidden layers inside the DAE-LSTM networks to either reduce run-time or to improve on the quality of solutions found during the evolutionary search.
+Pre-Training has shown to be especially useful for the lower layers of DAEs therefore it could be worthwhile to further experiment with only adapting the weights for a subset of the hidden layers inside the DAE-LSTM networks to either reduce run-time or to improve on the quality of solutions found during the evolutionary search.
 

### 2023-02-15 paper.Rmd
autogit

@@
 This thesis investigates the possible benefits and downsides of using pre-training for the DAE-LSTM networks of DAE-GP for four real world symbolic regression problems.
-The experiments conducted did show that pre-training can drastically reduce the number of training epochs that are necessary at each generation of the DAE-GP search while also increasing the diversity of the population.
+The experiments conducted did show that pre-training can drastically reduce the number of epochs that are necessary for the DAE-LSTM training at each generation of the DAE-GP search.
+Another interesting finding was that pre-training increases the levenshtein edit distance between individual solutions inside the population which is a metric for the diversity of a population.
 Unfortunately pre-training did not show any improvements for both final fitness and the final size of solutions while largely increasing the total run-time for DAE-GP.

### 2023-02-15 paper.Rmd
Wed Feb 15 09:00:06 CET 2023: update-remote.sh

@@
 The experiments conducted did show that pre-training can drastically reduce the number of epochs that are necessary for the DAE-LSTM training at each generation of the DAE-GP search.
-Another interesting finding was that pre-training increases the levenshtein edit distance between individual solutions inside the population which is a metric for the diversity of a population.
+Another interesting finding was that pre-training also increases the levenshtein edit distance between individual solutions inside the population which is a metric for the diversity of a population.
 Unfortunately pre-training did not show any improvements for both final fitness and the final size of solutions while largely increasing the total run-time for DAE-GP.

### 2023-02-15 paper.Rmd
Wed Feb 15 10:00:06 CET 2023: update-remote.sh

@@
 
-The pre-training strategy used in this thesis is applied to the DAE-LSTM models $M_g$ that are used in each DAE-GP generation $g$ for $g\in{1,...,g_{max}}$.
+The pre-training strategy used in this thesis is applied to the DAE-LSTM models $M_g$ that are used in each DAE-GP generation $g$ for $g \in \{{x \mid x \text{ is a number and } 1 \leq x \leq g_{max}}\}$.
 Instead of initializing and optimizing each model from scratch as done in previous work (e.g.
@@
 
-The trainable parameters $\theta_{\hat{M}}$ obtained after finishing the training procedure of $\hat{M}$ are then used as the starting parameters for each following DAE-LSTM model $M_g$ for $g\in{1,...,g_{max}}$.
+The trainable parameters $\theta_{\hat{M}}$ obtained after finishing the training procedure of $\hat{M}$ are then used as the starting parameters for each following DAE-LSTM model $M_g$.
 
@@
 For a pre-training implementation, this strategy can not be easily adapted since it leads to a changing number of neurons at each generation resulting in different dimensions of the DAE-LSTM.
-To allow the sharing of the state of the pre-trained parameters $\theta_{\hat{M}}$ from the pre-trained model $\hat{M}$ to each $M_g$ for $g\in{1,...,g_{max}}$ the number of hidden neurons per hidden layer for all DAE-GP runs, pre-trained as well as regular, is statically set to a predefined value.
+To allow the sharing of pre-trained parameters $\theta_{\hat{M}}$ from the pre-trained model $\hat{M}$ to each $M_g$, the number of hidden neurons per hidden layer for all DAE-GP runs, pre-trained as well as regular, is statically set to a predefined value.
 
@@
 
-\newpage
+<!-- \newpage -->
 
-# Statutory Declaration {#IV .unnumbered}
+<!-- # Statutory Declaration {#IV .unnumbered} -->
 
-```{r statutory_declaration, echo=FALSE}
-knitr::include_graphics("~/masterThesis/master_thesis/private/statutory_declaration.pdf")
-```
+<!-- ```{r statutory_declaration, echo=FALSE} -->
+
+<!-- knitr::include_graphics("~/masterThesis/master_thesis/private/statutory_declaration.pdf") -->
+
+<!-- ``` -->

### 2023-02-15 paper.tex
Wed Feb 15 10:00:06 CET 2023: update-remote.sh

@@
 \author{}
-\date{\vspace{-2.5em}Date of Submission: 2023-02-14}
+\date{\vspace{-2.5em}Date of Submission: 2023-02-15}
 
@@
 This thesis investigates the possible benefits and downsides of using pre-training for the DAE-LSTM networks of DAE-GP for four real world symbolic regression problems.
-The experiments conducted did show that pre-training can drastically reduce the number of training epochs that are necessary at each generation of the DAE-GP search while also increasing the diversity of the population.
+The experiments conducted did show that pre-training can drastically reduce the number of epochs that are necessary for the DAE-LSTM training at each generation of the DAE-GP search.
+Another interesting finding was that pre-training also increases the levenshtein edit distance between individual solutions inside the population which is a metric for the diversity of a population.
 Unfortunately pre-training did not show any improvements for both final fitness and the final size of solutions while largely increasing the total run-time for DAE-GP.
@@
 
-The pre-training strategy used in this thesis is applied to the DAE-LSTM models \(M_g\) that are used in each DAE-GP generation \(g\) for \(g\in{1,...,g_{max}}\).
+The pre-training strategy used in this thesis is applied to the DAE-LSTM models \(M_g\) that are used in each DAE-GP generation \(g\) for \(g \in \{{x \mid x \text{ is a number and } 1 \leq x \leq g_{max}}\}\).
 Instead of initializing and optimizing each model from scratch as done in previous work (e.g.
@@
 
-The trainable parameters \(\theta_{\hat{M}}\) obtained after finishing the training procedure of \(\hat{M}\) are then used as the starting parameters for each following DAE-LSTM model \(M_g\) for \(g\in{1,...,g_{max}}\).
+The trainable parameters \(\theta_{\hat{M}}\) obtained after finishing the training procedure of \(\hat{M}\) are then used as the starting parameters for each following DAE-LSTM model \(M_g\).
 
@@
 For a pre-training implementation, this strategy can not be easily adapted since it leads to a changing number of neurons at each generation resulting in different dimensions of the DAE-LSTM.
-To allow the sharing of the state of the pre-trained parameters \(\theta_{\hat{M}}\) from the pre-trained model \(\hat{M}\) to each \(M_g\) for \(g\in{1,...,g_{max}}\) the number of hidden neurons per hidden layer for all DAE-GP runs, pre-trained as well as regular, is statically set to a predefined value.
+To allow the sharing of pre-trained parameters \(\theta_{\hat{M}}\) from the pre-trained model \(\hat{M}\) to each \(M_g\), the number of hidden neurons per hidden layer for all DAE-GP runs, pre-trained as well as regular, is statically set to a predefined value.
 

### 2023-02-15 paper.Rmd
autogit

@@
 
+Denoising Autoencoder Genetic Programming (DAE-GP) ist ein Estimation of Distribution Algorithmus (EDA) aus dem Forschungfeld der genetischen Programmierung (GP).
+In DAE-GP werden Denoising Autoencoders Long Short-Term Memory Networks (DAE-LSTM) als probabilistische Modelle verwendet um neue Populationen von Lösungen im zu erzeugen.
+Diese Masterarbeit untersucht die Vor- und Nachteile des Einsatzes einer Pre-Training Strategie für die DAE-LSTM Netzwerke von DAE-GP an vier Daten für symbolische Regression.
+Die durchgefüh
+
 # Keywords {.unlisted .unnumbered}

### 2023-02-15 paper.Rmd
autogit

--- 2023-02-15 paper.Rmd
@@
 
+## English {#Ia .unnumbered .unlisted}
+
 Denoising Autoencoder Genetic Programming (DAE-GP) is an Estimation of Distribution Algorithm in the domain of Genetic Programming that uses Denoising Autoencoders Long Short-Term Memory Networks (DAE-LSTM) as probabilistic models for sampling new populations of solutions.
@@
 Unfortunately pre-training did not show any improvements for both final fitness and the final size of solutions while largely increasing the total run-time for DAE-GP.
+
+## Deutsch {#Ib .unnumbered .unlisted}
+
+Denoising Autoencoder Genetic Programming (DAE-GP) ist ein Estimation of Distribution Algorithmus (EDA) aus dem Forschungfeld der genetischen Programmierung (GP).
+In DAE-GP werden Denoising Autoencoders Long Short-Term Memory Netzwerke (DAE-LSTM) als probabilistische Modelle verwendet um neue Populationen von Lösungen für eine evolutionäre Suche zu erzeugen.
+Diese Masterarbeit untersucht die Vor- und Nachteile des Einsatzes einer Pre-Training Strategie für die DAE-LSTM Netzwerke von DAE-GP an vier Datensätzen für symbolische Regression.
+Die durchgeführten Experimente haben gezeigt, dass Pre-Training die Anzahl von Trainingsepochen für die DAE-LSTM Netzwerke in jeder Generation statistisch signifikant reduzieren können.
+Außerdem zeigt sich, dass Pre-Training die Levenshtein Editierdistanz, ein Maß für die Populationsdiversität, signifikant erhöhen konnte.
+Leider konnte Pre-Training die Qualität der jeweils besten gefundenen Lösungen, weder im Bezug auf ihre Fitness noch auf ihre Größe, verbessern.
+Auch führte Pre-Training in den durchgeführten Experimenten zu einer starken Erhöhung der Laufzeit des DAE-GP Algorithmus.
 
@@
 
-<!-- \newpage -->
-
-<!-- # Statutory Declaration {#IV .unnumbered} -->
-
-<!-- ```{r statutory_declaration, echo=FALSE} -->
-
-<!-- knitr::include_graphics("~/masterThesis/master_thesis/private/statutory_declaration.pdf") -->
-
-<!-- ``` -->
+\newpage
+
+# Statutory Declaration {#IV .unnumbered}
+
+```{r statutory_declaration, echo=FALSE}
+knitr::include_graphics("~/masterThesis/master_thesis/private/statutory_declaration_signed.pdf")
+```

### 2023-02-15 paper.tex
autogit

@@
 
+\hypertarget{Ia}{%
+\subsection*{English}\label{Ia}}
+
 Denoising Autoencoder Genetic Programming (DAE-GP) is an Estimation of Distribution Algorithm in the domain of Genetic Programming that uses Denoising Autoencoders Long Short-Term Memory Networks (DAE-LSTM) as probabilistic models for sampling new populations of solutions.
@@
 Unfortunately pre-training did not show any improvements for both final fitness and the final size of solutions while largely increasing the total run-time for DAE-GP.
+
+\hypertarget{Ib}{%
+\subsection*{Deutsch}\label{Ib}}
+
+Denoising Autoencoder Genetic Programming (DAE-GP) ist ein Estimation of Distribution Algorithmus (EDA) aus dem Forschungfeld der genetischen Programmierung (GP).
+In DAE-GP werden Denoising Autoencoders Long Short-Term Memory Netzwerke (DAE-LSTM) als probabilistische Modelle verwendet um neue Populationen von Lösungen für eine evolutionäre Suche zu erzeugen.
+Diese Masterarbeit untersucht die Vor- und Nachteile des Einsatzes einer Pre-Training Strategie für die DAE-LSTM Netzwerke von DAE-GP an vier Datensätzen für symbolische Regression.
+Die durchgeführten Experimente haben gezeigt, dass Pre-Training die Anzahl von Trainingsepochen für die DAE-LSTM Netzwerke in jeder Generation statistisch signifikant reduzieren können.
+Außerdem zeigt sich, dass Pre-Training die Levenshtein Editierdistanz, ein Maß für die Populationsdiversität, signifikant erhöhen konnte.
+Leider konnte Pre-Training die Qualität der jeweils besten gefundenen Lösungen, weder im Bezug auf ihre Fitness noch auf ihre Größe, verbessern.
+Auch führte Pre-Training in den durchgeführten Experimenten zu einer starken Erhöhung der Laufzeit des DAE-GP Algorithmus.
 
@@
 
-\begin{center}\includegraphics{private/statutory_declaration} \end{center}
+\begin{center}\includegraphics{private/statutory_declaration_signed} \end{center}
 

### 2023-02-15 statutory_declaration/statutory_declaration.Rmd
autogit

---
output: pdf_document
pagestyle: empty
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Statutory Declaration

I hereby assure that I have independently and without using sources or resources other than those indicated, written this paper.
Directly quoted sentences or sentence parts are marked as citations, while other sources of inspiration, in terms of content and scope, are identified with references.
The work has not been presented to any examination board in the same or a similar form and has not been published.
It has not been used, even partially, for any other examination or academic achievement.
I am aware of the regulations for ensuring good scientific practice in research and teaching, as well as the procedures for dealing with scientific misconduct.

Place, Date:

------------------------------------------------------------------------

Signature:

------------------------------------------------------------------------

## Ehrenwörtliche Erklärung

Ich versichere hiermit, dass ich die vorliegende Arbeit selbständig und ohne Benutzung anderer als der angegebenen Quellen und Hilfsmittel verfasst habe.
Wörtlich übernommene Sätze oder Satzteile sind als Zitat belegt, andere Anlehnungen, hinsichtlich Aussage und Umfang, unter Quellenangabe kenntlich gemacht.
Die Arbeit hat in gleicher oder ähnlicher Form noch keiner Prüfungsbehörde vorgelegen und ist nicht veröffentlicht.
Sie wurde nicht, auch nicht auszugsweise, für eine andere Prüfungs- oder Studienleistung verwendet.
Von der Ordnung zur Sicherung guter wissenschaftlicher Praxis in Forschung und Lehre und zum Verfahren zum Umgang mit wissenschaftlichem Fehlverhalten wurde Kenntnis genommen wurde.

Ort, Datum:

------------------------------------------------------------------------

Unterschrift:

------------------------------------------------------------------------

### 2023-02-15 paper.Rmd
autogit

@@
 The framework that was used to run all experiments of this thesis was provided by the supervisor of this thesis David Wittenberg and uses the python programming language [@python3] in conjunction with the baseline libraries `Keras` [@chollet2015keras] for deep neural networks and `deap` [@DEAP_JMLR2012] for evolutionary computation.
-For data visualizations the python package `matplotlib` [@matplotlib] was used, statistical analysis and evaluations were conducted using primarily the data science libraries `scipy` [@scipy], `numpy` [@numpy] and `pandas` [@pandas].
+For data visualizations the python package `matplotlib` [@matplotlib] was used, statistical analysis and evaluations were conducted using primarily the data science libraries `numpy` [@numpy], `scipy` [@scipy] and `pandas` [@pandas].
 
@@
 
-```{r statutory_declaration, echo=FALSE}
+```{r statutory_declaration, echo=FALSE, fig.align = 'center', out.width='500px'}
 knitr::include_graphics("~/masterThesis/master_thesis/private/statutory_declaration_signed.pdf")

### 2023-02-15 paper.tex
autogit

@@
 The framework that was used to run all experiments of this thesis was provided by the supervisor of this thesis David Wittenberg and uses the python programming language {[}\protect\hyperlink{ref-python3}{27}{]} in conjunction with the baseline libraries \texttt{Keras} {[}\protect\hyperlink{ref-chollet2015keras}{2}{]} for deep neural networks and \texttt{deap} {[}\protect\hyperlink{ref-DEAP_JMLR2012}{6}{]} for evolutionary computation.
-For data visualizations the python package \texttt{matplotlib} {[}\protect\hyperlink{ref-matplotlib}{12}{]} was used, statistical analysis and evaluations were conducted using primarily the data science libraries \texttt{scipy} {[}\protect\hyperlink{ref-scipy}{30}{]}, \texttt{numpy} {[}\protect\hyperlink{ref-numpy}{8}{]} and \texttt{pandas} {[}\protect\hyperlink{ref-pandas}{25}{]}.
+For data visualizations the python package \texttt{matplotlib} {[}\protect\hyperlink{ref-matplotlib}{12}{]} was used, statistical analysis and evaluations were conducted using primarily the data science libraries \texttt{numpy} {[}\protect\hyperlink{ref-numpy}{8}{]}, \texttt{scipy} {[}\protect\hyperlink{ref-scipy}{30}{]} and \texttt{pandas} {[}\protect\hyperlink{ref-pandas}{25}{]}.
 
@@
 
-\begin{center}\includegraphics{private/statutory_declaration_signed} \end{center}
+\begin{center}\includegraphics[width=500px]{private/statutory_declaration_signed} \end{center}
 

### 2023-02-16 paper.Rmd
Thu Feb 16 15:00:07 CET 2023: update-remote.sh

@@
 
-\setcounter{page}{1}
+\setcounter{page}{2}
 
@@
 \listoftables
+\thispagestyle{empty}
+\newpage
 \listoffigures
@@
 In DAE-GP, DAE-LSTMs are trained with a subset of high-fitness solutions selected from a parent population with the aim to capture their promising qualities.
-The resulting model is then used to sample new offspring solutions by propagating partially mutated solutions from the parent population through the DAE-LSTMS [@dae-gp_2020_rtree].
+The resulting model is then used to sample new offspring solutions by propagating partially mutated solutions from the parent population through the DAE-LSTM [@dae-gp_2020_rtree].
 In previous work DAE-GP has been shown to outperform GP for both a generalized version of the royal tree problem [@dae-gp_2020_rtree] as well as for a real-world symbolic regression problem [@dae-gp_2022_symreg].
@@
 
-\newpage
-
 # Statutory Declaration {#IV .unnumbered}
@@
 ```{r statutory_declaration, echo=FALSE, fig.align = 'center', out.width='500px'}
-knitr::include_graphics("~/masterThesis/master_thesis/private/statutory_declaration_signed.pdf")
-```
+knitr::include_graphics("~/masterThesis/master_thesis/private/statGood.pdf")
+```

### 2023-02-16 paper.Rmd
autogit
Thu Feb 16 16:00:07 CET 2023: update-remote.sh

@@
 
-Figure \@ref(fig:airfoil-median-fitness-30gens) shows both the mean and median the best fitness by generations for the airfoil dataset.
+Figure \@ref(fig:airfoil-median-fitness-30gens) shows both the mean and median best fitness by generations for the airfoil dataset.
 The results are aggregated for $10$ individuals runs per algorithm and also include regular GP as a benchmark.
@@
 
-Another interesting metric to examine in population-based optimization algorithms is the diversity of the current generations population.
+Another interesting metric to examine in population-based optimization algorithms is the population diversity over all generations of the search.
 Since individual solutions in DAE-GP are computer programs in the form of parse trees, the normalized Levenshtein edit distance [@NormLevDistance] was selected as a metric for tracking population diversity.
@@
 Again, the experiment was based on $10$ individual runs and otherwise used the same hyperparameters as listed in tables \@ref(tab:baseParams-GP) and \@ref(tab:baseParams-DAEGP).
-It clearly shows that DAE-GP can strongly reduce the number of trainable parameters by adjusting the number of hidden neurons per hidden layers to the current generations population which largely reduces the amount of training necessary and likely reduce the run-time of DAE-GP.
+It clearly shows that DAE-GP can strongly reduce the number of trainable parameters by adjusting the number of hidden neurons per hidden layers to the current generations population which largely reduces the amount of training necessary and likely reduces the run-time of DAE-GP.
 
@@
 The basic idea is to run regular DAE-GP for the first generation without any interference, after completing the first generation, the pre-training population $\hat{P}$ is created by sampling the DAE-LSTM model $M_1$.
-The pre-training model $\hat{M}$ is then trained using $\hat{P}$, after finishing, DAE-GP continues running and initializes each new DAE-LSTM model $M_g$ for $g\in{2,...,g_{max}}$ using the trainable parameters $\theta_{\hat{M}}$.
+The pre-training model $\hat{M}$ is then trained using $\hat{P}$, after finishing, DAE-GP continues running and initializes each new DAE-LSTM model $M_g$ for $g \in \{{x \mid x \text{ is a number and } 2 \leq x \leq g_{max}}\}$ using the trainable parameters $\theta_{\hat{M}}$.
 
@@
 An alternative method of initializing $\hat{P}$ is to use the `grow` initialization method where solutions are not guaranteed to be full trees since every leaf below the maximum depth of a tree can also be a terminal expression.
-This different structural composition of solutions inside the pre-training population might be another way to improve the pre-training strategy by training on a population that more closely resembles the structural properties of further generations population where a lower percentage of full grown trees is to be expected.
+This different structural composition of solutions inside the pre-training population might be another way to improve the pre-training strategy by training on a population that more closely resembles the structural properties of further generation's populations where a lower percentage of full grown trees is to be expected.
 
@@
 
-Finally, figure \@ref(fig:symReg-EpochsPerGen) shows the median number of epochs spend on training each generations DAE-LSTM network $M_g$ until the termination criterion was satisfied.
+Finally, figure \@ref(fig:symReg-EpochsPerGen) shows the median number of epochs spend on training each generation's DAE-LSTM network $M_g$ until the termination criterion was satisfied.
 Pre-Training shows to be very effective at reducing the number of training epochs in comparison to regular DAE-GP.

### 2023-02-16 paper.tex
Thu Feb 16 15:00:07 CET 2023: update-remote.sh
Thu Feb 16 16:00:07 CET 2023: update-remote.sh

@@
 \author{}
-\date{\vspace{-2.5em}Date of Submission: 2023-02-15}
+\date{\vspace{-2.5em}Date of Submission: 2023-02-16}
 
@@
 }
-\setcounter{page}{1}
+\setcounter{page}{2}
 
@@
 In DAE-GP, DAE-LSTMs are trained with a subset of high-fitness solutions selected from a parent population with the aim to capture their promising qualities.
-The resulting model is then used to sample new offspring solutions by propagating partially mutated solutions from the parent population through the DAE-LSTMS {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{35}{]}.
+The resulting model is then used to sample new offspring solutions by propagating partially mutated solutions from the parent population through the DAE-LSTM {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{35}{]}.
 In previous work DAE-GP has been shown to outperform GP for both a generalized version of the royal tree problem {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{35}{]} as well as for a real-world symbolic regression problem {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{34}{]}.
@@
 
-Figure \ref{fig:airfoil-median-fitness-30gens} shows both the mean and median the best fitness by generations for the airfoil dataset.
+Figure \ref{fig:airfoil-median-fitness-30gens} shows both the mean and median best fitness by generations for the airfoil dataset.
 The results are aggregated for \(10\) individuals runs per algorithm and also include regular GP as a benchmark.
@@
 
-Another interesting metric to examine in population-based optimization algorithms is the diversity of the current generations population.
+Another interesting metric to examine in population-based optimization algorithms is the population diversity over all generations of the search.
 Since individual solutions in DAE-GP are computer programs in the form of parse trees, the normalized Levenshtein edit distance {[}\protect\hyperlink{ref-NormLevDistance}{16}{]} was selected as a metric for tracking population diversity.
@@
 Again, the experiment was based on \(10\) individual runs and otherwise used the same hyperparameters as listed in tables \ref{tab:baseParams-GP} and \ref{tab:baseParams-DAEGP}.
-It clearly shows that DAE-GP can strongly reduce the number of trainable parameters by adjusting the number of hidden neurons per hidden layers to the current generations population which largely reduces the amount of training necessary and likely reduce the run-time of DAE-GP.
+It clearly shows that DAE-GP can strongly reduce the number of trainable parameters by adjusting the number of hidden neurons per hidden layers to the current generations population which largely reduces the amount of training necessary and likely reduces the run-time of DAE-GP.
 
@@
 The basic idea is to run regular DAE-GP for the first generation without any interference, after completing the first generation, the pre-training population \(\hat{P}\) is created by sampling the DAE-LSTM model \(M_1\).
-The pre-training model \(\hat{M}\) is then trained using \(\hat{P}\), after finishing, DAE-GP continues running and initializes each new DAE-LSTM model \(M_g\) for \(g\in{2,...,g_{max}}\) using the trainable parameters \(\theta_{\hat{M}}\).
+The pre-training model \(\hat{M}\) is then trained using \(\hat{P}\), after finishing, DAE-GP continues running and initializes each new DAE-LSTM model \(M_g\) for \(g \in \{{x \mid x \text{ is a number and } 2 \leq x \leq g_{max}}\}\) using the trainable parameters \(\theta_{\hat{M}}\).
 
@@
 
-\newpage
-
 \hypertarget{IV}{%
@@
 
-\begin{center}\includegraphics[width=500px]{private/statutory_declaration_signed} \end{center}
+\begin{center}\includegraphics[width=500px]{private/statGood} \end{center}
 

### 2023-02-16 paper.Rmd
final commit :-)

@@
 Another interesting finding was that pre-training also increases the levenshtein edit distance between individual solutions inside the population which is a metric for the diversity of a population.
-Unfortunately pre-training did not show any improvements for both final fitness and the final size of solutions while largely increasing the total run-time for DAE-GP.
+Unfortunately, pre-training did not yield any improvements in the final fitness or size of solutions, despite significantly increasing the total run-time for DAE-GP.
 
@@
 Diese Masterarbeit untersucht die Vor- und Nachteile des Einsatzes einer Pre-Training Strategie für die DAE-LSTM Netzwerke von DAE-GP an vier Datensätzen für symbolische Regression.
-Die durchgeführten Experimente haben gezeigt, dass Pre-Training die Anzahl von Trainingsepochen für die DAE-LSTM Netzwerke in jeder Generation statistisch signifikant reduzieren können.
+Die durchgeführten Experimente haben gezeigt, dass Pre-Training die Anzahl von Trainingsepochen für die DAE-LSTM Netzwerke in jeder Generation statistisch signifikant reduzieren konnte.
 Außerdem zeigt sich, dass Pre-Training die Levenshtein Editierdistanz, ein Maß für die Populationsdiversität, signifikant erhöhen konnte.
@@
 For 100 hidden neurons, regular DAE-GP not only starts to overfit to the training data earlier than pre-trained DAE-GP, also it can be seen that pre-trained DAE-GP starts to overfit at a lower reconstruction error which would result in a higher model performance if training strategies such as early stopping would be employed.
-Doubling the number of hidden neurons to 200 results in very strong overfitting behavior for both DAE-GP variants starting around the 200.th epoch.
+Doubling the number of hidden neurons to 200 results in very strong overfitting behavior for both DAE-GP variants starting around the 200th epoch.
 
@@
 While for a single hidden layer overfitting begins at around the 400th.
-epoch, doubling the number of hidden layers to two leads to both algorithms overfitting at around the 200.th epoch.
+epoch, doubling the number of hidden layers to two leads to both algorithms overfitting at around the 200th epoch.
 In general over all three sub experiments, both algorithms start to overfit after approximately the same number of epochs but pre-training leads to reaching the point of overfitting at a higher reconstruction error which would result in a inferior model performance if the early stopping training strategy would be used during model training.
@@
 Regarding the size of the best solution per generation, the results show that the pre-trained version of algorithm did produce solutions with the same median size as traditional DAE-GP.
-The median size of the best solutions found after terminating the search is the same for both algorithms (DAE-GP: `r round(REG_MED_BESTSIZE, 3)`, Pre-Trained DAE-GP: `r round(PT_MED_BESTSIZE, 3)`)
+The median size of the best solutions found after terminating the search is the same for both algorithms (DAE-GP: `r round(REG_MED_BESTSIZE, 3)`, Pre-Trained DAE-GP: `r round(PT_MED_BESTSIZE, 3)`).
 

### 2023-02-16 paper.tex
final commit :-)

@@
 Regarding the size of the best solution per generation, the results show that the pre-trained version of algorithm did produce solutions with the same median size as traditional DAE-GP.
-The median size of the best solutions found after terminating the search is the same for both algorithms (DAE-GP: 5, Pre-Trained DAE-GP: 5)
+The median size of the best solutions found after terminating the search is the same for both algorithms (DAE-GP: 5, Pre-Trained DAE-GP: 5).
 
@@
 An alternative method of initializing \(\hat{P}\) is to use the \texttt{grow} initialization method where solutions are not guaranteed to be full trees since every leaf below the maximum depth of a tree can also be a terminal expression.
-This different structural composition of solutions inside the pre-training population might be another way to improve the pre-training strategy by training on a population that more closely resembles the structural properties of further generations population where a lower percentage of full grown trees is to be expected.
+This different structural composition of solutions inside the pre-training population might be another way to improve the pre-training strategy by training on a population that more closely resembles the structural properties of further generation's populations where a lower percentage of full grown trees is to be expected.
 
@@
 
-Finally, figure \ref{fig:symReg-EpochsPerGen} shows the median number of epochs spend on training each generations DAE-LSTM network \(M_g\) until the termination criterion was satisfied.
+Finally, figure \ref{fig:symReg-EpochsPerGen} shows the median number of epochs spend on training each generation's DAE-LSTM network \(M_g\) until the termination criterion was satisfied.
 Pre-Training shows to be very effective at reducing the number of training epochs in comparison to regular DAE-GP.
